### 2019-08-06 paper/paper.tex
included zeroth draft of paper texfile and bibtex file

% Whisker barrel generation paper equations, for pasting into Google
% docs draft.

% While writing, don't stop for errors
%\nonstopmode

\documentclass[a4paper,11pt]{article}
\usepackage{amsmath,amssymb,mathtools}
\usepackage{enumerate}
\usepackage{times}
\usepackage{color}
\usepackage{parskip}
\usepackage{afterpage}
\usepackage{setspace}
\usepackage[modulo]{lineno}
%\usepackage{xltxtra}
%\usepackage{xunicode}
\usepackage{gensymb}

\setcounter{tocdepth}{3}
\usepackage{graphicx}
\usepackage{comment}
\usepackage{textcomp}
\usepackage{natbib}

\usepackage[labelfont=bf,labelsep=period,justification=justified]{caption}

\doublespacing
\topmargin 0.0cm
\oddsidemargin 0.5cm
\evensidemargin 0.5cm
\textwidth 16cm
\textheight 21cm
\makeatletter
\makeatother
\date{}
\pagestyle{plain}

%\newcommand{\dvrg}{\nabla\vcdot\nabla}
\newcommand{\mb}[1]{\mathbf{#1}}
\makeatletter
\newcommand*\vcdot{\mathpalette\vcdot@{.35}}
\newcommand*\vcdot@[2]{\mathbin{\vcenter{\hbox{\scalebox{#2}{$\m@th#1\bullet$}}}}}
\makeatother

%\bibliographystyle{plos2015}
\bibliographystyle{apalike}

\begin{document}
%\begin{center}
\begin{flushleft}
{\huge
\textbf{Self-organization of whisker barrels in simulation requires competition between thalamocortical axons}
 \vspace{5em}
}
%\end{center}

Sebastian~S.~James$^{1*}$, Stuart~P.~Wilson$^{1}$ \vspace{1em}\\
$^{1}$Department of Psychology, The University of Sheffield, Sheffield, United Kingdom.\vspace{1em}\\
$^{*}$Corresponding author email: Seb.James@Sheffield.ac.uk\vspace{1em}\\
Short title: Self-organising whisker barrels\vspace{5em}\\
\end{flushleft}

\newpage{}
%\raggedright

\section*{Abstract}

Understanding how genetic information and spontaneous pattern formation together give rise to the whisker barrels is an important challenge for developmental neurobiology. The barrel patterning that coalesces in layer IV has been described formally as a Voronoi tesselation, which is indicative of an underlying developmental process based on local competitive interactions. If pattern formation relies only on such intrinsic self-organising processes, how can the somatotopic order at the periphery be maintained in the projections from brainstem to thalamus and from thalamus to cortex? To investigate, we began by extending a computational model of neocortical arealization by Karbowski \& Ermentrout (2004; J. Comp. Neurosci.) to generate fields on a two-dimensional cortical sheet. Central to this model is an assumption that axonal branching and synaptogenesis are coupled, and that thalamic projections compete for cortical territory under the influence of molecular signalling gradients. We found that these processes alone are incapable of generating patterns of cortical fields that are more intricate than the patterns already present in molecular guidance signals. However, a full barrel-field pattern emerges under the guidance of as few as four signalling gradients when the model is extended to introduce competition between the branches of axons that originate from different thalamic barreloids. We show that the key requirements for the emergence of realistic (Voronoi tesselation) patterning are i) at each cortical location thalamocortical projections compete for a limited number of available connections, ii) at each location the axonal branching rate of each thalamocortical projection is reduced by the total branching density of the other projections, and iii) the density of axon branching for each thalamocortical projection is conserved. We demonstrate how a purely self-organising system can initialise a pattern which can then be faithfully replicated in multiple neural structures. We show that this minimal model of barrel self-organisation provides a parsimonious account of data from a range of classic studies in which intrinsic and extrinsic factors have been experimentally manipulated.

\section*{Introduction}

\modulolinenumbers{}
\linenumbers

Spatial patterns in neural connectivity provide clues about the constraints under which brains evolve and develop. Whether or not patterning governs information processing, models of the development of information processing circuits can be validated in terms of their ability to account for biological patterning (\citealp{Purves1992}; see also \citealp{Wilson2015,Bednar2016}). One of the most distinctive patterns of neural connectivity is found in the rodent barrel cortex (\citealp{Woolsey1948,Woolsey1970,Welker1974}, see \citealp{Fox2008}). Understanding how this particular pattern forms might shed light on the interplay of genetic and extrinsic factors that shape cortical circuits, and on a wealth of anatomical, physiological, and behavioural data obtained using this model system.

The barrels are large cortical columns, identifiable from birth in layer 4 of rodent primary somatosensory cortex as dense clusters of thalamocortical axons, which are enclosed by borders a few neurons thick from postnatal day 3 \citep{Erzurumlu2012}. In the plane tangential to the cortical surface the barrels constitute a map of the arrangement of the whiskers on the face, such that the cells of adjacent barrels respond most strongly and most quickly to deflection of adjacent whiskers \citep{Armstrong-James1992}. The barrels reflect subcortical whisker maps comprising cellular aggregates known as barrelettes in the brainstem and barreloids in the thalamus \citep{Killackey1980}.

Each whisker and its corresponding barrelette, barreloid, and barrel can be identified by its position on a two-dimensional grid \citep{Zucker1969,Killackey1980,Loos1976}. The barrels are normally up to twice as long in the axis aligned to the grid ‘rows’ compared to the ‘arcs’, and the rows are visible from day 2 \citep{Rebsam2002} whereas the arcs become distinct by day 3 \citep{Erzurumlu2012,Rebsam2002}. In metabotropic glutamate receptor 5  (mGlu5) knockouts the barrel rows develop but the arcs do not (\citealp{Hannan2001} see also \citealp{Fox2008}), thus distinct processes likely guide development of the barrels with respect to the orthogonal grid axes.

Barrel formation requires afferent input, driven by whisker stimulation and by thalamic calcium waves that propagate via mGlu5-receptor activation (\citealp{Anton2019}, see also \citealp{Anton2018}). Patterning also depends on a complex network of axon guidance processes, which includes interactions between guidance molecules like ephrin-A5 and A7 \citep{Miller2006} and between adhesion molecules like cadherin-6 and 8 \citep{Greig2013,Bishop2000}, and the binding of p75 to neurotropins (\citealp{Shimogori2005,Bishop2002}; see also \citealp{Dye2011a,Dye2011b}). This network is orchestrated by interactions between morphogens Fgf8 and Fgf17 and transcription factors Emx2, Pax6, Sp8, and Coup-tf1 \citep{Shimogori2005,Bishop2000}, which are expressed in gradients that mark orthogonal axes and can be manipulated to cause the barrels to stretch, shrink, shift, skew, and even duplicate \citep{Shimogori2005,Assimacopoulos2012,Borello2014,Sahara2007,Ypsilanti2016,Sur2005}.

The boundaries of the barrels form polygonal shapes that together create a Voronoi tesselation across the cortical sheet, wherein all points enclosed by a given boundary are closer to its center than to the center of any other boundary \citep{Senft1991}. This patterning suggests that the barrel topology may be preserved in the initial projection of the thalamocortical axons into the cortex, and that a barrel forms by the lateral branching of the axons from their initial centres of mass, which ceases upon contact with the axons branching from adjacent centres. According to this model the barrel pattern could be entirely determined by the distribution of the initial centre points, which would itself be determined by guidance cues. However, the assumption of pre-arranged centres is difficult to resolve with the observation that thalamocortical axons arrive in the cortical plate as an undifferentiated bundle prior to barreloid formation (see \citealp{Erzurumlu2012} for a review).

Alternatively, reaction-diffusion dynamics could generate a Voronoi-like tesselation in the cortex without the need for pre-arranged centres, by amplifying characteristic modes in the noisy initial distribution of bundled axon branches, as a net effect of short-range cooperative and longer-range competitive interactions between cells. According to this model the barrel pattern would be determined by the relative strength of these interactions and by the shape of the cortical field boundary. However, reaction-diffusion dynamics alone are unlikely to be able to account for the precise co-registration and topographic connectivity between emergent domains in thalamus and cortex, the non-uniform sizes and spacing of the barrels to form rows versus arcs, and the contribution of gene expression gradients.

As theories of barrel patterning, the centre-point and reaction-diffusion models are not mutually exclusive. Pre-organised centres in the initial conditions could bias reaction-diffusion processes to generate the specific somatotopic arrangement of barrels more reliably, and similarly the mechanisms by which axons branch laterally from barrel centres may involve a tension between cooperative and competitive interactions. However, a proof that barrel patterning can in principle emerge within an undifferentiated bundle of thalamocortical axons, based only on local interactions between cells, would remove the need to invoke a global supervisor, i.e., to assume a separate stage and/or extrinsic mechanism for pre-organising thalamocortical connections.

To this end, the aim of the present study is to establish whether it is possible for the observed arrangement of barrels to self-organise in a system governed by reaction-diffusion dynamics, under the guidance of molecular signalling gradients, and in the absence of pre-defined centre-points. Our approach is to represent in a computational model the minimal set of biological assumptions that might lead to the emergence of barrel patterning in simulation, and, if necessary, to add to this minimal set only assumptions that can be represented by the local exchange of information between adjacent cells. The emergence of barrel patterning under such conditions would serve as an existence proof of the plausibility of the hypothesis that barrel development is a self-organising process.

\section*{Models}

\cite{Karbowski2004} developed what remains, to the best of our knowledge, the only dynamical model of cortical arealization capable of generating predictions about the influence of signalling molecules on the size, shape, and topological organization of emergent cortical fields. This model describes competition between thalamocortical projections for territory along a one-dimensional string of cortical cells, e.g., along the anterior-posterior axis. The equations of the model define how the density of connections, $c$, and the density of axon branches, $a$, interact at time $t$ at each cortical location $x$, for $N$ thalamocortical projections indexed by $i$. The tesselation of the barrels that we investigate is a two-dimensional structure, hence we first extended the original model for simulation in two spatial dimensions, and describe it here as such (i.e., suibstituting $x$ for $\mb{x}$ and $\frac{d}{dx}$ for $\nabla$).

The model was derived from the assumption that the rates of synaptogenesis and thalamocortical axon branching are reciprocally coupled, resulting in an expression for the change in $c_i$ as a function of $a_i$, and for the change in $a_i$ as a function of $c_i$. Synaptogenesis is modelled  as

\begin{equation} \label{eq:Karb2DExt_dc}
\frac{\partial c_i(\mb{x},t)}{\partial t} =-\alpha c_i(\mb{x},t) +\beta  \left(1 - \sum_{j=1}^{N} c_j(\mb{x}, t)\right).[a_i(\mb{x},t)]^k.
\end{equation}

Accordingly when the final term evaluates to zero, the connection density decays exponentially at a rate given by the decay constant $\alpha$. This occurs, at a given location, when the sum of connections is one. Otherwise, the connection density is assumed to vary non-linearly ($k>1$) with changes in the axon branching density.

The change in density of axon branches is assumed to depend also on the influence of guidance molecules and on the density of branches at adjacent cortical locations:

%
\begin{equation} \label{eq:Karb2DExt_da}
\frac{\partial a_i(\mb{x},t)}{\partial t} = \nabla\vcdot\left(D \nabla a_i(\mb{x},t)-a_i(\mb{x},t)\sum_{j=1}^{M} \gamma_{i,j}\nabla \rho_j(\mb{x})\right) - \frac{\partial c_i(\mb{x},t)}{\partial t} + \chi
\end{equation}

The first term on the right describes the spatial gradient (indicated by $\nabla$) of a quantity referred to in models of heat diffusion as the `flux' (enclosed by parentheses). The central idea in such models is that the change in temperature (over time) at a given point on a surface is proportional to the gradient in the temperature (over space) at that location. Substituting heat for the number of branches, axons from a given thalamic projection are thus assumed to branch laterally across the cortical sheet at a rate given by the diffusion constant $D$.

Lateral branching is augmented by a measure of the local alignment between the density of branches and the gradient in the concentration of each of $M$ distinct molecular signals, $\rho(\mb{x})$. The influence of a given signal (indexed by $j$) on a given projection (indexed by $i$), is determined by the parameter $\gamma_i,j$. Interactions for which this value is positive encourage axon branching in the direction in which the concentration of the signal is increasing, whereas negative interactions encourage branching in the opposite direction. Hence the axons of a given projection are assumed to branch towards either higher or lower concentrations of each signalling molecule.

The second term on the right of equation 2 represents an assumption that the number of axon branches decreases when the number of connections being made increases, and vice versa, i.e., that branches are required to form connections. Note that we introduce the term $\chi=0$ here, so that we might subsequently redefine $\chi$ in order to introduce the competition between thalamic projections that we found to be necessary for the emergence of a Voronoi tesselation.

Two points are important to emphasise here. First, notice that the only difference between thalamic projections is the $M$ interaction strengths ($\gamma$) that are chosen to define each projection. Hence any (reliable) differences between the size, shape, and location of simulated cortical fields are due to differences in the choice of these values only. Second, the formulation of the model is at this point entirely local, in the sense that simulating the dynamics does not require any information to be communicated from one cortical cell to any but its immediate neighbours.

In addition to the choice of initial conditions (for the fields $c(\mb{x},t=0)$ and $a(\mb{x},t=0)$) and the axon guidance fields $\rho(\mb{x})$, the free parameters of the model are the number of thalamocortical projections ($N$), constants $\alpha$, $\beta$, $k$, and $D$, and the $N\times M$ interaction strengths, $\gamma$.

Equations 1 and 2 were solved on a hexagonal lattice of grid elements represent the cortical sheet ($\mb{x}$), under the assumption that axon branches cannot extend beyond the boundary of the tissue, i.e., $\mb{J}_i(\mb{x},t) \bigg\rvert_{boundary} = 0$. We developed software tools for imposing arbitrary boundary shapes, specified in a scalable vector graphics (svg) file format, on the lattice. The dynamics were integrated using fourth-order Runge-Kutta for ordinary differential equation 1, and using a {\color{red} finite difference method} for the flux term (see Supplementary Materials).

\section*{Results}

%\begin{equation}
%\chi=\frac{\epsilon_i  a_i}{N-1} \sum_{j \ne i}^{N} a_j(\mb{x}, t)^l
%\end{equation}

%And normalization step

%\begin{equation}\label{eq:norm_comp8}
%a_i' = n_e \; \frac {a_i} {\textstyle \sum_j^{n_e} a_{i,j} }
%\end{equation}

%($a_i'$ becomes $a_i$ in the next computational step.)

\section*{Discussion}

We set out to test the hypothesis that barrel patterning may emerge via self-organising processes. We first extended an existing model of neocortical arealization to simulate pattern formation in two dimensions, inheriting from this model the assumption that axonal branching and synaptogenesis are coupled, and that thalamic projections compete for cortical territory under the influence of molecular signalling gradients. Realistic barrel patterning emerged only when terms were added to represent the assumption that  thalamocortical projections compete for a limited supply of connections with cortical cells, that their branching rates are reduced by the branching of other thalamocortical projections, and that the density of axon branching is conserved for each projection. The net effect of these local interactions is a Voronoi tesselation in which localised clusters of thalamic projections become somatotopically organised on the cortical sheet according to the relative values of coefficients specifying the rates at which axons are attracted and repelled by orthogonal pairs of molecular signal gradients. According to this model, the pattern of barrels is determined by the competitive and cooperative interactions between cells relative to the shape of the cortical field boundary. In simulation, we systematically investigated the origin of the asymmetry between the barrel rows and arcs, and identified {\color{red} XXXXX} as the most likely candidate for the elongation of barrel shapes along the rows.

\bibliography{bibfile}

\end{document}

%%%%% ORIGINAL EQUATIONS MAINTAINED AS NOTES HERE FOR REFERENCE

\section{Karbowski system with N TCs and M gradients}
%
\begin{equation} \label{eq:Karb2D_dc}
\frac{\partial c_i(\mb{x},t)}{\partial t} = -\alpha c_i(\mb{x},t) + \beta n(\mb{x},t)
[a_i(\mb{x},t)]^k
\end{equation}
%
\begin{equation} \label{eq:Karb2D_conserve}
n(\mb{x},t) = 1 - \sum_{i=1}^{N} c_i(\mb{x}, t)
\end{equation}
%
\begin{equation} \label{eq:Karb2D_da}
\frac{\partial a_i(\mb{x},t)}{\partial t}
= \nabla\vcdot\mb{J}_i(\mb{x},t) - \frac{\partial c_i(\mb{x},t)}{\partial t}
\end{equation}

with the flux current

\begin{equation} \label{eq:Karb2D_J}
\mb{J}_i(\mb{x},t) = D \nabla a_i(\mb{x},t) - a_i
\sum_{j=1}^M \big(\gamma_{i,j} \nabla\rho_j(\mb{x}) \big)
\end{equation}

The boundary condition applied to these equations is:

\begin{equation}
\mb{J}_i(\mb{x},t) \bigg\rvert_{boundary} = 0
\end{equation}

\section{Karbowski system with N TCs and M gradients plus competition}

%
\begin{equation} \label{eq:Karb2DExt_dc}
\frac{\partial c_i(\mb{x},t)}{\partial t} = -\alpha c_i(\mb{x},t) + \beta n(\mb{x},t)
[a_i(\mb{x},t)]^k
\end{equation}
%
\begin{equation} \label{eq:Karb2DExt_conserve}
n(\mb{x},t) = 1 - \sum_{i=1}^{N} c_i(\mb{x}, t)
\end{equation}
%
\begin{equation} \label{eq:Karb2DExt_da}
\frac{\partial a_i(\mb{x},t)}{\partial t}
= \nabla\vcdot\mb{J}_i(\mb{x},t) - \frac{\partial c_i(\mb{x},t)}{\partial t}- \frac{\epsilon_i  a_i}{N-1} \sum_{j \ne i}^{N} a_j(\mb{x}, t)^l
\end{equation}

And normalization step

\begin{equation}\label{eq:norm_comp8}
a_i' = n_e \; \frac {a_i} {\textstyle \sum_j^{n_e} a_{i,j} }
\end{equation}

($a_i'$ becomes $a_i$ in the next computational step.)

### 2019-08-07 paper/paper.tex
revised the methods a little

@@
 According to this model the barrel pattern would be determined by the relative strength of these interactions and by the shape of the cortical field boundary.
-However, reaction-diffusion dynamics alone are unlikely to be able to account for the precise co-registration and topographic connectivity between emergent domains in thalamus and cortex, the non-uniform sizes and spacing of the barrels to form rows versus arcs, and the contribution of gene expression gradients.
+However, reaction-diffusion dynamics alone are unlikely to be able to account for the precise co-registration and topographic connectivity between emergent domains in thalamus and cortex, the non-uniform sizes and spacing of the barrels to form rows versus arcs, and the influence of gene expression gradients.
 
@@
 \cite{Karbowski2004} developed what remains, to the best of our knowledge, the only dynamical model of cortical arealization capable of generating predictions about the influence of signalling molecules on the size, shape, and topological organization of emergent cortical fields.
-This model describes competition between thalamocortical projections for territory along a one-dimensional string of cortical cells, e.g., along the anterior-posterior axis.
+This model describes competition between thalamocortical projections for territory along a one-dimensional string of cortical cells, e.g., along an anterior-posterior axis.
 The equations of the model define how the density of connections, $c$, and the density of axon branches, $a$, interact at time $t$ at each cortical location $x$, for $N$ thalamocortical projections indexed by $i$.
-The tesselation of the barrels that we investigate is a two-dimensional structure, hence we first extended the original model for simulation in two spatial dimensions, and describe it here as such (i.e., suibstituting $x$ for $\mb{x}$ and $\frac{d}{dx}$ for $\nabla$).
+The Voronoi tesselation used to describe barrel patterning is a two-dimensional structure, hence we first extended the original model for simulation on a two-dimensional cortical sheet, and describe it here as such (i.e., substituting $x$ for the two-dimensional position vector $\mb{x}$ and substituting terms of the form $\frac{d}{dx}$ for the two-dimensional spatial operator $\nabla$).
 
-The model was derived from the assumption that the rates of synaptogenesis and thalamocortical axon branching are reciprocally coupled, resulting in an expression for the change in $c_i$ as a function of $a_i$, and for the change in $a_i$ as a function of $c_i$.
+The original model was derived from the assumption that the rates at which thalamocortical synapses form and thalamocortical axons branch are reciprocally coupled, resulting in an expression for the change in $c_i$ as a function of $a_i$, and for the change in $a_i$ as a function of $c_i$.
 Synaptogenesis is modelled  as
@@
 \begin{equation} \label{eq:Karb2DExt_dc}
-\frac{\partial c_i(\mb{x},t)}{\partial t} =-\alpha c_i(\mb{x},t) +\beta  \left(1 - \sum_{j=1}^{N} c_j(\mb{x}, t)\right).[a_i(\mb{x},t)]^k.
+\frac{\partial c_i(\mb{x},t)}{\partial t} =-\alpha c_i(\mb{x},t) +\beta  \left(1 - \sum_{j=1}^{N} c_j(\mb{x}, t)\right)[a_i(\mb{x},t)]^k.
 \end{equation}
 
-Accordingly when the final term evaluates to zero, the connection density decays exponentially at a rate given by the decay constant $\alpha$.
-This occurs, at a given location, when the sum of connections is one.
+Accordingly, when the final term evaluates to zero, the density of thalamocortical connections decays exponentially at a rate given by the decay constant $\alpha$.
+This occurs where the density of connections over all thalamic projections sum to one.
 Otherwise, the connection density is assumed to vary non-linearly ($k>1$) with changes in the axon branching density.
 
-The change in density of axon branches is assumed to depend also on the influence of guidance molecules and on the density of branches at adjacent cortical locations:
+Changes in axon brtanching density are assumed to depend also on the influence of guidance molecules and on the density of branches at adjacent cortical locations:
 
@@
 \begin{equation} \label{eq:Karb2DExt_da}
-\frac{\partial a_i(\mb{x},t)}{\partial t} = \nabla\vcdot\left(D \nabla a_i(\mb{x},t)-a_i(\mb{x},t)\sum_{j=1}^{M} \gamma_{i,j}\nabla \rho_j(\mb{x})\right) - \frac{\partial c_i(\mb{x},t)}{\partial t} + \chi
+\frac{\partial a_i(\mb{x},t)}{\partial t} = \nabla\vcdot\left(D \nabla a_i(\mb{x},t)-a_i(\mb{x},t)\sum_{j=1}^{M} \gamma_{i,j}\nabla \rho_j(\mb{x})\right) - \frac{\partial c_i(\mb{x},t)}{\partial t} + \chi.
 \end{equation}
 
-The first term on the right describes the spatial gradient (indicated by $\nabla$) of a quantity referred to in models of heat diffusion as the `flux' (enclosed by parentheses).
-The central idea in such models is that the change in temperature (over time) at a given point on a surface is proportional to the gradient in the temperature (over space) at that location.
-Substituting heat for the number of branches, axons from a given thalamic projection are thus assumed to branch laterally across the cortical sheet at a rate given by the diffusion constant $D$.
+The first term on the right describes the spatial gradient (indicated by $\nabla\vcdot$) of a quantity referred to in models of heat diffusion as the `flux' (the term enclosed by parentheses).
+The central idea in heat diffusion models is that the change in temperature (over time) at a given location is proportional to the gradient in the temperature (over space) at that location.
+Substituting heat for the branching density, axons from a given thalamic projection are thus assumed to branch laterally across the cortical sheet at a rate given by the diffusion constant $D$.
+Lateral branching is augmented by a measure of the local alignment between the branching density and the gradient in the concentration of each of $M$ distinct molecular signals, $\rho(\mb{x})$.
+The influence of a given signal (indexed by $j$) on a given thalamic projection (indexed by $i$), is determined by the parameter $\gamma_{i,j}$.
+Interactions for which this value is positive encourage axon branching in the direction in which the concentration of the signal is increasing, whereas negative interactions encourage branching in the opposite direction.
+Hence the axons of a given thalamic projection are assumed to branch towards either higher or lower concentrations of each signalling molecule.
 
-Lateral branching is augmented by a measure of the local alignment between the density of branches and the gradient in the concentration of each of $M$ distinct molecular signals, $\rho(\mb{x})$.
-The influence of a given signal (indexed by $j$) on a given projection (indexed by $i$), is determined by the parameter $\gamma_i,j$.
-Interactions for which this value is positive encourage axon branching in the direction in which the concentration of the signal is increasing, whereas negative interactions encourage branching in the opposite direction.
-Hence the axons of a given projection are assumed to branch towards either higher or lower concentrations of each signalling molecule.
-
-The second term on the right of equation 2 represents an assumption that the number of axon branches decreases when the number of connections being made increases, and vice versa, i.e., that branches are required to form connections.
-Note that we introduce the term $\chi=0$ here, so that we might subsequently redefine $\chi$ in order to introduce the competition between thalamic projections that we found to be necessary for the emergence of a Voronoi tesselation.
+The second term on the right of equation 2 represents an assumption that the axon branching density decreases when the number of connections being made increases, and vice versa, i.e., that branches are required to form synaptic connections.
+Note that we introduce the term $\chi=0$ here, so that we might subsequently redefine $\chi$ to introduce the competition between thalamic projections that we will show to be necessary for the emergence of a Voronoi tesselation (see Results).
 
 Two points are important to emphasise here.
-First, notice that the only difference between thalamic projections is the $M$ interaction strengths ($\gamma$) that are chosen to define each projection.
-Hence any (reliable) differences between the size, shape, and location of simulated cortical fields are due to differences in the choice of these values only.
-Second, the formulation of the model is at this point entirely local, in the sense that simulating the dynamics does not require any information to be communicated from one cortical cell to any but its immediate neighbours.
+First, notice that the only difference between thalamic projections is the interaction strengths, $\gamma$.
+Hence any differences between the size, shape, and location of cortical fields that reliably emerge from the dynamics of the model must be due to differences in the choice of these values only.
+Second, the formulation of the model is at this point entirely local, in the sense that simulating the dynamics requires no information to be communicated from cortical locations to any but the immediately adjacent locations.
 
-In addition to the choice of initial conditions (for the fields $c(\mb{x},t=0)$ and $a(\mb{x},t=0)$) and the axon guidance fields $\rho(\mb{x})$, the free parameters of the model are the number of thalamocortical projections ($N$), constants $\alpha$, $\beta$, $k$, and $D$, and the $N\times M$ interaction strengths, $\gamma$.
+In addition to the choice of initial conditions (for the fields $c(\mb{x},t=0)$ and $a(\mb{x},t=0)$) and the axon guidance fields $\rho(\mb{x})$, the free parameters of the model are the number of thalamocortical projections, $N$, constants $\alpha$, $\beta$, $k$, and $D$, and the $N\times M$ interaction strengths, $\gamma$.
 
-Equations 1 and 2 were solved on a hexagonal lattice of grid elements represent the cortical sheet ($\mb{x}$), under the assumption that axon branches cannot extend beyond the boundary of the tissue, i.e., $\mb{J}_i(\mb{x},t) \bigg\rvert_{boundary} = 0$.
-We developed software tools for imposing arbitrary boundary shapes, specified in a scalable vector graphics (svg) file format, on the lattice.
-The dynamics were integrated using fourth-order Runge-Kutta for ordinary differential equation 1, and using a {\color{red} finite difference method} for the flux term (see Supplementary Materials).
+The dynamics of the model (equations 1 and 2) were solved on a hexagonal lattice of grid elements representing the cortical sheet, under the assumption that axon branches cannot extend beyond the boundary of the sheet, i.e., $\mb{J}_i(\mb{x},t) \bigg\rvert_{boundary} = 0$.
+We developed software tools for imposing arbitrary boundary shapes, specified in a scalable vector graphics (.svg) file format, on the lattice.
+The dynamics were numerically integrated using fourth-order Runge-Kutta, and using a {\color{red} finite difference method} to compute the flux term (see Supplementary Materials).
 

### 2019-10-02 paper/paper.tex
Wraps paper to 78 chars. Adds some analysis code

@@
 
-Understanding how genetic information and spontaneous pattern formation together give rise to the whisker barrels is an important challenge for developmental neurobiology.
-The barrel patterning that coalesces in layer IV has been described formally as a Voronoi tesselation, which is indicative of an underlying developmental process based on local competitive interactions.
-If pattern formation relies only on such intrinsic self-organising processes, how can the somatotopic order at the periphery be maintained in the projections from brainstem to thalamus and from thalamus to cortex?
-To investigate, we began by extending a computational model of neocortical arealization by Karbowski \& Ermentrout (2004; J.
-Comp.
-Neurosci.) to generate fields on a two-dimensional cortical sheet.
-Central to this model is an assumption that axonal branching and synaptogenesis are coupled, and that thalamic projections compete for cortical territory under the influence of molecular signalling gradients.
-We found that these processes alone are incapable of generating patterns of cortical fields that are more intricate than the patterns already present in molecular guidance signals.
-However, a full barrel-field pattern emerges under the guidance of as few as four signalling gradients when the model is extended to introduce competition between the branches of axons that originate from different thalamic barreloids.
-We show that the key requirements for the emergence of realistic (Voronoi tesselation) patterning are i) at each cortical location thalamocortical projections compete for a limited number of available connections, ii) at each location the axonal branching rate of each thalamocortical projection is reduced by the total branching density of the other projections, and iii) the density of axon branching for each thalamocortical projection is conserved.
-We demonstrate how a purely self-organising system can initialise a pattern which can then be faithfully replicated in multiple neural structures.
-We show that this minimal model of barrel self-organisation provides a parsimonious account of data from a range of classic studies in which intrinsic and extrinsic factors have been experimentally manipulated.
+Understanding how genetic information and spontaneous pattern formation
+together give rise to the whisker barrels is an important challenge for
+developmental neurobiology. The barrel patterning that coalesces in layer IV
+has been described formally as a Voronoi tesselation, which is indicative of
+an underlying developmental process based on local competitive
+interactions. If pattern formation relies only on such intrinsic
+self-organising processes, how can the somatotopic order at the periphery be
+maintained in the projections from brainstem to thalamus and from thalamus to
+cortex? To investigate, we began by extending a computational model of
+neocortical arealization by Karbowski \& Ermentrout (2004; J. Comp. Neurosci.)
+to generate fields on a two-dimensional cortical sheet. Central to this model
+is an assumption that axonal branching and synaptogenesis are coupled, and
+that thalamic projections compete for cortical territory under the influence
+of molecular signalling gradients. We found that these processes alone are
+incapable of generating patterns of cortical fields that are more intricate
+than the patterns already present in molecular guidance signals. However, a
+full barrel-field pattern emerges under the guidance of as few as four
+signalling gradients when the model is extended to introduce competition
+between the branches of axons that originate from different thalamic
+barreloids. We show that the key requirements for the emergence of realistic
+(Voronoi tesselation) patterning are i) at each cortical location
+thalamocortical projections compete for a limited number of available
+connections, ii) at each location the axonal branching rate of each
+thalamocortical projection is reduced by the total branching density of the
+other projections, and iii) the density of axon branching for each
+thalamocortical projection is conserved. We demonstrate how a purely
+self-organising system can initialise a pattern which can then be faithfully
+replicated in multiple neural structures. We show that this minimal model of
+barrel self-organisation provides a parsimonious account of data from a range
+of classic studies in which intrinsic and extrinsic factors have been
+experimentally manipulated.
 
@@
 
-Spatial patterns in neural connectivity provide clues about the constraints under which brains evolve and develop.
-Whether or not patterning governs information processing, models of the development of information processing circuits can be validated in terms of their ability to account for biological patterning (\citealp{Purves1992}; see also \citealp{Wilson2015,Bednar2016}).
-One of the most distinctive patterns of neural connectivity is found in the rodent barrel cortex (\citealp{Woolsey1948,Woolsey1970,Welker1974}, see \citealp{Fox2008}).
-Understanding how this particular pattern forms might shed light on the interplay of genetic and extrinsic factors that shape cortical circuits, and on a wealth of anatomical, physiological, and behavioural data obtained using this model system.
+Spatial patterns in neural connectivity provide clues about the constraints
+under which brains evolve and develop. Whether or not patterning governs
+information processing, models of the development of information processing
+circuits can be validated in terms of their ability to account for biological
+patterning (\citealp{Purves1992}; see also
+\citealp{Wilson2015,Bednar2016}). One of the most distinctive patterns of
+neural connectivity is found in the rodent barrel cortex
+(\citealp{Woolsey1948,Woolsey1970,Welker1974}, see
+\citealp{Fox2008}). Understanding how this particular pattern forms might shed
+light on the interplay of genetic and extrinsic factors that shape cortical
+circuits, and on a wealth of anatomical, physiological, and behavioural data
+obtained using this model system.
 
-The barrels are large cortical columns, identifiable from birth in layer 4 of rodent primary somatosensory cortex as dense clusters of thalamocortical axons, which are enclosed by borders a few neurons thick from postnatal day 3 \citep{Erzurumlu2012}.
-In the plane tangential to the cortical surface the barrels constitute a map of the arrangement of the whiskers on the face, such that the cells of adjacent barrels respond most strongly and most quickly to deflection of adjacent whiskers \citep{Armstrong-James1992}.
-The barrels reflect subcortical whisker maps comprising cellular aggregates known as barrelettes in the brainstem and barreloids in the thalamus \citep{Killackey1980}.
+The barrels are large cortical columns, identifiable from birth in layer 4 of
+rodent primary somatosensory cortex as dense clusters of thalamocortical
+axons, which are enclosed by borders a few neurons thick from postnatal day 3
+\citep{Erzurumlu2012}. In the plane tangential to the cortical surface the
+barrels constitute a map of the arrangement of the whiskers on the face, such
+that the cells of adjacent barrels respond most strongly and most quickly to
+deflection of adjacent whiskers \citep{Armstrong-James1992}. The barrels
+reflect subcortical whisker maps comprising cellular aggregates known as
+barrelettes in the brainstem and barreloids in the thalamus
+\citep{Killackey1980}.
 
-Each whisker and its corresponding barrelette, barreloid, and barrel can be identified by its position on a two-dimensional grid \citep{Zucker1969,Killackey1980,Loos1976}.
-The barrels are normally up to twice as long in the axis aligned to the grid ‘rows’ compared to the ‘arcs’, and the rows are visible from day 2 \citep{Rebsam2002} whereas the arcs become distinct by day 3 \citep{Erzurumlu2012,Rebsam2002}.
-In metabotropic glutamate receptor 5  (mGlu5) knockouts the barrel rows develop but the arcs do not (\citealp{Hannan2001} see also \citealp{Fox2008}), thus distinct processes likely guide development of the barrels with respect to the orthogonal grid axes.
+Each whisker and its corresponding barrelette, barreloid, and barrel can be
+identified by its position on a two-dimensional grid
+\citep{Zucker1969,Killackey1980,Loos1976}. The barrels are normally up to
+twice as long in the axis aligned to the grid ‘rows’ compared to the ‘arcs’,
+and the rows are visible from day 2 \citep{Rebsam2002} whereas the arcs become
+distinct by day 3 \citep{Erzurumlu2012,Rebsam2002}. In metabotropic glutamate
+receptor 5 (mGlu5) knockouts the barrel rows develop but the arcs do not
+(\citealp{Hannan2001} see also \citealp{Fox2008}), thus distinct processes
+likely guide development of the barrels with respect to the orthogonal grid
+axes.
 
-Barrel formation requires afferent input, driven by whisker stimulation and by thalamic calcium waves that propagate via mGlu5-receptor activation (\citealp{Anton2019}, see also \citealp{Anton2018}).
-Patterning also depends on a complex network of axon guidance processes, which includes interactions between guidance molecules like ephrin-A5 and A7 \citep{Miller2006} and between adhesion molecules like cadherin-6 and 8 \citep{Greig2013,Bishop2000}, and the binding of p75 to neurotropins (\citealp{Shimogori2005,Bishop2002}; see also \citealp{Dye2011a,Dye2011b}).
-This network is orchestrated by interactions between morphogens Fgf8 and Fgf17 and transcription factors Emx2, Pax6, Sp8, and Coup-tf1 \citep{Shimogori2005,Bishop2000}, which are expressed in gradients that mark orthogonal axes and can be manipulated to cause the barrels to stretch, shrink, shift, skew, and even duplicate \citep{Shimogori2005,Assimacopoulos2012,Borello2014,Sahara2007,Ypsilanti2016,Sur2005}.
+Barrel formation requires afferent input, driven by whisker stimulation and by
+thalamic calcium waves that propagate via mGlu5-receptor activation
+(\citealp{Anton2019}, see also \citealp{Anton2018}). Patterning also depends
+on a complex network of axon guidance processes, which includes interactions
+between guidance molecules like ephrin-A5 and A7 \citep{Miller2006} and
+between adhesion molecules like cadherin-6 and 8 \citep{Greig2013,Bishop2000},
+and the binding of p75 to neurotropins (\citealp{Shimogori2005,Bishop2002};
+see also \citealp{Dye2011a,Dye2011b}). This network is orchestrated by
+interactions between morphogens Fgf8 and Fgf17 and transcription factors Emx2,
+Pax6, Sp8, and Coup-tf1 \citep{Shimogori2005,Bishop2000}, which are expressed
+in gradients that mark orthogonal axes and can be manipulated to cause the
+barrels to stretch, shrink, shift, skew, and even duplicate
+\citep{Shimogori2005,Assimacopoulos2012,Borello2014,Sahara2007,Ypsilanti2016,Sur2005}.
 
-The boundaries of the barrels form polygonal shapes that together create a Voronoi tesselation across the cortical sheet, wherein all points enclosed by a given boundary are closer to its center than to the center of any other boundary \citep{Senft1991}.
-This patterning suggests that the barrel topology may be preserved in the initial projection of the thalamocortical axons into the cortex, and that a barrel forms by the lateral branching of the axons from their initial centres of mass, which ceases upon contact with the axons branching from adjacent centres.
-According to this model the barrel pattern could be entirely determined by the distribution of the initial centre points, which would itself be determined by guidance cues.
-However, the assumption of pre-arranged centres is difficult to resolve with the observation that thalamocortical axons arrive in the cortical plate as an undifferentiated bundle prior to barreloid formation (see \citealp{Erzurumlu2012} for a review).
+The boundaries of the barrels form polygonal shapes that together create a
+Voronoi tesselation across the cortical sheet, wherein all points enclosed by
+a given boundary are closer to its center than to the center of any other
+boundary \citep{Senft1991}. This patterning suggests that the barrel topology
+may be preserved in the initial projection of the thalamocortical axons into
+the cortex, and that a barrel forms by the lateral branching of the axons from
+their initial centres of mass, which ceases upon contact with the axons
+branching from adjacent centres. According to this model the barrel pattern
+could be entirely determined by the distribution of the initial centre points,
+which would itself be determined by guidance cues. However, the assumption of
+pre-arranged centres is difficult to resolve with the observation that
+thalamocortical axons arrive in the cortical plate as an undifferentiated
+bundle prior to barreloid formation (see \citealp{Erzurumlu2012} for a
+review).
 
-Alternatively, reaction-diffusion dynamics could generate a Voronoi-like tesselation in the cortex without the need for pre-arranged centres, by amplifying characteristic modes in the noisy initial distribution of bundled axon branches, as a net effect of short-range cooperative and longer-range competitive interactions between cells.
-According to this model the barrel pattern would be determined by the relative strength of these interactions and by the shape of the cortical field boundary.
-However, reaction-diffusion dynamics alone are unlikely to be able to account for the precise co-registration and topographic connectivity between emergent domains in thalamus and cortex, the non-uniform sizes and spacing of the barrels to form rows versus arcs, and the influence of gene expression gradients.
+Alternatively, reaction-diffusion dynamics could generate a Voronoi-like
+tesselation in the cortex without the need for pre-arranged centres, by
+amplifying characteristic modes in the noisy initial distribution of bundled
+axon branches, as a net effect of short-range cooperative and longer-range
+competitive interactions between cells. According to this model the barrel
+pattern would be determined by the relative strength of these interactions and
+by the shape of the cortical field boundary. However, reaction-diffusion
+dynamics alone are unlikely to be able to account for the precise
+co-registration and topographic connectivity between emergent domains in
+thalamus and cortex, the non-uniform sizes and spacing of the barrels to form
+rows versus arcs, and the influence of gene expression gradients.
 
-As theories of barrel patterning, the centre-point and reaction-diffusion models are not mutually exclusive.
-Pre-organised centres in the initial conditions could bias reaction-diffusion processes to generate the specific somatotopic arrangement of barrels more reliably, and similarly the mechanisms by which axons branch laterally from barrel centres may involve a tension between cooperative and competitive interactions.
-However, a proof that barrel patterning can in principle emerge within an undifferentiated bundle of thalamocortical axons, based only on local interactions between cells, would remove the need to invoke a global supervisor, i.e., to assume a separate stage and/or extrinsic mechanism for pre-organising thalamocortical connections.
+As theories of barrel patterning, the centre-point and reaction-diffusion
+models are not mutually exclusive. Pre-organised centres in the initial
+conditions could bias reaction-diffusion processes to generate the specific
+somatotopic arrangement of barrels more reliably, and similarly the mechanisms
+by which axons branch laterally from barrel centres may involve a tension
+between cooperative and competitive interactions. However, a proof that barrel
+patterning can in principle emerge within an undifferentiated bundle of
+thalamocortical axons, based only on local interactions between cells, would
+remove the need to invoke a global supervisor, i.e., to assume a separate
+stage and/or extrinsic mechanism for pre-organising thalamocortical
+connections.
 
-To this end, the aim of the present study is to establish whether it is possible for the observed arrangement of barrels to self-organise in a system governed by reaction-diffusion dynamics, under the guidance of molecular signalling gradients, and in the absence of pre-defined centre-points.
-Our approach is to represent in a computational model the minimal set of biological assumptions that might lead to the emergence of barrel patterning in simulation, and, if necessary, to add to this minimal set only assumptions that can be represented by the local exchange of information between adjacent cells.
-The emergence of barrel patterning under such conditions would serve as an existence proof of the plausibility of the hypothesis that barrel development is a self-organising process.
+To this end, the aim of the present study is to establish whether it is
+possible for the observed arrangement of barrels to self-organise in a system
+governed by reaction-diffusion dynamics, under the guidance of molecular
+signalling gradients, and in the absence of pre-defined centre-points. Our
+approach is to represent in a computational model the minimal set of
+biological assumptions that might lead to the emergence of barrel patterning
+in simulation, and, if necessary, to add to this minimal set only assumptions
+that can be represented by the local exchange of information between adjacent
+cells. The emergence of barrel patterning under such conditions would serve as
+an existence proof of the plausibility of the hypothesis that barrel
+development is a self-organising process.
 
@@
 
-\cite{Karbowski2004} developed what remains, to the best of our knowledge, the only dynamical model of cortical arealization capable of generating predictions about the influence of signalling molecules on the size, shape, and topological organization of emergent cortical fields.
-This model describes competition between thalamocortical projections for territory along a one-dimensional string of cortical cells, e.g., along an anterior-posterior axis.
-The equations of the model define how the density of connections, $c$, and the density of axon branches, $a$, interact at time $t$ at each cortical location $x$, for $N$ thalamocortical projections indexed by $i$.
-The Voronoi tesselation used to describe barrel patterning is a two-dimensional structure, hence we first extended the original model for simulation on a two-dimensional cortical sheet, and describe it here as such (i.e., substituting $x$ for the two-dimensional position vector $\mb{x}$ and substituting terms of the form $\frac{d}{dx}$ for the two-dimensional spatial operator $\nabla$).
+\cite{Karbowski2004} developed what remains, to the best of our knowledge, the
+only dynamical model of cortical arealization capable of generating
+predictions about the influence of signalling molecules on the size, shape,
+and topological organization of emergent cortical fields. This model describes
+competition between thalamocortical projections for territory along a
+one-dimensional string of cortical cells, e.g., along an anterior-posterior
+axis. The equations of the model define how the density of connections, $c$,
+and the density of axon branches, $a$, interact at time $t$ at each cortical
+location $x$, for $N$ thalamocortical projections indexed by $i$. The Voronoi
+tesselation used to describe barrel patterning is a two-dimensional structure,
+hence we first extended the original model for simulation on a two-dimensional
+cortical sheet, and describe it here as such (i.e., substituting $x$ for the
+two-dimensional position vector $\mb{x}$ and substituting terms of the form
+$\frac{d}{dx}$ for the two-dimensional spatial operator $\nabla$).
 
-The original model was derived from the assumption that the rates at which thalamocortical synapses form and thalamocortical axons branch are reciprocally coupled, resulting in an expression for the change in $c_i$ as a function of $a_i$, and for the change in $a_i$ as a function of $c_i$.
-Synaptogenesis is modelled  as
+The original model was derived from the assumption that the rates at which
+thalamocortical synapses form and thalamocortical axons branch are
+reciprocally coupled, resulting in an expression for the change in $c_i$ as a
+function of $a_i$, and for the change in $a_i$ as a function of
+$c_i$. Synaptogenesis is modelled as
 
@@
 
-Accordingly, when the final term evaluates to zero, the density of thalamocortical connections decays exponentially at a rate given by the decay constant $\alpha$.
-This occurs where the density of connections over all thalamic projections sum to one.
-Otherwise, the connection density is assumed to vary non-linearly ($k>1$) with changes in the axon branching density.
+Accordingly, when the final term evaluates to zero, the density of
+thalamocortical connections decays exponentially at a rate given by the decay
+constant $\alpha$. This occurs where the density of connections over all
+thalamic projections sum to one. Otherwise, the connection density is assumed
+to vary non-linearly ($k>1$) with changes in the axon branching density.
 
-Changes in axon brtanching density are assumed to depend also on the influence of guidance molecules and on the density of branches at adjacent cortical locations:
+Changes in axon branching density are assumed to depend also on the
+influence of guidance molecules and on the density of branches at
+adjacent cortical locations:
 
@@
 
-The first term on the right describes the spatial gradient (indicated by $\nabla\vcdot$) of a quantity referred to in models of heat diffusion as the `flux' (the term enclosed by parentheses).
-The central idea in heat diffusion models is that the change in temperature (over time) at a given location is proportional to the gradient in the temperature (over space) at that location.
-Substituting heat for the branching density, axons from a given thalamic projection are thus assumed to branch laterally across the cortical sheet at a rate given by the diffusion constant $D$.
-Lateral branching is augmented by a measure of the local alignment between the branching density and the gradient in the concentration of each of $M$ distinct molecular signals, $\rho(\mb{x})$.
-The influence of a given signal (indexed by $j$) on a given thalamic projection (indexed by $i$), is determined by the parameter $\gamma_{i,j}$.
-Interactions for which this value is positive encourage axon branching in the direction in which the concentration of the signal is increasing, whereas negative interactions encourage branching in the opposite direction.
-Hence the axons of a given thalamic projection are assumed to branch towards either higher or lower concentrations of each signalling molecule.
+The first term on the right describes the spatial gradient (indicated by
+$\nabla\vcdot$) of a quantity referred to in models of heat diffusion as the
+`flux' (the term enclosed by parentheses). The central idea in heat diffusion
+models is that the change in temperature (over time) at a given location is
+proportional to the gradient in the temperature (over space) at that
+location. Substituting heat for the branching density, axons from a given
+thalamic projection are thus assumed to branch laterally across the cortical
+sheet at a rate given by the diffusion constant $D$. Lateral branching is
+augmented by a measure of the local alignment between the branching density
+and the gradient in the concentration of each of $M$ distinct molecular
+signals, $\rho(\mb{x})$. The influence of a given signal (indexed by $j$) on a
+given thalamic projection (indexed by $i$), is determined by the parameter
+$\gamma_{i,j}$. Interactions for which this value is positive encourage axon
+branching in the direction in which the concentration of the signal is
+increasing, whereas negative interactions encourage branching in the opposite
+direction. Hence the axons of a given thalamic projection are assumed to
+branch towards either higher or lower concentrations of each signalling
+molecule.
 
-The second term on the right of equation 2 represents an assumption that the axon branching density decreases when the number of connections being made increases, and vice versa, i.e., that branches are required to form synaptic connections.
-Note that we introduce the term $\chi=0$ here, so that we might subsequently redefine $\chi$ to introduce the competition between thalamic projections that we will show to be necessary for the emergence of a Voronoi tesselation (see Results).
+The second term on the right of equation 2 represents an assumption that the
+axon branching density decreases when the number of connections being made
+increases, and vice versa, i.e., that branches are required to form synaptic
+connections. Note that we introduce the term $\chi=0$ here, so that we might
+subsequently redefine $\chi$ to introduce the competition between thalamic
+projections that we will show to be necessary for the emergence of a Voronoi
+tesselation (see Results).
 
-Two points are important to emphasise here.
-First, notice that the only difference between thalamic projections is the interaction strengths, $\gamma$.
-Hence any differences between the size, shape, and location of cortical fields that reliably emerge from the dynamics of the model must be due to differences in the choice of these values only.
-Second, the formulation of the model is at this point entirely local, in the sense that simulating the dynamics requires no information to be communicated from cortical locations to any but the immediately adjacent locations.
+Two points are important to emphasise here. First, notice that the only
+difference between thalamic projections is the interaction strengths,
+$\gamma$. Hence any differences between the size, shape, and location of
+cortical fields that reliably emerge from the dynamics of the model must be
+due to differences in the choice of these values only. Second, the formulation
+of the model is at this point entirely local, in the sense that simulating the
+dynamics requires no information to be communicated from cortical locations to
+any but the immediately adjacent locations.
 
-In addition to the choice of initial conditions (for the fields $c(\mb{x},t=0)$ and $a(\mb{x},t=0)$) and the axon guidance fields $\rho(\mb{x})$, the free parameters of the model are the number of thalamocortical projections, $N$, constants $\alpha$, $\beta$, $k$, and $D$, and the $N\times M$ interaction strengths, $\gamma$.
+In addition to the choice of initial conditions (for the fields
+$c(\mb{x},t=0)$ and $a(\mb{x},t=0)$) and the axon guidance fields
+$\rho(\mb{x})$, the free parameters of the model are the number of
+thalamocortical projections, $N$, constants $\alpha$, $\beta$, $k$, and $D$,
+and the $N\times M$ interaction strengths, $\gamma$.
 
-The dynamics of the model (equations 1 and 2) were solved on a hexagonal lattice of grid elements representing the cortical sheet, under the assumption that axon branches cannot extend beyond the boundary of the sheet, i.e., $\mb{J}_i(\mb{x},t) \bigg\rvert_{boundary} = 0$.
-We developed software tools for imposing arbitrary boundary shapes, specified in a scalable vector graphics (.svg) file format, on the lattice.
-The dynamics were numerically integrated using fourth-order Runge-Kutta, and using a {\color{red} finite difference method} to compute the flux term (see Supplementary Materials).
+The dynamics of the model (equations 1 and 2) were solved on a hexagonal
+lattice of grid elements representing the cortical sheet, under the assumption
+that axon branches cannot extend beyond the boundary of the sheet, i.e.,
+$\mb{J}_i(\mb{x},t) \bigg\rvert_{boundary} = 0$. We developed software tools
+for imposing arbitrary boundary shapes, specified in a scalable vector
+graphics (.svg) file format, on the lattice. The dynamics were numerically
+integrated using fourth-order Runge-Kutta, and using a {\color{red} finite
+  difference method} to compute the flux term (see Supplementary Materials).
 
 \section*{Results}
+
+\subsection*{}
+
+\subsection*{Relative interaction strengths}
+
+One possibility is to modify the strength of the guidance interactions of the
+two, orthogonal guidance gradients. In the simulation we modify the magnitude
+of the expression of each guidance molecule, although this is equivalent to,
+and could stand for a more or less vigourous movement of axons in response to
+un-modified guidance gradients.
 
@@
 
-We set out to test the hypothesis that barrel patterning may emerge via self-organising processes.
-We first extended an existing model of neocortical arealization to simulate pattern formation in two dimensions, inheriting from this model the assumption that axonal branching and synaptogenesis are coupled, and that thalamic projections compete for cortical territory under the influence of molecular signalling gradients.
-Realistic barrel patterning emerged only when terms were added to represent the assumption that  thalamocortical projections compete for a limited supply of connections with cortical cells, that their branching rates are reduced by the branching of other thalamocortical projections, and that the density of axon branching is conserved for each projection.
-The net effect of these local interactions is a Voronoi tesselation in which localised clusters of thalamic projections become somatotopically organised on the cortical sheet according to the relative values of coefficients specifying the rates at which axons are attracted and repelled by orthogonal pairs of molecular signal gradients.
-According to this model, the pattern of barrels is determined by the competitive and cooperative interactions between cells relative to the shape of the cortical field boundary.
-In simulation, we systematically investigated the origin of the asymmetry between the barrel rows and arcs, and identified {\color{red} XXXXX} as the most likely candidate for the elongation of barrel shapes along the rows.
+We set out to test the hypothesis that barrel patterning may emerge via
+self-organising processes. We first extended an existing model of neocortical
+arealization to simulate pattern formation in two dimensions, inheriting from
+this model the assumption that axonal branching and synaptogenesis are
+coupled, and that thalamic projections compete for cortical territory under
+the influence of molecular signalling gradients. Realistic barrel patterning
+emerged only when terms were added to represent the assumption that
+thalamocortical projections compete for a limited supply of connections with
+cortical cells, that their branching rates are reduced by the branching of
+other thalamocortical projections, and that the density of axon branching is
+conserved for each projection. The net effect of these local interactions is a
+Voronoi tesselation in which localised clusters of thalamic projections become
+somatotopically organised on the cortical sheet according to the relative
+values of coefficients specifying the rates at which axons are attracted and
+repelled by orthogonal pairs of molecular signal gradients. According to this
+model, the pattern of barrels is determined by the competitive and cooperative
+interactions between cells relative to the shape of the cortical field
+boundary. In simulation, we systematically investigated the origin of the
+asymmetry between the barrel rows and arcs, and identified {\color{red} XXXXX}
+as the most likely candidate for the elongation of barrel shapes along the
+rows.
 

### 2019-10-02 analysis/include/domcentres.py

@@ -1,0 +1,73 @@
+import numpy as np
+
+# Is @apositiveint a square number?
+def is_square(apositiveint):
+    x = apositiveint // 2
+    seen = set([x])
+    while x * x != apositiveint:
+        x = (x + (apositiveint // x)) // 2
+        if x in seen:
+            return False
+        seen.add(x)
+    return True
+
+# Do a linear fit, returning the square of the gradient and the sum of the residuals
+def do_fit (lx, ly):
+    A = np.vstack([lx, np.ones(len(lx))]).T
+    out = np.linalg.lstsq (A, ly, rcond=None)
+    m, c = out[0]
+    gradsq = (m*m)
+    resid = 0.0
+    if out[1].size > 0:
+        resid = out[1][0]
+    return gradsq, resid
+
+# What does this want to do? Return, for a given timestep (or for every timestep), the average verticalness of the lines of best fit to groups of centres.
+def domcentres_analyse (dc, isvert=True):
+
+    sos_gradients = []
+    summed_residuals = []
+    numt = np.shape(dc)[0]
+    N = int(np.shape(dc)[1])
+    # is N a square number?
+    if is_square(N) == False:
+        print ('domcentres_analyse current coded only for square arrays of domains (could be fixed)')
+        return sos_gradients, summed_residuals
+    n = int(np.sqrt(N))
+
+    tt = 0 # timepoint
+    while tt < numt:
+        # sum of the square of the gradient for all 5 in a column
+        sos_gradient = 0.0
+        # The sum of the residuals for 5 linear fits
+        sum_resid = 0.0
+        if isvert:
+            for i in range(0,N,n):
+                # Note: Swap around x and y, so that truly vertical lines will come out with m=0 (approx)
+                ly = dc[tt,i:n+i,0] # x
+                lx = dc[tt,i:n+i,1] # y
+                gradsq, resid = do_fit (lx, ly)
+                sos_gradient = sos_gradient + gradsq
+                sum_resid = sum_resid + resid
+        else:
+            for i in range(0,n):
+                # For n=5; if i=0, indices: 0,5,10,15,20
+                #          if i=1, indices: 1,6,11,16,21 etc.
+                lx = dc[tt,i:N-n+1+i:n,0] # x
+                ly = dc[tt,i:N-n+1+i:n,1] # y
+                gradsq, resid = do_fit (lx, ly)
+                #print ('gradsq: {0}, resid: {1}'.format(gradsq, resid))
+                sos_gradient = sos_gradient + gradsq
[…]

### 2019-10-02 analysis/include/load.py

@@ -21,21 +21,21 @@ def readDirichData (logdir):
     numtimes = len(files)
 
     # Create the time series to return. Values to be filled in from dirich_*.h5 file names
-    t = np.empty([numtimes], dtype=int)
+    t = np.zeros([numtimes], dtype=int)
 
     # To hold the honda values
-    honda = np.empty([numtimes], dtype=float)
+    honda = np.zeros([numtimes], dtype=float)
     # To hold the edgedeviation
-    edgedev = np.empty([numtimes], dtype=float)
+    edgedev = np.zeros([numtimes], dtype=float)
     # To hold a count of number of domains
-    numdoms = np.empty([numtimes], dtype=float)
+    numdoms = np.zeros([numtimes], dtype=float)
     # The area of the grid that is detected Dirichlet domains
-    domarea = np.empty([numtimes], dtype=float)
+    domarea = np.zeros([numtimes], dtype=float)
     # Coordinates of the putative centre of the domain. Need to open first file to get N.
     #print ('h5py open {0}'.format(files[0]))
     f = h5py.File(files[0], 'r')
     N = list(f['N'])[0]
-    domcentre = np.empty([numtimes, N, 2], dtype=float)
+    domcentre = np.zeros([numtimes, N, 2], dtype=float)
 
     fi = 0
     for filename in files:
@@ -102,7 +102,7 @@ def readSimDataFiles (logdir):
     print ('Have {0} files/timepoints which are: {1}'.format(numtimes,files))
 
     # Create the time series to return. Values to be filled in from c_*.h5 file names
-    t = np.empty([numtimes], dtype=int)
+    t = np.zeros([numtimes], dtype=int)
 
     # Count up how many c files we have in each time point once only:
     f = h5py.File(files[0], 'r')
@@ -118,13 +118,13 @@ def readSimDataFiles (logdir):
     # c1 etc as cols and spatial index as rows, all relating to a
     # single time point.
     print ('Creating empty 3d matrix of dims [{0},{1},{2}]'.format (numcs, numhexes, numtimes))
-    cmatrix = np.empty([numcs, numhexes, numtimes], dtype=float)
+    cmatrix = np.zeros([numcs, numhexes, numtimes], dtype=float)
     # There are as many 'a's as 'c's:
-    amatrix = np.empty([numcs, numhexes, numtimes], dtype=float)
+    amatrix = np.zeros([numcs, numhexes, numtimes], dtype=float)
[…]

### 2019-10-02 analysis/plot1.py

@@ -1,3 +1,6 @@
+# To access argv and also include the include dir
+import sys
+sys.path.insert (0, './include')
 import numpy as np
 # Import data loading code
 import load as ld
@@ -5,91 +8,10 @@ import load as ld
 import plot as pt
 import matplotlib
 import matplotlib.pyplot as plt
-# To access argv:
-import sys
 # Direct HDF5 access
 import h5py
-
-#
-# Functions
-#
-
-# Is @apositiveint a square number?
-def is_square(apositiveint):
-    x = apositiveint // 2
-    seen = set([x])
-    while x * x != apositiveint:
-        x = (x + (apositiveint // x)) // 2
-        if x in seen:
-            return False
-        seen.add(x)
-    return True
-
-# Do a linear fit, returning the square of the gradient and the sum of the residuals
-def do_fit (lx, ly):
-    A = np.vstack([lx, np.ones(len(lx))]).T
-    out = np.linalg.lstsq (A, ly, rcond=None)
-    m, c = out[0]
-    gradsq = (m*m)
-    resid = 0.0
-    if out[1].size > 0:
-        resid = out[1][0]
-    return gradsq, resid
-
-# What does this want to do? Return, for a given timestep (or for every timestep), the average verticalness of the lines of best fit to groups of centres.
-def domcentres_analyse (dc, isvert=True):
-
-    sos_gradients = []
-    summed_residuals = []
-    numt = np.shape(dc)[0]
-    N = int(np.shape(dc)[1])
-    # is N a square number?
-    if is_square(N) == False:
-        print ('domcentres_analyse current coded only for square arrays of domains (could be fixed)')
-        return sos_gradients, summed_residuals
-    n = int(np.sqrt(N))
-
-    tt = 0 # timepoint
-    while tt < numt:
-        # sum of the square of the gradient for all 5 in a column
-        sos_gradient = 0.0
-        # The sum of the residuals for 5 linear fits
-        sum_resid = 0.0
-        if isvert:
-            for i in range(0,N,n):
-                # Note: Swap around x and y, so that truly vertical lines will come out with m=0 (approx)
-                ly = dc[tt,i:n+i,0] # x
-                lx = dc[tt,i:n+i,1] # y
-                gradsq, resid = do_fit (lx, ly)
-                sos_gradient = sos_gradient + gradsq
[…]

### 2019-10-03 paper/paper.tex
A bit more WIP

@@
 \usepackage{times}
-\usepackage{color}
+\usepackage[usenames]{color}
 \usepackage{parskip}
@@
 
+\usepackage{lmodern}
+
 \usepackage[labelfont=bf,labelsep=period,justification=justified]{caption}
@@
 
-%\newcommand{\dvrg}{\nabla\vcdot\nabla}
+% Blue for comments
+\newcommand{\cmnt}[1]{\textcolor{blue}{#1}}
+% Code in sans serif
+\newcommand{\code}[1]{\textsf{#1}}
+% Divergence symbol (Del dot Del)
+\newcommand{\dvrg}{\nabla\vcdot\nabla}
+% Emphasis shortcut
+\newcommand{\e}{\emph}
+% Bold shortcut
+\newcommand{\bol}{\textbf}
+% Math bold
 \newcommand{\mb}[1]{\mathbf{#1}}
@@
 hence we first extended the original model for simulation on a two-dimensional
-cortical sheet, and describe it here as such (i.e., substituting $x$ for the
+cortical sheet, and describe it here as such (i.e., substituting $x$ with the
 two-dimensional position vector $\mb{x}$ and substituting terms of the form
-$\frac{d}{dx}$ for the two-dimensional spatial operator $\nabla$).
+$\frac{d}{dx}$ with the two-dimensional spatial gradient operator $\nabla$).
 
@@
 proportional to the gradient in the temperature (over space) at that
-location. Substituting heat for the branching density, axons from a given
+location. Substituting the branching density for heat, axons from a given
 thalamic projection are thus assumed to branch laterally across the cortical
@@
 
-\subsection*{}
+\cmnt{If we need to present the results of the 2-D system with $\chi=0$, then
+  we'll do that first, showing a 2-D version of the simulation obtained by
+  running \code{james0 configs/karb.json}}
 
-\subsection*{Relative interaction strengths}
+We first verified that the system given by Eqns. \ref{eq:Karb2DExt_dc} \&
+\ref{eq:Karb2DExt_da} described a two dimensional extension of the axon
+branching model in \citet{Karbowski2004}. We showed that in the presence of
+three guidance gradients varying only in one dimension (along the long axis of
+an elliptical domain), five thalamocortical axon types, each interacting
+uniquely with the three guidance gradients, would segregate into five separate
+regions. We investigated a number of configurations of the guidance gradients,
+varying their shape, $\rho(\mb{x})$, and number, $M$, and their effect on a
+varying number of thalamocortical projections, $N$. In general, we were not
+able to produce structure in the resulting patterns of the connections,
+$c(\mb{x},t\rightarrow\infty)$, which was not present in the guidance
+patterns.
+
+Complex structure can indeed be observed in protein expression for some genes,
+such as the molecules Bhlhb5 \citep{Joshi2008} and Lmo4
+\citep{Cederquist2013,Greig2013}. These expression patterns can therefore
+provide a theoretical mechanism for the guidance of axons into distinct
+regions, shifting the explanation of how such structures are generated into
+the realm of gene transcriptional interactions and mRNA diffusion/movement
+(\cmnt{Fixme: find a ref on structure in gene transcription}). However,
+complex, stable structure in gene expression is less commonly reported than
+homogeneous or simply varying expression, such as Pax6, Emx2, Coup-TF1 which
+display graded expression across the entire isocortex
+(\cmnt{REFS}). Furthermore, the complex structure observed in the whisker
+barrel field has not, to our knowledge, been linked to a complex set of gene
+expression patterns matching the scale of the barrels. To investigate how the
+whisker barrel structure might form in the presence of only two
+coordinate-giving morphogen gradients, we turned our attention to the
+competition for connections built into the model.
+
+There exists an element of competition in the model system specified by
+Eqns. \ref{eq:Karb2DExt_dc} \& \ref{eq:Karb2DExt_da}.
+
+\subsection*{Basic results}
+
+Formation of a grid of domains, with somatotopic ordering.
+
+Fig of domains.
+
+\subsection*{With surrounding large domains}
+
+Maybe, put this in to remind myself that I did this. It looks nice.
+
+\subsection*{Plucking experiments}
+
+Include results of plucking a single whisker, a row of whiskers and a
+column of whiskers.
+
+\subsection*{Any parameter sweeps necessary to show behaviour is
+  robust to changes in params}
+
+Go here.
+
+\subsection*{Manipulating Competition: Group competition stronger/weaker than inter-group competition}
+
+Just an idea - have two separate epsilons.
+
+\subsection*{Manipulating Guidance: Relative interaction strengths}
 
@@
 un-modified guidance gradients.
+
+\subsection*{Manipulating Guidance: Delayed onset in one axis}
+
+Another idea; begin with only one guidance gradient, or one strong and
+one very weak guidance gradient, and begin the development in this
+form, so as to encourage the formation of rows. Subsequently, turn on
+the second strong guidance signal.
+
+\subsection*{Manipulating Guidance: Different guidance profiles}
+
+Linear, circ linear, exponential, etc.
 

### 2019-10-03 paper/paper.tex
More work in progress

@@
 \begin{equation} \label{eq:Karb2DExt_dc}
-\frac{\partial c_i(\mb{x},t)}{\partial t} =-\alpha c_i(\mb{x},t) +\beta  \left(1 - \sum_{j=1}^{N} c_j(\mb{x}, t)\right)[a_i(\mb{x},t)]^k.
+\frac{\partial c_i(\mb{x},t)}{\partial t} =-\alpha c_i(\mb{x},t) +\beta  \left(1 - \sum_{i'=1}^{N} c_{i'}(\mb{x}, t)\right)[a_i(\mb{x},t)]^k.
 \end{equation}
@@
 constant $\alpha$. This occurs where the density of connections over all
-thalamic projections sum to one. Otherwise, the connection density is assumed
-to vary non-linearly ($k>1$) with changes in the axon branching density.
+thalamic projections sum to one; interpreted as a state in which all of the
+available synaptic receptors within the cortical dendritic trees are
+occupied. Otherwise, the connection density is assumed to vary non-linearly
+($k>1$) with the axon branching density.
 
@@
 There exists an element of competition in the model system specified by
-Eqns. \ref{eq:Karb2DExt_dc} \& \ref{eq:Karb2DExt_da}.
+Eqns. \ref{eq:Karb2DExt_dc} \& \ref{eq:Karb2DExt_da}. The term $\left(1 -
+\sum_{i=1}^{N} c_i(\mb{x}, t)\right)$ ensures that once the available synaptic
+receptors in a region have been fully occupied, say by thalamocortical types
+$i=2$ and $i=3$, there is no longer the possibility for type $i=1$ to form any
+synapses. This competition turns out to be relatively weak, because it does
+not strongly affect the branching density in that region; two or more TC axon
+types can be branching equally in the region, leading to roughly equal numbers
+of connections being made up until the connection limit is reached.
+
+To introduce an element of competition between the branching of
+thalamocortical axon types, we introduced the new term $\chi$ given by
+%
+\begin{equation} \label{eq:branch_comp}
+\chi(\mb{x}, t) = - \frac{\epsilon  a_i}{N-1} \sum_{i' \ne i}^{N} a_{i'}(\mb{x}, t)
+\end{equation}
+%
+which leads to a reduction in the rate of increase of branching density for
+axon type $i$ in regions where the branching density of \e{other} axon
+types is high.
+
+This introduces a new parameter, $\epsilon$, which governs the
+strength of the branching competition. The factor $1/(N-1)$ ensures that the
+branching competition strength scales with the number of TC axons in the
+system.
 
@@
 
-\section{Karbowski system with N TCs and M gradients}
+\section*{App 1: Karbowski system with N TCs and M gradients}
 %
@@
 
-\section{Karbowski system with N TCs and M gradients plus competition}
+\section{App 2: Karbowski system with N TCs and M gradients plus competition}
 

### 2019-10-03 analysis/include/load.py

@@ -35,7 +35,10 @@ def readDirichData (logdir):
     #print ('h5py open {0}'.format(files[0]))
     f = h5py.File(files[0], 'r')
     N = list(f['N'])[0]
+    # Domain centres using centroid method:
     domcentre = np.zeros([numtimes, N, 2], dtype=float)
+    # Dirichlet domain putative centres, Honda method:
+    dirichcentre = np.zeros([numtimes, N, 2], dtype=float)
 
     fi = 0
     for filename in files:
@@ -51,6 +54,7 @@ def readDirichData (logdir):
         honda[fi] = list(f['honda'])[0]
 
         # Coordinates, but need to re-cast them so that they have all ids. Or do I do that in the C++?
+        dom_id_list = list(f['reg_centroids_id'])
         domcentre[fi, :, 0] = list(f['reg_centroids_x'])
         domcentre[fi, :, 1] = list(f['reg_centroids_y'])
         #print ('x coords: {0}'.format(domcentre[fi, :, 0]))
@@ -69,12 +73,16 @@ def readDirichData (logdir):
             domarea[fi] += list(f[dom]['area'])[0]
             #print('P: {0}'.format(list(f[dom]['P'])))
             # store domcentre for each domain domcentre[fi]
+            #print ('{0}'.format(dom_id_list))
+            domid = dom_id_list.index(list(f[dom]['f'])[0])
+            dirichcentre[fi, domid, 0] = f[dom]['P'][0]
+            dirichcentre[fi, domid, 1] = f[dom]['P'][1]
 
         edgedev[fi] = edgedev[fi] / len(domset)
 
         fi = fi + 1
 
-    return [t, honda, edgedev, numdoms, domarea, domcentre]
+    return [t, honda, edgedev, numdoms, domarea, domcentre, dirichcentre]
 #
 # Load and read all the c_NNNNN.h5 files in logdir which contain a, c,
 # etc. Also reads x and y data which is stored in positions.h5.
@@ -83,6 +91,7 @@ def readSimDataFiles (logdir):
 
     # Take off any trailing directory slash
     logdir = logdir.rstrip ('/')
+    print ('logdir: {0}'.format(logdir))
 
     # Read x and y first
     pf = h5py.File(logdir+'/positions.h5', 'r')
@@ -141,7 +150,7 @@ def readSimDataFiles (logdir):
         klist = list(f.keys())
 
         for k in klist:
-            #print ('Key: {0}'.format(k))
[…]

### 2019-10-03 analysis/rel_guide_strengths.py

@@ -1,0 +1,119 @@
+# To access argv and also include the include dir
+import sys
+sys.path.insert (0, './include')
+
+import numpy as np
+# Import data loading code
+import load as ld
+# Import MY plotting code:
+import plot as pt
+import matplotlib
+import matplotlib.pyplot as plt
+# Direct HDF5 access
+import h5py
+# My domcentres linear fit code:
+import domcentres as dc
+
+# Set plotting defaults
+fs = 12
+fnt = {'family' : 'DejaVu Sans',
+       'weight' : 'regular',
+       'size'   : fs}
+matplotlib.rc('font', **fnt)
+
+# Which time step to analyse for
+keytime = 22
+
+uds = []
+lrs = []
+hondadeltas = []
+s_resids_v = []
+gfits_v = []
+s_resids_h = []
+gfits_h = []
+
+# Get target x/y hex to show trace for and the time step to show the
+# map for from the arguments:
+logdirbase = "../logs/rel_guide_strengths_r1_UD"
+
+do_maps = 1
+fcount = 0
+if do_maps:
+    # Create a figure
+    fs = 12
+    fnt = {'family' : 'DejaVu Sans',
+           'weight' : 'regular',
+           'size'   : fs}
+    matplotlib.rc('font', **fnt)
+    F0 = plt.figure (figsize=(21,18))
+    fcount = 0
+
+for ud in np.linspace(0,1,11):
+
+    logdirname = "{0}{1:.1f}".format (logdirbase, ud)
+    print ('logdirname: {0}'.format(logdirname))
+
+    # Load the dirichlet data, domcentres, etc
+    [t1, hondadelta, edgedev, numdoms, domarea, domcentres, ddoms] = ld.readDirichData (logdirname)
+
+    print ('hondadeltas: {0}'.format(hondadelta))
+
+    # The ID colour maps
+    if do_maps:
+        fcount = fcount + 1
+        # Read the data
+        (x, y, t, cmatrix, amatrix, nmatrix, idmatrix, tarea) = ld.readSimDataFiles (logdirname)
+        # Plot one plot only:
+        ax0 = F0.add_subplot (4, 3, fcount)
+        ax0.set_title('UD{0:0.1f} / LR{1:0.1f}'.format(ud, 1-ud))
+        # Plot regions:
[…]

### 2019-10-03 sim/james1.cpp

@@ -382,7 +382,7 @@ int main (int argc, char **argv)
     if (plot_dr && do_dirichlet_analysis) {
         winTitle = worldName + ": dr"; //4
         displays.push_back (morph::Gdisplay (win_width_contours, win_height_contours, 100, 1800, winTitle.c_str(),
-                                             rhoInit*0.7, thetaInit, phiInit, (windowId==0?0:displays[0].win)));
+                                             rhoInit*1.2, thetaInit, phiInit, (windowId==0?0:displays[0].win)));
         displays.back().resetDisplay (fix, eye, rot);
         displays.back().redrawDisplay();
         dr_id = windowId++;
@@ -698,7 +698,11 @@ int main (int argc, char **argv)
                 }
             }
             if (plot_dr && do_dirichlet_analysis) {
-                plt.scalarfields (displays[dr_id], RD.hg, RD.regions);
+                if (plot_contours) {
+                    plt.plot_contour_and_scalar (displays[dr_id], RD.hg, ctrs, RD.regions);
+                } else {
+                    plt.scalarfields (displays[dr_id], RD.hg, RD.regions);
+                }
             }
             // Then add:
             //plt.plot_dirichlet_boundaries (displays[n_id], RD.hg, vv);

### 2019-10-04 paper/paper.tex
Adds some potential subsection titles for the results section.

@@
 \usepackage{gensymb}
+
+\usepackage{titlesec}
+\titleformat*{\section}{\LARGE\bfseries}
+\titleformat*{\subsection}{\Large\bfseries}
+\titleformat*{\subsubsection}{\large\bfseries}
+\titleformat*{\paragraph}{\large\bfseries}
+\titleformat*{\subparagraph}{\large\bfseries}
 
@@
 
-\subsection*{Basic results}
+\subsection*{Competition is required to form somatotopically ordered barrels}
 
@@
 
-\subsection*{With surrounding large domains}
+\subsection*{The barrel field reacts to changes in the expression pattern of guidance molecules}
 
-Maybe, put this in to remind myself that I did this. It looks nice.
+To cover the reaction to changes in the guidance molecules caused by
+experiments like those by Shimogori \& Grove.
 
-\subsection*{Plucking experiments}
+Linear, circ linear, exponential, etc.
 
-Include results of plucking a single whisker, a row of whiskers and a
-column of whiskers.
+\subsection*{The barrel field reacts to simulated whisker plucking}
 
-\subsection*{Any parameter sweeps necessary to show behaviour is
-  robust to changes in params}
+A section on whisker plucking?
 
-Go here.
+%\subsubsection*{With surrounding large domains}
+%Maybe, put this in to remind myself that I did this. It looks nice.
 
-\subsection*{Manipulating Competition: Group competition stronger/weaker than inter-group competition}
+%\subsection*{Plucking experiments}
+%Include results of plucking a single whisker, a row of whiskers and a
+%column of whiskers.
+
+%\subsection*{Any parameter sweeps necessary to show behaviour is
+%  robust to changes in params}
+
+\subsection*{Extending the model to capture row and arc development}
+
+\subsubsection*{Manipulating Competition: Group competition stronger/weaker than inter-group competition}
 
@@
 
-\subsection*{Manipulating Guidance: Relative interaction strengths}
+\subsubsection*{Manipulating Guidance: Relative interaction strengths}
 
@@
 
-\subsection*{Manipulating Guidance: Delayed onset in one axis}
+\subsubsection*{Manipulating Guidance: Delayed onset in one axis}
 
@@
 the second strong guidance signal.
-
-\subsection*{Manipulating Guidance: Different guidance profiles}
-
-Linear, circ linear, exponential, etc.
 

### 2019-10-04 paper/paper.tex
Minor

--- 2019-10-03 paper/paper.tex
@@
 complex, stable structure in gene expression is less commonly reported than
-homogeneous or simply varying expression, such as Pax6, Emx2, Coup-TF1 which
+homogeneous or simply varying expression, such as Pax6, Emx2, Sp8 and Coup-tf1 which
 display graded expression across the entire isocortex

### 2019-10-11 paper/poster_equations.sh
Adds 'poster' equations for posters or slides

@@ -1,0 +1,18 @@
+#!/bin/bash
+
+# Use convert to snip out bits of the rendered latex into png images
+# suitable for Google docs
+
+# Basic Karbowski
+convert -verbose -density 450 -trim "poster_equations.pdf[0]" -quality 100 -antialias -flatten -crop 1850x700+950+1275 "WBEqns_01_to_03.jpg" &
+
+convert -verbose -density 450 -trim "poster_equations.pdf[0]" -quality 100 -antialias -flatten -crop 1850x245+950+2040 "WBEqns_04_to_04.jpg" &
+
+convert -verbose -density 450 -trim "poster_equations.pdf[0]" -quality 100 -antialias -flatten -crop 1850x245+950+2400 "WBEqns_05_to_05.jpg" &
+
+# With competition
+convert -verbose -density 450 -trim "poster_equations.pdf[0]" -quality 100 -antialias -flatten -crop 1850x800+950+2900 "WBEqns_06_to_08.jpg" &
+
+convert -verbose -density 450 -trim "poster_equations.pdf[0]" -quality 100 -antialias -flatten -crop 1850x225+950+3850 "WBEqns_09_to_09.jpg" &
+
+wait

### 2019-10-11 analysis/hondadelta.py
Minor tweak

@@ -57,7 +57,7 @@ F1 = plt.figure (figsize=(9,8))
 
 ax1 = F1.add_subplot(1,1,1)
 l1, = ax1.plot(t1, hondadelta, 'o', markersize=12, color=col.springgreen2, label='Honda $\delta$')
-l2, = ax1.plot((0,200000), (0.003, 0.003), '--', color=col.orange, linewidth=3, label='good\npattern')
+l2, = ax1.plot((0,200000), (0.003, 0.003), '--', color=col.orange, linewidth=3, label="'good'\npattern")
 #l2, = ax1.plot(t1, edgedev, 'o', label='Edge deviation')
 #l3, = ax1.plot(t1, domarea/totalarea[0], 'go', label='Domain area proportion')
 #l5, = ax1.plot(t1, s_resid, 's', label='Summed residuals to vert. line fits')

### 2019-10-29 misc/movie2.sh
Higher framerate movies

@@ -1,0 +1,10 @@
+#!/bin/bash
+
+if [ ! -z ${1} ]; then
+    pushd ${1}
+fi
+
+for i in contours maxval; do
+    touch ${i}2.mp4
+    ffmpeg -framerate 60 -i ${i}_%05d.png -vb 5MB -vcodec mpeg4 ${i}2.mp4 -y
+done

### 2019-10-29 paper/paper.tex
Insert the barreloid figure

@@
 
+\cmnt{Start from the thalamic barreloids, to generate the interaction
+  parameters}
+
+% Include figure ../boundaries/barreloids_vanderloos_graph.png
+\begin{figure*}
+\begin{center}
+\includegraphics[width=\textwidth]{../boundaries/barreloids_vanderloos_graph.png}
+\end{center}
+\vspace{-2em}
+\caption{Barreloid centres taken from \cite{van_der_loos_1976}, Fig.~2
+  are shown as black circles. It is assumed that there are at least
+  two linearly increasing, orthogonal signalling gradients in the
+  thalamus (side graphs - addme!), such that the location of each
+  barreloid within the graded thalamic signal field can differentiate
+  the efferent axons, with the result that axons from each barreloid
+  interact to varying degrees with cortical signalling
+  gradients. Coloured arrows show, for each barreloid, the magnitude
+  of the interaction that its thalamocortical projection will have
+  with the four signalling gradients in the cortex. The mappings from
+  barreloid location to interaction strength are linear.}
+\label{fig:barreloids}
+\end{figure*}
+
 \section*{Results}

### 2019-11-01 paper/paper.tex
Adds Zotero-generated bibfile

@@
 circuits can be validated in terms of their ability to account for biological
-patterning (\citealp{Purves1992}; see also
-\citealp{Wilson2015,Bednar2016}). One of the most distinctive patterns of
+patterning (\citealp{purves_iterated_1992}; see also
+\citealp{wilson_what_2015,bednar_cortical_2016}). One of the most distinctive patterns of
 neural connectivity is found in the rodent barrel cortex
-(\citealp{Woolsey1948,Woolsey1970,Welker1974}, see
+(\citealp{woolsey_pattern_1948,woolsey_structural_1970,welker_structure_1974}, see
 \citealp{Fox2008}). Understanding how this particular pattern forms might shed
@@
 between guidance molecules like ephrin-A5 and A7 \citep{Miller2006} and
-between adhesion molecules like cadherin-6 and 8 \citep{Greig2013,Bishop2000},
+between adhesion molecules like cadherin-6 and 8 \citep{greig_molecular_2013,Bishop2000},
 and the binding of p75 to neurotropins (\citealp{Shimogori2005,Bishop2002};
@@
 
-\begin{equation} \label{eq:Karb2DExt_dc}
+\begin{equation} \label{eq:dc}
 \frac{\partial c_i(\mb{x},t)}{\partial t} =-\alpha c_i(\mb{x},t) +\beta  \left(1 - \sum_{i'=1}^{N} c_{i'}(\mb{x}, t)\right)[a_i(\mb{x},t)]^k.
@@
 %
-\begin{equation} \label{eq:Karb2DExt_da}
+\begin{equation} \label{eq:da}
 \frac{\partial a_i(\mb{x},t)}{\partial t} = \nabla\vcdot\left(D \nabla a_i(\mb{x},t)-a_i(\mb{x},t)\sum_{j=1}^{M} \gamma_{i,j}\nabla \rho_j(\mb{x})\right) - \frac{\partial c_i(\mb{x},t)}{\partial t} + \chi.
@@
 
-\cmnt{Start from the thalamic barreloids, to generate the interaction
-  parameters}
-
-% Include figure ../boundaries/barreloids_vanderloos_graph.png
-\begin{figure*}
-\begin{center}
-\includegraphics[width=\textwidth]{../boundaries/barreloids_vanderloos_graph.png}
-\end{center}
-\vspace{-2em}
-\caption{Barreloid centres taken from \cite{van_der_loos_1976}, Fig.~2
-  are shown as black circles. It is assumed that there are at least
-  two linearly increasing, orthogonal signalling gradients in the
-  thalamus (side graphs - addme!), such that the location of each
-  barreloid within the graded thalamic signal field can differentiate
-  the efferent axons, with the result that axons from each barreloid
-  interact to varying degrees with cortical signalling
-  gradients. Coloured arrows show, for each barreloid, the magnitude
-  of the interaction that its thalamocortical projection will have
-  with the four signalling gradients in the cortex. The mappings from
-  barreloid location to interaction strength are linear.}
-\label{fig:barreloids}
-\end{figure*}
-
 \section*{Results}
@@
 
-We first verified that the system given by Eqns. \ref{eq:Karb2DExt_dc} \&
-\ref{eq:Karb2DExt_da} described a two dimensional extension of the axon
+We first verified that the system given by Eqns. \ref{eq:dc} \&
+\ref{eq:da} described a two dimensional extension of the axon
 branching model in \citet{Karbowski2004}. We showed that in the presence of
@@
 such as the molecules Bhlhb5 \citep{Joshi2008} and Lmo4
-\citep{Cederquist2013,Greig2013}. These expression patterns can therefore
+\citep{Cederquist2013,greig_molecular_2013}. These expression patterns can therefore
 provide a theoretical mechanism for the guidance of axons into distinct
-regions, shifting the explanation of how such structures are generated into
-the realm of gene transcriptional interactions and mRNA diffusion/movement
-(\cmnt{Fixme: find a ref on structure in gene transcription}). However,
-complex, stable structure in gene expression is less commonly reported than
-homogeneous or simply varying expression, such as Pax6, Emx2, Sp8 and Coup-tf1 which
-display graded expression across the entire isocortex
-(\cmnt{REFS}). Furthermore, the complex structure observed in the whisker
-barrel field has not, to our knowledge, been linked to a complex set of gene
-expression patterns matching the scale of the barrels. To investigate how the
-whisker barrel structure might form in the presence of only two
-coordinate-giving morphogen gradients, we turned our attention to the
+regions. The question of how such structures are generated becomes one of of
+how gene transcriptional interactions and mRNA diffusion and movement achieves
+this (\cmnt{Fixme: find a ref on structure in gene transcription}). However,
+complex and stable structure in gene expression data is commonly reported;
+monotonically varying expression patterns, such as those for Pax6, Emx2, Sp8
+and Coup-tf1 which display graded expression across the entire isocortex
+(\cmnt{REFS}) are more typical. Furthermore, the complex structure observed in
+the whisker barrel field has not, to our knowledge, been linked to a complex
+set of gene expression patterns matching the scale of the barrels. To
+investigate how the whisker barrel structure might form in the presence of
+only monotonically varying morphogen gradients, we turned our attention to the
 competition for connections built into the model.
 
+\subsection*{Competition is required to form somatotopically ordered barrels}
+
 There exists an element of competition in the model system specified by
-Eqns. \ref{eq:Karb2DExt_dc} \& \ref{eq:Karb2DExt_da}. The term $\left(1 -
-\sum_{i=1}^{N} c_i(\mb{x}, t)\right)$ ensures that once the available synaptic
+Eqns. \ref{eq:dc} \& \ref{eq:da}. The term $\left(1 -
+\sum_{i=1}^{N} c_i(\mb{x}, t)\right)$ in Eq.~\ref{eq:dc} ensures that once the available synaptic
 receptors in a region have been fully occupied, say by thalamocortical types
@@
 
-\subsection*{Competition is required to form somatotopically ordered barrels}
+\cmnt{Start from the thalamic barreloids, to generate the interaction
+  parameters}
 
-Formation of a grid of domains, with somatotopic ordering.
-
-Fig of domains.
+% Include figure ../boundaries/barreloids_vanderloos_graph.png
+\begin{figure*}
+\begin{center}
+\includegraphics[width=0.7\textwidth]{../boundaries/barreloids_vanderloos_graph.png}
+\end{center}
+%\vspace{-0em}
+\caption{Barreloid centres taken from \cite{van_der_loos_1976}, Fig.~2
+  are shown as black circles. It is assumed that there are at least
+  two linearly increasing, orthogonal signalling gradients in the
+  thalamus (side graphs - addme!), such that the location of each
+  barreloid within the graded thalamic signal field can differentiate
+  the efferent axons, with the result that axons from each barreloid
+  interact to varying degrees with cortical signalling
+  gradients. Coloured arrows show, for each barreloid, the magnitude
+  of the interaction that its thalamocortical projection will have
+  with the four signalling gradients in the cortex. The mappings from
+  barreloid location to interaction strength are linear.}
+\label{fig:barreloids}
+\end{figure*}
 
@@
 
-\bibliography{bibfile}
+\bibliography{BarrelEmerge}
 
@@
 %
-\begin{equation} \label{eq:Karb2DExt_dc}
+\begin{equation} \label{eq:dc2}
 \frac{\partial c_i(\mb{x},t)}{\partial t} = -\alpha c_i(\mb{x},t) + \beta n(\mb{x},t)
@@
 %
-\begin{equation} \label{eq:Karb2DExt_conserve}
+\begin{equation} \label{eq:conserve}
 n(\mb{x},t) = 1 - \sum_{i=1}^{N} c_i(\mb{x}, t)
@@
 %
-\begin{equation} \label{eq:Karb2DExt_da}
+\begin{equation} \label{eq:da2}
 \frac{\partial a_i(\mb{x},t)}{\partial t}

### 2019-11-01 paper/paper.tex
Finish updating citations

@@
 (\citealp{woolsey_pattern_1948,woolsey_structural_1970,welker_structure_1974}, see
-\citealp{Fox2008}). Understanding how this particular pattern forms might shed
+\citealp{fox_barrel_2008}). Understanding how this particular pattern forms might shed
 light on the interplay of genetic and extrinsic factors that shape cortical
@@
 axons, which are enclosed by borders a few neurons thick from postnatal day 3
-\citep{Erzurumlu2012}. In the plane tangential to the cortical surface the
+\citep{erzurumlu_development_2012}. In the plane tangential to the cortical surface the
 barrels constitute a map of the arrangement of the whiskers on the face, such
 that the cells of adjacent barrels respond most strongly and most quickly to
-deflection of adjacent whiskers \citep{Armstrong-James1992}. The barrels
+deflection of adjacent whiskers \citep{armstrong-james_flow_1992}. The barrels
 reflect subcortical whisker maps comprising cellular aggregates known as
 barrelettes in the brainstem and barreloids in the thalamus
-\citep{Killackey1980}.
+\citep{killackey_pattern_1980}.
 
@@
 identified by its position on a two-dimensional grid
-\citep{Zucker1969,Killackey1980,Loos1976}. The barrels are normally up to
+\citep{zucker_coding_1969,killackey_pattern_1980,van_der_loos_barreloids_1976}. The barrels are normally up to
 twice as long in the axis aligned to the grid ‘rows’ compared to the ‘arcs’,
-and the rows are visible from day 2 \citep{Rebsam2002} whereas the arcs become
-distinct by day 3 \citep{Erzurumlu2012,Rebsam2002}. In metabotropic glutamate
+and the rows are visible from day 2 \citep{rebsam_refinement_2002} whereas the arcs become
+distinct by day 3 \citep{erzurumlu_development_2012,rebsam_refinement_2002}. In metabotropic glutamate
 receptor 5 (mGlu5) knockouts the barrel rows develop but the arcs do not
-(\citealp{Hannan2001} see also \citealp{Fox2008}), thus distinct processes
+(\citealp{hannan_plc-1_2001} see also \citealp{fox_barrel_2008}), thus distinct processes
 likely guide development of the barrels with respect to the orthogonal grid
@@
 thalamic calcium waves that propagate via mGlu5-receptor activation
-(\citealp{Anton2019}, see also \citealp{Anton2018}). Patterning also depends
+(\citealp{anton-bolanos_prenatal_2019}, see also \citealp{anton-bolanos_developmental_2018}). Patterning also depends
 on a complex network of axon guidance processes, which includes interactions
-between guidance molecules like ephrin-A5 and A7 \citep{Miller2006} and
-between adhesion molecules like cadherin-6 and 8 \citep{greig_molecular_2013,Bishop2000},
-and the binding of p75 to neurotropins (\citealp{Shimogori2005,Bishop2002};
-see also \citealp{Dye2011a,Dye2011b}). This network is orchestrated by
+between guidance molecules like ephrin-A5 and A7 \citep{miller_epha7-ephrin-a5_2006} and
+between adhesion molecules like cadherin-6 and 8 \citep{greig_molecular_2013,bishop_regulation_2000},
+and the binding of p75 to neurotropins (\citealp{shimogori_fibroblast_2005,bishop_distinct_2002};
+see also \citealp{dye_lifespan_2011,dye_lifespan_2011-1}). This network is orchestrated by
 interactions between morphogens Fgf8 and Fgf17 and transcription factors Emx2,
-Pax6, Sp8, and Coup-tf1 \citep{Shimogori2005,Bishop2000}, which are expressed
+Pax6, Sp8, and Coup-tf1 \citep{shimogori_fibroblast_2005,bishop_regulation_2000}, which are expressed
 in gradients that mark orthogonal axes and can be manipulated to cause the
 barrels to stretch, shrink, shift, skew, and even duplicate
-\citep{Shimogori2005,Assimacopoulos2012,Borello2014,Sahara2007,Ypsilanti2016,Sur2005}.
+\citep{shimogori_fibroblast_2005,assimacopoulos_fibroblast_2012,borello_sp8_2014,sahara_sp8_2007,ypsilanti_transcriptional_2016,sur_patterning_2005}.
 
-The boundaries of the barrels form polygonal shapes that together create a
-Voronoi tesselation across the cortical sheet, wherein all points enclosed by
-a given boundary are closer to its center than to the center of any other
-boundary \citep{Senft1991}. This patterning suggests that the barrel topology
-may be preserved in the initial projection of the thalamocortical axons into
-the cortex, and that a barrel forms by the lateral branching of the axons from
-their initial centres of mass, which ceases upon contact with the axons
-branching from adjacent centres. According to this model the barrel pattern
-could be entirely determined by the distribution of the initial centre points,
-which would itself be determined by guidance cues. However, the assumption of
-pre-arranged centres is difficult to resolve with the observation that
-thalamocortical axons arrive in the cortical plate as an undifferentiated
-bundle prior to barreloid formation (see \citealp{Erzurumlu2012} for a
-review).
+The boundaries of the barrels form polygonal shapes that together
+create a Voronoi tesselation across the cortical sheet, wherein all
+points enclosed by a given boundary are closer to its center than to
+the center of any other boundary \citep{senft_mouse_1991}. This
+patterning suggests that the barrel topology may be preserved in the
+initial projection of the thalamocortical axons into the cortex, and
+that a barrel forms by the lateral branching of the axons from their
+initial centres of mass, which ceases upon contact with the axons
+branching from adjacent centres. According to this model the barrel
+pattern could be entirely determined by the distribution of the
+initial centre points, which would itself be determined by guidance
+cues. However, the assumption of pre-arranged centres is difficult to
+resolve with the observation that thalamocortical axons arrive in the
+cortical plate as an undifferentiated bundle prior to barreloid
+formation (see \citealp{erzurumlu_development_2012} for a review).
 
@@
 
-\cite{Karbowski2004} developed what remains, to the best of our knowledge, the
+\cite{karbowski_model_2004} developed what remains, to the best of our knowledge, the
 only dynamical model of cortical arealization capable of generating
@@
 \ref{eq:da} described a two dimensional extension of the axon
-branching model in \citet{Karbowski2004}. We showed that in the presence of
+branching model in \citet{karbowski_model_2004}. We showed that in the presence of
 three guidance gradients varying only in one dimension (along the long axis of
@@
 Complex structure can indeed be observed in protein expression for some genes,
-such as the molecules Bhlhb5 \citep{Joshi2008} and Lmo4
-\citep{Cederquist2013,greig_molecular_2013}. These expression patterns can therefore
+such as the molecules Bhlhb5 \citep{joshi_bhlhb5_2008} and Lmo4
+\citep{cederquist_lmo4_2013,greig_molecular_2013}. These expression patterns can therefore
 provide a theoretical mechanism for the guidance of axons into distinct
@@
 %\vspace{-0em}
-\caption{Barreloid centres taken from \cite{van_der_loos_1976}, Fig.~2
+\caption{Barreloid centres taken from \cite{van_der_loos_barreloids_1976}, Fig.~2
   are shown as black circles. It is assumed that there are at least

### 2019-11-12 paper/paper.tex
Adds a config and associated binary to set the competition with a logistic function

@@
 
-The first term on the right describes the spatial gradient (indicated by
+The first term on the right describes the divergence (indicated by
 $\nabla\vcdot$) of a quantity referred to in models of heat diffusion as the
@@
 
-There exists an element of competition in the model system specified by
-Eqns. \ref{eq:dc} \& \ref{eq:da}. The term $\left(1 -
-\sum_{i=1}^{N} c_i(\mb{x}, t)\right)$ in Eq.~\ref{eq:dc} ensures that once the available synaptic
-receptors in a region have been fully occupied, say by thalamocortical types
-$i=2$ and $i=3$, there is no longer the possibility for type $i=1$ to form any
-synapses. This competition turns out to be relatively weak, because it does
-not strongly affect the branching density in that region; two or more TC axon
-types can be branching equally in the region, leading to roughly equal numbers
-of connections being made up until the connection limit is reached.
+There exists an element of competition in the model system specified
+by Eqns. \ref{eq:dc} \& \ref{eq:da}. The term $\left(1 -
+\sum_{i=1}^{N} c_i(\mb{x}, t)\right)$ in Eq.~\ref{eq:dc} ensures that
+once the available synaptic receptors in a region have been fully
+occupied, say by thalamocortical type $i=1$, there is no
+longer the possibility for type $i=2$ to form any synapses. This
+competition turns out to be relatively weak, because it does not
+strongly affect the branching density in that region. In the model, it
+is possible for two or more TC axon types to have roughly equal
+branching density in some region, leading to roughly equal numbers of
+connections for each TC type being made,  up until the `hard' connection
+limit is reached.
 
-To introduce an element of competition between the branching of
-thalamocortical axon types, we introduced the new term $\chi$ given by
+To introduce competition between the \emph{branching} of thalamocortical axon
+types, we introduced the new term $\chi$ given by
 %
@@
 system.
+
+To increase the strength of the competition, we could take the power
+of the branching density:
+%
+\begin{equation} \label{eq:branch_comp_pow}
+\chi(\mb{x}, t) = - \frac{\epsilon  a_i}{N-1} \sum_{i' \ne i}^{N} (a_{i'})^l
+\end{equation}
+%
+with $l>1$. This competition term can quickly become excessively
+large, so to preserve the geometric shape of the function for small
+$a$, whilst representing the idea that the amount of branching
+competition has an upper bound, we pass the function through a
+logistic function.  This leads to the following branching equation:
+
+%
+\begin{equation} \label{eq:branch_comp_pow_logistic}
+\chi(\mb{x}, t) = - \frac{\epsilon a_i}{N-1} \sum_{i' \ne i}^{N}
+\frac{1}{1 + e^{-m (a_{i'})^l}} - \frac{1}{2}
+\end{equation}
+%
+$\epsilon$ represents the maximum value of the competition term for TC
+type $i$. The parameter $m$ controls the steepness of the function.
 

### 2019-11-12 sim/james1.cpp

@@ -253,6 +253,7 @@ int main (int argc, char **argv)
 
 #if defined DNCOMP || defined DNCOMP2
     const FLT l = root.get ("l", 1).asDouble();
+    const FLT m = root.get ("m", 1e-8).asDouble();
     const double E = root.get ("E", 0.1).asDouble();
     cout << "E is set to " << E << endl;
 #endif
@@ -480,6 +481,13 @@ int main (int argc, char **argv)
 #if defined DNCOMP || defined DNCOMP2
     cout << "Setting RD.l to " << l << endl;
     RD.l = l;
+    RD.m = m;
+#ifndef E_A_DIVN
+    if (E > 0.0) {
+        cout << "ERROR: You have E>0.0, but you are using a binary without the code to compute E div n" << endl;
+        exit (1);
+    }
+#endif
     RD.E = E;
 #endif
 

### 2019-11-12 sim/rd_james_dncomp.h

@@ -54,6 +54,8 @@ public:
     //@{
     //! The power to which a_j is raised for the inter-TC axon competition term.
     alignas(Flt) Flt l = 3.0;
+    //! The steepness of the logistic function
+    alignas(Flt) Flt m = 1e-8;
     //! epsilon_i parameters. axon competition parameter
     alignas(alignof(vector<Flt>))
     vector<Flt> epsilon;
@@ -161,6 +163,8 @@ public:
             // Multiply sum by 2D/3d^2 to give term1
             Flt term1 = this->twoDover3dd * thesum;
 
+            // This is required if E > 0:
+#ifdef E_A_DIVN
             // Term 1.1 is E a div(n)
             Flt term1_1 = this->E * fa[hi] * this->div_n[hi];
             if (isnan(term1_1)) {
@@ -175,6 +179,8 @@ public:
                 cerr << "term1_2 isnan" << endl;
                 exit (21);
             }
+#endif
+
 #if 0
             // 2. The (a div(g)) term.
             Flt term2 = fa[hi] * this->divg_over3d[i][hi];
@@ -197,7 +203,11 @@ public:
                 }
             }
 
-            this->divJ[i][hi] = term1 - term1_1 - term1_2  - term2 - term3;
+            this->divJ[i][hi] = term1
+#ifdef E_A_DIVN
+                - term1_1 - term1_2
+#endif
+                - term2 - term3;
         }
     }
 
@@ -222,15 +232,33 @@ public:
         milliseconds ms2 = duration_cast<milliseconds>(system_clock::now().time_since_epoch());
         this->codetimes.back().a_precompute += (ms2-ms1);
 
+#if defined BRANCHCOMP_LOGISTIC_FN
+        Flt one = static_cast<Flt>(1.0);
+        Flt mhalf = static_cast<Flt>(-0.5);
+#endif
+
         // Compute eps_all once only
         this->zero_vector_variable (this->eps_all);
         for (unsigned int j=0; j<this->N; ++j) {
-#pragma omp parallel for
+#if defined BRANCHCOMP_A_TO_POWER_L
+# pragma omp parallel for
             for (unsigned int h=0; h<this->nhex; ++h) {
-                eps_all[h] += static_cast<Flt>(pow (this->a[j][h], this->l));
+                Flt ea = static_cast<Flt>(pow (this->a[j][h], this->l));
+                //ea = ea > 1e6 ? 1e6 : ea;
+                eps_all[h] += ea;
             }
[…]

### 2019-11-13 paper/paper.tex
Adding figures for results section of paper

@@
 \citep{erzurumlu_development_2012}. In the plane tangential to the cortical surface the
-barrels constitute a map of the arrangement of the whiskers on the face, such
+barrels constitute a map of the arrangement of the whiskers on the
+face (see \citealp{yamakado_subdivision_1979}), such
 that the cells of adjacent barrels respond most strongly and most quickly to
@@
 identified by its position on a two-dimensional grid
-\citep{zucker_coding_1969,killackey_pattern_1980,van_der_loos_barreloids_1976}. The barrels are normally up to
+\citep{zucker_coding_1969,killackey_pattern_1980,van_der_loos_barreloids_1976,haidarliu_size_2001}. The barrels are normally up to
 twice as long in the axis aligned to the grid ‘rows’ compared to the ‘arcs’,
@@
 cortical plate as an undifferentiated bundle prior to barreloid
-formation (see \citealp{erzurumlu_development_2012} for a review).
+formation (\citealp{agmon_organized_1993}; see \citealp{erzurumlu_development_2012} for a review).
 
@@
 two-dimensional position vector $\mb{x}$ and substituting terms of the form
-$\frac{d}{dx}$ with the two-dimensional spatial gradient operator $\nabla$).
+$\frac{d}{dx}$ with the spatial gradient operator $\nabla$).
 
@@
 
-\cmnt{If we need to present the results of the 2-D system with $\chi=0$, then
-  we'll do that first, showing a 2-D version of the simulation obtained by
-  running \code{james0 configs/karb.json}}
+%\cmnt{If we need to present the results of the 2-D system with $\chi=0$, then
+%  we'll do that first, showing a 2-D version of the simulation obtained by
+%  running \code{james0 configs/karb.json}}
 
-We first verified that the system given by Eqns. \ref{eq:dc} \&
-\ref{eq:da} described a two dimensional extension of the axon
-branching model in \citet{karbowski_model_2004}. We showed that in the presence of
-three guidance gradients varying only in one dimension (along the long axis of
-an elliptical domain), five thalamocortical axon types, each interacting
-uniquely with the three guidance gradients, would segregate into five separate
-regions. We investigated a number of configurations of the guidance gradients,
-varying their shape, $\rho(\mb{x})$, and number, $M$, and their effect on a
-varying number of thalamocortical projections, $N$. In general, we were not
-able to produce structure in the resulting patterns of the connections,
-$c(\mb{x},t\rightarrow\infty)$, which was not present in the guidance
-patterns.
+We first verified that the system given by Eqns. \ref{eq:dc} \& \ref{eq:da}
+described a two dimensional extension of the axon branching model in
+\citet{karbowski_model_2004}. We showed that in the presence of three guidance
+gradients varying only in one dimension (along the long axis of an elliptical
+domain), five thalamocortical axon types, each interacting uniquely with the
+three guidance gradients, would segregate into five separate regions. We
+investigated a number of configurations of the guidance gradients, varying
+their shape, $\rho(\mb{x})$, and number, $M$, and their effect on a varying
+number of thalamocortical projections, $N$. In general, we were not able to
+producen a complexity of structure in the resulting patterns of the
+connections, $c(\mb{x},t\rightarrow\infty)$, which was not present in the
+guidance patterns.
 
@@
 such as the molecules Bhlhb5 \citep{joshi_bhlhb5_2008} and Lmo4
-\citep{cederquist_lmo4_2013,greig_molecular_2013}. These expression patterns can therefore
-provide a theoretical mechanism for the guidance of axons into distinct
-regions. The question of how such structures are generated becomes one of of
-how gene transcriptional interactions and mRNA diffusion and movement achieves
-this (\cmnt{Fixme: find a ref on structure in gene transcription}). However,
-complex and stable structure in gene expression data is commonly reported;
-monotonically varying expression patterns, such as those for Pax6, Emx2, Sp8
-and Coup-tf1 which display graded expression across the entire isocortex
-(\cmnt{REFS}) are more typical. Furthermore, the complex structure observed in
-the whisker barrel field has not, to our knowledge, been linked to a complex
-set of gene expression patterns matching the scale of the barrels. To
+\citep{cederquist_lmo4_2013,greig_molecular_2013}. These expression patterns
+can therefore provide a theoretical mechanism for the guidance of axons into
+distinct regions.
+%The question of how such structures are generated becomes
+%one of of how gene transcriptional interactions and mRNA diffusion and
+%movement achieves this (\cmnt{Fixme: find a ref on structure in gene
+%  transcription}).
+However, complex and stable structure in gene expression data is not commonly
+reported; monotonically varying expression patterns, such as those for Pax6,
+Emx2, Sp8 and Coup-tf1, which display graded expression across the entire
+isocortex (\cmnt{REFS}) are more typical. Furthermore, the complex structure
+observed in the whisker barrel field has not, to our knowledge, been linked to
+a set of gene expression patterns matching the scale of the barrels. To
 investigate how the whisker barrel structure might form in the presence of
@@
 
-\subsection*{Competition is required to form somatotopically ordered barrels}
+\subsection*{Branching competition is required to form somatotopically ordered barrels}
 
-There exists an element of competition in the model system specified
-by Eqns. \ref{eq:dc} \& \ref{eq:da}. The term $\left(1 -
-\sum_{i=1}^{N} c_i(\mb{x}, t)\right)$ in Eq.~\ref{eq:dc} ensures that
-once the available synaptic receptors in a region have been fully
-occupied, say by thalamocortical type $i=1$, there is no
-longer the possibility for type $i=2$ to form any synapses. This
-competition turns out to be relatively weak, because it does not
-strongly affect the branching density in that region. In the model, it
-is possible for two or more TC axon types to have roughly equal
+There exists an element of competition in the model system specified by
+Eqns. \ref{eq:dc} \& \ref{eq:da}. The term $\left(1 - \sum_{i=1}^{N}
+c_i(\mb{x}, t)\right)$ in Eq.~\ref{eq:dc} ensures that once the available
+synaptic receptors in a region have been fully occupied, say by
+thalamocortical type $i=1$, there is no longer the possibility for type $i=2$
+to form any synapses. This competition turns out to be relatively weak,
+because it does not strongly affect the branching density in that region. In
+the model, it is possible for two or more TC axon types to have roughly equal
 branching density in some region, leading to roughly equal numbers of
-connections for each TC type being made,  up until the `hard' connection
-limit is reached.
+connections for each TC type being made, up until the `hard' connection limit
+is reached.
 
@@
 which leads to a reduction in the rate of increase of branching density for
-axon type $i$ in regions where the branching density of \e{other} axon
-types is high.
+axon type $i$ in regions where the branching density of \e{other} axon types
+is high.
 
-This introduces a new parameter, $\epsilon$, which governs the
-strength of the branching competition. The factor $1/(N-1)$ ensures that the
-branching competition strength scales with the number of TC axons in the
-system.
+This introduces a new parameter, $\epsilon$, which governs the strength of the
+branching competition. The factor $1/(N-1)$ ensures that the branching
+competition strength scales with the number of TC axon types in the system.
+
+We first investigate the effect of the competition by examining a system of 25
+thalamocortical axon types simulated on a square domain. The axons interact
+with two orgthogonal signalling gradients, one of which increases from the top
+edge to the bottom edge, the other of which increases from the right edge to
+the left edge. The values of $\gamma$ are chosen as a grid.
+
+\begin{figure*}
+\begin{center}
+\includegraphics[width=\textwidth]{./Fig_25N2M_square/Fig_25N2M_square.png}
+\end{center}
+%\vspace{-0em}
+\caption{\textbf{(a)--(e)} Connection densities of 25 thalamocortical axons
+  simulated on a square domain. \textbf{(a)} Connection density at simulation
+  time t=0. The left image shows a map of the identity of the thalamocortical
+  axon type with the highest absolute connection density. Hexagonal elements
+  coloured the darkest blue have the highest connection density for TC type
+  $i=1$; those coloured the darkest red have maximal connection density in TC
+  type $i=25$. The right hand image of each pair shows the hexes for which the
+  connection density passes 0.5 for each TC type, giving a contour map. At
+  t=0, connection densities are randomly distributed. \textbf{(b)} t=0.14. The
+  grid-like organisation of the TC connections is beginning to emerge.
+  \textbf{(c)} t=0.2. The 5x5 grid is now clear in the left map. Branching
+  competition and differential interaction with the two guidance gradients is
+  causing the branching, and therefore the connection densities, to localise in
+  25 different locations. \textbf{(d)} t=0.64. Connection densities for each
+  TC type have increased sufficiently for the contours to form. In (d), the
+  mapping between colour and TC type $i$ is indicated. \textbf{(e)}
+  t=1.41. Contours are well formed. In (e), the $\gamma$ values for each TC
+  type, $i$, are shown. $\gamma_{i,1}$ gives the interaction with molecule `1'
+  [red map in (f)]; $\gamma_{i,2}$ is the interaction with molecule
+  `2'. \textbf{(f)} The two guidance gradients (`1' and `2') for this
+  simulation.}
+\label{fig:square}
+\end{figure*}
+
+\subsection*{The model can transfer realistic barrel patterns}
+
+In which we show that a realistic barrel pattern in the thalamus can be
+recreated in barrel cortex.
+
+\subsubsection*{Stronger branching competition}
+
+\cmnt{The realistic barrel pattern experiment is enhanced by stronger
+  branching competition. This section may be required here or elsewhere.}
 
@@
 %
-with $l>1$. This competition term can quickly become excessively
-large, so to preserve the geometric shape of the function for small
-$a$, whilst representing the idea that the amount of branching
-competition has an upper bound, we pass the function through a
-logistic function.  This leads to the following branching equation:
+with $l>1$. This competition term can quickly become excessively large, so to
+preserve the geometric shape of the function for small $a$, whilst
+representing the idea that the amount of branching competition has an upper
+bound, we pass the function through a logistic function.  This leads to the
+following branching equation:
 
@@
 %
-$\epsilon$ represents the maximum value of the competition term for TC
-type $i$. The parameter $m$ controls the steepness of the function.
+$\epsilon$ represents the maximum value of the competition term for TC type
+$i$. The parameters $l$ and $m$ control the steepness and shape of the
+function.
 
-\cmnt{Start from the thalamic barreloids, to generate the interaction
-  parameters}
+%%% End of stronger branching competition %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-% Include figure ../boundaries/barreloids_vanderloos_graph.png
 \begin{figure*}
 \begin{center}
-\includegraphics[width=0.7\textwidth]{../boundaries/barreloids_vanderloos_graph.png}
+\includegraphics[width=0.7\textwidth]{../boundaries/barreloids_haidarliu_graph.png}
 \end{center}
 %\vspace{-0em}
-\caption{Barreloid centres taken from \cite{van_der_loos_barreloids_1976}, Fig.~2
-  are shown as black circles. It is assumed that there are at least
-  two linearly increasing, orthogonal signalling gradients in the
-  thalamus (side graphs - addme!), such that the location of each
-  barreloid within the graded thalamic signal field can differentiate
-  the efferent axons, with the result that axons from each barreloid
-  interact to varying degrees with cortical signalling
-  gradients. Coloured arrows show, for each barreloid, the magnitude
-  of the interaction that its thalamocortical projection will have
-  with the four signalling gradients in the cortex. The mappings from
+\caption{Barreloid centres taken from \cite{haidarliu_size_2001} are shown as
+  black circles. It is assumed that there are two linearly increasing,
+  orthogonal signalling gradients (molecules A \& B) in the thalamus (side
+  graphs), such that the location of each barreloid within the graded thalamic
+  signal field can differentiate the efferent axons, with the result that
+  axons from each barreloid interact to varying degrees with cortical
+  signalling gradients. Coloured arrows show, for each barreloid, the
+  magnitude (and sign) of the interaction that its thalamocortical projection
+  will have with the signalling gradients in the cortex. The mappings from
   barreloid location to interaction strength are linear.}
 \label{fig:barreloids}
+\end{figure*}
+
+\begin{figure*}
+\begin{center}
+\includegraphics[width=\textwidth]{./realistic_barrels.png}
+\end{center}
+%\vspace{-0em}
+\caption{Realistic barrel pattern}
+\label{fig:barrels}
 \end{figure*}

### 2019-11-13 analysis/include/plot.py

@@ -71,7 +71,7 @@ def surface (dmatrix, x, y, ix, title):
 
 # ...with names
 def surface_withnames (dmatrix, x, y, ix, title, idnames, domcentres):
-    fs = 16
+    fs = 20
     fnt = {'family' : 'DejaVu Sans',
            'weight' : 'regular',
            'size'   : fs}
@@ -116,7 +116,7 @@ def surface_withnames (dmatrix, x, y, ix, title, idnames, domcentres):
         else:
             thechar = idn_arr[count]
 
-        f1.text (dc[0], dc[1], thechar, fontsize=14, verticalalignment='center', horizontalalignment='center', color=cmap(cidx))
+        f1.text (dc[0], dc[1], thechar, fontsize=24, verticalalignment='center', horizontalalignment='center', color=cmap(cidx))
 
         count = count + 1
     return f1

### 2019-11-13 analysis/plot_singlemap.py

@@ -12,6 +12,7 @@ import matplotlib.pyplot as plt
 import h5py
 # My domcentres linear fit code:
 import domcentres as dc
+import os
 
 # Set plotting defaults
 fs = 12
@@ -46,6 +47,7 @@ if ti == -1:
 print ('ti = {0}'.format(ti))
 f1 = pt.surface_withnames (idmatrix[:,ti], x, y, 0, idstring, idnames, domcentres[ti,:,:])
 
-plt.savefig('map.png')
+mapname = '{0}_{1}'.format(os.path.basename(logdirname), ti)
+plt.savefig(mapname)
 
 plt.show()

### 2019-11-13 paper/paper.sh

@@ -1,0 +1,6 @@
+#!/bin/bash
+
+xelatex paper.tex
+bibtex paper.aux
+xelatex paper.tex
+xelatex paper.tex

### 2019-11-15 paper/paper.tex
Friday work on the paper

@@
 
-We first verified that the system given by Eqns. \ref{eq:dc} \& \ref{eq:da}
+We first verified that the system given by Eqs.~\ref{eq:dc} \& \ref{eq:da}
 described a two dimensional extension of the axon branching model in
@@
 reported; monotonically varying expression patterns, such as those for Pax6,
-Emx2, Sp8 and Coup-tf1, which display graded expression across the entire
-isocortex (\cmnt{REFS}) are more typical. Furthermore, the complex structure
-observed in the whisker barrel field has not, to our knowledge, been linked to
-a set of gene expression patterns matching the scale of the barrels. To
-investigate how the whisker barrel structure might form in the presence of
-only monotonically varying morphogen gradients, we turned our attention to the
-competition for connections built into the model.
+Emx2 \citep{shimogori_fibroblast_2005,bishop_regulation_2000},
+Sp8~\citep{sahara_sp8_2007} and Coup-tf1~\citep{borello_sp8_2014}, which
+display graded expression across the entire isocortex are more
+typical. Furthermore, the complex structure observed in the whisker barrel
+field has not, to our knowledge, been linked to a set of gene expression
+patterns matching the scale of the barrels. To investigate how the whisker
+barrel structure might form in the presence of only monotonically varying
+morphogen gradients, we turned our attention to the competition for
+connections built into the model.
 
@@
 There exists an element of competition in the model system specified by
-Eqns. \ref{eq:dc} \& \ref{eq:da}. The term $\left(1 - \sum_{i=1}^{N}
+Eqs.~\ref{eq:dc} \& \ref{eq:da}. The term $\left(1 - \sum_{i=1}^{N}
 c_i(\mb{x}, t)\right)$ in Eq.~\ref{eq:dc} ensures that once the available
@@
 branching density in some region, leading to roughly equal numbers of
-connections for each TC type being made, up until the `hard' connection limit
+connections forming for each TC type, up until the `hard' connection limit
 is reached.
@@
 \begin{equation} \label{eq:branch_comp}
-\chi(\mb{x}, t) = - \frac{\epsilon  a_i}{N-1} \sum_{i' \ne i}^{N} a_{i'}(\mb{x}, t)
+\chi(\mb{x}, t) = - \frac{\epsilon  a_i(\mb{x}, t)}{N-1} \sum_{i' \ne i}^{N} a_{i'}(\mb{x}, t).
 \end{equation}
@@
 
-We first investigate the effect of the competition by examining a system of 25
-thalamocortical axon types simulated on a square domain. The axons interact
-with two orgthogonal signalling gradients, one of which increases from the top
-edge to the bottom edge, the other of which increases from the right edge to
-the left edge. The values of $\gamma$ are chosen as a grid.
+With the addition of the competition term, $\chi$, it is necessary to
+introduce a normalisation step, in order to prevent one TC type from
+out-competing every other TC type. Normalisation conforms to the idea that for
+a fixed number of afferent thalamocortical axons, the overall number of
+branches must have some upper limit; given the presumably finite resources
+available in each axon. We divisively normalize the branching density,
+dividing $a_i$ by the mean of $a_i$ per unit area; formalised as:
+%
+\begin{equation} \label{eq:norm}
+a_i(\mathbf{x}, t) = A \; \frac {a'_i(\mathbf{x}, t)} {\int_\Omega a'_i(\mathbf{x}, t) d\sigma}
+\end{equation}
+%
+where the prime symbols indicate the use of the intermediate value of $a_i$
+after it has been computed according to Eqs.~\ref{eq:dc}--\ref{eq:branch_comp}
+on the basis of the system state at time $t-{\delta}t$. $A$ is the area of the
+domain. This normalises $a_i(\mathbf{x}, t)$ to 1, integrated over the domain. To
+account for the possibility that there might be more axons afferent for TC
+type $1$ cf.~TC type $2$, we normalise to the integral of $a_i$ at time 0.
+%
+\begin{equation} \label{eq:norm_time0}
+\begin{split}
+a_i(\mathbf{x}, t) & =
+\frac{\int_\Omega  a'_i(\mathbf{x}, 0) d\sigma}{A} \; A \; \frac {a'_i(\mathbf{x}, t)} {\int_\Omega
+  a'_i(\mathbf{x}, t) d\sigma} \\
+%
+& = \int_\Omega  a'_i(\mathbf{x}, 0) d\sigma \; \frac {a'_i(\mathbf{x}, t)} {\int_\Omega
+  a'_i(\mathbf{x}, t) d\sigma}
+\end{split}
+\end{equation}
+%
+In the simulation, in which space is represented as $n_e$ discrete
+hexagonal tiles, the equivalent expression is:
+%
+\begin{equation} \label{eq:norm_discrete}
+{a_{i,p}}(t) = \textstyle \sum_j^{n_e} a_{i,j}(0) \; \frac {a'_{i,p}(t)} {\textstyle \sum_j^{n_e} a'_{i,j}(t) }
+\end{equation}
+%
+where $a_{i,j}(t)$ is the value of the variable $a_i$ in the $j^{\mathrm{th}}$
+hexagonal element at time t.
+
+We first investigated the effect of the competition by examining a system of
+25 thalamocortical axon types simulated on a square domain, as in
+Fig.~\ref{fig:square}. The hex to hex distance was 0.02 mm and the
+square's side length was 1.16 mm. The axons interacted with two orthogonal
+signalling gradients, one of which increased from the top edge to the bottom
+edge, the other of which had its gradient increasing from the right edge to
+the left edge [Fig.~\ref{fig:square}(f)]. For each thalamocortical type, $i$,
+two interaction parameters were set; $\gamma_{i,1}$ was the interaction with
+guidance molecule 1 and $\gamma_{i,2}$ was the interaction with molecule
+2. The branching of TC types with a negative interaction with molecule 2
+($\gamma_{i,2}<0$) will be pushed to the left; those with $\gamma_{i,2}>0$ are
+pushed to the right. TC types that interact negatively with molecule 1
+($\gamma_{i,1}<0$) are pushed up; those with $\gamma_{i,1}>0$ are pushed down.
+
+Fig.~\ref{fig:square}(a) shows the connection densities after the first 100
+simulation time steps (t=0.005 s). Each colour uniquely identifies the TC
+type, $i$ [the full mapping between index $i$ and the colour is shown in
+  Fig.~\ref{fig:square}(d)]. The left-hand map colours each hex according to
+which TC type has the highest connection density. At this early stage of the
+simulation, the colouring is somewhat random because the branching density is
+initialised stochastically ($a_{i,j}(0) = 0.2 + U([0,0.2])$), which results in
+a randomised pattern for ${c_i}^{max}$. The right-hand map of
+Fig.~\ref{fig:square}(a) colours those hexes which have $c_i \geq 0.5$ and an
+adjoining hex for which $c_i < 0.5$ according to the mapping shown in
+Fig.~\ref{fig:square}(d)]. Due to the initial stochasticity, there are many
+  such crossings of the threshold. The predominance of blue (for $i<10$)
+  results from the order in which the hexes have been rendered.
+Fig.~\ref{fig:square}(b) shows the simulation at t=0.14. \cmnt{CONTINUE FROM HERE.}
+
+% Later
+The full set of these interaction parameters are shown in
+Fig.~\ref{fig:square}(e). The values of $\gamma$ were chosen with the
+intention of forming a grid; the interactions for $i =$ 1, 5, 21 \& 25 (the
+corners) were strongest and the interaction parameters for $i=$13 (the centre)
+were zero. Fig.~\ref{fig:square}(d) shows the mapping between colour and TC
+type $i$ and can be used, in conjunction with Fig.~\ref{fig:square}(e), to see
+that $\gamma_{1,1}=-2$, $\gamma_{1,2}=-2$, $\gamma_{3,1}=0$,
+$\gamma_{3,2}=-2$, and so on.
 

### 2019-11-18 paper/paper.tex
Finished section on Fig 1

@@
 introduce a normalisation step, in order to prevent one TC type from
-out-competing every other TC type. Normalisation conforms to the idea that for
+out-competing every other TC type. Normalisation represents the idea that for
 a fixed number of afferent thalamocortical axons, the overall number of
-branches must have some upper limit; given the presumably finite resources
-available in each axon. We divisively normalize the branching density,
-dividing $a_i$ by the mean of $a_i$ per unit area; formalised as:
+branches must have some upper limit; given the finite resources available in
+each axon for building new processes. We divisively normalize the branching
+density, dividing it at each computational step by the mean branching density
+per unit area; formalised as:
 %
@@
 account for the possibility that there might be more axons afferent for TC
-type $1$ cf.~TC type $2$, we normalise to the integral of $a_i$ at time 0.
+type $1$ cf.~TC type $2$, we normalise to the integral of $a_i$ at time 0, so
+that Eq.~\ref{eq:norm} becomes:
 %
@@
 a_i(\mathbf{x}, t) & =
-\frac{\int_\Omega  a'_i(\mathbf{x}, 0) d\sigma}{A} \; A \; \frac {a'_i(\mathbf{x}, t)} {\int_\Omega
+\frac{\int_\Omega  a_i(\mathbf{x}, 0) d\sigma}{A} \; A \; \frac {a'_i(\mathbf{x}, t)} {\int_\Omega
   a'_i(\mathbf{x}, t) d\sigma} \\
 %
-& = \int_\Omega  a'_i(\mathbf{x}, 0) d\sigma \; \frac {a'_i(\mathbf{x}, t)} {\int_\Omega
+& = \int_\Omega  a_i(\mathbf{x}, 0) d\sigma \; \frac {a'_i(\mathbf{x}, t)} {\int_\Omega
   a'_i(\mathbf{x}, t) d\sigma}
+%\\
+%& = a_0 \frac {a'_i(\mathbf{x}, t)} {\int_\Omega
+%  a'_i(\mathbf{x}, t) d\sigma}
 \end{split}
@@
 \begin{equation} \label{eq:norm_discrete}
-{a_{i,p}}(t) = \textstyle \sum_j^{n_e} a_{i,j}(0) \; \frac {a'_{i,p}(t)} {\textstyle \sum_j^{n_e} a'_{i,j}(t) }
+  {a_{i,p}}(t) = \textstyle \sum_j^{n_e} a_{i,j}(0) \; \frac {a'_{i,p}(t)}
+  {\textstyle \sum_j^{n_e} a'_{i,j}(t) }
 \end{equation}
@@
 25 thalamocortical axon types simulated on a square domain, as in
-Fig.~\ref{fig:square}. The hex to hex distance was 0.02 mm and the
-square's side length was 1.16 mm. The axons interacted with two orthogonal
-signalling gradients, one of which increased from the top edge to the bottom
-edge, the other of which had its gradient increasing from the right edge to
-the left edge [Fig.~\ref{fig:square}(f)]. For each thalamocortical type, $i$,
-two interaction parameters were set; $\gamma_{i,1}$ was the interaction with
+Fig.~\ref{fig:sq}, to determine if a complex pattern could emerge. The hex
+to hex distance was 0.02 mm and the square's side length was 1.16 mm (3919 hex
+elements). The axons interacted with two orthogonal signalling gradients, one
+of which increased from the top edge to the bottom edge, the other of which
+had its expression increasing from the right edge to the left edge
+[Fig.~\ref{fig:sq}(f)]. For each thalamocortical type, $i$, two
+interaction parameters were set; $\gamma_{i,1}$ was the interaction with
 guidance molecule 1 and $\gamma_{i,2}$ was the interaction with molecule
 2. The branching of TC types with a negative interaction with molecule 2
-($\gamma_{i,2}<0$) will be pushed to the left; those with $\gamma_{i,2}>0$ are
-pushed to the right. TC types that interact negatively with molecule 1
-($\gamma_{i,1}<0$) are pushed up; those with $\gamma_{i,1}>0$ are pushed down.
-
-Fig.~\ref{fig:square}(a) shows the connection densities after the first 100
-simulation time steps (t=0.005 s). Each colour uniquely identifies the TC
-type, $i$ [the full mapping between index $i$ and the colour is shown in
-  Fig.~\ref{fig:square}(d)]. The left-hand map colours each hex according to
-which TC type has the highest connection density. At this early stage of the
-simulation, the colouring is somewhat random because the branching density is
-initialised stochastically ($a_{i,j}(0) = 0.2 + U([0,0.2])$), which results in
-a randomised pattern for ${c_i}^{max}$. The right-hand map of
-Fig.~\ref{fig:square}(a) colours those hexes which have $c_i \geq 0.5$ and an
-adjoining hex for which $c_i < 0.5$ according to the mapping shown in
-Fig.~\ref{fig:square}(d)]. Due to the initial stochasticity, there are many
-  such crossings of the threshold. The predominance of blue (for $i<10$)
-  results from the order in which the hexes have been rendered.
-Fig.~\ref{fig:square}(b) shows the simulation at t=0.14. \cmnt{CONTINUE FROM HERE.}
-
-% Later
-The full set of these interaction parameters are shown in
-Fig.~\ref{fig:square}(e). The values of $\gamma$ were chosen with the
-intention of forming a grid; the interactions for $i =$ 1, 5, 21 \& 25 (the
-corners) were strongest and the interaction parameters for $i=$13 (the centre)
-were zero. Fig.~\ref{fig:square}(d) shows the mapping between colour and TC
-type $i$ and can be used, in conjunction with Fig.~\ref{fig:square}(e), to see
-that $\gamma_{1,1}=-2$, $\gamma_{1,2}=-2$, $\gamma_{3,1}=0$,
-$\gamma_{3,2}=-2$, and so on.
+($\gamma_{i,2}<0$) were pushed to the left; those with $\gamma_{i,2}>0$ were
+pushed to the right. TC types that interacted negatively with molecule 1
+($\gamma_{i,1}<0$) were pushed up; those with $\gamma_{i,1}>0$ were pushed
+down.
 
@@
   simulated on a square domain. \textbf{(a)} Connection density at simulation
-  time t=0. The left image shows a map of the identity of the thalamocortical
+  time t=0.005. The left image shows a map of the identity of the thalamocortical
   axon type with the highest absolute connection density. Hexagonal elements
@@
   connection density passes 0.5 for each TC type, giving a contour map. At
-  t=0, connection densities are randomly distributed. \textbf{(b)} t=0.14. The
+  t=0, connection densities are randomly distributed. \textbf{(b)} t=0.18. The
   grid-like organisation of the TC connections is beginning to emerge.
-  \textbf{(c)} t=0.2. The 5x5 grid is now clear in the left map. Branching
+  \textbf{(c)} t=0.25. The 5x5 grid is now clear in the left map. Branching
   competition and differential interaction with the two guidance gradients is
-  causing the branching, and therefore the connection densities, to localise in
-  25 different locations. \textbf{(d)} t=0.64. Connection densities for each
-  TC type have increased sufficiently for the contours to form. In (d), the
-  mapping between colour and TC type $i$ is indicated. \textbf{(e)}
-  t=1.41. Contours are well formed. In (e), the $\gamma$ values for each TC
+  causing the branching, and therefore the connection densities, to localise
+  in 25 different locations. \textbf{(d)} t=0.81. Connection densities for
+  each TC type have increased sufficiently for the contours to form. In (d),
+  the mapping between colour and TC type $i$ is indicated. \textbf{(e)}
+  t=1.77. Contours are well formed. In (e), the $\gamma$ values for each TC
   type, $i$, are shown. $\gamma_{i,1}$ gives the interaction with molecule `1'
   [red map in (f)]; $\gamma_{i,2}$ is the interaction with molecule
-  `2'. \textbf{(f)} The two guidance gradients (`1' and `2') for this
-  simulation.}
-\label{fig:square}
+  `2'. \textbf{(f)} Expression maps for the two guidance molecules (`1' and
+  `2') for this simulation. Colour saturation indicates expression intensity,
+  which for each molecule varies linearly in one direction.}
+\label{fig:sq}
 \end{figure*}
+
+Fig.~\ref{fig:sq}(a) shows the connection densities after the first 100
+simulation time steps (t=0.005 s). Each colour uniquely identifies the TC
+type, $i$ [the full mapping between index $i$ and the colour is shown in
+  Fig.~\ref{fig:sq}(d)]. The left-hand map colours each hex according to
+which TC type has the highest connection density. At this early stage of the
+simulation, the colouring is somewhat random because the branching density was
+initialised stochastically ($a_{i,j}(0) = 0.2 + U([0,0.2])$), which resulted
+in a randomised pattern for ${c_i}^{max}$. The right-hand map of
+Fig.~\ref{fig:sq}(a) colours those hexes which have $c_i \geq 0.5$ and an
+adjoining hex for which $c_i < 0.5$ according to the mapping shown in
+Fig.~\ref{fig:sq}(d)]. Due to the initial stochasticity, there are many
+  such crossings of the threshold. The predominance of blue (for $i<10$)
+  results from the order in which the hexes were rendered, with blue hexes
+  being coloured last.  Fig.~\ref{fig:sq}(b) shows the simulation at
+  t=0.14.
+%
+%
+Fig.~\ref{fig:sq}(b) shows the connection density after 0.18 s of simulation
+time (3600 steps of 5x10$^{-5}$ s). The initial randomness in the pattern has
+disappearing, to be replaced by coalesced regions in which each TC type $i$ is
+maximal.  There is a clear blue/red division from left to right. The threshold
+plot indicates that the guidance molecules have pushed the branching towards
+the corners of the square for the 4 TC types with the strongest interactions
+with the guidance molecules; these exceed the threshold of 0.5. For the other
+21 TC types, the connection densities are not yet sufficiently localised to
+cross the threshold.
+%
+0.08 s of simulation time later, in Fig.~\ref{fig:sq}(c), the 5x5 grid is
+clearly visible in the maximal value map, though the localisation of each
+connection density map is still not sufficient for the contour map to emerge.
+%
+The contour map emerges by 0.81 s and is well formed by 1.77 s
+[Figs.~\ref{fig:sq}(d) \& (e)]. Fig.~\ref{fig:sq}(e) shows the interaction
+parameters for each TC type ($\gamma$). The values of $\gamma$ were chosen
+with the intention of forming a grid; the interactions for $i =$ 1, 5, 21 \&
+25 (the corners) were strongest and the interaction parameters for $i=$13 (the
+centre) were zero. Fig.~\ref{fig:sq}(d) shows the mapping between colour and
+TC type $i$ and can be used, in conjunction with Fig.~\ref{fig:sq}(e), to see
+that $\gamma_{1,1}=-2$, $\gamma_{1,2}=-2$, $\gamma_{3,1}=0$,
+$\gamma_{3,2}=-2$, and so on.
+
+The simulation shown in Fig.~\ref{fig:sq} demonstrates that complex structure
+can form in a system with just two guidance gradients \cmnt{(provided that
+  interactions with each gradient occur in both senses; positive and negative)}.
 

### 2019-11-19 paper/briefpaper/brief.tex
Housekeeping, creating longpaper and briefpaper folders.

--- 2019-11-13 paper/paper.tex
@@
-% Whisker barrel generation paper equations, for pasting into Google
-% docs draft.
+% Brief report style paper
 
+% It's about how a somatotopically ordered pattern can be
+% re-established in the cortex based on information carried with
+% thalamocortical axons afferent from the ventrobasal thalamus, but
+% without specifying absolute positions as centre-points for the
+% barrels. Everything in the paper should be tailored towards arguing
+% this idea.
+
+\documentclass[9pt,twocolumn,twoside,lineno]{pnas-new}
+\templatetype{pnasresearcharticle}
+
+\title{Self-organization of whisker barrels in simulation requires competition between thalamocortical axons}
 % While writing, don't stop for errors
@@
 
-\documentclass[a4paper,11pt]{article}
-\usepackage{amsmath,amssymb,mathtools}
-\usepackage{enumerate}
-\usepackage{times}
-\usepackage[usenames]{color}
-\usepackage{parskip}
-\usepackage{afterpage}
-\usepackage{setspace}
-\usepackage[modulo]{lineno}
-%\usepackage{xltxtra}
-%\usepackage{xunicode}
-\usepackage{gensymb}
+% Use letters for affiliations, numbers to show equal authorship (if applicable) and to indicate the corresponding author
+\author[a,1]{Sebastian~S.~James}
+\author[a]{Stuart~P.~Wilson}
 
-\usepackage{titlesec}
-\titleformat*{\section}{\LARGE\bfseries}
-\titleformat*{\subsection}{\Large\bfseries}
-\titleformat*{\subsubsection}{\large\bfseries}
-\titleformat*{\paragraph}{\large\bfseries}
-\titleformat*{\subparagraph}{\large\bfseries}
+\affil[a]{Department of Psychology, The University of Sheffield, Sheffield, United Kingdom.}
 
-\setcounter{tocdepth}{3}
-\usepackage{graphicx}
-\usepackage{comment}
-\usepackage{textcomp}
-\usepackage{natbib}
+% $^{*}$Corresponding author email: Seb.James@Sheffield.ac.uk\vspace{1em}\\
 
-\usepackage{lmodern}
+% Please give the surname of the lead author for the running footer
+\leadauthor{James}
 
-\usepackage[labelfont=bf,labelsep=period,justification=justified]{caption}
+% Please add a significance statement to explain the relevance of your work
+\significancestatement{The mystacial whiskers of the rat are a complex, ordered set of structures whose pattern is repeated several times in separate locations in the brain.
+We contribute a mechanism by which their spatially ordered pattern can be recreated in a self-organising process which requires minimal genetic pre-determination.}
 
-\doublespacing
-\topmargin 0.0cm
-\oddsidemargin 0.5cm
-\evensidemargin 0.5cm
-\textwidth 16cm
-\textheight 21cm
-\makeatletter
-\makeatother
-\date{}
-\pagestyle{plain}
+\correspondingauthor{\textsuperscript{1}To whom correspondence should be addressed. E-mail: seb.james\@sheffield.ac.uk}
 
-% Blue for comments
-\newcommand{\cmnt}[1]{\textcolor{blue}{#1}}
-% Code in sans serif
-\newcommand{\code}[1]{\textsf{#1}}
-% Divergence symbol (Del dot Del)
-\newcommand{\dvrg}{\nabla\vcdot\nabla}
-% Emphasis shortcut
-\newcommand{\e}{\emph}
-% Bold shortcut
-\newcommand{\bol}{\textbf}
-% Math bold
-\newcommand{\mb}[1]{\mathbf{#1}}
-\makeatletter
-\newcommand*\vcdot{\mathpalette\vcdot@{.35}}
-\newcommand*\vcdot@[2]{\mathbin{\vcenter{\hbox{\scalebox{#2}{$\m@th#1\bullet$}}}}}
-\makeatother
+% At least three keywords are required at submission. Please provide three to five keywords, separated by the pipe symbol.
+\keywords{Keyword 1 $|$ Keyword 2 $|$ Keyword 3 $|$ ...}
 
-%\bibliographystyle{plos2015}
-\bibliographystyle{apalike}
-
-\begin{document}
-%\begin{center}
-\begin{flushleft}
-{\huge
-\textbf{Self-organization of whisker barrels in simulation requires competition between thalamocortical axons}
- \vspace{5em}
-}
-%\end{center}
-
-Sebastian~S.~James$^{1*}$, Stuart~P.~Wilson$^{1}$ \vspace{1em}\\
-$^{1}$Department of Psychology, The University of Sheffield, Sheffield, United Kingdom.\vspace{1em}\\
-$^{*}$Corresponding author email: Seb.James@Sheffield.ac.uk\vspace{1em}\\
-Short title: Self-organising whisker barrels\vspace{5em}\\
-\end{flushleft}
-
-\newpage{}
-%\raggedright
-
-\section*{Abstract}
-
+\begin{abstract}
 Understanding how genetic information and spontaneous pattern formation
@@
 experimentally manipulated.
+\end{abstract}
 
-\section*{Introduction}
+\dates{This manuscript was compiled on \today}
+\doi{\url{www.pnas.org/cgi/doi/10.1073/pnas.XXXXXXXXXX}}
+
+\begin{document}
+
+% Blue for comments
+\newcommand{\cmnt}[1]{\textcolor{blue}{#1}}
+% Divergence symbol (Del dot Del)
+\newcommand{\dvrg}{\nabla\vcdot\nabla}
+% Emphasis shortcut
+\newcommand{\e}{\emph}
+% Bold shortcut
+\newcommand{\bol}{\textbf}
+% Math bold
+\newcommand{\mb}[1]{\mathbf{#1}}
+\makeatletter
+\newcommand*\vcdot{\mathpalette\vcdot@{.35}}
+\newcommand*\vcdot@[2]{\mathbin{\vcenter{\hbox{\scalebox{#2}{$\m@th#1\bullet$}}}}}
+\makeatother
+
+\maketitle
+\thispagestyle{firststyle}
+\ifthenelse{\boolean{shortarticle}}{\ifthenelse{\boolean{singlecolumn}}{\abscontentformatted}{\abscontent}}{}
 
@@
 
-Spatial patterns in neural connectivity provide clues about the constraints
-under which brains evolve and develop. Whether or not patterning governs
-information processing, models of the development of information processing
-circuits can be validated in terms of their ability to account for biological
-patterning (\citealp{purves_iterated_1992}; see also
-\citealp{wilson_what_2015,bednar_cortical_2016}). One of the most distinctive patterns of
-neural connectivity is found in the rodent barrel cortex
-(\citealp{woolsey_pattern_1948,woolsey_structural_1970,welker_structure_1974}, see
-\citealp{fox_barrel_2008}). Understanding how this particular pattern forms might shed
-light on the interplay of genetic and extrinsic factors that shape cortical
-circuits, and on a wealth of anatomical, physiological, and behavioural data
-obtained using this model system.
+\dropcap{S}patial patterns in neural connectivity provide clues about
+the constraints under which brains evolve and develop. Whether or not
+patterning governs information processing, models of the development
+of information processing circuits can be validated in terms of their
+ability to account for biological patterning
+(\citealp{purves_iterated_1992}; see also
+\citealp{wilson_what_2015,bednar_cortical_2016}). One of the most
+distinctive patterns of neural connectivity is found in the rodent
+barrel cortex
+(\citealp{woolsey_pattern_1948,woolsey_structural_1970,welker_structure_1974},
+see \citealp{fox_barrel_2008}). Understanding how this particular
+pattern forms might shed light on the interplay of genetic and
+extrinsic factors that shape cortical circuits, and on a wealth of
+anatomical, physiological, and behavioural data obtained using this
+model system.
 
@@
 number of thalamocortical projections, $N$. In general, we were not able to
-producen a complexity of structure in the resulting patterns of the
+produce a complexity of structure in the resulting patterns of the
 connections, $c(\mb{x},t\rightarrow\infty)$, which was not present in the
@@
 which leads to a reduction in the rate of increase of branching density for
-axon type $i$ in regions where the branching density of \e{other} axon types
+axon type $i$ in regions where the branching density of \emph{other} axon types
 is high.
@@
 
-We first investigate the effect of the competition by examining a system of 25
-thalamocortical axon types simulated on a square domain. The axons interact
-with two orgthogonal signalling gradients, one of which increases from the top
-edge to the bottom edge, the other of which increases from the right edge to
-the left edge. The values of $\gamma$ are chosen as a grid.
-
-\begin{figure*}
-\begin{center}
-\includegraphics[width=\textwidth]{./Fig_25N2M_square/Fig_25N2M_square.png}
-\end{center}
-%\vspace{-0em}
-\caption{\textbf{(a)--(e)} Connection densities of 25 thalamocortical axons
-  simulated on a square domain. \textbf{(a)} Connection density at simulation
-  time t=0. The left image shows a map of the identity of the thalamocortical
-  axon type with the highest absolute connection density. Hexagonal elements
-  coloured the darkest blue have the highest connection density for TC type
-  $i=1$; those coloured the darkest red have maximal connection density in TC
-  type $i=25$. The right hand image of each pair shows the hexes for which the
-  connection density passes 0.5 for each TC type, giving a contour map. At
-  t=0, connection densities are randomly distributed. \textbf{(b)} t=0.14. The
-  grid-like organisation of the TC connections is beginning to emerge.
-  \textbf{(c)} t=0.2. The 5x5 grid is now clear in the left map. Branching
-  competition and differential interaction with the two guidance gradients is
-  causing the branching, and therefore the connection densities, to localise in
-  25 different locations. \textbf{(d)} t=0.64. Connection densities for each
-  TC type have increased sufficiently for the contours to form. In (d), the
-  mapping between colour and TC type $i$ is indicated. \textbf{(e)}
-  t=1.41. Contours are well formed. In (e), the $\gamma$ values for each TC
-  type, $i$, are shown. $\gamma_{i,1}$ gives the interaction with molecule `1'
-  [red map in (f)]; $\gamma_{i,2}$ is the interaction with molecule
-  `2'. \textbf{(f)} The two guidance gradients (`1' and `2') for this
-  simulation.}
-\label{fig:square}
-\end{figure*}
+With the addition of the competition term, $\chi$, it is necessary to
+introduce a normalisation step, in order to prevent one TC type from
+{out-competing every other TC type. Normalisation represents the idea
+that for a fixed number of afferent thalamocortical axons, the overall
+number of branches must have some upper limit; given the finite
+resources available in each axon for building new processes. We
+divisively normalize the branching density, dividing it at each
+computational step by the mean branching density per unit area. It is
+then multiplied by the branching density per unit area at time 0, to
+account for the possibility that there might be more axons afferent
+for TC type $1$ cf.~TC type $2$:
+%
+\begin{equation} \label{eq:norm}
+  a_i(\mathbf{x}, t) = \int_\Omega  a_i(\mathbf{x}, 0) d\sigma \; \frac {a'_i(\mathbf{x}, t)} {\int_\Omega
+  a'_i(\mathbf{x}, t) d\sigma}
+\end{equation}
+%
+The prime symbols indicate the use of the intermediate value of $a_i$
+after it has been computed according to Eqs.~\ref{eq:dc}--\ref{eq:branch_comp}
+on the basis of the system state at time $t-{\delta}t$.
 
@@
 recreated in barrel cortex.
-
-\subsubsection*{Stronger branching competition}
 
@@
 \begin{center}
-\includegraphics[width=0.7\textwidth]{../boundaries/barreloids_haidarliu_graph.png}
+\includegraphics[width=0.7\textwidth]{../../boundaries/barreloids_haidarliu_graph.png}
 \end{center}
@@
 \begin{center}
-\includegraphics[width=\textwidth]{./realistic_barrels.png}
+\includegraphics[width=\textwidth]{../longpaper/realistic_barrels.png}
 \end{center}
@@
 experiments like those by Shimogori \& Grove.
-
-Linear, circ linear, exponential, etc.
-
-\subsection*{The barrel field reacts to simulated whisker plucking}
-
-A section on whisker plucking?
-
-%\subsubsection*{With surrounding large domains}
-%Maybe, put this in to remind myself that I did this. It looks nice.
-
-%\subsection*{Plucking experiments}
-%Include results of plucking a single whisker, a row of whiskers and a
-%column of whiskers.
-
-%\subsection*{Any parameter sweeps necessary to show behaviour is
-%  robust to changes in params}
-
-\subsection*{Extending the model to capture row and arc development}
-
-\subsubsection*{Manipulating Competition: Group competition stronger/weaker than inter-group competition}
-
-Just an idea - have two separate epsilons.
-
-\subsubsection*{Manipulating Guidance: Relative interaction strengths}
-
-One possibility is to modify the strength of the guidance interactions of the
-two, orthogonal guidance gradients. In the simulation we modify the magnitude
-of the expression of each guidance molecule, although this is equivalent to,
-and could stand for a more or less vigourous movement of axons in response to
-un-modified guidance gradients.
-
-\subsubsection*{Manipulating Guidance: Delayed onset in one axis}
-
-Another idea; begin with only one guidance gradient, or one strong and
-one very weak guidance gradient, and begin the development in this
-form, so as to encourage the formation of rows. Subsequently, turn on
-the second strong guidance signal.
-
-%\begin{equation}
-%\chi=\frac{\epsilon_i  a_i}{N-1} \sum_{j \ne i}^{N} a_j(\mb{x}, t)^l
-%\end{equation}
-
-%And normalization step
-
-%\begin{equation}\label{eq:norm_comp8}
-%a_i' = n_e \; \frac {a_i} {\textstyle \sum_j^{n_e} a_{i,j} }
-%\end{equation}
-
-%($a_i'$ becomes $a_i$ in the next computational step.)
 
@@
 
-\bibliography{BarrelEmerge}
+\bibliography{../BarrelEmerge}
 
 \end{document}
-
-%%%%% ORIGINAL EQUATIONS MAINTAINED AS NOTES HERE FOR REFERENCE
-
-\section*{App 1: Karbowski system with N TCs and M gradients}
-%
-\begin{equation} \label{eq:Karb2D_dc}
-\frac{\partial c_i(\mb{x},t)}{\partial t} = -\alpha c_i(\mb{x},t) + \beta n(\mb{x},t)
-[a_i(\mb{x},t)]^k
-\end{equation}
-%
-\begin{equation} \label{eq:Karb2D_conserve}
-n(\mb{x},t) = 1 - \sum_{i=1}^{N} c_i(\mb{x}, t)
-\end{equation}
-%
-\begin{equation} \label{eq:Karb2D_da}
-\frac{\partial a_i(\mb{x},t)}{\partial t}
-= \nabla\vcdot\mb{J}_i(\mb{x},t) - \frac{\partial c_i(\mb{x},t)}{\partial t}
-\end{equation}
-
-with the flux current
-
-\begin{equation} \label{eq:Karb2D_J}
-\mb{J}_i(\mb{x},t) = D \nabla a_i(\mb{x},t) - a_i
-\sum_{j=1}^M \big(\gamma_{i,j} \nabla\rho_j(\mb{x}) \big)
-\end{equation}
-
-The boundary condition applied to these equations is:
-
-\begin{equation}
-\mb{J}_i(\mb{x},t) \bigg\rvert_{boundary} = 0
-\end{equation}
-
-\section{App 2: Karbowski system with N TCs and M gradients plus competition}
-
-%
-\begin{equation} \label{eq:dc2}
-\frac{\partial c_i(\mb{x},t)}{\partial t} = -\alpha c_i(\mb{x},t) + \beta n(\mb{x},t)
-[a_i(\mb{x},t)]^k
-\end{equation}
-%
-\begin{equation} \label{eq:conserve}
-n(\mb{x},t) = 1 - \sum_{i=1}^{N} c_i(\mb{x}, t)
-\end{equation}
-%
-\begin{equation} \label{eq:da2}
-\frac{\partial a_i(\mb{x},t)}{\partial t}
-= \nabla\vcdot\mb{J}_i(\mb{x},t) - \frac{\partial c_i(\mb{x},t)}{\partial t}- \frac{\epsilon_i  a_i}{N-1} \sum_{j \ne i}^{N} a_j(\mb{x}, t)^l
-\end{equation}
-
-And normalization step
-
-\begin{equation}\label{eq:norm_comp8}
-a_i' = n_e \; \frac {a_i} {\textstyle \sum_j^{n_e} a_{i,j} }
-\end{equation}
-
-($a_i'$ becomes $a_i$ in the next computational step.)

### 2019-11-19 paper/briefpaper/brief.sh

@@ -1,0 +1,6 @@
+#!/bin/bash
+
+xelatex brief.tex
+bibtex brief.aux
+xelatex brief.tex
+xelatex brief.tex

### 2019-11-19 paper/briefpaper/brief.tex
Reformatting paragrapohs with a set-fill-column value of 78

@@
 
-\title{Self-organization of whisker barrels in simulation requires competition between thalamocortical axons}
+%\title{Self-organization of whisker barrels in simulation requires competition between thalamocortical axons}
+%\title{Simulated barrel cortex self-organizes using information carried by thalamocortical axons}
+\title{Information carried by thalamocortical axons enables self-organization of the barrel field}
 % While writing, don't stop for errors
@@
 % Please add a significance statement to explain the relevance of your work
-\significancestatement{The mystacial whiskers of the rat are a complex, ordered set of structures whose pattern is repeated several times in separate locations in the brain.
-We contribute a mechanism by which their spatially ordered pattern can be recreated in a self-organising process which requires minimal genetic pre-determination.}
+\significancestatement{The mystacial whiskers of the rat are a complex,
+  ordered set of structures whose pattern is repeated several times in
+  separate locations in the brain. We contribute a mechanism by which their
+  spatially ordered pattern can be recreated in a self-organising process
+  which requires minimal genetic pre-determination.}
 
-\correspondingauthor{\textsuperscript{1}To whom correspondence should be addressed. E-mail: seb.james\@sheffield.ac.uk}
+\correspondingauthor{\textsuperscript{1}To whom correspondence should be addressed. E-mail: \href{mailto:seb.james@sheffield.ac.uk}{\nolinkurl{seb.james@sheffield.ac.uk}}}
 
 % At least three keywords are required at submission. Please provide three to five keywords, separated by the pipe symbol.
-\keywords{Keyword 1 $|$ Keyword 2 $|$ Keyword 3 $|$ ...}
+\keywords{Barrel cortex $|$ Self-organization $|$ Somatotopy $|$ Center-point $|$ Pattern $|$}
 
@@
 
-\dropcap{S}patial patterns in neural connectivity provide clues about
-the constraints under which brains evolve and develop. Whether or not
-patterning governs information processing, models of the development
-of information processing circuits can be validated in terms of their
-ability to account for biological patterning
-(\citealp{purves_iterated_1992}; see also
-\citealp{wilson_what_2015,bednar_cortical_2016}). One of the most
+\dropcap{S}patial patterns in neural connectivity provide clues about the
+constraints under which brains evolve and develop. Whether or not patterning
+governs information processing, models of the development of information
+processing circuits can be validated in terms of their ability to account for
+biological patterning (\citealp{purves_iterated_1992}
+%; see also \citealp{wilson_what_2015,bednar_cortical_2016}
+). One of the most
 distinctive patterns of neural connectivity is found in the rodent
-barrel cortex
-(\citealp{woolsey_pattern_1948,woolsey_structural_1970,welker_structure_1974},
-see \citealp{fox_barrel_2008}). Understanding how this particular
+barrel cortex \cite{woolsey_structural_1970}.
+%
+%(\citealp{woolsey_pattern_1948,woolsey_structural_1970,welker_structure_1974},
+%see \citealp{fox_barrel_2008}).
+%
+Understanding how this particular
 pattern forms might shed light on the interplay of genetic and
-extrinsic factors that shape cortical circuits, and on a wealth of
-anatomical, physiological, and behavioural data obtained using this
-model system.
+extrinsic factors that shape cortical circuits.
+%, and on a wealth of
+%anatomical, physiological, and behavioural data obtained using this
+%model system.
 
@@
 axons, which are enclosed by borders a few neurons thick from postnatal day 3
-\citep{erzurumlu_development_2012}. In the plane tangential to the cortical surface the
-barrels constitute a map of the arrangement of the whiskers on the
-face (see \citealp{yamakado_subdivision_1979}), such
-that the cells of adjacent barrels respond most strongly and most quickly to
-deflection of adjacent whiskers \citep{armstrong-james_flow_1992}. The barrels
-reflect subcortical whisker maps comprising cellular aggregates known as
-barrelettes in the brainstem and barreloids in the thalamus
-\citep{killackey_pattern_1980}.
+\citep{erzurumlu_development_2012}. In the plane tangential to the cortical
+surface the barrels constitute a map of the arrangement of the whiskers on the
+face
+%(see \citealp{yamakado_subdivision_1979})
+, such that the cells of adjacent barrels respond most strongly and most
+quickly to deflection of adjacent whiskers
+\citep{armstrong-james_flow_1992}. The barrels reflect subcortical whisker
+maps comprising cellular aggregates known as barreloids in the thalamus and
+barrelettes in the brainstem.
+%\citep{killackey_pattern_1980}.
 
@@
 identified by its position on a two-dimensional grid
-\citep{zucker_coding_1969,killackey_pattern_1980,van_der_loos_barreloids_1976,haidarliu_size_2001}. The barrels are normally up to
-twice as long in the axis aligned to the grid ‘rows’ compared to the ‘arcs’,
-and the rows are visible from day 2 \citep{rebsam_refinement_2002} whereas the arcs become
-distinct by day 3 \citep{erzurumlu_development_2012,rebsam_refinement_2002}. In metabotropic glutamate
-receptor 5 (mGlu5) knockouts the barrel rows develop but the arcs do not
-(\citealp{hannan_plc-1_2001} see also \citealp{fox_barrel_2008}), thus distinct processes
-likely guide development of the barrels with respect to the orthogonal grid
-axes.
+\citep{zucker_coding_1969,killackey_pattern_1980,van_der_loos_barreloids_1976,haidarliu_size_2001}. The
+barrels are normally up to twice as long in the axis aligned to the grid
+‘rows’ compared to the ‘arcs’, and the rows are visible from day 2
+\citep{rebsam_refinement_2002} whereas the arcs become distinct by day 3
+\citep{erzurumlu_development_2012,rebsam_refinement_2002}. In metabotropic
+glutamate receptor 5 (mGlu5) knockouts the barrel rows develop but the arcs do
+not (\citealp{hannan_plc-1_2001} see also \citealp{fox_barrel_2008}), thus
+distinct processes likely guide development of the barrels with respect to the
+orthogonal grid axes.
 
@@
 thalamic calcium waves that propagate via mGlu5-receptor activation
-(\citealp{anton-bolanos_prenatal_2019}, see also \citealp{anton-bolanos_developmental_2018}). Patterning also depends
-on a complex network of axon guidance processes, which includes interactions
-between guidance molecules like ephrin-A5 and A7 \citep{miller_epha7-ephrin-a5_2006} and
-between adhesion molecules like cadherin-6 and 8 \citep{greig_molecular_2013,bishop_regulation_2000},
-and the binding of p75 to neurotropins (\citealp{shimogori_fibroblast_2005,bishop_distinct_2002};
-see also \citealp{dye_lifespan_2011,dye_lifespan_2011-1}). This network is orchestrated by
-interactions between morphogens Fgf8 and Fgf17 and transcription factors Emx2,
-Pax6, Sp8, and Coup-tf1 \citep{shimogori_fibroblast_2005,bishop_regulation_2000}, which are expressed
+(\citealp{anton-bolanos_prenatal_2019}, see also
+\citealp{anton-bolanos_developmental_2018}). Patterning also depends on a
+complex network of axon guidance processes, which includes interactions
+between guidance molecules like ephrin-A5 and A7
+\citep{miller_epha7-ephrin-a5_2006} and between adhesion molecules like
+cadherin-6 and 8 \citep{greig_molecular_2013,bishop_regulation_2000}, and the
+binding of p75 to neurotropins
+(\citealp{shimogori_fibroblast_2005,bishop_distinct_2002}; see also
+\citealp{dye_lifespan_2011,dye_lifespan_2011-1}). This network is orchestrated
+by interactions between morphogens Fgf8 and Fgf17 and transcription factors
+Emx2, Pax6, Sp8, and Coup-tf1
+\citep{shimogori_fibroblast_2005,bishop_regulation_2000}, which are expressed
 in gradients that mark orthogonal axes and can be manipulated to cause the
@@
 
-The boundaries of the barrels form polygonal shapes that together
-create a Voronoi tesselation across the cortical sheet, wherein all
-points enclosed by a given boundary are closer to its center than to
-the center of any other boundary \citep{senft_mouse_1991}. This
-patterning suggests that the barrel topology may be preserved in the
-initial projection of the thalamocortical axons into the cortex, and
-that a barrel forms by the lateral branching of the axons from their
-initial centres of mass, which ceases upon contact with the axons
-branching from adjacent centres. According to this model the barrel
-pattern could be entirely determined by the distribution of the
-initial centre points, which would itself be determined by guidance
-cues. However, the assumption of pre-arranged centres is difficult to
-resolve with the observation that thalamocortical axons arrive in the
-cortical plate as an undifferentiated bundle prior to barreloid
-formation (\citealp{agmon_organized_1993}; see \citealp{erzurumlu_development_2012} for a review).
+The boundaries of the barrels form polygonal shapes that together create a
+Voronoi tesselation across the cortical sheet, wherein all points enclosed by
+a given boundary are closer to its center than to the center of any other
+boundary \citep{senft_mouse_1991}. This patterning suggests that the barrel
+topology may be preserved in the initial projection of the thalamocortical
+axons into the cortex, and that a barrel forms by the lateral branching of the
+axons from their initial centres of mass, which ceases upon contact with the
+axons branching from adjacent centres. According to this model the barrel
+pattern could be entirely determined by the distribution of the initial centre
+points, which would itself be determined by guidance cues. However, the
+assumption of pre-arranged centres is difficult to resolve with the
+observation that thalamocortical axons arrive in the cortical plate as an
+undifferentiated bundle prior to barreloid formation
+(\citealp{agmon_organized_1993}; see \citealp{erzurumlu_development_2012} for
+a review).
 
@@
 
-\cite{karbowski_model_2004} developed what remains, to the best of our knowledge, the
-only dynamical model of cortical arealization capable of generating
-predictions about the influence of signalling molecules on the size, shape,
-and topological organization of emergent cortical fields. This model describes
-competition between thalamocortical projections for territory along a
-one-dimensional string of cortical cells, e.g., along an anterior-posterior
-axis. The equations of the model define how the density of connections, $c$,
-and the density of axon branches, $a$, interact at time $t$ at each cortical
-location $x$, for $N$ thalamocortical projections indexed by $i$. The Voronoi
-tesselation used to describe barrel patterning is a two-dimensional structure,
-hence we first extended the original model for simulation on a two-dimensional
-cortical sheet, and describe it here as such (i.e., substituting $x$ with the
-two-dimensional position vector $\mb{x}$ and substituting terms of the form
-$\frac{d}{dx}$ with the spatial gradient operator $\nabla$).
+\cite{karbowski_model_2004} developed what remains, to the best of our
+knowledge, the only dynamical model of cortical arealization capable of
+generating predictions about the influence of signalling molecules on the
+size, shape, and topological organization of emergent cortical fields. This
+model describes competition between thalamocortical projections for territory
+along a one-dimensional string of cortical cells, e.g., along an
+anterior-posterior axis. The equations of the model define how the density of
+connections, $c$, and the density of axon branches, $a$, interact at time $t$
+at each cortical location $x$, for $N$ thalamocortical projections indexed by
+$i$. The Voronoi tesselation used to describe barrel patterning is a
+two-dimensional structure, hence we first extended the original model for
+simulation on a two-dimensional cortical sheet, and describe it here as such
+(i.e., substituting $x$ with the two-dimensional position vector $\mb{x}$ and
+substituting terms of the form $\frac{d}{dx}$ with the spatial gradient
+operator $\nabla$).
 
@@
 
-Changes in axon branching density are assumed to depend also on the
-influence of guidance molecules and on the density of branches at
-adjacent cortical locations:
+Changes in axon branching density are assumed to depend also on the influence
+of guidance molecules and on the density of branches at adjacent cortical
+locations:
 
@@
 branching density in some region, leading to roughly equal numbers of
-connections forming for each TC type, up until the `hard' connection limit
-is reached.
+connections forming for each TC type, up until the `hard' connection limit is
+reached.
 
@@
 which leads to a reduction in the rate of increase of branching density for
-axon type $i$ in regions where the branching density of \emph{other} axon types
-is high.
+axon type $i$ in regions where the branching density of \emph{other} axon
+types is high.
 
@@
 introduce a normalisation step, in order to prevent one TC type from
-{out-competing every other TC type. Normalisation represents the idea
-that for a fixed number of afferent thalamocortical axons, the overall
-number of branches must have some upper limit; given the finite
-resources available in each axon for building new processes. We
-divisively normalize the branching density, dividing it at each
-computational step by the mean branching density per unit area. It is
-then multiplied by the branching density per unit area at time 0, to
-account for the possibility that there might be more axons afferent
+out-competing every other TC type. Normalisation represents the idea that for
+a fixed number of afferent thalamocortical axons, the overall number of
+branches must have some upper limit; given the finite resources available in
+each axon for building new processes. We divisively normalize the branching
+density, dividing it at each computational step by the mean branching density
+per unit area. It is then multiplied by the branching density per unit area at
+time 0, to account for the possibility that there might be more axons afferent
 for TC type $1$ cf.~TC type $2$:
@@
 %
-The prime symbols indicate the use of the intermediate value of $a_i$
-after it has been computed according to Eqs.~\ref{eq:dc}--\ref{eq:branch_comp}
-on the basis of the system state at time $t-{\delta}t$.
+The prime symbols indicate the use of the intermediate value of $a_i$ after it
+has been computed according to Eqs.~\ref{eq:dc}--\ref{eq:branch_comp} on the
+basis of the system state at time $t-{\delta}t$.
 

### 2019-11-19 paper/briefpaper/brief.tex
Some work reducing down the text in the paper

@@
 % At least three keywords are required at submission. Please provide three to five keywords, separated by the pipe symbol.
-\keywords{Barrel cortex $|$ Self-organization $|$ Somatotopy $|$ Center-point $|$ Pattern $|$}
+\keywords{Barrel cortex $|$ Self-organization $|$ Somatotopy $|$ Center-point $|$ Pattern}
 
@@
 maintained in the projections from brainstem to thalamus and from thalamus to
-cortex? To investigate, we began by extending a computational model of
-neocortical arealization by Karbowski \& Ermentrout (2004; J. Comp. Neurosci.)
-to generate fields on a two-dimensional cortical sheet. Central to this model
-is an assumption that axonal branching and synaptogenesis are coupled, and
-that thalamic projections compete for cortical territory under the influence
-of molecular signalling gradients. We found that these processes alone are
-incapable of generating patterns of cortical fields that are more intricate
-than the patterns already present in molecular guidance signals. However, a
-full barrel-field pattern emerges under the guidance of as few as four
-signalling gradients when the model is extended to introduce competition
-between the branches of axons that originate from different thalamic
-barreloids. We show that the key requirements for the emergence of realistic
-(Voronoi tesselation) patterning are i) at each cortical location
-thalamocortical projections compete for a limited number of available
-connections, ii) at each location the axonal branching rate of each
-thalamocortical projection is reduced by the total branching density of the
-other projections, and iii) the density of axon branching for each
-thalamocortical projection is conserved. We demonstrate how a purely
+cortex?
+%
+To investigate, we began by extending a existing computational model of axon
+branching and synaptogenesis in thalamocortical projections under the
+influence of molecular signalling gradients.  We found that, although these
+processes alone are incapable of generating patterns of cortical fields that
+are more intricate than the patterns already present in the signalling
+gradients, the addition of competition between the branches of axons that
+originate from different thalamic barreloids allowed a full barrel-field
+pattern to emerge in the presence of only two gradients.
+%
+%However, a
+%full barrel-field pattern emerges under the guidance of as few as four
+%signalling gradients when the model is extended to introduce competition
+%between the branches of axons that originate from different thalamic
+%barreloids.
+%
+%We show that the key requirements for the emergence of realistic
+%(Voronoi tesselation) patterning are i) at each cortical location
+%thalamocortical projections compete for a limited number of available
+%connections, ii) at each location the axonal branching rate of each
+%thalamocortical projection is reduced by the total branching density of the
+%other projections, and iii) the density of axon branching for each
+%thalamocortical projection is conserved.
+%
+We demonstrate how a purely
 self-organising system can initialise a pattern which can then be faithfully
-replicated in multiple neural structures. We show that this minimal model of
-barrel self-organisation provides a parsimonious account of data from a range
-of classic studies in which intrinsic and extrinsic factors have been
-experimentally manipulated.
+replicated in multiple neural structures.
+%We show that this minimal model of
+%barrel self-organisation provides a parsimonious account of data from a range
+%of classic studies in which intrinsic and extrinsic factors have been
+%experimentally manipulated.
 \end{abstract}
@@
 \dropcap{S}patial patterns in neural connectivity provide clues about the
-constraints under which brains evolve and develop. Whether or not patterning
-governs information processing, models of the development of information
-processing circuits can be validated in terms of their ability to account for
-biological patterning (\citealp{purves_iterated_1992}
-%; see also \citealp{wilson_what_2015,bednar_cortical_2016}
-). One of the most
+constraints under which brains evolve and develop.
+%Whether or not patterning
+%governs information processing, models of the development of information
+%processing circuits can be validated in terms of their ability to account for
+%biological patterning \citep{purves_iterated_1992}.
+%%see also \citealp{wilson_what_2015,bednar_cortical_2016}
+One of the most
 distinctive patterns of neural connectivity is found in the rodent
 barrel cortex \cite{woolsey_structural_1970}.
-%
 %(\citealp{woolsey_pattern_1948,woolsey_structural_1970,welker_structure_1974},
 %see \citealp{fox_barrel_2008}).
-%
-Understanding how this particular
-pattern forms might shed light on the interplay of genetic and
-extrinsic factors that shape cortical circuits.
+Understanding how this particular pattern forms might shed light on the
+interplay of genetic and extrinsic factors that shape cortical circuits.
 %, and on a wealth of
@@
 The barrels are large cortical columns, identifiable from birth in layer 4 of
-rodent primary somatosensory cortex as dense clusters of thalamocortical
-axons, which are enclosed by borders a few neurons thick from postnatal day 3
+primary somatosensory cortex as dense clusters of thalamocortical axons, which
+are enclosed by borders a few neurons thick from postnatal day 3
 \citep{erzurumlu_development_2012}. In the plane tangential to the cortical
 surface the barrels constitute a map of the arrangement of the whiskers on the
-face
+face,
 %(see \citealp{yamakado_subdivision_1979})
-, such that the cells of adjacent barrels respond most strongly and most
-quickly to deflection of adjacent whiskers
-\citep{armstrong-james_flow_1992}. The barrels reflect subcortical whisker
-maps comprising cellular aggregates known as barreloids in the thalamus and
+such that the cells of adjacent barrels respond most strongly and most quickly
+to deflection of adjacent whiskers \citep{armstrong-james_flow_1992}. The
+barrels reflect subcortical whisker maps comprising cellular aggregates known
+as barreloids in the thalamus \cite{van_der_loos_barreloids_1976} and
 barrelettes in the brainstem.
@@
 
-Each whisker and its corresponding barrelette, barreloid, and barrel can be
-identified by its position on a two-dimensional grid
-\citep{zucker_coding_1969,killackey_pattern_1980,van_der_loos_barreloids_1976,haidarliu_size_2001}. The
-barrels are normally up to twice as long in the axis aligned to the grid
-‘rows’ compared to the ‘arcs’, and the rows are visible from day 2
-\citep{rebsam_refinement_2002} whereas the arcs become distinct by day 3
-\citep{erzurumlu_development_2012,rebsam_refinement_2002}. In metabotropic
-glutamate receptor 5 (mGlu5) knockouts the barrel rows develop but the arcs do
-not (\citealp{hannan_plc-1_2001} see also \citealp{fox_barrel_2008}), thus
-distinct processes likely guide development of the barrels with respect to the
-orthogonal grid axes.
+%Each whisker and its corresponding barrelette, barreloid, and barrel can be
+%identified by its position on a two-dimensional grid.
+%\citep{zucker_coding_1969,killackey_pattern_1980,van_der_loos_barreloids_1976,haidarliu_size_2001}.
+%
+%The barrels are normally up to twice as long in the axis aligned to the grid
+%‘rows’ compared to the ‘arcs’, and the rows are visible from day 2
+%\citep{rebsam_refinement_2002} whereas the arcs become distinct by day 3
+%\citep{erzurumlu_development_2012,rebsam_refinement_2002}.
+%In metabotropic glutamate receptor 5 (mGlu5) knockouts the barrel rows develop
+%but the arcs do not (\citealp{hannan_plc-1_2001} see also
+%\citealp{fox_barrel_2008}), thus distinct processes likely guide development
+%of the barrels with respect to the orthogonal grid axes.
 
@@
 thalamic calcium waves that propagate via mGlu5-receptor activation
-(\citealp{anton-bolanos_prenatal_2019}, see also
-\citealp{anton-bolanos_developmental_2018}). Patterning also depends on a
-complex network of axon guidance processes, which includes interactions
-between guidance molecules like ephrin-A5 and A7
+\citep{anton-bolanos_prenatal_2019}.
+%(\citealp{anton-bolanos_prenatal_2019}, see also
+%\citealp{anton-bolanos_developmental_2018}).
+Patterning also depends on a complex network of axon guidance processes, which
+includes interactions between guidance molecules like ephrin-A5 and A7
 \citep{miller_epha7-ephrin-a5_2006} and between adhesion molecules like
-cadherin-6 and 8 \citep{greig_molecular_2013,bishop_regulation_2000}, and the
+cadherin-6 and 8 \citep{bishop_regulation_2000}, and the
 binding of p75 to neurotropins
-(\citealp{shimogori_fibroblast_2005,bishop_distinct_2002}; see also
-\citealp{dye_lifespan_2011,dye_lifespan_2011-1}). This network is orchestrated
-by interactions between morphogens Fgf8 and Fgf17 and transcription factors
-Emx2, Pax6, Sp8, and Coup-tf1
+\citep{bishop_distinct_2002}.
+% or both citations?
+%\citep{shimogori_fibroblast_2005,bishop_distinct_2002}.
+%
+%(\citealp{shimogori_fibroblast_2005,bishop_distinct_2002}
+%; see also \citealp{dye_lifespan_2011,dye_lifespan_2011-1}).
+This network is orchestrated by interactions between morphogens Fgf8 and Fgf17
+and transcription factors Emx2, Pax6, Sp8, and Coup-tf1
 \citep{shimogori_fibroblast_2005,bishop_regulation_2000}, which are expressed
 in gradients that mark orthogonal axes and can be manipulated to cause the
-barrels to stretch, shrink, shift, skew, and even duplicate
-\citep{shimogori_fibroblast_2005,assimacopoulos_fibroblast_2012,borello_sp8_2014,sahara_sp8_2007,ypsilanti_transcriptional_2016,sur_patterning_2005}.
+barrels to stretch, shrink, shift, skew, and even duplicate \cite{assimacopoulos_fibroblast_2012}.
+% A longer list of citations:
+%\citep{shimogori_fibroblast_2005,assimacopoulos_fibroblast_2012,borello_sp8_2014,sahara_sp8_2007,ypsilanti_transcriptional_2016,sur_patterning_2005}.
+%
+While complex structure in protein expression can be observed for some genes,
+such as the molecules Bhlhb5 \citep{joshi_bhlhb5_2008}, a complex gene
+expression pattern with the complexity of the barrel field has not, to our
+knowledge been reported and so the mechanism by which the barrels form seems
+likely to be one in which the structure forms by self-organization.
 
-The boundaries of the barrels form polygonal shapes that together create a
-Voronoi tesselation across the cortical sheet, wherein all points enclosed by
-a given boundary are closer to its center than to the center of any other
-boundary \citep{senft_mouse_1991}. This patterning suggests that the barrel
+The boundaries of the barrels form a Voronoi tesselation across the cortical
+sheet \citep{senft_mouse_1991}. This patterning suggests that the barreloid
 topology may be preserved in the initial projection of the thalamocortical
@@
 observation that thalamocortical axons arrive in the cortical plate as an
-undifferentiated bundle prior to barreloid formation
-(\citealp{agmon_organized_1993}; see \citealp{erzurumlu_development_2012} for
-a review).
+undifferentiated bundle prior to barreloid formation within the thalamus
+\cite{agmon_organized_1993}. % Plus erzurumli, probably
 
@@
 
-To this end, the aim of the present study is to establish whether it is
-possible for the observed arrangement of barrels to self-organise in a system
-governed by reaction-diffusion dynamics, under the guidance of molecular
-signalling gradients, and in the absence of pre-defined centre-points. Our
-approach is to represent in a computational model the minimal set of
+We seek to establish whether it is possible for the observed arrangement of
+barrels to self-organise in a system governed by reaction-diffusion dynamics,
+under the guidance of molecular signalling gradients, and in the absence of
+pre-defined centre-points.
+%
+Our approach is to represent in a computational model the minimal set of
 biological assumptions that might lead to the emergence of barrel patterning
@@
 
-\cite{karbowski_model_2004} developed what remains, to the best of our
-knowledge, the only dynamical model of cortical arealization capable of
-generating predictions about the influence of signalling molecules on the
-size, shape, and topological organization of emergent cortical fields. This
-model describes competition between thalamocortical projections for territory
-along a one-dimensional string of cortical cells, e.g., along an
-anterior-posterior axis. The equations of the model define how the density of
-connections, $c$, and the density of axon branches, $a$, interact at time $t$
+Karbowski \& Ermentrout \citep{karbowski_model_2004} developed a dynamical
+model of cortical arealization capable of generating predictions about the
+influence of signalling molecules on the size, shape, and topological
+organization of emergent cortical fields. Their model describes competition
+between thalamocortical projections for territory along an anterior-posterior
+cortical axis. The equations of the model define how the density of
+connections, $c(x,t)$, and the density of axon branches, $a(x,t)$, interact at time $t$
 at each cortical location $x$, for $N$ thalamocortical projections indexed by
-$i$. The Voronoi tesselation used to describe barrel patterning is a
-two-dimensional structure, hence we first extended the original model for
-simulation on a two-dimensional cortical sheet, and describe it here as such
-(i.e., substituting $x$ with the two-dimensional position vector $\mb{x}$ and
-substituting terms of the form $\frac{d}{dx}$ with the spatial gradient
-operator $\nabla$).
-
-The original model was derived from the assumption that the rates at which
+$i$.
+%The Voronoi tesselation used to describe barrel patterning is a
+%two-dimensional structure, hence we first extended the original model for
+%simulation on a two-dimensional cortical sheet, and describe it here as such
+%(i.e., substituting $x$ with the two-dimensional position vector $\mb{x}$ and
+%substituting terms of the form $\frac{d}{dx}$ with the spatial gradient
+%operator $\nabla$).
+%
+The model was derived from the assumption that the rates at which
 thalamocortical synapses form and thalamocortical axons branch are
-reciprocally coupled, resulting in an expression for the change in $c_i$ as a
+reciprocally coupled, resulting in expressions for the change in $c_i$ as a
 function of $a_i$, and for the change in $a_i$ as a function of
-$c_i$. Synaptogenesis is modelled as
-
+$c_i$. Synaptogenesis is modelled [in two dimensions, so the variables are
+$a_i(\mb{x},t)$ \& $c_i(\mb{x},t)$] as
+%
 \begin{equation} \label{eq:dc}
-\frac{\partial c_i(\mb{x},t)}{\partial t} =-\alpha c_i(\mb{x},t) +\beta  \left(1 - \sum_{i'=1}^{N} c_{i'}(\mb{x}, t)\right)[a_i(\mb{x},t)]^k.
+\frac{\partial c_i}{\partial t} =-\alpha c_i +\beta  \left(1 - \sum_{i'=1}^{N} c_{i'}\right)[a_i]^k.
 \end{equation}
-
-Accordingly, when the final term evaluates to zero, the density of
-thalamocortical connections decays exponentially at a rate given by the decay
-constant $\alpha$. This occurs where the density of connections over all
-thalamic projections sum to one; interpreted as a state in which all of the
-available synaptic receptors within the cortical dendritic trees are
+%
+When the final term evaluates to zero, the density of thalamocortical
+connections decays exponentially at a rate given by $\alpha$. This occurs when
+the density of connections over all thalamic projections sums to one;
+interpreted as a state in which all available cortical synaptic receptors are
 occupied. Otherwise, the connection density is assumed to vary non-linearly
@@
 locations:
-
 %
 \begin{equation} \label{eq:da}
-\frac{\partial a_i(\mb{x},t)}{\partial t} = \nabla\vcdot\left(D \nabla a_i(\mb{x},t)-a_i(\mb{x},t)\sum_{j=1}^{M} \gamma_{i,j}\nabla \rho_j(\mb{x})\right) - \frac{\partial c_i(\mb{x},t)}{\partial t} + \chi.
+\frac{\partial a_i}{\partial t} = \nabla\vcdot\left(D \nabla a_i-a_i\sum_{j=1}^{M} \gamma_{i,j}\nabla \rho_j(\mb{x})\right) - \frac{\partial c_i}{\partial t} + \chi.
 \end{equation}
+%
+The first term on the right describes the divergence (indicated by
+$\nabla\vcdot$) of a quantity referred to as the `flux of axonal branching'
+(the term enclosed by parentheses, called $\mb{J}_i(\mb{x},t)$).
+%The central idea in heat diffusion
+%models is that the change in temperature (over time) at a given location is
+%proportional to the gradient in the temperature (over space) at that
+%location. Substituting the branching density for heat, axons from a given
+The flux consists of lateral diffusion across the cortical sheet at a rate
+given by the diffusion constant $D$, and an interaction term, in which
+branching is influenced by gradients in the concentration of $M$ distinct
+molecular signals, $\rho(\mb{x})$. The influence of a given signal (indexed by
+$j$) on a given thalamic projection (indexed by $i$), is determined by the
+parameter $\gamma_{i,j}$ which may be positive or negative such that axons of
+a given thalamic projection may branch preferentially towards either higher or
+lower concentrations of each signalling molecule.
+%Interactions for
+%which this value is positive encourage axon branching in the direction in
+%which the concentration of the signal is increasing, whereas negative
+%interactions encourage branching in the opposite direction. Hence the axons of
+%a given thalamic projection are assumed to branch towards either higher or
+%lower concentrations of each signalling molecule.
 
-The first term on the right describes the divergence (indicated by
-$\nabla\vcdot$) of a quantity referred to in models of heat diffusion as the
-`flux' (the term enclosed by parentheses). The central idea in heat diffusion
-models is that the change in temperature (over time) at a given location is
-proportional to the gradient in the temperature (over space) at that
-location. Substituting the branching density for heat, axons from a given
-thalamic projection are thus assumed to branch laterally across the cortical
-sheet at a rate given by the diffusion constant $D$. Lateral branching is
-augmented by a measure of the local alignment between the branching density
-and the gradient in the concentration of each of $M$ distinct molecular
-signals, $\rho(\mb{x})$. The influence of a given signal (indexed by $j$) on a
-given thalamic projection (indexed by $i$), is determined by the parameter
-$\gamma_{i,j}$. Interactions for which this value is positive encourage axon
-branching in the direction in which the concentration of the signal is
-increasing, whereas negative interactions encourage branching in the opposite
-direction. Hence the axons of a given thalamic projection are assumed to
-branch towards either higher or lower concentrations of each signalling
-molecule.
-
-The second term on the right of equation 2 represents an assumption that the
+The second term on the right of Eq.~\ref{eq:da} represents an assumption that the
 axon branching density decreases when the number of connections being made
@@
 connections. Note that we introduce the term $\chi=0$ here, so that we might
-subsequently redefine $\chi$ to introduce the competition between thalamic
-projections that we will show to be necessary for the emergence of a Voronoi
-tesselation (see Results).
+subsequently redefine $\chi$ to introduce competition between thalamic
+projections.
+%that we will show to be necessary for the emergence of a
+%somatotopic pattern.
 
-Two points are important to emphasise here. First, notice that the only
-difference between thalamic projections is the interaction strengths,
-$\gamma$. Hence any differences between the size, shape, and location of
+Two points are important to emphasise: First, notice that the only
+difference between thalamic projections are the interaction strengths,
+$\gamma_{i,j}$. Any differences between the size, shape, and location of
 cortical fields that reliably emerge from the dynamics of the model must be
-due to differences in the choice of these values only. Second, the formulation
+due to differences in the choice of these values only; Second, the formulation
 of the model is at this point entirely local, in the sense that simulating the
@@
 
-The dynamics of the model (equations 1 and 2) were solved on a hexagonal
-lattice of grid elements representing the cortical sheet, under the assumption
-that axon branches cannot extend beyond the boundary of the sheet, i.e.,
-$\mb{J}_i(\mb{x},t) \bigg\rvert_{boundary} = 0$. We developed software tools
-for imposing arbitrary boundary shapes, specified in a scalable vector
-graphics (.svg) file format, on the lattice. The dynamics were numerically
-integrated using fourth-order Runge-Kutta, and using a {\color{red} finite
-  difference method} to compute the flux term (see Supplementary Materials).
+The dynamics of the model (Eqs.~\ref{eq:dc} \& \ref{eq:da}) were solved on a
+hexagonal lattice of grid elements representing the cortical sheet, under the
+assumption that axon branches cannot extend beyond the boundary of the sheet,
+i.e., $\mb{J}_i(\mb{x},t) \big\rvert_{boundary} = 0$. We developed software
+tools for imposing arbitrary boundary shapes on the lattice. The dynamics were
+numerically integrated using fourth-order Runge-Kutta, and using a
+{\color{red} finite difference method} to compute the flux term (see
+Supplementary Materials).
 
@@
 described a two dimensional extension of the axon branching model in
-\citet{karbowski_model_2004}. We showed that in the presence of three guidance
+\cite{karbowski_model_2004}. We showed that in the presence of three guidance
 gradients varying only in one dimension (along the long axis of an elliptical
@@
 
-Complex structure can indeed be observed in protein expression for some genes,
-such as the molecules Bhlhb5 \citep{joshi_bhlhb5_2008} and Lmo4
-\citep{cederquist_lmo4_2013,greig_molecular_2013}. These expression patterns
-can therefore provide a theoretical mechanism for the guidance of axons into
-distinct regions.
-%The question of how such structures are generated becomes
-%one of of how gene transcriptional interactions and mRNA diffusion and
-%movement achieves this (\cmnt{Fixme: find a ref on structure in gene
-%  transcription}).
-However, complex and stable structure in gene expression data is not commonly
-reported; monotonically varying expression patterns, such as those for Pax6,
-Emx2 \citep{shimogori_fibroblast_2005,bishop_regulation_2000},
-Sp8~\citep{sahara_sp8_2007} and Coup-tf1~\citep{borello_sp8_2014}, which
-display graded expression across the entire isocortex are more
-typical. Furthermore, the complex structure observed in the whisker barrel
-field has not, to our knowledge, been linked to a set of gene expression
-patterns matching the scale of the barrels. To investigate how the whisker
-barrel structure might form in the presence of only monotonically varying
-morphogen gradients, we turned our attention to the competition for
-connections built into the model.
-
 \subsection*{Branching competition is required to form somatotopically ordered barrels}
 
-There exists an element of competition in the model system specified by
-Eqs.~\ref{eq:dc} \& \ref{eq:da}. The term $\left(1 - \sum_{i=1}^{N}
-c_i(\mb{x}, t)\right)$ in Eq.~\ref{eq:dc} ensures that once the available
-synaptic receptors in a region have been fully occupied, say by
-thalamocortical type $i=1$, there is no longer the possibility for type $i=2$
-to form any synapses. This competition turns out to be relatively weak,
-because it does not strongly affect the branching density in that region. In
-the model, it is possible for two or more TC axon types to have roughly equal
-branching density in some region, leading to roughly equal numbers of
-connections forming for each TC type, up until the `hard' connection limit is
-reached.
+To investigate how the whisker barrel structure might form in the presence of
+only monotonically varying morphogen gradients, we turned our attention to the
+competition for connections built into the model. There does exist an element
+of competition in the original model (Eqs.~\ref{eq:dc} \& \ref{eq:da}). The
+term $\left(1 - \sum_{i=1}^{N} c_i\right)$ in Eq.~\ref{eq:dc} ensures that
+once the available synaptic receptors in a region have been fully occupied by
+one thalamocortical axon type, there is no longer the possibility for other
+types to form synapses. This competition turns out to be relatively weak,
+because it does not strongly affect the branching density in that region; it
+is possible for two or more TC axon types to have roughly equal branching
+density in some region, leading to roughly equal numbers of connections
+forming for each TC type, up until the `hard' connection limit is reached.
 
@@
 %
-\begin{equation} \label{eq:branch_comp}
-\chi(\mb{x}, t) = - \frac{\epsilon  a_i(\mb{x}, t)}{N-1} \sum_{i' \ne i}^{N} a_{i'}(\mb{x}, t).
+\begin{equation} \label{eq:comp}
+\chi(\mb{x}, t) = - \frac{\epsilon  a_i}{N-1} \sum_{i' \ne i}^{N} f(a_{i'}).
 \end{equation}
 %
-which leads to a reduction in the rate of increase of branching density for
-axon type $i$ in regions where the branching density of \emph{other} axon
-types is high.
-
-This introduces a new parameter, $\epsilon$, which governs the strength of the
-branching competition. The factor $1/(N-1)$ ensures that the branching
-competition strength scales with the number of TC axon types in the system.
+with the choice $f(a) = a$, which leads to a reduction in the rate of increase
+of branching density for axon type $i$ in regions where the branching density
+of \emph{other} axon types is high. The new parameter $\epsilon$ governs the
+strength of the branching competition and the factor $1/(N-1)$ ensures that the
+strength scales with $N$.
 
@@
 The prime symbols indicate the use of the intermediate value of $a_i$ after it
-has been computed according to Eqs.~\ref{eq:dc}--\ref{eq:branch_comp} on the
+has been computed according to Eqs.~\ref{eq:dc}--\ref{eq:comp} on the
 basis of the system state at time $t-{\delta}t$.
@@
 \cmnt{The realistic barrel pattern experiment is enhanced by stronger
-  branching competition. This section may be required here or elsewhere.}
+  branching competition:}
 
-To increase the strength of the competition, we could take the power
-of the branching density:
-%
-\begin{equation} \label{eq:branch_comp_pow}
-\chi(\mb{x}, t) = - \frac{\epsilon  a_i}{N-1} \sum_{i' \ne i}^{N} (a_{i'})^l
-\end{equation}
-%
-with $l>1$. This competition term can quickly become excessively large, so to
-preserve the geometric shape of the function for small $a$, whilst
-representing the idea that the amount of branching competition has an upper
-bound, we pass the function through a logistic function.  This leads to the
-following branching equation:
-
-%
-\begin{equation} \label{eq:branch_comp_pow_logistic}
-\chi(\mb{x}, t) = - \frac{\epsilon a_i}{N-1} \sum_{i' \ne i}^{N}
-\frac{1}{1 + e^{-m (a_{i'})^l}} - \frac{1}{2}
-\end{equation}
-%
-$\epsilon$ represents the maximum value of the competition term for TC type
-$i$. The parameters $l$ and $m$ control the steepness and shape of the
+To increase the strength of the competition, we choose an alternative form for
+$f(a)$ in Eq.~\ref{eq:comp} which has a geometric form for small $a$: $f(a) = \frac{1}{1 + e^{-m[a]^l}} - \frac{1}{2}$. The parameters $l$ and $m$ control the steepness and shape of the
 function.
@@
 \begin{center}
-\includegraphics[width=0.7\textwidth]{../../boundaries/barreloids_haidarliu_graph.png}
+\includegraphics[width=0.5\textwidth]{../../boundaries/barreloids_haidarliu_graph.png}
 \end{center}

### 2019-11-25 paper/briefpaper/brief.tex
Work on significance statement

@@
 % Please add a significance statement to explain the relevance of your work
-\significancestatement{The mystacial whiskers of the rat are a complex,
-  ordered set of structures whose pattern is repeated several times in
-  separate locations in the brain. We contribute a mechanism by which their
-  spatially ordered pattern can be recreated in a self-organising process
-  which requires minimal genetic pre-determination.}
+\significancestatement{The mystacial whiskers of rodents are a complex,
+  spatially ordered set of structures whose pattern is repeated in several
+  locations in the brain. We contribute a self-organising mechanism by which
+  the pattern can be recreated with minimal genetic pre-determination at each
+  location; only two pairs of axon guidance gradients are required in each
+  tissue. The pattern's togography is carried by the axons which grow from one
+  tissue to another and interact variably (depending upon which whisker they
+  are associated with) with the guidance gradients in the target tissue whilst
+  competing with each other for space. We demonstrate that this mechanism
+  recreates a realistic whisker barrel field and also accounts for the results
+  of well-known gene misexpression experiments.}
 

### 2019-11-27 paper/briefpaper/brief.tex
A new figure for the paper

@@
 \begin{center}
-\includegraphics[width=0.5\textwidth]{../../boundaries/barreloids_haidarliu_graph.png}
+\includegraphics[width=\textwidth]{./MainFig.png}
 \end{center}
 %\vspace{-0em}
-\caption{Barreloid centres taken from \cite{haidarliu_size_2001} are shown as
+\caption{A. Barreloid centres taken from \cite{haidarliu_size_2001} are shown as
   black circles. It is assumed that there are two linearly increasing,
@@
   will have with the signalling gradients in the cortex. The mappings from
-  barreloid location to interaction strength are linear.}
-\label{fig:barreloids}
-\end{figure*}
-
-\begin{figure*}
-\begin{center}
-\includegraphics[width=\textwidth]{../longpaper/realistic_barrels.png}
-\end{center}
-%\vspace{-0em}
-\caption{Realistic barrel pattern}
-\label{fig:barrels}
+  barreloid location to interaction strength are linear. B. }
+\label{fig:main}
 \end{figure*}

### 2019-11-27 analysis/hondadelta.py

@@ -55,9 +55,10 @@ print ('total area: {0}'.format (totalarea))
 # And plot this longhand here:
 F1 = plt.figure (figsize=(9,8))
 
+xmax = 12000
 ax1 = F1.add_subplot(1,1,1)
 l1, = ax1.plot(t1, hondadelta, 'o', markersize=12, color=col.springgreen2, label='Honda $\delta$')
-l2, = ax1.plot((0,200000), (0.003, 0.003), '--', color=col.orange, linewidth=3, label="'good'\npattern")
+l2, = ax1.plot((0,xmax), (0.003, 0.003), '--', color=col.orange, linewidth=3, label="'good'\npattern")
 #l2, = ax1.plot(t1, edgedev, 'o', label='Edge deviation')
 #l3, = ax1.plot(t1, domarea/totalarea[0], 'go', label='Domain area proportion')
 #l5, = ax1.plot(t1, s_resid, 's', label='Summed residuals to vert. line fits')
@@ -65,9 +66,9 @@ l2, = ax1.plot((0,200000), (0.003, 0.003), '--', color=col.orange, linewidth=3, 
 #ax1.set_title ('Honda Dirichletiform measure');
 ax1.set_xlabel ('Simulation step')
 ax1.set_ylabel ('Honda $\delta$ measure')
-ax1.set_xlim ((0,200000))
+ax1.set_xlim ((0,xmax))
 ax1.set_ylim ((0,0.05))
-ax1.set_xticks ((0,100000,200000))
+ax1.set_xticks ((0,0.5*xmax,xmax))
 plt.legend()
 
 #ax2 = F1.add_subplot(1,1,1, sharex=ax1, frameon=False)
@@ -79,6 +80,6 @@ plt.legend()
 
 plt.tight_layout()
 
-plt.savefig('hondadelta.png')
+plt.savefig('hondadelta.svg', transparent=True)
 
 plt.show()

### 2019-11-27 analysis/include/load.py

@@ -192,6 +192,34 @@ def readSimDataFiles (logdir):
 
     return (x, y, t, cmatrix, amatrix, nmatrix, idmatrix, totalarea, idnames, domcentre)
 
+# Read in the positions and the guidance molecules
+def readGuidance (logdir):
+
+    # Take off any trailing directory slash
+    logdir = logdir.rstrip ('/')
+
+    # Read x and y first
+    pf = h5py.File(logdir+'/positions.h5', 'r')
+    x = np.array(pf['x']); # HDF5 dataset object converted into numpy array
+    y = np.array(pf['y']);
+
+    # Read guidance expression
+    gf = h5py.File(logdir+'/guidance.h5', 'r')
+    # Count up how many c files we have in each time point once only:
+    gklist = list(gf.keys())
+    M = 0
+    numhexes = 0
+    for k in gklist:
+        if k[0] == 'r': # rh0, rh1 etc
+            M = M + 1
+            numhexes = len(gf[k])
+
+    gmatrix = np.zeros([numhexes, M], dtype=float)
+    for m in range (0, M):
+        gmatrix[:,m] = np.array(gf['rh{0}'.format(m)]);
+
+    return (x, y, gmatrix)
+
 #
 # targ is a container of a target x,y coordinate. x and y are the
 # vectors of positions of the hexes in the hexgrid. This returns the

### 2019-11-27 analysis/include/plot.py

@@ -1,6 +1,9 @@
 #
 # A small library of plotting functions for the RD systems
 #
+# In fact, this needs to be a class, with members such as overall
+# width, height etc.
+#
 
 import numpy as np
 import matplotlib
@@ -8,155 +11,246 @@ matplotlib.use ('TKAgg', warn=False, force=True)
 from mpl_toolkits.mplot3d import Axes3D
 import matplotlib.pyplot as plt
 
-def trace (dmatrix, ix, t, title):
+class RDPlot:
+
+    # Note: Can't have multiple __init__ overloads.
+    def __init__ (self, width, height):
+        # Overall figure width and height
+        self.width = width
+        self.height = height
+
+    # Member attributes
+
+    # Should the axes be shown or not?
+    showAxes = True
+
+    # Flag to show ID names or not, for some methods
+    showNames = True
+
+    # If the graph wants a title
+    title = ''
+
+    # Default colour map
+    cmap = plt.cm.jet
+
+    # Default fontsize
     fs = 12
-    fnt = {'family' : 'DejaVu Sans',
-           'weight' : 'regular',
-           'size'   : fs}
-    matplotlib.rc('font', **fnt)
-    trace0 = dmatrix[0,ix,:] # c0
-    trace1 = dmatrix[1,ix,:] # c1
-    F1 = plt.figure (figsize=(8,8))
-    f1 = F1.add_subplot(1,1,1)
-    f1.plot (t,trace0)
-    f1.plot (t,trace1)
-    f1.set_title(title)
+    # A second fontsize
+    fs2 = 20
 
-def trace2 (nmatrix, ix, t, title):
-    fs = 12
-    fnt = {'family' : 'DejaVu Sans',
-           'weight' : 'regular',
-           'size'   : fs}
-    matplotlib.rc('font', **fnt)
-    trace0 = nmatrix[ix,:]
-    F1 = plt.figure (figsize=(8,8))
-    f1 = F1.add_subplot(1,1,1)
-    f1.plot (t,trace0)
-    f1.set_title(title)
+    def trace (self, dmatrix, ix, t, title):
+        fnt = {'family' : 'DejaVu Sans',
+               'weight' : 'regular',
+               'size'   : self.fs}
+        matplotlib.rc('font', **fnt)
+        trace0 = dmatrix[0,ix,:] # c0
+        trace1 = dmatrix[1,ix,:] # c1
+        F1 = plt.figure (figsize=(self.width,self.height))
+        f1 = F1.add_subplot(1,1,1)
[…]

### 2019-11-27 analysis/plot_guidance.py

@@ -1,0 +1,59 @@
+# To access argv and also include the include dir
+import sys
+sys.path.insert (0, './include')
+import numpy as np
+# Import data loading code
+import load as ld
+# Import MY plotting code:
+import plot as pt
+import matplotlib
+import matplotlib.pyplot as plt
+# Direct HDF5 access
+import h5py
+# My domcentres linear fit code:
+import domcentres as dc
+import os
+
+# Get target x/y hex to show trace for and the time step to show the
+# map for from the arguments:
+if len(sys.argv) < 2:
+    print('Provide logdirname on cmd line please. Optionally provide molecule index.')
+    exit(1)
+logdirname = sys.argv[1]
+
+# molecule index
+if len(sys.argv) > 2:
+    mi = int(sys.argv[2])
+else:
+    mi = 0
+
+# Read data
+(x, y, gmatrix) = ld.readGuidance (logdirname)
+
+# Plot guidance...
+shp = np.shape(gmatrix)
+print ('gmatrix shape: {0}'.format(shp))
+
+print ('basename: {0}'.format(os.path.basename(logdirname)))
+winwidth = 12
+winheight = 11
+if os.path.basename(logdirname) == '52N2M_thalguide_fgfdup':
+    winwidth = 22
+    winheight = 14.3
+pl = pt.RDPlot (winwidth, winheight)
+pl.fs = 16
+pl.fs2 = 24
+pl.showAxes = False
+pl.showNames = False
+if mi==0:
+    pl.cmap = plt.cm.Blues
+else:
+    pl.cmap = plt.cm.Reds
+
+f1 = pl.surface (gmatrix[:,mi], x, y)
+
+# Put saving into the class?
+mapname = '{0}_guide{1}.png'.format(os.path.basename(logdirname), mi)
+plt.savefig (mapname, dpi=300, transparent=True)
+
+plt.show()

### 2019-11-27 analysis/plot_singlemap.py

@@ -14,13 +14,6 @@ import h5py
 import domcentres as dc
 import os
 
-# Set plotting defaults
-fs = 12
-fnt = {'family' : 'DejaVu Sans',
-       'weight' : 'regular',
-       'size'   : fs}
-matplotlib.rc('font', **fnt)
-
 # Get target x/y hex to show trace for and the time step to show the
 # map for from the arguments:
 if len(sys.argv) < 2:
@@ -28,10 +21,16 @@ if len(sys.argv) < 2:
     exit(1)
 logdirname = sys.argv[1]
 
+# time index
 if len(sys.argv) > 2:
     ti = int(sys.argv[2])
 else:
     ti = -1
+
+shownames = 0
+if len(sys.argv) > 3:
+    shownames = int(sys.argv[3])
+
 
 # Read the data
 (x, y, t, cmatrix, amatrix, nmatrix, idmatrix, tarea, idnames, domcentres) = ld.readSimDataFiles (logdirname)
@@ -39,15 +38,31 @@ else:
 idstring = 'id{0}'.format(0);
 print ('idstring: {0}'.format(idstring))
 
-# Plot one of the a matrices:
+# Plot one of the matrices:
 shp = np.shape(idmatrix)
 print ('idmatrix shape: {0}'.format(shp))
 if ti == -1:
     ti = shp[1]-1
 print ('ti = {0}'.format(ti))
-f1 = pt.surface_withnames (idmatrix[:,ti], x, y, 0, idstring, idnames, domcentres[ti,:,:])
 
-mapname = '{0}_{1}'.format(os.path.basename(logdirname), ti)
-plt.savefig(mapname)
+print ('basename: {0}'.format(os.path.basename(logdirname)))
+winwidth = 12
+winheight = 11
+if os.path.basename(logdirname) == '52N2M_thalguide_fgfdup':
+    winwidth = 22
+    winheight = 14.3
+pl = pt.RDPlot (winwidth, winheight)
+pl.fs = 16
+pl.fs2 = 24
+pl.showAxes = False
+pl.showNames = False
+if shownames > 0:
+    pl.showNames = True
 
-plt.show()
+f1 = pl.surface_withnames (idmatrix[:,ti], x, y, 0, idstring, idnames, domcentres[ti,:,:])
+
+# Put saving into the class?
+mapname = '{0}_{1}.png'.format(os.path.basename(logdirname), ti)
+plt.savefig (mapname, dpi=300, transparent=True)
+
+#plt.show()

### 2019-11-27 boundaries/barreloids_haidarliu.py

@@ -146,7 +146,7 @@ g3qui = np.array([])
 g4qui = np.array([])
 xqui = np.array([])
 yqui = np.array([])
-small_size = ["E6","E7","E8","E9","E9","E10","E11","E12","E13","D6","D7","D8","D9","D9","D10","D11","D12","C10"]
+small_size = ["E6","E7","E8","E9","E9","E10","E11","E12","E13","D6","D7","D8","D9","D9","D10","D11","D12","C10","C9","C8","C7","C6"]
 for d in D: # Care, requires that there are no duplicate labels
     if d == 'a':
         ax0.text (D[d][0]+txt_xoff, D[d][1]+txt_yoff, r'$\alpha$')
@@ -193,29 +193,30 @@ ap[:,0] = x__
 ap[:,1] = y__
 ax0.plot (ap[:,0], ap[:,1], c='grey', marker='None', linestyle='-.', linewidth=2)
 
+fs2 = 42
 # Anterior
-txt1 = [-0.08, -0.01]
+txt1 = [-0.09, -0.06]
 x__ = txt1[0] * cosphi - txt1[1] * sinphi
 y__ = txt1[1] * cosphi + txt1[0] * sinphi
-ax0.text (x__,y__,'ant.')
+ax0.text (x__,y__,'ant.',fontsize=fs2)
 
 # Posterior
 txt2 = [0.36, 0.12]
 x__ = txt2[0] * cosphi - txt2[1] * sinphi
 y__ = txt2[1] * cosphi + txt2[0] * sinphi
-ax0.text (x__,y__,'post.')
+ax0.text (x__,y__,'post.',fontsize=fs2)
 
 # Lateral by the stragglers 0.14v-0.1
-txt3 = [0.14, -0.125]
+txt3 = [0.12, -0.135]
 x__ = txt3[0] * cosphi - txt3[1] * sinphi
 y__ = txt3[1] * cosphi + txt3[0] * sinphi
-ax0.text (x__,y__,'lat.')
+ax0.text (x__,y__,'lat.',fontsize=fs2)
 
 # Medial
 txt4 = [0.165, 0.4]
 x__ = txt4[0] * cosphi - txt4[1] * sinphi
 y__ = txt4[1] * cosphi + txt4[0] * sinphi
-ax0.text (x__,y__,'med.')
+ax0.text (x__,y__,'med.',fontsize=fs2)
 
 # Plot voronoi
 vpts = np.vstack((xqui,yqui)).T
[…]

### 2019-11-28 paper/briefpaper/brief.tex
Now have a _medium_ length paper, mostly finished

@@
 While complex structure in protein expression can be observed for some genes,
-such as the molecules Bhlhb5 \citep{joshi_bhlhb5_2008}, a complex gene
-expression pattern with the complexity of the barrel field has not, to our
-knowledge been reported and so the mechanism by which the barrels form seems
-likely to be one in which the structure forms by self-organization.
+such as the molecules Bhlhb5 \citep{joshi_bhlhb5_2008}, we are not aware of
+any report of gene expression patterns with the complexity of the barrel field
+and so the mechanism by which the barrels form seems likely to be one in which
+the structure forms by self-organization.
 
@@
 tools for imposing arbitrary boundary shapes on the lattice. The dynamics were
-numerically integrated using fourth-order Runge-Kutta, and using a
-{\color{red} finite difference method} to compute the flux term (see
-Supplementary Materials).
+numerically integrated using fourth-order Runge-Kutta (with fixed time-step
+${\delta}t$), and using Gauss's Theorem to compute the flux term.
 
 \section*{Results}
+
+\begin{figure*}
+\begin{center}
+\includegraphics[width=\textwidth]{./MainFig.png}
+\end{center}
+%\vspace{-0em}
+\caption{Simulated barrel fields. \textbf{A.} Barreloid centres taken from
+  \cite{haidarliu_size_2001} are shown as black circles. Barreloids are
+  labelled according to the usual convention. Labels R1 to R5 are the rhinal
+  whisker barreloids. The dashed grey line represents the boundary of the
+  barreloid field; the dot-dash line shows the anterior-posterior axis. It is
+  assumed that there are two linearly increasing, orthogonal signalling
+  gradients (molecules T1 \& T2) in the thalamus (side graphs), such that the
+  location of each barreloid within the graded thalamic signal field can
+  differentiate the efferent axons, with the result that axons from each
+  barreloid interact to varying degrees with cortical signalling gradients
+  (shown in panel B). Coloured arrows show, for each barreloid, the magnitude
+  (and sign) of the interaction that its thalamocortical projection will have
+  with the signalling gradients in the cortex. The mappings from barreloid
+  location to interaction strength are linear. \textbf{B.} Guidance molecules
+  in the simulated cortex. Molecule 1 is expressed high to low in an
+  approximately anterior to posterior gradient. Molecule 2 is expressed
+  lateral to medial (analogously to ephrin - Vanderhaegen et al) \textbf{C.}
+  Whisker barrels from \cite{shimogori_fibroblast_2005}, with annotations for
+  barrel identity and approximate axes.  \textbf{D-F.} Barrel cortex
+  simulation. The maps indicate which thalomocortical axons make maximum
+  connection density. Each hex element is coloured one of 52 unique colours
+  depending on which e thalamocortical axon makes the most connections in that
+  element. The mapping between barrel identity and colour is given in F.
+  \textbf{G.} The Honda Voronoi measure as a function of simulation time for
+  results shown in D--F. \textbf{H.}  Guidance molecules in the simulated Fgf8
+  misexpression experiment. Molecule 1 is mirrored over the anterior-posterior
+  axis \textbf{I.} The simulation indicates that the mirroring of the barrel
+  field can occur as a result of the mirroring of guidance molecule 1 and the
+  field boundary.}
+\label{fig} % There's only one fig.
+\end{figure*}
 
@@
 
-To investigate how the whisker barrel structure might form in the presence of
-only monotonically varying morphogen gradients, we turned our attention to the
-competition for connections built into the model. There does exist an element
-of competition in the original model (Eqs.~\ref{eq:dc} \& \ref{eq:da}). The
-term $\left(1 - \sum_{i=1}^{N} c_i\right)$ in Eq.~\ref{eq:dc} ensures that
-once the available synaptic receptors in a region have been fully occupied by
-one thalamocortical axon type, there is no longer the possibility for other
-types to form synapses. This competition turns out to be relatively weak,
-because it does not strongly affect the branching density in that region; it
-is possible for two or more TC axon types to have roughly equal branching
-density in some region, leading to roughly equal numbers of connections
-forming for each TC type, up until the `hard' connection limit is reached.
+To investigate how the whisker barrel structure (Fig.~\ref{fig}C) might
+form in the presence of only monotonically varying morphogen gradients, we
+turned our attention to the competition for connections built into the
+model. There \emph{does} exist an element of competition in the original model
+(Eqs.~\ref{eq:dc} \& \ref{eq:da}). The term $\left(1 - \sum_{i=1}^{N}
+c_i\right)$ in Eq.~\ref{eq:dc} ensures that once the available synaptic
+receptors in a region have been fully occupied by one thalamocortical axon
+type, there is no longer the possibility for other types to form
+synapses. This competition turns out to be relatively weak, because it does
+not strongly affect the branching density in that region; it is possible for
+two or more TC axon types to have roughly equal branching density in some
+region, leading to roughly equal numbers of connections forming for each TC
+type, up until the `hard' connection limit is reached.
 
@@
 \begin{equation} \label{eq:comp}
-\chi(\mb{x}, t) = - \frac{\epsilon  a_i}{N-1} \sum_{i' \ne i}^{N} f(a_{i'}).
+\chi(\mb{x}, t) = - \frac{\epsilon  a_i}{N-1} \sum_{i' \ne i}^{N} a_{i'}
 \end{equation}
 %
-with the choice $f(a) = a$, which leads to a reduction in the rate of increase
-of branching density for axon type $i$ in regions where the branching density
+%with the choice $f(a) = a$,
+which leads to a reduction in
+%the rate of increase of branching density for axon type $i$
+%$\frac{\partial a_i}{\partial t}$
+${\partial a_i}/{\partial t}$
+in regions where the branching density
 of \emph{other} axon types is high. The new parameter $\epsilon$ governs the
@@
 
-With the addition of the competition term, $\chi$, it is necessary to
-introduce a normalisation step, in order to prevent one TC type from
-out-competing every other TC type. Normalisation represents the idea that for
-a fixed number of afferent thalamocortical axons, the overall number of
-branches must have some upper limit; given the finite resources available in
-each axon for building new processes. We divisively normalize the branching
-density, dividing it at each computational step by the mean branching density
-per unit area. It is then multiplied by the branching density per unit area at
-time 0, to account for the possibility that there might be more axons afferent
-for TC type $1$ cf.~TC type $2$:
+With the addition of branching competition, it is necessary to introduce a
+normalisation step, in order to prevent one TC type from out-competing every
+other type. This represents the idea that for a fixed number of afferent
+thalamocortical axons, the overall number of branches must have some upper
+limit, given the finite resources available in each cell for building new
+processes. We divisively normalize the branching density, dividing it at each
+computational step by the mean branching density per unit area. It is then
+multiplied by the branching density per unit area at time 0, which preserves
+the effect of any disparity in the numbers of axons of each type which
+innervate the cortical sub-plate:
 %
@@
 The prime symbols indicate the use of the intermediate value of $a_i$ after it
-has been computed according to Eqs.~\ref{eq:dc}--\ref{eq:comp} on the
-basis of the system state at time $t-{\delta}t$.
+has been computed according to Eqs.~\ref{eq:dc}--\ref{eq:comp} on the basis of
+the system state at time $t-{\delta}t$. Integrals over $\Omega$ denote a
+spatial integral over the domain of interest.
 
@@
 
-In which we show that a realistic barrel pattern in the thalamus can be
-recreated in barrel cortex.
+With these changes we were able to generate grid-like patterns with
+appropriate (linearly spaced) choices of the interaction parameters,
+$\gamma$. To determine the $\gamma$ values suitable to recreate the cortical
+barrel field by a principled method, we postulated that the interaction of
+each thalamocortical axon growth cone with the ephrin gradients in the
+cortical sub-plate might be determined by the location of the \emph{barreloid}
+containing the axon's soma within signalling gradents \emph{in the
+  thalamus}.
+%That is, that interaction is grouped by barreloid identity.
+%
+%We therefore sought a map of the thalamic barreloids from the
+%literature.
+Ref.~\cite{haidarliu_size_2001} provided a comprehensive map of the
+layout of the barreloids in rat, which is similar to mouse
+\cite{van_der_loos_barreloids_1976}. We used these data to create the map of
+52 barreloids shown in Fig.~\ref{fig}A. We postulated two linearly
+varying molecules and used the barreloids' positions within this field to
+generate the interaction parameters represented by the blue and red arrows.
 
-\cmnt{The realistic barrel pattern experiment is enhanced by stronger
-  branching competition:}
+We configured a simulation based on Eqs.~\ref{eq:dc}--\ref{eq:norm} to run on
+a domain whose shape was matched to data from \cite{shimogori_fibroblast_2005}
+(Fig.~\ref{fig}C). 52 separate axon branching fields were initialised with
+random noise and given the $\gamma$ parameters from Fig.~\ref{fig}A. Two
+orthogonal gradients were specified; molecule 1 with a posterior to anterior
+gradient and molecule 2 with a medial-lateral gradient (resembling ephrin
+A5). Fig.~\ref{fig}D--E shows snapshots of the connection density at
+simulation times 0.08, 0.2 and 0.32 (${\delta}t=10^{-4}$), with colour
+indicating the identity of the barreloid whose axons make maximum connections
+at each location and \cmnt{contours showing where connection density passes
+  0.5, indicating that each axon group is well localized}. A pattern emerges
+which strongly resembles the barrel cortex in Fig.~\ref{fig}C, with `simulated
+barrels' forming for whisker rows A to E, along with a suitably located group
+for the rhinal whiskers.
+%
+Fig.~\ref{fig}G shows (using the Honda measure) that the simulated mechanism
+does generate Voronoi patterns, in agreement with
+\cite{senft_mouse_1991}.
 
-To increase the strength of the competition, we choose an alternative form for
-$f(a)$ in Eq.~\ref{eq:comp} which has a geometric form for small $a$: $f(a) = \frac{1}{1 + e^{-m[a]^l}} - \frac{1}{2}$. The parameters $l$ and $m$ control the steepness and shape of the
-function.
-
+% On stronger branching; not used in the results presented here.
+% \cmnt{The realistic barrel pattern experiment is enhanced by stronger
+%  branching competition:}
+%
+% To increase the strength of the competition, we choose an alternative form for
+% $f(a)$ in Eq.~\ref{eq:comp} which has a geometric form for small $a$: $f(a) =
+% \frac{1}{1 + e^{-m[a]^l}} - \frac{1}{2}$. The parameters $l$ and $m$ control
+% the steepness and shape of the function.
 %%% End of stronger branching competition %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\begin{figure*}
-\begin{center}
-\includegraphics[width=\textwidth]{./MainFig.png}
-\end{center}
-%\vspace{-0em}
-\caption{A. Barreloid centres taken from \cite{haidarliu_size_2001} are shown as
-  black circles. It is assumed that there are two linearly increasing,
-  orthogonal signalling gradients (molecules A \& B) in the thalamus (side
-  graphs), such that the location of each barreloid within the graded thalamic
-  signal field can differentiate the efferent axons, with the result that
-  axons from each barreloid interact to varying degrees with cortical
-  signalling gradients. Coloured arrows show, for each barreloid, the
-  magnitude (and sign) of the interaction that its thalamocortical projection
-  will have with the signalling gradients in the cortex. The mappings from
-  barreloid location to interaction strength are linear. B. }
-\label{fig:main}
-\end{figure*}
 
@@
 
-To cover the reaction to changes in the guidance molecules caused by
-experiments like those by Shimogori \& Grove.
+A plausible theory of barrel patterning should reproduce the results of
+experimental whisker barrel manipulations.
+%
+Here, we show that our mechanism reproduces the mirrored whisker barrel fields
+reported in \cite{shimogori_fibroblast_2005,assimacopoulos_fibroblast_2012}
+when an additional rostro-caudal source of Fgf8 is caused by
+electroporation. We assume that a mechanism separate to ours governs the
+definition of the boundary within which the barrel field will form and that
+the additional Fgf8 source causes the boundary to become mirrored, with an
+additional, posterior region forming. We also assume that under this
+manipulation, the anteror-posterior guidance molecule (1) takes on a 'v'
+shaped profile, with high expression at the far anterior and posterior of the
+double barrel field domain, and low expression around the centre
+(Fig.~\ref{fig}H). We assume molecule 2 is unchanged in its expression
+pattern. Fig.~\ref{fig}I shows the pattern which forms with the simulation of
+52 TC types having the same $\gamma$ parameters as set for the original
+simulation (i.e.~from Fig.~\ref{fig}A). Similar, mirrored patterns form in
+each side of the expanded domain. Note that for many of the 52 axon types, 2
+fields form in spatially distinct locations (selected examples are shown),
+though the straggler barrels $\alpha$ to $\delta$ and a few others (incl.~A1,
+C1) appear in only single locations.
 
@@
 We set out to test the hypothesis that barrel patterning may emerge via
-self-organising processes. We first extended an existing model of neocortical
-arealization to simulate pattern formation in two dimensions, inheriting from
-this model the assumption that axonal branching and synaptogenesis are
-coupled, and that thalamic projections compete for cortical territory under
-the influence of molecular signalling gradients. Realistic barrel patterning
-emerged only when terms were added to represent the assumption that
-thalamocortical projections compete for a limited supply of connections with
-cortical cells, that their branching rates are reduced by the branching of
-other thalamocortical projections, and that the density of axon branching is
-conserved for each projection. The net effect of these local interactions is a
-Voronoi tesselation in which localised clusters of thalamic projections become
-somatotopically organised on the cortical sheet according to the relative
-values of coefficients specifying the rates at which axons are attracted and
-repelled by orthogonal pairs of molecular signal gradients. According to this
-model, the pattern of barrels is determined by the competitive and cooperative
-interactions between cells relative to the shape of the cortical field
-boundary. In simulation, we systematically investigated the origin of the
-asymmetry between the barrel rows and arcs, and identified {\color{red} XXXXX}
-as the most likely candidate for the elongation of barrel shapes along the
-rows.
+self-organising processes.
+%
+We first extended an existing model of neocortical arealization to simulate
+pattern formation in two dimensions, inheriting from this model the assumption
+that axonal branching and synaptogenesis are coupled, and that thalamic
+projections compete for cortical territory under the influence of molecular
+signalling gradients.
+%
+Realistic barrel patterning emerged only when terms were added to represent
+the assumption that thalamocortical projections compete for a limited supply
+of connections with cortical cells, that their branching rates are reduced by
+the branching of other thalamocortical projections, and that the density of
+axon branching is conserved for each projection.
+%
+The net effect of these local interactions is a Voronoi tesselation in which
+localised clusters of thalamic projections become somatotopically organised on
+the cortical sheet according to the relative values of coefficients specifying
+the rates at which axons are attracted or repelled by orthogonal molecular
+signal gradients.
+%
+According to this model, the pattern of barrels is determined by the
+competitive and cooperative interactions between cells relative to the shape
+of the cortical field boundary. By computing this simulation on a realistic
+barrel field, we provide an existence proof that the somatotopic order of the
+barrels can be transferred without the need for absolute positional
+information on their centre-point locations. Instead, relative interaction
+information, extracted from the source tissue of the afferent axons (the
+thalamus) can reproduce the somatotopy regardless of changes to the boundary
+shape or size. A similar process could enable the transfer of peripheral
+pattern information (the ultimate source of the pattern?) to the barrelettes
+in the brainstem and from the barrelettes to the ventro basal thalamus.
+
+The requirement for two orthogonal guidance gradients is easy to understand;
+this provides a coordinate system to differentiate locations within the
+cortical sub-plate. It is interesting to note that in order to successfully
+reproduce the barrel patterns of Fig.~\ref{fig}C, we found it necessary for
+axon types to experience \emph{both} positive \emph{and} negative interactions
+with the two gradients. If only positive interactions are permitted (that is
+growth cones only move up the gradient) then the pattern is heavily skewed
+towards one side of the domain.
+%
+If empirical data showed that only positive (or negative) interactions were
+possible, then by having \emph{four} guidance gradients, as two, orthogonal,
+opposing pairs, the same, balancing effect could be achieved. Such a pattern
+of opposing pairs of expression gradients is known to exist in the cortical
+transcription factors that are related to Fgf8.
+
+% Can of worms?
+We think of this process as taking place in the cortical sub-plate and have
+reported that the pattern resembles a Voronoi diagram. Anatomical data
+suggests that the somatotopy is determined within the sub-plate. However, the
+actual barrel pattern forms in layer IV of the cortex, to which axons are seen
+to rise, and so the final Voronoi layout of the barrels is likely to be
+influenced by mechanisms taking place within layer IV. This may explain why,
+as the Voronoi measure of the simulated barrel pattern improves, the
+correspondence between the simulated, and measured barrel layout diverges.
+
+% Anything to say about the misexpression experiment?
+
+\subsection*{Supplementary Material}
+
+The code and configuration required to reproduce the results in this paper are
+available at \url{https://github.com/ABRG-Models/BarrelEmerge}.
 

### 2019-11-28 paper/supplementary/supp.tex
Adds the lab nodes/rd_karbowski document for modification into a supplementary text

% This is an example of using latex for a paper/report of specified
% size/layout. It's useful if you want to provide a PDF that looks
% like it was made in a normal word processor.

% While writing, don't stop for errors
\nonstopmode

% Use the article doc class, with an 11 pt basic font size
\documentclass[11pt, a4paper]{article}

% Makes the main font Nimbus Roman, a Times New Roman lookalike:
%\usepackage{mathptmx}% http://ctan.org/pkg/mathptmx
% OR use this for proper Times New Roman (from msttcorefonts package
% on Ubuntu). Use xelatex instead of pdflatex to compile:
\usepackage{fontspec}
\usepackage{xltxtra}
\usepackage{xunicode}
\defaultfontfeatures{Scale=MatchLowercase,Mapping=tex-text}
\setmainfont{Times New Roman}

% Set margins
\usepackage[margin=2.5cm]{geometry}

% Multilingual support
\usepackage[english]{babel}

% Nice mathematics
\usepackage{amsmath}

% Left right harpoons for kinetic equations
\usepackage{mathtools}

% Control over maketitle
\usepackage{titling}

% Section styling
\usepackage{titlesec}

% Ability to use colour in text
\usepackage[usenames]{color}

% For the \degree symbol
\usepackage{gensymb}

% Allow includegraphics and nice wrapped figures
\usepackage{graphicx}
\usepackage{wrapfig}
\usepackage[outercaption]{sidecap}

% Set formats using titlesec
\titleformat*{\section}{\bfseries\rmfamily}
\titleformat*{\subsection}{\bfseries\itshape\rmfamily}

% thetitle is the number of the section. This sets the distance from
% the number to the section text.
\titlelabel{\thetitle.\hskip0.3em\relax}

% Set title spacing with titlesec, too.  The first {1.0ex plus .2ex
% minus .7ex} sets the spacing above the section title. The second
% {-1.0ex plus 0.2ex} sets the spacing the section title to the
% paragraph.
\titlespacing{\section}{0pc}{1.0ex plus .2ex minus .7ex}{-1.1ex plus 0.2ex}

%% Trick to define a language alias and permit language = {en} in the .bib file.
% From: http://tex.stackexchange.com/questions/199254/babel-define-language-synonym
\usepackage{letltxmacro}
\LetLtxMacro{\ORIGselectlanguage}{\selectlanguage}
\makeatletter
\DeclareRobustCommand{\selectlanguage}[1]{%
  \@ifundefined{alias@\string#1}
    {\ORIGselectlanguage{#1}}
    {\begingroup\edef\x{\endgroup
       \noexpand\ORIGselectlanguage{\@nameuse{alias@#1}}}\x}%
}
\newcommand{\definelanguagealias}[2]{%
  \@namedef{alias@#1}{#2}%
}
\makeatother
\definelanguagealias{en}{english}
\definelanguagealias{eng}{english}
%% End language alias trick

%% Any aliases here
\newcommand{\mb}[1]{\mathbf{#1}} % this won't work?
% Emphasis and bold.
\newcommand{\e}{\emph}
\newcommand{\mycite}[1]{\cite{#1}}
\newcommand{\code}[1]{\textsf{#1}}
\newcommand{\dvrg}{\nabla\vcdot\nabla}
%% END aliases

% Custom font defs
% fontsize is \fontsize{fontsize}{linespacesize}
\def\authorListFont{\fontsize{11}{11} }
\def\corrAuthorFont{\fontsize{10}{10} }
\def\affiliationListFont{\fontsize{11}{11}\itshape }
\def\titleFont{\fontsize{14}{11} \bfseries }
\def\textFont{\fontsize{11}{11} }
\def\sectionHdrFont{\fontsize{11}{11}\bfseries}
\def\bibFont{\fontsize{10}{10} }
\def\captionFont{\fontsize{10}{10} }

% Caption font size to be small.
\usepackage[font=small,labelfont=bf]{caption}

% Make a dot for the dot product, call it vcdot for 'vector calculus
% dot'. Bigger than \cdot, smaller than \bullet.
\makeatletter
\newcommand*\vcdot{\mathpalette\vcdot@{.35}}
\newcommand*\vcdot@[2]{\mathbin{\vcenter{\hbox{\scalebox{#2}{$\m@th#1\bullet$}}}}}
\makeatother

\def\firstAuthorLast{James}

% Affiliations
\def\Address{\\
\affiliationListFont Adaptive Behaviour Research Group, Department of Psychology,
  The University of Sheffield, Sheffield, UK \\
}

% The Corresponding Author should be marked with an asterisk. Provide
% the exact contact address (this time including street name and city
% zip code) and email of the corresponding author
\def\corrAuthor{Seb James}
\def\corrAddress{Department of Psychology, The University of Sheffield,
  Western Bank, Sheffield, S10 2TP, UK}
\def\corrEmail{seb.james@sheffield.ac.uk}

% Figure out the font for the author list..
\def\Authors{\authorListFont Sebastian James\\[1 ex]  \Address \\
  \corrAuthorFont $^{*}$ Correspondence: \corrEmail}

% No page numbering please
\pagenumbering{gobble}

% A trick to get the bibliography to show up with 1. 2. etc in place
% of [1], [2] etc.:
\makeatletter
\renewcommand\@biblabel[1]{#1.}
\makeatother

% reduce separation between bibliography items if not using natbib:
\let\OLDthebibliography\thebibliography
\renewcommand\thebibliography[1]{
  \OLDthebibliography{#1}
  \setlength{\parskip}{0pt}
  \setlength{\itemsep}{0pt plus 0.3ex}
}

% Set correct font for bibliography (doesn't work yet)
%\renewcommand*{\bibfont}{\bibFont}

% No paragraph indenting to match the VPH format
\setlength{\parindent}{0pt}

% Skip a line after paragraphs
\setlength{\parskip}{0.5\baselineskip}
\onecolumn

% titling definitions
\pretitle{\begin{center}\titleFont}
\posttitle{\par\end{center}\vskip 0em}
\preauthor{ % Fonts are set within \Authors
        \vspace{-1.1cm} % Bring authors up towards title
        \begin{center}
        \begin{tabular}[t]{c}
}
\postauthor{\end{tabular}\par\end{center}}

% Define title, empty date and authors
\title {
  Karbowski model in 2-D
}
\date{} % No date please
\author{\Authors}

%% END OF PREAMBLE

\begin{document}

\setlength{\droptitle}{-1.8cm} % move the title up a suitable amount
\maketitle

\vspace{-1.8cm} % HACK bring the introduction up towards the title. It
                % would be better to do this with titling in \maketitle

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}

This is a working document starting from the model described in Eqs.~2
and 4 in ``Model of the Early Development of Thalamo-Cortical
Connections and Area Patterning via Signaling Molecules'',
Karbowski \& Ermentrout, 2004~\cite{karbowski_model_2004}. I then
consider extending this model into two dimensions.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{One dimensional system}

Equations 1-4 of \cite{karbowski_model_2004} is the full set of
differential equations describing their system:

\begin{equation} \label{eq:Karb1D_dc}
\frac{\partial c_i(x, t)}{\partial t} = -\alpha_i c_i(x, t) + \beta_i
n(x, t)[a_i(x, t)]^k
\end{equation}
%
\begin{equation} \label{eq:Karb1D_da}
\frac{\partial a_i(x, t)}{\partial t} = \frac{\partial J_i(x, t)}{\partial
x} + \alpha_i c_i(x, t) - \beta_i
n(x, t)[a_i(x, t)]^k
\end{equation}
%
\begin{equation} \label{eq:Karb1D_conserve}
n(x, t) + \sum_{i=1}^{N} c_i(x, t) = 1
\end{equation}
%
with the flux current
%
\begin{equation} \label{eq:Karb1D_J}
J_i(x, t) = D \frac{\partial a_i(x, t)}{\partial x} - a_i
\bigg(\gamma_{Ai} \frac{\partial \rho_A(x)}{\partial x} +\gamma_{Bi} \frac{\partial \rho_B(x)}{\partial x} + \gamma_{Ci} \frac{\partial \rho_C(x)}{\partial x} \bigg)
\end{equation}

Note that (unlike in the original paper) I have been explicit about
the space and time dependence of the state variables $a_i$, $c_i$,
$J_i$, $n$, $\rho_A$, $\rho_B$ and $\rho_C$. Note also that
$i$ \emph{always} refers to the thalamo-cortial connection type; there
is one equation system for each connection type. Keep this in mind
when reading the sums in the maths below. Note finally that I've not
defined all the terms in this working document, but I've used the same
variable names as in the paper~\cite{karbowski_model_2004}.

By inspection of Eqs.~\ref{eq:Karb1D_dc} and \ref{eq:Karb1D_da}, it
can be seen that Eq.~\ref{eq:Karb1D_da} can also be written:
%
\begin{equation} \label{eq:Karb1D_da_alt}
\frac{\partial a_i(x, t)}{\partial t} = \frac{\partial J_i(x, t)}{\partial
x} - \frac{\partial c_i(x, t)}{\partial t}
\end{equation}
%
Both forms of the equation are useful; Eq.~\ref{eq:Karb1D_da} is
closer to the way the numerical solution is found in the code, in
which a Runge-Kutta computation is carried out first for $a$ and then
for $c$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Boundary condition}

Karbowski and Ermentrout apply the `sealed-end' boundary condition (a
term that I've found used in cable theory):
%
$J(0,t) = J(L,t) = 0$. $x$ is in the range $[0,L]$.

\begin{equation}
J_i(x, t) \bigg\rvert_{boundary} = D \frac{\partial a_i(x, t)}{\partial x} - a_i
\bigg(\gamma_{Ai} \frac{\partial \rho_A(x)}{\partial x}
+\gamma_{Bi} \frac{\partial \rho_B(x)}{\partial x}
+ \gamma_{Ci} \frac{\partial \rho_C(x)}{\partial x} \bigg) \bigg\rvert_{boundary} = 0
\end{equation}
%
which is a `Robin boundary condition'; a linear combination of the
derivative of $a_i$ and the value $a_i$. More specifically, it's an
`insulating boundary condition' (because it is set to 0) (from the
diffusion-convection equation world).

K \& E claim
that requiring $J(0)=J(L)=0$ ``implies that the total number of axonal
branches and connections of any given type in the system is
constant'' and that this statement is equivalent to saying that
%
\begin{equation}
\int_{0}^{L} [a_i(x,t) + c_i(x,t)] dx
\end{equation}
%
is not a function of time. To show this, start from
Eq.~\ref{eq:Karb1D_da_alt}:
%
\begin{equation}
\frac{\partial a_i}{\partial t} = \frac{\partial J_i}{\partial
x} - \frac{\partial c_i}{\partial t}
\end{equation}
%
\begin{equation}
\begin{split}
\implies \frac{\partial J_i}{\partial x} & =
\frac{\partial a_i}{\partial t}
+ \frac{\partial c_i}{\partial t} \\
& = \frac{\partial}{\partial t}(a_i + c_i)
\end{split}
\end{equation}
%
(because differentiation is linear). So, integrate this system
spatially from $0$ to $L$:
%
\begin{equation}
\begin{split}
\int_0^L \frac{\partial J_i}{\partial x} dx & =
\int_0^L  \frac{\partial}{\partial t}(a_i + c_i) dx \\
\implies \int_0^L \partial J_i & = \int_0^L  \frac{\partial}{\partial
t}(a_i + c_i) dx \\
\implies J(L) - J(0) & = \frac{\partial}{\partial t} \int_0^L (a_i + c_i) dx
\end{split}
\end{equation}
%
because $\int_0^L \partial J_i $ is just the sum of the changes in the
value of $J$, which is equal to the difference between the final value
and the initial value. Finally, because our no-flux boundary
conditions state that $J(0) = J(L) = 0$, then
%
\begin{equation}
 \frac{\partial}{\partial t} \int_0^L (a_i + c_i) dx = 0
\end{equation}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Two dimensional system}

Extending the system above to two spatial dimensions changes $x$ into
a two-dimensional $\mb{x}$ and changes the flux term, $J_i(x,t)$ as follows:
%
\begin{equation} \label{eq:dc}
\frac{\partial c_i(\mb{x},t)}{\partial t} = -\alpha_i c_i(\mb{x},t) + \beta_i n(\mb{x},t)
[a_i(\mb{x},t)]^k
\end{equation}
%
\begin{equation} \label{eq:da}
\frac{\partial a_i(\mb{x},t)}{\partial t}
= \nabla\vcdot\mb{J}_i(\mb{x},t) - \frac{\partial c_i(\mb{x},t)}{\partial t}
\end{equation}
%
\begin{equation} \label{eq:conserve}
n(\mb{x},t) + \sum_{i=1}^{N} c_i(\mb{x}, t) = 1
\end{equation}
%
with the flux current
%
\begin{equation} \label{eq:J}
\mb{J}_i(\mb{x},t) = D \nabla a_i(\mb{x},t) - a_i
\big(\gamma_{Ai} \nabla\rho_A(\mb{x}) +\gamma_{Bi} \nabla\rho_B(\mb{x}) + \gamma_{Ci} \nabla\rho_C(\mb{x}) \big)
\end{equation}

The boundary condition that I'd like to apply to these equations is:
%
\begin{equation}
\mb{J}_i(\mb{x},t) \bigg\rvert_{boundary} = 0
\end{equation}
%
to agree with the boundary condition applied in the paper for the 1D system.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Computing div J: approach 1}

In order to compute $\nabla\vcdot\mb{J}_i(\mb{x},t)$, I need to do
a little work. First, I point out that
%
\begin{equation}
\mb{g}(\mb{x}) \equiv \big(\gamma_{Ai} \nabla\rho_A(\mb{x}) +\gamma_{Bi} \nabla\rho_B(\mb{x})
+ \gamma_{Ci} \nabla\rho_C(\mb{x}) \big)
\end{equation}
%
is a static vector field; once computed at the start of the
simulation, it is time-invariant. $\mb{J}_i$ is thus:
%
\begin{equation}\label{eq:J_again}
\mb{J}_i(\mb{x},t) = D \nabla a_i(\mb{x},t) - a_i \mb{g}(\mb{x})
\end{equation}

Taking the divergence of Eq.~\ref{eq:J_again}:
%
\begin{equation} \label{eq:divJ}
\nabla\vcdot\mb{J}_i(\mb{x},t) = \nabla\vcdot\bigg(D \nabla
a_i(\mb{x},t) - a_i \mb{g}(\mb{x}) \bigg)
\end{equation}

Now simply multiply the current values of $a_i$ and $\mb{g}$ and
compute an approximation to $\nabla a_i$, resulting in a vector field
which is $\mb{J}$. The boundary condition can be forced by setting
this to 0 for boundary hexes (ugly, and quite possibly completely
fallacious). Finally, compute $\nabla\vcdot\mb{J}_i(\mb{x},t)$ using
the result that divergences can be simplified by means of Gauss's
Theorem. This states that (in a two-dimensional system) the area
integral of the divergence of a vector field $\mb{f}$ is equal to the
integral around the boundary of the field dotted with the normal to
the boundary:
%
\begin{equation}
\iint_A \nabla\vcdot\mb{f}\;dxdy = \oint_c \mb{f}\vcdot d\hat{\mb{n}}
\end{equation}

See, for example \cite{george_b._arfken_mathematical_1995},
Sect.~1.11, p 58.

This can be used to find the average value of the divergence of
$\mb{f}$ over the area of a hexagon; $\mb{f}_0$ is the average value
of the vector field at the centre of the hexagon;
$\mb{f}_1$--$\mb{f}_6$ are the average values of the field in the 6
neighbouring hexagons. The contour integral can be discretized for
this hexagonal region, which has area $\Omega$ and a side length $v$:
%
\begin{equation}
\begin{split}
\frac{1}{\Omega} \oint_c \mb{f}\vcdot d\hat{\mathbf{n}} & \approx \frac{1}{\Omega} \sum_{j=1}^{6} \frac{\mb{f}_j + \mb{f}_0}{2}\vcdot \hat{\mb{n}}\;v \\
& = \frac{1}{\Omega} \sum_{j=1}^{6} \frac{\mb{f}_j^x + \mb{f}_0^x}{2} \vcdot \hat{\mb{n}}\;v +  \sum_{j=1}^{6} \frac{\mb{f}_j^y + \mb{f}_0^y}{2} \vcdot \hat{\mb{n}}\;v \\
& = \frac{1}{\Omega} \sum_{j=1}^{6} \frac{\mb{f}_j^x + \mb{f}_0^x}{2} \cos (60\times(j-1))\;v + \frac{1}{\Omega} \sum_{j=1}^{6} \frac{\mb{f}_j^y + \mb{f}_0^y}{2} \sin (60\times(j-1))\;v \\
\end{split}
\end{equation}

An advantage to using this approach is that if it is required that the
\emph{divergence} be 0 across the boundary then for those edges of the hex
that lie on the boundary $\mb{f}_j$ can be set equal to $\mb{f}_0$
(this is the ghost cell method), satisfying this boundary condition,
whilst permitting flow through the hex parallel to the
boundary. However, it is not so easy to choose a ghost cell with
properties that will force the \emph{value} of $\mb{f}_0$ to 0.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Computing div J: approach 2}

%
An alternative approach is to apply some vector calculus identities to
work out the divergence of $\mb{J}_i(\mb{x})$. Because the divergence
operator is distributive, Eq.~\ref{eq:divJ} can be expanded:
%
\begin{equation}
\nabla\vcdot\mb{J}_i(\mb{x}) = \nabla\vcdot\big(D \nabla
a_i(\mb{x},t)\big) - \nabla\vcdot\big(a_i \mb{g}(\mb{x})\big)
\end{equation}
%
$D$ is a constant and applying the vector calculus product rule
identity to $\nabla\vcdot\big(a_i \mb{g}(\mb{x})\big)$:
%
\begin{equation}
%
\nabla\vcdot\mb{J}_i(\mb{x}) =
D \dvrg a_i(\mb{x},t)
-
\big(
a_i(\mb{x},t)\nabla\vcdot\mb{g}(\mb{x})
+
\mb{g}(\mb{x})\vcdot\nabla a_i(\mb{x},t)
\big)
%
\end{equation}
%
(to prove $\nabla\vcdot D\nabla a = D \dvrg a$, apply the
product rule and note that $\nabla D = 0$ for constant $D$). Expanding
the brackets we have:
%
\begin{equation} \label{eq:divJExpanded}
%
\boxed{
\nabla\vcdot\mb{J}_i(\mb{x}) =
%
D \dvrg a_i(\mb{x},t) % term1
-
a_i(\mb{x},t)\nabla\vcdot\mb{g}(\mb{x}) % term2
-
\mb{g}(\mb{x})\vcdot\nabla a_i(\mb{x},t)} % term3
%
\end{equation}
%
which has three elements to compute: the Laplacian of
$a_i(\mb{x},t)$; a time-independent modulator of
$a_i(\mb{x},t)$ [because $\nabla\vcdot\mb{g}(\mb{x})$ is a
time-independent static field]; and the dot product of the static
vector field $\mb{g}(\mb{x})$ and the gradient of
$a_i(\mb{x},t)$. These three terms (named, very sensibly \code{term1},
\code{term2} and \code{term3}) are computed in the \code{compute\_divJ()}
method of both the \code{RD\_2D\_Karb} and \code{RD\_James} classes.

Each of the divergences can be simplified by means of Gauss's Theorem.
Reference \cite{lee_hexagonal_2014} presents an application of Gauss's
Theorem to the solution of Poisson's equation, which contains a
Laplacian term, on an hexagonal grid. The computation of the mean
value of our Laplacian, $D \dvrg a_i(\mb{x},t)$, across
the area of one hexagon located at position $\mb{p}_0$, with
neighbours at positions $\mb{p}_1$--$\mb{p}_6$ is:
%
\begin{equation} \label{eq:computeDivJTerm1}
\boxed{
D \dvrg a_i(\mb{p}_0,t) = D\frac{2}{3d^2} \sum_{j=1}^{6}\big(a_i(\mb{p}_j) - a_i(\mb{p}_0)\big)
}
\end{equation}
%
where $d$ is the centre-to-centre distance between hexes in the
grid. This is \code{term1}. Let's work it through. In the following, $v$ is the length of
each of the 6 edges of the hexagon's perimeter and $v =
d/\sqrt{3}$. $d\gamma$ is an infinitessimally small distance along the
perimeter of the hexagon. The area, $\Omega$, of each hexagon in the lattice is
$\frac{\sqrt{3}}{2}d^2$.

\begin{equation}
\begin{split}
\frac{1}{\Omega} \iint_A \dvrg a_i(\mb{x}) dxdy & = \frac{1}{\Omega} \oint_c \frac{\partial a_i}{\partial \hat{\mb{n}}} d\gamma \\
& \approx \frac{1}{\Omega} \sum_{j=1}^6 \frac{\partial a_i(\mb{p}_j)}{\partial \hat{\mb{n}}} \bigg\rvert_{mid} v \\
& = \frac{2}{\sqrt{3} d^2} \sum_{j=1}^6 \frac{a_i(\mb{p}_j) - a_i(\mb{p}_0)}{d} \frac{d}{\sqrt{3}} \\
& = \frac{2}{3 d^2} \sum_{j=1}^6 \big(a_i(\mb{p}_j) - a_i(\mb{p}_0)\big)
\end{split}
\end{equation}

Applying Gauss's Theorem to the second term, the computation of
$a_i(\mb{p}_0,t)\nabla\vcdot\mb{g}(\mb{p}_0)$ can be written out in a
similar way:
%
\begin{equation} \label{eq:divg}
\begin{split}
%
\frac{1}{\Omega} \iint_A a_i\nabla\vcdot\mb{g}\;dxdy & = \frac{a_i(\mb{p}_0)}{\Omega}  \oint_c \mb{g}\vcdot d\hat{\mathbf{n}} \\
%
& \approx \frac{a_i(\mb{p}_0,t)}{\Omega} \sum_{j=1}^{6} \frac{\mb{g}_j + \mb{g}_0}{2}\vcdot \hat{\mb{n}}\;v \\
%
& = \frac{2 a_i(\mb{p}_0,t)v}{\sqrt{3}d^2} \bigg( \sum_{j=1}^{6} \frac{\mb{g}_j^x + \mb{g}_0^x}{2} \vcdot  \hat{\mb{n}} + \sum_{j=1}^{6} \frac{\mb{g}_j^y + \mb{g}_0^y}{2} \vcdot  \hat{\mb{n}} \bigg) \\
%
& = \frac{2 a_i(\mb{p}_0,t)d}{\sqrt{3}d^2\sqrt{3}} \bigg( \sum_{j=1}^{6} \frac{\mb{g}_j^x + \mb{g}_0^x}{2} \vcdot  \hat{\mb{n}} + \sum_{j=1}^{6} \frac{\mb{g}_j^y + \mb{g}_0^y}{2} \vcdot  \hat{\mb{n}} \bigg) \\
%
& = \frac{a_i(\mb{p}_0,t)}{3d} \bigg( \sum_{j=1}^{6} \big(\mb{g}_j^x + \mb{g}_0^x\big) \cos (60\times(j-1)) + \sum_{j=1}^{6} \big({\mb{g}_j^y + \mb{g}_0^y}\big) \sin (60\times(j-1)) \bigg) \\
%
\end{split}
\end{equation}
%

\begin{equation} \label{eq:computeDivJTerm2}
\boxed{
a_i(\mb{x},t)\nabla\vcdot\mb{g}(\mb{x}) = \frac{a_i(\mb{p}_0,t)}{3d} \bigg( \sum_{j=1}^{6} \big(\mb{g}_j^x + \mb{g}_0^x\big) \cos (60\times(j-1)) + \sum_{j=1}^{6} \big({\mb{g}_j^y + \mb{g}_0^y}\big) \sin (60\times(j-1)) \bigg)
}
\end{equation}

The final expression is suitable for a numerical computation and can
be found in \code{compute\_divJ()} as \code{term2}. Lastly, the final
term in Eq.~\ref{eq:divJExpanded} is the scalar product of two vector
fields which I'll carry out simply with Cartesian $x$ and $y$
components - so I'll re-use my spacegrad2D function to find $\nabla
a_i$. That can be found as \code{term3}.

I think that the advantage of separating this computation out is that
I can ensure that, at the boundary, $\mb{J}$ resulting from the first
term remains 0, by the ghost cell method, then I can tailor
$\mb{g}(\mb{x})$ so that it, and its normal derivative go to 0 at the
boundary, ensuring that the second and third terms also contribute
nothing to $\mb{J}$. To do this, I'll apply a fairly sharp logistic
function based on distance to the boundary to $\mb{g}(\mb{x})$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Computing $a$ and $c$}

See code in \code{rd\_james.h} to inspect the Runge-Kutta numerical
computation. Note that I include some tests to ensure that $a_i$ does
not become negative, $c_i$ does not exceed 1 and $n$ does not become
negative (i.e. that $\textstyle \sum c_i$ does not exceed 1).

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section {Signalling}

Gao et al. show ephrin-A5 inhibits the growth of axons in neurons of
the medial thalamus, while allowing neurons from the lateral thalamus
to branch. ``The EphA5 receptor and its ligand, ephrin-A5 are
expressed in complementary patterns in the thalamus and their cortical
targets. \emph{Additionally}, ephrin-A5 specifically inhibits neurite
outgrowth of medial thalamic neurons from limbic nuclei and sustains
neurite outgrowth of lateral thalamic neurons from primary sensory and
motor nuclei. ... Further analysis using hippocampal neurons indicated
that ephrin-A5 inhibits primarily the growth of axons detected with
antibody against the axon-specific marker $\tau$(60--70\% reduction in
axonal length), although a mild inhibition of dentritic growth also
was observed ($\approx$20\% reduction).

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Two dimensional system with N TCs and M gradients}

We're modifying the system above so that we have N TC populations and
M curated gradients.
%
\begin{equation} \label{eq:dc_NM}
\frac{\partial c_i(\mb{x},t)}{\partial t} = -\alpha_i c_i(\mb{x},t) + \beta_i n(\mb{x},t)
[a_i(\mb{x},t)]^k
\end{equation}
%
\begin{equation} \label{eq:da_NM}
\frac{\partial a_i(\mb{x},t)}{\partial t}
= \nabla\vcdot\mb{J}_i(\mb{x},t) - \frac{\partial c_i(\mb{x},t)}{\partial t}
\end{equation}
%
\begin{equation} \label{eq:conserve_NM}
n(\mb{x},t) + \sum_{i=1}^{N} c_i(\mb{x}, t) = 1
\end{equation}
%
with the flux current
%
\begin{equation} \label{eq:J_NM}
\mb{J}_i(\mb{x},t) = D \nabla a_i(\mb{x},t) - a_i
\sum_{j=1}^M \big(\gamma_{i,j} \nabla\rho_j(\mb{x}) \big)
\end{equation}

The boundary condition that I'd like to apply to these equations is:
%
\begin{equation}
\mb{J}_i(\mb{x},t) \bigg\rvert_{boundary} = 0
\end{equation}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Analyse competition in the 2D Karbowski system}

First sight of the 2D Karbowski system, as implemented in
\code{rd\_james.h/james1.cpp} with N=2 and M=0 is that competition is very
weak indeed. Over the simulation timescale during which the guidance
gradients act (about a thousand steps) pretty much nothing happens to
the competition between the two TC connection types. If one waits a
very long time, then eventually, there is a split between the two,
with one winning over the other, but we're talking timescales on the
order of 30000 and this effect is entirely dwarfed by the effect of
the guidance molecules in the system.

To see if competition is weak across the parameter space, I did a grid
search, varying the parameter $\alpha$ [2:0.15:5], $\beta$ [2:0.15:5]
and $k$ [2:0.1:4]. I left $D$ fixed at 0.1 and as there were no
guidance molecules, the $\gamma$s were all 0.

\begin{figure}[htb!]
\centering
\includegraphics[width=0.8\textwidth]{./plots/ps_2N0M.png}
\caption[Parameter space search.]
{A search of a region of parameter space around $k$=3, $\alpha$=3,
$\beta$=3. The colour of each point represents a normalised scaling of
the sum of squared differences between elements of $c_0$ and $c_1$
after 5000 iterations of the simulation. The maximum value of this
metric was 0.0066; the minimum was 1.7 $\times$ 10$^{-5}$. For
comparison, the value of the metric for a guided system which
separated into two clear regions at 5000 steps was 722.}
\label{fig:ps_2N0M}
\end{figure}

I created two methods to run through the parameter space. One, called
\code{analysis/scanspace.py}, creates a json file for each parameter set, then
calls the program \code{james1c}, and finally
calls \code{sos\_centroid\_analyse()} to add each result
to \code{logs/scanspace.h5}. The problem with this method is that it
repeatedly creates the \code{morph::HexGrid}, adding lots of worthless
computation. To fix this, I wrote \code{ps\_2N0M.cpp}, which uses a
single \code{RD\_James} instance, which it re-sets after each
simulation, without having to re-compute the HexGrid. This can
evaluate 5000 timesteps of one parameter set with the Hex to Hex
spacing set to 0.02 in 1.3 seconds (on
threadbeast). \code{ps\_2N0M.cpp} runs the simulations, but it doesn't
apply the final analysis step of computing the sum of squared
differences between c0 and c1. This is carried out by
\code{analysis/ps\_2N0M\_postproc.py} and plotting by
\code{analysis/ps\_2N0M\_plot.py}. See Fig.~\ref{fig:ps_2N0M}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Two dimensional system with changed competition}

The symmetry of the terms for $\pm \alpha_i c_i$ and
$\pm \beta_i n a_i$ in
Equations \ref{eq:Karb1D_dc} \& \ref{eq:Karb1D_da} results directly
from the simple kinetic reaction:

\begin{equation} \label{eq:kinetic}
n + a_i \xrightleftharpoons[\alpha_{i}]{\beta_{i}} c_i
\end{equation}

If I want to change the competition between axons from different
thalamocortical regions, I need to consder exactly what it is about
Eq.~\ref{eq:kinetic} that I want to modify, so that the model is
changed with some sort of guiding principle.

Here are some modifications to the equations to enable competition
between regions dreamed up in just such a flagrant, un-principled way:

In the following, I'll use these definitions:

Let $\hat{a}_i$ be the sum of the branching densities of all axon
types except $i$.
%
\begin{equation}
\hat{a}_i(\mb{x},t) = \sum_{p\ne i}^N a_p(\mb{x},t)
\end{equation}
%
Define the static vector field $\mb{g}$ as:
%
\begin{equation} \label{eq:g_i}
\mb{g}_i(\mb{x}) = \sum_{j=1}^M \big(\gamma_{i,j} \nabla\rho_j(\mb{x}) \big)
\end{equation}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Allow the branching rate of all other TC axon types to change $\dot{a}_i$}
\label{sec:comp1}

Add a term to the Karbowski equations which reduces the branching rate
change based on the proximity of branching of other TC axon types:
%
\begin{equation} \label{eq:dc_compo_1}
\frac{\partial c_i(\mb{x},t)}{\partial t} = -\alpha_i c_i(\mb{x},t)
+ \beta_i n(\mb{x},t)
[a_i(\mb{x},t)]^k
\end{equation}
%
\begin{equation} \label{eq:da_compo_1}
\frac{\partial a_i(\mb{x},t)}{\partial t}
= \nabla\vcdot\mb{J}_i(\mb{x},t) - \frac{\partial
c_i(\mb{x},t)}{\partial t} - \frac{\epsilon_i  a_i}{N-1} \sum_{p \ne i}^{N} a_p(\mb{x}, t)^l
\end{equation}
%
\begin{equation}
n(\mb{x},t) = 1 - \sum_{i=1}^{N} c_i(\mb{x}, t)
\end{equation}
%
with the flux current as given by Eq.~\ref{eq:J} and Neumann
boundary conditions as usual. This adds two parameter sets; $\epsilon$
and $l$. The factor $\frac{1}{N-1}$ ensures that the value of
$\epsilon$ scales with $N$.

See the binary \code{james1} for the implementation.

The additional term could result from some sort of axon-axon repulsion
as the axons grow into the cortex, or an effect that occurs only once
they are growing within the cortical layer.

An implementation of this mechanism in \code{rd\_james\_comp1.cpp}
does introduce competition quite nicely. Especially evident if $\beta$
is large.

The problem with using this mechanism on its own, is that the
diffusion can easily swamp the $\epsilon_i$ attenuation term, and if
one level of branching for one TC type runs away, then the attenuation
of all the others runs away.

One solution is to pass $a_i$ through a sigmoid, suggesting that there
might be a maximal branching density allowable.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Add an interaction with $\nabla \hat{a}_i$}
\label{sec:comp2}

There could be an additional term to interact with the other branching
densities, to prevent the \emph{branching} overlapping (rather than
just relying on the competition between connections implied by the
definition of $n(\mb{x},t)$ in Eq.~\ref{eq:conserve_NM}).

Change the flux current to:
%
\begin{equation} \label{eq:J_NM_with_comp}
\mb{J}_i(\mb{x},t) = D \nabla a_i(\mb{x},t)
+ \frac{\bar{a}_i(\mb{x}, t) F}{N-1} \nabla \hat{a}_i(\mb{x}, t) -
a_i(\mb{x}, t) \, \mb{g}(\mb{x})
\end{equation}
%
where $\bar{a}_i$ is a sigmoid of height $h$, offset $o$ and
sharpness, $s$:
\begin{equation}
\bar{a}_i = \frac{h}{1 + e^{(o - s \, a_i)}}
\end{equation}
%
The use of the sigmoid prevents the flux current from blowing up due
to the $\nabla \hat{a}_i$ term. In this term (only), we model the idea
that there must be an upper bound on the axonal branching density.

which should be another way to add competition to the system. The idea
here is that branching density should flow away from regions where
there is a high level of branching of other TC axon types.

Note that the $\nabla \hat{a}_i$ term is positive; if branching is
high for the TC type $i$ \emph{and} for all other TC types, then
diffusion of branching away from that location should be increased.

Dropping references to space and time arguments:

\begin{equation} \label{eq:J_NM_with_comp_noargs}
\mb{J}_i = D \nabla a_i
+ \frac{\bar{a}_i F}{N-1} \nabla \hat{a}_i - a_i \, \mb{g}
\end{equation}

The divergence of $\mb{J}_i$ is
%
\begin{equation}
\begin{split}
\nabla\vcdot\mb{J}_i & = \nabla\vcdot \big[ D \nabla a_i + \frac{\bar{a}_i
F}{N-1} \nabla \hat{a}_i - a_i \mb{g}_i \big] \\
%
& =
D \dvrg a_i
+ \frac{F}{N-1} \nabla\vcdot \bar{a}_i \nabla \hat{a}_i
- \nabla\vcdot a_i \mb{g}_i
\end{split}
\end{equation}
%
because the divergence operator is distributive (and $D$ \& $F$ are
scalar constants).  Applying the vector calculus product rule identity to
$\nabla\vcdot\big(\bar{a}_i\nabla \hat{a}_i\big)$ and
$\nabla\vcdot\big(a_i \mb{g}_i\big)$:
%
\begin{equation}
%
\nabla\vcdot\mb{J}_i = D \dvrg a_i
%
+ \frac{F}{N-1} \big(
\bar{a}_i \dvrg \hat{a}_i + \nabla \hat{a}_i \vcdot \nabla \bar{a}_i
\big)
%
- \big(
a_i\nabla\vcdot\mb{g}_i
+
\mb{g}_i\vcdot\nabla a_i
\big)
%
\end{equation}
%
Expanding the brackets we have:
%
\begin{equation} \label{eq:divJExpanded_comp2}
\boxed{
%
\nabla\vcdot\mb{J}_i = D \dvrg a_i
%
+ \frac{F}{N-1} \bar{a}_i \dvrg \hat{a}_i
+ \frac{F}{N-1} \nabla \hat{a}_i \vcdot \nabla \bar{a}_i
%
- a_i\nabla\vcdot\mb{g}_i
- \mb{g}_i\vcdot\nabla a_i
}
\end{equation}
which gives the five terms to compute in \code{rd\_james\_comp2.h}.

Result: With the sigmoid parameters set to $h=2, o=5, s=0.5$, the
range of stable values for $F$ were $[0--6]$. At the upper end of this
range, the competition effect was very slight.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Add an interaction with $\nabla n$}
\label{sec:comp3}

Imagine that as connections are made to the cortical dendrites, a
molecule is expressed. Axon branching should expand into areas with
low connections, so the `flux of branching' should follow the gradient
of this `connectivity molecule' from low $n$ (region with few
connection sites) to high $n$ (region with many connection sites). I've
expressed this by adding a new term to Eq.~\ref{eq:J_NM}:
%
\begin{equation} \label{eq:J_NM_with_comp3}
\mb{J}_i(\mb{x},t) = D \nabla a_i(\mb{x},t)
\boxed{-a_i(\mb{x}, t) E \nabla n(\mb{x}, t)}
 - a_i(\mb{x},t) \, \mb{g}(\mb{x})
\end{equation}
%
With the new term (boxed), it's probably worth going back to look at
the method for computing $\nabla\vcdot\mb{J}_i$. Let's drop references
to the space and time arguments ($\mb{g}_i(\mb{x})$ becomes
$\mb{g}_i$; $n(\mb{x}, t)$ becomes $n$; and so on). This allows a
tidier version of Eq.~\ref{eq:J_NM_with_comp3} to be written
down:
%
\begin{equation} \label{eq:J_NM_with_comp3_simpler}
\mb{J}_i = D \nabla a_i - a_i E \nabla n - a_i(\mb{x}, t) \mb{g}_i
\end{equation}
%
The divergence of $\mb{J}_i$ is
%%%%
\begin{equation}
\begin{split}
\nabla\vcdot\mb{J}_i & = \nabla\vcdot \big[ D \nabla a_i - a_i
E \nabla n - a_i \mb{g}_i \big] \\
%
& =
D \dvrg a_i
- E \nabla\vcdot a_i \nabla n
- \nabla\vcdot a_i \mb{g}_i
\end{split}
\end{equation}
%
because the divergence operator is distributive (and $D$ \& $E$ are
scalar constants).  Applying the vector calculus product rule identity to
$\nabla\vcdot\big(a_i\nabla n\big)$ and
$\nabla\vcdot\big(a_i \mb{g}_i\big)$:
%
\begin{equation}
%
\nabla\vcdot\mb{J}_i = D \dvrg a_i
%
- E \big(
a_i \dvrg n + \nabla n \vcdot \nabla a_i
\big)
%
- \big(
a_i\nabla\vcdot\mb{g}_i
+
\mb{g}_i\vcdot\nabla a_i
\big)
%
\end{equation}
%
Expanding the brackets we have:
%
\begin{equation} \label{eq:divJExpanded_comp3}
\boxed{
%
\nabla\vcdot\mb{J}_i = D \dvrg a_i
%
- E a_i \dvrg n
- E \nabla n \vcdot \nabla a_i
%
- a_i\nabla\vcdot\mb{g}_i
- \mb{g}_i\vcdot\nabla a_i
}
%
\end{equation}
%
which gives us 5 terms to compute in \code{compute\_divJ()} in place
of the original 3 terms in Eq.~\ref{eq:divJExpanded}.

This approach simply increases how fast branching diffuses into free
space; the $D\nabla a_i$ term already provides diffusion of branching.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Interaction with $\nabla{n}$ and $\hat{a}_i$ inhibits $\dot{a_i}$}
\label{sec:comp4}

Combines sections~\ref{sec:comp1} and \ref{sec:comp3}. Descending the
gradient $a_i\nabla{n}$ expands axon branching into new regions;
competition between $a_i$ and $\hat{a}_i$ should prevent excessive
branching into neighbouring regions.

%
\begin{equation} \label{eq:dc_comp4}
\frac{\partial c_i}{\partial t} = -\alpha_i c_i
+ \beta_i n
[a_i]^k
\end{equation}
%
\begin{equation} \label{eq:da_comp4}
\frac{\partial a_i}{\partial t}
= \nabla\vcdot\mb{J}_i + \alpha_i c_i - \beta_i n
[a_i]^k - \frac{\epsilon_i}{N-1} a_i \sum_{p \ne i}^{N} a_p^l
\end{equation}
%
\begin{equation} \label{eq:conserve_NM_comp4}
n = 1 - \sum_{i=1}^{N} c_i(\mb{x}, t)
\end{equation}
%
\begin{equation} \label{eq:divJExpanded_comp4}
%
\nabla\vcdot\mb{J}_i = D \dvrg a_i
%
- E a_i \dvrg n
- E \nabla n \vcdot \nabla a_i
%
- a_i\nabla\vcdot\mb{g}_i
- \mb{g}_i\vcdot\nabla a_i
%
\end{equation}

Note that
Eqs.~\ref{eq:dc_comp4}--\ref{eq:conserve_NM_comp4} are
not quite identical those in section~\ref{sec:comp1}. The difference
is the $\frac{1}{N-1}$ factor and the inclusion of $a_i$ in the third term of
Eq.~\ref{eq:da_comp4}.

This is very promising! Initial branching conditions produce reliable
regions. Initial conditions can be modified to simulate denucleated
animal; region becomes smaller, but not non-existent.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{$\hat{a}_i$ inhibits $\dot{a_i}$ \bf{more}}
\label{sec:comp5}

In the config where I'm trying to make 5 barrel rows, the competition
is less competitive than I imagine it could be. The initial condition
is a blob of each of the 5 TC types of roughly the same size and I
think the problem might be that the $\epsilon$ term is inhibiting all
5 TC types more or less equally.

This idea (to avoid attenuating the biggest $a$) didn't work but lead
me to think about bounding $a$ and the problem of a winner-takes-all
situation that is easily seen.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Add interactions with $\nabla {a}_{{p}\ne{i}}$}
\label{sec:comp6}

A scheme similar to comp2 (section \ref{sec:comp2}). However, rather
than using the gradient $\nabla \hat{a}_i$, use all the gradients of
the other TC types, with a $\zeta$ matrix to set the strengths of
interaction.

Change the flux current to:
%
\begin{equation} \label{eq:J_NM_with_comp6}
\mb{J}_i(\mb{x},t) = D \nabla a_i(\mb{x},t)
+ \bar{a}_i(\mb{x}, t)  \sum_{p \ne i}  \zeta_{i,p} \nabla {a}_p(\mb{x}, t) -
a_i(\mb{x}, t) \, \mb{g}(\mb{x})
\end{equation}
%
where $\bar{a}_i$ is a sigmoid of height $h$, offset $o$ and
sharpness, $s$:
\begin{equation}
\bar{a}_i = \frac{h}{1 + e^{(o - s \, a_i)}}
\end{equation}
%
used for the same reason as in section \ref{sec:comp2}.

Dropping references to space and time arguments:

\begin{equation} \label{eq:J_NM_with_comp6_noargs}
\mb{J}_i = D \nabla a_i + \bar{a}_i \sum_{p \ne i} \zeta_{i,p} \nabla {a}_p - a_i \, \mb{g}
\end{equation}

The divergence of $\mb{J}_i$ is
%
\begin{equation}
\begin{split}
\nabla\vcdot\mb{J}_i & = \nabla\vcdot \big[ D \nabla a_i
+ \sum_{p \ne i} \zeta_{i,p} \bar{a}_i  \nabla {a}_p - a_i \, \mb{g} \big] \\
%
& =
D \dvrg a_i
+ \sum_{p \ne i} \zeta_{i,p} \nabla\vcdot \bar{a}_i \nabla {a}_p - \nabla\vcdot a_i \mb{g}_i
\end{split}
\end{equation}
%
because the divergence operator is distributive (and $D$ \& $\zeta$ are
scalar).  Applying the vector calculus product rule identity to
$\nabla\vcdot\big(\bar{a}_i\nabla {a}_p\big)$ and
$\nabla\vcdot\big(a_i \mb{g}_i\big)$:
%
\begin{equation}
\boxed{
%
\nabla\vcdot\mb{J}_i = D \dvrg a_i
%
+ \sum_{p \ne i} \zeta_{i,p} \big(
\bar{a}_i \dvrg {a}_p + \nabla {a}_p \vcdot \nabla \bar{a}_i
\big)
%
- \big(
a_i\nabla\vcdot\mb{g}_i
+
\mb{g}_i\vcdot\nabla a_i
\big)
%
} % boxed
\end{equation}
%
which on expansion gives $3 + 2(N-1)$ terms terms to compute
in \code{rd\_james\_comp6.h} (not yet implemented).

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{$\hat{a}_i$ multiplicatively reduces diffusion of $a_i$}
\label{sec:comp7}

Change the flux current to:
%
\begin{equation} \label{eq:J_NM_with_comp7}
\mb{J}_i = D (1 - \sigma(\hat{a}_i)) \nabla a_i - a_i \, \mb{g}
\end{equation}

where $\sigma(\hat{a}_i)$ denotes a sigmoid function:

\begin{equation}
\sigma(a) = \frac{1}{1 + e^{(o - s \, a)}}
\end{equation}

Let $\lambda = 1 - \sigma$. Then $\nabla \vcdot \mb{J}_i$ is

\begin{equation}
\nabla\vcdot \mb{J}_i = \nabla\vcdot \big( D \lambda \nabla a_i - a_i \, \mb{g} \big)
\end{equation}

Product rules all round to give:

\begin{equation}
\boxed{
\nabla\vcdot \mb{J}_i = D\big( \lambda \dvrg a_i + \nabla
a_i \vcdot \nabla \lambda \big) - \big( a_i \nabla.\mb{g} +
\mb{g} \vcdot \nabla a_i \big)
}
\end{equation}

Ah, I added in the direct competition mechanism of
section \ref{sec:comp1} (though as refined in section \ref{sec:comp4})
too (see \code{rd\_james\_comp7.h}).

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Divisive normalization of $a_i$}
\label{sec:comp8}

Taking inspiration from the divisive normalization idea in neural
networks~\cite{carandini_normalization_2011}, where the overall
activity of a group of neurons modulates the output response of an
individual member of the group, we will use the overall branching
level of all of the axons in a layer to modulate the response, $a_i'$,
of the branching level in each element.

The response can be written out with the normalization
equation~\cite{carandini_normalization_2011}:
%
\begin{equation}\label{eq:norm}
a_i' = h \; \frac {a_i^q} {\xi^q + \textstyle \sum_j^h a_{i,j}^q }
\end{equation}
%
in which $\xi$ and $q$ are free parameters and $n_e$ is the number of
hex elements in the domain of computation (although $n_e$ could also be
chosen as a free parameter if required). Here, I'll let $\xi=0$ and $q=1$.

$a_i'$, the `branching response', then has to be used to drive
connections, in practice, it is written back into $a_i$ on every
timestep.

That gives the following set of equations to express in the code:

%
\begin{equation} \label{eq:dc_comp8}
\frac{\partial c_i}{\partial t} = -\alpha_i c_i
+ \beta_i n
[a_i]^k
\end{equation}
%
\begin{equation} \label{eq:da_comp8}
\frac{\partial a_i}{\partial t}
= \nabla\vcdot\mb{J}_i + \alpha_i c_i - \beta_i n [a_i]^k
\end{equation}
%
\begin{equation}\label{eq:norm_comp8}
a_i' = n_e \; \frac {a_i} {\textstyle \sum_j^{n_e} a_{i,j} }
\end{equation}
%
($a_i'$ becomes $a_i$ in the next computational step.)
%
\begin{equation} \label{eq:conserve_NM_comp8}
n = 1 - \sum_{i=1}^{N} c_i
\end{equation}
%
\begin{equation} \label{eq:divJExpanded_comp8}
\nabla\vcdot\mb{J}_i(\mb{x}) =
D \dvrg a_i - a_i\nabla\vcdot\mb{g} - \mb{g}\vcdot\nabla a_i
\end{equation}

One issue with divisive normalization is that each TC type is
normalized to have the same overall branching, and so changed initial
conditions, such as reduced initial innervation for one field over
another will not result is very different simulations.

An idea to get around this is to introduce an additional multiplier to
the normalization, which is the mean initial branching density per
element, $d_i/n_e$, so that Eq.~\ref{eq:norm_comp8} becomes

\begin{equation}\label{eq:norm_comp8plus}
a_i' = d_i \; \frac {a_i} {\textstyle \sum_j^{n_e} a_{i,j} }
\end{equation}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Divisive normalization of $a_i$ and $\hat{a}_i$ inhibits $\dot{a}_i$}
\label{sec:comp9}

This is a combination of code in section \ref{sec:comp4} (the
competition part) and \ref{sec:comp8}.

The separate regions compete evenly, as they are all normalized to
have the same branching density.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Piecewise linear transfer function normalizes $a_i$}
\label{sec:comp10}

%
\begin{equation} \label{eq:dc_comp10}
\frac{\partial c_i}{\partial t} = -\alpha_i c_i
+ \beta_i n
[a_i]^k
\end{equation}
%
\begin{equation} \label{eq:da_comp10}
\frac{\partial a_i}{\partial t}
= \nabla\vcdot\mb{J}_i + \alpha_i c_i - \beta_i n [a_i]^k
\end{equation}
%
\begin{equation}\label{eq:norm_comp1}
a_i' = \begin{cases}
     0 & a_i < 0 \\
     a_i & a_i\leq a_{max} \\
     a_{max} & a_i > a_{max}
     \end{cases}
\end{equation}
%
($a_i'$ becomes $a_i$ in the next computational step.)
%
\begin{equation} \label{eq:conserve_NM_comp10}
n = 1 - \sum_{i=1}^{N} c_i
\end{equation}
%
\begin{equation} \label{eq:divJExpanded_comp10}
\nabla\vcdot\mb{J}_i(\mb{x}) =
D \dvrg a_i - a_i\nabla\vcdot\mb{g} - \mb{g}\vcdot\nabla a_i
\end{equation}

This is implemented in \code{rd\_james\_comp10.h}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Piecewise linear transfer function normalizes $a_i$ plus competition}
\label{sec:comp11}

This is implemented in \code{rd\_james\_comp11.h}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Divisive normalization of $a_i$ plus competition with
extended $\epsilon_{i,j}$}
\label{sec:comp12}

Allow each TC type to interact with the other $N-1$ TC types with
varying magnitudes. This is an extension to the competition type in
section \ref{sec:comp9}. The $N$ sized vector $\epsilon$ becomes an
$N \times N$ matrix $\epsilon$ and Eq. \ref{eq:da_comp4} becomes
%
\begin{equation} \label{eq:da_comp12}
\frac{\partial a_i}{\partial t}
= \nabla\vcdot\mb{J}_i + \alpha_i c_i - \beta_i n
[a_i]^k - \frac{1}{N-1} a_i \sum_{p \ne i}^{N} \epsilon_{i,p} \; a_p^l
\end{equation}
%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Divisive normalization, additional competition mechanisms}
\label{sec:comp13}

Because the divisive normalization is so helpful in keeping the $N$
branching variables balanced with respect to each other, I want to
reintroduce the additional diffusion ideas developed earlier.  First,
introduce section \ref{sec:comp2} (flux current affected by gradient
of other branching densities).

Note that the mechanism from section \ref{sec:comp2} shouldn't need
the sigmoid, as $a_i$ is already divisively normalized in this
scheme. The computation of $\mb{J}_i$ is thus:
%
\begin{equation} \label{eq:J_comp13}
\mb{J}_i = D \nabla a_i
+ \frac{a_i F}{N-1} \nabla \hat{a}_i - a_i \, \mb{g}
\end{equation}
%
The divergence of $\mb{J}_i$ can be computed in the same was as in
section \ref{sec:comp2}:
%
\begin{equation} \label{eq:divJExpanded_comp13}
%
\nabla\vcdot\mb{J}_i = D \dvrg a_i
%
+ \frac{F}{N-1} {a}_i \dvrg \hat{a}_i
+ \frac{F}{N-1} \nabla \hat{a}_i \vcdot \nabla {a}_i
%
- a_i\nabla\vcdot\mb{g}_i
- \mb{g}_i\vcdot\nabla a_i
\end{equation}

This is implemented in \code{rd\_james\_comp13}. It seems to be very
unstable, as it was before.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Divisive normalization, additional competition mechanisms 2}
\label{sec:comp14}

I've added\ref{sec:comp3} (diffusion is increased into
regions with low $n$). This mechanism is successful in keeping
branching outside regions where connections have been made.

I'm also going to try to add \ref{sec:comp7} (multiplicative reduction
of diffusion of $a_i$), which I should be able to turn on and off with
a parameter.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Unimplemented ideas}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Upper bound on $a_i$ (sigmoid) and $\hat{a}_i$ inhibits $\dot{a}_i$}

The alternative is to let $n_c \equiv n$ and introduce a new $n_a$
such that
%
\begin{equation}
n_c = 1 - \sum_i c_i
\end{equation}
%
\begin{equation}
n_{a,i} = 1 - \sum_i a_i
\end{equation}
%
and then introduce $n_{a,i}$ into the RD equations. The quantity
$n_{a,i}$ is then the maximum density of branching permissible for TC
type $i$, due to physical and/or chemical considerations.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Initial conditions}

Initial reading of the literature suggests that axons grow towards
cortex in a `guideposted manner' and then start branching once within
the cortex. This would suggest that initial conditions should be more
targeted than in the Karbowski model, with axons of type $i$ arriving
in a region expected to be region $i$, but with the possibility of
modifying the model to simulate an enucleated animal by reducing the
density of incident visual TC axons, or reducing their branching rate.

I've implemented a scheme, using 2D Gaussians of varying (symmetrical)
width, gain and location, to make masks for the initial branching
densities for the different thalamocortical axon types.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Axon guidance reading}

Erzurumlu \& Gaspar~\cite{erzurumlu_development_2012} have a great
table of chronological events in the development of the whisker
barrels.

Gross topographic order is present in the projections from the
trigeminal ganglion to the whisker pad and brainstem even from E12.

Thalamocortical axons are on their way to the neocortex by E13,
passing a ``corridor'' by E13.5.

Layer IV neurons in cortex are born around E14-E15.

Thalamocortical axon pioneers arrive at prospective somatosensory
cortex around E15.

At around birth (E19 to P0), barrelette patterns emerge in principle
sensory nucleus of the trigeminal nerve.

At P1, axons of the principle sensory nucleus of the trigeminal nerve
form bands in the ventro-postero-lateral nucleus the thalamic complex
(VPM).

By P3, barreloids are visible in VPM.

Sequential ordering of pattern formation first in brainstem, then in
thalamus and finally in cortex is reviewed in Sehara \& Kawasaki, 2011.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Agent based model}

It may be helpful to create a more realistic axon model, where axons
grow in a tree manner, with guidance and interactions, from physically
plausible locations in a virtual thalamus.

\section{Applying the model to whisker barrels}

Exploring how the model can speak to Whisker barrels.

Axons on their way from ventrobasal thalamus (VB) to somatosensory
cortex arrive in Layer VIb and form a `mass' of axons, with axons
branching in every direction.

I'm going to first try to model the formation of rows of axons within
this mass. It's plausible that there is an ordering within this
region, because from this region can come both the wildtype barrel
cortex \emph{or} the reversed duplicate barrel field.

\subsection{Barrel rows: $N$ 1D Gaussian waves}

This mechanism makes nice
rows. See \code{configs/whiskers/5N5M\_rows.json}. The problem is that
it requires bands of expressed guidance molecules and banded
expression isn't seen in barrel cortex by the experimentalists.

\subsection{Barrel rows: 1 guidance molecule plus competition}

With the competition described in section~\ref{sec:comp4}, it is
difficult for separate, well delineated regions to form. I'm trying to
think of an idea to make the competition between regions stronger.

One possibility is to have each TC type inhibit only selected other TC
types. This might separate the centroids of effect, leading to better
row definition.

Another idea is to introduce a chemical generated in proportion to
axon branching, which then has some finite decay time.

Or, whichever a is the highest sees no attenuation. (Implemented
as \code{rd\_james\_comp5.h}).

I think the issue is that whatever inter-TC-type attenuation there is
is swamped by the diffusion term. That branching just keeps on
spreading out...

I'll re-visit the concept of using the gradient of branching in types $j \ne i$
to influence the branching flux of type $i$ (That's in section~\ref{sec:comp2}).

\subsection{Barrel rows; 1 guidance molecule pushes right, $N$ others
push left, giving rows.}

The mechanism works quite nicely, and has the advantage of using only
graded guidance molecule expression.
See \code{configs/whiskers/5N\_boundedorder.json}
%
% BIBLIOGRAPHY
%
\selectlanguage{English}
\bibliographystyle{abbrvnotitle}
% The bibliography NoTremor.bib is the one exported from Zotero. It
% may be necessary to run my UTF-8 cleanup script, bbl_utf8_to_latex.sh
%%\bibliography{NoTremor}
%
% It produces this:
\begin{thebibliography}{1}

\bibitem{george_b._arfken_mathematical_1995}
{George B.~Arfken} and H.~J.~Weber.
\newblock {\em Mathematical {Methods} for {Physicists}}.
\newblock Academic Press, fourth edition edition, 1995.

\bibitem{karbowski_model_2004}
J.~Karbowski and G.~Ermentrout.
\newblock {\em Journal of Computational Neuroscience}, 17(3):347--363, Nov.
  2004.

\bibitem{lee_hexagonal_2014}
D.~Lee, H.~Tien, C.~Luo, et~al.
\newblock {\em International Journal of Computer Mathematics},
  91(9):1986--2009, Sept. 2014.

\bibitem{erzurumlu_development_2012}
R.~Erzurumlu and P.~Gaspar.
\newblock {\em European Journal of Neuroscience},
  35:1540--1553, Feb. 2012.

\bibitem{carandini_normalization_2011}
M.~Carandini and D.~J.~Heeger
\newblock {\em Nature Reviews Neuroscience},
  13:51--62, Nov. 2011.

\end{thebibliography}

%%% Upload the *bib file along with the *tex file and PDF on
%%% submission if the bibliography is not in the main *tex file

\end{document}

### 2019-11-28 paper/supplementary/supp.sh

@@ -1,0 +1,6 @@
+#!/bin/bash
+
+xelatex supp.tex
+bibtex supp.aux
+xelatex supp.tex
+xelatex supp.tex

### 2019-11-28 paper/medpaper/medium.sh
Adds a copy of the paper in its 'medium length' format prior to taking a hatchet to the briefpaper

@@ -1,0 +1,6 @@
+#!/bin/bash
+
+xelatex medium.tex
+bibtex medium.aux
+xelatex medium.tex
+xelatex medium.tex

### 2019-12-02 paper/briefpaper/brief.tex
WIP reducing text down

@@
 \begin{abstract}
-Understanding how genetic information and spontaneous pattern formation
-together give rise to the whisker barrels is an important challenge for
-developmental neurobiology. The barrel patterning that coalesces in layer IV
-has been described formally as a Voronoi tesselation, which is indicative of
-an underlying developmental process based on local competitive
-interactions. If pattern formation relies only on such intrinsic
-self-organising processes, how can the somatotopic order at the periphery be
-maintained in the projections from brainstem to thalamus and from thalamus to
-cortex?
+The whisker barrel patterning that coalesces in cortical layer IV has been
+described formally as a Voronoi tesselation, which is indicative of an
+underlying developmental process based on local competitive interactions. If
+pattern formation relies only on such intrinsic self-organising processes, how
+can the somatotopic order at the periphery be maintained in the projections
+from brainstem to thalamus and from thalamus to cortex?
 %
@@
 %
+
 %However, a
@@
 
-\dropcap{S}patial patterns in neural connectivity provide clues about the
-constraints under which brains evolve and develop.
-%Whether or not patterning
-%governs information processing, models of the development of information
-%processing circuits can be validated in terms of their ability to account for
-%biological patterning \citep{purves_iterated_1992}.
-%%see also \citealp{wilson_what_2015,bednar_cortical_2016}
-One of the most
-distinctive patterns of neural connectivity is found in the rodent
-barrel cortex \cite{woolsey_structural_1970}.
-%(\citealp{woolsey_pattern_1948,woolsey_structural_1970,welker_structure_1974},
-%see \citealp{fox_barrel_2008}).
-Understanding how this particular pattern forms might shed light on the
-interplay of genetic and extrinsic factors that shape cortical circuits.
-%, and on a wealth of
-%anatomical, physiological, and behavioural data obtained using this
-%model system.
-
-The barrels are large cortical columns, identifiable from birth in layer 4 of
-primary somatosensory cortex as dense clusters of thalamocortical axons, which
-are enclosed by borders a few neurons thick from postnatal day 3
+\dropcap{O}ne of the most distinctive patterns of neural connectivity is found
+in the rodent \emph{barrel cortex} \cite{woolsey_structural_1970}, which is
+comprised of large cortical columns which develop from birth
 \citep{erzurumlu_development_2012}. In the plane tangential to the cortical
 surface the barrels constitute a map of the arrangement of the whiskers on the
-face,
-%(see \citealp{yamakado_subdivision_1979})
-such that the cells of adjacent barrels respond most strongly and most quickly
-to deflection of adjacent whiskers \citep{armstrong-james_flow_1992}. The
-barrels reflect subcortical whisker maps comprising cellular aggregates known
-as barreloids in the thalamus \cite{van_der_loos_barreloids_1976} and
-barrelettes in the brainstem.
-%\citep{killackey_pattern_1980}.
+face, such that the cells of one barrel respond most strongly to the
+deflection of its corresponding whisker \citep{armstrong-james_flow_1992}. The
+barrel field reflects similar, subcortical whisker maps comprising cellular
+aggregates known as \emph{barreloids} in the thalamus
+\cite{van_der_loos_barreloids_1976} and \emph{barrelettes} in the brainstem.
+%
+In this paper, we consider the mechanism by which the map can be reliably
+re-created in these separate tissues. Any candidate mechanism must maintain
+the somatotopic ordering of the facial whiskers; it must operate in tissues
+comprised of differing neuronal cell species and which develop at several
+different length scales. We focus on the thalamic barreloid map and how
+thalamocortical axons become ordered in their terminations in the barrel
+field, but point out that the mechanism should be transferable to the creation
+of the barreloids from the barrelettes and the formation of the barrelette
+fields from the mystacial whiskers themselves.
 
-%Each whisker and its corresponding barrelette, barreloid, and barrel can be
-%identified by its position on a two-dimensional grid.
-%\citep{zucker_coding_1969,killackey_pattern_1980,van_der_loos_barreloids_1976,haidarliu_size_2001}.
-%
-%The barrels are normally up to twice as long in the axis aligned to the grid
-%‘rows’ compared to the ‘arcs’, and the rows are visible from day 2
-%\citep{rebsam_refinement_2002} whereas the arcs become distinct by day 3
-%\citep{erzurumlu_development_2012,rebsam_refinement_2002}.
-%In metabotropic glutamate receptor 5 (mGlu5) knockouts the barrel rows develop
-%but the arcs do not (\citealp{hannan_plc-1_2001} see also
-%\citealp{fox_barrel_2008}), thus distinct processes likely guide development
-%of the barrels with respect to the orthogonal grid axes.
-
-Barrel formation requires afferent input, driven by whisker stimulation and by
-thalamic calcium waves that propagate via mGlu5-receptor activation
-\citep{anton-bolanos_prenatal_2019}.
-%(\citealp{anton-bolanos_prenatal_2019}, see also
-%\citealp{anton-bolanos_developmental_2018}).
-Patterning also depends on a complex network of axon guidance processes, which
-includes interactions between guidance molecules like ephrin-A5 and A7
-\citep{miller_epha7-ephrin-a5_2006} and between adhesion molecules like
-cadherin-6 and 8 \citep{bishop_regulation_2000}, and the
-binding of p75 to neurotropins
-\citep{bishop_distinct_2002}.
-% or both citations?
-%\citep{shimogori_fibroblast_2005,bishop_distinct_2002}.
-%
-%(\citealp{shimogori_fibroblast_2005,bishop_distinct_2002}
-%; see also \citealp{dye_lifespan_2011,dye_lifespan_2011-1}).
-This network is orchestrated by interactions between morphogens Fgf8 and Fgf17
-and transcription factors Emx2, Pax6, Sp8, and Coup-tf1
-\citep{shimogori_fibroblast_2005,bishop_regulation_2000}, which are expressed
-in gradients that mark orthogonal axes and can be manipulated to cause the
-barrels to stretch, shrink, shift, skew, and even duplicate \cite{assimacopoulos_fibroblast_2012}.
-% A longer list of citations:
-%\citep{shimogori_fibroblast_2005,assimacopoulos_fibroblast_2012,borello_sp8_2014,sahara_sp8_2007,ypsilanti_transcriptional_2016,sur_patterning_2005}.
+The formation of the cortical barrels is related to the guidance of axons
+which grow from the thalamus, via the internal capsule and into the developing
+cortical sub-plate \citep{erzurumlu_development_2012}. Here, they form an
+undifferentiated bundle \citep{agmon_organized_1993}, out of which they ascend
+in an ordered pattern to arrive in cortical layer IV, where they form
+synapses. Axon guidance depends on a network of actors including guidance
+molecules such as ephrin-A5 and A7 and their receptors the Ephs
+\citep{miller_epha7-ephrin-a5_2006} and adhesion molecules like cadherin-6 and
+8 \citep{bishop_regulation_2000}. This network is orchestrated by interactions
+between morphogens Fgf8 and Fgf17 and transcription factors Emx2, Pax6, Sp8,
+and Coup-tf1 \citep{shimogori_fibroblast_2005,bishop_regulation_2000}, which
+are expressed in gradients that mark orthogonal axes and can be manipulated to
+cause the barrels to stretch, shrink, shift, skew, and even duplicate
+\cite{assimacopoulos_fibroblast_2012}.
 %
 While complex structure in protein expression can be observed for some genes,
-such as the molecules Bhlhb5 \citep{joshi_bhlhb5_2008}, we are not aware of
-any report of gene expression patterns with the complexity of the barrel field
-and so the mechanism by which the barrels form seems likely to be one in which
-the structure forms by self-organization.
+such as the molecule Bhlhb5 \citep{joshi_bhlhb5_2008}, all of the factors
+mentioned above are expressed in monotonic, slowly varying patterns. This
+suggests that the barrel field is not created according to a genetically
+specified `stencil' but is instead formed by a self-organising process.
+%
+% Can I get this to be less of an orphan sentence?
+Barrel formation also requires afferent input, driven by whisker stimulation
+and by thalamic calcium waves that propagate via mGlu5-receptor activation
+\citep{anton-bolanos_prenatal_2019}.
 
 The boundaries of the barrels form a Voronoi tesselation across the cortical
-sheet \citep{senft_mouse_1991}. This patterning suggests that the barreloid
-topology may be preserved in the initial projection of the thalamocortical
-axons into the cortex, and that a barrel forms by the lateral branching of the
-axons from their initial centres of mass, which ceases upon contact with the
-axons branching from adjacent centres. According to this model the barrel
-pattern could be entirely determined by the distribution of the initial centre
-points, which would itself be determined by guidance cues. However, the
-assumption of pre-arranged centres is difficult to resolve with the
-observation that thalamocortical axons arrive in the cortical plate as an
-undifferentiated bundle prior to barreloid formation within the thalamus
-\cite{agmon_organized_1993}. % Plus erzurumli, probably
+sheet \citep{senft_mouse_1991} in which each \emph{Voronoi domain} (a barrel)
+contains a geometrically defined centre point. A Voronoi tesselation can be
+formed by a process which is initialized with the locations of all the centre
+points and then builds the Voronoi domains. The advantage of this, in
+understanding barrel formation, is that by pre-specifying the centre-point
+locations, the map is guaranteed to maintain its ordering. However, this
+theory requires that some initial process defines the location of each barrel,
+prior to the arrival of the thalamocortical axons. If this were so, then some
+factor should be detectable which would indicate the structural complexity of
+the barrel field.
 
-Alternatively, reaction-diffusion dynamics could generate a Voronoi-like
-tesselation in the cortex without the need for pre-arranged centres, by
-amplifying characteristic modes in the noisy initial distribution of bundled
-axon branches, as a net effect of short-range cooperative and longer-range
-competitive interactions between cells. According to this model the barrel
-pattern would be determined by the relative strength of these interactions and
-by the shape of the cortical field boundary. However, reaction-diffusion
-dynamics alone are unlikely to be able to account for the precise
-co-registration and topographic connectivity between emergent domains in
-thalamus and cortex, the non-uniform sizes and spacing of the barrels to form
-rows versus arcs, and the influence of gene expression gradients.
+The presence of the Voronoi tesselation could alternatively suggest a process
+which forms the Voronoi domains according to local interactions between axon
+branches, dendrites and the surrounding cells - such as in a reaction
+diffusion system \cmnt{finish me}. For the barrels, this leaves
+unanswered the question of how the somatotopy is preserved, as fluctuations in
+such a process would likely cause a different pattern to form in each
+individual.
 
-As theories of barrel patterning, the centre-point and reaction-diffusion
-models are not mutually exclusive. Pre-organised centres in the initial
-conditions could bias reaction-diffusion processes to generate the specific
-somatotopic arrangement of barrels more reliably, and similarly the mechanisms
-by which axons branch laterally from barrel centres may involve a tension
-between cooperative and competitive interactions. However, a proof that barrel
-patterning can in principle emerge within an undifferentiated bundle of
-thalamocortical axons, based only on local interactions between cells, would
-remove the need to invoke a global supervisor, i.e., to assume a separate
-stage and/or extrinsic mechanism for pre-organising thalamocortical
-connections.
-
-We seek to establish whether it is possible for the observed arrangement of
-barrels to self-organise in a system governed by reaction-diffusion dynamics,
-under the guidance of molecular signalling gradients, and in the absence of
-pre-defined centre-points.
-%
-Our approach is to represent in a computational model the minimal set of
-biological assumptions that might lead to the emergence of barrel patterning
-in simulation, and, if necessary, to add to this minimal set only assumptions
-that can be represented by the local exchange of information between adjacent
-cells. The emergence of barrel patterning under such conditions would serve as
-an existence proof of the plausibility of the hypothesis that barrel
-development is a self-organising process.
+We seek to establish whether it is possible, and if so, the minimal set of
+biological assumptions required, for the observed arrangement of barrels to
+self-organise in a system governed by reaction-diffusion dynamics, under the
+guidance of molecular signalling gradients, and in the absence of pre-defined
+centre-points.
 

### 2019-12-03 paper/briefpaper_stuedits/brief copy.tex
added sreamlined draft

--- 2019-11-27 paper/briefpaper/brief.tex
@@
 
+\begin{figure*}
+\begin{center}
+\includegraphics[width=\textwidth]{./MainFig.png}
+\end{center}
+%\vspace{-0em}
+\caption{\textbf{A.} Barreloid centres taken from \cite{haidarliu_size_2001}
+  are shown as black circles. Barreloids are labelled according to the usual
+  convention. Labels R1 to R5 are the rhinal whisker barreloids. The dashed
+  grey line represents the boundary of the barreloid field; the dot-dash line
+  shows the anterior-posterior axis. It is assumed that there are two linearly
+  increasing, orthogonal signalling gradients (molecules T1 \& T2) in the
+  thalamus (side graphs), such that the location of each barreloid within the
+  graded thalamic signal field can differentiate the efferent axons, with the
+  result that axons from each barreloid interact to varying degrees with
+  cortical signalling gradients (shown in panel B). Coloured arrows show, for
+  each barreloid, the magnitude (and sign) of the interaction that its
+  thalamocortical projection will have with the signalling gradients in the
+  cortex. The mappings from barreloid location to interaction strength are
+  linear. \textbf{B.} Guidance molecules in the simulated cortex. Molecule 1
+  is expressed high to low in an approximately anterior to posterior
+  gradient. Molecule 2 is expressed lateral to medial (analogously to ephrin -
+  Vanderhaegen et al) \textbf{C.}  Whisker barrels from
+  \cite{shimogori_fibroblast_2005}, with annotations for barrel identity and
+  approximate axes.  \textbf{D-F.} Barrel cortex simulation. The maps indicate
+  which thalomocortical axons make maximum connection density. Each hex
+  element is coloured one of 52 unique colours depending on which e
+  thalamocortical axon makes the most connections in that element. The mapping
+  between barrel identity and colour is given in F.  \textbf{G.} The Honda
+  Dirichlet measure as a function of simluation time for the simulations show
+  in D-F. \textbf{H.}  Guidance molecules in the simulated Fgf8 misexpression
+  experiment. Molecule 1 is mirrored over the anterior-posterior axis
+  \textbf{I.} The simulation indicates that the mirroring of the barrel field
+  can occur as a result of the mirroring of guidance molecule 1.}
+\label{fig:main}
+\end{figure*}
+
 %\cmnt{If we need to present the results of the 2-D system with $\chi=0$, then
@@
 
-To investigate how the whisker barrel structure might form in the presence of
-only monotonically varying morphogen gradients, we turned our attention to the
-competition for connections built into the model. There does exist an element
-of competition in the original model (Eqs.~\ref{eq:dc} \& \ref{eq:da}). The
-term $\left(1 - \sum_{i=1}^{N} c_i\right)$ in Eq.~\ref{eq:dc} ensures that
-once the available synaptic receptors in a region have been fully occupied by
-one thalamocortical axon type, there is no longer the possibility for other
-types to form synapses. This competition turns out to be relatively weak,
-because it does not strongly affect the branching density in that region; it
-is possible for two or more TC axon types to have roughly equal branching
-density in some region, leading to roughly equal numbers of connections
-forming for each TC type, up until the `hard' connection limit is reached.
+To investigate how the whisker barrel structure (Fig.~\ref{fig:main}.\,C) might
+form in the presence of only monotonically varying morphogen gradients, we
+turned our attention to the competition for connections built into the
+model. There \emph{does} exist an element of competition in the original model
+(Eqs.~\ref{eq:dc} \& \ref{eq:da}). The term $\left(1 - \sum_{i=1}^{N}
+c_i\right)$ in Eq.~\ref{eq:dc} ensures that once the available synaptic
+receptors in a region have been fully occupied by one thalamocortical axon
+type, there is no longer the possibility for other types to form
+synapses. This competition turns out to be relatively weak, because it does
+not strongly affect the branching density in that region; it is possible for
+two or more TC axon types to have roughly equal branching density in some
+region, leading to roughly equal numbers of connections forming for each TC
+type, up until the `hard' connection limit is reached.
 
@@
 \begin{equation} \label{eq:comp}
-\chi(\mb{x}, t) = - \frac{\epsilon  a_i}{N-1} \sum_{i' \ne i}^{N} f(a_{i'}).
+\chi(\mb{x}, t) = - \frac{\epsilon  a_i}{N-1} \sum_{i' \ne i}^{N} a_{i'}
 \end{equation}
 %
-with the choice $f(a) = a$, which leads to a reduction in the rate of increase
+%with the choice $f(a) = a$,
+which leads to a reduction in the rate of increase
 of branching density for axon type $i$ in regions where the branching density
@@
 
-With the addition of the competition term, $\chi$, it is necessary to
-introduce a normalisation step, in order to prevent one TC type from
-out-competing every other TC type. Normalisation represents the idea that for
-a fixed number of afferent thalamocortical axons, the overall number of
-branches must have some upper limit; given the finite resources available in
-each axon for building new processes. We divisively normalize the branching
-density, dividing it at each computational step by the mean branching density
-per unit area. It is then multiplied by the branching density per unit area at
-time 0, to account for the possibility that there might be more axons afferent
-for TC type $1$ cf.~TC type $2$:
+With the addition of branching competition, it is necessary to introduce a
+normalisation step, in order to prevent one TC type from out-competing every
+other type. This represents the idea that for a fixed number of afferent
+thalamocortical axons, the overall number of branches must have some upper
+limit, given the finite resources available in each axon for building new
+processes. We divisively normalize the branching density, dividing it at each
+computational step by the mean branching density per unit area. It is then
+multiplied by the branching density per unit area at time 0, to account for
+the possibility that there might be more axons afferent for TC type $1$ cf.~TC
+type $2$:
 %
@@
 
-In which we show that a realistic barrel pattern in the thalamus can be
-recreated in barrel cortex.
+With these changes we were able to generate grid-like patterns with
+appropriate (linearly spaced) choices of the interaction parameters,
+$\gamma$. To determine the $\gamma$ values suitable to recreate the cortical
+barrel field by a principled method, we postulated that the interaction of
+each thalamocortical axon group with the ephrin gradients in the cortical
+sub-plate might be determined by the location of the barreloid within
+signalling gradents \emph{in the thalamus}. We therefore sought a map of the
+thalamic barreloids from the literature. \cite{haidarliu_size_2001} provided a
+comprehensive map of the layout of the barreloids in rat, which is similar to
+mouse \cite{van_der_loos_barreloids_1976}. We used these data to create the
+map of 52 barreloids shown in Fig.~\ref{fig:main}A. We postulated two linearly
+varying molecules and used the barreloids' positions within this field to
+generate the interaction parameters represented by the blue and red arrows.
 
-\cmnt{The realistic barrel pattern experiment is enhanced by stronger
-  branching competition:}
+We configured a simulation based on Eqs.~\ref{eq:dc}--\ref{eq:norm} to run on
+a domain whose shape was matched to data from \cite{shimogori_fibroblast_2005}
+(Fig.~\ref{fig:main}A). 52 separate thalamocortical fields were initialised
+with random noise in the branching density and given our thalamic interaction
+parameters. Two orthogonal gradients were specified; molecule 1 with an
+posterior to anterior gradient and molecule 2 resembling ephrin A5, with a
+medial-lateral gradient. Fig.~\ref{fig:main}D--E shows snapshots of the
+connection density at simulation times x (800 steps), y (2000) and z (3200).
 
-To increase the strength of the competition, we choose an alternative form for
-$f(a)$ in Eq.~\ref{eq:comp} which has a geometric form for small $a$: $f(a) = \frac{1}{1 + e^{-m[a]^l}} - \frac{1}{2}$. The parameters $l$ and $m$ control the steepness and shape of the
-function.
-
+% On stronger branching; not used in the results presented here.
+% \cmnt{The realistic barrel pattern experiment is enhanced by stronger
+%  branching competition:}
+%
+% To increase the strength of the competition, we choose an alternative form for
+% $f(a)$ in Eq.~\ref{eq:comp} which has a geometric form for small $a$: $f(a) =
+% \frac{1}{1 + e^{-m[a]^l}} - \frac{1}{2}$. The parameters $l$ and $m$ control
+% the steepness and shape of the function.
 %%% End of stronger branching competition %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\begin{figure*}
-\begin{center}
-\includegraphics[width=\textwidth]{./MainFig.png}
-\end{center}
-%\vspace{-0em}
-\caption{A. Barreloid centres taken from \cite{haidarliu_size_2001} are shown as
-  black circles. It is assumed that there are two linearly increasing,
-  orthogonal signalling gradients (molecules A \& B) in the thalamus (side
-  graphs), such that the location of each barreloid within the graded thalamic
-  signal field can differentiate the efferent axons, with the result that
-  axons from each barreloid interact to varying degrees with cortical
-  signalling gradients. Coloured arrows show, for each barreloid, the
-  magnitude (and sign) of the interaction that its thalamocortical projection
-  will have with the signalling gradients in the cortex. The mappings from
-  barreloid location to interaction strength are linear. B. }
-\label{fig:main}
-\end{figure*}
 
@@
 experiments like those by Shimogori \& Grove.
+
+The code and configuration required to reproduce the results in this paper are
+available at \url{https://github.com/ABRG-Models/BarrelEmerge}.
 

### 2019-12-03 paper/briefpaper_stuedits/brief.tex
added sreamlined draft

% Brief report style paper

% It's about how a somatotopically ordered pattern can be
% re-established in the cortex based on information carried with
% thalamocortical axons afferent from the ventrobasal thalamus, but
% without specifying absolute positions as centre-points for the
% barrels. Everything in the paper should be tailored towards arguing
% this idea.

\documentclass[9pt,twocolumn,twoside,lineno]{pnas-new}
\templatetype{pnasresearcharticle}

\title{Modelling the emergence of whisker barrels}
% While writing, don't stop for errors
%\nonstopmode

% Use letters for affiliations, numbers to show equal authorship (if applicable) and to indicate the corresponding author
\author[a,1]{Sebastian~S.~James}
\author[a]{Stuart~P.~Wilson}

\affil[a]{Department of Psychology, The University of Sheffield, Sheffield, United Kingdom.}

% $^{*}$Corresponding author email: Seb.James@Sheffield.ac.uk\vspace{1em}\\

\leadauthor{James}

\correspondingauthor{\textsuperscript{1}To whom correspondence should be addressed. E-mail: \href{mailto:seb.james@sheffield.ac.uk}{\nolinkurl{seb.james@sheffield.ac.uk}}}

\keywords{Barrel cortex $|$ Self-organization $|$ Somatotopic map $|$ Axon guidance}

\begin{abstract}
Brain development relies on an interplay of genetic specification and self-organisation. Striking examples of this interplay can be found in the rodent somatosensory brainstem, thalamus, and cortex, where the arrangement of the facial whiskers is preserved in the arrangement of cellular aggregates to form precise somatotopic maps. We show in simulation how realistic whisker maps can self-organize, by assuming that information is exchanged between adjacent cells only, under the guidance of gene expression gradients. The resulting model provides a simple account of how genetic information can constrain spontaneous pattern formation to faithfully reproduce functional maps in subsequent brain structures.
\end{abstract}

\dates{This manuscript was compiled on \today}
\doi{\url{www.pnas.org/cgi/doi/10.1073/pnas.XXXXXXXXXX}}

\begin{document}

% Blue for comments
\newcommand{\cmnt}[1]{\textcolor{blue}{#1}}
\newcommand{\dvrg}{\nabla\vcdot\nabla}
\newcommand{\e}{\emph}
\newcommand{\bol}{\textbf}
\newcommand{\mb}[1]{\mathbf{#1}}
\makeatletter
\newcommand*\vcdot{\mathpalette\vcdot@{.35}}
\newcommand*\vcdot@[2]{\mathbin{\vcenter{\hbox{\scalebox{#2}{$\m@th#1\bullet$}}}}}
\makeatother

\maketitle
\thispagestyle{firststyle}
\ifthenelse{\boolean{shortarticle}}{\ifthenelse{\boolean{singlecolumn}}{\abscontentformatted}{\abscontent}}{}

\modulolinenumbers{}
\linenumbers

\dropcap{S}patial patterns in neural connectivity provide clues about the constraints under which brains evolve and develop \citep{purves_iterated_1992}. Perhaps the most distinctive pattern of neural connectivity can be found in the rodent barrel cortex \cite{woolsey_structural_1970}. The barrels are large cortical columns, identifiable from birth in layer 4 of primary somatosensory cortex as dense clusters of thalamocortical axons, which are enclosed by borders a few neurons thick from postnatal day 3 \citep{erzurumlu_development_2012}. In the plane tangential to the cortical surface the barrels constitute a somatotopic map of the whiskers, with cells of adjacent barrels responding most strongly and quickly to deflection of adjacent whiskers \citep{armstrong-james_flow_1992}. Barrel patterning reflects subcortical whisker maps comprising cell aggregates called barrelettes in the brainstem and barreloids in the thalamus \cite{van_der_loos_barreloids_1976}.

Barrel patterning requires afferent input from whisker stimulation and thalamic calcium waves \citep{anton-bolanos_prenatal_2019}, and depends on a complex network of axon guidance molecules such as ephrin-A5 and A7 and adhesion molecules such as cadherin-6 and 8 \citep{miller_epha7-ephrin-a5_2006}.
This network is orchestrated by interactions between morphogens Fgf8 and Fgf17 and transcription factors Emx2, Pax6, Sp8, and Coup-tf1 \citep{shimogori_fibroblast_2005,bishop_regulation_2000}, which are expressed in gradients that mark orthogonal axes and can be manipulated to stretch, shrink, shift, and even duplicate barrels \cite{assimacopoulos_fibroblast_2012}.

The barrel boundaries form a Voronoi tessellation across the cortical sheet \citep{senft_mouse_1991}, suggesting that barreloid topology is preserved in the projection of thalamocortical axons into the cortex, and that a barrel forms by lateral axon branching from an initial centre-point that ceases upon contact with axons branching from adjacent centres. But the assumption of pre-arranged centres is difficult to resolve with the observation that  axons arrive in the cortical plate as an undifferentiated bundle, prior to barreloid formation \cite{agmon_organized_1993}.

Alternatively, reaction-diffusion dynamics could generate a Voronoi tessellation without pre-arranged centres, by amplifying characteristic modes in a noisy initial distribution of axon branches, as a net effect of short-range cooperative and longer-range competitive interactions. Accordingly, the barrel pattern would be determined by the relative strength of these interactions and by the shape of the cortical field boundary. However, intrinsic cortical dynamics alone cannot account for the topographic correspondence between domains in the thalamus and cortex, or the irregular sizes and specific arrangement of the barrels in rows and arcs, or the influence of gene expression gradients.

The centre-point and reaction-diffusion models are not mutually exclusive. Pre-organised centres could bias reaction-diffusion processes to generate specific  arrangements more reliably, and mechanisms of lateral axon branching may constitute the tension between cooperation and competition required for self-organization. However, proof that barrel patterning can emerge from an undifferentiated bundle of  axons, based only on local interactions, would show that a separate stage and/or extrinsic mechanism for pre-organising thalamocortical connections need not be assumed. To this end, we ask whether barrel maps can emerge in a system with reaction-diffusion dynamics, under the guidance of signalling gradients, and in the absence of pre-defined centres.

\section*{Models}

Karbowski \& Ermentrout \citep{karbowski_model_2004} developed a reaction-diffusion style model of how extrinsic signalling gradients can constrain the emergence of distinct fields from intrinsic cortical dynamics. Their model defines how the density of connections $c(x,t)$, and axon branches $a(x,t)$, interact at time $t$, along a 1D anterior-posterior axis $x$, for $N$ thalamocortical projections indexed by $i$. The model was derived from the assumption that the rates at which $a_i$ and $c_i$ grow are reciprocally coupled. Extending the original 1D model to simulate arealization on a 2D cortical sheet, we use $a_i(\mb{x},t)$ and $c_i(\mb{x},t)$, and  model synaptogenesis as
%
\begin{equation} \label{eq:dc}
\frac{\partial c_i}{\partial t} =-\alpha c_i +\beta  \left(1 - \sum_{k=1}^{N} c_{k}\right)[a_i]^l.
\end{equation}
%
When the final term evaluates to zero, i.e., when the total density of connections sums to one, the density of thalamocortical connections decays at rate $\alpha$. Otherwise the synaptic connection density increases non-linearly ($l>1$) with the axon branching density. Changes in axon branching are modelled as 
%
\begin{equation} \label{eq:da}
\frac{\partial a_i}{\partial t} = \nabla\vcdot\left(D \nabla a_i-a_i\sum_{j=1}^{M} \gamma_{i,j}\nabla \rho_j(\mb{x})\right) - \frac{\partial c_i}{\partial t} + \chi_i.
\end{equation}
%
The first term on the right describes the divergence (indicated by $\nabla\vcdot$) of the quantity in parentheses, which is referred to as the `flux of axonal branching'. The flux represents diffusion across the cortical sheet, at rate $D$, and the influence of $M$ molecular signalling fields, $\rho(\mb{x})$. The influence of a given field (indexed by $j$) on a given thalamic projection (indexed by $i$), is determined by $\gamma_{i,j}$, which may be positive or negative in order that axons may branch in the direction of either higher or lower  concentrations. The second term on the right represents an assumption that axon branching decreases when synaptogenesis increases, and vice versa, i.e., that axon branches are required for synaptic connections to form. Here $\chi_i=0$ is a placeholder.

Software tools for modelling the cortical sheet as a hexagonal lattice with no-flux boundary conditions and an arbitrary boundary shape, for computing the flux using a finite difference method, for numerical integration by the fourth-order Runge-Kutta method, and for reproducing the reported results are available at \url{https://github.com/ABRG-Models/BarrelEmerge}.

\section*{Results}

\begin{figure*}
\begin{center}
\includegraphics[width=\textwidth]{./MainFig.png}
\end{center}

\caption{\textbf{A.} Barreloid centres taken from \cite{haidarliu_size_2001} are shown as black circles. Barreloids are labelled according to the usual convention. Labels R1 to R5 are the rhinal whisker barreloids. The dashed grey line represents the boundary of the barreloid field; the dot-dash line shows the anterior-posterior axis. It is assumed that there are two linearly increasing, orthogonal signalling gradients (molecules T1 \& T2) in the thalamus (side graphs), such that the location of each barreloid within the graded thalamic signal field can differentiate the efferent axons, with the result that axons from each barreloid interact to varying degrees with cortical signalling gradients (shown in panel B). Coloured arrows show, for each barreloid, the magnitude (and sign) of the interaction that its thalamocortical projection will have with the signalling gradients in the cortex. The mappings from barreloid location to interaction strength are linear. \textbf{B.} Guidance molecules in the simulated cortex. Molecule 1 is expressed high to low in an approximately anterior to posterior gradient. Molecule 2 is expressed lateral to medial (analogously to ephrin - Vanderhaegen et al) \textbf{C.}  Whisker barrels from \cite{shimogori_fibroblast_2005}, with annotations for barrel identity and approximate axes.  \textbf{D-F.} Barrel cortex simulation. The maps indicate which thalomocortical axons make maximum connection density. Each hex element is coloured one of 52 unique colours depending on which thalamocortical axon makes the most connections in that element. The mapping between barrel identity and colour is given in F.  \textbf{G.} The Honda Dirichlet measure as a function of simluation time for the simulations show in D-F. \textbf{H.}  Guidance molecules in the simulated Fgf8 misexpression experiment. Molecule 1 is mirrored over the anterior-posterior axis \textbf{I.} The simulation indicates that the mirroring of the barrel field can occur as a result of the mirroring of guidance molecule 1.}
\label{fig:main}
\end{figure*}

First we verified that all results established for a 1D axis could be reproduced using our extension to a 2D cortical sheet. Using an elliptical boundary with $M=3$ guidance gradients aligned to the longer axis, $N=5$ thalamocortical projections gave rise to five distinct cortical fields at locations that preserved the topographic ordering defined by the original $\gamma$ values of \cite{karbowski_model_2004}. However, we found that specifying $N$ ordered areas required $M\approx N$ signalling fields. As the number of distinct guidance fields is unlikely to approach the number of barrels, further extensions were required.

The term in parentheses in Eq.~\ref{eq:dc} represents competition between thalamic projections for a limited availability of cortical connections. To introduce competition also in terms of axon branching we introduced
%
\begin{equation} \label{eq:comp}
\chi_i(\mb{x}, t) = - \frac{\epsilon  a_i}{N-1} \sum_{j \ne i}^{N} a_{j},
\end{equation}
%
which reduces branching for each projection where the branching of other projections is high. Divisive normalization is used to keep the axon branch density bounded on each iteration to the initial density  at $t=0$,

\begin{equation} \label{eq:norm}
  a_i(\mathbf{x}, t) = \int_\Omega  a_i(\mathbf{x}, 0) d\sigma \; \frac {a'_i(\mathbf{x}, t)} {\int_\Omega
  a'_i(\mathbf{x}, t) d\sigma},
\end{equation}
%
where prime symbols indicate the use of intermediate values computed from Eqs.~\ref{eq:dc}--\ref{eq:comp} at time $t-{\delta}t$.

The only differences between thalamic projections are the interaction strengths, $\gamma$, hence any reliable differences between emergent cortical fields must be due to differences in these values only. We chose $\gamma$ values by speculating that interaction strengths with ephrin gradients in the cortical subplate are determined by sampling from similar thalamic gradients at the origin of each thalamic projection, i.e., at each barreloid centre. Two orthogonal linear thalamic gradients were defined, from which 52 pairs of $\gamma$ values were sampled, at the coordinates of 52 barreloid centers recreated from  \cite{haidarliu_size_2001}.

A cortical boundary enclosing 52 corresponding barrels was traced from a cytochrome oxidase stain from  \cite{shimogori_fibroblast_2005}, and Eqs.~\ref{eq:dc}--\ref{eq:norm} were solved for $N=52$ projections on the resulting domain, using $M=2$ linear signalling gradients aligned with the anterior-posterior and  medial-lateral axes. From random initial conditions for $a({\bf x},0)$ and $c({\bf x},0)$, a clear Voronoi-like tesselation emerged. A reduction in the Honda delta metric (see \cite{senft_mouse_1991}) confirmed that a `good' Voronoi pattern emerged within $\approx 3000$ iterations (see Figure 1). The sum of squared differences between the real and simulated barrel centres revealed a strong correspondence.

To further investigate the interplay of genetic and intrinsic mechanisms we simulated a seminal barrel duplication experiment \cite{shimogori_fibroblast_2005},  first by solving on a cortical domain comprising two separate, identically shaped cortical boundaries. In this case one distinct cortical field per thalamic projection emerged, within one of the two boundaries only, i.e., no duplication. However, merging the two fields to create an extended mirror-symmetric boundary shape, with an anterior-posterior guidance field reflected at the join, gave rise to two mirror-symmetrical barrel fields comprising 104 barrels, i.e., two cortical fields for each thalamic projection (see Figure 2). Together these results suggest that by the misexpression of Fgf8, Shimogori \& Grove \citep{shimogori_fibroblast_2005} effectively created one large barrel field rather than two distinct fields.

\section*{Discussion}

The present results suggest that the key requirements for the emergence of realistic barrel patterning are i) at each cortical location thalamocortical projections compete for a limited number of available synaptic connections (Eqs.~\ref{eq:dc}--\ref{eq:da}), ii) at each location the branching rate of a given projection is reduced by the density of other projections (Eqs.~\ref{eq:comp}), and iii) the branch density of each projection is conserved (Eq.~\ref{eq:norm}).

It is important to emphasize that the formulation of the model is entirely local, insofar as simulation requires no information to be communicated between thalamic projections, or from a given cortical location to any but those immediately adjacent. Hence the simulations demonstrate how a self-organising system, constrained by genetically specified guidance cues, can faithfully reproduce an arrangement of cell aggregates in one neural structure as a topographic map in another.

The model also shows that the shape of the barrels may be primarily determined by the competitive and cooperative interactions between cells relative to the shape of the cortical field boundary. And that somatotopic map formation does not require the pre-specification of centre-points by as yet undetermined additional developmental mechanisms.

\bibliography{../BarrelEmerge}

\end{document}

### 2019-12-06 sim/rd_james.h
Use the centroids obtained from the barrel outlines, instead of from the centroid circles

@@ -308,7 +308,7 @@ public:
     bool dirichletComputed = false;
 
     //! Key-mapped coordinates of experimental barrels
-    map<string, pair<float, float>> identified_coords;
+    map<string, pair<float, float>> expt_centroids;
 
     /*!
      * From the contour information in the SVG, determine experimental barrel identity for each Hex,
@@ -465,15 +465,15 @@ public:
     virtual void allocate (void) {
 
         morph::RD_Base<Flt>::allocate();
-
+#ifdef USE_USER_SUPPLIED_CIRCLES
         // Copy the list of circles from the ReadCurves object
-        this->identified_coords = this->r.circles;
+        this->expt_centroids = this->r.circles;
         // Invert the y axis of these coordinates, just as the y axis is inverted in void
         // morph::HexGrid::setBoundary (const BezCurvePath& p)
-        for (auto& c : this->identified_coords) {
+        for (auto& c : this->expt_centroids) {
             c.second.second = -c.second.second;
         }
-
+#endif
         // Resize and zero-initialise the various containers
         this->resize_vector_vector (this->c, this->N);
         this->resize_vector_vector (this->dc, this->N);
@@ -624,6 +624,8 @@ public:
             DBG ("Barrel " << idstr << "/" << er.name << " contains " << regHexes.size() << " hexes");
             Flt theid = this->tc_name_to_id (idstr);
             this->expt_areas[theid] = static_cast<int>(regHexes.size());
+            // This line instead of the code in the ifdef block "USE_USER_SUPPLIED_CIRCLES":
+            this->expt_centroids[idstr] = regCentroid;
             for (auto rh : regHexes) {
                 this->expt_barrel_id[rh->vi] = theid;
             }
@@ -1320,8 +1322,8 @@ public:
         this->sos_distances = 0.0;
         for (unsigned int i = 0; i < this->N; ++i) {
             Flt idx = (Flt)i/(Flt)this->N;
-            Flt dx = this->reg_centroids[idx].first - (this->identified_coords[tcnames[idx]].first - this->hg->originalBoundaryCentroid.first);
-            Flt dy = this->reg_centroids[idx].second - (this->identified_coords[tcnames[idx]].second - this->hg->originalBoundaryCentroid.second);
+            Flt dx = this->reg_centroids[idx].first - (this->expt_centroids[tcnames[idx]].first - this->hg->originalBoundaryCentroid.first);
[…]

### 2019-12-06 paper/briefpaper_stuedits/brief.tex

@@
 Brain development relies on an interplay of genetic specification and self-organisation.
-Striking examples of this interplay can be found in the rodent somatosensory brainstem, thalamus, and cortex, where the arrangement of the facial whiskers is preserved in the arrangement of cellular aggregates to form precise somatotopic maps.
+Striking examples of this interplay can be found in the rodent somatosensory brainstem, thalamus, and cortex, where the arrangement of the facial whiskers is preserved in the arrangement of cell aggregates to form precise somatotopic maps.
 We show in simulation how realistic whisker maps can self-organize, by assuming that information is exchanged between adjacent cells only, under the guidance of gene expression gradients.
@@
 %
-When the final term evaluates to zero, i.e., when the total density of connections sums to one, the density of thalamocortical connections decays at rate $\alpha$.
+Accordingly, when the total density of connections sums to one, the density of thalamocortical connections decays at rate $\alpha$.
 Otherwise the synaptic connection density increases non-linearly ($l>1$) with the axon branching density.
@@
 The term in parentheses in Eq.~\ref{eq:dc} represents competition between thalamic projections for a limited availability of cortical connections.
-To introduce competition also in terms of axon branching we introduced
+To introduce competition also in terms of axon branching we redefined
 %

### 2019-12-06 boundaries/barreloids_haidarliu_Fig5d.py
Reorder boundary numbers. Updates to plotting of Fig5d barreloids

@@ -17,7 +17,7 @@ import math
 D = {}
 
 # Params for the output text
-epsilon = 1200
+epsilon = 200
 alpha = 3
 beta = 20
 xinit = -0.16
@@ -34,13 +34,7 @@ xar = []
 yar = []
 
 # A rotational transformation is applied with this many degrees:
-phi = 0#-40
-
-# -110 rotates the barreloids so that they appear in the same (though mirrored)
-# orientation as in the Erzurumli 2012 review paper
-# phi = -110
-
-#phi = 0
+phi = 105
 
 # Compute cos(phi) and sin(phi)
 pr = phi/360.0 * 2.0 * math.pi # "phi in radians"
@@ -103,16 +97,19 @@ with open('barreloids_haidarliu_Fig5d.csv') as csvDataFile:
 # Read the boundary points
 bndry_x = []
 bndry_y = []
-with open('barreloids_haidarliu_boundary_ordered.csv') as csvDataFile:
+bndry_idstr = []
+with open('barreloids_haidarliu_Fig5d_boundary.csv') as csvDataFile:
     csvReader = csv.reader (csvDataFile)
     for row in csvReader:
         if row[0] == 'op': continue # skip header
+        idstr = row[0]
         x = float(row[1])
         y = float(row[2])
         x_ = x * cosphi - y * sinphi
         y_ = y * cosphi + x * sinphi
         bndry_x.append(x_)
         bndry_y.append(y_)
+        bndry_idstr.append(idstr)
 
 # Produce guidance interactions for 4 gradients (two opposing pairs, orthogonally
 # arranged) or two orthogonal gradients for which interactions will be positive or
@@ -184,6 +181,9 @@ ax0.scatter (xqui, yqui, c='k', s=70, marker='o')
 
 # Plot boundary
 ax0.plot (bndry_x, bndry_y, c='grey', marker='None', linestyle='--', linewidth=2, label='boundary')
+# Optionally add labels
+#for ii in range(0,len(bndry_x)):
+#    ax0.text (bndry_x[ii]+txt_xoff, bndry_y[ii]+txt_yoff, bndry_idstr[ii])
 
 # Plot AP axis. Maybe use pyplot.arrow to make these
 ap = np.array(([[-0.04185,-0.0133],[0.3425,0.1232]]))
[…]

### 2019-12-13 paper/briefpaper_stuedits/brief.tex
paper draft updates

@@
 \dropcap{S}patial patterns in neural connectivity provide clues about the constraints under which brains evolve and develop \citep{purves_iterated_1992}.
-Perhaps the most distinctive pattern of neural connectivity can be found in the rodent barrel cortex \cite{woolsey_structural_1970}.
+Perhaps the most distinctive pattern can be found in the rodent barrel cortex \cite{woolsey_structural_1970}.
 The barrels are large cortical columns, identifiable from birth in layer 4 of primary somatosensory cortex as dense clusters of thalamocortical axons, which are enclosed by borders a few neurons thick from postnatal day 3 \citep{erzurumlu_development_2012}.
@@
 
-Barrel patterning requires afferent input from whisker stimulation and thalamic calcium waves \citep{anton-bolanos_prenatal_2019}, and depends on a complex network of axon guidance molecules such as ephrin-A5 and A7 and adhesion molecules such as cadherin-6 and 8 \citep{miller_epha7-ephrin-a5_2006}.
+Barrel formation requires afferent input from whisker stimulation and thalamic calcium waves \citep{anton-bolanos_prenatal_2019}, and depends on a complex network of axon guidance molecules such as ephrin-A5 and A7 and adhesion molecules such as cadherin-6 and 8 \citep{miller_epha7-ephrin-a5_2006}.
 This network is orchestrated by interactions between morphogens Fgf8 and Fgf17 and transcription factors Emx2, Pax6, Sp8, and Coup-tf1 \citep{shimogori_fibroblast_2005,bishop_regulation_2000}, which are expressed in gradients that mark orthogonal axes and can be manipulated to stretch, shrink, shift, and even duplicate barrels \cite{assimacopoulos_fibroblast_2012}.
@@
 Accordingly, the barrel pattern would be determined by the relative strength of these interactions and by the shape of the cortical field boundary.
-However, intrinsic cortical dynamics alone cannot account for the topographic correspondence between domains in the thalamus and cortex, or the irregular sizes and specific arrangement of the barrels in rows and arcs, or the influence of gene expression gradients.
+However, intrinsic cortical dynamics alone cannot account for the topographic correspondence between thalamic and cortical domains, the irregular sizes and specific arrangement of the barrels in rows and arcs, or the influence of gene expression gradients.
 
@@
 Karbowski \& Ermentrout \citep{karbowski_model_2004} developed a reaction-diffusion style model of how extrinsic signalling gradients can constrain the emergence of distinct fields from intrinsic cortical dynamics.
-Their model defines how the density of connections $c(x,t)$, and axon branches $a(x,t)$, interact at time $t$, along a 1D anterior-posterior axis $x$, for $N$ thalamocortical projections indexed by $i$.
+Their model defines how the density of connections $c(x,t)$ and axon branches $a(x,t)$ interact at time $t$, along a 1D anterior-posterior axis $x$, for $N$ thalamocortical projections indexed by $i$.
 The model was derived from the assumption that the rates at which $a_i$ and $c_i$ grow are reciprocally coupled.
@@
 %
-Accordingly, when the total density of connections sums to one, the density of thalamocortical connections decays at rate $\alpha$.
-Otherwise the synaptic connection density increases non-linearly ($l>1$) with the axon branching density.
-Changes in axon branching are modelled as
+Accordingly, where the total density of synaptic connections sums to one, connections decay at rate $\alpha$.
+Otherwise the connection density increases non-linearly ($l>1$) with the density of axon branching.
+Axon branching is modelled as
 %
@@
 %
-The first term on the right describes the divergence (indicated by $\nabla\vcdot$) of the quantity in parentheses, which is referred to as the `flux of axonal branching'.
+The first term on the right describes the divergence (indicated by $\nabla\vcdot$) of the quantity in parentheses, which is referred to as the `flux' of axonal branching.
 The flux represents diffusion across the cortical sheet, at rate $D$, and the influence of $M$ molecular signalling fields, $\rho(\mb{x})$.
@@
 
-First we verified that all results established for a 1D axis could be reproduced using our extension to a 2D cortical sheet.
-Using an elliptical boundary with $M=3$ guidance gradients aligned to the longer axis, $N=5$ thalamocortical projections gave rise to five distinct cortical fields at locations that preserved the topographic ordering defined by the original $\gamma$ values of \cite{karbowski_model_2004}.
+First we verified that all results established by \cite{karbowski_model_2004} for a 1D axis could be reproduced using our extension to a 2D cortical sheet.
+Using an elliptical boundary with $M=3$ offset guidance gradients aligned to the longer axis, $N=5$ thalamocortical projections gave rise to five distinct cortical fields at locations that preserved the topographic ordering defined by the original $\gamma$ values.
 However, we found that specifying $N$ ordered areas required $M\approx N$ signalling fields.
@@
 
-The term in parentheses in Eq.~\ref{eq:dc} represents competition between thalamic projections for a limited availability of cortical connections.
+The term in parentheses in Eq.~\ref{eq:dc} represents competition between thalamocortical projections for a limited availability of cortical connections.
 To introduce competition also in terms of axon branching we redefined
@@
 %
-which reduces branching for each projection where the branching of other projections is high.
-Divisive normalization is used to keep the axon branch density bounded on each iteration to the initial density  at $t=0$,
+which reduces branching for each projection where the branches of other projections are dense.
+Divisive normalization keeps the axon branch density bounded on each iteration to the initial density  at $t=0$;
 
@@
 
-The only differences between thalamic projections are the interaction strengths, $\gamma$, hence any reliable differences between emergent cortical fields must be due to differences in these values only.
-We chose $\gamma$ values by speculating that interaction strengths with ephrin gradients in the cortical subplate are determined by sampling from similar thalamic gradients at the origin of each thalamic projection, i.e., at each barreloid centre.
-Two orthogonal linear thalamic gradients were defined, from which 52 pairs of $\gamma$ values were sampled, at the coordinates of 52 barreloid centers recreated from  \cite{haidarliu_size_2001}.
+The only differences between thalamic projections are their strengths of interaction with the guidance fields, $\gamma$, hence any reliable differences between emergent cortical fields must be due to differences in these values only.
+We speculate that a projection interacts with a given ephrin field in the cortical subplate with a strength determined by the concentration of a similar molecule at its thalamic origin, i.e., the putative barreloid center.
+As such, two orthogonal linear thalamic gradients were defined, from which 52 pairs of $\gamma$ values were sampled, at the coordinates of 52 barreloid centers recreated from  \cite{haidarliu_size_2001}.
 

### 2019-12-13 paper/briefpaper_stuedits/brief.tex
A few minor edits to paper

@@
 \begin{equation} \label{eq:dc}
-\frac{\partial c_i}{\partial t} =-\alpha c_i +\beta  \left(1 - \sum_{k=1}^{N} c_{k}\right)[a_i]^l.
+\frac{\partial c_i}{\partial t} =-\alpha c_i +\beta  \left(1 - \sum_{k=1}^{N} c_{k}\right)[a_i]^k.
 \end{equation}
@@
 Accordingly, where the total density of synaptic connections sums to one, connections decay at rate $\alpha$.
-Otherwise the connection density increases non-linearly ($l>1$) with the density of axon branching.
+Otherwise the connection density increases non-linearly ($k>1$) with the density of axon branching.
 Axon branching is modelled as
@@
 \begin{center}
-\includegraphics[width=\textwidth]{./MainFig.png}
+\includegraphics[width=\textwidth]{../briefpaper/MainFig.png}
 \end{center}
@@
 We speculate that a projection interacts with a given ephrin field in the cortical subplate with a strength determined by the concentration of a similar molecule at its thalamic origin, i.e., the putative barreloid center.
-As such, two orthogonal linear thalamic gradients were defined, from which 52 pairs of $\gamma$ values were sampled, at the coordinates of 52 barreloid centers recreated from  \cite{haidarliu_size_2001}.
+As such, two orthogonal linear thalamic gradients were defined, from which 41 pairs of $\gamma$ values were sampled, at the coordinates of 41 barreloid centers recreated from \cite{haidarliu_size_2001}.
 
-A cortical boundary enclosing 52 corresponding barrels was traced from a cytochrome oxidase stain from  \cite{shimogori_fibroblast_2005}, and Eqs.~\ref{eq:dc}--\ref{eq:norm} were solved for $N=52$ projections on the resulting domain, using $M=2$ linear signalling gradients aligned with the anterior-posterior and  medial-lateral axes.
+A cortical boundary enclosing 41 corresponding barrels was traced from a cytochrome oxidase stain from \cite{zheng_signal_2001}, and Eqs.~\ref{eq:dc}--\ref{eq:norm} were solved for $N=41$ projections on the resulting domain, using $M=2$ linear signalling gradients aligned with the anterior-posterior and  medial-lateral axes.
 From random initial conditions for $a({\bf x},0)$ and $c({\bf x},0)$, a clear Voronoi-like tesselation emerged.
@@
 In this case one distinct cortical field per thalamic projection emerged, within one of the two boundaries only, i.e., no duplication.
-However, merging the two fields to create an extended mirror-symmetric boundary shape, with an anterior-posterior guidance field reflected at the join, gave rise to two mirror-symmetrical barrel fields comprising 104 barrels, i.e., two cortical fields for each thalamic projection (see Figure 2).
+However, merging the two fields to create an extended mirror-symmetric boundary shape, with an anterior-posterior guidance field reflected at the join, gave rise to two mirror-symmetrical barrel fields comprising NN barrels, i.e., two cortical fields for each thalamic projection (see Figure 2).
 Together these results suggest that by the misexpression of Fgf8, Shimogori \& Grove \citep{shimogori_fibroblast_2005} effectively created one large barrel field rather than two distinct fields.
@@
 
-The present results suggest that the key requirements for the emergence of realistic barrel patterning are i) at each cortical location thalamocortical projections compete for a limited number of available synaptic connections (Eqs.~\ref{eq:dc}--\ref{eq:da}), ii) at each location the branching rate of a given projection is reduced by the density of other projections (Eqs.~\ref{eq:comp}), and iii) the branch density of each projection is conserved (Eq.~\ref{eq:norm}).
+The present results suggest that the key requirements for the emergence of realistic barrel patterning are i) at each cortical location thalamocortical projections compete for a limited number of available synaptic connections (Eqs.~\ref{eq:dc}--\ref{eq:da}), ii) at each location the branching rate of a given projection is reduced by the density of other projections (Eq.~\ref{eq:comp}), and iii) the branch density of each projection is conserved (Eq.~\ref{eq:norm}).
 

### 2019-12-16 paper/briefpaper_stuedits/brief.tex
WIP on figure caption

@@
 \begin{equation} \label{eq:dc}
-\frac{\partial c_i}{\partial t} =-\alpha c_i +\beta  \left(1 - \sum_{k=1}^{N} c_{k}\right)[a_i]^k.
+\frac{\partial c_i}{\partial t} =-\alpha c_i +\beta  \left(1 - \sum_{j=1}^{N} c_{j}\right)[a_i]^k.
 \end{equation}
@@
 
-\caption{\textbf{A.} Barreloid centres taken from \cite{haidarliu_size_2001} are shown as black circles.
-Barreloids are labelled according to the usual convention.
-Labels R1 to R5 are the rhinal whisker barreloids.
-The dashed grey line represents the boundary of the barreloid field; the dot-dash line shows the anterior-posterior axis.
-It is assumed that there are two linearly increasing, orthogonal signalling gradients (molecules T1 \& T2) in the thalamus (side graphs), such that the location of each barreloid within the graded thalamic signal field can differentiate the efferent axons, with the result that axons from each barreloid interact to varying degrees with cortical signalling gradients (shown in panel B).
-Coloured arrows show, for each barreloid, the magnitude (and sign) of the interaction that its thalamocortical projection will have with the signalling gradients in the cortex.
-The mappings from barreloid location to interaction strength are linear.
-\textbf{B.} Guidance molecules in the simulated cortex.
-Molecule 1 is expressed high to low in an approximately anterior to posterior gradient.
-Molecule 2 is expressed lateral to medial (analogously to ephrin - Vanderhaegen et al) \textbf{C.}  Whisker barrels from \cite{shimogori_fibroblast_2005}, with annotations for barrel identity and approximate axes.
-\textbf{D-F.} Barrel cortex simulation.
-The maps indicate which thalomocortical axons make maximum connection density.
-Each hex element is coloured one of 52 unique colours depending on which thalamocortical axon makes the most connections in that element.
-The mapping between barrel identity and colour is given in F.
-\textbf{G.} The Honda Dirichlet measure as a function of simluation time for the simulations show in D-F.
-\textbf{H.}  Guidance molecules in the simulated Fgf8 misexpression experiment.
-Molecule 1 is mirrored over the anterior-posterior axis \textbf{I.} The simulation indicates that the mirroring of the barrel field can occur as a result of the mirroring of guidance molecule 1.}
+\caption{\textbf{C.} Cortical whisker barrels from \cite{zheng_signal_2001}
+  with annotations for approximate axes and lines drawn to join the
+  interstices into a putative Voronoi diagram.  \textbf{B.}  Top: Stochastic
+  initial condition for $a_i$ with $i=13$ (barrel C4). The roll-off towards
+  the boundary (thin line) ensures the boundary conditions are
+  satisfied. Bottom: Guidance molecules in the simulated cortex. Molecule 1 is
+  expressed high to low in an approximately anterior to posterior
+  gradient. Molecule 2 is expressed lateral to medial (analogously to ephrin -
+  Vanderhaegen et al) \textbf{C.}  Barreloid centres for mystacial whiskers
+  taken from \cite{haidarliu_size_2001} are shown as black circles. The dashed
+  grey line shows the boundary of the barreloid field. It is assumed that
+  there are two linearly increasing, orthogonal signalling molecules in the
+  thalamus, such that the location of each barreloid within the graded
+  thalamic signal field can differentiate the efferent axons, with the result
+  that axons from each barreloid interact to varying degrees with the cortical
+  signalling gradients shown in B. Coloured arrows show, for each barreloid,
+  the magnitude (and sign) of the interaction that its thalamocortical
+  projection will have with each signalling gradient in the cortex. The
+  mappings from barreloid location to interaction strength are
+  linear. \textbf{D.}  Barrel cortex simulation. The maps indicate which
+  thalomocortical axons make maximum connection density. Each hex element is
+  coloured one of 41 \cmnt{FIXME: Got to here} unique colours depending on which thalamocortical axon
+  makes the most connections in that element. The mapping between barrel
+  identity and colour is given in F.  \textbf{G.} The Honda Dirichlet measure
+  as a function of simluation time for the simulations show in
+  D-F. \textbf{H.}  Guidance molecules in the simulated Fgf8 misexpression
+  experiment. Molecule 1 is mirrored over the anterior-posterior axis
+  \textbf{I.} The simulation indicates that the mirroring of the barrel field
+  can occur as a result of the mirroring of guidance molecule 1.}
 \label{fig:main}

### 2019-12-16 paper/briefpaper/brief.tex
Changes to figure and caption with Stuart

--- 2019-11-27 paper/briefpaper/brief.tex
@@
 
-%\title{Self-organization of whisker barrels in simulation requires competition between thalamocortical axons}
-%\title{Simulated barrel cortex self-organizes using information carried by thalamocortical axons}
-\title{Information carried by thalamocortical axons enables self-organization of the barrel field}
+\title{Modelling the emergence of whisker barrels}
 % While writing, don't stop for errors
@@
 
-% Please give the surname of the lead author for the running footer
 \leadauthor{James}
-
-% Please add a significance statement to explain the relevance of your work
-\significancestatement{The mystacial whiskers of rodents are a complex,
-  spatially ordered set of structures whose pattern is repeated in several
-  locations in the brain. We contribute a self-organising mechanism by which
-  the pattern can be recreated with minimal genetic pre-determination at each
-  location; only two pairs of axon guidance gradients are required in each
-  tissue. The pattern's togography is carried by the axons which grow from one
-  tissue to another and interact variably (depending upon which whisker they
-  are associated with) with the guidance gradients in the target tissue whilst
-  competing with each other for space. We demonstrate that this mechanism
-  recreates a realistic whisker barrel field and also accounts for the results
-  of well-known gene misexpression experiments.}
 
@@
 
-% At least three keywords are required at submission. Please provide three to five keywords, separated by the pipe symbol.
-\keywords{Barrel cortex $|$ Self-organization $|$ Somatotopy $|$ Center-point $|$ Pattern}
+\keywords{Barrel cortex $|$ Self-organization $|$ Somatotopic map $|$ Axon guidance}
 
 \begin{abstract}
-Understanding how genetic information and spontaneous pattern formation
-together give rise to the whisker barrels is an important challenge for
-developmental neurobiology. The barrel patterning that coalesces in layer IV
-has been described formally as a Voronoi tesselation, which is indicative of
-an underlying developmental process based on local competitive
-interactions. If pattern formation relies only on such intrinsic
-self-organising processes, how can the somatotopic order at the periphery be
-maintained in the projections from brainstem to thalamus and from thalamus to
-cortex?
-%
-To investigate, we began by extending a existing computational model of axon
-branching and synaptogenesis in thalamocortical projections under the
-influence of molecular signalling gradients.  We found that, although these
-processes alone are incapable of generating patterns of cortical fields that
-are more intricate than the patterns already present in the signalling
-gradients, the addition of competition between the branches of axons that
-originate from different thalamic barreloids allowed a full barrel-field
-pattern to emerge in the presence of only two gradients.
-%
-%However, a
-%full barrel-field pattern emerges under the guidance of as few as four
-%signalling gradients when the model is extended to introduce competition
-%between the branches of axons that originate from different thalamic
-%barreloids.
-%
-%We show that the key requirements for the emergence of realistic
-%(Voronoi tesselation) patterning are i) at each cortical location
-%thalamocortical projections compete for a limited number of available
-%connections, ii) at each location the axonal branching rate of each
-%thalamocortical projection is reduced by the total branching density of the
-%other projections, and iii) the density of axon branching for each
-%thalamocortical projection is conserved.
-%
-We demonstrate how a purely
-self-organising system can initialise a pattern which can then be faithfully
-replicated in multiple neural structures.
-%We show that this minimal model of
-%barrel self-organisation provides a parsimonious account of data from a range
-%of classic studies in which intrinsic and extrinsic factors have been
-%experimentally manipulated.
+Brain development relies on an interplay of genetic specification and self-organisation.
+Striking examples of this interplay can be found in the rodent somatosensory brainstem, thalamus, and cortex, where the arrangement of the facial whiskers is preserved in the arrangement of cell aggregates to form precise somatotopic maps.
+We show in simulation how realistic whisker maps can self-organize, by assuming that information is exchanged between adjacent cells only, under the guidance of gene expression gradients.
+The resulting model provides a simple account of how genetic information can constrain spontaneous pattern formation to faithfully reproduce functional maps in subsequent brain structures.
 \end{abstract}
@@
 \newcommand{\cmnt}[1]{\textcolor{blue}{#1}}
-% Divergence symbol (Del dot Del)
 \newcommand{\dvrg}{\nabla\vcdot\nabla}
-% Emphasis shortcut
 \newcommand{\e}{\emph}
-% Bold shortcut
 \newcommand{\bol}{\textbf}
-% Math bold
 \newcommand{\mb}[1]{\mathbf{#1}}
@@
 
-\dropcap{S}patial patterns in neural connectivity provide clues about the
-constraints under which brains evolve and develop.
-%Whether or not patterning
-%governs information processing, models of the development of information
-%processing circuits can be validated in terms of their ability to account for
-%biological patterning \citep{purves_iterated_1992}.
-%%see also \citealp{wilson_what_2015,bednar_cortical_2016}
-One of the most
-distinctive patterns of neural connectivity is found in the rodent
-barrel cortex \cite{woolsey_structural_1970}.
-%(\citealp{woolsey_pattern_1948,woolsey_structural_1970,welker_structure_1974},
-%see \citealp{fox_barrel_2008}).
-Understanding how this particular pattern forms might shed light on the
-interplay of genetic and extrinsic factors that shape cortical circuits.
-%, and on a wealth of
-%anatomical, physiological, and behavioural data obtained using this
-%model system.
+\dropcap{S}patial patterns in neural connectivity provide clues about the constraints under which brains evolve and develop \citep{purves_iterated_1992}.
+Perhaps the most distinctive pattern can be found in the rodent barrel cortex \cite{woolsey_structural_1970}.
+The barrels are large cortical columns, identifiable from birth in layer 4 of primary somatosensory cortex as dense clusters of thalamocortical axons, which are enclosed by borders a few neurons thick from postnatal day 3 \citep{erzurumlu_development_2012}.
+In the plane tangential to the cortical surface the barrels constitute a somatotopic map of the whiskers, with cells of adjacent barrels responding most strongly and quickly to deflection of adjacent whiskers \citep{armstrong-james_flow_1992}.
+Barrel patterning reflects subcortical whisker maps comprising cell aggregates called barrelettes in the brainstem and barreloids in the thalamus \cite{van_der_loos_barreloids_1976}.
 
-The barrels are large cortical columns, identifiable from birth in layer 4 of
-primary somatosensory cortex as dense clusters of thalamocortical axons, which
-are enclosed by borders a few neurons thick from postnatal day 3
-\citep{erzurumlu_development_2012}. In the plane tangential to the cortical
-surface the barrels constitute a map of the arrangement of the whiskers on the
-face,
-%(see \citealp{yamakado_subdivision_1979})
-such that the cells of adjacent barrels respond most strongly and most quickly
-to deflection of adjacent whiskers \citep{armstrong-james_flow_1992}. The
-barrels reflect subcortical whisker maps comprising cellular aggregates known
-as barreloids in the thalamus \cite{van_der_loos_barreloids_1976} and
-barrelettes in the brainstem.
-%\citep{killackey_pattern_1980}.
+Barrel formation requires afferent input from whisker stimulation and thalamic calcium waves \citep{anton-bolanos_prenatal_2019}, and depends on a complex network of axon guidance molecules such as ephrin-A5 and A7 and adhesion molecules such as cadherin-6 and 8 \citep{miller_epha7-ephrin-a5_2006}.
+This network is orchestrated by interactions between morphogens Fgf8 and Fgf17 and transcription factors Emx2, Pax6, Sp8, and Coup-tf1 \citep{shimogori_fibroblast_2005,bishop_regulation_2000}, which are expressed in gradients that mark orthogonal axes and can be manipulated to stretch, shrink, shift, and even duplicate barrels \cite{assimacopoulos_fibroblast_2012}.
 
-%Each whisker and its corresponding barrelette, barreloid, and barrel can be
-%identified by its position on a two-dimensional grid.
-%\citep{zucker_coding_1969,killackey_pattern_1980,van_der_loos_barreloids_1976,haidarliu_size_2001}.
-%
-%The barrels are normally up to twice as long in the axis aligned to the grid
-%‘rows’ compared to the ‘arcs’, and the rows are visible from day 2
-%\citep{rebsam_refinement_2002} whereas the arcs become distinct by day 3
-%\citep{erzurumlu_development_2012,rebsam_refinement_2002}.
-%In metabotropic glutamate receptor 5 (mGlu5) knockouts the barrel rows develop
-%but the arcs do not (\citealp{hannan_plc-1_2001} see also
-%\citealp{fox_barrel_2008}), thus distinct processes likely guide development
-%of the barrels with respect to the orthogonal grid axes.
+The barrel boundaries form a Voronoi tessellation across the cortical sheet \citep{senft_mouse_1991}, suggesting that barreloid topology is preserved in the projection of thalamocortical axons into the cortex, and that a barrel forms by lateral axon branching from an initial centre-point that ceases upon contact with axons branching from adjacent centres.
+But the assumption of pre-arranged centres is difficult to resolve with the observation that  axons arrive in the cortical plate as an undifferentiated bundle, prior to barreloid formation \cite{agmon_organized_1993}.
 
-Barrel formation requires afferent input, driven by whisker stimulation and by
-thalamic calcium waves that propagate via mGlu5-receptor activation
-\citep{anton-bolanos_prenatal_2019}.
-%(\citealp{anton-bolanos_prenatal_2019}, see also
-%\citealp{anton-bolanos_developmental_2018}).
-Patterning also depends on a complex network of axon guidance processes, which
-includes interactions between guidance molecules like ephrin-A5 and A7
-\citep{miller_epha7-ephrin-a5_2006} and between adhesion molecules like
-cadherin-6 and 8 \citep{bishop_regulation_2000}, and the
-binding of p75 to neurotropins
-\citep{bishop_distinct_2002}.
-% or both citations?
-%\citep{shimogori_fibroblast_2005,bishop_distinct_2002}.
-%
-%(\citealp{shimogori_fibroblast_2005,bishop_distinct_2002}
-%; see also \citealp{dye_lifespan_2011,dye_lifespan_2011-1}).
-This network is orchestrated by interactions between morphogens Fgf8 and Fgf17
-and transcription factors Emx2, Pax6, Sp8, and Coup-tf1
-\citep{shimogori_fibroblast_2005,bishop_regulation_2000}, which are expressed
-in gradients that mark orthogonal axes and can be manipulated to cause the
-barrels to stretch, shrink, shift, skew, and even duplicate \cite{assimacopoulos_fibroblast_2012}.
-% A longer list of citations:
-%\citep{shimogori_fibroblast_2005,assimacopoulos_fibroblast_2012,borello_sp8_2014,sahara_sp8_2007,ypsilanti_transcriptional_2016,sur_patterning_2005}.
-%
-While complex structure in protein expression can be observed for some genes,
-such as the molecules Bhlhb5 \citep{joshi_bhlhb5_2008}, a complex gene
-expression pattern with the complexity of the barrel field has not, to our
-knowledge been reported and so the mechanism by which the barrels form seems
-likely to be one in which the structure forms by self-organization.
+Alternatively, reaction-diffusion dynamics could generate a Voronoi tessellation without pre-arranged centres, by amplifying characteristic modes in a noisy initial distribution of axon branches, as a net effect of short-range cooperative and longer-range competitive interactions.
+Accordingly, the barrel pattern would be determined by the relative strength of these interactions and by the shape of the cortical field boundary.
+However, intrinsic cortical dynamics alone cannot account for the topographic correspondence between thalamic and cortical domains, the irregular sizes and specific arrangement of the barrels in rows and arcs, or the influence of gene expression gradients.
 
-The boundaries of the barrels form a Voronoi tesselation across the cortical
-sheet \citep{senft_mouse_1991}. This patterning suggests that the barreloid
-topology may be preserved in the initial projection of the thalamocortical
-axons into the cortex, and that a barrel forms by the lateral branching of the
-axons from their initial centres of mass, which ceases upon contact with the
-axons branching from adjacent centres. According to this model the barrel
-pattern could be entirely determined by the distribution of the initial centre
-points, which would itself be determined by guidance cues. However, the
-assumption of pre-arranged centres is difficult to resolve with the
-observation that thalamocortical axons arrive in the cortical plate as an
-undifferentiated bundle prior to barreloid formation within the thalamus
-\cite{agmon_organized_1993}. % Plus erzurumli, probably
-
-Alternatively, reaction-diffusion dynamics could generate a Voronoi-like
-tesselation in the cortex without the need for pre-arranged centres, by
-amplifying characteristic modes in the noisy initial distribution of bundled
-axon branches, as a net effect of short-range cooperative and longer-range
-competitive interactions between cells. According to this model the barrel
-pattern would be determined by the relative strength of these interactions and
-by the shape of the cortical field boundary. However, reaction-diffusion
-dynamics alone are unlikely to be able to account for the precise
-co-registration and topographic connectivity between emergent domains in
-thalamus and cortex, the non-uniform sizes and spacing of the barrels to form
-rows versus arcs, and the influence of gene expression gradients.
-
-As theories of barrel patterning, the centre-point and reaction-diffusion
-models are not mutually exclusive. Pre-organised centres in the initial
-conditions could bias reaction-diffusion processes to generate the specific
-somatotopic arrangement of barrels more reliably, and similarly the mechanisms
-by which axons branch laterally from barrel centres may involve a tension
-between cooperative and competitive interactions. However, a proof that barrel
-patterning can in principle emerge within an undifferentiated bundle of
-thalamocortical axons, based only on local interactions between cells, would
-remove the need to invoke a global supervisor, i.e., to assume a separate
-stage and/or extrinsic mechanism for pre-organising thalamocortical
-connections.
-
-We seek to establish whether it is possible for the observed arrangement of
-barrels to self-organise in a system governed by reaction-diffusion dynamics,
-under the guidance of molecular signalling gradients, and in the absence of
-pre-defined centre-points.
-%
-Our approach is to represent in a computational model the minimal set of
-biological assumptions that might lead to the emergence of barrel patterning
-in simulation, and, if necessary, to add to this minimal set only assumptions
-that can be represented by the local exchange of information between adjacent
-cells. The emergence of barrel patterning under such conditions would serve as
-an existence proof of the plausibility of the hypothesis that barrel
-development is a self-organising process.
+The centre-point and reaction-diffusion models are not mutually exclusive.
+Pre-organised centres could bias reaction-diffusion processes to generate specific  arrangements more reliably, and mechanisms of lateral axon branching may constitute the tension between cooperation and competition required for self-organization.
+However, proof that barrel patterning can emerge from an undifferentiated bundle of  axons, based only on local interactions, would show that a separate stage and/or extrinsic mechanism for pre-organising thalamocortical connections need not be assumed.
+To this end, we ask whether barrel maps can emerge in a system with reaction-diffusion dynamics, under the guidance of signalling gradients, and in the absence of pre-defined centres.
 
@@
 
-Karbowski \& Ermentrout \citep{karbowski_model_2004} developed a dynamical
-model of cortical arealization capable of generating predictions about the
-influence of signalling molecules on the size, shape, and topological
-organization of emergent cortical fields. Their model describes competition
-between thalamocortical projections for territory along an anterior-posterior
-cortical axis. The equations of the model define how the density of
-connections, $c(x,t)$, and the density of axon branches, $a(x,t)$, interact at time $t$
-at each cortical location $x$, for $N$ thalamocortical projections indexed by
-$i$.
-%The Voronoi tesselation used to describe barrel patterning is a
-%two-dimensional structure, hence we first extended the original model for
-%simulation on a two-dimensional cortical sheet, and describe it here as such
-%(i.e., substituting $x$ with the two-dimensional position vector $\mb{x}$ and
-%substituting terms of the form $\frac{d}{dx}$ with the spatial gradient
-%operator $\nabla$).
-%
-The model was derived from the assumption that the rates at which
-thalamocortical synapses form and thalamocortical axons branch are
-reciprocally coupled, resulting in expressions for the change in $c_i$ as a
-function of $a_i$, and for the change in $a_i$ as a function of
-$c_i$. Synaptogenesis is modelled [in two dimensions, so the variables are
-$a_i(\mb{x},t)$ \& $c_i(\mb{x},t)$] as
+Karbowski \& Ermentrout \citep{karbowski_model_2004} developed a reaction-diffusion style model of how extrinsic signalling gradients can constrain the emergence of distinct fields from intrinsic cortical dynamics.
+Their model defines how the density of connections $c(x,t)$ and axon branches $a(x,t)$ interact at time $t$, along a 1D anterior-posterior axis $x$, for $N$ thalamocortical projections indexed by $i$.
+The model was derived from the assumption that the rates at which $a_i$ and $c_i$ grow are reciprocally coupled.
+Extending the original 1D model to simulate arealization on a 2D cortical sheet, we use $a_i(\mb{x},t)$ and $c_i(\mb{x},t)$, and  model synaptogenesis as
 %
 \begin{equation} \label{eq:dc}
-\frac{\partial c_i}{\partial t} =-\alpha c_i +\beta  \left(1 - \sum_{i'=1}^{N} c_{i'}\right)[a_i]^k.
+\frac{\partial c_i}{\partial t} =-\alpha c_i +\beta  \left(1 - \sum_{j=1}^{N} c_{j}\right)[a_i]^k.
 \end{equation}
 %
-When the final term evaluates to zero, the density of thalamocortical
-connections decays exponentially at a rate given by $\alpha$. This occurs when
-the density of connections over all thalamic projections sums to one;
-interpreted as a state in which all available cortical synaptic receptors are
-occupied. Otherwise, the connection density is assumed to vary non-linearly
-($k>1$) with the axon branching density.
-
-Changes in axon branching density are assumed to depend also on the influence
-of guidance molecules and on the density of branches at adjacent cortical
-locations:
+Accordingly, where the total density of synaptic connections sums to one, connections decay at rate $\alpha$.
+Otherwise the connection density increases non-linearly ($k>1$) with the density of axon branching.
+Axon branching is modelled as
 %
 \begin{equation} \label{eq:da}
-\frac{\partial a_i}{\partial t} = \nabla\vcdot\left(D \nabla a_i-a_i\sum_{j=1}^{M} \gamma_{i,j}\nabla \rho_j(\mb{x})\right) - \frac{\partial c_i}{\partial t} + \chi.
+\frac{\partial a_i}{\partial t} = \nabla\vcdot\left(D \nabla a_i-a_i\sum_{j=1}^{M} \gamma_{i,j}\nabla \rho_j(\mb{x})\right) - \frac{\partial c_i}{\partial t} + \chi_i.
 \end{equation}
 %
-The first term on the right describes the divergence (indicated by
-$\nabla\vcdot$) of a quantity referred to as the `flux of axonal branching'
-(the term enclosed by parentheses, called $\mb{J}_i(\mb{x},t)$).
-%The central idea in heat diffusion
-%models is that the change in temperature (over time) at a given location is
-%proportional to the gradient in the temperature (over space) at that
-%location. Substituting the branching density for heat, axons from a given
-The flux consists of lateral diffusion across the cortical sheet at a rate
-given by the diffusion constant $D$, and an interaction term, in which
-branching is influenced by gradients in the concentration of $M$ distinct
-molecular signals, $\rho(\mb{x})$. The influence of a given signal (indexed by
-$j$) on a given thalamic projection (indexed by $i$), is determined by the
-parameter $\gamma_{i,j}$ which may be positive or negative such that axons of
-a given thalamic projection may branch preferentially towards either higher or
-lower concentrations of each signalling molecule.
-%Interactions for
-%which this value is positive encourage axon branching in the direction in
-%which the concentration of the signal is increasing, whereas negative
-%interactions encourage branching in the opposite direction. Hence the axons of
-%a given thalamic projection are assumed to branch towards either higher or
-%lower concentrations of each signalling molecule.
+The first term on the right describes the divergence (indicated by $\nabla\vcdot$) of the quantity in parentheses, which is referred to as the `flux' of axonal branching.
+The flux represents diffusion across the cortical sheet, at rate $D$, and the influence of $M$ molecular signalling fields, $\rho(\mb{x})$.
+The influence of a given field (indexed by $j$) on a given thalamic projection (indexed by $i$), is determined by $\gamma_{i,j}$, which may be positive or negative in order that axons may branch in the direction of either higher or lower  concentrations.
+The second term on the right represents an assumption that axon branching decreases when synaptogenesis increases, and vice versa, i.e., that axon branches are required for synaptic connections to form.
+Here $\chi_i=0$ is a placeholder.
 
-The second term on the right of Eq.~\ref{eq:da} represents an assumption that the
-axon branching density decreases when the number of connections being made
-increases, and vice versa, i.e., that branches are required to form synaptic
-connections. Note that we introduce the term $\chi=0$ here, so that we might
-subsequently redefine $\chi$ to introduce competition between thalamic
-projections.
-%that we will show to be necessary for the emergence of a
-%somatotopic pattern.
-
-Two points are important to emphasise: First, notice that the only
-difference between thalamic projections are the interaction strengths,
-$\gamma_{i,j}$. Any differences between the size, shape, and location of
-cortical fields that reliably emerge from the dynamics of the model must be
-due to differences in the choice of these values only; Second, the formulation
-of the model is at this point entirely local, in the sense that simulating the
-dynamics requires no information to be communicated from cortical locations to
-any but the immediately adjacent locations.
-
-In addition to the choice of initial conditions (for the fields
-$c(\mb{x},t=0)$ and $a(\mb{x},t=0)$) and the axon guidance fields
-$\rho(\mb{x})$, the free parameters of the model are the number of
-thalamocortical projections, $N$, constants $\alpha$, $\beta$, $k$, and $D$,
-and the $N\times M$ interaction strengths, $\gamma$.
-
-The dynamics of the model (Eqs.~\ref{eq:dc} \& \ref{eq:da}) were solved on a
-hexagonal lattice of grid elements representing the cortical sheet, under the
-assumption that axon branches cannot extend beyond the boundary of the sheet,
-i.e., $\mb{J}_i(\mb{x},t) \big\rvert_{boundary} = 0$. We developed software
-tools for imposing arbitrary boundary shapes on the lattice. The dynamics were
-numerically integrated using fourth-order Runge-Kutta, and using a
-{\color{red} finite difference method} to compute the flux term (see
-Supplementary Materials).
+Software tools for modelling the cortical sheet as a hexagonal lattice with no-flux boundary conditions and an arbitrary boundary shape, for computing the flux using a finite difference method, for numerical integration by the fourth-order Runge-Kutta method, and for reproducing the reported results are available at \url{https://github.com/ABRG-Models/BarrelEmerge}.
 
@@
 
-%\cmnt{If we need to present the results of the 2-D system with $\chi=0$, then
-%  we'll do that first, showing a 2-D version of the simulation obtained by
-%  running \code{james0 configs/karb.json}}
-
-We first verified that the system given by Eqs.~\ref{eq:dc} \& \ref{eq:da}
-described a two dimensional extension of the axon branching model in
-\cite{karbowski_model_2004}. We showed that in the presence of three guidance
-gradients varying only in one dimension (along the long axis of an elliptical
-domain), five thalamocortical axon types, each interacting uniquely with the
-three guidance gradients, would segregate into five separate regions. We
-investigated a number of configurations of the guidance gradients, varying
-their shape, $\rho(\mb{x})$, and number, $M$, and their effect on a varying
-number of thalamocortical projections, $N$. In general, we were not able to
-produce a complexity of structure in the resulting patterns of the
-connections, $c(\mb{x},t\rightarrow\infty)$, which was not present in the
-guidance patterns.
-
-\subsection*{Branching competition is required to form somatotopically ordered barrels}
-
-To investigate how the whisker barrel structure might form in the presence of
-only monotonically varying morphogen gradients, we turned our attention to the
-competition for connections built into the model. There does exist an element
-of competition in the original model (Eqs.~\ref{eq:dc} \& \ref{eq:da}). The
-term $\left(1 - \sum_{i=1}^{N} c_i\right)$ in Eq.~\ref{eq:dc} ensures that
-once the available synaptic receptors in a region have been fully occupied by
-one thalamocortical axon type, there is no longer the possibility for other
-types to form synapses. This competition turns out to be relatively weak,
-because it does not strongly affect the branching density in that region; it
-is possible for two or more TC axon types to have roughly equal branching
-density in some region, leading to roughly equal numbers of connections
-forming for each TC type, up until the `hard' connection limit is reached.
-
-To introduce competition between the \emph{branching} of thalamocortical axon
-types, we introduced the new term $\chi$ given by
-%
-\begin{equation} \label{eq:comp}
-\chi(\mb{x}, t) = - \frac{\epsilon  a_i}{N-1} \sum_{i' \ne i}^{N} f(a_{i'}).
-\end{equation}
-%
-with the choice $f(a) = a$, which leads to a reduction in the rate of increase
-of branching density for axon type $i$ in regions where the branching density
-of \emph{other} axon types is high. The new parameter $\epsilon$ governs the
-strength of the branching competition and the factor $1/(N-1)$ ensures that the
-strength scales with $N$.
-
-With the addition of the competition term, $\chi$, it is necessary to
-introduce a normalisation step, in order to prevent one TC type from
-out-competing every other TC type. Normalisation represents the idea that for
-a fixed number of afferent thalamocortical axons, the overall number of
-branches must have some upper limit; given the finite resources available in
-each axon for building new processes. We divisively normalize the branching
-density, dividing it at each computational step by the mean branching density
-per unit area. It is then multiplied by the branching density per unit area at
-time 0, to account for the possibility that there might be more axons afferent
-for TC type $1$ cf.~TC type $2$:
-%
-\begin{equation} \label{eq:norm}
-  a_i(\mathbf{x}, t) = \int_\Omega  a_i(\mathbf{x}, 0) d\sigma \; \frac {a'_i(\mathbf{x}, t)} {\int_\Omega
-  a'_i(\mathbf{x}, t) d\sigma}
-\end{equation}
-%
-The prime symbols indicate the use of the intermediate value of $a_i$ after it
-has been computed according to Eqs.~\ref{eq:dc}--\ref{eq:comp} on the
-basis of the system state at time $t-{\delta}t$.
-
-\subsection*{The model can transfer realistic barrel patterns}
-
-In which we show that a realistic barrel pattern in the thalamus can be
-recreated in barrel cortex.
-
-\cmnt{The realistic barrel pattern experiment is enhanced by stronger
-  branching competition:}
-
-To increase the strength of the competition, we choose an alternative form for
-$f(a)$ in Eq.~\ref{eq:comp} which has a geometric form for small $a$: $f(a) = \frac{1}{1 + e^{-m[a]^l}} - \frac{1}{2}$. The parameters $l$ and $m$ control the steepness and shape of the
-function.
-
-%%% End of stronger branching competition %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
 \begin{figure*}
 \begin{center}
-\includegraphics[width=\textwidth]{./MainFig.png}
+\includegraphics[width=\textwidth]{../briefpaper/MainFig.png}
 \end{center}
-%\vspace{-0em}
-\caption{A. Barreloid centres taken from \cite{haidarliu_size_2001} are shown as
-  black circles. It is assumed that there are two linearly increasing,
-  orthogonal signalling gradients (molecules A \& B) in the thalamus (side
-  graphs), such that the location of each barreloid within the graded thalamic
-  signal field can differentiate the efferent axons, with the result that
-  axons from each barreloid interact to varying degrees with cortical
-  signalling gradients. Coloured arrows show, for each barreloid, the
-  magnitude (and sign) of the interaction that its thalamocortical projection
-  will have with the signalling gradients in the cortex. The mappings from
-  barreloid location to interaction strength are linear. B. }
+
+\caption{\textbf{C.} Cortical whisker barrels from \cite{zheng_signal_2001}
+  with annotations for approximate axes and lines drawn to join the
+  interstices into a putative Voronoi diagram.  \textbf{B.}  Top: Stochastic
+  initial condition for $a_i$ with $i=13$ (barrel C4). The roll-off towards
+  the boundary (thin line) ensures the boundary conditions are
+  satisfied. Bottom: Guidance molecules in the simulated cortex. Molecule 1 is
+  expressed high to low in an approximately anterior to posterior
+  gradient. Molecule 2 is expressed lateral to medial (analogously to ephrin -
+  Vanderhaegen et al) \textbf{C.}  Barreloid centres for mystacial whiskers
+  taken from \cite{haidarliu_size_2001} are shown as black circles. The dashed
+  grey line shows the boundary of the barreloid field. It is assumed that
+  there are two linearly increasing, orthogonal signalling molecules in the
+  thalamus, such that the location of each barreloid within the graded
+  thalamic signal field can differentiate the efferent axons, with the result
+  that axons from each barreloid interact to varying degrees with the cortical
+  signalling gradients shown in B. Coloured arrows show, for each barreloid,
+  the magnitude (and sign) of the interaction that its thalamocortical
+  projection will have with each signalling gradient in the cortex. The
+  mappings from barreloid location to interaction strength are
+  linear. \textbf{D.}  Barrel cortex simulation. Colours indicate
+  $\arg\max_i(c_i)$ for 41 is. Parameters:
+  $\alpha=3$, $\beta=20$, $k=3$, $D=0.2$, $\gamma\in\pm 2$,
+  $\delta{t}=0.0001$. \textbf{E.}  The Honda Dirichlet measure, $\delta$, as a
+  function of simulation time for the simulations show in D. \textbf{F.}
+  Guidance molecules in the simulated Fgf8 misexpression experiment. Molecule
+  1 is mirrored over the anterior-posterior axis The simulation indicates that
+  the mirroring of the barrel field can occur as a result of the mirroring of
+  guidance molecule 1.}
 \label{fig:main}
@@
 
-\subsection*{The barrel field reacts to changes in the expression pattern of guidance molecules}
+First we verified that all results established by \cite{karbowski_model_2004} for a 1D axis could be reproduced using our extension to a 2D cortical sheet.
+Using an elliptical boundary with $M=3$ offset guidance gradients aligned to the longer axis, $N=5$ thalamocortical projections gave rise to five distinct cortical fields at locations that preserved the topographic ordering defined by the original $\gamma$ values.
+However, we found that specifying $N$ ordered areas required $M\approx N$ signalling fields.
+As the number of distinct guidance fields is unlikely to approach the number of barrels, further extensions were required.
 
-To cover the reaction to changes in the guidance molecules caused by
-experiments like those by Shimogori \& Grove.
+The term in parentheses in Eq.~\ref{eq:dc} represents competition between thalamocortical projections for a limited availability of cortical connections.
+To introduce competition also in terms of axon branching we redefined
+%
+\begin{equation} \label{eq:comp}
+\chi_i(\mb{x}, t) = - \frac{\epsilon  a_i}{N-1} \sum_{j \ne i}^{N} a_{j},
+\end{equation}
+%
+which reduces branching for each projection where the branches of other projections are dense.
+Divisive normalization keeps the axon branch density bounded on each iteration to the initial density  at $t=0$;
+
+\begin{equation} \label{eq:norm}
+  a_i(\mathbf{x}, t) = \int_\Omega  a_i(\mathbf{x}, 0) d\sigma \; \frac {a'_i(\mathbf{x}, t)} {\int_\Omega
+  a'_i(\mathbf{x}, t) d\sigma},
+\end{equation}
+%
+where prime symbols indicate the use of intermediate values computed from Eqs.~\ref{eq:dc}--\ref{eq:comp} at time $t-{\delta}t$.
+
+The only differences between thalamic projections are their strengths of interaction with the guidance fields, $\gamma$, hence any reliable differences between emergent cortical fields must be due to differences in these values only.
+We speculate that a projection interacts with a given ephrin field in the cortical subplate with a strength determined by the concentration of a similar molecule at its thalamic origin, i.e., the putative barreloid center.
+As such, two orthogonal linear thalamic gradients were defined, from which 41 pairs of $\gamma$ values were sampled, at the coordinates of 41 barreloid centers recreated from \cite{haidarliu_size_2001}.
+
+A cortical boundary enclosing 41 corresponding barrels was traced from a cytochrome oxidase stain from \cite{zheng_signal_2001}, and Eqs.~\ref{eq:dc}--\ref{eq:norm} were solved for $N=41$ projections on the resulting domain, using $M=2$ linear signalling gradients aligned with the anterior-posterior and  medial-lateral axes.
+From random initial conditions for $a({\bf x},0)$ and $c({\bf x},0)$, a clear Voronoi-like tesselation emerged.
+A reduction in the Honda delta metric (see \cite{senft_mouse_1991}) confirmed that a `good' Voronoi pattern emerged within $\approx 3000$ iterations (see Figure 1).
+The sum of squared differences between the real and simulated barrel centres revealed a strong correspondence.
+
+To further investigate the interplay of genetic and intrinsic mechanisms we simulated a seminal barrel duplication experiment \cite{shimogori_fibroblast_2005},  first by solving on a cortical domain comprising two separate, identically shaped cortical boundaries.
+In this case one distinct cortical field per thalamic projection emerged, within one of the two boundaries only, i.e., no duplication.
+However, merging the two fields to create an extended mirror-symmetric boundary shape, with an anterior-posterior guidance field reflected at the join, gave rise to two mirror-symmetrical barrel fields comprising NN barrels, i.e., two cortical fields for each thalamic projection (see Figure 2).
+Together these results suggest that by the misexpression of Fgf8, Shimogori \& Grove \citep{shimogori_fibroblast_2005} effectively created one large barrel field rather than two distinct fields.
 
@@
 
-We set out to test the hypothesis that barrel patterning may emerge via
-self-organising processes. We first extended an existing model of neocortical
-arealization to simulate pattern formation in two dimensions, inheriting from
-this model the assumption that axonal branching and synaptogenesis are
-coupled, and that thalamic projections compete for cortical territory under
-the influence of molecular signalling gradients. Realistic barrel patterning
-emerged only when terms were added to represent the assumption that
-thalamocortical projections compete for a limited supply of connections with
-cortical cells, that their branching rates are reduced by the branching of
-other thalamocortical projections, and that the density of axon branching is
-conserved for each projection. The net effect of these local interactions is a
-Voronoi tesselation in which localised clusters of thalamic projections become
-somatotopically organised on the cortical sheet according to the relative
-values of coefficients specifying the rates at which axons are attracted and
-repelled by orthogonal pairs of molecular signal gradients. According to this
-model, the pattern of barrels is determined by the competitive and cooperative
-interactions between cells relative to the shape of the cortical field
-boundary. In simulation, we systematically investigated the origin of the
-asymmetry between the barrel rows and arcs, and identified {\color{red} XXXXX}
-as the most likely candidate for the elongation of barrel shapes along the
-rows.
+The present results suggest that the key requirements for the emergence of realistic barrel patterning are i) at each cortical location thalamocortical projections compete for a limited number of available synaptic connections (Eqs.~\ref{eq:dc}--\ref{eq:da}), ii) at each location the branching rate of a given projection is reduced by the density of other projections (Eq.~\ref{eq:comp}), and iii) the branch density of each projection is conserved (Eq.~\ref{eq:norm}).
+
+It is important to emphasize that the formulation of the model is entirely local, insofar as simulation requires no information to be communicated between thalamic projections, or from a given cortical location to any but those immediately adjacent.
+Hence the simulations demonstrate how a self-organising system, constrained by genetically specified guidance cues, can faithfully reproduce an arrangement of cell aggregates in one neural structure as a topographic map in another.
+
+The model also shows that the shape of the barrels may be primarily determined by the competitive and cooperative interactions between cells relative to the shape of the cortical field boundary.
+And that somatotopic map formation does not require the pre-specification of centre-points by as yet undetermined additional developmental mechanisms.
 

### 2019-12-16 analysis/plot_honda_quick.py

@@ -26,7 +26,7 @@ matplotlib.rc('font', **fnt)
 # And plot this longhand here:
 F1 = plt.figure (figsize=(9,8))
 
-t1_masked = np.load ('postproc/honda_t.npy')
+t1_masked = np.load ('postproc/honda_t.npy')/10000
 hondadelta = np.load ('postproc/honda_delta.npy')
 area_measure =  np.load ('postproc/area_measure.npy')
 
@@ -34,23 +34,23 @@ xmax = max(t1_masked)
 xmax = xmax[0]
 print ('xmax = {0}'.format(xmax))
 ax1 = F1.add_subplot(1,1,1)
-l1, = ax1.plot(t1_masked, hondadelta, 'o', markersize=12, color=col.black, label='Honda $\delta$')
+l1, = ax1.plot(t1_masked, hondadelta, 'o', markersize=12, color=col.red, label='Honda $\delta$')
 
-l2, = ax1.plot((0,xmax), (0.055, 0.055), '--', color=col.black, linewidth=3, label="good (S&W)")
+l2, = ax1.plot((0,xmax), (0.055, 0.055), '--', color=col.red, linewidth=3, label="good (S&W)")
 #l3, = ax1.plot((0,xmax), (0.15, 0.15), '-.', color=col.black, linewidth=3, label="awful (non Dirichlet)")
 
 ax2 = ax1.twinx()
 
-l4, = ax2.plot(t1_masked, area_measure, 's', markersize=12, color=col.blue)
+l4, = ax2.plot(t1_masked, area_measure, 's', markersize=12, color=col.black)
 
-ax1.set_xlabel ('Steps')
-ax1.set_ylabel ('$\delta$')
-ax2.set_ylabel ('$\zeta$')
-ax2.tick_params (axis='y', labelcolor=col.blue)
+ax1.set_xlabel ('10k steps')
+ax1.set_ylabel ('$\delta$',rotation=0)
+ax2.set_ylabel ('$\sigma$',rotation=0)
+ax2.tick_params (axis='y', labelcolor=col.black)
 ax1.set_xlim ((0,xmax))
 ax1.set_ylim ((0,0.3))
 ax2.set_ylim ((0,500))
-ax1.set_xticks ((0,10000,20000))
+ax1.set_xticks ((0,1,2))
 ax1.set_yticks ((0,0.1,0.2,0.3))
 ax2.set_yticks ((0,200,400))
 

### 2019-12-17 paper/briefpaper/brief.tex
Updated figure (corrected black boundaries between regions and unified font sizes)

@@
 
-\caption{\textbf{C.} Cortical whisker barrels from \cite{zheng_signal_2001}
-  with annotations for approximate axes and lines drawn to join the
-  interstices into a putative Voronoi diagram.  \textbf{B.}  Top: Stochastic
-  initial condition for $a_i$ with $i=13$ (barrel C4). The roll-off towards
-  the boundary (thin line) ensures the boundary conditions are
-  satisfied. Bottom: Guidance molecules in the simulated cortex. Molecule 1 is
-  expressed high to low in an approximately anterior to posterior
-  gradient. Molecule 2 is expressed lateral to medial (analogously to ephrin -
-  Vanderhaegen et al) \textbf{C.}  Barreloid centres for mystacial whiskers
-  taken from \cite{haidarliu_size_2001} are shown as black circles. The dashed
-  grey line shows the boundary of the barreloid field. It is assumed that
-  there are two linearly increasing, orthogonal signalling molecules in the
-  thalamus, such that the location of each barreloid within the graded
-  thalamic signal field can differentiate the efferent axons, with the result
-  that axons from each barreloid interact to varying degrees with the cortical
-  signalling gradients shown in B. Coloured arrows show, for each barreloid,
-  the magnitude (and sign) of the interaction that its thalamocortical
-  projection will have with each signalling gradient in the cortex. The
-  mappings from barreloid location to interaction strength are
-  linear. \textbf{D.}  Barrel cortex simulation. Colours indicate
-  $\arg\max_i(c_i)$ for 41 is. Parameters:
-  $\alpha=3$, $\beta=20$, $k=3$, $D=0.2$, $\gamma\in\pm 2$,
-  $\delta{t}=0.0001$. \textbf{E.}  The Honda Dirichlet measure, $\delta$, as a
-  function of simulation time for the simulations show in D. \textbf{F.}
-  Guidance molecules in the simulated Fgf8 misexpression experiment. Molecule
-  1 is mirrored over the anterior-posterior axis The simulation indicates that
-  the mirroring of the barrel field can occur as a result of the mirroring of
-  guidance molecule 1.}
+\caption{\textbf{A.} Cortical whisker barrels from \cite{zheng_signal_2001}
+  annotated with lines joining interstices into a putative Voronoi diagram
+  ($\delta=0.022$). \textbf{B.}  Top: Stochastic initial condition for $a_i$
+  with $i=1$ (barrel `$\alpha$'). The roll-off towards the boundary (thin line)
+  ensures the boundary conditions are satisfied. Bottom: Guidance molecules
+  $\rho_1(\mathbf{x})$ and $\rho_2(\mathbf{x})$ in the simulated
+  cortex. $\rho_1(\mathbf{x})$ varies along the anterior-posterior axis;
+  $\rho_2(\mathbf{x})$ is expressed along the lateral/medial axis, analogous
+  to ephrin A5 \citep{vanderhaeghen_mapping_2000}. \textbf{C.}  Barreloid
+  centres for mystacial whiskers taken from \cite{haidarliu_size_2001} are
+  shown as black circles. The dashed grey line shows the boundary of the
+  barreloid field. It is assumed that there are two linearly increasing,
+  orthogonal signalling molecules in the thalamus, such that the location of
+  each barreloid within the graded thalamic signal field can differentiate the
+  efferent axons, with the result that axons from each barreloid interact to
+  varying degrees with the cortical signalling gradients $\rho_1$ and
+  $\rho_2$. Blue (green) arrows show, for each barreloid, the magnitude (and
+  sign) of the interaction that its thalamocortical projection will have with
+  $\rho_1$ ($\rho_2$). The mappings from barreloid location to interaction
+  strength are linear. \textbf{D.}  Barrel cortex simulation with parameters
+  $\alpha=3$, $\beta=20$, $k=3$, $D=0.2$, $\gamma\in\pm 2$ and
+  $\delta{t}=0.0001$. Colours indicate $\arg\max_i(c_i)$ for 41 TC
+  projections.  \textbf{E.}  The Honda Voronoi measure, $\delta$, (red) and a
+  pattern fitting metric based on the distance, $d_i$, between simulated and
+  experimental barrel centroids and the absolute difference in their areas,
+  $\Delta A_i$ given by $\sigma = \sum_i^N d_i^2 \times \sum_i^N \Delta A_i$
+  (black) plotted as a function of time for the simulations shown in
+  D. \textbf{F.}  Guidance molecules in the simulated Fgf8 misexpression
+  experiment. $\rho_1$ is mirrored over the anterior-posterior axis. The
+  simulation indicates that the mirroring of the barrel field can occur as a
+  result of the mirroring of guidance molecule 1.}
 \label{fig:main}
@@
 Using an elliptical boundary with $M=3$ offset guidance gradients aligned to the longer axis, $N=5$ thalamocortical projections gave rise to five distinct cortical fields at locations that preserved the topographic ordering defined by the original $\gamma$ values.
-However, we found that specifying $N$ ordered areas required $M\approx N$ signalling fields.
+However, we found that specifying $N$ ordered areas required $M\approx (N+1)/2$ signalling fields.
 As the number of distinct guidance fields is unlikely to approach the number of barrels, further extensions were required.

### 2019-12-17 paper/briefpaper/brief.tex
Change sigma to eta, because we have dsigma in Eq 4

@@
 In the plane tangential to the cortical surface the barrels constitute a somatotopic map of the whiskers, with cells of adjacent barrels responding most strongly and quickly to deflection of adjacent whiskers \citep{armstrong-james_flow_1992}.
-Barrel patterning reflects subcortical whisker maps comprising cell aggregates called barrelettes in the brainstem and barreloids in the thalamus \cite{van_der_loos_barreloids_1976}.
+Barrel patterning reflects subcortical whisker maps comprising cell aggregates called barrelettes in the brainstem and barreloids in the thalamus.
+%\cite{van_der_loos_barreloids_1976}.
 
 Barrel formation requires afferent input from whisker stimulation and thalamic calcium waves \citep{anton-bolanos_prenatal_2019}, and depends on a complex network of axon guidance molecules such as ephrin-A5 and A7 and adhesion molecules such as cadherin-6 and 8 \citep{miller_epha7-ephrin-a5_2006}.
-This network is orchestrated by interactions between morphogens Fgf8 and Fgf17 and transcription factors Emx2, Pax6, Sp8, and Coup-tf1 \citep{shimogori_fibroblast_2005,bishop_regulation_2000}, which are expressed in gradients that mark orthogonal axes and can be manipulated to stretch, shrink, shift, and even duplicate barrels \cite{assimacopoulos_fibroblast_2012}.
+This network is orchestrated by interactions between morphogens Fgf8 and Fgf17 and transcription factors Emx2, Pax6, Sp8, and Coup-tf1 \citep{shimogori_fibroblast_2005}, which are expressed in gradients that mark orthogonal axes and can be manipulated to stretch, shrink, shift, and even duplicate barrels \cite{assimacopoulos_fibroblast_2012}.
+% Removed ref: ,bishop_regulation_2000 after ..Coup-tf1
 
@@
   ($\delta=0.022$). \textbf{B.}  Top: Stochastic initial condition for $a_i$
-  with $i=1$ (barrel `$\alpha$'). The roll-off towards the boundary (thin line)
-  ensures the boundary conditions are satisfied. Bottom: Guidance molecules
-  $\rho_1(\mathbf{x})$ and $\rho_2(\mathbf{x})$ in the simulated
+  with $i=1$ (barrel `$\alpha$'). The roll-off towards the boundary (thin
+  line) ensures the boundary conditions are satisfied. Bottom: Guidance
+  molecules $\rho_1(\mathbf{x})$ and $\rho_2(\mathbf{x})$ in the simulated
   cortex. $\rho_1(\mathbf{x})$ varies along the anterior-posterior axis;
@@
   experimental barrel centroids and the absolute difference in their areas,
-  $\Delta A_i$ given by $\sigma = \sum_i^N d_i^2 \times \sum_i^N \Delta A_i$
-  (black) plotted as a function of time for the simulations shown in
-  D. \textbf{F.}  Guidance molecules in the simulated Fgf8 misexpression
+  $\Delta A_i$ given by $\eta = \sum_i^N d_i^2 \times \sum_i^N \Delta A_i$
+  (black, units L$^4$) plotted as a function of time for the simulations shown
+  in D. \textbf{F.}  Guidance molecules in the simulated Fgf8 misexpression
   experiment. $\rho_1$ is mirrored over the anterior-posterior axis. The

### 2019-12-17 analysis/plot_honda_quick.py

@@ -45,7 +45,7 @@ l4, = ax2.plot(t1_masked, area_measure, 's', markersize=12, color=col.black)
 
 ax1.set_xlabel ('10k steps')
 ax1.set_ylabel ('$\delta$',rotation=0)
-ax2.set_ylabel ('$\sigma$',rotation=0)
+ax2.set_ylabel ('$\eta$',rotation=0)
 ax2.tick_params (axis='y', labelcolor=col.black)
 ax1.set_xlim ((0,xmax))
 ax1.set_ylim ((0,0.3))

### 2019-12-17 paper/briefpaper/brief.tex
Adds citation in intro

@@
 
-Barrel formation requires afferent input from whisker stimulation and thalamic calcium waves \citep{anton-bolanos_prenatal_2019}, and depends on a complex network of axon guidance molecules such as ephrin-A5 and A7 and adhesion molecules such as cadherin-6 and 8 \citep{miller_epha7-ephrin-a5_2006}.
+Barrel formation requires afferent input from whisker stimulation and thalamic calcium waves \citep{anton-bolanos_prenatal_2019}, and depends on a complex network of axon guidance molecules such as ephrin-A5 and A7 and adhesion molecules such as cadherin-6 and 8 \citep{vanderhaeghen_mapping_2000,miller_epha7-ephrin-a5_2006}.
 This network is orchestrated by interactions between morphogens Fgf8 and Fgf17 and transcription factors Emx2, Pax6, Sp8, and Coup-tf1 \citep{shimogori_fibroblast_2005}, which are expressed in gradients that mark orthogonal axes and can be manipulated to stretch, shrink, shift, and even duplicate barrels \cite{assimacopoulos_fibroblast_2012}.

### 2019-12-17 paper/briefpaper/brief.tex
Fixes a mis-placed white background box in figure, also tweaks to caption

@@
 
-Software tools for modelling the cortical sheet as a hexagonal lattice with no-flux boundary conditions and an arbitrary boundary shape, for computing the flux using a finite difference method, for numerical integration by the fourth-order Runge-Kutta method, and for reproducing the reported results are available at \url{https://github.com/ABRG-Models/BarrelEmerge}.
+Software for modelling the cortical sheet as a hexagonal lattice with no-flux
+boundary conditions and an arbitrary boundary shape, for computing the flux
+using a finite difference method, for numerical integration by the
+fourth-order Runge-Kutta method, and for reproducing these results is available at \url{https://github.com/ABRG-Models/BarrelEmerge}.
 
@@
   annotated with lines joining interstices into a putative Voronoi diagram
-  ($\delta=0.022$). \textbf{B.}  Top: Stochastic initial condition for $a_i$
-  with $i=1$ (barrel `$\alpha$'). The roll-off towards the boundary (thin
-  line) ensures the boundary conditions are satisfied. Bottom: Guidance
-  molecules $\rho_1(\mathbf{x})$ and $\rho_2(\mathbf{x})$ in the simulated
-  cortex. $\rho_1(\mathbf{x})$ varies along the anterior-posterior axis;
-  $\rho_2(\mathbf{x})$ is expressed along the lateral/medial axis, analogous
-  to ephrin A5 \citep{vanderhaeghen_mapping_2000}. \textbf{C.}  Barreloid
-  centres for mystacial whiskers taken from \cite{haidarliu_size_2001} are
-  shown as black circles. The dashed grey line shows the boundary of the
-  barreloid field. It is assumed that there are two linearly increasing,
-  orthogonal signalling molecules in the thalamus, such that the location of
-  each barreloid within the graded thalamic signal field can differentiate the
-  efferent axons, with the result that axons from each barreloid interact to
-  varying degrees with the cortical signalling gradients $\rho_1$ and
-  $\rho_2$. Blue (green) arrows show, for each barreloid, the magnitude (and
-  sign) of the interaction that its thalamocortical projection will have with
-  $\rho_1$ ($\rho_2$). The mappings from barreloid location to interaction
-  strength are linear. \textbf{D.}  Barrel cortex simulation with parameters
-  $\alpha=3$, $\beta=20$, $k=3$, $D=0.2$, $\gamma\in\pm 2$ and
+  ($\delta=0.022$). \textbf{B.}  Top: Stochastic initial condition for $a_1$
+  (barrel `$\alpha$'). The roll-off towards the boundary (thin line) ensures
+  the boundary conditions are satisfied. Bottom: Guidance molecules
+  $\rho_1(\mathbf{x})$ and $\rho_2(\mathbf{x})$ in the simulated
+  cortex. $\rho_1(\mathbf{x})$ varies along the anterior-posterior axis,
+  $\rho_2(\mathbf{x})$ along the lateral/medial axis (cf.~ephrin A5,
+  \cite{vanderhaeghen_mapping_2000}). \textbf{C.}  Barreloid centres for
+  mystacial whiskers taken from \cite{haidarliu_size_2001} are shown as black
+  circles; the dashed grey line indicates the boundary. Assuming two linearly
+  increasing, orthogonal signalling molecules in the thalamus, the location of
+  each barreloid within this signal field can differentiate the efferent
+  axons. Axons from each barreloid interact to varying degrees with the
+  cortical signals $\rho_1$ and $\rho_2$. Blue (green) arrows show, for each
+  barreloid's projection, the magnitude (and sign) of the interaction that it
+  will have with $\rho_1$ ($\rho_2$). \textbf{D.}  Barrel cortex simulation
+  with parameters $\alpha=3$, $\beta=20$, $k=3$, $D=0.2$, $\gamma\in\pm 2$ and
   $\delta{t}=0.0001$. Colours indicate $\arg\max_i(c_i)$ for 41 TC
-  projections.  \textbf{E.}  The Honda Voronoi measure, $\delta$, (red) and a
-  pattern fitting metric based on the distance, $d_i$, between simulated and
-  experimental barrel centroids and the absolute difference in their areas,
-  $\Delta A_i$ given by $\eta = \sum_i^N d_i^2 \times \sum_i^N \Delta A_i$
-  (black, units L$^4$) plotted as a function of time for the simulations shown
-  in D. \textbf{F.}  Guidance molecules in the simulated Fgf8 misexpression
-  experiment. $\rho_1$ is mirrored over the anterior-posterior axis. The
-  simulation indicates that the mirroring of the barrel field can occur as a
-  result of the mirroring of guidance molecule 1.}
+  projections; black lines indicate boundaries. \textbf{E.}
+  The Honda Voronoi measure, $\delta$, (red) and a pattern fitness metric
+  based on the distance, $d_i$, between simulated and observed barrel
+  centroids and the absolute difference in their areas, $\Delta A_i$ given by
+  $\eta = \sum_i^N d_i^2 \times \sum_i^N \Delta A_i$ (black, units mm$^4$)
+  plotted as a function of time for the simulations shown in D. \textbf{F.}
+  Guidance molecules and resulting pattern in the simulated Fgf8 misexpression
+  experiment. Parameters as in D, but $\rho_1$ is mirrored over the
+  anterior-posterior axis. The simulation indicates that the mirroring of the
+  barrel field can occur as a result of the mirroring of guidance molecule 1.
+  All scale bars are 1~mm.}
 \label{fig:main}
@@
 Divisive normalization keeps the axon branch density bounded on each iteration to the initial density  at $t=0$;
-
+%
 \begin{equation} \label{eq:norm}

### 2019-12-17 paper/briefpaper/brief.tex
Tweaks to final para

@@
 Hence the simulations demonstrate how a self-organising system, constrained by genetically specified guidance cues, can faithfully reproduce an arrangement of cell aggregates in one neural structure as a topographic map in another.
-
-The model also shows that the shape of the barrels may be primarily determined by the competitive and cooperative interactions between cells relative to the shape of the cortical field boundary.
-And that somatotopic map formation does not require the pre-specification of centre-points by as yet undetermined additional developmental mechanisms.
+%
+The model also shows that the shape of the barrels may be primarily determined
+by the competitive and cooperative interactions between cells relative to the
+shape of the cortical field boundary and that somatotopic map formation does
+not require the pre-specification of centre-points by as yet undetermined
+additional developmental mechanisms.
 

### 2019-12-19 paper/briefpaper/brief.tex
Adds Leah to author list
commiting version sent to Leah

--- 2019-12-17 paper/briefpaper/brief.tex
@@
 \author[a,1]{Sebastian~S.~James}
+\author[b]{Leah A. Krubitzer}
 \author[a]{Stuart~P.~Wilson}
@@
 \affil[a]{Department of Psychology, The University of Sheffield, Sheffield, United Kingdom.}
-
+\affil[b]{Center for Neuroscience, The University of California, Davis, United States.}
 % $^{*}$Corresponding author email: Seb.James@Sheffield.ac.uk\vspace{1em}\\
@@
 
-The barrel boundaries form a Voronoi tessellation across the cortical sheet \citep{senft_mouse_1991}, suggesting that barreloid topology is preserved in the projection of thalamocortical axons into the cortex, and that a barrel forms by lateral axon branching from an initial centre-point that ceases upon contact with axons branching from adjacent centres.
+The barrel boundaries form a Voronoi tessellation across the cortical sheet \citep{senft_mouse_1991} (Fig.\,1A), suggesting that barreloid topology is preserved in the projection of thalamocortical axons into the cortex, and that a barrel forms by lateral axon branching from an initial centre-point that ceases upon contact with axons branching from adjacent centres.
 But the assumption of pre-arranged centres is difficult to resolve with the observation that  axons arrive in the cortical plate as an undifferentiated bundle, prior to barreloid formation \cite{agmon_organized_1993}.
@@
 \begin{center}
-\includegraphics[width=\textwidth]{../briefpaper/MainFig.png}
+\includegraphics[width=0.8\textwidth]{../briefpaper/MainFig.pdf}
 \end{center}
-
-\caption{\textbf{A.} Cortical whisker barrels from \cite{zheng_signal_2001}
-  annotated with lines joining interstices into a putative Voronoi diagram
-  ($\delta=0.022$). \textbf{B.}  Top: Stochastic initial condition for $a_1$
-  (barrel `$\alpha$'). The roll-off towards the boundary (thin line) ensures
-  the boundary conditions are satisfied. Bottom: Guidance molecules
-  $\rho_1(\mathbf{x})$ and $\rho_2(\mathbf{x})$ in the simulated
-  cortex. $\rho_1(\mathbf{x})$ varies along the anterior-posterior axis,
-  $\rho_2(\mathbf{x})$ along the lateral/medial axis (cf.~ephrin A5,
-  \cite{vanderhaeghen_mapping_2000}). \textbf{C.}  Barreloid centres for
-  mystacial whiskers taken from \cite{haidarliu_size_2001} are shown as black
-  circles; the dashed grey line indicates the boundary. Assuming two linearly
-  increasing, orthogonal signalling molecules in the thalamus, the location of
-  each barreloid within this signal field can differentiate the efferent
-  axons. Axons from each barreloid interact to varying degrees with the
-  cortical signals $\rho_1$ and $\rho_2$. Blue (green) arrows show, for each
-  barreloid's projection, the magnitude (and sign) of the interaction that it
-  will have with $\rho_1$ ($\rho_2$). \textbf{D.}  Barrel cortex simulation
-  with parameters $\alpha=3$, $\beta=20$, $k=3$, $D=0.2$, $\gamma\in\pm 2$ and
-  $\delta{t}=0.0001$. Colours indicate $\arg\max_i(c_i)$ for 41 TC
-  projections; black lines indicate boundaries. \textbf{E.}
-  The Honda Voronoi measure, $\delta$, (red) and a pattern fitness metric
-  based on the distance, $d_i$, between simulated and observed barrel
-  centroids and the absolute difference in their areas, $\Delta A_i$ given by
-  $\eta = \sum_i^N d_i^2 \times \sum_i^N \Delta A_i$ (black, units mm$^4$)
-  plotted as a function of time for the simulations shown in D. \textbf{F.}
-  Guidance molecules and resulting pattern in the simulated Fgf8 misexpression
-  experiment. Parameters as in D, but $\rho_1$ is mirrored over the
-  anterior-posterior axis. The simulation indicates that the mirroring of the
-  barrel field can occur as a result of the mirroring of guidance molecule 1.
-  All scale bars are 1~mm.}
+\caption{\textbf{A} Left shows a cytochrome oxidase stain obtained from rat S1 by \cite{zheng_signal_2001}, with black lines to delineate barrels and to measure departure ($\delta$) from a perfect Voronoi tesselation.
+Right shows the initial distribution of axon branching density ($a$) for one thalamocortical projection and two molecular guidance fields ($\rho$).
+\textbf{B} The strength of interaction $\gamma$ for each of 41 projections with field $\rho_1$ and $\rho_2$ are indicated by the lengths of green and blue arrows respectively, assuming that similar fields aligned to the posterior-anterior and medial-lateral axes in the ventroposterior medial nucleus of the thalamus are sampled at the locations of barreloid centers (reconstructed from \cite{haidarliu_size_2001}).
+\textbf{C} Simulation results for parameters $N=41$, $\alpha=3$, $\beta=20$, $k=3$, $D=0.2$, $\gamma\in\pm 2$ and $\delta{t}=0.0001$.
+Colours indicate the thalamic projection for which the connection density is maximal, black lines delineate boundaries, and overlaid contours show $c>0.2$.
+\textbf{D} The Honda $\delta$ metric (red dots; see \cite{senft_mouse_1991}) measures departure from a Voronoi tesselation (dotted line corresponds to the barrels in $A$), and $\eta$ (black squares) measures the correspondence between the real and simulated barrel shapes (the product of the sum of squared differences between real and simulated centers and the sum of squared differences in area; units mm$^4$).
+\textbf{F.} Guidance fields and emergent barrel pattern in a Fgf8 misexpression experiment (c.f.
+\cite{assimacopoulos_fibroblast_2012}), simulated by reflecting $\rho_1$ about the anterior-posterior axis.
+All scale bars 1mm.}
 \label{fig:main}
@@
 We speculate that a projection interacts with a given ephrin field in the cortical subplate with a strength determined by the concentration of a similar molecule at its thalamic origin, i.e., the putative barreloid center.
-As such, two orthogonal linear thalamic gradients were defined, from which 41 pairs of $\gamma$ values were sampled, at the coordinates of 41 barreloid centers recreated from \cite{haidarliu_size_2001}.
+As such, two orthogonal linear thalamic gradients were defined, from which 41 pairs of $\gamma$ values were sampled, at the coordinates of 41 barreloid centers recreated from \cite{haidarliu_size_2001} (Fig.\,1B).
 
 A cortical boundary enclosing 41 corresponding barrels was traced from a cytochrome oxidase stain from \cite{zheng_signal_2001}, and Eqs.~\ref{eq:dc}--\ref{eq:norm} were solved for $N=41$ projections on the resulting domain, using $M=2$ linear signalling gradients aligned with the anterior-posterior and medial-lateral axes.
-From random initial conditions for $a({\bf x},0)$ and $c({\bf x},0)$, a clear Voronoi-like tesselation emerged.
-A reduction in the Honda delta metric (see \cite{senft_mouse_1991}) confirmed that a `good' Voronoi pattern emerged within $\approx 3000$ iterations (see Figure 1).
-The sum of squared differences between the real and simulated barrel centres revealed a strong correspondence.
+From random initial conditions for $a({\bf x},0)$ and $c({\bf x},0)$, a clear Voronoi-like tesselation emerged (Fig.\,1C).
+A reduction in the Honda delta metric (see \cite{senft_mouse_1991}) confirmed that a `good' Voronoi pattern emerged within $\approx 3000$ iterations (see Fig.\,1D).
+A measure of the difference in shape between real and simulated barrels revealed a strong correspondence (see Fig.\,1D).
 
@@
 In this case one distinct cortical field per thalamic projection emerged, within one of the two boundaries only, i.e., no duplication.
-However, merging the two fields to create an extended mirror-symmetric boundary shape, with an anterior-posterior guidance field reflected at the join, gave rise to two mirror-symmetrical barrel fields comprising NN barrels, i.e., two cortical fields for each thalamic projection (see Figure 2).
-Together these results suggest that by the misexpression of Fgf8, Shimogori \& Grove \citep{shimogori_fibroblast_2005} effectively created one large barrel field rather than two distinct fields.
+However, merging the two fields to create an extended mirror-symmetric boundary shape, with an anterior-posterior guidance field reflected at the join, gave rise to two mirror-symmetrical barrel fields comprising 2N barrels, i.e., two cortical fields for each thalamic projection (Fig.\,1E).
+Together these results suggest that by the misexpression of Fgf8, Shimogori \& Grove \citep{shimogori_fibroblast_2005} effectively created one large barrel field rather than two distinct cortical fields.
 
@@
 
-It is important to emphasize that the formulation of the model is entirely local, insofar as simulation requires no information to be communicated between thalamic projections, or from a given cortical location to any but those immediately adjacent.
+It is important to emphasize that the formulation of the model is entirely local, insofar as simulation requires no information to be communicated from a given cortical grid cell to any but those immediately adjacent (via diffusion).
 Hence the simulations demonstrate how a self-organising system, constrained by genetically specified guidance cues, can faithfully reproduce an arrangement of cell aggregates in one neural structure as a topographic map in another.
-
-The model also shows that the shape of the barrels may be primarily determined by the competitive and cooperative interactions between cells relative to the shape of the cortical field boundary.
-And that somatotopic map formation does not require the pre-specification of centre-points by as yet undetermined additional developmental mechanisms.
+%
+The model also shows that the shape of the barrels may be primarily determined
+by the competitive and cooperative interactions between cells relative to the
+shape of the cortical field boundary. And that somatotopic map formation does
+not require the pre-specification of centre-points by as yet undetermined
+additional developmental mechanisms.
 

### 2019-12-19 paper/briefpaper/brief.tex
Corrects white background, creates png and only changes brief.tex to use the phng
Fixes small error

@@
 \begin{center}
-\includegraphics[width=0.8\textwidth]{../briefpaper/MainFig.pdf}
+\includegraphics[width=0.8\textwidth]{./MainFig.png}
 \end{center}
@@
 Colours indicate the thalamic projection for which the connection density is maximal, black lines delineate boundaries, and overlaid contours show $c>0.2$.
-\textbf{D} The Honda $\delta$ metric (red dots; see \cite{senft_mouse_1991}) measures departure from a Voronoi tesselation (dotted line corresponds to the barrels in $A$), and $\eta$ (black squares) measures the correspondence between the real and simulated barrel shapes (the product of the sum of squared differences between real and simulated centers and the sum of squared differences in area; units mm$^4$).
+\textbf{D} The Honda $\delta$ metric (red dots; see \cite{senft_mouse_1991}) measures departure from a Voronoi tesselation (dotted line corresponds to the barrels in $A$), and $\eta$ (black squares) measures the correspondence between the real and simulated barrel shapes (the product of the sum of squared differences between real and simulated centers and the sum of differences in area; units mm$^4$).
 \textbf{F.} Guidance fields and emergent barrel pattern in a Fgf8 misexpression experiment (c.f.

### 2019-12-19 paper/briefpaper/brief.tex
E, not F
Contours show c>0.5, not >0.2

@@
 \textbf{C} Simulation results for parameters $N=41$, $\alpha=3$, $\beta=20$, $k=3$, $D=0.2$, $\gamma\in\pm 2$ and $\delta{t}=0.0001$.
-Colours indicate the thalamic projection for which the connection density is maximal, black lines delineate boundaries, and overlaid contours show $c>0.2$.
+Colours indicate the thalamic projection for which the connection density is maximal, black lines delineate boundaries, and overlaid contours show $c>0.5$.
 \textbf{D} The Honda $\delta$ metric (red dots; see \cite{senft_mouse_1991}) measures departure from a Voronoi tesselation (dotted line corresponds to the barrels in $A$), and $\eta$ (black squares) measures the correspondence between the real and simulated barrel shapes (the product of the sum of squared differences between real and simulated centers and the sum of differences in area; units mm$^4$).
-\textbf{F.} Guidance fields and emergent barrel pattern in a Fgf8 misexpression experiment (c.f.
+\textbf{E.} Guidance fields and emergent barrel pattern in a Fgf8 misexpression experiment (c.f.
 \cite{assimacopoulos_fibroblast_2012}), simulated by reflecting $\rho_1$ about the anterior-posterior axis.

### 2019-12-19 paper/briefpaper/brief.tex
sfs

--- 2019-12-19 paper/briefpaper/brief.tex
@@
 \begin{abstract}
-Brain development relies on an interplay of genetic specification and self-organisation.
-Striking examples of this interplay can be found in the rodent somatosensory brainstem, thalamus, and cortex, where the arrangement of the facial whiskers is preserved in the arrangement of cell aggregates to form precise somatotopic maps.
+Brain development relies on an interplay between genetic specification and self-organisation.
+Striking examples of this relationship can be found in the  somatosensory brainstem, thalamus, and cortex of rats and mice, where the arrangement of the facial whiskers is preserved in the arrangement of cell aggregates to form precise somatotopic maps.
 We show in simulation how realistic whisker maps can self-organize, by assuming that information is exchanged between adjacent cells only, under the guidance of gene expression gradients.
-The resulting model provides a simple account of how genetic information can constrain spontaneous pattern formation to faithfully reproduce functional maps in subsequent brain structures.
+The resulting model provides a simple account of how patterns of gene expression can constrain spontaneous pattern formation to faithfully reproduce functional maps in subsequent brain structures.
 \end{abstract}
@@
 \dropcap{S}patial patterns in neural connectivity provide clues about the constraints under which brains evolve and develop \citep{purves_iterated_1992}.
-Perhaps the most distinctive pattern can be found in the rodent barrel cortex \cite{woolsey_structural_1970}.
-The barrels are large cortical columns, identifiable from birth in layer 4 of primary somatosensory cortex as dense clusters of thalamocortical axons, which are enclosed by borders a few neurons thick from postnatal day 3 \citep{erzurumlu_development_2012}.
-In the plane tangential to the cortical surface the barrels constitute a somatotopic map of the whiskers, with cells of adjacent barrels responding most strongly and quickly to deflection of adjacent whiskers \citep{armstrong-james_flow_1992}.
+Perhaps the most distinctive pattern can be found in the barrel cortex of many rodent species \cite{woolsey_structural_1970}.
+The barrels are identifiable soon after birth in layer 4 of primary somatosensory cortex as dense clusters of thalamocortical axons, which are enclosed by borders a few neurons thick from postnatal day 3 \citep{erzurumlu_development_2012}.
+In the plane tangential to the cortical surface the barrels constitute a somatotopic map of the whiskers, with cells within adjacent barrels responding most strongly and quickly to deflection of adjacent whiskers \citep{armstrong-james_flow_1992}.
 Barrel patterning reflects subcortical whisker maps comprising cell aggregates called barrelettes in the brainstem and barreloids in the thalamus.
@@
 
-The present results suggest that the key requirements for the emergence of realistic barrel patterning are i) at each cortical location thalamocortical projections compete for a limited number of available synaptic connections (Eqs.~\ref{eq:dc}--\ref{eq:da}), ii) at each location the branching rate of a given projection is reduced by the density of other projections (Eq.~\ref{eq:comp}), and iii) the branch density of each projection is conserved (Eq.~\ref{eq:norm}).
+The present results suggest that the key requirements for the emergence of realistic barrel patterning are i) at each cortical location thalamocortical projections compete for a limited number of available synaptic connections (Eqs.~\ref{eq:dc}--\ref{eq:da}), ii) at each location the branching rate of a given projection is reduced by the density of other projections (Eq.~\ref{eq:comp}), and iii) the branch density of each projection is conserved over time (Eq.~\ref{eq:norm}).
 
 It is important to emphasize that the formulation of the model is entirely local, insofar as simulation requires no information to be communicated from a given cortical grid cell to any but those immediately adjacent (via diffusion).
-Hence the simulations demonstrate how a self-organising system, constrained by genetically specified guidance cues, can faithfully reproduce an arrangement of cell aggregates in one neural structure as a topographic map in another.
-%
-The model also shows that the shape of the barrels may be primarily determined
-by the competitive and cooperative interactions between cells relative to the
-shape of the cortical field boundary. And that somatotopic map formation does
-not require the pre-specification of centre-points by as yet undetermined
-additional developmental mechanisms.
+Hence the simulations demonstrate how a self-organising system, constrained by genetically specified guidance cues and by the shape of the cortical field boundary, can faithfully reproduce an arrangement of cell aggregates in one neural structure as a topographic map in another.
+
+The emergence of barrels in simulation required competition between thalamic projections in terms of synaptic connectivity, and also in terms of cortical space (as represented by $\chi$).
+This requirement may account for the absence of barrels in rodents with larger brains, such as capybara, for which competition for space is presumably weaker \citep{Woolsey1975}.
+Hence, irrespective of whether barrels are necessary for adaptive whisker function, the emergence of somatotopically ordered modular structures may be an inevitable consequence of local competition for cortical territory driven by input from an array of discrete sensory organs \citep{purves_iterated_1992}.
+
+Moreover, the present results confirm that such structures can emerge without requiring the pre-specification of centre-points by as yet undetermined additional developmental mechanisms.
+
+%The model also shows that the shape of the barrels may be primarily determined by the competitive and cooperative interactions between cells relative to the shape of the cortical field boundary.
+And that somatotopic map formation does not require the pre-specification of centre-points by as yet undetermined additional developmental mechanisms.
 

### 2019-12-19 paper/briefpaper/brief.tex

@@
 \textbf{C} Simulation results for parameters $N=41$, $\alpha=3$, $\beta=20$, $k=3$, $D=0.2$, $\gamma\in\pm 2$ and $\delta{t}=0.0001$.
-Colours indicate the thalamic projection for which the connection density is maximal, black lines delineate boundaries, and overlaid contours show $c>0.2$.
-\textbf{D} The Honda $\delta$ metric (red dots; see \cite{senft_mouse_1991}) measures departure from a Voronoi tesselation (dotted line corresponds to the barrels in $A$), and $\eta$ (black squares) measures the correspondence between the real and simulated barrel shapes (the product of the sum of squared differences between real and simulated centers and the sum of squared differences in area; units mm$^4$).
-\textbf{F.} Guidance fields and emergent barrel pattern in a Fgf8 misexpression experiment (c.f.
+Colours indicate the thalamic projection for which the connection density is maximal, black lines delineate boundaries, and overlaid contours show $c>0.5$.
+\textbf{D} The Honda $\delta$ metric (red dots; see \cite{senft_mouse_1991}) measures departure from a Voronoi tesselation (dotted line corresponds to the barrels in $A$), and $\eta$ (black squares) measures the correspondence between the real and simulated barrel shapes (the product of the sum of squared differences between real and simulated centers and the sum of differences in area; units mm$^4$).
+\textbf{E.} Guidance fields and emergent barrel pattern in a Fgf8 misexpression experiment (c.f.
 \cite{assimacopoulos_fibroblast_2012}), simulated by reflecting $\rho_1$ about the anterior-posterior axis.

### 2019-12-19 paper/briefpaper/brief.tex
fdsf

@@
 \begin{center}
-\includegraphics[width=0.8\textwidth]{./MainFig.png}
+\includegraphics[width=0.8\textwidth]{./MainFig.pdf}
 \end{center}
-\caption{\textbf{A} Left shows a cytochrome oxidase stain obtained from rat S1 by \cite{zheng_signal_2001}, with black lines to delineate barrels and to measure departure ($\delta$) from a perfect Voronoi tesselation.
-Right shows the initial distribution of axon branching density ($a$) for one thalamocortical projection and two molecular guidance fields ($\rho$).
-\textbf{B} The strength of interaction $\gamma$ for each of 41 projections with field $\rho_1$ and $\rho_2$ are indicated by the lengths of green and blue arrows respectively, assuming that similar fields aligned to the posterior-anterior and medial-lateral axes in the ventroposterior medial nucleus of the thalamus are sampled at the locations of barreloid centers (reconstructed from \cite{haidarliu_size_2001}).
-\textbf{C} Simulation results for parameters $N=41$, $\alpha=3$, $\beta=20$, $k=3$, $D=0.2$, $\gamma\in\pm 2$ and $\delta{t}=0.0001$.
+\caption{\textbf{A} Left shows a cytochrome oxidase stain obtained from rat S1 by \cite{zheng_signal_2001}, with black lines to delineate barrels and to measure departure (Honda-$\delta$; see \cite{senft_mouse_1991}) from a perfect Voronoi tesselation.
+Right shows the initial distribution of axon branching density ($a$) for one thalamocortical projection, and two molecular guidance fields ($\rho$).
+\textbf{B} The strengths of interaction $\gamma$ with fields $\rho_1$ and $\rho_2$ are indicated for each of 41 projections by the lengths of green and blue arrows respectively, assuming that similar fields aligned to the posterior-anterior and medial-lateral axes in the ventroposterior medial nucleus of the thalamus are sampled at the locations of putative barreloid centers (reconstructed from \cite{haidarliu_size_2001}).
+\textbf{C} Simulation results for parameters $N=41$, $\alpha=3$, $\beta=20$, $k=3$, $D=0.2$, $\gamma\in\pm 2$, $\epsilon=150$ and $\delta{t}=0.0001$.
 Colours indicate the thalamic projection for which the connection density is maximal, black lines delineate boundaries, and overlaid contours show $c>0.5$.
-\textbf{D} The Honda $\delta$ metric (red dots; see \cite{senft_mouse_1991}) measures departure from a Voronoi tesselation (dotted line corresponds to the barrels in $A$), and $\eta$ (black squares) measures the correspondence between the real and simulated barrel shapes (the product of the sum of squared differences between real and simulated centers and the sum of differences in area; units mm$^4$).
+\textbf{D} Red dots show the Honda-$\delta$ metric obtained from simulation approaching that obtained from barrels in \textbf{A} (dotted line);  black squares measure the correspondence between the real and simulated barrel shapes, $\eta$ (the product of the sum of squared differences between real and simulated centers and the sum of differences in area; units mm$^4$).
 \textbf{E.} Guidance fields and emergent barrel pattern in a Fgf8 misexpression experiment (c.f.
-\cite{assimacopoulos_fibroblast_2012}), simulated by reflecting $\rho_1$ about the anterior-posterior axis.
+\cite{assimacopoulos_fibroblast_2012}), simulated by reflecting $\rho_1$ at the join ($\epsilon=80$).
 All scale bars 1mm.}
@@
 
-It is important to emphasize that the formulation of the model is entirely local, insofar as simulation requires no information to be communicated from a given cortical grid cell to any but those immediately adjacent (via diffusion).
-Hence the simulations demonstrate how a self-organising system, constrained by genetically specified guidance cues and by the shape of the cortical field boundary, can faithfully reproduce an arrangement of cell aggregates in one neural structure as a topographic map in another.
-
 The emergence of barrels in simulation required competition between thalamic projections in terms of synaptic connectivity, and also in terms of cortical space (as represented by $\chi$).
@@
 
-Moreover, the present results confirm that such structures can emerge without requiring the pre-specification of centre-points by as yet undetermined additional developmental mechanisms.
+It is important to emphasize that the formulation of the model is entirely local, insofar as simulation requires no information to be communicated from a given cortical grid cell to any but those immediately adjacent (via diffusion).
+Hence the simulations demonstrate how a self-organising system, constrained by genetically specified guidance cues and by the shape of the cortical field boundary, can faithfully reproduce an arrangement of cell aggregates in one neural structure as a topographic map in another.
 
-%The model also shows that the shape of the barrels may be primarily determined by the competitive and cooperative interactions between cells relative to the shape of the cortical field boundary.
-And that somatotopic map formation does not require the pre-specification of centre-points by as yet undetermined additional developmental mechanisms.
+Moreover, the present results confirm that somatotopic map formation does not require the pre-specification of centre-points by as yet undetermined additional developmental mechanisms.
 

### 2019-12-19 paper/briefpaper/brief.tex
fdsf

@@
 \textbf{C} Simulation results for parameters $N=41$, $\alpha=3$, $\beta=20$, $k=3$, $D=0.2$, $\gamma\in\pm 2$, $\epsilon=150$ and $\delta{t}=0.0001$.
-Colours indicate the thalamic projection for which the connection density is maximal, black lines delineate boundaries, and overlaid contours show $c>0.5$.
+Colours indicate the thalamic projection for which the connection density is maximal, black lines delineate boundaries, and overlaid contours show $c>0.5$ (see Movie S1).
 \textbf{D} Red dots show the Honda-$\delta$ metric obtained from simulation approaching that obtained from barrels in \textbf{A} (dotted line);  black squares measure the correspondence between the real and simulated barrel shapes, $\eta$ (the product of the sum of squared differences between real and simulated centers and the sum of differences in area; units mm$^4$).
-\textbf{E.} Guidance fields and emergent barrel pattern in a Fgf8 misexpression experiment (c.f.
+\textbf{E} Guidance fields and emergent barrel pattern in a Fgf8 misexpression experiment (c.f.
 \cite{assimacopoulos_fibroblast_2012}), simulated by reflecting $\rho_1$ at the join ($\epsilon=80$).
@@
 A cortical boundary enclosing 41 corresponding barrels was traced from a cytochrome oxidase stain from \cite{zheng_signal_2001}, and Eqs.~\ref{eq:dc}--\ref{eq:norm} were solved for $N=41$ projections on the resulting domain, using $M=2$ linear signalling gradients aligned with the anterior-posterior and medial-lateral axes.
-From random initial conditions for $a({\bf x},0)$ and $c({\bf x},0)$, a clear Voronoi-like tesselation emerged (Fig.\,1C).
+From random initial conditions for $a({\bf x},0)$ and $c({\bf x},0)$, a clear Voronoi-like tesselation emerged (Fig.\,1C; see Movie S1).
 A reduction in the Honda delta metric (see \cite{senft_mouse_1991}) confirmed that a `good' Voronoi pattern emerged within $\approx 3000$ iterations (see Fig.\,1D).

### 2019-12-20 paper/briefpaper/brief.tex
Use png for submission PDF
Adds acknowledgement and movie caption

@@
 \begin{center}
-\includegraphics[width=0.8\textwidth]{./MainFig.pdf}
+\includegraphics[width=0.8\textwidth]{./MainFig.png}
 \end{center}
@@
 
+\subsection*{Movie S1 caption}
+
+Movie corresponding to Fig.\,1C in the main paper. Simulation parameters were
+$N=41$, $\alpha=3$, $\beta=20$, $k=3$, $D=0.2$, $\gamma\in\pm 2$,
+$\epsilon=150$ and $\delta{t}=0.0001$. Colours indicate the thalamic
+projection for which the connection density is maximal, black lines delineate
+boundaries, and overlaid contours show $c>0.5$. The final frame in the movie
+is step 25000 of the simulation.
+
 \section*{Discussion}
@@
 
+\acknow{The authors thank Jason Berwick at the University of Sheffield for
+  access to original rat barrel stains. This work was supported by a
+  Collaborative Activity Award, \emph{Cortical Plasticity Within and Across
+  Lifetimes}, from the James S.~McDonnell Foundation (grant 220020516).}
+
+\showacknow{} % Display the acknowledgments section
+
 \bibliography{../BarrelEmerge}

### 2019-12-20 paper/briefpaper/brief_pp.sh
Explicitly choose 17.8 cm as per guidelines

@@ -1,0 +1,6 @@
+#!/bin/bash
+
+xelatex brief_pp.tex
+bibtex brief_pp.aux
+xelatex brief_pp.tex
+xelatex brief_pp.tex

### 2019-12-20 paper/briefpaper/brief.tex
Explicitly choose 17.8 cm as per guidelines
Tweak acknowl

@@
 \begin{center}
-\includegraphics[width=0.8\textwidth]{./MainFig.png}
+\includegraphics[width=17.8cm]{./MainFig.png}
 \end{center}
@@
 \acknow{The authors thank Jason Berwick at the University of Sheffield for
-  access to original rat barrel stains. This work was supported by a
-  Collaborative Activity Award, \emph{Cortical Plasticity Within and Across
-  Lifetimes}, from the James S.~McDonnell Foundation (grant 220020516).}
+  advice and for access to the rat barrel stains used to construct
+  Fig.\,1A. This work was supported by a Collaborative Activity Award,
+  \emph{Cortical Plasticity Within and Across Lifetimes}, from the James
+  S.~McDonnell Foundation (grant 220020516).}
 

### 2019-12-20 misc/movie_reduce.sh
The incantation to reduce the size of the movie from 109 MB to about 12 MB

@@ -1,0 +1,1 @@
+ffmpeg -i 41N2M_thalguide_eps150_movie_c_id.mp4  -s 900x825 -b:v 2000k -vcodec mpeg4 -acodec copy reduced_movie.mp4

### 2019-12-21 paper/briefpaper/brief.tex
Minor whitespace wrangling
Squeezed some more blood out

--- 2019-12-19 paper/briefpaper/brief.tex
@@
 % thalamocortical axons afferent from the ventrobasal thalamus, but
-% without specifying absolute positions as centre-points for the
+% without specifying absolute positions as center-points for the
 % barrels. Everything in the paper should be tailored towards arguing
@@
 \author[a,1]{Sebastian~S.~James}
-\author[b]{Leah A. Krubitzer}
+\author[b]{Leah~A.~Krubitzer}
 \author[a]{Stuart~P.~Wilson}
@@
 
-\dropcap{S}patial patterns in neural connectivity provide clues about the constraints under which brains evolve and develop \citep{purves_iterated_1992}.
-Perhaps the most distinctive pattern can be found in the barrel cortex of many rodent species \cite{woolsey_structural_1970}.
-The barrels are identifiable soon after birth in layer 4 of primary somatosensory cortex as dense clusters of thalamocortical axons, which are enclosed by borders a few neurons thick from postnatal day 3 \citep{erzurumlu_development_2012}.
-In the plane tangential to the cortical surface the barrels constitute a somatotopic map of the whiskers, with cells within adjacent barrels responding most strongly and quickly to deflection of adjacent whiskers \citep{armstrong-james_flow_1992}.
-Barrel patterning reflects subcortical whisker maps comprising cell aggregates called barrelettes in the brainstem and barreloids in the thalamus.
+\dropcap{S}patial patterns in neural connectivity provide clues about the
+constraints under which brains evolve and develop
+\citep{purves_iterated_1992}. A distinctive example is found in the barrel
+cortex of many rodent species. The barrels
+become identifiable in layer 4 of primary somatosensory cortex as clusters of
+thalamocortical axons, enclosed by borders a few neurons thick from postnatal
+day 3 \citep{erzurumlu_development_2012}. Across the
+cortical surface the barrels constitute a somatotopic map of the whiskers,
+with cells within adjacent barrels responding most strongly and quickly to
+deflection of adjacent whiskers \citep{armstrong-james_flow_1992}. Barrel
+patterning reflects subcortical whisker maps comprising cell aggregates called
+barrelettes in the brainstem and barreloids in the thalamus.
 %\cite{van_der_loos_barreloids_1976}.
 
-Barrel formation requires afferent input from whisker stimulation and thalamic calcium waves \citep{anton-bolanos_prenatal_2019}, and depends on a complex network of axon guidance molecules such as ephrin-A5 and A7 and adhesion molecules such as cadherin-6 and 8 \citep{vanderhaeghen_mapping_2000,miller_epha7-ephrin-a5_2006}.
+Barrel formation requires afferent input from whisker stimulation and thalamic calcium waves \citep{anton-bolanos_prenatal_2019}, and depends on a complex network of axon guidance molecules such as ephrin-A5 and A7 and adhesion molecules such as cadherin-6 and 8 \citep{miller_epha7-ephrin-a5_2006}.
 This network is orchestrated by interactions between morphogens Fgf8 and Fgf17 and transcription factors Emx2, Pax6, Sp8, and Coup-tf1 \citep{shimogori_fibroblast_2005}, which are expressed in gradients that mark orthogonal axes and can be manipulated to stretch, shrink, shift, and even duplicate barrels \cite{assimacopoulos_fibroblast_2012}.
@@
 
-The barrel boundaries form a Voronoi tessellation across the cortical sheet \citep{senft_mouse_1991} (Fig.\,1A), suggesting that barreloid topology is preserved in the projection of thalamocortical axons into the cortex, and that a barrel forms by lateral axon branching from an initial centre-point that ceases upon contact with axons branching from adjacent centres.
-But the assumption of pre-arranged centres is difficult to resolve with the observation that axons arrive in the cortical plate as an undifferentiated bundle, prior to barreloid formation \cite{agmon_organized_1993}.
+The barrel boundaries form a Voronoi tessellation across the cortical sheet \citep{senft_mouse_1991} (Fig.\,1A), suggesting that barreloid topology is preserved in the projection of thalamocortical axons into the cortex, and that a barrel forms by lateral axon branching from an initial center-point that ceases upon contact with axons branching from adjacent centers.
+But the assumption of pre-arranged centers is difficult to resolve with the observation that axons arrive in the cortical plate as an undifferentiated bundle, prior to barreloid formation \cite{agmon_organized_1993}.
 
-Alternatively, reaction-diffusion dynamics could generate a Voronoi tessellation without pre-arranged centres, by amplifying characteristic modes in a noisy initial distribution of axon branches, as a net effect of short-range cooperative and longer-range competitive interactions.
+Alternatively, reaction-diffusion dynamics could generate a Voronoi tessellation without pre-arranged centers, by amplifying characteristic modes in a noisy initial distribution of axon branches, as a net effect of short-range cooperative and longer-range competitive interactions.
 Accordingly, the barrel pattern would be determined by the relative strength of these interactions and by the shape of the cortical field boundary.
-However, intrinsic cortical dynamics alone cannot account for the topographic correspondence between thalamic and cortical domains, the irregular sizes and specific arrangement of the barrels in rows and arcs, or the influence of gene expression gradients.
+However, intrinsic cortical dynamics alone cannot account for the topographic correspondence between thalamic and cortical domains, the irregular sizes and specific arrangement of the barrels, or the influence of gene expression gradients.
 
-The centre-point and reaction-diffusion models are not mutually exclusive.
-Pre-organised centres could bias reaction-diffusion processes to generate specific  arrangements more reliably, and mechanisms of lateral axon branching may constitute the tension between cooperation and competition required for self-organization.
-However, proof that barrel patterning can emerge from an undifferentiated bundle of  axons, based only on local interactions, would show that a separate stage and/or extrinsic mechanism for pre-organising thalamocortical connections need not be assumed.
-To this end, we ask whether barrel maps can emerge in a system with reaction-diffusion dynamics, under the guidance of signalling gradients, and in the absence of pre-defined centres.
+The center-point and reaction-diffusion models are not mutually exclusive.
+Pre-organised centers could bias reaction-diffusion processes to generate specific  arrangements more reliably, and mechanisms of lateral axon branching may constitute the tension between cooperation and competition required for self-organization.
+However, proof that barrel patterning can emerge from an undifferentiated bundle of  axons, based only on local interactions, would show that a separate mechanism for pre-organising thalamocortical connections need not be assumed.
+To this end, we ask whether barrel maps can emerge in a system with reaction-diffusion dynamics, under the guidance of signalling gradients, and in the absence of pre-defined centers.
 
@@
 
-Software for modelling the cortical sheet as a hexagonal lattice with no-flux
-boundary conditions and an arbitrary boundary shape, for computing the flux
-using a finite difference method, for numerical integration by the
-fourth-order Runge-Kutta method, and for reproducing these results is available at \url{https://github.com/ABRG-Models/BarrelEmerge}.
+%All software required to reproduce these results is available at \url{https://github.com/ABRG-Models/BarrelEmerge}.
 
@@
 \begin{center}
-\includegraphics[width=0.8\textwidth]{./MainFig.pdf}
+\includegraphics[width=17.8cm]{./MainFig.png}
 \end{center}
-\caption{\textbf{A} Left shows a cytochrome oxidase stain obtained from rat S1 by \cite{zheng_signal_2001}, with black lines to delineate barrels and to measure departure (Honda-$\delta$; see \cite{senft_mouse_1991}) from a perfect Voronoi tesselation.
-Right shows the initial distribution of axon branching density ($a$) for one thalamocortical projection, and two molecular guidance fields ($\rho$).
-\textbf{B} The strengths of interaction $\gamma$ with fields $\rho_1$ and $\rho_2$ are indicated for each of 41 projections by the lengths of green and blue arrows respectively, assuming that similar fields aligned to the posterior-anterior and medial-lateral axes in the ventroposterior medial nucleus of the thalamus are sampled at the locations of putative barreloid centers (reconstructed from \cite{haidarliu_size_2001}).
-\textbf{C} Simulation results for parameters $N=41$, $\alpha=3$, $\beta=20$, $k=3$, $D=0.2$, $\gamma\in\pm 2$, $\epsilon=150$ and $\delta{t}=0.0001$.
-Colours indicate the thalamic projection for which the connection density is maximal, black lines delineate boundaries, and overlaid contours show $c>0.5$ (see Movie S1).
-\textbf{D} Red dots show the Honda-$\delta$ metric obtained from simulation approaching that obtained from barrels in \textbf{A} (dotted line);  black squares measure the correspondence between the real and simulated barrel shapes, $\eta$ (the product of the sum of squared differences between real and simulated centers and the sum of differences in area; units mm$^4$).
-\textbf{E} Guidance fields and emergent barrel pattern in a Fgf8 misexpression experiment (c.f.
-\cite{assimacopoulos_fibroblast_2012}), simulated by reflecting $\rho_1$ at the join ($\epsilon=80$).
-All scale bars 1mm.}
+\caption{\textbf{A} Left shows a cytochrome oxidase stain obtained from rat S1
+  by \cite{zheng_signal_2001}, with black lines to delineate barrels and to
+  measure departure (Honda-$\delta$; see \cite{senft_mouse_1991}) from a
+  perfect Voronoi tesselation. Right shows the initial distribution of axon
+  branching density ($a$) for one thalamocortical projection, and two
+  molecular guidance fields ($\rho$). \textbf{B} The strengths of interaction
+  $\gamma$ with fields $\rho_1$ and $\rho_2$ are indicated for each of 41
+  projections by the lengths of green and blue arrows respectively, assuming
+  that similar fields aligned to the posterior-anterior and medial-lateral
+  axes in the ventroposterior medial nucleus of the thalamus are sampled at
+  the locations of putative barreloid centers (reconstructed from
+  \cite{haidarliu_size_2001}). \textbf{C} Simulation results for parameters
+  $N=41$, $\alpha=3$, $\beta=20$, $k=3$, $D=0.2$, $\gamma\in\pm 2$,
+  $\epsilon=150$ and $\delta{t}=0.0001$. Colours indicate the thalamic
+  projection for which the connection density is maximal, black lines
+  delineate boundaries, and overlaid contours show $c>0.5$ (see Movie
+  S1). Software for modelling the cortical sheet as a hexagonal lattice with
+  no-flux boundary conditions and an arbitrary boundary shape, for computing
+  the flux using a finite difference method, for numerical integration by the
+  fourth-order Runge-Kutta method, and for reproducing these results is
+  available at \url{https://github.com/ABRG-Models/BarrelEmerge} \textbf{D} Red
+  dots show the Honda-$\delta$ metric obtained from simulation approaching
+  that obtained from barrels in \textbf{A} (dotted line); black squares
+  measure the correspondence between the real and simulated barrel shapes,
+  $\eta$ (the product of the sum of squared differences between real and
+  simulated centers and the sum of differences in area; units
+  mm$^4$). \textbf{E} Guidance fields and emergent barrel pattern in a Fgf8
+  misexpression experiment (c.f. \cite{assimacopoulos_fibroblast_2012}),
+  simulated by reflecting $\rho_1$ at the join ($\epsilon=80$). All scale bars
+  1mm.}
 \label{fig:main}
@@
 
-First we verified that all results established by \cite{karbowski_model_2004} for a 1D axis could be reproduced using our extension to a 2D cortical sheet.
-Using an elliptical boundary with $M=3$ offset guidance gradients aligned to the longer axis, $N=5$ thalamocortical projections gave rise to five distinct cortical fields at locations that preserved the topographic ordering defined by the original $\gamma$ values.
+First we verified that all results established by \cite{karbowski_model_2004}
+for a 1D axis could be reproduced in our 2D implementation.
+Using an elliptical boundary with $M=3$ offset guidance gradients aligned to the longer axis, $N=5$ projections gave rise to five distinct cortical fields at locations that preserved the topographic ordering defined by the original $\gamma$ values.
 However, we found that specifying $N$ ordered areas required $M\approx (N+1)/2$ signalling fields.
@@
 
-The term in parentheses in Eq.~\ref{eq:dc} represents competition between thalamocortical projections for a limited availability of cortical connections.
-To introduce competition also in terms of axon branching we redefined
+The term in parentheses in Eq.~\ref{eq:dc} represents competition between
+projections for limited cortical connections. To introduce axon branching
+competition we redefined
 %
@@
 As such, two orthogonal linear thalamic gradients were defined, from which 41 pairs of $\gamma$ values were sampled, at the coordinates of 41 barreloid centers recreated from \cite{haidarliu_size_2001} (Fig.\,1B).
+%
+A cortical boundary enclosing 41 corresponding barrels was traced from a
+cytochrome oxidase stain from \cite{zheng_signal_2001}, and
+Eqs.~\ref{eq:dc}--\ref{eq:norm} were solved for $N=41$ projections on the
+resulting domain, using $M=2$ linear signalling gradients aligned with the
+anterior-posterior and medial-lateral axes. From random initial conditions for
+$a({\bf x},0)$, a clear Voronoi-like tesselation emerged (Fig.\,1C; Movie
+S1). The Honda $\delta$ metric (see \cite{senft_mouse_1991}) confirmed that a
+`good' Voronoi pattern emerged within $\approx 3000$ iterations (Fig.\,1D). A
+measure of the difference in shape between real and simulated barrels revealed
+a strong correspondence (see Fig.\,1D).
 
-A cortical boundary enclosing 41 corresponding barrels was traced from a cytochrome oxidase stain from \cite{zheng_signal_2001}, and Eqs.~\ref{eq:dc}--\ref{eq:norm} were solved for $N=41$ projections on the resulting domain, using $M=2$ linear signalling gradients aligned with the anterior-posterior and medial-lateral axes.
-From random initial conditions for $a({\bf x},0)$ and $c({\bf x},0)$, a clear Voronoi-like tesselation emerged (Fig.\,1C; see Movie S1).
-A reduction in the Honda delta metric (see \cite{senft_mouse_1991}) confirmed that a `good' Voronoi pattern emerged within $\approx 3000$ iterations (see Fig.\,1D).
-A measure of the difference in shape between real and simulated barrels revealed a strong correspondence (see Fig.\,1D).
+To further investigate the interplay of genetic and intrinsic mechanisms we
+simulated a seminal barrel duplication experiment
+\cite{shimogori_fibroblast_2005}, by solving on a domain comprising two
+separate, identically shaped cortical boundaries. In this case one distinct
+cortical field per thalamic projection emerged, within one of the two
+boundaries only, i.e., no duplication. However, merging the two fields to
+create an extended mirror-symmetric boundary shape, with an anterior-posterior
+guidance field reflected at the join, gave rise to two mirror-symmetrical
+barrel fields comprising $\approx 2N$ barrels, i.e., two cortical fields for
+each projection (Fig.\,1E). Together these results suggest that by the
+misexpression of Fgf8, Shimogori \& Grove \citep{shimogori_fibroblast_2005}
+effectively created one large, rather than two distinct, cortical fields.
 
-To further investigate the interplay of genetic and intrinsic mechanisms we simulated a seminal barrel duplication experiment \cite{shimogori_fibroblast_2005},  first by solving on a cortical domain comprising two separate, identically shaped cortical boundaries.
-In this case one distinct cortical field per thalamic projection emerged, within one of the two boundaries only, i.e., no duplication.
-However, merging the two fields to create an extended mirror-symmetric boundary shape, with an anterior-posterior guidance field reflected at the join, gave rise to two mirror-symmetrical barrel fields comprising 2N barrels, i.e., two cortical fields for each thalamic projection (Fig.\,1E).
-Together these results suggest that by the misexpression of Fgf8, Shimogori \& Grove \citep{shimogori_fibroblast_2005} effectively created one large barrel field rather than two distinct cortical fields.
+% Omit to shorten manuscript
+%\subsection*{Movie S1 caption}
+%
+%Movie corresponding to Fig.\,1C in the main paper. Simulation parameters were
+%$N=41$, $\alpha=3$, $\beta=20$, $k=3$, $D=0.2$, $\gamma\in\pm 2$,
+%$\epsilon=150$ and $\delta{t}=0.0001$. Colours indicate the thalamic
+%projection for which the connection density is maximal, black lines delineate
+%boundaries, and overlaid contours show $c>0.5$. The final frame in the movie
+%is step 25000 of the simulation.
 
@@
 
-The present results suggest that the key requirements for the emergence of realistic barrel patterning are i) at each cortical location thalamocortical projections compete for a limited number of available synaptic connections (Eqs.~\ref{eq:dc}--\ref{eq:da}), ii) at each location the branching rate of a given projection is reduced by the density of other projections (Eq.~\ref{eq:comp}), and iii) the branch density of each projection is conserved over time (Eq.~\ref{eq:norm}).
+These results suggest that the emergence of realistic barrel patterning
+requires that i) at each cortical location thalamocortical projections compete for a limited number of available synaptic connections (Eqs.~\ref{eq:dc}--\ref{eq:da}), ii) at each location the branching rate of a given projection is reduced by the density of other projections (Eq.~\ref{eq:comp}), and iii) the branch density of each projection is conserved over time (Eq.~\ref{eq:norm}).
 
-The emergence of barrels in simulation required competition between thalamic projections in terms of synaptic connectivity, and also in terms of cortical space (as represented by $\chi$).
+The emergence of barrels in simulation required competition between projections for synaptic connectivity, and also for cortical space (represented by $\chi$).
 This requirement may account for the absence of barrels in rodents with larger brains, such as capybara, for which competition for space is presumably weaker \citep{Woolsey1975}.
@@
 
-It is important to emphasize that the formulation of the model is entirely local, insofar as simulation requires no information to be communicated from a given cortical grid cell to any but those immediately adjacent (via diffusion).
-Hence the simulations demonstrate how a self-organising system, constrained by genetically specified guidance cues and by the shape of the cortical field boundary, can faithfully reproduce an arrangement of cell aggregates in one neural structure as a topographic map in another.
+We emphasize that the formulation of the model is entirely local; the
+simulation permits information to be communicated from a given cortical grid
+cell only to those immediately adjacent (via diffusion). Hence the simulations
+demonstrate how a self-organising system, constrained by genetically specified
+guidance cues and by the shape of the boundary, can faithfully reproduce an
+arrangement of cell aggregates in one neural structure as a topographic map in
+another.
 
-Moreover, the present results confirm that somatotopic map formation does not require the pre-specification of centre-points by as yet undetermined additional developmental mechanisms.
+Moreover, the present results confirm that somatotopic map formation does not require the pre-specification of center-points by as yet undetermined additional developmental mechanisms.
+
+\acknow{The authors thank Jason Berwick for advice and data used to construct
+  Fig.\,1A. This work was supported by the James S.~McDonnell Foundation
+  (grant 220020516).}
+
+\showacknow{} % Display the acknowledgments section
 

### 2019-12-21 paper/briefpaper/brief.tex
Justifies paragraphs

@@
 \begin{abstract}
-Brain development relies on an interplay between genetic specification and self-organisation.
-Striking examples of this relationship can be found in the  somatosensory brainstem, thalamus, and cortex of rats and mice, where the arrangement of the facial whiskers is preserved in the arrangement of cell aggregates to form precise somatotopic maps.
-We show in simulation how realistic whisker maps can self-organize, by assuming that information is exchanged between adjacent cells only, under the guidance of gene expression gradients.
-The resulting model provides a simple account of how patterns of gene expression can constrain spontaneous pattern formation to faithfully reproduce functional maps in subsequent brain structures.
+Brain development relies on an interplay between genetic specification and
+self-organisation. Striking examples of this relationship can be found in the
+somatosensory brainstem, thalamus, and cortex of rats and mice, where the
+arrangement of the facial whiskers is preserved in the arrangement of cell
+aggregates to form precise somatotopic maps. We show in simulation how
+realistic whisker maps can self-organize, by assuming that information is
+exchanged between adjacent cells only, under the guidance of gene expression
+gradients. The resulting model provides a simple account of how patterns of
+gene expression can constrain spontaneous pattern formation to faithfully
+reproduce functional maps in subsequent brain structures.
 \end{abstract}
@@
 \citep{purves_iterated_1992}. A distinctive example is found in the barrel
-cortex of many rodent species. The barrels
-become identifiable in layer 4 of primary somatosensory cortex as clusters of
-thalamocortical axons, enclosed by borders a few neurons thick from postnatal
-day 3 \citep{erzurumlu_development_2012}. Across the
-cortical surface the barrels constitute a somatotopic map of the whiskers,
-with cells within adjacent barrels responding most strongly and quickly to
-deflection of adjacent whiskers \citep{armstrong-james_flow_1992}. Barrel
-patterning reflects subcortical whisker maps comprising cell aggregates called
-barrelettes in the brainstem and barreloids in the thalamus.
+cortex of many rodent species. The barrels become identifiable in layer 4 of
+primary somatosensory cortex as clusters of thalamocortical axons, enclosed by
+borders a few neurons thick from postnatal day 3
+\citep{erzurumlu_development_2012}. Across the cortical surface the barrels
+constitute a somatotopic map of the whiskers, with cells within adjacent
+barrels responding most strongly and quickly to deflection of adjacent
+whiskers \citep{armstrong-james_flow_1992}. Barrel patterning reflects
+subcortical whisker maps comprising cell aggregates called barrelettes in the
+brainstem and barreloids in the thalamus.
 %\cite{van_der_loos_barreloids_1976}.
 
-Barrel formation requires afferent input from whisker stimulation and thalamic calcium waves \citep{anton-bolanos_prenatal_2019}, and depends on a complex network of axon guidance molecules such as ephrin-A5 and A7 and adhesion molecules such as cadherin-6 and 8 \citep{miller_epha7-ephrin-a5_2006}.
-This network is orchestrated by interactions between morphogens Fgf8 and Fgf17 and transcription factors Emx2, Pax6, Sp8, and Coup-tf1 \citep{shimogori_fibroblast_2005}, which are expressed in gradients that mark orthogonal axes and can be manipulated to stretch, shrink, shift, and even duplicate barrels \cite{assimacopoulos_fibroblast_2012}.
+Barrel formation requires afferent input from whisker stimulation and thalamic
+calcium waves \citep{anton-bolanos_prenatal_2019}, and depends on a complex
+network of axon guidance molecules such as ephrin-A5 and A7 and adhesion
+molecules such as cadherin-6 and 8 \citep{miller_epha7-ephrin-a5_2006}.  This
+network is orchestrated by interactions between morphogens Fgf8 and Fgf17 and
+transcription factors Emx2, Pax6, Sp8, and Coup-tf1
+\citep{shimogori_fibroblast_2005}, which are expressed in gradients that mark
+orthogonal axes and can be manipulated to stretch, shrink, shift, and even
+duplicate barrels \cite{assimacopoulos_fibroblast_2012}.
 % Removed ref: ,bishop_regulation_2000 after ..Coup-tf1
 
-The barrel boundaries form a Voronoi tessellation across the cortical sheet \citep{senft_mouse_1991} (Fig.\,1A), suggesting that barreloid topology is preserved in the projection of thalamocortical axons into the cortex, and that a barrel forms by lateral axon branching from an initial center-point that ceases upon contact with axons branching from adjacent centers.
-But the assumption of pre-arranged centers is difficult to resolve with the observation that axons arrive in the cortical plate as an undifferentiated bundle, prior to barreloid formation \cite{agmon_organized_1993}.
+The barrel boundaries form a Voronoi tessellation across the cortical sheet
+\citep{senft_mouse_1991} (Fig.\,1A), suggesting that barreloid topology is
+preserved in the projection of thalamocortical axons into the cortex, and that
+a barrel forms by lateral axon branching from an initial center-point that
+ceases upon contact with axons branching from adjacent centers. But the
+assumption of pre-arranged centers is difficult to resolve with the
+observation that axons arrive in the cortical plate as an undifferentiated
+bundle, prior to barreloid formation \cite{agmon_organized_1993}.
 
-Alternatively, reaction-diffusion dynamics could generate a Voronoi tessellation without pre-arranged centers, by amplifying characteristic modes in a noisy initial distribution of axon branches, as a net effect of short-range cooperative and longer-range competitive interactions.
-Accordingly, the barrel pattern would be determined by the relative strength of these interactions and by the shape of the cortical field boundary.
-However, intrinsic cortical dynamics alone cannot account for the topographic correspondence between thalamic and cortical domains, the irregular sizes and specific arrangement of the barrels, or the influence of gene expression gradients.
+Alternatively, reaction-diffusion dynamics could generate a Voronoi
+tessellation without pre-arranged centers, by amplifying characteristic modes
+in a noisy initial distribution of axon branches, as a net effect of
+short-range cooperative and longer-range competitive
+interactions. Accordingly, the barrel pattern would be determined by the
+relative strength of these interactions and by the shape of the cortical field
+boundary. However, intrinsic cortical dynamics alone cannot account for the
+topographic correspondence between thalamic and cortical domains, the
+irregular sizes and specific arrangement of the barrels, or the influence of
+gene expression gradients.
 
-The center-point and reaction-diffusion models are not mutually exclusive.
-Pre-organised centers could bias reaction-diffusion processes to generate specific  arrangements more reliably, and mechanisms of lateral axon branching may constitute the tension between cooperation and competition required for self-organization.
-However, proof that barrel patterning can emerge from an undifferentiated bundle of  axons, based only on local interactions, would show that a separate mechanism for pre-organising thalamocortical connections need not be assumed.
-To this end, we ask whether barrel maps can emerge in a system with reaction-diffusion dynamics, under the guidance of signalling gradients, and in the absence of pre-defined centers.
+The center-point and reaction-diffusion models are not mutually
+exclusive. Pre-organised centers could bias reaction-diffusion processes to
+generate specific arrangements more reliably, and mechanisms of lateral axon
+branching may constitute the tension between cooperation and competition
+required for self-organization. However, proof that barrel patterning can
+emerge from an undifferentiated bundle of axons, based only on local
+interactions, would show that a separate mechanism for pre-organising
+thalamocortical connections need not be assumed. To this end, we ask whether
+barrel maps can emerge in a system with reaction-diffusion dynamics, under the
+guidance of signalling gradients, and in the absence of pre-defined centers.
 
@@
 
-Karbowski \& Ermentrout \citep{karbowski_model_2004} developed a reaction-diffusion style model of how extrinsic signalling gradients can constrain the emergence of distinct fields from intrinsic cortical dynamics.
-Their model defines how the density of connections $c(x,t)$ and axon branches $a(x,t)$ interact at time $t$, along a 1D anterior-posterior axis $x$, for $N$ thalamocortical projections indexed by $i$.
-The model was derived from the assumption that the rates at which $a_i$ and $c_i$ grow are reciprocally coupled.
-Extending the original 1D model to simulate arealization on a 2D cortical sheet, we use $a_i(\mb{x},t)$ and $c_i(\mb{x},t)$, and model synaptogenesis as
+Karbowski \& Ermentrout \citep{karbowski_model_2004} developed a
+reaction-diffusion style model of how extrinsic signalling gradients can
+constrain the emergence of distinct fields from intrinsic cortical
+dynamics. Their model defines how the density of connections $c(x,t)$ and axon
+branches $a(x,t)$ interact at time $t$, along a 1D anterior-posterior axis
+$x$, for $N$ thalamocortical projections indexed by $i$. The model was derived
+from the assumption that the rates at which $a_i$ and $c_i$ grow are
+reciprocally coupled. Extending the original 1D model to simulate arealization
+on a 2D cortical sheet, we use $a_i(\mb{x},t)$ and $c_i(\mb{x},t)$, and model
+synaptogenesis as
 %
@@
 %
-Accordingly, where the total density of synaptic connections sums to one, connections decay at rate $\alpha$.
-Otherwise the connection density increases non-linearly ($k>1$) with the density of axon branching.
-Axon branching is modelled as
+Accordingly, where the total density of synaptic connections sums to one,
+connections decay at rate $\alpha$. Otherwise the connection density increases
+non-linearly ($k>1$) with the density of axon branching. Axon branching is
+modelled as
 %
@@
 %
-The first term on the right describes the divergence (indicated by $\nabla\vcdot$) of the quantity in parentheses, which is referred to as the `flux' of axonal branching.
-The flux represents diffusion across the cortical sheet, at rate $D$, and the influence of $M$ molecular signalling fields, $\rho(\mb{x})$.
-The influence of a given field (indexed by $j$) on a given thalamic projection (indexed by $i$), is determined by $\gamma_{i,j}$, which may be positive or negative in order that axons may branch in the direction of either higher or lower  concentrations.
-The second term on the right represents an assumption that axon branching decreases when synaptogenesis increases, and vice versa, i.e., that axon branches are required for synaptic connections to form.
-Here $\chi_i=0$ is a placeholder.
+The first term on the right describes the divergence (indicated by
+$\nabla\vcdot$) of the quantity in parentheses, which is referred to as the
+`flux' of axonal branching. The flux represents diffusion across the cortical
+sheet, at rate $D$, and the influence of $M$ molecular signalling fields,
+$\rho(\mb{x})$. The influence of a given field (indexed by $j$) on a given
+thalamic projection (indexed by $i$), is determined by $\gamma_{i,j}$, which
+may be positive or negative in order that axons may branch in the direction of
+either higher or lower concentrations. The second term on the right represents
+an assumption that axon branching decreases when synaptogenesis increases, and
+vice versa, i.e., that axon branches are required for synaptic connections to
+form. Here $\chi_i=0$ is a placeholder.
 
@@
 First we verified that all results established by \cite{karbowski_model_2004}
-for a 1D axis could be reproduced in our 2D implementation.
-Using an elliptical boundary with $M=3$ offset guidance gradients aligned to the longer axis, $N=5$ projections gave rise to five distinct cortical fields at locations that preserved the topographic ordering defined by the original $\gamma$ values.
-However, we found that specifying $N$ ordered areas required $M\approx (N+1)/2$ signalling fields.
-As the number of distinct guidance fields is unlikely to approach the number of barrels, further extensions were required.
+for a 1D axis could be reproduced in our 2D implementation. Using an
+elliptical boundary with $M=3$ offset guidance gradients aligned to the longer
+axis, $N=5$ projections gave rise to five distinct cortical fields at
+locations that preserved the topographic ordering defined by the original
+$\gamma$ values. However, we found that specifying $N$ ordered areas required
+$M\approx (N+1)/2$ signalling fields. As the number of distinct guidance
+fields is unlikely to approach the number of barrels, further extensions were
+required.
 
@@
 %
-which reduces branching for each projection where the branches of other projections are dense.
-Divisive normalization keeps the axon branch density bounded on each iteration to the initial density  at $t=0$;
+which reduces branching for each projection where the branches of other
+projections are dense. Divisive normalization keeps the axon branch density
+bounded on each iteration to the initial density at $t=0$;
 %
@@
 
-The only differences between thalamic projections are their strengths of interaction with the guidance fields, $\gamma$, hence any reliable differences between emergent cortical fields must be due to differences in these values only.
-We speculate that a projection interacts with a given ephrin field in the cortical subplate with a strength determined by the concentration of a similar molecule at its thalamic origin, i.e., the putative barreloid center.
-As such, two orthogonal linear thalamic gradients were defined, from which 41 pairs of $\gamma$ values were sampled, at the coordinates of 41 barreloid centers recreated from \cite{haidarliu_size_2001} (Fig.\,1B).
+The only differences between thalamic projections are their strengths of
+interaction with the guidance fields, $\gamma$, hence any reliable differences
+between emergent cortical fields must be due to differences in these values
+only. We speculate that a projection interacts with a given ephrin field in
+the cortical subplate with a strength determined by the concentration of a
+similar molecule at its thalamic origin, i.e., the putative barreloid
+center. As such, two orthogonal linear thalamic gradients were defined, from
+which 41 pairs of $\gamma$ values were sampled, at the coordinates of 41
+barreloid centers recreated from \cite{haidarliu_size_2001} (Fig.\,1B).
 %
@@
 These results suggest that the emergence of realistic barrel patterning
-requires that i) at each cortical location thalamocortical projections compete for a limited number of available synaptic connections (Eqs.~\ref{eq:dc}--\ref{eq:da}), ii) at each location the branching rate of a given projection is reduced by the density of other projections (Eq.~\ref{eq:comp}), and iii) the branch density of each projection is conserved over time (Eq.~\ref{eq:norm}).
+requires that i) at each cortical location thalamocortical projections compete
+for a limited number of available synaptic connections
+(Eqs.~\ref{eq:dc}--\ref{eq:da}), ii) at each location the branching rate of a
+given projection is reduced by the density of other projections
+(Eq.~\ref{eq:comp}), and iii) the branch density of each projection is
+conserved over time (Eq.~\ref{eq:norm}).
 
-The emergence of barrels in simulation required competition between projections for synaptic connectivity, and also for cortical space (represented by $\chi$).
-This requirement may account for the absence of barrels in rodents with larger brains, such as capybara, for which competition for space is presumably weaker \citep{Woolsey1975}.
-Hence, irrespective of whether barrels are necessary for adaptive whisker function, the emergence of somatotopically ordered modular structures may be an inevitable consequence of local competition for cortical territory driven by input from an array of discrete sensory organs \citep{purves_iterated_1992}.
+The emergence of barrels in simulation required competition between
+projections for synaptic connectivity, and also for cortical space
+(represented by $\chi$). This requirement may account for the absence of
+barrels in rodents with larger brains, such as capybara, for which competition
+for space is presumably weaker \citep{Woolsey1975}. Hence, irrespective of
+whether barrels are necessary for adaptive whisker function, the emergence of
+somatotopically ordered modular structures may be an inevitable consequence of
+local competition for cortical territory driven by input from an array of
+discrete sensory organs \citep{purves_iterated_1992}.
 
@@
 
-Moreover, the present results confirm that somatotopic map formation does not require the pre-specification of center-points by as yet undetermined additional developmental mechanisms.
+Moreover, the present results confirm that somatotopic map formation does not
+require the pre-specification of center-points by as yet undetermined
+additional developmental mechanisms.
 

### 2020-01-16 paper/supplementary/supp.tex
Adds some notes on PNAS Reviewer 1's suggestion for an additional flux term.

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\subsection{PNAS Reviewer 1 suggestions}
+\label{sec:comp15}
+
+This mechanism was suggested by PNAS reviewer 1 who asked if we'd tried
+anything similar to this as a way around requiring divisive
+normalization.
+
+This is similar to the ideas tried in Sections ~\ref{sec:comp2}
+and \ref{sec:comp6}. The PNAS brief report submission Eq. 2 was:
+%
+\begin{equation}
+\label{eq:comp15_eq0}
+\frac{\partial a_i}{\partial t} = \nabla\vcdot\mb{J}_i - \frac{\partial
+c_i}{\partial t} + \chi
+\end{equation}
+%
+Reviewer 1 suggested (where we might assume the equals sign in red was missed by typo):
+%
+\begin{equation}
+\label{eq:comp15_eq1}
+\chi \color{red}=\color{black} a_i \nabla \frac{1}{N-1} \sum_{j \ne i}^{N} a_j
+\end{equation}
+%
+This isn't really sensible, because the gradient, which has direction, will be
+positive on one side of a `hill of $a_{j\ne i}$' and negative on the other
+side. Thus, $a$ would accumulate on one side and de-accumulate on the other,
+rather than reducing all the way around the hill. However, if you took
+the \emph{magnitude} of the gradient (along with a little rearranging, and the
+incorporation of our parameter $\epsilon$) you get:
+%
+\begin{equation}
+\label{eq:comp15_eq2}
+\chi = - \frac{\epsilon_i  a_i}{N-1} \left\| \nabla  \sum_{j \ne i}^{N} a_j \right\|
+\end{equation}
+%
+which may work, although it is unclear how this resolves the need for the
+normalization that the reviewer (and we too) would like to avoid.
+
+Although the reviewer prepended $\chi$ to this equation, they probably meant
+adding to the flux term, which was the terminology they used. In that case, we
+would have exactly Eq.~\ref{eq:J_NM_with_comp} and so yes, it was something we
+tried, and I will try it again (Jan 2020) to refresh my memory of how it
+worked.
+
 \section{Unimplemented ideas}

### 2020-01-16 paper/briefpaper/brief.tex
Addresses R1's minor point

@@
 \templatetype{pnasresearcharticle}
+
+\usepackage{amsmath,esint}
+\newcommand*\dif{\mathop{}\!\mathrm{d}}
 
@@
 \begin{equation} \label{eq:dc}
-\frac{\partial c_i}{\partial t} =-\alpha c_i +\beta  \left(1 - \sum_{j=1}^{N} c_{j}\right)[a_i]^k.
+\frac{\partial c_i}{\partial t} =-\alpha c_i +\beta  \left(1 - \sum_{j=1}^{N} c_{j}\right)[a_i]^k
 \end{equation}
@@
 \begin{equation} \label{eq:da}
-\frac{\partial a_i}{\partial t} = \nabla\vcdot\left(D \nabla a_i-a_i\sum_{j=1}^{M} \gamma_{i,j}\nabla \rho_j(\mb{x})\right) - \frac{\partial c_i}{\partial t} + \chi_i.
+\frac{\partial a_i}{\partial t} = \nabla\vcdot\left(D \nabla a_i-a_i\sum_{j=1}^{M} \gamma_{i,j}\nabla \rho_j(\mb{x})\right) - \frac{\partial c_i}{\partial t} + \chi_i
 \end{equation}
@@
 for a 1D axis could be reproduced in our 2D implementation. Using an
-elliptical boundary with $M=3$ offset guidance gradients aligned to the longer
+elliptical domain, $S$, with $M=3$ offset guidance gradients aligned to the longer
 axis, $N=5$ projections gave rise to five distinct cortical fields at
@@
 \begin{equation} \label{eq:comp}
-\chi_i(\mb{x}, t) = - \frac{\epsilon  a_i}{N-1} \sum_{j \ne i}^{N} a_{j},
+\chi_i(\mb{x}, t) = - \frac{\epsilon  a_i}{N-1} \sum_{j \ne i}^{N} a_{j}
 \end{equation}
@@
 \begin{equation} \label{eq:norm}
-  a_i(\mathbf{x}, t) = \int_\Omega  a_i(\mathbf{x}, 0) d\sigma \; \frac {a'_i(\mathbf{x}, t)} {\int_\Omega
-  a'_i(\mathbf{x}, t) d\sigma},
+  a_i(\mathbf{x}, t) = \oiint_{S}  a_i(\mathbf{x}, 0) \dif S \; \frac {a'_i(\mathbf{x}, t)} {\oiint_{S}
+  a'_i(\mathbf{x}, t) \dif S}
 \end{equation}
@@
 A cortical boundary enclosing 41 corresponding barrels was traced from a
-cytochrome oxidase stain from \cite{zheng_signal_2001}, and
+cytochrome oxidase stain from \cite{zheng_signal_2001} to define $S$, and
 Eqs.~\ref{eq:dc}--\ref{eq:norm} were solved for $N=41$ projections on the

### 2020-01-24 sim/rd_james_comp2.h
Work on testing out the comp2 method again, after PNAS Rev 1 suggested this.

@@ -93,8 +93,10 @@ public:
      *
      * Stable with dt = 0.0001;
      */
-    void compute_divJ (vector<Flt>& fa, unsigned int i) {
+#define SIGMOID_ROLLOFF_FOR_A 1
+    virtual void compute_divJ (vector<Flt>& fa, unsigned int i) {
 
+#ifdef SIGMOID_ROLLOFF_FOR_A
         // Compute \bar{a}_i and its spatial gradient
         Flt h = 2.0; // height parameter for sigmoid
         Flt o = 5.0; // offset
@@ -104,6 +106,7 @@ public:
             this->abar[hi] = h / (1 - exp (o - s * fa[hi]));
         }
         this->spacegrad2D (this->abar, this->grad_abar);
+#endif
 
         // Compute gradient of a_i(x), for use computing the third term, below.
         this->spacegrad2D (fa, this->grad_a[i]);
@@ -138,39 +141,60 @@ public:
                 exit (21);
             }
 
+            //if (hi==5055) {
+            //    cout << "term1[5055]="  << term1 << endl;
+            //}
+
+#ifdef SIGMOID_ROLLOFF_FOR_A
             // Term 1.1 is F/N-1 abar div(ahat)
             Flt term1_1 = this->FOverNm1 * abar[hi] * this->div_ahat[hi];
+#else
+            // Term 1.1 is F/N-1 a div(ahat)
+            Flt term1_1 = this->FOverNm1 * fa[hi] * this->div_ahat[hi];
+#endif
             if (isnan(term1_1)) {
                 cerr << "term1_1 isnan" << endl;
                 cerr << "fa[hi="<<hi<<"] = " << fa[hi] << ", this->div_ahat[hi] = " << this->div_ahat[hi] << endl;
                 exit (21);
             }
 
+#ifdef SIGMOID_ROLLOFF_FOR_A
             // Term 1.2 is F/N-1 grad(ahat) . grad(abar)
             Flt term1_2 = this->FOverNm1 * (this->grad_ahat[0][hi] * this->grad_abar[0][hi]
                                             + this->grad_ahat[1][hi] * this->grad_abar[1][hi]);
+#else
+            // Term 1.2 is F/N-1 grad(ahat) . grad(a)
+            Flt term1_2 = this->FOverNm1 * (this->grad_ahat[0][hi] * this->grad_a[i][0][hi]
+                                            + this->grad_ahat[1][hi] * this->grad_a[i][1][hi]);
+#endif
+
+            //if (hi==1000) {
[…]

### 2020-01-24 paper/briefpaper/brief.tex
Re-wraps the bioRXiv version of the paper, then applies eLife formatting and some initial changes

--- 2019-12-21 paper/briefpaper/brief.tex
@@
 % thalamocortical axons afferent from the ventrobasal thalamus, but
-% without specifying absolute positions as center-points for the
+% without specifying absolute positions as centre-points for the
 % barrels. Everything in the paper should be tailored towards arguing
@@
 
-\documentclass[9pt,twocolumn,twoside,lineno]{pnas-new}
-\templatetype{pnasresearcharticle}
+\documentclass[9pt,lineno]{elife}
+
+\usepackage{amsmath,esint}
+\newcommand*\dif{\mathop{}\!\mathrm{d}}
 
 \title{Modelling the emergence of whisker barrels}
-% While writing, don't stop for errors
-%\nonstopmode
 
-% Use letters for affiliations, numbers to show equal authorship (if applicable) and to indicate the corresponding author
-\author[a,1]{Sebastian~S.~James}
-\author[b]{Leah~A.~Krubitzer}
-\author[a]{Stuart~P.~Wilson}
+\author[1*]{Sebastian~S.~James}
+\author[2]{Leah~A.~Krubitzer}
+\author[1]{Stuart~P.~Wilson}
 
-\affil[a]{Department of Psychology, The University of Sheffield, Sheffield, United Kingdom.}
-\affil[b]{Center for Neuroscience, The University of California, Davis, United States.}
-% $^{*}$Corresponding author email: Seb.James@Sheffield.ac.uk\vspace{1em}\\
+\affil[1]{Department of Psychology, The University of Sheffield, Sheffield, United Kingdom.}
+\affil[2]{Center for Neuroscience, The University of California, Davis, United States.}
+\corr{seb.james@sheffield.ac.uk}{SSJ}
 
-\leadauthor{James}
+%\keywords{Barrel cortex $|$ Self-organization $|$ Somatotopic map $|$ Axon guidance}
 
-\correspondingauthor{\textsuperscript{1}To whom correspondence should be addressed. E-mail: \href{mailto:seb.james@sheffield.ac.uk}{\nolinkurl{seb.james@sheffield.ac.uk}}}
+\begin{document}
 
-\keywords{Barrel cortex $|$ Self-organization $|$ Somatotopic map $|$ Axon guidance}
+\maketitle
 
@@
 
-\dates{This manuscript was compiled on \today}
-\doi{\url{www.pnas.org/cgi/doi/10.1073/pnas.XXXXXXXXXX}}
-
-\begin{document}
-
 % Blue for comments
@@
 \newcommand*\vcdot@[2]{\mathbin{\vcenter{\hbox{\scalebox{#2}{$\m@th#1\bullet$}}}}}
-\makeatother
 
-\maketitle
-\thispagestyle{firststyle}
-\ifthenelse{\boolean{shortarticle}}{\ifthenelse{\boolean{singlecolumn}}{\abscontentformatted}{\abscontent}}{}
+\section{Introduction}
 
-\modulolinenumbers{}
-\linenumbers
-
-\dropcap{S}patial patterns in neural connectivity provide clues about the
-constraints under which brains evolve and develop
-\citep{purves_iterated_1992}. A distinctive example is found in the barrel
-cortex of many rodent species. The barrels become identifiable in layer 4 of
-primary somatosensory cortex as clusters of thalamocortical axons, enclosed by
-borders a few neurons thick from postnatal day 3
-\citep{erzurumlu_development_2012}. Across the cortical surface the barrels
-constitute a somatotopic map of the whiskers, with cells within adjacent
-barrels responding most strongly and quickly to deflection of adjacent
-whiskers \citep{armstrong-james_flow_1992}. Barrel patterning reflects
-subcortical whisker maps comprising cell aggregates called barrelettes in the
-brainstem and barreloids in the thalamus.
+Spatial patterns in neural connectivity provide clues about the constraints
+under which brains evolve and develop \citep{purves_iterated_1992}. Perhaps
+the most distinctive pattern can be found in the barrel cortex of many rodent
+species \cite{woolsey_structural_1970}. The barrels are identifiable soon
+after birth in layer 4 of primary somatosensory cortex as dense clusters of
+thalamocortical axons, which are enclosed by borders a few neurons thick from
+postnatal day 3 \citep{erzurumlu_development_2012}. In the plane tangential to
+the cortical surface the barrels constitute a somatotopic map of the whiskers,
+with cells within adjacent barrels responding most strongly and quickly to
+deflection of adjacent whiskers \citep{armstrong-james_flow_1992}. Barrel
+patterning reflects subcortical whisker maps comprising cell aggregates called
+barrelettes in the brainstem and barreloids in the thalamus.
 %\cite{van_der_loos_barreloids_1976}.
@@
 network of axon guidance molecules such as ephrin-A5 and A7 and adhesion
-molecules such as cadherin-6 and 8 \citep{miller_epha7-ephrin-a5_2006}.  This
-network is orchestrated by interactions between morphogens Fgf8 and Fgf17 and
+molecules such as cadherin-6 and 8
+\citep{vanderhaeghen_mapping_2000,miller_epha7-ephrin-a5_2006}.  This network
+is orchestrated by interactions between morphogens Fgf8 and Fgf17 and
 transcription factors Emx2, Pax6, Sp8, and Coup-tf1
@@
 preserved in the projection of thalamocortical axons into the cortex, and that
-a barrel forms by lateral axon branching from an initial center-point that
-ceases upon contact with axons branching from adjacent centers. But the
-assumption of pre-arranged centers is difficult to resolve with the
-observation that axons arrive in the cortical plate as an undifferentiated
-bundle, prior to barreloid formation \cite{agmon_organized_1993}.
+a barrel forms by lateral axon branching from an initial centre-point that
+ceases upon contact with axons branching from adjacent centres.  \cmnt{Another
+  sentence about center-points here.} But the assumption of pre-arranged
+centres is difficult to resolve with the observation that axons arrive in the
+cortical plate as an undifferentiated bundle, prior to barreloid formation
+\cite{agmon_organized_1993}.
 
 Alternatively, reaction-diffusion dynamics could generate a Voronoi
-tessellation without pre-arranged centers, by amplifying characteristic modes
+tessellation without pre-arranged centres, by amplifying characteristic modes
 in a noisy initial distribution of axon branches, as a net effect of
@@
 topographic correspondence between thalamic and cortical domains, the
-irregular sizes and specific arrangement of the barrels, or the influence of
-gene expression gradients.
+irregular sizes and specific arrangement of the barrels in rows and arcs, or
+the influence of gene expression gradients.
 
-The center-point and reaction-diffusion models are not mutually
-exclusive. Pre-organised centers could bias reaction-diffusion processes to
+The centre-point and reaction-diffusion models are not mutually
+exclusive. Pre-organised centres could bias reaction-diffusion processes to
 generate specific arrangements more reliably, and mechanisms of lateral axon
@@
 emerge from an undifferentiated bundle of axons, based only on local
-interactions, would show that a separate mechanism for pre-organising
-thalamocortical connections need not be assumed. To this end, we ask whether
-barrel maps can emerge in a system with reaction-diffusion dynamics, under the
-guidance of signalling gradients, and in the absence of pre-defined centers.
+interactions, would show that a separate stage and/or extrinsic mechanism for
+pre-organising thalamocortical connections need not be assumed. To this end,
+we ask whether barrel maps can emerge in a system with reaction-diffusion
+dynamics, under the guidance of signalling gradients, and in the absence of
+pre-defined centres.
 
-\section*{Models}
+\section{Models}
 
@@
 an assumption that axon branching decreases when synaptogenesis increases, and
-vice versa, i.e., that axon branches are required for synaptic connections to
-form. Here $\chi_i=0$ is a placeholder.
+vice versa, i.e., \cmnt{when a connection is made, the axon branching
+  decreases by one branch and when a connection is lost, one branch is added
+  to the branching density}
+%that axon branches are required for synaptic connections to form.
+Here $\chi_i=0$ is a placeholder.
 
-%All software required to reproduce these results is available at \url{https://github.com/ABRG-Models/BarrelEmerge}.
+Software for modelling the cortical sheet as a hexagonal lattice with no-flux
+boundary conditions and an arbitrary boundary shape, for computing the flux
+using a finite difference method, for numerical integration by the
+fourth-order Runge-Kutta method, and for reproducing these results is
+available at \url{https://github.com/ABRG-Models/BarrelEmerge}.
 
-\section*{Results}
+\section{Results}
 
-\begin{figure*}
-\begin{center}
-\includegraphics[width=17.8cm]{./MainFig.png}
-\end{center}
+\begin{figure}
+\begin{fullwidth}
+\includegraphics[width=\linewidth]{./MainFig_superlowres.png}
 \caption{\textbf{A} Left shows a cytochrome oxidase stain obtained from rat S1
@@
   delineate boundaries, and overlaid contours show $c>0.5$ (see Movie
-  S1). Software for modelling the cortical sheet as a hexagonal lattice with
-  no-flux boundary conditions and an arbitrary boundary shape, for computing
-  the flux using a finite difference method, for numerical integration by the
-  fourth-order Runge-Kutta method, and for reproducing these results is
-  available at \url{https://github.com/ABRG-Models/BarrelEmerge} \textbf{D} Red
-  dots show the Honda-$\delta$ metric obtained from simulation approaching
-  that obtained from barrels in \textbf{A} (dotted line); black squares
-  measure the correspondence between the real and simulated barrel shapes,
-  $\eta$ (the product of the sum of squared differences between real and
-  simulated centers and the sum of differences in area; units
-  mm$^4$). \textbf{E} Guidance fields and emergent barrel pattern in a Fgf8
-  misexpression experiment (c.f. \cite{assimacopoulos_fibroblast_2012}),
-  simulated by reflecting $\rho_1$ at the join ($\epsilon=80$). All scale bars
-  1mm.}
+  S1). \textbf{D} Red dots show the Honda-$\delta$ metric obtained from
+  simulation approaching that obtained from barrels in \textbf{A} (dotted
+  line); black squares measure the correspondence between the real and
+  simulated barrel shapes, $\eta$ (the product of the sum of squared
+  differences between real and simulated centers and the sum of differences in
+  area; units mm$^4$). \textbf{E} Guidance fields and emergent barrel pattern
+  in a Fgf8 misexpression experiment
+  (c.f. \cite{assimacopoulos_fibroblast_2012}), simulated by reflecting
+  $\rho_1$ at the join ($\epsilon=80$). All scale bars 1mm.}
 \label{fig:main}
-\end{figure*}
+\end{fullwidth}
+\end{figure}
 
 First we verified that all results established by \cite{karbowski_model_2004}
-for a 1D axis could be reproduced in our 2D implementation. Using an
-elliptical boundary with $M=3$ offset guidance gradients aligned to the longer
-axis, $N=5$ projections gave rise to five distinct cortical fields at
-locations that preserved the topographic ordering defined by the original
-$\gamma$ values. However, we found that specifying $N$ ordered areas required
-$M\approx (N+1)/2$ signalling fields. As the number of distinct guidance
-fields is unlikely to approach the number of barrels, further extensions were
-required.
+for a 1D axis could be reproduced using our extension to a 2D cortical
+sheet. Using an elliptical domain, $S$, with $M=3$ offset guidance gradients
+aligned to the longer axis, $N=5$ thalamocortical projections gave rise to
+five distinct cortical fields at locations that preserved the topographic
+ordering defined by the original $\gamma$ values. However, we found that
+specifying $N$ ordered areas required $M\approx (N+1)/2$ signalling fields. As
+the number of distinct guidance fields is unlikely to approach the number of
+barrels, further extensions were required.
 
 The term in parentheses in Eq.~\ref{eq:dc} represents competition between
-projections for limited cortical connections. To introduce axon branching
-competition we redefined
+thalamocortical projections for a limited availability of cortical
+connections. To introduce competition also in terms of axon branching we
+redefined
 %
@@
 \begin{equation} \label{eq:norm}
-  a_i(\mathbf{x}, t) = \int_\Omega  a_i(\mathbf{x}, 0) d\sigma \; \frac {a'_i(\mathbf{x}, t)} {\int_\Omega
-  a'_i(\mathbf{x}, t) d\sigma},
+  a_i(\mathbf{x}, t) = \oiint_{S}  a_i(\mathbf{x}, 0) \dif S \; \frac {a'_i(\mathbf{x}, t)} {\oiint_{S}
+  a'_i(\mathbf{x}, t) \dif S}
 \end{equation}
 %
-where prime symbols indicate the use of intermediate values computed from Eqs.~\ref{eq:dc}--\ref{eq:comp} at time $t-{\delta}t$.
+where prime symbols indicate the use of intermediate values computed from
+Eqs.~\ref{eq:dc}--\ref{eq:comp} at time $t-{\delta}t$. \cmnt{Normalization
+  corresponds to the biological constraint that the branching and growth of
+  thalamocortical axons is backed by a finite resource provided by their
+  parent cell bodies.}
 
@@
 barreloid centers recreated from \cite{haidarliu_size_2001} (Fig.\,1B).
-%
+
 A cortical boundary enclosing 41 corresponding barrels was traced from a
@@
 anterior-posterior and medial-lateral axes. From random initial conditions for
-$a({\bf x},0)$, a clear Voronoi-like tesselation emerged (Fig.\,1C; Movie
-S1). The Honda $\delta$ metric (see \cite{senft_mouse_1991}) confirmed that a
-`good' Voronoi pattern emerged within $\approx 3000$ iterations (Fig.\,1D). A
-measure of the difference in shape between real and simulated barrels revealed
-a strong correspondence (see Fig.\,1D).
+$a({\bf x},0)$ and $c({\bf x},0)$, a clear Voronoi-like tesselation emerged
+(Fig.\,1C; see Movie S1). A reduction in the Honda delta metric (see
+\cite{senft_mouse_1991}) confirmed that a `good' Voronoi pattern emerged
+within $\approx 3000$ iterations (see Fig.\,1D). A measure of the difference
+in shape between real and simulated barrels revealed a strong correspondence
+(see Fig.\,1D).
 
@@
 simulated a seminal barrel duplication experiment
-\cite{shimogori_fibroblast_2005}, by solving on a domain comprising two
-separate, identically shaped cortical boundaries. In this case one distinct
-cortical field per thalamic projection emerged, within one of the two
-boundaries only, i.e., no duplication. However, merging the two fields to
+\cite{shimogori_fibroblast_2005}, first by solving on a cortical domain
+comprising two separate, identically shaped cortical boundaries. In this case
+one distinct cortical field per thalamic projection emerged, within one of the
+two boundaries only, i.e., no duplication. However, merging the two fields to
 create an extended mirror-symmetric boundary shape, with an anterior-posterior
 guidance field reflected at the join, gave rise to two mirror-symmetrical
-barrel fields comprising $\approx 2N$ barrels, i.e., two cortical fields for
-each projection (Fig.\,1E). Together these results suggest that by the
+barrel fields comprising 2N barrels, i.e., two cortical fields for each
+thalamic projection (Fig.\,1E). Together these results suggest that by the
 misexpression of Fgf8, Shimogori \& Grove \citep{shimogori_fibroblast_2005}
-effectively created one large, rather than two distinct, cortical fields.
+effectively created one large barrel field rather than two distinct cortical
+fields.
 
-% Omit to shorten manuscript
-%\subsection*{Movie S1 caption}
-%
-%Movie corresponding to Fig.\,1C in the main paper. Simulation parameters were
-%$N=41$, $\alpha=3$, $\beta=20$, $k=3$, $D=0.2$, $\gamma\in\pm 2$,
-%$\epsilon=150$ and $\delta{t}=0.0001$. Colours indicate the thalamic
-%projection for which the connection density is maximal, black lines delineate
-%boundaries, and overlaid contours show $c>0.5$. The final frame in the movie
-%is step 25000 of the simulation.
+\section{Discussion}
 
-\section*{Discussion}
-
-These results suggest that the emergence of realistic barrel patterning
-requires that i) at each cortical location thalamocortical projections compete
-for a limited number of available synaptic connections
+The present results suggest that the key requirements for the emergence of
+realistic barrel patterning are i) at each cortical location thalamocortical
+projections compete for a limited number of available synaptic connections
 (Eqs.~\ref{eq:dc}--\ref{eq:da}), ii) at each location the branching rate of a
@@
 
-The emergence of barrels in simulation required competition between
-projections for synaptic connectivity, and also for cortical space
-(represented by $\chi$). This requirement may account for the absence of
-barrels in rodents with larger brains, such as capybara, for which competition
-for space is presumably weaker \citep{Woolsey1975}. Hence, irrespective of
-whether barrels are necessary for adaptive whisker function, the emergence of
-somatotopically ordered modular structures may be an inevitable consequence of
-local competition for cortical territory driven by input from an array of
-discrete sensory organs \citep{purves_iterated_1992}.
+The emergence of barrels in simulation required competition between thalamic
+projections in terms of synaptic connectivity \cmnt{(and an associated
+  in-group/out-group axonal identification mechanism)}, and also in terms of
+cortical space (as represented by $\chi$). This requirement may account for
+the absence of barrels in rodents with larger brains, such as capybara, for
+which competition for space is presumably weaker
+\citep{woolsey_comparative_1975}. Hence, irrespective of whether barrels are
+necessary for adaptive whisker function, the emergence of somatotopically
+ordered modular structures may be an inevitable consequence of local
+competition for cortical territory driven by input from an array of discrete
+sensory organs \citep{purves_iterated_1992}.
 
-We emphasize that the formulation of the model is entirely local; the
-simulation permits information to be communicated from a given cortical grid
-cell only to those immediately adjacent (via diffusion). Hence the simulations
-demonstrate how a self-organising system, constrained by genetically specified
-guidance cues and by the shape of the boundary, can faithfully reproduce an
-arrangement of cell aggregates in one neural structure as a topographic map in
-another.
+\cmnt{The emergence of elongated pre-barrel regions at earlier stages of the
+  simulation could be taken as a prediction of the model although where there
+  is an absence of the white contours seen in Fig.\,1C/25,000 steps the
+  barrels have not yet become well localized and would not be expected to form
+  distinct structures.}
+
+It is important to emphasize that the formulation of the model is entirely
+local, insofar as simulation requires no information to be communicated from a
+given cortical grid cell to any but those immediately adjacent (via
+diffusion). Hence the simulations demonstrate how a self-organising system,
+constrained by genetically specified guidance cues and by the shape of the
+cortical field boundary, can faithfully reproduce an arrangement of cell
+aggregates in one neural structure as a topographic map in another.
 
 Moreover, the present results confirm that somatotopic map formation does not
-require the pre-specification of center-points by as yet undetermined
+require the pre-specification of centre-points by as yet undetermined
 additional developmental mechanisms.
 
-\acknow{The authors thank Jason Berwick for advice and data used to construct
-  Fig.\,1A. This work was supported by the James S.~McDonnell Foundation
-  (grant 220020516).}
+\section{Materials \& Methods}
 
-\showacknow{} % Display the acknowledgments section
+\cmnt{It's still a short report if I add a condensed version of supp.tex
+  here.}
+
+\cmnt{Software for modelling the cortical sheet as a hexagonal lattice with
+  no-flux boundary conditions and an arbitrary boundary shape, for computing
+  the flux using a finite difference method, for numerical integration by the
+  fourth-order Runge-Kutta method, and for reproducing these results is
+  available at \url{https://github.com/ABRG-Models/BarrelEmerge}}
+
+\section{Movie S1 caption}
+
+Movie corresponding to Fig.\,1C in the main paper. Simulation parameters were
+$N=41$, $\alpha=3$, $\beta=20$, $k=3$, $D=0.2$, $\gamma\in\pm 2$,
+$\epsilon=150$ and $\delta{t}=0.0001$. Colours indicate the thalamic
+projection for which the connection density is maximal, black lines delineate
+boundaries, and overlaid contours show $c>0.5$. The final frame in the movie
+is step 25,000 of the simulation.
+
+\section{Acknowledgments}
+
+The authors thank Jason Berwick at the University of Sheffield for advice and
+for access to the rat barrel stains used to construct Fig.\,1A. This work was
+supported by a Collaborative Activity Award, \emph{Cortical Plasticity Within
+  and Across Lifetimes}, from the James S.~McDonnell Foundation (grant
+220020516).
 

### 2020-01-24 paper/supplementary/supp.tex
Work on testing out the comp2 method again, after PNAS Rev 1 suggested this.
Adds an idea about divisive normalization

--- 2019-11-28 paper/supplementary/supp.tex
@@
 \usepackage{amsmath}
+\usepackage{esint}
 
@@
 Taking inspiration from the divisive normalization idea in neural
-networks~\cite{carandini_normalization_2011}, where the overall
-activity of a group of neurons modulates the output response of an
-individual member of the group, we will use the overall branching
-level of all of the axons in a layer to modulate the response, $a_i'$,
-of the branching level in each element.
+networks~\cite{carandini_normalization_2011}, where the overall activity of a
+group of neurons modulates the output response of an individual member of the
+group, we will use the overall branching level of all of the axons in a layer
+to modulate the response, $a_i[j]'$, of the branching level in each element
+$j$ (using this notation ($a_i[j]$) makes explicit that we're talking about a
+system containing $n_e$ discrete elements indexed by $j$).
 
@@
 \begin{equation}\label{eq:norm}
-a_i' = h \; \frac {a_i^q} {\xi^q + \textstyle \sum_j^h a_{i,j}^q }
+a_i[j]' = h \; \frac {(a_i[j])^q} {\xi^q + \textstyle \sum_j^h (a_{i}[j])^q }
 \end{equation}
@@
 
-$a_i'$, the `branching response', then has to be used to drive
-connections, in practice, it is written back into $a_i$ on every
+$a_i[j]'$, the `branching response' at element $j$, then has to be used to
+drive connections, in practice, it is written back into $a_i[j]$ on every
 timestep.
@@
 \begin{equation}\label{eq:norm_comp8}
-a_i' = n_e \; \frac {a_i} {\textstyle \sum_j^{n_e} a_{i,j} }
+a_i[j]' = n_e \; \frac {a_i[j]} {\textstyle \sum_j^{n_e} a_i[j] }
 \end{equation}
@@
 element, $d_i/n_e$, so that Eq.~\ref{eq:norm_comp8} becomes
-
+%
 \begin{equation}\label{eq:norm_comp8plus}
-a_i' = d_i \; \frac {a_i} {\textstyle \sum_j^{n_e} a_{i,j} }
+a_i[j]' = d_i \; \frac {a_i[j]} {\textstyle \sum_j^{n_e} a_i[j] }
+\end{equation}
+%
+which is (making time explicit now):
+%
+\begin{equation}\label{eq:norm_comp8plus1}
+        a_i[j](t)' = \sum_j^{n_e} a_i[j](0) \; \frac {a_i[j](t)} {\textstyle \sum_j^{n_e} a_i[j](t) }
+\end{equation}
+
+One further extension is to note that in the K\&E simulation, the quantity of
+axon branching in the system is bounded and the number of connections and
+branching quantity is related. So we could normalize the `proportion of $a_i$
+which is not participating in connections'.
+%
+\begin{equation}\label{eq:norm_comp8plus2}
+        a_i[j](t)' = \sum_j^{n_e} \big[ a_i[j](0) - c_i[j](t) \big] \; \frac {a_i[j](t)} {\textstyle \sum_j^{n_e} a_i[j](t) }
+\end{equation}
+%
+The continuous equivalent, if the surface is called $S$, would be:
+%
+\begin{equation} \label{eq:norm_comp8plus3}
+  a_i(\mathbf{x}, t) = \oiint_{S} \big[  a_i(\mathbf{x}, 0) - c_i(\mathbf{x},
+  t) \big]  dS \; \frac {a'_i(\mathbf{x}, t)} {\oiint_{S}
+  a'_i(\mathbf{x}, t) dS}
 \end{equation}

### 2020-01-24 sim/rd_james_divnorm.h
Implements idea to subtract c(t) from a(0) when divisively normalizing

@@ -5,6 +5,10 @@
 
 #include "rd_james.h"
 
+//! Use the normalization method in which we account for those branches which have
+//! made connections (see supp.tex. 2020/01/24)
+#define SUBTRACT_C_FROM_A 1
+
 template <class Flt>
 class RD_James_divnorm : public RD_James<Flt>
 {
@@ -20,6 +24,13 @@ public:
      */
     alignas(vector<Flt>) vector<Flt> sum_a_init;
 
+#ifdef SUBTRACT_C_FROM_A
+    /*!
+     * An N element vector holding the sum of c_i for each TC type.
+     */
+    alignas(vector<Flt>) vector<Flt> sum_c;
+#endif
+
     /*!
      * Simple constructor; no arguments. Just calls base constructor.
      */
@@ -31,6 +42,9 @@ public:
         RD_James<Flt>::allocate();
         this->resize_vector_param (this->sum_a, this->N);
         this->resize_vector_param (this->sum_a_init, this->N);
+#ifdef SUBTRACT_C_FROM_A
+        this->resize_vector_param (this->sum_c, this->N);
+#endif
     }
 
     virtual void init (void) {
@@ -66,15 +80,28 @@ public:
             sum_tmp += this->a[_i][h];
         }
         this->sum_a[_i] = sum_tmp;
+
+#ifdef SUBTRACT_C_FROM_A
+        // Compute the sum of c[i] across the sheet.
+        this->sum_c[_i] = 0.0;
+        sum_tmp = 0.0;
+#pragma omp parallel for reduction(+:sum_tmp)
+        for (unsigned int h=0; h<this->nhex; ++h) {
+            sum_tmp += this->c[_i][h];
+        }
+        this->sum_c[_i] = sum_tmp;
+#endif
     }
 
     /*!
      * The normalization/transfer function.
      */
     virtual inline Flt transfer_a (const Flt& _a, const unsigned int _i) {
-#ifdef NORMALIZE_TO_ONE
+#if defined NORMALIZE_TO_ONE
         // Divisive normalization to one
         Flt a_rtn = this->nhex * _a / this->sum_a[_i];
+#elif defined SUBTRACT_C_FROM_A
+        Flt a_rtn = (this->sum_a_init[_i] - this->sum_c[_i]) * _a / this->sum_a[_i];
 #else
         // Divisive norm with initial sum multiplier
         Flt a_rtn = this->sum_a_init[_i] * _a / this->sum_a[_i];

### 2020-01-24 paper/briefpaper/brief.tex
Friday afternoon changes

@@
 the most distinctive pattern can be found in the barrel cortex of many rodent
-species \cite{woolsey_structural_1970}. The barrels are identifiable soon
+species \citep{woolsey_structural_1970}. The barrels are identifiable soon
 after birth in layer 4 of primary somatosensory cortex as dense clusters of
@@
 patterning reflects subcortical whisker maps comprising cell aggregates called
-barrelettes in the brainstem and barreloids in the thalamus.
-%\cite{van_der_loos_barreloids_1976}.
+barrelettes in the brainstem and barreloids in the thalamus
+\citep{ma_barrelettesarchitectonic_1991,van_der_loos_barreloids_1976}.
 
@@
 transcription factors Emx2, Pax6, Sp8, and Coup-tf1
-\citep{shimogori_fibroblast_2005}, which are expressed in gradients that mark
-orthogonal axes and can be manipulated to stretch, shrink, shift, and even
-duplicate barrels \cite{assimacopoulos_fibroblast_2012}.
-% Removed ref: ,bishop_regulation_2000 after ..Coup-tf1
+\citep{shimogori_fibroblast_2005,bishop_regulation_2000}, which are expressed
+in gradients that mark orthogonal axes and can be manipulated to stretch,
+shrink, shift, and even duplicate barrels
+\citep{assimacopoulos_fibroblast_2012}.
 
@@
 preserved in the projection of thalamocortical axons into the cortex, and that
-a barrel forms by lateral axon branching from an initial centre-point that
-ceases upon contact with axons branching from adjacent centres.  \cmnt{Another
-  sentence about center-points here.} But the assumption of pre-arranged
-centres is difficult to resolve with the observation that axons arrive in the
-cortical plate as an undifferentiated bundle, prior to barreloid formation
-\cite{agmon_organized_1993}.
+a barrel forms by lateral axon branching from an initial center-point that
+ceases upon contact with axons branching from adjacent centers.  The
+assumption of pre-arranged center-points \cmnt{(an assumption shared with
+  chemoaffinity theories following \citealp{sperry_chemoaffinity_1963})} is
+difficult to resolve with the observation that axons arrive in the cortical
+plate as an undifferentiated bundle, \emph{prior} to barreloid formation
+\citep{agmon_organized_1993}.  \cmnt{But it's not that hard to resolve, if, as
+  in the paper, we assume orthogonal gradients in the barreloid field.}
+
+% Note: The key difference between this theory and the chemospecificity theory
+% is that of group membership.
 
 Alternatively, reaction-diffusion dynamics could generate a Voronoi
-tessellation without pre-arranged centres, by amplifying characteristic modes
+tessellation without pre-arranged centers, by amplifying characteristic modes
 in a noisy initial distribution of axon branches, as a net effect of
@@
 
-The centre-point and reaction-diffusion models are not mutually
-exclusive. Pre-organised centres could bias reaction-diffusion processes to
+The center-point and reaction-diffusion models are not mutually
+exclusive. Pre-organised centers could bias reaction-diffusion processes to
 generate specific arrangements more reliably, and mechanisms of lateral axon
@@
 dynamics, under the guidance of signalling gradients, and in the absence of
-pre-defined centres.
+pre-defined centers.
 
@@
 
-Karbowski \& Ermentrout \citep{karbowski_model_2004} developed a
-reaction-diffusion style model of how extrinsic signalling gradients can
-constrain the emergence of distinct fields from intrinsic cortical
-dynamics. Their model defines how the density of connections $c(x,t)$ and axon
-branches $a(x,t)$ interact at time $t$, along a 1D anterior-posterior axis
-$x$, for $N$ thalamocortical projections indexed by $i$. The model was derived
-from the assumption that the rates at which $a_i$ and $c_i$ grow are
-reciprocally coupled. Extending the original 1D model to simulate arealization
-on a 2D cortical sheet, we use $a_i(\mb{x},t)$ and $c_i(\mb{x},t)$, and model
-synaptogenesis as
+\cite{karbowski_model_2004} developed a reaction-diffusion style model of how
+extrinsic signalling gradients can constrain the emergence of distinct fields
+from intrinsic cortical dynamics. Their model defines how the density of
+connections $c(x,t)$ and axon branches $a(x,t)$ interact at time $t$, along a
+1D anterior-posterior axis $x$, for $N$ thalamocortical projections indexed by
+$i$. The model was derived from the assumption that the rates at which $a_i$
+and $c_i$ grow are reciprocally coupled. Extending the original 1D model to
+simulate arealization on a 2D cortical sheet, we use $a_i(\mb{x},t)$ and
+$c_i(\mb{x},t)$, and model synaptogenesis as
 %
@@
   by \cite{zheng_signal_2001}, with black lines to delineate barrels and to
-  measure departure (Honda-$\delta$; see \cite{senft_mouse_1991}) from a
+  measure departure (Honda-$\delta$; see \citealp{senft_mouse_1991}) from a
   perfect Voronoi tesselation. Right shows the initial distribution of axon
@@
   the locations of putative barreloid centers (reconstructed from
-  \cite{haidarliu_size_2001}). \textbf{C} Simulation results for parameters
+  \citealp{haidarliu_size_2001}). \textbf{C} Simulation results for parameters
   $N=41$, $\alpha=3$, $\beta=20$, $k=3$, $D=0.2$, $\gamma\in\pm 2$,
@@
   in a Fgf8 misexpression experiment
-  (c.f. \cite{assimacopoulos_fibroblast_2012}), simulated by reflecting
+  (c.f. \citealp{assimacopoulos_fibroblast_2012}), simulated by reflecting
   $\rho_1$ at the join ($\epsilon=80$). All scale bars 1mm.}
@@
 projections are dense. Divisive normalization keeps the axon branch density
-bounded on each iteration to the initial density at $t=0$;
+bounded on each iteration to the initial density at $t=0$ \cmnt{(less the
+  number of connections that have been made).};
 %
 \begin{equation} \label{eq:norm}
-  a_i(\mathbf{x}, t) = \oiint_{S}  a_i(\mathbf{x}, 0) \dif S \; \frac {a'_i(\mathbf{x}, t)} {\oiint_{S}
+  a_i(\mathbf{x}, t) = \oiint_{S} \big[ a_i(\mathbf{x}, 0) - c_i(\mathbf{x}, t) \big] \dif S \; \frac {a'_i(\mathbf{x}, t)} {\oiint_{S}
   a'_i(\mathbf{x}, t) \dif S}
@@
 which 41 pairs of $\gamma$ values were sampled, at the coordinates of 41
-barreloid centers recreated from \cite{haidarliu_size_2001} (Fig.\,1B).
+barreloid centers recreated from \citealp{haidarliu_size_2001} (Fig.\,1B).
 
@@
 (Fig.\,1C; see Movie S1). A reduction in the Honda delta metric (see
-\cite{senft_mouse_1991}) confirmed that a `good' Voronoi pattern emerged
+\citealp{senft_mouse_1991}) confirmed that a `good' Voronoi pattern emerged
 within $\approx 3000$ iterations (see Fig.\,1D). A measure of the difference
@@
 simulated a seminal barrel duplication experiment
-\cite{shimogori_fibroblast_2005}, first by solving on a cortical domain
+\citep{shimogori_fibroblast_2005}, first by solving on a cortical domain
 comprising two separate, identically shaped cortical boundaries. In this case
@@
 thalamic projection (Fig.\,1E). Together these results suggest that by the
-misexpression of Fgf8, Shimogori \& Grove \citep{shimogori_fibroblast_2005}
+misexpression of Fgf8, \cite{shimogori_fibroblast_2005}
 effectively created one large barrel field rather than two distinct cortical
@@
 Moreover, the present results confirm that somatotopic map formation does not
-require the pre-specification of centre-points by as yet undetermined
+require the pre-specification of center-points by as yet undetermined
 additional developmental mechanisms.

### 2020-01-24 paper/supplementary/supp.tex
Friday afternoon changes

@@
 \begin{equation} \label{eq:norm_comp8plus3}
-  a_i(\mathbf{x}, t) = \oiint_{S} \big[  a_i(\mathbf{x}, 0) - c_i(\mathbf{x},
+\begin{split}
+  a_i(\mathbf{x}, t) & = \oiint_{S} \big[  a_i(\mathbf{x}, 0) - c_i(\mathbf{x},
   t) \big]  dS \; \frac {a'_i(\mathbf{x}, t)} {\oiint_{S}
+  a'_i(\mathbf{x}, t) dS} \\
+  & = \bigg[ \oiint_{S} a_i(\mathbf{x}, 0) dS - \oiint_{S} c_i(\mathbf{x}, t)
+  dS  \bigg] \; \frac {a'_i(\mathbf{x}, t)} {\oiint_{S}
   a'_i(\mathbf{x}, t) dS}
-\end{equation}
+\end{split}
+\end{equation}
+which is a little more expensive as $\oiint_{S} c_i(\mathbf{x}, t) dS$ has to
+be computed on each iteration.
 

### 2020-01-27 paper/briefpaper/brief.tex
Adds methods section covering computation of divJ

@@
 \newcommand*\vcdot@[2]{\mathbin{\vcenter{\hbox{\scalebox{#2}{$\m@th#1\bullet$}}}}}
+\newcommand{\code}[1]{\textsf{#1}}
 
@@
 
-\cmnt{It's still a short report if I add a condensed version of supp.tex
-  here.}
+\cmnt{I've left this section in black, although it is all new, cf the PNAS
+  draft.}
+% Be sure to include/mention:
+%
+% The no-flux boundary conditions applied in the \cite{karbowski_model_2004}
+% model are retained.
+%
+% Arbitrary boundary shapes were applied
+%
+% computing the flux using a finite difference method,
+%
+% numerical integration by the fourth-order Runge-Kutta method,
 
-\cmnt{Software for modelling the cortical sheet as a hexagonal lattice with
-  no-flux boundary conditions and an arbitrary boundary shape, for computing
-  the flux using a finite difference method, for numerical integration by the
-  fourth-order Runge-Kutta method, and for reproducing these results is
-  available at \url{https://github.com/ABRG-Models/BarrelEmerge}}
+In these simulations, the cortex was modelled as a two dimensional lattice in
+which each element represented a small area of the cortical sheet. We chose to
+use hexagonal elements to simplify the computation of the 2D Laplacian, a
+spatial operator used in reaction diffusion systems. Within a boundary traced
+from rat primary barrel cortex (Fig.\,1A) we set the hex-to-hex distance
+$d=0.03$\,mm, which resulted in a lattice containing 6515 hexes for the
+simulations shown in Figs.\,1A,C \& D and 12739 hexes for the Fgf8
+misexpression study shown in Fig.\,1E. Each hex contained 82 time-dependent
+variables: 41 branching densities ($a_i$) and 41 connection densities
+($c_i$). Each hex also contained 82 time-independent variables; the $x$ and
+$y$ components of the sum of the product of the interaction paramters with the
+two molecular signalling gradients.
+
+The rates of change of each of the time-dependent variables (Eqs.\,\ref{eq:dc}
+\& \ref{eq:da}) was computed using a fourth-order Runge-Kutta method. The most
+involved part of this computation was finding the divergence of the flux of
+axonal branching, $\mb{J}_i(\mb{x},t)$---the term in parentheses in
+Eq.\,\ref{eq:da}:
+%
+\begin{equation}
+  \label{eq:divJ}
+  \nabla\vcdot\mb{J}_i(\mb{x},t) \equiv \nabla\vcdot\left(D \nabla a_i-a_i\sum_{j=1}^{M} \gamma_{i,j}\nabla \rho_j(\mb{x})\right)
+\end{equation}
+%
+We can define the time-independent sum of the guidance gradients
+$\mb{g}_i(\mb{x}) \equiv \sum_{j=1}^{M} \gamma_{i,j} \nabla \rho_j(\mb{x})$
+and then apply vector calculus identities to compute this term. Because
+the divergence operator is distributive, Eq.~\ref{eq:divJ} can be expanded
+(dropping all explicit references to space or time dependence):
+%
+\begin{equation}
+\nabla\vcdot\mb{J}_i = \nabla\vcdot\big(D \nabla a_i\big) - \nabla\vcdot\big(a_i \mb{g}_i\big)
+\end{equation}
+%
+Because $D$ is a constant and by applying the vector calculus product rule
+identity to $\nabla\vcdot\big(a_i \mb{g}_i\big)$ we obtain:
+%
+%%\begin{equation}
+%%\nabla\vcdot\mb{J}_i = D \dvrg a_i
+%%-
+%%\big(
+%%a_i\nabla\vcdot\mb{g}_i
+%%+
+%%\mb{g}_i\vcdot\nabla a_i
+%%\big)
+%%\end{equation}
+%
+%%(to prove $\nabla\vcdot D\nabla a = D \dvrg a$, apply the
+%%product rule and note that $\nabla D = 0$). Expanding
+%%the brackets we have:
+%
+\begin{equation} \label{eq:divJExpanded}
+%
+\nabla\vcdot\mb{J}_i =
+%
+D \dvrg a_i % term1
+-
+a_i\nabla\vcdot\mb{g}_i % term2
+-
+\mb{g}_i\vcdot\nabla a_i % term3
+%
+\end{equation}
+%
+which has three elements to compute: i)~$D \dvrg a_i$ (the Laplacian of
+$a_i(\mb{x},t)$); ii)~a time-independent modulator of $a_i(\mb{x},t)$ (because
+$\nabla\vcdot\mb{g}_i(\mb{x})$ is a time-independent static field); and iii)~the
+scalar product of the static vector field $\mb{g}_i$ and the gradient of
+$a_i$.
+
+Each of the divergences can be simplified by means of Gauss's Theorem
+following \cite{lee_hexagonal_2014}. The computation of the mean value of the
+Laplacian across the area of one hexagon of area $\Omega$ located at position
+$\mb{p}_0$, with neighbours at positions $\mb{p}_1$--$\mb{p}_6$ is:
+%
+%%\begin{equation} \label{eq:computeDivJTerm1}
+%%\boxed{
+%%D \dvrg a_i(\mb{p}_0,t) = D\frac{2}{3d^2} \sum_{j=1}^{6}\big(a_i(\mb{p}_j) - a_i(\mb{p}_0)\big)
+%%}
+%%\end{equation}
+
+\begin{equation}
+\begin{split}
+\langle D \dvrg a_i(\mb{p}_0,t) \rangle  = \frac{1}{\Omega} \iint_A \dvrg a_i(\mb{x}, t) dxdy & = \frac{1}{\Omega} \oint_c \frac{\partial a_i}{\partial \hat{\mb{n}}} d\gamma \\
+& \approx \frac{1}{\Omega} \sum_{j=1}^6 \frac{\partial a_i(\mb{p}_j)}{\partial \hat{\mb{n}}} \bigg\rvert_{mid} v \\
+& = \frac{2}{\sqrt{3} d^2} \sum_{j=1}^6 \frac{a_i(\mb{p}_j) - a_i(\mb{p}_0)}{d} \frac{d}{\sqrt{3}} \\
+& = \frac{2}{3 d^2} \sum_{j=1}^6 \big(a_i(\mb{p}_j) - a_i(\mb{p}_0)\big)
+\end{split}
+\end{equation}
+
+where $d$ is the centre-to-centre distance between hexes in the grid, $v =
+d/\sqrt{3}$ is the length of each of the 6 edges of the hexagon's perimeter
+and $d\gamma$ is an infinitessimally small distance along the perimeter of the
+hexagon. The area, $\Omega$, of each hexagon in the lattice is
+$\frac{\sqrt{3}}{2}d^2$.
+
+The computation of the second term in Eq.\,\ref{eq:divJExpanded},
+$\langle a_i(\mb{p}_0,t)\nabla\vcdot\mb{g}(\mb{p}_0)\rangle$, can be written out similarly:
+
+\begin{equation} \label{eq:divg}
+\begin{split}
+%
+\frac{1}{\Omega} \iint_A a_i\nabla\vcdot\mb{g}\;dxdy & = \frac{a_i(\mb{p}_0)}{\Omega}  \oint_c \mb{g}\vcdot d\hat{\mathbf{n}} \\
+%
+& \approx \frac{a_i(\mb{p}_0,t)}{\Omega} \sum_{j=1}^{6} \frac{\mb{g}_j + \mb{g}_0}{2}\vcdot \hat{\mb{n}}\;v \\
+%
+& = \frac{2 a_i(\mb{p}_0,t)v}{\sqrt{3}d^2} \bigg( \sum_{j=1}^{6} \frac{\mb{g}_j^x + \mb{g}_0^x}{2} \vcdot  \hat{\mb{n}} + \sum_{j=1}^{6} \frac{\mb{g}_j^y + \mb{g}_0^y}{2} \vcdot  \hat{\mb{n}} \bigg) \\
+%
+%%& = \frac{2 a_i(\mb{p}_0,t)d}{\sqrt{3}d^2\sqrt{3}} \bigg( \sum_{j=1}^{6} \frac{\mb{g}_j^x + \mb{g}_0^x}{2} \vcdot  \hat{\mb{n}} + \sum_{j=1}^{6} \frac{\mb{g}_j^y + \mb{g}_0^y}{2} \vcdot  \hat{\mb{n}} \bigg) \\
+%
+\Rightarrow \langle a_i(\mb{p}_0,t)\nabla\vcdot\mb{g}(\mb{p}_0)\rangle & = \frac{a_i(\mb{p}_0,t)}{3d} \bigg( \sum_{j=1}^{6} \big(\mb{g}_j^x + \mb{g}_0^x\big) \cos (60\times(j-1)) + \sum_{j=1}^{6} \big({\mb{g}_j^y + \mb{g}_0^y}\big) \sin (60\times(j-1)) \bigg) \\
+%
+\end{split}
+\end{equation}
+
+The final expression is suitable for a numerical computation and can
+be found in \code{compute\_divJ()} as \code{term2}.
+
+The final term in Eq.~\ref{eq:divJExpanded} is the scalar product of two
+vector fields which is straighforward to compute from their Cartesian
+components.
+
+The advantage of separating this computation out is that we can ensure that,
+at the boundary, $\mb{J}$ resulting from the first term remains 0, by the
+ghost cell method, then $\mb{g}_i(\mb{x})$ can be tailored so that it, and its
+normal derivative approach 0 at the boundary, ensuring that the second and
+third terms of Eq.~\ref{eq:divJExpanded} also contribute nothing to
+$\mb{J}$. This is achieved by applying a fairly sharp logistic function based
+on distance to the boundary to $\mb{g}(\mb{x})$.
+
+All code required to reproduce these results is available at
+\url{https://github.com/ABRG-Models/BarrelEmerge}
 

### 2020-01-27 paper/briefpaper/brief.tex
Changes following meeting with Stuart

--- 2020-01-24 paper/briefpaper/brief.tex
@@
 ceases upon contact with axons branching from adjacent centers.  The
-assumption of pre-arranged center-points \cmnt{(an assumption shared with
-  chemoaffinity theories following \citealp{sperry_chemoaffinity_1963})} is
-difficult to resolve with the observation that axons arrive in the cortical
-plate as an undifferentiated bundle, \emph{prior} to barreloid formation
-\citep{agmon_organized_1993}.  \cmnt{But it's not that hard to resolve, if, as
-  in the paper, we assume orthogonal gradients in the barreloid field.}
+assumption of pre-arranged center-points is difficult to resolve with the
+observation that axons arrive in the cortical plate as an undifferentiated
+bundle, \emph{prior} to barreloid formation \citep{agmon_organized_1993}.
 
@@
 may be positive or negative in order that axons may branch in the direction of
-either higher or lower concentrations. The second term on the right represents
-an assumption that axon branching decreases when synaptogenesis increases, and
-vice versa, i.e., \cmnt{when a connection is made, the axon branching
-  decreases by one branch and when a connection is lost, one branch is added
-  to the branching density}
-%that axon branches are required for synaptic connections to form.
-Here $\chi_i=0$ is a placeholder.
+either higher or lower concentrations. \cmnt{The second term on the right
+  quantifies the coupling between axon branching and synaptogenesis}. Here $\chi_i=0$ is a placeholder.
 
@@
 projections are dense. Divisive normalization keeps the axon branch density
-bounded on each iteration to the initial density at $t=0$ \cmnt{(less the
-  number of connections that have been made).};
+bounded on each iteration to the initial density \cmnt{of unconnected
+  branches} at $t=0$:
 %
@@
 where prime symbols indicate the use of intermediate values computed from
-Eqs.~\ref{eq:dc}--\ref{eq:comp} at time $t-{\delta}t$. \cmnt{Normalization
-  corresponds to the biological constraint that the branching and growth of
-  thalamocortical axons is backed by a finite resource provided by their
-  parent cell bodies.}
+Eqs.~\ref{eq:dc}--\ref{eq:comp} at time $t-{\delta}t$. \cmnt{Note that this
+  operation is local to individual afferent projections.}
 
@@
 The emergence of barrels in simulation required competition between thalamic
-projections in terms of synaptic connectivity \cmnt{(and an associated
-  in-group/out-group axonal identification mechanism)}, and also in terms of
-cortical space (as represented by $\chi$). This requirement may account for
-the absence of barrels in rodents with larger brains, such as capybara, for
-which competition for space is presumably weaker
-\citep{woolsey_comparative_1975}. Hence, irrespective of whether barrels are
-necessary for adaptive whisker function, the emergence of somatotopically
-ordered modular structures may be an inevitable consequence of local
-competition for cortical territory driven by input from an array of discrete
-sensory organs \citep{purves_iterated_1992}.
-
-\cmnt{The emergence of elongated pre-barrel regions at earlier stages of the
-  simulation could be taken as a prediction of the model although where there
-  is an absence of the white contours seen in Fig.\,1C/25,000 steps the
-  barrels have not yet become well localized and would not be expected to form
-  distinct structures.}
+projections in terms of synaptic connectivity \cmnt{and an implicit self/other
+  identifier amoungst projections}, and also in terms of cortical space (as
+represented by $\chi$). This requirement may account for the absence of
+barrels in rodents with larger brains, such as capybara, for which competition
+for space is presumably weaker \citep{woolsey_comparative_1975}. Hence,
+irrespective of whether barrels are necessary for adaptive whisker function,
+the emergence of somatotopically ordered modular structures may be an
+inevitable consequence of local competition for cortical territory driven by
+input from an array of discrete sensory organs \citep{purves_iterated_1992}.
 
@@
 
-\cmnt{It's still a short report if I add a condensed version of supp.tex
-  here.}
+% Be sure to include/mention:
+%
+% The no-flux boundary conditions applied in the \cite{karbowski_model_2004}
+% model are retained.
+%
+% Arbitrary boundary shapes were applied
 
-\cmnt{Software for modelling the cortical sheet as a hexagonal lattice with
-  no-flux boundary conditions and an arbitrary boundary shape, for computing
-  the flux using a finite difference method, for numerical integration by the
-  fourth-order Runge-Kutta method, and for reproducing these results is
-  available at \url{https://github.com/ABRG-Models/BarrelEmerge}}
+The cortex was modelled as a two dimensional hexagonal lattice which
+simplifies the computation of the 2D Laplacian. Within a boundary traced from
+rat primary barrel cortex (Fig.\,1A) we set the hex-to-hex distance
+$d=0.03$\,mm, which resulted in a lattice containing 6515 hexes for the
+simulations shown in Figs.\,1A,C \& D and 12739 hexes for the Fgf8
+misexpression study shown in Fig.\,1E. Each hex contained 82 time-dependent
+variables: 41 branching densities ($a_i$) and 41 connection densities ($c_i$).
+The rate of change of each of the time-dependent variables (Eqs.\,\ref{eq:dc}
+\& \ref{eq:da}) was computed using a fourth-order Runge-Kutta method.
+
+The most involved part of this computation is to find the divergence of the
+flux of axonal branching, $\mb{J}_i(\mb{x},t)$, the term in parentheses in
+Eq.\,\ref{eq:da}:
+%
+\begin{equation}
+  \label{eq:divJ}
+  \nabla\vcdot\mb{J}_i(\mb{x},t) \equiv \nabla\vcdot\left(D \nabla a_i-a_i\sum_{j=1}^{M} \gamma_{i,j}\nabla \rho_j(\mb{x})\right).
+\end{equation}
+%
+We can define the time-independent sum of the guidance gradients
+$\mb{g}_i(\mb{x}) \equiv \sum_{j=1}^{M} \gamma_{i,j} \nabla \rho_j(\mb{x})$
+and then apply vector calculus identities to compute
+$\nabla\vcdot\mb{J}_i$. Because the divergence operator is distributive,
+Eq.~\ref{eq:divJ} can be expanded (dropping references to $\mb{x}$ and $t$ for
+clarity):
+%
+\begin{equation}
+\nabla\vcdot\mb{J}_i = \nabla\vcdot\big(D \nabla a_i\big) - \nabla\vcdot\big(a_i \mb{g}_i\big).
+\end{equation}
+%
+Applying the vector calculus product rule identity yields
+%
+\begin{equation} \label{eq:divJExpanded}
+%
+\nabla\vcdot\mb{J}_i =
+%
+D \dvrg a_i % term1
+-
+a_i\nabla\vcdot\mb{g}_i % term2
+-
+\mb{g}_i\vcdot\nabla a_i % term3
+,
+\end{equation}
+%
+which has three elements to compute: i)~$D \dvrg a_i$ (the Laplacian of
+$a_i(\mb{x},t)$); ii)~a time-independent modulator of $a_i(\mb{x},t)$ (because
+$\nabla\vcdot\mb{g}_i(\mb{x})$ is a time-independent static field); and iii)~the
+scalar product of the static vector field $\mb{g}_i$ and the gradient of
+$a_i$. Each of the divergences can be simplified by means of Gauss's Theorem
+following \cite{lee_hexagonal_2014}.
+
+(i) The computation of the mean value of the Laplacian across the area of one
+hexagon of area $\Omega$ located at position $\mb{p}_0$, with neighbours at
+positions $\mb{p}_1$--$\mb{p}_6$ is
+%
+\begin{equation} \label{eq:lapl}
+\begin{split}
+\langle D \dvrg a_i(\mb{p}_0,t) \rangle  = \frac{1}{\Omega} \iint_A \dvrg a_i(\mb{x}, t) dxdy & = \frac{1}{\Omega} \oint_c \frac{\partial a_i}{\partial \hat{\mb{n}}} d\gamma \\
+& \approx \frac{1}{\Omega} \sum_{j=1}^6 \frac{\partial a_i(\mb{p}_j)}{\partial \hat{\mb{n}}} \bigg\rvert_{mid} v \\
+& = \frac{2}{\sqrt{3} d^2} \sum_{j=1}^6 \frac{a_i(\mb{p}_j) - a_i(\mb{p}_0)}{d} \frac{d}{\sqrt{3}} \\
+& = \frac{2}{3 d^2} \sum_{j=1}^6 \big(a_i(\mb{p}_j) - a_i(\mb{p}_0)\big)
+\end{split}
+\end{equation}
+%
+where $d$ is the centre-to-centre distance between hexes in the grid, $v =
+d/\sqrt{3}$ is the length of each of the 6 edges of the hexagon's perimeter
+and $d\gamma$ is an infinitessimally small distance along the perimeter of the
+hexagon. The area, $\Omega$, of each hexagon in the lattice is
+$\frac{\sqrt{3}}{2}d^2$.
+
+ii) The computation of the second term in Eq.\,\ref{eq:divJExpanded},
+$\langle a_i(\mb{p}_0,t)\nabla\vcdot\mb{g}(\mb{p}_0)\rangle$, can be written out similarly:
+
+\begin{equation} \label{eq:divg}
+\begin{split}
+%
+\frac{1}{\Omega} \iint_A a_i\nabla\vcdot\mb{g}\;dxdy & = \frac{a_i(\mb{p}_0)}{\Omega}  \oint_c \mb{g}\vcdot d\hat{\mathbf{n}} \\
+%
+& \approx \frac{a_i(\mb{p}_0,t)}{\Omega} \sum_{j=1}^{6} \frac{\mb{g}_j + \mb{g}_0}{2}\vcdot \hat{\mb{n}}\;v \\
+%
+& = \frac{2 a_i(\mb{p}_0,t)v}{\sqrt{3}d^2} \bigg( \sum_{j=1}^{6} \frac{\mb{g}_j^x + \mb{g}_0^x}{2} \vcdot  \hat{\mb{n}} + \sum_{j=1}^{6} \frac{\mb{g}_j^y + \mb{g}_0^y}{2} \vcdot  \hat{\mb{n}} \bigg) \\
+%
+%%& = \frac{2 a_i(\mb{p}_0,t)d}{\sqrt{3}d^2\sqrt{3}} \bigg( \sum_{j=1}^{6} \frac{\mb{g}_j^x + \mb{g}_0^x}{2} \vcdot  \hat{\mb{n}} + \sum_{j=1}^{6} \frac{\mb{g}_j^y + \mb{g}_0^y}{2} \vcdot  \hat{\mb{n}} \bigg) \\
+%
+\Rightarrow \langle a_i(\mb{p}_0,t)\nabla\vcdot\mb{g}(\mb{p}_0)\rangle & \approx \frac{a_i(\mb{p}_0,t)}{3d} \bigg( \sum_{j=1}^{6} \big(\mb{g}_j^x + \mb{g}_0^x\big) \cos \big(\frac{\pi}{3}(j-1)\big) + \sum_{j=1}^{6} \big({\mb{g}_j^y + \mb{g}_0^y}\big) \sin \big(\frac{\pi}{3}(j-1)\big) \bigg).
+\\
+%
+\end{split}
+\end{equation}
+%
+Both this last expression, and the final expression of Eq.\,\ref{eq:lapl} are
+easily computed, comprising short sums of nearest-neighbour relationships.
+
+iii) The final term in Eq.~\ref{eq:divJExpanded} is the scalar product of two
+vector fields which is straightforward to compute from their Cartesian
+components.
+
+The advantage of separating the computation of Eq.~\ref{eq:divJ} in (i), (ii)
+\& (iii) is that we can ensure that the no-flux boundary condition is
+fulfilled. On the boundary, the contribution to $\mb{J}$ resulting from the
+first term of Eq.~\ref{eq:divJExpanded} can be held at 0 by the ghost cell
+method (in which a hex outside the boundary containing the same value as the
+hex inside the boundary is imagined such that the flux of $J$ across the
+boundary is 0). Then, $\mb{g}_i(\mb{x})$ can be tailored so that it, and its
+normal derivative approach 0 at the boundary, ensuring that the second and
+third terms of Eq.~\ref{eq:divJExpanded} also contribute nothing to
+$\mb{J}$. This is achieved by applying to $\mb{g}(\mb{x})$ a sharp logistic
+function of the distance from $\mb{x}$ to the boundary.
+
+All code required to reproduce these results is available at
+\url{https://github.com/ABRG-Models/BarrelEmerge}. The computations described
+in (i), (ii) and (iii) may be found in the code \code{compute\_divJ()} as
+\code{term1}, \code{term2} and \code{term3}, respectively.
 

### 2020-01-27 paper/briefpaper/brief.tex
A few more tweaks

@@
 
-Software for modelling the cortical sheet as a hexagonal lattice with no-flux
-boundary conditions and an arbitrary boundary shape, for computing the flux
-using a finite difference method, for numerical integration by the
-fourth-order Runge-Kutta method, and for reproducing these results is
-available at \url{https://github.com/ABRG-Models/BarrelEmerge}.
-
 \section{Results}
@@
 
-The term in parentheses in Eq.~\ref{eq:dc} represents competition between
+The term in parentheses in Eq.\,\ref{eq:dc} represents competition between
 thalamocortical projections for a limited availability of cortical
@@
 given projection is reduced by the density of other projections
-(Eq.~\ref{eq:comp}), and iii) the branch density of each projection is
-conserved over time (Eq.~\ref{eq:norm}).
+(Eq.\,\ref{eq:comp}), and iii) the branch density of each projection is
+conserved over time (Eq.\,\ref{eq:norm}).
 
@@
 rat primary barrel cortex (Fig.\,1A) we set the hex-to-hex distance
-$d=0.03$\,mm, which resulted in a lattice containing 6515 hexes for the
+$d$ to 0.03\,mm, which resulted in a lattice containing 6515 hexes for the
 simulations shown in Figs.\,1A,C \& D and 12739 hexes for the Fgf8
@@
   \label{eq:divJ}
-  \nabla\vcdot\mb{J}_i(\mb{x},t) \equiv \nabla\vcdot\left(D \nabla a_i-a_i\sum_{j=1}^{M} \gamma_{i,j}\nabla \rho_j(\mb{x})\right).
+  \nabla\vcdot\mb{J}_i(\mb{x},t) = \nabla\vcdot\left(D \nabla a_i-a_i\sum_{j=1}^{M} \gamma_{i,j}\nabla \rho_j(\mb{x})\right).
 \end{equation}
 %
-We can define the time-independent sum of the guidance gradients
-$\mb{g}_i(\mb{x}) \equiv \sum_{j=1}^{M} \gamma_{i,j} \nabla \rho_j(\mb{x})$
-and then apply vector calculus identities to compute
+We can note that the sum of the guidance gradients is time-independent and
+define $\mb{g}_i(\mb{x}) \equiv \sum_{j=1}^{M} \gamma_{i,j} \nabla
+\rho_j(\mb{x})$ and then apply vector calculus identities to compute
 $\nabla\vcdot\mb{J}_i$. Because the divergence operator is distributive,
-Eq.~\ref{eq:divJ} can be expanded (dropping references to $\mb{x}$ and $t$ for
-clarity):
+Eq.\,\ref{eq:divJ} can be expanded (dropping references to $\mb{x}$ and $t$
+for clarity):
 %
@@
 which has three elements to compute: i)~$D \dvrg a_i$ (the Laplacian of
-$a_i(\mb{x},t)$); ii)~a time-independent modulator of $a_i(\mb{x},t)$ (because
-$\nabla\vcdot\mb{g}_i(\mb{x})$ is a time-independent static field); and iii)~the
+$a_i$); ii)~a time-independent modulator of $a_i$ (because
+$\nabla\vcdot\mb{g}_i$ is a time-independent static field); and iii)~the
 scalar product of the static vector field $\mb{g}_i$ and the gradient of
@@
 
-(i) The computation of the mean value of the Laplacian across the area of one
-hexagon of area $\Omega$ located at position $\mb{p}_0$, with neighbours at
-positions $\mb{p}_1$--$\mb{p}_6$ is
+(i) The computation of the mean value of the Laplacian across one hexagon of
+area $\Omega = \frac{\sqrt{3}}{2}d^2$, located at position $\mb{p}_0$, with
+neighbours at positions $\mb{p}_1$--$\mb{p}_6$ is
 %
@@
 \begin{split}
-\langle D \dvrg a_i(\mb{p}_0,t) \rangle  = \frac{1}{\Omega} \iint_A \dvrg a_i(\mb{x}, t) dxdy & = \frac{1}{\Omega} \oint_c \frac{\partial a_i}{\partial \hat{\mb{n}}} d\gamma \\
+\langle D \dvrg a_i(\mb{p}_0,t) \rangle  = \frac{1}{\Omega} \iint_\Omega \dvrg a_i(\mb{x}, t) dxdy & = \frac{1}{\Omega} \oint_c \frac{\partial a_i}{\partial \hat{\mb{n}}} d\gamma \\
 & \approx \frac{1}{\Omega} \sum_{j=1}^6 \frac{\partial a_i(\mb{p}_j)}{\partial \hat{\mb{n}}} \bigg\rvert_{mid} v \\
@@
 and $d\gamma$ is an infinitessimally small distance along the perimeter of the
-hexagon. The area, $\Omega$, of each hexagon in the lattice is
-$\frac{\sqrt{3}}{2}d^2$.
+hexagon.
 
@@
 %
-\frac{1}{\Omega} \iint_A a_i\nabla\vcdot\mb{g}\;dxdy & = \frac{a_i(\mb{p}_0)}{\Omega}  \oint_c \mb{g}\vcdot d\hat{\mathbf{n}} \\
+\frac{1}{\Omega} \iint_\Omega a_i\nabla\vcdot\mb{g}\;dxdy & = \frac{a_i(\mb{p}_0,t)}{\Omega}  \oint_c \mb{g}\vcdot d\hat{\mathbf{n}} \\
 %
@@
 
-iii) The final term in Eq.~\ref{eq:divJExpanded} is the scalar product of two
+iii) The final term in Eq.\,\ref{eq:divJExpanded} is the scalar product of two
 vector fields which is straightforward to compute from their Cartesian
@@
 
-The advantage of separating the computation of Eq.~\ref{eq:divJ} in (i), (ii)
-\& (iii) is that we can ensure that the no-flux boundary condition is
-fulfilled. On the boundary, the contribution to $\mb{J}$ resulting from the
-first term of Eq.~\ref{eq:divJExpanded} can be held at 0 by the ghost cell
-method (in which a hex outside the boundary containing the same value as the
-hex inside the boundary is imagined such that the flux of $J$ across the
-boundary is 0). Then, $\mb{g}_i(\mb{x})$ can be tailored so that it, and its
-normal derivative approach 0 at the boundary, ensuring that the second and
-third terms of Eq.~\ref{eq:divJExpanded} also contribute nothing to
-$\mb{J}$. This is achieved by applying to $\mb{g}(\mb{x})$ a sharp logistic
-function of the distance from $\mb{x}$ to the boundary.
+By separating the computation of Eq.\,\ref{eq:divJ} into parts (i), (ii) \&
+(iii), we can ensure that the no-flux boundary condition,
+$\mb{J}_i(\mb{x},t)\big\rvert_{boundary} = 0$, is fulfilled. On the boundary,
+the contribution to $\mb{J}$ resulting from the first term of
+Eq.\,\ref{eq:divJExpanded} can be fixed to 0 by the `ghost cell method' in
+which, during the evaluation of (i), a hex outside the boundary containing the
+same value as the hex inside the boundary is imagined to exist such that the
+flux of $\mb{J}$ across the boundary is 0. Then, $\mb{g}_i(\mb{x})$ can be
+tailored so that it, and its normal derivative approach 0 at the boundary,
+ensuring that the second and third terms of Eq.\,\ref{eq:divJExpanded} also
+contribute nothing to $\mb{J}$. This is achieved by applying to
+$\mb{g}_i(\mb{x})$ a sharp logistic function of the distance from $\mb{x}$ to
+the boundary.
 
 All code required to reproduce these results is available at
-\url{https://github.com/ABRG-Models/BarrelEmerge}. The computations described
-in (i), (ii) and (iii) may be found in the code \code{compute\_divJ()} as
-\code{term1}, \code{term2} and \code{term3}, respectively.
+\url{https://github.com/ABRG-Models/BarrelEmerge/tree/eLife_submission1}. The
+computations described in (i), (ii) and (iii) may be found in the class method
+\code{RD\_James::compute\_divJ()} as \code{term1}, \code{term2} and \code{term3},
+respectively.
 

### 2020-01-28 paper/briefpaper/brief.tex
a coupld of commas

@@
 bounded on each iteration to the initial density \cmnt{of unconnected
-  branches} at $t=0$:
+  branches} at $t=0$
 %
@@
   a_i(\mathbf{x}, t) = \oiint_{S} \big[ a_i(\mathbf{x}, 0) - c_i(\mathbf{x}, t) \big] \dif S \; \frac {a'_i(\mathbf{x}, t)} {\oiint_{S}
-  a'_i(\mathbf{x}, t) \dif S}
+  a'_i(\mathbf{x}, t) \dif S}\,,
 \end{equation}
@@
 %
-We can note that the sum of the guidance gradients is time-independent and
-define $\mb{g}_i(\mb{x}) \equiv \sum_{j=1}^{M} \gamma_{i,j} \nabla
-\rho_j(\mb{x})$ and then apply vector calculus identities to compute
-$\nabla\vcdot\mb{J}_i$. Because the divergence operator is distributive,
-Eq.\,\ref{eq:divJ} can be expanded (dropping references to $\mb{x}$ and $t$
-for clarity):
+Note that the sum of the guidance gradients is time-independent and define
+$\mb{g}_i(\mb{x}) \equiv \sum_{j=1}^{M} \gamma_{i,j} \nabla\rho_j(\mb{x})$.
+Because the divergence operator is distributive, Eq.\,\ref{eq:divJ} can be
+expanded using vector calculus identities (dropping references to $\mb{x}$ and
+$t$ for clarity):
 %
@@
 & = \frac{2}{\sqrt{3} d^2} \sum_{j=1}^6 \frac{a_i(\mb{p}_j) - a_i(\mb{p}_0)}{d} \frac{d}{\sqrt{3}} \\
-& = \frac{2}{3 d^2} \sum_{j=1}^6 \big(a_i(\mb{p}_j) - a_i(\mb{p}_0)\big)
+& = \frac{2}{3 d^2} \sum_{j=1}^6 \big(a_i(\mb{p}_j) - a_i(\mb{p}_0)\big),
 \end{split}
@@
 By separating the computation of Eq.\,\ref{eq:divJ} into parts (i), (ii) \&
-(iii), we can ensure that the no-flux boundary condition,
-$\mb{J}_i(\mb{x},t)\big\rvert_{boundary} = 0$, is fulfilled. On the boundary,
-the contribution to $\mb{J}$ resulting from the first term of
+(iii), the no-flux boundary condition,
+$\mb{J}_i(\mb{x},t)\big\rvert_{\mathrm{boundary}} = 0$, can be fulfilled. On
+the boundary, the contribution to $\mb{J}$ resulting from the first term of
 Eq.\,\ref{eq:divJExpanded} can be fixed to 0 by the `ghost cell method' in

### 2020-01-28 paper/briefpaper/brief.tex
Last changes to text

@@
 a barrel forms by lateral axon branching from an initial center-point that
-ceases upon contact with axons branching from adjacent centers.  The
+ceases upon contact with axons branching from adjacent centers.  However, the
 assumption of pre-arranged center-points is difficult to resolve with the
@@
 may be positive or negative in order that axons may branch in the direction of
-either higher or lower concentrations. \cmnt{The second term on the right
-  quantifies the coupling between axon branching and synaptogenesis}. Here $\chi_i=0$ is a placeholder.
+either higher or lower concentrations. Note that computing the divergence in
+simulation requires cells on the cortical sheet to communicate with immedately
+adjacent cells only (see \emph{Materials \& Methods}) The second term on the
+right quantifies the coupling between axon branching and synaptogenesis. Here
+  $\chi_i=0$ is a placeholder.
 
@@
 \begin{fullwidth}
-\includegraphics[width=\linewidth]{./MainFig_superlowres.png}
+\includegraphics[width=\linewidth]{./MainFig.png}
 \caption{\textbf{A} Left shows a cytochrome oxidase stain obtained from rat S1
@@
 ordering defined by the original $\gamma$ values. However, we found that
-specifying $N$ ordered areas required $M\approx (N+1)/2$ signalling fields. As
-the number of distinct guidance fields is unlikely to approach the number of
-barrels, further extensions were required.
+specifying $N$ ordered areas required $M\approx (N+1)/2$ signalling
+fields. This is because localization of axon densities occurs only when
+projections are influenced by interactions with two or more signalling
+gradients that encourage migration in opposing directions. Note that these
+dynamics are quite unline classic chemospecificity models
+\citep{sperry_chemoaffinity_1963}, which essentially assume center-points
+i.e., conditions in the target tissue that instruct pre-identified afferents
+to stop growing. As the number of distinct guidance fields is unlikely to
+approach the number of barrels, further extensions were required.
 
@@
 projections are dense. Divisive normalization keeps the axon branch density
-bounded on each iteration to the initial density \cmnt{of unconnected
-  branches} at $t=0$
+bounded on each iteration to the initial (random) density of unconnected
+  branches at $t=0$
 %
@@
 where prime symbols indicate the use of intermediate values computed from
-Eqs.~\ref{eq:dc}--\ref{eq:comp} at time $t-{\delta}t$. \cmnt{Note that this
-  operation is local to individual afferent projections.}
+Eqs.~\ref{eq:dc}--\ref{eq:comp} at time $t-{\delta}t$. Note that this
+  operation is local to individual afferent projections.
 
@@
 between emergent cortical fields must be due to differences in these values
-only. We speculate that a projection interacts with a given ephrin field in
-the cortical subplate with a strength determined by the concentration of a
-similar molecule at its thalamic origin, i.e., the putative barreloid
-center. As such, two orthogonal linear thalamic gradients were defined, from
-which 41 pairs of $\gamma$ values were sampled, at the coordinates of 41
-barreloid centers recreated from \citealp{haidarliu_size_2001} (Fig.\,1B).
+only. We speculate that the contribution of a given ephrin field, to the
+velocity at which a projection migrates across the cortical subplate, is
+determined by the concentration of a similar molecule at its thalamic origin,
+i.e., the putative barreloid center. As such, two orthogonal linear thalamic
+gradients were defined, from which 41 pairs of $\gamma$ values were sampled,
+at the coordinates of 41 barreloid centers recreated from
+\citealp{haidarliu_size_2001} (Fig.\,1B).
 
@@
 projections compete for a limited number of available synaptic connections
-(Eqs.~\ref{eq:dc}--\ref{eq:da}), ii) at each location the branching rate of a
+(Eqs.~\ref{eq:dc}--\ref{eq:da}), ii) at each We speculate that a projection interacts with a given ephrin field in
+the cortical subplate with a strength determined by the concentration of a
+similar molecule at its thalamic origin, i.e., the putative barreloid
+centerlocation the branching rate of a
 given projection is reduced by the density of other projections
@@
 
-The emergence of barrels in simulation required competition between thalamic
-projections in terms of synaptic connectivity \cmnt{and an implicit self/other
-  identifier amoungst projections}, and also in terms of cortical space (as
-represented by $\chi$). This requirement may account for the absence of
+The emergence of barrels in simulation required cWe speculate that a projection interacts with a given ephrin field in
+the cortical subplate with a strength determined by the concentration of a
+similar molecule at its thalamic origin, i.e., the putative barreloid
+centerompetition between thalamic
+projections in terms of synaptic connectivity and an implicit self/other
+identifier amoungst projections, and also competition in terms of cortical
+space (as represented by $\chi$). This latter requirement may account for the
+absence of
 barrels in rodents with larger brains, such as capybara, for which competition
@@
 
-The cortex was modelled as a two dimensional hexagonal lattice which
-simplifies the computation of the 2D Laplacian. Within a boundary traced from
-rat primary barrel cortex (Fig.\,1A) we set the hex-to-hex distance
+The cortical sheet was modelled as a two dimensional hexagonal lattice, which
+simplifies the computation of the 2D Laplacian. Within a boundary traced
+around the edge of a
+rat barrel field (Fig.\,1A) we set the hex-to-hex distance
 $d$ to 0.03\,mm, which resulted in a lattice containing 6515 hexes for the
@@
 \begin{split}
-\langle D \dvrg a_i(\mb{p}_0,t) \rangle  = \frac{1}{\Omega} \iint_\Omega \dvrg a_i(\mb{x}, t) dxdy & = \frac{1}{\Omega} \oint_c \frac{\partial a_i}{\partial \hat{\mb{n}}} d\gamma \\
+\langle D \dvrg a_i(\mb{p}_0,t) \rangle  = \frac{1}{\Omega} \oiint_\Omega \dvrg a_i(\mb{x}, t) \dif\Omega & = \frac{1}{\Omega} \oint \frac{\partial a_i}{\partial \hat{\mb{n}}} d\gamma \\
 & \approx \frac{1}{\Omega} \sum_{j=1}^6 \frac{\partial a_i(\mb{p}_j)}{\partial \hat{\mb{n}}} \bigg\rvert_{mid} v \\
@@
 %
-where $d$ is the centre-to-centre distance between hexes in the grid, $v =
-d/\sqrt{3}$ is the length of each of the 6 edges of the hexagon's perimeter
-and $d\gamma$ is an infinitessimally small distance along the perimeter of the
-hexagon.
+where $v = d/\sqrt{3}$ is the length of each edge of the hexagon and $d\gamma$
+is an infinitesimally small distance along its perimeter.
 
@@
 %
-\frac{1}{\Omega} \iint_\Omega a_i\nabla\vcdot\mb{g}\;dxdy & = \frac{a_i(\mb{p}_0,t)}{\Omega}  \oint_c \mb{g}\vcdot d\hat{\mathbf{n}} \\
+\frac{1}{\Omega} \oiint_\Omega a_i\nabla\vcdot\mb{g}\;\dif\Omega & = \frac{a_i(\mb{p}_0,t)}{\Omega}  \oint \mb{g}\vcdot d\hat{\mathbf{n}} \\
 %
@@
 %
-Both this last expression, and the final expression of Eq.\,\ref{eq:lapl} are
-easily computed, comprising short sums of nearest-neighbour relationships.
+Both this last expression, and the final expression of Eq.\,\ref{eq:lapl} can
+be computed locally, by summing over values of the nearest neighbours.
 
@@
 computations described in (i), (ii) and (iii) may be found in the class method
-\code{RD\_James::compute\_divJ()} as \code{term1}, \code{term2} and \code{term3},
-respectively.
+\code{RD\_James::compute\_divJ()} which calculates \code{term1}, \code{term2}
+and \code{term3}, respectively.
 

### 2020-01-29 paper/briefpaper/brief.tex
Tweaks to maths notation

@@
 \begin{split}
-\langle D \dvrg a_i(\mb{p}_0,t) \rangle  = \frac{1}{\Omega} \oiint_\Omega \dvrg a_i(\mb{x}, t) \dif\Omega & = \frac{1}{\Omega} \oint \frac{\partial a_i}{\partial \hat{\mb{n}}} d\gamma \\
-& \approx \frac{1}{\Omega} \sum_{j=1}^6 \frac{\partial a_i(\mb{p}_j)}{\partial \hat{\mb{n}}} \bigg\rvert_{mid} v \\
+\langle D \dvrg a_i(\mb{p}_0,t) \rangle  = \frac{1}{\Omega} \oiint_\Omega \dvrg a_i(\mb{x}, t) \dif\Omega & = \frac{1}{\Omega} \oint \frac{\partial a_i}{\partial \hat{\mb{n}}} \dif\gamma \\
+& \approx \frac{1}{\Omega} \sum_{j=1}^6 \frac{\partial a_i(\mb{p}_j)}{\partial \hat{\mb{n}}} \bigg\rvert_{\mathrm{mid}} v \\
 & = \frac{2}{\sqrt{3} d^2} \sum_{j=1}^6 \frac{a_i(\mb{p}_j) - a_i(\mb{p}_0)}{d} \frac{d}{\sqrt{3}} \\
@@
 %
-where $v = d/\sqrt{3}$ is the length of each edge of the hexagon and $d\gamma$
+where $v = d/\sqrt{3}$ is the length of each edge of the hexagon and $\dif\gamma$
 is an infinitesimally small distance along its perimeter.
@@
 ii) The computation of the second term in Eq.\,\ref{eq:divJExpanded},
-$\langle a_i(\mb{p}_0,t)\nabla\vcdot\mb{g}(\mb{p}_0)\rangle$, can be written out similarly:
+$\langle a_i(\mb{p}_0,t)\nabla\vcdot\mb{g}_i(\mb{p}_0)\rangle$, can be written out similarly:
 
@@
 %
-\frac{1}{\Omega} \oiint_\Omega a_i\nabla\vcdot\mb{g}\;\dif\Omega & = \frac{a_i(\mb{p}_0,t)}{\Omega}  \oint \mb{g}\vcdot d\hat{\mathbf{n}} \\
+\frac{1}{\Omega} \oiint_\Omega a_i\nabla\vcdot\mb{g}_i\;\dif\Omega & = \frac{a_i(\mb{p}_0,t)}{\Omega}  \oint \mb{g}_i\vcdot \dif\hat{\mathbf{n}} \\
 %
-& \approx \frac{a_i(\mb{p}_0,t)}{\Omega} \sum_{j=1}^{6} \frac{\mb{g}_j + \mb{g}_0}{2}\vcdot \hat{\mb{n}}\;v \\
+& \approx \frac{a_i(\mb{p}_0,t)}{\Omega} \sum_{j=1}^{6} \frac{\mb{g}_i(\mb{p}_j) + \mb{g}_i(\mb{p}_0)}{2}\vcdot \hat{\mb{n}}\;v \\
 %
-& = \frac{2 a_i(\mb{p}_0,t)v}{\sqrt{3}d^2} \bigg( \sum_{j=1}^{6} \frac{\mb{g}_j^x + \mb{g}_0^x}{2} \vcdot  \hat{\mb{n}} + \sum_{j=1}^{6} \frac{\mb{g}_j^y + \mb{g}_0^y}{2} \vcdot  \hat{\mb{n}} \bigg) \\
+& = \frac{2 a_i(\mb{p}_0,t)v}{\sqrt{3}d^2} \sum_{j=1}^{6} \bigg[ \frac{g_i^x(\mb{p}_j) + g_i^x(\mb{p}_0)}{2} \vcdot  \hat{\mb{n}} + \frac{g_i^y(\mb{p}_j) + g_i^y(\mb{p}_0)}{2} \vcdot  \hat{\mb{n}} \bigg] \\
 %
@@
 %
-\Rightarrow \langle a_i(\mb{p}_0,t)\nabla\vcdot\mb{g}(\mb{p}_0)\rangle & \approx \frac{a_i(\mb{p}_0,t)}{3d} \bigg( \sum_{j=1}^{6} \big(\mb{g}_j^x + \mb{g}_0^x\big) \cos \big(\frac{\pi}{3}(j-1)\big) + \sum_{j=1}^{6} \big({\mb{g}_j^y + \mb{g}_0^y}\big) \sin \big(\frac{\pi}{3}(j-1)\big) \bigg).
-\\
+%\Rightarrow \langle a_i(\mb{p}_0,t)\nabla\vcdot\mb{g}(\mb{p}_0)\rangle &
+%\approx \frac{a_i(\mb{p}_0,t)}{3d} \bigg( \sum_{j=1}^{6} \big(g_i^x(\mb{p}_j)
+%+ g_i^x(\mb{p}_0)\big) \cos \big(\frac{\pi}{3}(j-1)\big) + \sum_{j=1}^{6}
+%\big({g_i^y(\mb{p}_j) + g_i^y(\mb{p}_0)}\big) \sin
+%\big(\frac{\pi}{3}(j-1)\big) \bigg), \\
+%% Condensing into a single sum:
+\Rightarrow \langle a_i(\mb{p}_0,t)\nabla\vcdot\mb{g}(\mb{p}_0)\rangle & \approx \frac{a_i(\mb{p}_0,t)}{3d} \sum_{j=1}^{6} \bigg[ \big(g_i^x(\mb{p}_j) + g_i^x(\mb{p}_0)\big) \cos \big(\frac{\pi}{3}(j-1)\big) + \big({g_i^y(\mb{p}_j) + g_i^y(\mb{p}_0)}\big) \sin \big(\frac{\pi}{3}(j-1)\big) \bigg], \\
 %
@@
 %
-Both this last expression, and the final expression of Eq.\,\ref{eq:lapl} can
+where $g_i^x$ and $g_i^y$ are the Cartesian components of $\mb{g}_i$. Both this
+last expression, and the final expression of Eq.\,\ref{eq:lapl} can
 be computed locally, by summing over values of the nearest neighbours.

### 2020-01-29 paper/briefpaper/brief.tex
Fixes a mis-pasted tex error in the manuscript

@@
 projections compete for a limited number of available synaptic connections
-(Eqs.~\ref{eq:dc}--\ref{eq:da}), ii) at each We speculate that a projection interacts with a given ephrin field in
-the cortical subplate with a strength determined by the concentration of a
-similar molecule at its thalamic origin, i.e., the putative barreloid
-centerlocation the branching rate of a
+(Eqs.~\ref{eq:dc}--\ref{eq:da}), ii) at each location the branching rate of a
 given projection is reduced by the density of other projections
@@
 
-The emergence of barrels in simulation required cWe speculate that a projection interacts with a given ephrin field in
-the cortical subplate with a strength determined by the concentration of a
-similar molecule at its thalamic origin, i.e., the putative barreloid
-centerompetition between thalamic
+The emergence of barrels in simulation required competition between thalamic
 projections in terms of synaptic connectivity and an implicit self/other

### 2020-04-21 paper/briefpaper/brief.tex

@@
 Brain development relies on an interplay between genetic specification and
-self-organisation. Striking examples of this relationship can be found in the
+self-organization. Striking examples of this relationship can be found in the
 somatosensory brainstem, thalamus, and cortex of rats and mice, where the
@@
 The center-point and reaction-diffusion models are not mutually
-exclusive. Pre-organised centers could bias reaction-diffusion processes to
+exclusive. Pre-organized centers could bias reaction-diffusion processes to
 generate specific arrangements more reliably, and mechanisms of lateral axon
@@
 interactions, would show that a separate stage and/or extrinsic mechanism for
-pre-organising thalamocortical connections need not be assumed. To this end,
+pre-organizing thalamocortical connections need not be assumed. To this end,
 we ask whether barrel maps can emerge in a system with reaction-diffusion
@@
 gradients that encourage migration in opposing directions. Note that these
-dynamics are quite unline classic chemospecificity models
+dynamics are quite unlike classic chemospecificity models
 \citep{sperry_chemoaffinity_1963}, which essentially assume center-points
@@
 where prime symbols indicate the use of intermediate values computed from
-Eqs.~\ref{eq:dc}--\ref{eq:comp} at time $t-{\delta}t$. Note that this
+Eqs.~\ref{eq:dc}--\ref{eq:comp} at time $t-{\delta}t$ (${\delta}t$ is the time
+represented by one simulation step). Note that this
   operation is local to individual afferent projections.
@@
 at the coordinates of 41 barreloid centers recreated from
-\citealp{haidarliu_size_2001} (Fig.\,1B).
+\cite{haidarliu_size_2001} (Fig.\,1B).
 
@@
 anterior-posterior and medial-lateral axes. From random initial conditions for
-$a({\bf x},0)$ and $c({\bf x},0)$, a clear Voronoi-like tesselation emerged
+$a(\mb{x},0)$ and $c(\mb{x},0)$, a clear Voronoi-like tesselation emerged
 (Fig.\,1C; see Movie S1). A reduction in the Honda delta metric (see
 \citealp{senft_mouse_1991}) confirmed that a `good' Voronoi pattern emerged
-within $\approx 3000$ iterations (see Fig.\,1D). A measure of the difference
+within $\approx 3000$ iterations (Fig.\,1D). A measure of the difference
 in shape between real and simulated barrels revealed a strong correspondence
@@
 The emergence of barrels in simulation required competition between thalamic
-projections in terms of synaptic connectivity and an implicit self/other
-identifier amoungst projections, and also competition in terms of cortical
-space (as represented by $\chi$). This latter requirement may account for the
-absence of
-barrels in rodents with larger brains, such as capybara, for which competition
-for space is presumably weaker \citep{woolsey_comparative_1975}. Hence,
-irrespective of whether barrels are necessary for adaptive whisker function,
-the emergence of somatotopically ordered modular structures may be an
-inevitable consequence of local competition for cortical territory driven by
-input from an array of discrete sensory organs \citep{purves_iterated_1992}.
+projections in terms of synaptic connectivity and also competition in terms of
+cortical space, as represented by $\chi$, with an implicit requirement for a
+self/other identifier amoungst projections. This latter form of competition may
+account for the absence of barrels in rodents with larger brains, such as
+capybara, for which competition for space is presumably weaker
+\citep{woolsey_comparative_1975}. Hence, irrespective of whether barrels are
+necessary for adaptive whisker function, the emergence of somatotopically
+ordered modular structures may be an inevitable consequence of local
+competition for cortical territory driven by input from an array of discrete
+sensory organs \citep{purves_iterated_1992}.
 
@@
 given cortical grid cell to any but those immediately adjacent (via
-diffusion). Hence the simulations demonstrate how a self-organising system,
+diffusion). Hence the simulations demonstrate how a self-organizing system,
 constrained by genetically specified guidance cues and by the shape of the

### 2020-05-19 paper/supplementary/supp.tex
Some more notes on the Post-eLife review work

@@
 %
-\begin{equation}
-\hat{a}_i(\mb{x},t) = \sum_{p\ne i}^N a_p(\mb{x},t)
+\begin{equation}\label{eq:ahat}
+\hat{a}_i(\mb{x},t) = \sum_{j\ne i}^N a_j(\mb{x},t)
 \end{equation}
@@
 = \nabla\vcdot\mb{J}_i(\mb{x},t) - \frac{\partial
-c_i(\mb{x},t)}{\partial t} - \frac{\epsilon_i  a_i}{N-1} \sum_{p \ne i}^{N} a_p(\mb{x}, t)^l
+c_i(\mb{x},t)}{\partial t} - \frac{\epsilon_i  a_i}{N-1} \sum_{j \ne i}^{N} a_j(\mb{x}, t)^l
 \end{equation}
@@
 = \nabla\vcdot\mb{J}_i + \alpha_i c_i - \beta_i n
-[a_i]^k - \frac{\epsilon_i}{N-1} a_i \sum_{p \ne i}^{N} a_p^l
+[a_i]^k - \frac{\epsilon_i}{N-1} a_i \sum_{j \ne i}^{N} (a_j)^l
 \end{equation}
@@
 See \code{configs/whiskers/5N\_boundedorder.json}
+
+\section {Post eLife Review work}
+
+One of the reviewers, Bard Ermentrout, suggested an alternative formulation of
+the branching competition, to avoid the need to divisively normalize the
+branching density at each simulation step.
+
+The suggestion was to let the $\chi$ term from equation 3 from the reviewed
+paper be given by:
+
+\begin{equation}\label{eq:chibard}
+\chi_i = \frac{a_i}{N-1} \nabla\hat{a_i}
+\end{equation}
+
+where $\hat{a}_i$, the sum of branching over all other TC types, is given by
+Eq.~\ref{eq:ahat}.
+
+Compare this with Eq.~\ref{eq:J_NM_with_comp} from Section \ref{sec:comp2},
+which was one of the competition types I tried fairly early on. In fact, I
+think it was about my first idea for competition. It's similar, but not
+identical. I passed it over at that stage, because the competition it produced
+seemed weak, and I was relying on the development of contours and `well
+localized regions' to indicate success. In Eq.~\ref{eq:J_NM_with_comp}, the
+additional term is in the expression for $\mb{J}_i$, so the \emph{flux of
+branching density for TC type $i$ is a function of the gradient of the
+branching of other TC types, $j\ne i$}. Eq.~\ref{eq:chibard} says something
+subtly different: \emph{the rate of change of branching density is modulated
+by the gradient of branching of other TC types}.
+
+Implementing Eq.~\ref{eq:chibard} gives:
+
+\begin{equation} \label{eq:dabard}
+\frac{\partial a_i}{\partial t} = \nabla\vcdot\left(D \nabla a_i-a_i\sum_{j=1}^{M} \gamma_{i,j}\nabla \rho_j(\mb{x})\right) - \frac{\partial c_i}{\partial t} + \frac{a_i}{N-1} \nabla\hat{a_i}.
+\end{equation}
+
+in the style of the preprint.
+
+Although I have re-implemented rd\_james\_comp2, which expresses
+Eq.~\ref{eq:J_NM_with_comp},and found that it does provide competition enough
+to generate a barrel field like ordering, I've not yet implemented Bard's
+suggestion precisely, so I think that's up next. However, I imagine that this
+may be unstable without the sigmoiding of $a_i$ in the expression for
+$\chi_i$.
+
 %

### 2020-05-19 paper/supplementary/supp.tex
Adds text based on realization that Bard's suggestion can't be implemented as suggested#

@@
 
-in the style of the preprint.
-
-Although I have re-implemented rd\_james\_comp2, which expresses
-Eq.~\ref{eq:J_NM_with_comp},and found that it does provide competition enough
-to generate a barrel field like ordering, I've not yet implemented Bard's
-suggestion precisely, so I think that's up next. However, I imagine that this
-may be unstable without the sigmoiding of $a_i$ in the expression for
-$\chi_i$.
+in the style of the preprint. There's a problem here. The problem is that the
+gradient of $\hat{a}_i$ is a \emph{vector} quantity. So in fact, Bard's
+suggestion doesn't make good sense, and Eq.~\ref{eq:J_NM_with_comp} is more
+sensible.
+
+I have re-implemented rd\_james\_comp2, which expresses
+Eq.~\ref{eq:J_NM_with_comp}, and found that it does provide competition enough
+to generate a barrel-field-like ordering.
 

### 2020-05-20 paper/supplementary/supp.tex
Updates to supplementary, plus pdf for convenience

--- 2020-01-24 paper/supplementary/supp.tex
@@
 sharpness, $s$:
-\begin{equation}
+\begin{equation}\label{eq:abar}
 \bar{a}_i = \frac{h}{1 + e^{(o - s \, a_i)}}
@@
 See \code{configs/whiskers/5N\_boundedorder.json}
+
+\section {Post eLife Review work}
+
+One of the reviewers, Bard Ermentrout, suggested an alternative formulation of
+the branching competition, to avoid the undesirable need to divisively
+normalize the branching density at each simulation step.
+
+The suggestion was to let the $\chi$ term from Eq.~3 of the preprint be given
+by:
+
+\begin{equation}\label{eq:chibard}
+\chi_i = \frac{a_i}{N-1} \nabla\hat{a}_i,
+\end{equation}
+
+where $\hat{a}_i$, the sum of branching over all other TC types, is given by
+Eq.~\ref{eq:ahat}. Implementing Eq.~\ref{eq:chibard} gives:
+
+\begin{equation} \label{eq:dabard}
+\frac{\partial a_i}{\partial t} =
+\nabla\vcdot\left(D \nabla a_i-a_i\sum_{j=1}^{M} \gamma_{i,j}\nabla \rho_j(\mb{x})\right)
++ \frac{a_i}{N-1} \nabla\hat{a}_i
+- \frac{\partial c_i}{\partial t}.
+\end{equation}
+
+There's a problem here. The problem is that the gradient of $\hat{a}_i$ is
+a \emph{vector} quantity whereas $\dot{a}_i$ is scalar. So in fact, Bard's
+suggestion doesn't make good sense as written.
+
+Compare this with Eq.~\ref{eq:J_NM_with_comp} from Section \ref{sec:comp2},
+which was one of the competition types I tried fairly early on. In fact, it
+was my second idea for competition. It's similar, but not identical. Written
+out in the same style as Eq.~\ref{eq:dabard} it is:
+
+\begin{equation} \label{eq:da_comp2_elife}
+\frac{\partial a_i}{\partial t}
+= \nabla\vcdot  \left(D \nabla a_i
+- a_i\sum_{j=1}^{M} \gamma_{i,j}\nabla \rho_j(\mb{x})
++ \frac{\bar{a}_i F}{N-1} \nabla \hat{a}_i \right)
+- \frac{\partial c_i}{\partial t}.
+\end{equation}
+
+This modifies the axon branching with a similar term to Eq.~\ref{eq:chibard},
+depending on the gradient of $\hat{a}_i = \sum_{j\ne i}{a_j}$. However, here,
+the term is included in the \emph{flux of axonal branching}
+(i.e.~\emph{inside} the brackets) and so its divergence is taken and this
+contributes a scalar quantity to the change of $a_i$. Two other points to note
+about Eq.~\ref{eq:da_comp2_elife}: I included a scalar parameter $F$ to modify
+the relative contribution of this term; and $\bar{a}_i$ is the result of
+passing $a_i$ through a sigmoidal transfer function (see Eq.~\ref{eq:abar})
+which I found necessary to prevent the term from becoming numerically
+unstable.
+
+I passed Eq.~\ref{eq:da_comp2_elife} over for the preprint, because the
+competition it produced seemed weak, and I was relying on the development of
+contours and `well localized regions' to indicate success.
+
+After re-examining the effect of \code{rd\_james\_comp2}, which expresses
+Eq.~\ref{eq:da_comp2_elife}, I found that it does provide competition enough
+to generate a barrel-field-like ordering, when the argmaxes of the $N$ $c_i$
+values are plotted, as in the preprint.
+
+\begin{equation}\label{eq:chibard_mag}
+\chi_i = \frac{a_i}{N-1} \| \nabla\hat{a}_i\|.
+\end{equation}
+
 %

### 2020-05-22 paper/briefpaper/brief.tex
The final tag in the repo will just be eLife
WIP on paper

@@
 
+\usepackage{color}
 \usepackage{amsmath,esint}
+
+% Some document-defined commands
 \newcommand*\dif{\mathop{}\!\mathrm{d}}
+\newcommand{\cmnt}[1]{\textcolor{blue}{#1}}
+\newcommand{\dvrg}{\nabla\vcdot\nabla}
+\newcommand{\e}{\emph}
+\newcommand{\bol}{\textbf}
+\newcommand{\mb}[1]{\mathbf{#1}}
+\makeatletter
+\newcommand*\vcdot{\mathpalette\vcdot@{.35}}
+\newcommand*\vcdot@[2]{\mathbin{\vcenter{\hbox{\scalebox{#2}{$\m@th#1\bullet$}}}}}
+\newcommand{\code}[1]{\textsf{#1}}
 
@@
 \end{abstract}
-
-% Blue for comments
-\newcommand{\cmnt}[1]{\textcolor{blue}{#1}}
-\newcommand{\dvrg}{\nabla\vcdot\nabla}
-\newcommand{\e}{\emph}
-\newcommand{\bol}{\textbf}
-\newcommand{\mb}[1]{\mathbf{#1}}
-\makeatletter
-\newcommand*\vcdot{\mathpalette\vcdot@{.35}}
-\newcommand*\vcdot@[2]{\mathbin{\vcenter{\hbox{\scalebox{#2}{$\m@th#1\bullet$}}}}}
-\newcommand{\code}[1]{\textsf{#1}}
 
@@
 \begin{equation} \label{eq:da}
-\frac{\partial a_i}{\partial t} = \nabla\vcdot\left(D \nabla a_i-a_i\sum_{j=1}^{M} \gamma_{i,j}\nabla \rho_j(\mb{x})\right) - \frac{\partial c_i}{\partial t} + \chi_i.
+\frac{\partial a_i}{\partial t} = \nabla\vcdot\left(D \nabla a_i-a_i\sum_{j=1}^{M} \gamma_{i,j}\nabla \rho_j(\mb{x}) \cmnt{+ \chi_i}\right) - \frac{\partial c_i}{\partial t}.
 \end{equation}
@@
 \begin{fullwidth}
-\includegraphics[width=\linewidth]{./MainFig.png}
+%\includegraphics[width=\linewidth]{./MainFig.png}
+\includegraphics[width=\linewidth]{./MainFig_superlowres.png}
 \caption{\textbf{A} Left shows a cytochrome oxidase stain obtained from rat S1
@@
 connections. To introduce competition also in terms of axon branching we
-redefined
+% EITHER
+%%redefined
+%%%
+%%\begin{equation} \label{eq:comp}
+%%\color{blue}
+%%\chi_i(\mb{x}, t) = \frac{\bar{a}_i F}{N-1} \nabla \sum_{j\ne i}{a_j}
+%%\color{black}
+%%\end{equation}
+% OR
+\cmnt{first define a quantity to represent the branching density of all
+  \emph{other} TC types; $\hat{a}_i = \sum_{j\ne i}{a_j}$; then set the term
+  $\chi_i$ to be}
 %
 \begin{equation} \label{eq:comp}
-\chi_i(\mb{x}, t) = - \frac{\epsilon  a_i}{N-1} \sum_{j \ne i}^{N} a_{j},
+\color{blue}
+\chi_i(\mb{x}, t) = \frac{\bar{a}_i F}{N-1} \nabla \hat{a}_i
+\color{black}
+\end{equation}
+
+%
+\cmnt{where $F$ is a parameter and $\bar{a}_i$ is a bounded version of $a_i$
+  obtained by passing the branching density through the sigmoid function}
+%
+\begin{equation} \label{eq:abar}
+\color{blue}
+  \bar{a}_i = \frac{2}{1+\mathrm{e}^{5 - 0.5 a_i}}.
+\color{black}
 \end{equation}
 %
-which reduces branching for each projection where the branches of other
-projections are dense. Divisive normalization keeps the axon branch density
-bounded on each iteration to the initial (random) density of unconnected
-  branches at $t=0$
-%
-\begin{equation} \label{eq:norm}
-  a_i(\mathbf{x}, t) = \oiint_{S} \big[ a_i(\mathbf{x}, 0) - c_i(\mathbf{x}, t) \big] \dif S \; \frac {a'_i(\mathbf{x}, t)} {\oiint_{S}
-  a'_i(\mathbf{x}, t) \dif S}\,,
-\end{equation}
-%
-where prime symbols indicate the use of intermediate values computed from
-Eqs.~\ref{eq:dc}--\ref{eq:comp} at time $t-{\delta}t$ (${\delta}t$ is the time
-represented by one simulation step). Note that this
-  operation is local to individual afferent projections.
+\cmnt{In this expression for $\bar{a}_i$ (only), we model the idea
+ that there must be an upper bound on the axonal branching density.}
+
+\cmnt{which contributes to the \emph{flux of axonal branching} by making branching
+  density for each projection `flow away' from regions} where the branches of
+other projections are dense. Note that this operation is local to individual
+afferent projections.
 
@@
 All code required to reproduce these results is available at
-\url{https://github.com/ABRG-Models/BarrelEmerge/tree/eLife_submission1}. The
+\url{https://github.com/ABRG-Models/BarrelEmerge/tree/eLife}. The
 computations described in (i), (ii) and (iii) may be found in the class method

### 2020-05-23 paper/briefpaper/brief.tex
Back down to 3 main equations, because the sigmoid transfer function turns out to be unnecessary in comop2

@@
 \color{blue}
-\chi_i(\mb{x}, t) = \frac{\bar{a}_i F}{N-1} \nabla \hat{a}_i
-\color{black}
-\end{equation}
-
-%
-\cmnt{where $F$ is a parameter and $\bar{a}_i$ is a bounded version of $a_i$
-  obtained by passing the branching density through the sigmoid function}
-%
-\begin{equation} \label{eq:abar}
-\color{blue}
-  \bar{a}_i = \frac{2}{1+\mathrm{e}^{5 - 0.5 a_i}}.
+\chi_i(\mb{x}, t) = \frac{a_i F}{N-1} \nabla \hat{a}_i,
 \color{black}
@@
 %
-\cmnt{In this expression for $\bar{a}_i$ (only), we model the idea
- that there must be an upper bound on the axonal branching density.}
-
-\cmnt{which contributes to the \emph{flux of axonal branching} by making branching
-  density for each projection `flow away' from regions} where the branches of
-other projections are dense. Note that this operation is local to individual
-afferent projections.
+\cmnt{which contributes to the \emph{flux of axonal branching} by making
+  branching density for each projection `flow away' from regions} where the
+branches of other projections are dense. \cmnt{$F$ parameterizes the strength
+  of this competitive interaction.} Note that this operation is local to
+individual afferent projections.
 
@@
 cytochrome oxidase stain from \cite{zheng_signal_2001}, and
-Eqs.~\ref{eq:dc}--\ref{eq:norm} were solved for $N=41$ projections on the
+Eqs.~\ref{eq:dc}--\ref{eq:comp} were solved for $N=41$ projections on the
 resulting domain, using $M=2$ linear signalling gradients aligned with the

### 2020-05-26 analysis/process_paramsearch_comp2_linmax.py
Updates to processing scripts for model with/without the max bounding limit on a in the competition term

@@ -47,8 +47,7 @@ k_vals = [ 3.0 ]
 #
 # Filename format is like: pe_dncomp_D0.0631_ep300_ab0.01_k3
 # Filename format is like: pe_comp2_D0.0631_F1_ab0.01_k3
-#basedir = '/home/seb/gdrive_usfd/data/BarrelEmerge/paramexplore_comp2/'
-basedir = '/home/seb/paramexplore_comp2/'
+basedir = '/home/seb/paramexplore_comp2_linmax/'
 table = []
 for logdirname in os.listdir(basedir):
 
@@ -201,7 +200,7 @@ for logdirname in os.listdir(basedir):
         table.append (tableline)
 
 import csv
-with open(('postproc/paramsearch_k{0}_comp2.csv'.format(k_vals[0])), 'w', newline='\n') as csvfile:
+with open(('postproc/paramsearch_k{0}_comp2_linmax.csv'.format(ff_k)), 'w', newline='\n') as csvfile:
     cw = csv.writer (csvfile, delimiter=',')
     cw.writerow (['k','D','alphabeta','alpha','beta','epsilon','t','hondadelta','sos_dist','area_diff','F','localization'])
     for tableline in table:

### 2020-05-26 paper/supplementary/supp.tex
Removes irrelevant eqn
Adds note about no need for sigmoid

@@
 which I found necessary to prevent the term from becoming numerically
-unstable.
+unstable.*
 
@@
 
-\begin{equation}\label{eq:chibard_mag}
-\chi_i = \frac{a_i}{N-1} \| \nabla\hat{a}_i\|.
+*: Ah, no sigmoid transfer function required. With $\bar{a}_i = a_i$ the
+ system runs successfully, although for a limited parameter range. The system
+ can be made slightly more stable by making
+%
+\begin{equation}
+\begin{split}
+\bar{a}_i & = a_i, a<a_{max} \\
+          & = a_{max}, a\geq a_{max}
+\end{split}
 \end{equation}

### 2020-05-26 scripts/paramexplore_comp2_linmax.sh
Sorts out the full range of parameter exploration

@@ -47,8 +47,8 @@ EPSILON=150 # Doesn't do anything, but is incorporated into the JSON anyway
 DT=0.0001
 HEXHEXD=0.03
 
-for D in 0.12 0.06 0.03; do # 0.25 0.5 1.0; do
-    for F in 0.48; do # 0.03 0.08 0.19 0.48 1.2 3.0; do
+for D in 0.03 0.06 0.12 0.25 0.5 1.0; do
+    for F in 0.03 0.08 0.19 0.48 1.2 3.0; do
         for ALPHABETA in 0.06 0.18 0.55 1.6 5.0 15.0; do
 
             ((BETA=3/ALPHABETA))

### 2020-05-27 paper/briefpaper/brief.tex
Wednesday work on figures

@@
 connections. To introduce competition also in terms of axon branching we
-% EITHER
-%%redefined
-%%%
-%%\begin{equation} \label{eq:comp}
-%%\color{blue}
-%%\chi_i(\mb{x}, t) = \frac{\bar{a}_i F}{N-1} \nabla \sum_{j\ne i}{a_j}
-%%\color{black}
-%%\end{equation}
-% OR
-\cmnt{first define a quantity to represent the branching density of all
-  \emph{other} TC types; $\hat{a}_i = \sum_{j\ne i}{a_j}$; then set the term
-  $\chi_i$ to be}
+redefined
 %
@@
 \color{blue}
-\chi_i(\mb{x}, t) = \frac{a_i F}{N-1} \nabla \hat{a}_i,
+\chi_i(\mb{x}, t) = \frac{\epsilon a_i}{N-1} \nabla \textstyle\sum_{j\ne i}{a_j}
 \color{black}
@@
 %
-\cmnt{which contributes to the \emph{flux of axonal branching} by making
-  branching density for each projection `flow away' from regions} where the
-branches of other projections are dense. \cmnt{$F$ parameterizes the strength
-  of this competitive interaction.} Note that this operation is local to
-individual afferent projections.
+\cmnt{which contributes to the \emph{flux of axonal branching} as an
+  additional diffusion-like term which makes the branching density for a given
+  projection `flow away' from regions} where the branches of other projections
+become dense. \cmnt{$\epsilon$ parameterizes the strength of this competitive
+  interaction.} Note that this operation is local to individual afferent
+projections.
 
@@
 
+\begin{figure}
+\begin{fullwidth}
+\includegraphics[width=\linewidth]{./paramsweep.png}
+\caption{\cmnt{\textbf{A} Parameter sweep caption goes here}}
+\label{fig:paramsweep}
+\end{fullwidth}
+\end{figure}
+
+\cmnt{Parameter sweep paragraph goes here.}
+
 To further investigate the interplay of genetic and intrinsic mechanisms we
@@
 projections compete for a limited number of available synaptic connections
-(Eqs.~\ref{eq:dc}--\ref{eq:da}), ii) at each location the branching rate of a
+(Eqs.~\ref{eq:dc}--\ref{eq:da}), ii) \cmnt{review} at each location the branching rate of a
 given projection is reduced by the density of other projections
-(Eq.\,\ref{eq:comp}), and iii) the branch density of each projection is
+(Eq.\,\ref{eq:comp}), and iii)  \cmnt{review} the branch density of each projection is
 conserved over time (Eq.\,\ref{eq:norm}).

### 2020-05-27 analysis/include/Surface.py

@@ -35,7 +35,7 @@ class Surface:
         # If the graph wants a title
         self.title = ''
         # The colourmap to use when plotting z values vs x,y (i.e., when c is empty)
-        self.cmap = plt.cm.jet
+        self.cmap = plt.cm.viridis
         # Scale bar parameters
         self.showScalebar = False
         # Scale bar coordinate 1
@@ -67,7 +67,7 @@ class Surface:
         # Set True to draw a box or something around the map in an ID colour
         self.drawid = False
         self.idcolour = 0
-        self.textid = True
+        self.textid = False
         self.idlabel = ''
 
         # The data to plot

### 2020-05-27 analysis/include/paramplot.py

@@ -576,14 +576,14 @@ def paramplot_pub (sdata, F, column_tag, x_tag, y_tag, ktarg, ttarg, param_tuple
             for j in range(0,len(y_all)): tlist.append(''.format(y6x6[j,0]))
         plt.yticks (y_tick_list, tlist)
         if coltarg == colall[0]:
-            ax.set_ylabel('{0}'.format(_y_tag))
-            ax.text(-3.5,2.5,'Honda $\delta$',rotation=90,va='center')
+            ax.set_ylabel('{0}'.format(_y_tag),rotation=0)
+            ax.text(-3.5,2.5,'$\delta$',rotation=0)
 
         if column_tag == 'alphabeta':
             alphbet = coltarg
             alph = 20.0*alphbet
             bet = 3.0/alphbet
-            ax.set_title ('$\\alpha$={2:.2f} $\\beta$={1:.2f}'.format(column_tag, bet, alph))
+            ax.set_title ('$\\alpha$ = {2:.2f}   $\\beta$ = {1:.2f}'.format(column_tag, bet, alph))
         else:
             ax.set_title ('{0}={1:.1f}'.format(column_tag,coltarg))
 
@@ -600,8 +600,8 @@ def paramplot_pub (sdata, F, column_tag, x_tag, y_tag, ktarg, ttarg, param_tuple
             for j in range(0,len(y_all)): tlist.append(''.format(y6x6[j,0]))
         plt.yticks (y_tick_list, tlist)
         if coltarg == colall[0]:
-            ax2.set_ylabel('{0}'.format(_y_tag))
-            ax2.text(-3.5,2.5,'Area difference',rotation=90,va='center')
+            ax2.set_ylabel('{0}'.format(_y_tag), rotation=0)
+            ax2.text(-3.5,2.5,'$\eta$',rotation=0)
 
         ax3 = F.add_subplot(graph_rows,len(colall),(2*len(colall))+i)
         im3 = ax3.imshow (locn6x6, cmap=cmapstr, vmin=locn_min, vmax=locn_max,
@@ -633,12 +633,8 @@ def paramplot_pub (sdata, F, column_tag, x_tag, y_tag, ktarg, ttarg, param_tuple
         plt.yticks (y_tick_list, tlist)
 
         if coltarg == colall[0]:
-            ax3.set_ylabel('{0}'.format(_y_tag))
-            # Not going to use lines to show stuff:
[…]

### 2020-05-27 analysis/plot_honda_quick.py

@@ -18,15 +18,18 @@ col = sebcolour.Colour
 
 # Set plotting defaults
 fs = 32
-fnt = {'family' : 'DejaVu Sans',
+fnt = {'family' : 'Arial',
        'weight' : 'regular',
        'size'   : fs}
 matplotlib.rc('font', **fnt)
 
+# IMPORTANT for svg output of text as things that can be edited in inkscape
+plt.rcParams['svg.fonttype'] = 'none'
+
 # And plot this longhand here:
 F1 = plt.figure (figsize=(9,8))
 
-t1_masked = np.load ('postproc/honda_t.npy')/10000
+t1_masked = np.load ('postproc/honda_t.npy')
 hondadelta = np.load ('postproc/honda_delta.npy')
 area_diff =  np.load ('postproc/area_diff.npy')
 locn_vs_t =  np.load ('postproc/locn_vs_t.npy')
@@ -44,21 +47,30 @@ l2, = ax1.plot((0,xmax), (0.055, 0.055), '--', color=col.red, linewidth=3, label
 
 ax2 = F1.add_subplot(2,1,2)
 
-l3, = ax2.plot(t1_masked, area_diff, 'v', markersize=12, color=col.black)
+l3, = ax2.plot(t1_masked, area_diff, 's', markersize=12, color=col.black)
 
 ax3 = ax2.twinx()
 l4, = ax3.plot(t1_masked, locn_vs_t, 'h', markersize=12, color=col.gray60)
 
 ax2.set_xlabel ('time (10k steps)')
-ax1.set_ylabel ('$\delta$', rotation=0, labelpad=10)
-ax2.set_ylabel ('$\eta$', rotation=0)
-ax3.set_ylabel ('$\omega$', rotation=0)
+ax1.set_ylabel ('$\delta$', rotation=0, labelpad=30)
+ax2.set_ylabel ('$\eta$', rotation=0, labelpad=30)
+ax3.set_ylabel ('$\omega$', rotation=0, labelpad=30)
 ax2.tick_params (axis='y', labelcolor=col.black)
 ax1.set_xlim ((0,xmax))
-ax1.set_ylim ((0,0.3))
-#ax2.set_ylim ((0,500))
-#ax1.set_xticks ((0,1,2,3,4,5))
-#plt.legend()
+ax1.set_ylim ((0,0.35))
+ax2.set_ylim ((0,1.2))
+ax3.set_ylim ((0,8.5))
[…]

### 2020-05-27 analysis/plot_locn.py

@@ -79,7 +79,8 @@ sf.plotPoly()
 # Or single contour for each field
 for ii in range(0,bdo.N):
     c = bdo.c[ii,:,0]
-    sf.addContour (c, 0.2, 'white', 1.0, ii, False);
+    ccontour = 0.95*np.max(c)
+    sf.addContour (c, ccontour, 'white', 1.0, ii, False);
 
 sf.addOuterBoundary()
 

### 2020-05-27 analysis/plot_paramsearch_paper.py

@@ -15,8 +15,11 @@ import sebcolour as sc
 
 # Set plotting defaults
 matplotlib.use('TkAgg') # cleaner likely to work
-fnt = {'family' : 'DejaVu Sans', 'weight' : 'regular', 'size' : 14 }
+fnt = {'family' : 'Arial', 'weight' : 'regular', 'size' : 14 }
 matplotlib.rc('font', **fnt)
+
+# IMPORTANT for svg output of text as things that can be edited in inkscape
+plt.rcParams['svg.fonttype'] = 'none'
 
 # Cols in sdata are: k,D,alphabeta,alpha,beta,epsilon,t,hondadelta,sos_dist,area_diff
 sdata = np.genfromtxt ('postproc/paramsearch_k3.0_comp2.csv', delimiter=",", names=True)
@@ -34,63 +37,65 @@ F1 = plt.figure (figsize=(20,10))
 pp.paramplot_pub (sdata, F1, 'alphabeta', 'D', 'F', 3, timepoint, param_tuples, 0, 0, 'viridis');
 plt.savefig('plots/paramsweep.svg')
 
-# map plots to go under the param sweep colour grids
-map_i = 1 # map index
-for pt in param_tuples:
+do_maps = 0
+if do_maps:
+    # map plots to go under the param sweep colour grids
+    map_i = 1 # map index
+    for pt in param_tuples:
 
-    logdirname = '/home/seb/paramexplore_comp2/pe_comp2_D{2}_F{0}_ab{1}_k3'.format (pt[0], pt[1], pt[2])
-    print ('Plotting image from log dir {0}'.format(logdirname))
-    # Read the data
-    bdo = bd.BarrelData()
-    # Set True for inter-lines:
-    bdo.loadAnalysisData = True
-    bdo.loadDivisions = False
-    # If loadGuidance is True, then expt id map will be plotted:
-    bdo.loadGuidance = False
-    bdo.loadSimData = True
-    bdo.loadTimeStep = timepoint
-    bdo.loadHexFlags = True
-    bdo.load (logdirname)
-    # Compute max of c
-    maxc = np.max (bdo.c, axis=0)
-    # Use the precomputed ID map:
-    c_id = bdo.id_c[:,0]
-    # Compute the colour map
-    colmap = np.zeros([bdo.nhex,3], dtype=float)
-    ii = 0
-    for oneid in c_id:
-        colmap[ii] = bdo.gammaColour_byid[oneid]
-        ii = ii + 1
-    # Plot a surface
-    import Surface as surf
[…]

### 2020-05-28 paper/briefpaper/brief.tex
Thursday work on figures and paper

@@
 The barrel boundaries form a Voronoi tessellation across the cortical sheet
-\citep{senft_mouse_1991} (Fig.\,1A), suggesting that barreloid topology is
+\citep{senft_mouse_1991} (Fig.\,\ref{fig:main}A), suggesting that barreloid topology is
 preserved in the projection of thalamocortical axons into the cortex, and that
@@
       branching density ($a$) for one thalamocortical projection, and two
-  molecular guidance fields ($\rho$). \textbf{B} The strengths of interaction
+      molecular guidance fields ($\rho$), \cmnt{where the domain $S$ has been
+        traced from \textbf{A}}. \textbf{B} The strengths of interaction
       $\gamma$ with fields $\rho_1$ and $\rho_2$ are indicated for each of 41
@@
       \citealp{haidarliu_size_2001}). \textbf{C} Simulation results for parameters
-  $N=41$, $\alpha=3$, $\beta=20$, $k=3$, $D=0.2$, $\gamma\in\pm 2$,
-  $\epsilon=150$ and $\delta{t}=0.0001$. Colours indicate the thalamic
-  projection for which the connection density is maximal, black lines
-  delineate boundaries, and overlaid contours show $c>0.5$ (see Movie
+      \cmnt{$N=41$, $\alpha=3.6$, $\beta=16.67$, $k=3$, $D=0.5$, $\gamma\in\pm 2$,
+      $\epsilon=1.2$} and $\delta{t}=0.0001$. Colours indicate the thalamic
+      projection for which the connection density is maximal and black lines
+      delineate boundaries (see Movie
       S1). \textbf{D} Red dots show the Honda-$\delta$ metric obtained from
       simulation approaching that obtained from barrels in \textbf{A} (dotted
-  line); black squares measure the correspondence between the real and
-  simulated barrel shapes, $\eta$ (the product of the sum of squared
-  differences between real and simulated centers and the sum of differences in
-  area; units mm$^4$). \textbf{E} Guidance fields and emergent barrel pattern
-  in a Fgf8 misexpression experiment
-  (c.f. \citealp{assimacopoulos_fibroblast_2012}), simulated by reflecting
-  $\rho_1$ at the join ($\epsilon=80$). All scale bars 1mm.}
+      line); black squares \cmnt{give a} measure \cmnt{of} the correspondence
+      between the real and
+      simulated barrel shapes, $\eta$ (the sum of differences in
+      area; units \cmnt{mm$^2$});
+      \cmnt{grey hexagons show the measure of localization, $\omega =
+        \oiint_{S} \left( \mathrm{max}_i(c_i)/\sum_{i=1}^{N} c_i\right) \mathrm{d}S$. }
+      \textbf{E} \cmnt{The localization measure plotted across the domain
+        shows that the structure in the connection density which has developed
+        by $t=50000$ is largely absent at $t=1000$. The
+        colour map
+        indicates the values of  $\mathrm{max}_i(c_i)/\sum_{i=1}^{N} c_i$. Overlaid
+        contours show $c_i > 0.95 \mathrm{max}_i(c_i)$.} All scale bars 1mm.}
     \label{fig:main}
@@
 at the coordinates of 41 barreloid centers recreated from
-\cite{haidarliu_size_2001} (Fig.\,1B).
+\cite{haidarliu_size_2001} (Fig.\,\ref{fig:main}B).
 
@@
 $a(\mb{x},0)$ and $c(\mb{x},0)$, a clear Voronoi-like tesselation emerged
-(Fig.\,1C; see Movie S1). A reduction in the Honda delta metric (see
+(Fig.\,\ref{fig:main}C; see Movie S1). A reduction in the Honda delta metric (see
 \citealp{senft_mouse_1991}) confirmed that a `good' Voronoi pattern emerged
-within $\approx 3000$ iterations (Fig.\,1D). A measure of the difference
+within \cmnt{$\approx 10000$} iterations (Fig.\,\ref{fig:main}D). A measure of the difference
 in shape between real and simulated barrels revealed a strong correspondence
-(see Fig.\,1D).
+(see Fig.\,\ref{fig:main}D).
 
+% Full page width
 \begin{figure}
@@
 
+% This figure is text width
+\begin{center}
+  \includegraphics[width=\linewidth]{./Fig3.png}
+  \captionof{figure}{Guidance
+    fields and emergent barrel pattern in a Fgf8 misexpression experiment
+    (c.f. \citealp{assimacopoulos_fibroblast_2012}), simulated by reflecting
+    $\rho_1$ from Fig.\,\ref{fig:main}A at the join ($\epsilon=80$).}
+  \label{fig:fgf8}
+\end{center}
+
 To further investigate the interplay of genetic and intrinsic mechanisms we
@@
 barrel fields comprising 2N barrels, i.e., two cortical fields for each
-thalamic projection (Fig.\,1E). Together these results suggest that by the
+thalamic projection (Fig.\,\ref{fig:main}E). Together these results suggest that by the
 misexpression of Fgf8, \cite{shimogori_fibroblast_2005}
@@
 projections compete for a limited number of available synaptic connections
-(Eqs.~\ref{eq:dc}--\ref{eq:da}), ii) \cmnt{review} at each location the branching rate of a
+(Eqs.~\ref{eq:dc}--\ref{eq:da}), ii) at each location the branching rate of a
 given projection is reduced by the density of other projections
-(Eq.\,\ref{eq:comp}), and iii)  \cmnt{review} the branch density of each projection is
-conserved over time (Eq.\,\ref{eq:norm}).
+(Eq.\,\ref{eq:comp}), and iii) \cmnt{REVIEW: No longer true. Try plotting $a$
+  vs $t$} the branch density of each projection is conserved over time.
 
@@
 around the edge of a
-rat barrel field (Fig.\,1A) we set the hex-to-hex distance
+rat barrel field (Fig.\,\ref{fig:main}A) we set the hex-to-hex distance
 $d$ to 0.03\,mm, which resulted in a lattice containing 6515 hexes for the
-simulations shown in Figs.\,1A,C \& D and 12739 hexes for the Fgf8
-misexpression study shown in Fig.\,1E. Each hex contained 82 time-dependent
+simulations shown in Figs.\,\ref{fig:main}A,C \& D and 12739 hexes for the Fgf8
+misexpression study shown in Fig.\,\ref{fig:fgf8}. Each hex contained 82 time-dependent
 variables: 41 branching densities ($a_i$) and 41 connection densities ($c_i$).
@@
 computations described in (i), (ii) and (iii) may be found in the class method
-\code{RD\_James::compute\_divJ()} which calculates \code{term1}, \code{term2}
+\code{RD\_James\cmnt{\_comp2}::compute\_divJ()} which calculates \code{term1}, \code{term2}
 and \code{term3}, respectively.
@@
 
-Movie corresponding to Fig.\,1C in the main paper. Simulation parameters were
-$N=41$, $\alpha=3$, $\beta=20$, $k=3$, $D=0.2$, $\gamma\in\pm 2$,
-$\epsilon=150$ and $\delta{t}=0.0001$. Colours indicate the thalamic
+Movie corresponding to Fig.\,\ref{fig:main}C in the main paper. Simulation parameters were
+$N=41$, $\alpha=3.6$, $\beta=16.67$, $k=3$, $D=0.5$, $\gamma\in\pm 2$,
+$\epsilon=1.2$ and $\delta{t}=0.0001$. Colours indicate the thalamic
 projection for which the connection density is maximal, black lines delineate

### 2020-05-28 analysis/include/BarrelData.py

@@ -264,11 +264,43 @@ class BarrelData:
 
     #
     # New analysis. For each hex, compute a localization variable which is
-    # N*c[i_max] - sum(c[i!=i_max])
+    # N*c[i_max] - sum(c[i!=i_max]) WHERE sum(c) was first normalized
     #
     # Call this function after loading data (being sure to set loadSimData
     # True)
     #
+    def computeLocalization_subtractive (self):
+        if self.loadSimData == False:
+            print ('Error: You have to ensure that you load simulation data by setting loadSimData True before calling load()')
+            return
+        numtimes = np.shape(self.c)[2]
+        self.locn = np.zeros([self.nhex, numtimes], dtype=float)
+        self.locn_vs_t = np.zeros((numtimes), dtype=float)
+        for tidx in range(0, numtimes): # Loop over all times available
+            # This loop runs for each TIME
+            cs = self.c[:,:,tidx]
+            csum = 0.0
+            allhexes_csum = 0.0 # A sum of the "localization measure" over all hexes
+            hnum = int(0);
+            for cc in cs.T:
+                # This loop runs for each HEX
+                imax = int(np.argmax(cc))
+                sum_cc = np.sum(cc)
+                csum = cc[imax] * self.N / sum_cc # do 41 * c_max - sum(c_others)
+                for i in range(0, self.N): # np.shape(self.c)[0] should be 41
+                    if i == imax:
+                        continue
+                    # Have to guard against the odd trash value in cc
+                    if (cc[i] >= 0.0) and (cc[i] <= 1.0):
+                        csum -= (cc[i] / sum_cc)
+                self.locn[hnum, tidx] = csum
+                hnum += 1
+                # Add csum to tcsum
+                allhexes_csum += csum
+            # Now have tcsum for this particular t value; append it to the object:
+            self.locn_vs_t[tidx] = allhexes_csum/float(self.nhex)
+
+    # Localisation measure:
     def computeLocalization (self):
         if self.loadSimData == False:
             print ('Error: You have to ensure that you load simulation data by setting loadSimData True before calling load()')
@@ -285,13 +317,8 @@ class BarrelData:
             for cc in cs.T:
                 # This loop runs for each HEX
                 imax = int(np.argmax(cc))
-                csum = cc[imax] * self.N # do 41 * c_max - sum(c_others)
-                for i in range(0, self.N): # np.shape(self.c)[0] should be 41
-                    if i == imax:
-                        continue
[…]

### 2020-05-28 analysis/include/Surface.py

@@ -115,7 +115,7 @@ class Surface:
     #
     def initFig (self):
         # Create the actual figure.
-        fnt = {'family' : 'DejaVu Sans',
+        fnt = {'family' : 'Arial',
                'weight' : 'regular',
                'size'   : self.fs}
         matplotlib.rc ('font', **fnt)
@@ -265,12 +265,20 @@ class Surface:
             count = 0
             cmap_ = matplotlib.cm.get_cmap('Greys')
 
-            print ('In Surface. domcentres shape: {0}'.format (self.domcentres))
+            #print ('In Surface. domcentres: {0}'.format (self.domcentres))
             for dc in self.domcentres:
                 print('dc: {0}'.format(dc))
 
                 # Compute a greyscale colour for the text from the raw index:
-                cidx = np.float32(count)/np.float32(N)
+                if len(self.z) > 0:
+                    # loop through x and y, checking for index for which x==domcentres.x and y==domcentres.y
+                    cidx = 0
+                    for i in range(0, self.nhex):
+                        if self.x[i] == dc[0] and self.y[i] == dc[1]:
+                            # Hit
+                            cidx = c[i]
+                else:
+                    cidx = np.float32(count)/np.float32(N)
                 # or using the sum of the rgb values
                 cidx = (self.gammaColour_byid[cidx][0] + self.gammaColour_byid[cidx][1] + self.gammaColour_byid[cidx][2]) / 3.0
 
@@ -324,6 +332,11 @@ class Surface:
 
         self.f1.set_aspect ('equal')
         self.F1.tight_layout()
+
+    def addColorBar (self):
+        cb_ax = self.F1.add_axes([0.05, 0.05, 0.05, 0.3])
+        cbar = self.F1.colorbar (plt.cm.ScalarMappable(norm=matplotlib.colors.Normalize(), cmap=self.cmap), cax=cb_ax)
+
 
     def addOuterBoundary (self):
         for i in range(self.nhex):

### 2020-05-28 analysis/plot_cid.py

@@ -85,9 +85,11 @@ def content():
     #sf.addContour (maxc[:,0], 0.4, 'grey', 1.6);
 
     # Or single contour for each field
-    for ii in range(0,bdo.N):
-        c = bdo.c[ii,:,0]
-        sf.addContour (c, 0.5, 'white', 1.0, ii, False);
+    do_contour = 0
+    if do_contour:
+        for ii in range(0,bdo.N):
+            c = bdo.c[ii,:,0]
+            sf.addContour (c, 0.5, 'white', 1.0, ii, False);
 
     sf.addOuterBoundary()
 

### 2020-05-28 analysis/plot_honda_quick.py

@@ -60,7 +60,7 @@ ax2.tick_params (axis='y', labelcolor=col.black)
 ax1.set_xlim ((0,xmax))
 ax1.set_ylim ((0,0.35))
 ax2.set_ylim ((0,1.2))
-ax3.set_ylim ((0,8.5))
+ax3.set_ylim ((0,0.28))
 ax2.set_xlim ((0,xmax))
 
 lw = 2

### 2020-05-28 analysis/plot_locn.py

@@ -9,6 +9,9 @@ import BarrelData as bd
 import plot as pt
 import sebcolour as sc
 
+# IMPORTANT for svg output of text as things that can be edited in inkscape
+plt.rcParams['svg.fonttype'] = 'none'
+
 # Get target x/y hex to show trace for and the time step to show the
 # map for from the arguments:
 if len(sys.argv) < 3:
@@ -77,14 +80,20 @@ sf.plotPoly()
 #sf.addContour (maxc[:,0], 0.4, 'grey', 1.6);
 
 # Or single contour for each field
-for ii in range(0,bdo.N):
-    c = bdo.c[ii,:,0]
-    ccontour = 0.95*np.max(c)
-    sf.addContour (c, ccontour, 'white', 1.0, ii, False);
+do_contours = 1
+if do_contours:
+    for ii in range(0,bdo.N):
+        c = bdo.c[ii,:,0]
+        ccontour = 0.95*np.max(c)
+        sf.addContour (c, ccontour, 'white', 1.0, ii, False);
 
 sf.addOuterBoundary()
 
 mapname = 'plots/{0}_locn_{1:06d}.png'.format(os.path.basename(logdirname), ti)
+plt.savefig (mapname, dpi=300, transparent=True)
+
+sf.addColorBar()
+mapname = 'plots/{0}_locn_{1:06d}.svg'.format(os.path.basename(logdirname), ti)
 plt.savefig (mapname, dpi=300, transparent=True)
 
 plt.show()

### 2020-05-28 analysis/plot_paramsearch_paper.py

@@ -37,7 +37,7 @@ F1 = plt.figure (figsize=(20,10))
 pp.paramplot_pub (sdata, F1, 'alphabeta', 'D', 'F', 3, timepoint, param_tuples, 0, 0, 'viridis');
 plt.savefig('plots/paramsweep.svg')
 
-do_maps = 0
+do_maps = 1
 if do_maps:
     # map plots to go under the param sweep colour grids
     map_i = 1 # map index
@@ -89,9 +89,9 @@ if do_maps:
         sf.textid = False
         sf.plotPoly()
         # A single contour for each field
-        for ii in range(0,bdo.N):
-            c = bdo.c[ii,:,0]
-            sf.addContour (c, 0.5, 'white', 1.0, ii, False);
+        #for ii in range(0,bdo.N):
+        #    c = bdo.c[ii,:,0]
+        #    sf.addContour (c, 0.5, 'white', 1.0, ii, False);
         # The nice grey outer boundary
         sf.addOuterBoundary()
 

### 2020-05-28 analysis/process_honda.py

@@ -80,7 +80,6 @@ print ('t shape {0}, t1_masked shape {1}'.format(np.shape(bdo.t),np.shape(t1_mas
 
 sos_dist = sos_dist.compressed()
 hondadelta = hondadelta.compressed()
-area_diff = area_diff.compressed()
 
 print ('sos_dist shape: {0}, area_diff shape: {1}'.format (np.shape (sos_dist), np.shape (area_diff[:,0])))
 

### 2020-05-29 paper/briefpaper/brief.tex
Notes about conservation of a

@@
 given projection is reduced by the density of other projections
-(Eq.\,\ref{eq:comp}), and iii) \cmnt{REVIEW: No longer true. Try plotting $a$
-  vs $t$} the branch density of each projection is conserved over time.
+(Eq.\,\ref{eq:comp}), and iii) \cmnt{REVIEW: plotting $a(t)$, $c(t)$ and
+  $a(t)-c(t)$ showed that there was some leaking of $a$ going on at the
+  boundary; increasing the boundary fall off distance to 3 times hex to hex
+  distance reduces the leaking to a negligible amount. So the branchdesnity of
+each projections IS conserved over time.} the branch density of
+each projection is conserved over time.
 

### 2020-06-01 paper/briefpaper/brief.tex
Figure name refactoring
More text for paper

--- 2020-05-28 paper/briefpaper/brief.tex
@@
   \begin{fullwidth}
-    %\includegraphics[width=\linewidth]{./MainFig.png}
-    \includegraphics[width=\linewidth]{./MainFig_superlowres.png}
+    \includegraphics[width=\linewidth]{./Fig1.png}
     \caption{\textbf{A} Left shows a cytochrome oxidase stain obtained from rat S1
@@
       S1). \textbf{D} Red dots show the Honda-$\delta$ metric obtained from
-      simulation approaching that obtained from barrels in \textbf{A} (dotted
+      simulation approaching that obtained from \cmnt{real} barrels in \textbf{A} (dotted
       line); black squares \cmnt{give a} measure \cmnt{of} the correspondence
@@
         shows that the structure in the connection density which has developed
-        by $t=50000$ is largely absent at $t=1000$. The
+        by $t=30000$ is largely absent at $t=1000$. The
         colour map
@@
 \citealp{senft_mouse_1991}) confirmed that a `good' Voronoi pattern emerged
-within \cmnt{$\approx 10000$} iterations (Fig.\,\ref{fig:main}D). A measure of the difference
-in shape between real and simulated barrels revealed a strong correspondence
-(see Fig.\,\ref{fig:main}D).
+within \cmnt{$\approx 20000$} iterations (Fig.\,\ref{fig:main}D). A measure of the difference
+in \cmnt{barrel area} between real and simulated barrels revealed a strong correspondence
+(see Fig.\,\ref{fig:main}D). \cmnt{In a chosen region, the ratio of the connection density
+  of the TC type $i$ which makes the densest connections to the sum of all
+  connection densities gives a measure of the localization of the connection
+density. We computed this measure across the entire barrel domain and named it $\omega$
+(Fig.\,\ref{fig:main}D). $\omega$ increases during the first $20000$ simulation
+steps and converges on a stable value. Fig.\,{fig:main}E shows
+the localization measure plotted as a luminance map, giving a more graded
+impression of the structure which develops in the system.}
 
@@
   \begin{fullwidth}
-    \includegraphics[width=\linewidth]{./paramsweep.png}
+    \includegraphics[width=\linewidth]{./Fig2.png}
     \caption{\cmnt{\textbf{A} Parameter sweep caption goes here}}
@@
 
-\cmnt{Parameter sweep paragraph goes here.}
+\cmnt{To determine the robustness of the model to its parameters, we
+  investigated its behaviour with respect to changes in the parameters $D$,
+  $\epsilon$ and the ratio of $\alpha$ to $\beta$. Fig.\,\ref{fig:paramsweep}A
+  shows the value of the metrics $\delta$, $\eta$ and $\omega$ after $30000$
+  iterations plotted as colour maps for $180$ parameter sets. Each row of maps
+  shows the results for a separate metric. Each column of maps shows
+  results for a different value of the ratio $\alpha / \beta$. The diffusion
+  parameter, $D$ varies with the $x$ axes and the strength of competition,
+  $\epsilon$ varies with the $y$ axes. The colour maps are arranged so that
+  `yellow is good'. Lower values of $\delta$ indicate good Voroni patterns and
+  are yellow. Lower values of the difference between real and simulated barrel
+  area indicates good pattern matching. Higher values of $\omega$ indicate
+  better localization and are coloured yellow. White squares indicate that the
+  simulation was not stable for that parameter set.}
+
+\cmnt{The maps indicate that the pattern is not strongly influenced by the
+  choice of $\alpha/\beta$, which was varied by 5 orders of magnitude. The
+  most Voronoi-like patterns were achieved for parameter sets for which $D$
+  and $\epsilon$ were high. The barrel area metric, $\eta$ indicates a strong
+  preference for $D\approx0.5$, although this does not give the maximally
+  localized solution, which for $D$ should be small. Thus, there is a
+  trade-off between good localization and good Voroni pattern conformation. We
+  chose the parameter set indicated by the asterisk in
+  Fig.\,\ref{fig:paramsweep}A to present in Fig.\,\ref{fig:main} because it
+  provides the best barrel area correspondence and produces a good Voronoi
+  pattern. In Fig.\,\ref{fig:paramsweep}B, we present some patterns
+  representing alternative parameter choices. Fig.\,\ref{fig:paramsweep}B i)
+  shows the simulation result with low diffusion constant, $D$. Reducing the
+  diffusion constant is equivalent to increasing the size of the domain $S$
+  and so this simulation can be thought to represent an animal with a large
+  cortical area but with unchanged cortical complexity. In this simulation, a
+  large proportion of the domain is occupied by the projections with
+  intermediate interaction parameters; those with strong interaction
+  parameters are bunched up towards the edge of the domain. A barrel pattern
+  fails to form. Fig.\,\ref{fig:paramsweep}B ii) and iii) indicate the effect
+  of varying $D$ in either direction from its optimum. ii) is a less
+  exaggerated version of the pattern in i), whereas iii) shows an expansion of
+  the outer barrels and a corresponding compression of the barrels within.}
 
@@
 barrel fields comprising 2N barrels, i.e., two cortical fields for each
-thalamic projection (Fig.\,\ref{fig:main}E). Together these results suggest that by the
+thalamic projection (Fig.\,\ref{fig:fgf8}). Together these results suggest that by the
 misexpression of Fgf8, \cite{shimogori_fibroblast_2005}
@@
 given projection is reduced by the density of other projections
-(Eq.\,\ref{eq:comp}), and iii) \cmnt{REVIEW: No longer true. Try plotting $a$
-  vs $t$} the branch density of each projection is conserved over time.
+(Eq.\,\ref{eq:comp}), and iii) the branch density of each projection is
+conserved over time.
 

### 2020-06-02 paper/briefpaper/brief.tex
Parameter sweep caption and methods updates
Methods needed updating with the newer competition terms.

@@
         indicates the values of  $\mathrm{max}_i(c_i)/\sum_{i=1}^{N} c_i$. Overlaid
-        contours show $c_i > 0.95 \mathrm{max}_i(c_i)$.} All scale bars 1mm.}
+        contours show $c_i > 0.95 \mathrm{max}_i(c_i)$.} All scale bars 1\,mm.}
     \label{fig:main}
@@
   \color{blue}
-  \chi_i(\mb{x}, t) = \frac{\epsilon a_i}{N-1} \nabla \textstyle\sum_{j\ne i}{a_j}
+  \chi_i(\mb{x}, t) = \frac{\epsilon a_i}{N-1} \nabla \textstyle\sum_{j\ne i}^{N}{a_j}
   \color{black}
@@
     \includegraphics[width=\linewidth]{./Fig2.png}
-    \caption{\cmnt{\textbf{A} Parameter sweep caption goes here}}
+    \caption{\cmnt{\textbf{A} Exploring the parameter space of the model given
+        in Eqs.\,1--3. Graphs show the three pattern metrics for $t=30000$
+        steps as colourmaps: Honda $\delta$ (top row) shows conformance to a
+        Voronoi pattern; $\eta$ (middle row) gives the simulated/experimental
+        barrel area difference and $\omega$ (bottom row) is a measure of
+        localization (separation) of the connection densities for the 41
+        thalamocortical axon types. For each row, lighter (orange/yellow)
+        colours imply a better model. The parameter space explored is three
+        dimensional with the parameter ratio $\alpha/\beta$ varied along the
+        columns of graphs, the competition parameter $\epsilon$ varied on the
+        $y$-axes and the diffusion constant, $D$ varied along the $x$-axes. An
+        asterisk (*) marks the parameter set displayed in
+        Fig.\,\ref{fig:main}. Boxes i), ii) and iii) mark parameter sets for which
+        corresponding patterns are shown in \textbf{B} (for $t=30000$ steps).}}
     \label{fig:paramsweep}
@@
     (c.f. \citealp{assimacopoulos_fibroblast_2012}), simulated by reflecting
-    $\rho_1$ from Fig.\,\ref{fig:main}A at the join ($\epsilon=80$).}
+    $\rho_1$ from Fig.\,\ref{fig:main}A at the join. \cmnt{The model
+      parameters match those in Fig.\,\ref{fig:main}C. Scale
+      bars are 1\,mm.}}
   \label{fig:fgf8}
@@
   \label{eq:divJ}
-  \nabla\vcdot\mb{J}_i(\mb{x},t) = \nabla\vcdot\left(D \nabla a_i-a_i\sum_{j=1}^{M} \gamma_{i,j}\nabla \rho_j(\mb{x})\right).
+  \nabla\vcdot\mb{J}_i(\mb{x},t) = \nabla\vcdot\left(D \nabla
+  a_i-a_i\sum_{j=1}^{M} \gamma_{i,j}\nabla \rho_j(\mb{x}) \color{blue}+
+  \frac{\epsilon a_i}{N-1} \nabla \hat{a}_i \color{black}\right),
 \end{equation}
 %
-Note that the sum of the guidance gradients is time-independent and define
-$\mb{g}_i(\mb{x}) \equiv \sum_{j=1}^{M} \gamma_{i,j} \nabla\rho_j(\mb{x})$.
-Because the divergence operator is distributive, Eq.\,\ref{eq:divJ} can be
-expanded using vector calculus identities (dropping references to $\mb{x}$ and
-$t$ for clarity):
+\cmnt{where we have defined $\hat{a}_i=\sum_{j\ne i}^{N}a_j$.} Note that the sum of the
+guidance gradients is time-independent and define $\mb{g}_i(\mb{x}) \equiv
+\sum_{j=1}^{M} \gamma_{i,j} \nabla\rho_j(\mb{x})$.  Because the divergence
+operator is distributive, Eq.\,\ref{eq:divJ} can be expanded using vector
+calculus identities (dropping references to $\mb{x}$ and $t$ for clarity):
 %
 \begin{equation}
-\nabla\vcdot\mb{J}_i = \nabla\vcdot\big(D \nabla a_i\big) - \nabla\vcdot\big(a_i \mb{g}_i\big).
+\nabla\vcdot\mb{J}_i = \nabla\vcdot\big(D \nabla a_i\big) -
+\nabla\vcdot\big(a_i \mb{g}_i\big) \color{blue} + \frac{\epsilon}{N-1}\nabla\vcdot\big({a}_i\nabla\hat{a}_i\big)\color{black}.
 \end{equation}
@@
 \mb{g}_i\vcdot\nabla a_i % term3
+\color{blue}
++ \frac{\epsilon a_i}{N-1} \dvrg \hat{a}_i % term1_1
++ \frac{\epsilon}{N-1} \nabla \hat{a}_i\vcdot\nabla{a}_i% term1_2
+\color{black}
 ,
@@
 %
-which has three elements to compute: i)~$D \dvrg a_i$ (the Laplacian of
+which has \cmnt{five} elements to compute: i)~$D \dvrg a_i$ (the Laplacian of
 $a_i$); ii)~a time-independent modulator of $a_i$ (because
-$\nabla\vcdot\mb{g}_i$ is a time-independent static field); and iii)~the
-scalar product of the static vector field $\mb{g}_i$ and the gradient of
-$a_i$. Each of the divergences can be simplified by means of Gauss's Theorem
-following \cite{lee_hexagonal_2014}.
+$\nabla\vcdot\mb{g}_i$ is a time-independent static field); iii)~the scalar
+product of the static vector field $\mb{g}_i$ and the gradient of $a_i$;
+\cmnt{iv)~the Laplacian of $\hat{a}_i$; and v)~a term involving the gradients
+  of $a_i$ and $\hat{a}_i$}. Each of the divergences can be simplified by
+means of Gauss's Theorem following \cite{lee_hexagonal_2014}.
 
@@
 \begin{split}
-\langle D \dvrg a_i(\mb{p}_0,t) \rangle  = \frac{1}{\Omega} \oiint_\Omega \dvrg a_i(\mb{x}, t) \dif\Omega & = \frac{1}{\Omega} \oint \frac{\partial a_i}{\partial \hat{\mb{n}}} \dif\gamma \\
-& \approx \frac{1}{\Omega} \sum_{j=1}^6 \frac{\partial a_i(\mb{p}_j)}{\partial \hat{\mb{n}}} \bigg\rvert_{\mathrm{mid}} v \\
-& = \frac{2}{\sqrt{3} d^2} \sum_{j=1}^6 \frac{a_i(\mb{p}_j) - a_i(\mb{p}_0)}{d} \frac{d}{\sqrt{3}} \\
-& = \frac{2}{3 d^2} \sum_{j=1}^6 \big(a_i(\mb{p}_j) - a_i(\mb{p}_0)\big),
+\langle D \dvrg a_i(\mb{p}_0,t) \rangle  = \frac{D}{\Omega} \oiint_\Omega \dvrg a_i(\mb{x}, t) \dif\Omega & = \frac{D}{\Omega} \oint \frac{\partial a_i}{\partial \hat{\mb{n}}} \dif\gamma \\
+& \approx \frac{D}{\Omega} \sum_{j=1}^6 \frac{\partial a_i(\mb{p}_j)}{\partial \hat{\mb{n}}} \bigg\rvert_{\mathrm{mid}} v \\
+& = \frac{2D}{\sqrt{3} d^2} \sum_{j=1}^6 \frac{a_i(\mb{p}_j) - a_i(\mb{p}_0)}{d} \frac{d}{\sqrt{3}} \\
+& = \frac{2D}{3 d^2} \sum_{j=1}^6 \big(a_i(\mb{p}_j) - a_i(\mb{p}_0)\big),
 \end{split}
@@
 
+\noindent
 ii) The computation of the second term in Eq.\,\ref{eq:divJExpanded},
@@
 
-iii) The final term in Eq.\,\ref{eq:divJExpanded} is the scalar product of two
+\noindent
+iii) The \cmnt{middle} term in Eq.\,\ref{eq:divJExpanded} is the scalar product of two
 vector fields which is straightforward to compute from their Cartesian
@@
 
-By separating the computation of Eq.\,\ref{eq:divJ} into parts (i), (ii) \&
-(iii), the no-flux boundary condition,
-$\mb{J}_i(\mb{x},t)\big\rvert_{\mathrm{boundary}} = 0$, can be fulfilled. On
+\noindent
+\cmnt{iv) The same method used to compute $\dvrg a_i$ in term (i) is used to compute
+$\dvrg \hat{a}_i$.}
+
+\noindent
+\cmnt{v) The final term is the scalar product of the two vector fields $\nabla
+  a_i$ and $\nabla \hat{a}_i$.}
+
+By separating the computation of Eq.\,\ref{eq:divJ} into parts (i)--(v), the
+no-flux boundary condition,
+%
+\begin{equation}
+  \label{eq:noflux}
+  \mb{J}_i(\mb{x},t)\big\rvert_{\mathrm{boundary}} = 0,
+\end{equation}
+%
+can be fulfilled. On
 the boundary, the contribution to $\mb{J}$ resulting from the first term of
@@
 ensuring that the second and third terms of Eq.\,\ref{eq:divJExpanded} also
-contribute nothing to $\mb{J}$. This is achieved by applying to
-$\mb{g}_i(\mb{x})$ a sharp logistic function of the distance from $\mb{x}$ to
-the boundary.
+contribute nothing to $\mb{J}$. This is achieved by \cmnt{multiplying}
+$\mb{g}_i(\mb{x})$ \cmnt{by} a sharp logistic function of the distance, $d_b$,
+from $\mb{x}$ to the boundary, \cmnt{of the form $1/[1 +
+  \exp(100(d_f-d_b))]$, where $d_f=0.1$\,mm is the boundary fall-off distance,
+  chosen to be approximately $3d$}.
 
 All code required to reproduce these results is available at
-\url{https://github.com/ABRG-Models/BarrelEmerge/tree/eLife}. The
-computations described in (i), (ii) and (iii) may be found in the class method
-\code{RD\_James\cmnt{\_comp2}::compute\_divJ()} which calculates \code{term1}, \code{term2}
-and \code{term3}, respectively.
+\url{https://github.com/ABRG-Models/BarrelEmerge/tree/eLife}. The computations
+described in (i)--(v) may be found in the class method
+\code{RD\_James\cmnt{\_comp2}::compute\_divJ()} which calculates \code{term1},
+\code{term2}, \code{term3}, \code{term1\_1} and \code{term1\_2}, respectively.
 
@@
 boundaries, and overlaid contours show $c>0.5$. The final frame in the movie
-is step 25,000 of the simulation.
+is step $30000$ of the simulation.
 

### 2020-06-02 paper/briefpaper/brief.tex
Tweaks on re-reading

@@
 in \cmnt{barrel area} between real and simulated barrels revealed a strong correspondence
-(see Fig.\,\ref{fig:main}D). \cmnt{In a chosen region, the ratio of the connection density
-  of the TC type $i$ which makes the densest connections to the sum of all
-  connection densities gives a measure of the localization of the connection
+(see Fig.\,\ref{fig:main}D, squares). \cmnt{In a chosen region, the ratio of \emph{the connection density
+  of the TC type $i$ which makes the densest connections} to \emph{the sum of all
+  connection densities} gives a measure of the localization of the connection
 density. We computed this measure across the entire barrel domain and named it $\omega$
-(Fig.\,\ref{fig:main}D). $\omega$ increases during the first $20000$ simulation
-steps and converges on a stable value. Fig.\,{fig:main}E shows
+(see Fig.\,\ref{fig:main}D, grey hexagons). $\omega$ increases during the first $20000$ simulation
+steps and converges on a stable value. Fig.\,\ref{fig:main}E shows
 the localization measure plotted as a luminance map, giving a more graded
@@
   localized solution, which for $D$ should be small. Thus, there is a
-  trade-off between good localization and good Voroni pattern conformation. We
-  chose the parameter set indicated by the asterisk in
-  Fig.\,\ref{fig:paramsweep}A to present in Fig.\,\ref{fig:main} because it
+  trade-off between good localization and good Voroni pattern
+  conformation. The
+  parameter set indicated by the asterisk in
+  Fig.\,\ref{fig:paramsweep}A is presented in Fig.\,\ref{fig:main} because it
   provides the best barrel area correspondence and produces a good Voronoi
@@
 %
-\cmnt{where we have defined $\hat{a}_i=\sum_{j\ne i}^{N}a_j$.} Note that the sum of the
+\cmnt{where $\hat{a}_i\equiv\sum_{j\ne i}^{N}a_j$.} Note that the sum of the
 guidance gradients is time-independent and define $\mb{g}_i(\mb{x}) \equiv

### 2020-06-02 paper/briefpaper/brief.tex
6*6*6 is not 180!

@@
   shows the value of the metrics $\delta$, $\eta$ and $\omega$ after $30000$
-  iterations plotted as colour maps for $180$ parameter sets. Each row of maps
+  iterations plotted as colour maps for $216$ parameter sets. Each row of maps
   shows the results for a separate metric. Each column of maps shows

### 2020-06-02 paper/briefpaper/brief.tex
Adds expanded section on how the gammas were set

@@
 thalamocortical projections for a limited availability of cortical
-connections. To introduce competition also in terms of axon branching we
+connections. To introduce competition also in terms of axon branching,
+\cmnt{whilst ensuring that $a_i$ is conserved over time,} we
 redefined
@@
   additional diffusion-like term which makes the branching density for a given
-  projection `flow away' from regions} where the branches of other projections
+  projection diffuse away from regions} where the branches of other projections
 become dense. \cmnt{$\epsilon$ parameterizes the strength of this competitive
@@
 determined by the concentration of a similar molecule at its thalamic origin,
-i.e., the putative barreloid center. As such, two orthogonal linear thalamic
-gradients were defined, from which 41 pairs of $\gamma$ values were sampled,
-at the coordinates of 41 barreloid centers recreated from
-\cite{haidarliu_size_2001} (Fig.\,\ref{fig:main}B).
+i.e., the putative barreloid center. \cmnt{Thus, we suggest that there exist a
+  pair of signalling molecules whose expression varies linearly (with
+  orthogonal gradients) across the thalamus, analgous to the ephrins, whose
+  existence is well known in the cortex. Thalamocortical axons growing from a
+  region in the thalamus with high expression of one of these molecules will
+  interact with a corresponding cortical signalling molecule more strongly
+  than an axon emanating from a region with low expression of the thalamic
+  signal.}  \cmnt{We defined two orthogonal, linear gradients oriented in the
+  anterior-posterior and medial-lateral directions and used the coordinates of
+  41 barreloid centers recreated from \cite{haidarliu_size_2001},
+  corresponding to the macrovibrissae, to populate 41 pairs of $\gamma$
+  values.}  (Fig.\,\ref{fig:main}B). \cmnt{In other words, we simply scaled
+  the barreloid coordinates in an anterior-posterior/medial-lateral coordinate
+  system to create the $\gamma$ values. Note that positive \emph{and} negative
+  $\gamma$ values were generated; the origin of the coordinate system was
+  located at the center of the thalamic barreloid field. This prevented an
+  overall drift of the cortical pattern away from the centroid of the barrel
+  field and may explain why coordinate-giving transcription factors and the
+  ephrins are often found as pairs acting in opposing directions.}
+\cmnt{We concentrated on macrovibrissae (excluding microvibrissae) because it
+  is easy to identify their corresponding structures both in the thalamus and
+  in the cortex. We also excluded the 5 rhinal whiskers because their cortical
+  representation is in a separate island, distinct, and somewhat removed from
+  the representation of the macrovibirissae.}
 
-A cortical boundary enclosing 41 corresponding barrels was traced from a
-cytochrome oxidase stain from \cite{zheng_signal_2001}, and
-Eqs.~\ref{eq:dc}--\ref{eq:comp} were solved for $N=41$ projections on the
-resulting domain, using $M=2$ linear signalling gradients aligned with the
-anterior-posterior and medial-lateral axes. From random initial conditions for
-$a(\mb{x},0)$ and $c(\mb{x},0)$, a clear Voronoi-like tesselation emerged
-(Fig.\,\ref{fig:main}C; see Movie S1). A reduction in the Honda delta metric (see
-\citealp{senft_mouse_1991}) confirmed that a `good' Voronoi pattern emerged
-within \cmnt{$\approx 20000$} iterations (Fig.\,\ref{fig:main}D). A measure of the difference
-in \cmnt{barrel area} between real and simulated barrels revealed a strong correspondence
-(see Fig.\,\ref{fig:main}D, squares). \cmnt{In a chosen region, the ratio of \emph{the connection density
-  of the TC type $i$ which makes the densest connections} to \emph{the sum of all
-  connection densities} gives a measure of the localization of the connection
-density. We computed this measure across the entire barrel domain and named it $\omega$
-(see Fig.\,\ref{fig:main}D, grey hexagons). $\omega$ increases during the first $20000$ simulation
-steps and converges on a stable value. Fig.\,\ref{fig:main}E shows
-the localization measure plotted as a luminance map, giving a more graded
-impression of the structure which develops in the system.}
+A cortical boundary enclosing 41 \cmnt{macrovibrissae} barrels was traced from
+a cytochrome oxidase stain from \cite{zheng_signal_2001} \cmnt{(using the
+  authors' original data)}, and Eqs.~\ref{eq:dc}--\ref{eq:comp} were solved
+for $N=$ 41 projections on the resulting domain, using $M=2$ linear signalling
+gradients aligned with the anterior-posterior and medial-lateral axes. From
+random initial conditions for $a(\mb{x},0)$ and $c(\mb{x},0)$, a clear
+Voronoi-like tesselation emerged (Fig.\,\ref{fig:main}C; see Movie S1). A
+reduction in the Honda delta metric (see \citealp{senft_mouse_1991}) confirmed
+that a `good' Voronoi pattern emerged within \cmnt{$\approx 20000$} iterations
+(Fig.\,\ref{fig:main}D).  A measure of the difference in \cmnt{barrel area}
+between real and simulated barrels revealed a strong correspondence (see
+Fig.\,\ref{fig:main}D, squares). \cmnt{In a chosen region, the ratio of
+  \emph{the connection density of the TC type $i$ which makes the densest
+    connections} to \emph{the sum of all connection densities} gives a measure
+  of the localization of the connection density. We computed this measure
+  across the entire barrel domain and named it $\omega$ (see
+  Fig.\,\ref{fig:main}D, grey hexagons). $\omega$ increases during the first
+  $20000$ simulation steps and converges on a stable
+  value. Fig.\,\ref{fig:main}E shows the localization measure plotted as a
+  luminance map, giving a more graded impression of the structure which
+  develops in the system.}
 
@@
 effectively created one large barrel field rather than two distinct cortical
-fields.
+fields. \cmnt{Add something about the slightly off-vertical arrangement of the
+  fields.}
 

### 2020-06-02 paper/briefpaper/brief.tex
Minor tweaks to text

@@
   anterior-posterior and medial-lateral directions and used the coordinates of
-  41 barreloid centers recreated from \cite{haidarliu_size_2001},
+  $41$ barreloid centers recreated from \cite{haidarliu_size_2001},
   corresponding to the macrovibrissae, to populate 41 pairs of $\gamma$
@@
   authors' original data)}, and Eqs.~\ref{eq:dc}--\ref{eq:comp} were solved
-for $N=$ 41 projections on the resulting domain, using $M=2$ linear signalling
+for $N=41$ projections on the resulting domain, using $M=2$ linear signalling
 gradients aligned with the anterior-posterior and medial-lateral axes. From
@@
 reduction in the Honda delta metric (see \citealp{senft_mouse_1991}) confirmed
-that a `good' Voronoi pattern emerged within \cmnt{$\approx 20000$} iterations
+that a `good' Voronoi pattern emerged within \cmnt{$\approx$~20000} iterations
 (Fig.\,\ref{fig:main}D).  A measure of the difference in \cmnt{barrel area}
@@
   Fig.\,\ref{fig:main}D, grey hexagons). $\omega$ increases during the first
-  $20000$ simulation steps and converges on a stable
+  20000 simulation steps and converges on a stable
   value. Fig.\,\ref{fig:main}E shows the localization measure plotted as a

### 2020-06-02 boundaries/barreloids/barreloids_haidarliu.py

@@ -1,5 +1,5 @@
-# Read barreloids_vanderloos_dia_only.csv and graph, along with
-# labels. Output 4 sets of gammas assuming two orthogonal signals.
+# Read barreloids_haidarliu.csv and graph, along with labels. Output 2 sets of
+# gammas assuming two orthogonal signals which can be +ve or -ve.
 
 import numpy as np
 import matplotlib

### 2020-06-02 boundaries/barreloids/barreloids_haidarliu_Fig5d.py

@@ -1,5 +1,6 @@
-# Read barreloids_vanderloos_dia_only.csv and graph, along with
-# labels. Output 4 sets of gammas assuming two orthogonal signals.
+# Read barreloids_haidarliu_Fig5d.csv and graph, along with labels. Output 2
+# sets of gammas assuming two orthogonal signals which can be +ve or -ve. This
+# is as used in the eLife submission.
 
 import numpy as np
 import matplotlib

### 2020-06-03 paper/briefpaper/brief.tex
Results of running model 10 times to get variation in outputs based on noisy initial conditions

@@
   anterior-posterior and medial-lateral directions and used the coordinates of
-  $41$ barreloid centers recreated from \cite{haidarliu_size_2001},
+  41 barreloid centers recreated from \cite{haidarliu_size_2001},
   corresponding to the macrovibrissae, to populate 41 pairs of $\gamma$
@@
   the barreloid coordinates in an anterior-posterior/medial-lateral coordinate
-  system to create the $\gamma$ values. Note that positive \emph{and} negative
-  $\gamma$ values were generated; the origin of the coordinate system was
-  located at the center of the thalamic barreloid field. This prevented an
-  overall drift of the cortical pattern away from the centroid of the barrel
-  field and may explain why coordinate-giving transcription factors and the
-  ephrins are often found as pairs acting in opposing directions.}
-\cmnt{We concentrated on macrovibrissae (excluding microvibrissae) because it
-  is easy to identify their corresponding structures both in the thalamus and
-  in the cortex. We also excluded the 5 rhinal whiskers because their cortical
-  representation is in a separate island, distinct, and somewhat removed from
-  the representation of the macrovibirissae.}
+  system to create the $\gamma$ values in the range [-2,2]. Note that positive
+  \emph{and} negative $\gamma$ values were generated; the origin of the
+  coordinate system was located at the approximate center of the thalamic
+  barreloid field. This prevented an overall drift of the cortical pattern
+  away from the centroid of the barrel field and may explain why
+  coordinate-giving transcription factors and the ephrins are often found as
+  pairs acting in opposing directions.}  \cmnt{We concentrated on
+  macrovibrissae (excluding microvibrissae) because it is easy to identify
+  their corresponding structures both in the thalamus and in the cortex. We
+  also excluded the 5 rhinal whiskers because their cortical representation is
+  in a separate island, distinct, and somewhat removed from the representation
+  of the macrovibirissae.}
 
@@
 gradients aligned with the anterior-posterior and medial-lateral axes. From
-random initial conditions for $a(\mb{x},0)$ and $c(\mb{x},0)$, a clear
-Voronoi-like tesselation emerged (Fig.\,\ref{fig:main}C; see Movie S1). A
-reduction in the Honda delta metric (see \citealp{senft_mouse_1991}) confirmed
-that a `good' Voronoi pattern emerged within \cmnt{$\approx$~20000} iterations
+random initial conditions for $a(\mb{x},0)$, \cmnt{with each hexagonal element
+  of $a_i$ set from a uniform distribution according to $0.2+{U}(0,0.2)$, and
+  $c(\mb{x},0)=0$}, a clear Voronoi-like tesselation
+emerged (Fig.\,\ref{fig:main}C; see Movie S1). A reduction in the Honda delta
+metric (see \citealp{senft_mouse_1991}) confirmed that a `good' Voronoi
+pattern emerged within \cmnt{$\approx$~20000} iterations
 (Fig.\,\ref{fig:main}D).  A measure of the difference in \cmnt{barrel area}
@@
         asterisk (*) marks the parameter set displayed in
-        Fig.\,\ref{fig:main}. Boxes i), ii) and iii) mark parameter sets for which
+        Fig.\,\ref{fig:main}. Boxes i), ii)~and iii)~mark parameter sets for which
         corresponding patterns are shown in \textbf{B} (for $t=30000$ steps).}}
@@
   parameters are bunched up towards the edge of the domain. A barrel pattern
-  fails to form. Fig.\,\ref{fig:paramsweep}B ii) and iii) indicate the effect
-  of varying $D$ in either direction from its optimum. ii) is a less
-  exaggerated version of the pattern in i), whereas iii) shows an expansion of
+  fails to form. Fig.\,\ref{fig:paramsweep}B ii)~and iii)~indicate the effect
+  of varying $D$ in either direction from its optimum. ii)~is a less
+  exaggerated version of the pattern in i), whereas iii)~shows an expansion of
   the outer barrels and a corresponding compression of the barrels within.}
+
+\cmnt{We performed a sensitivity analysis to determine if the quality of the
+  pattern was sensitive to any of the following inputs: i)~The angle and
+  magnitude of each of the two guidance fields (Fig.\,\ref{fig:main}A);
+  ii)~the magnitude, offset and characteristic length scale of the noise
+  applied to $a_i$ at $t=0$; and iii)~the magnitude of each of the interaction
+  parameters, $\gamma_{i,j}$. We analysed the effect on the three pattern
+  metrics, $\delta$, $\eta$ and $\omega$ evaluated after 30000 simulation
+  steps. First, we conducted 10 runs of the simulation to evaluate the mean
+  and standard deviation (as a result of the different random initial
+  conditions for $a_i$) of the three metrics in the default model
+  (Fig.\,\ref{fig:main}C--E) which were $\delta=0.089\pm{0.004}$,
+  $\eta=0.531\pm{0.005}$ and $\omega=0.2165\pm{0.0001}$). These results are
+  not strongly affected by the random initial conditions for $a_i$ but give a
+  baseline from which to evaluate whether changes in the other inputs have a
+  statistically significant effect.}
 
@@
 The present results suggest that the key requirements for the emergence of
-realistic barrel patterning are i) at each cortical location thalamocortical
+realistic barrel patterning are i)~at each cortical location thalamocortical
 projections compete for a limited number of available synaptic connections
-(Eqs.~\ref{eq:dc}--\ref{eq:da}), ii) at each location the branching rate of a
+(Eqs.~\ref{eq:dc}--\ref{eq:da}), ii)~at each location the branching rate of a
 given projection is reduced by the density of other projections
-(Eq.\,\ref{eq:comp}), and iii) the branch density of each projection is
+(Eq.\,\ref{eq:comp}), and iii)~the branch density of each projection is
 conserved over time.

### 2020-06-03 analysis/include/BarrelData.py

@@ -139,12 +139,13 @@ class BarrelData:
         # experimental barrels and the simulated barrels. len(t)
         self.sos_dist = np.array([])
 
-        # The 'map difference'. if Hex n expt. ID != Hex n sim. ID, then
-        # add one to this, len(t)
+        # The 'map difference'. if Hex n expt. ID != Hex n sim. ID, then add
+        # one to this, len(t). It's the proportion of hexes which are different
+        # in ID to the experimental map. Lower is better. Range [0,1].
         self.mapdiff = np.array([])
 
-        # The sum of the absolute difference in areas of each
-        # experimental domain and simulated domain. len(t)
+        # The sum (in hexes) of the absolute difference in areas of each
+        # experimental domain and simulated domain. len(t). Range [0,nhex].
         self.area_diff = np.array([])
 
         # The divisions between domains. All those paths that the C++ code
@@ -164,6 +165,7 @@ class BarrelData:
         self.meanbeta=0
         self.meanepsilon=0
 
+        self.debug = False
     #
     # Create self.t and self.t_steps, first checking to see if they've
     # already been created.
@@ -176,7 +178,8 @@ class BarrelData:
         else:
             t_cmp = np.zeros([numtimes], dtype=int)
             if np.shape(t_cmp) == np.shape(self.t_steps):
-                print ('dirich_*.h5 files and c_*.h5 files match up.')
+                if self.debug:
+                    print ('dirich_*.h5 files and c_*.h5 files match up.')
             else:
                 print ('WARNING: dirich_*.h5 files and c_*.h5 files DO NOT match up. Re-creating t...')
                 self.t_steps = np.zeros([numtimes], dtype=int)
@@ -198,7 +201,8 @@ class BarrelData:
                 self.F = 0
             self.k = jd["k"]
             tc = jd["tc"]
-            print ('tc length: {0}'.format(len(jd["tc"])))
+            if self.debug:
+                print ('tc length: {0}'.format(len(jd["tc"])))
             self.N = len(jd["tc"])
             sumalpha = 0
             sumbeta = 0
@@ -207,7 +211,10 @@ class BarrelData:
                 # Do a key-value thing
                 sumalpha = sumalpha + i["alpha"]
                 sumbeta = sumbeta + i["beta"]
-                sumepsilon = sumepsilon + i["epsilon"]
+                try:
+                    sumepsilon = sumepsilon + i["epsilon"]
+                except:
+                    pass
                 theid = count/np.float32(self.N)
[…]

### 2020-06-03 analysis/plot_honda_quick.py

@@ -32,6 +32,7 @@ F1 = plt.figure (figsize=(9,8))
 t1_masked = np.load ('postproc/honda_t.npy')
 hondadelta = np.load ('postproc/honda_delta.npy')
 area_diff =  np.load ('postproc/area_diff.npy')
+map_diff =  np.load ('postproc/map_diff.npy')
 locn_vs_t =  np.load ('postproc/locn_vs_t.npy')
 
 xmax = max(t1_masked)
@@ -45,12 +46,18 @@ l1, = ax1.plot(t1_masked, hondadelta, 'o', markersize=12, color=col.red, label='
 # 0.054 is Senft and Woolsey's result for barrels (mouse 0,054, other rodents about 0.055)
 l2, = ax1.plot((0,xmax), (0.055, 0.055), '--', color=col.red, linewidth=3, label="good (S&W)")
 
+show_mapdiff = 0
+if show_mapdiff:
+    ax4 = ax1.twinx()
+    l5, = ax4.plot(t1_masked, map_diff, 's', markersize=12, color=col.blue)
+
 ax2 = F1.add_subplot(2,1,2)
 
 l3, = ax2.plot(t1_masked, area_diff, 's', markersize=12, color=col.black)
 
 ax3 = ax2.twinx()
 l4, = ax3.plot(t1_masked, locn_vs_t, 'h', markersize=12, color=col.gray60)
+
 
 ax2.set_xlabel ('time (10k steps)')
 ax1.set_ylabel ('$\delta$', rotation=0, labelpad=30)

### 2020-06-03 analysis/process_honda.py

@@ -73,7 +73,8 @@ mask_combined = np.invert(hondadelta.mask | sos_dist.mask)
 t1_masked = bdo.t_steps[mask_combined].T
 mapdiff = bdo.mapdiff[mask_combined].T
 area_diff = bdo.area_diff[mask_combined].T
-area_diff = area_diff / np.max(area_diff)
+# area_diff makes sense normalized by nhex
+area_diff = area_diff / bdo.nhex
 
 # a vs t.
 print ('a shape {0}'.format(np.shape(bdo.a)))
@@ -103,4 +104,5 @@ t1_masked = t1_masked / 10000
 np.save ('postproc/honda_t.npy', t1_masked)
 np.save ('postproc/honda_delta.npy', hondadelta)
 np.save ('postproc/area_diff.npy', area_diff)
+np.save ('postproc/map_diff.npy', mapdiff)
 np.save ('postproc/locn_vs_t.npy', bdo.locn_vs_t)

### 2020-06-03 analysis/process_multirun.py

@@ -1,0 +1,63 @@
+# To access argv and also include the include dir
+import os
+import sys
+sys.path.insert (0, './include')
+import numpy as np
+import BarrelData as bd
+import h5py
+import domcentres as dc
+
+# Constants
+vert = True
+horz = False
+
+basedir = '/home/seb/gdrive_usfd/data/BarrelEmerge/multirun_comp2/'
+t=30000
+
+hondas = []
+sos = []
+area = []
+mapdiffs = []
+locns = []
+
+for logdirname in os.listdir(basedir):
+    print (logdirname)
+
+    # Load the dirichlet data, domcentres, etc
+    bdo = bd.BarrelData()
+    bdo.loadTimeStep = t
+    bdo.loadAnalaysisData = True
+    bdo.loadPositions = True # for totalarea
+    bdo.loadGuidance = False
+    bdo.loadSimData = True # For localization and a
+    bdo.loadDivisions = False
+    try:
+        print ('load')
+        bdo.load (basedir+logdirname)
+        print ('loaded')
+        bdo.computeLocalization() # can then get bdo.locn_vs_t
+
+        #gfits, s_resid = dc.domcentres_analyse (bdo.domcentres, vert)
+        #gfits_h, s_resid_h = dc.domcentres_analyse (bdo.domcentres, horz)
+
+        # I just want the honda delta, the area measure and the omega for t=30000
+        print ('bdo.honda: {0}'.format(bdo.honda))
+        print ('bdo.sos_dist: {0}'.format(bdo.sos_dist))
+        print ('bdo.area_diff: {0}'.format(bdo.area_diff))
+        print ('bdo.locn_vs_t: {0}'.format(bdo.locn_vs_t))
+        hondas.append(bdo.honda[0])
+        sos.append(bdo.sos_dist[0])
+        # area_diff = area_diff / np.max(area_diff) # I don't like this division by the max of area_diff.
+        area.append(bdo.area_diff[0]/bdo.nhex)
+        mapdiffs.append(bdo.mapdiff[0])
+        locns.append(bdo.locn_vs_t[0])
+
+    except:
+        print ('exception')
+        continue
+
+print ('honda: {0} +/- {1}'.format(np.mean(hondas), np.std(hondas)))
+print ('sos_dist: {0} +/- {1}'.format(np.mean(sos), np.std(sos)))
[…]

### 2020-06-03 scripts/multirun_comp2.sh

@@ -1,0 +1,168 @@
+#!/bin/zsh
+
+# Run the 41N2M system R times, saving results in sequentially named
+# directories.
+
+# Check we're running from BarrelEmerge dir
+CURDIR=$(pwd | awk -F '/' '{print $NF}')
+echo "CURDIR: $CURDIR"
+
+if [ ! $CURDIR = "BarrelEmerge" ]; then
+    echo "Run from the base BarrelEmerge directory; i.e. ./scripts/$0"
+    exit 1
+fi
+
+PROGTAG=comp2 # NOT dccomp because in comp2 we need to vary F, not epsilon
+
+# Model parameters - these should match model in Fig 1 of the paper.
+k=3
+D=0.5
+F=1.2
+ALPHA=3.6
+BETA=16.67
+DT=0.0001
+HEXHEXD=0.03
+# Boundary fall off distance - should be at least 3 times HEXHEXD
+BFD=0.1
+# Noise on a
+ANG=0.2 # gain
+ANO=0.2 # offset
+
+# How many runs to do?
+R=10
+
+CONFIG_DIR="configs/rat/multirun"
+
+# Some loops over some sensitivity variables
+for run in {1..${R}}; do  # 0.03 never works; should go one higher too
+
+    # mr for multirun
+    JSON="mr_${PROGTAG}_Fig1_${run}_of_${R}.json"
+
+    cat > ${CONFIG_DIR}/${JSON} <<EOF
+{
+    "steps" : 30000,
+    "logevery": 30000,
+    "overwrite_logs": true,
+    "logbase" : "/home/seb/gdrive_usfd/data/BarrelEmerge/multirun_comp2/",
+    //"logbase" : "/home/seb/paramexplore_comp2/",
+    "hextohex_d" : ${HEXHEXD}, // Hex to hex distance, determines num hexes
+    "svgpath" : "./boundaries/rat_barrels/wb_110405_Dirichlet.svg",
+    "boundaryFalloffDist" : ${BFD},
+    "G" : 1.0,     // gamma gain
+    "dt" : ${DT}, // Timestep. Defaults to 0.00001 if omitted here
+
+    // Initial conditions parameters
+    "aNoiseGain" : ${ANG},
+    "aInitialOffset" : ${ANO},
+
+    // Parameters that will vary
+    "D" : ${D},
+    "F" : ${F},
[…]

### 2020-06-12 analysis/plot_FigSensitivity.py
Adds subgraph for eta <0.1

@@ -60,6 +60,11 @@ l2, = ax2.plot(gmn_noise_gain[:-mo], gmn_locn[:-mo], 'h-', markersize=12, color=
 ax3 = F1.add_subplot(3,3,7)
 l3, = ax3.plot(gmn_noise_gain[:-mo], gmn_eta[:-mo], 's-', markersize=12, color=col.black, label='$\eta$')
 
+# this is an inset axes over the third axis
+ax3_1 = plt.axes([.17, .24, .1, .06])
+ax3_1.plot(gmn_noise_gain[:-10], gmn_eta[:-10], 's-', markersize=8, color=col.black, label='$\eta$')
+ax3_1.tick_params(axis='both', which='major', labelsize=14)
+
 #
 # Guidance noise
 #

### 2020-06-12 analysis/plot_gamma_noise.py
Just removes some code comments

@@ -17,9 +17,6 @@ plt.rcParams['svg.fonttype'] = 'none'
 
 sdata = np.genfromtxt ('postproc/gamma_noise.csv', delimiter=",", names=True)
 s = np.sort (sdata, order='noise_gain')
-# If required
-#xlbl = (['.03', '.06', '.15', '.6', '1.5'])
-#ylbl = (['.015', '.022', '.03', '.06', '.1'])
 
 noise_gain = s['noise_gain']
 honda = s['hondadelta']
@@ -50,8 +47,6 @@ ax2 = F1.add_subplot(3,1,2)
 l2, = ax2.plot(noise_gain, locn, 'h-', markersize=12, color=col.gray60, label='Locn $\omega$')
 ax3 = F1.add_subplot(3,1,3)
 l3, = ax3.plot(noise_gain, eta, 's-', markersize=12, color=col.black, label='$\eta$')
-##ax4 = ax3.twinx()
-##l4, = ax4.plot(noise_gain, sos, 'v-', markersize=12, color=col.blue, label='SOS')
 
 # Axis tweaking
 ax1.set_xlabel ('Noise mag.')
@@ -60,15 +55,8 @@ ax2.set_xlabel ('Noise mag.')
 ax2.set_ylabel ('$\omega$', rotation=0, labelpad=30)
 ax3.set_xlabel ('Noise mag.')
 ax3.set_ylabel ('$\eta$', rotation=0, labelpad=30)
-##ax4.set_ylabel ('SOS', rotation=0, labelpad=30)
-#ax2.tick_params (axis='y', labelcolor=col.black)
-#ax1.set_xlim((0,1.0))
 ax1.set_ylim ((honda_min,honda_max))
 ax2.set_ylim ((locn_min,locn_max)) # locn, omega
-#ax2.set_xlim((0,1.0))
-#ax3.set_ylim ((eta_min,eta_max)) # area, eta
-#ax3.set_ylim ((0.52,0.62))
-#ax4.set_ylim ((sos_min,3.6))
 
 lw = 2
 for axis in ['top','bottom','left','right']:
@@ -78,8 +66,6 @@ for axis in ['top','bottom','left','right']:
   ax2.tick_params(width=lw)
   ax3.spines[axis].set_linewidth(lw)
   ax3.tick_params(width=lw)
-  ##ax4.spines[axis].set_linewidth(lw)
-  ##ax4.tick_params(width=lw)
 
 plt.tight_layout()
 

### 2020-06-12 paper/briefpaper/brief.tex
Updated figure
Figure improvements

@@
       the locations of putative barreloid centers (reconstructed from
-      \citealp{haidarliu_size_2001}). \textbf{C} Simulation results for parameters
+      \citealp{haidarliu_size_2001}, \cmnt{Fig.\,5b}). \textbf{C} Simulation results for parameters
       \cmnt{$N=41$, $\alpha=3.6$, $\beta=16.67$, $k=3$, $D=0.5$, $\gamma\in\pm 2$,
@@
 
+% Full page width
+\begin{figure}
+  \begin{fullwidth}
+    \includegraphics[width=\linewidth]{./Fig3.png}
+    \caption{\cmnt{Sensitivity analysis. Column \textbf{A} shows...}}
+    \label{fig:sens}
+  \end{fullwidth}
+\end{figure}
+
 \cmnt{We performed a sensitivity analysis to determine if the quality of the
-  pattern was sensitive to any of the following inputs: i)~The angle and
-  magnitude of each of the two guidance fields (Fig.\,\ref{fig:main}A);
-  ii)~the magnitude, offset and characteristic length scale of the noise
-  applied to $a_i$ at $t=0$; and iii)~the magnitude of each of the interaction
-  parameters, $\gamma_{i,j}$. We analysed the effect on the three pattern
-  metrics, $\delta$, $\eta$ and $\omega$ evaluated after 30000 simulation
-  steps. First, we conducted 10 runs of the simulation to evaluate the mean
-  and standard deviation (as a result of the different random initial
-  conditions for $a_i$) of the three metrics in the default model
-  (Fig.\,\ref{fig:main}C--E) which were $\delta=0.089\pm{0.004}$,
-  $\eta=0.531\pm{0.005}$ and $\omega=0.2165\pm{0.0001}$). These results are
-  not strongly affected by the random initial conditions for $a_i$ but give a
-  baseline from which to evaluate whether changes in the other inputs have a
-  statistically significant effect.}
+  pattern was affected by perturbations to any of the following inputs:
+%
+  i)~the magnitude and offset of the noise applied to $a_i$ at $t=0$;
+%
+  ii)~random noise applied to the interaction parameters, $\gamma_{i,j}$;
+%
+  iii)~random noise (of various length scales) applied to the guidance fields;
+%
+  and
+%
+  iv)~The angle and magnitude of one of the guidance fields, ($\rho_1$;
+  Fig.\,\ref{fig:main}A).
+%
+  We analysed the effect on the three pattern metrics, $\delta$ (the Honda
+  Voronoi metric), $\eta$ (the pattern-match metric) and $\omega$ (which
+  measures localization of connections, $c_i$) evaluated after 30000
+  simulation steps.}
+
+\cmnt{First, we conducted 10 runs of the simulation shown in
+  Fig.\,\ref{fig:main}C to evaluate the mean and standard deviation of the
+  three metrics in the default model. The results (where deviations occur as a
+  result of the different random initial conditions for $a_i$) were
+  $\delta=0.089\pm{0.004}$, $\eta=113.5\pm{1.2}$ and
+  $\omega=0.2165\pm{0.0001}$. None of the three metrics were strongly affected
+  by the random initial conditions for $a_i$. The results give a baseline from
+  which to evaluate whether changes in the other inputs have a statistically
+  significant effect.}
+
+\cmnt{Next, we changed the initial condition for $a_i$, changing the
+  parameters of the generating uniform distribution to set each hex to
+  $0.4+{U}(0,0.4)$. We ran this system 10 times and obtain the result
+  $\delta=0.0\pm{0.000}$, $\eta=100\pm{0}$ and $\omega=0.2ish\pm{0.0000}$. A
+  t-test demonstrates that this is ????. A final test of a noise-free initial
+  condition with $a_i=0.2$ for every hex also ????.}
+
+\cmnt{Fig\,\ref{fig:sens}A shows the results for perturbations ii) to iv). The
+  first column shows the results for the three metrics where the interaction
+  parameters $\gamma_{i,j}$ are replaced by $\gamma_{i,j}'$ according to
+  $\gamma_{i,j}' = \gamma_{i,j} + U(0,4\nu_\gamma)$, where $\nu_\gamma$ is the
+  chosen magnitude of the noise (the proportion of the full range of
+  $\gamma_{i,j}$, which is 4). The Voronoi nature of the pattern ($\delta$)
+  and the localization of the fields are not strongly affected by interaction
+  parameter noise, but the pattern metric, $\eta$ shows a... ADD SUB GRAPH.}
 
@@
 \begin{center}
-  \includegraphics[width=\linewidth]{./Fig3.png}
+  \includegraphics[width=\linewidth]{./Fig4.png}
   \captionof{figure}{Guidance
@@
 
+\cmnt{Add a description of $\eta$, the pattern-matching measure, here.}
+
 All code required to reproduce these results is available at

### 2020-06-12 analysis/process_multirun_anoise.py
Last bit of Friday work

@@ -1,0 +1,74 @@
+# To access argv and also include the include dir
+import os
+import sys
+sys.path.insert (0, './include')
+import numpy as np
+import BarrelData as bd
+import h5py
+import domcentres as dc
+
+# Constants
+vert = True
+horz = False
+
+basedir = '/home/seb/gdrive_usfd/data/BarrelEmerge/multirun_anoise/'
+t=30000
+
+hondas = []
+sos = []
+area = []
+mapdiffs = []
+locns = []
+eta = []
+
+for logdirname in os.listdir(basedir):
+    print (logdirname)
+
+    # Load the dirichlet data, domcentres, etc
+    bdo = bd.BarrelData()
+    bdo.loadTimeStep = t
+    bdo.loadAnalaysisData = True
+    bdo.loadPositions = True # for totalarea
+    bdo.loadGuidance = True
+    bdo.loadHexFlags = True
+    bdo.loadSimData = True # For localization and a
+    bdo.loadDivisions = False
+    try:
+        print ('load')
+        bdo.load (basedir+logdirname)
+        print ('loaded')
+        bdo.computeLocalization() # can then get bdo.locn_vs_t
+        bdo.computeAdjacencyMeasure()
+
+        #gfits, s_resid = dc.domcentres_analyse (bdo.domcentres, vert)
+        #gfits_h, s_resid_h = dc.domcentres_analyse (bdo.domcentres, horz)
+
+        # I just want the honda delta, the area measure and the omega for t=30000
+        print ('bdo.honda: {0}'.format(bdo.honda))
+        print ('bdo.sos_dist: {0}'.format(bdo.sos_dist))
+        print ('bdo.area_diff: {0}'.format(bdo.area_diff))
+        print ('bdo.locn_vs_t: {0}'.format(bdo.locn_vs_t))
+        hondas.append(bdo.honda[0])
+        sos.append(bdo.sos_dist[0])
+        area_diff = bdo.area_diff[0] / bdo.nhex
+        area.append(area_diff)
+
+        mapdiffs.append(bdo.mapdiff[0])
+        locns.append(bdo.locn_vs_t[0])
+
+        print ('bdo.mean_adjacency_differencemag: {0}'.format(bdo.mean_adjacency_differencemag))
+        print ('bdo.mean_adjacency_arrangement: {0}'.format(bdo.mean_adjacency_arrangement))
[…]

### 2020-06-12 analysis/process_multirun_anoiseless.py

@@ -1,0 +1,74 @@
+# To access argv and also include the include dir
+import os
+import sys
+sys.path.insert (0, './include')
+import numpy as np
+import BarrelData as bd
+import h5py
+import domcentres as dc
+
+# Constants
+vert = True
+horz = False
+
+basedir = '/home/seb/gdrive_usfd/data/BarrelEmerge/multirun_anoiseless/'
+t=30000
+
+hondas = []
+sos = []
+area = []
+mapdiffs = []
+locns = []
+eta = []
+
+for logdirname in os.listdir(basedir):
+    print (logdirname)
+
+    # Load the dirichlet data, domcentres, etc
+    bdo = bd.BarrelData()
+    bdo.loadTimeStep = t
+    bdo.loadAnalaysisData = True
+    bdo.loadPositions = True # for totalarea
+    bdo.loadGuidance = True
+    bdo.loadHexFlags = True
+    bdo.loadSimData = True # For localization and a
+    bdo.loadDivisions = False
+    try:
+        print ('load')
+        bdo.load (basedir+logdirname)
+        print ('loaded')
+        bdo.computeLocalization() # can then get bdo.locn_vs_t
+        bdo.computeAdjacencyMeasure()
+
+        #gfits, s_resid = dc.domcentres_analyse (bdo.domcentres, vert)
+        #gfits_h, s_resid_h = dc.domcentres_analyse (bdo.domcentres, horz)
+
+        # I just want the honda delta, the area measure and the omega for t=30000
+        print ('bdo.honda: {0}'.format(bdo.honda))
+        print ('bdo.sos_dist: {0}'.format(bdo.sos_dist))
+        print ('bdo.area_diff: {0}'.format(bdo.area_diff))
+        print ('bdo.locn_vs_t: {0}'.format(bdo.locn_vs_t))
+        hondas.append(bdo.honda[0])
+        sos.append(bdo.sos_dist[0])
+        area_diff = bdo.area_diff[0] / bdo.nhex
+        area.append(area_diff)
+
+        mapdiffs.append(bdo.mapdiff[0])
+        locns.append(bdo.locn_vs_t[0])
+
+        print ('bdo.mean_adjacency_differencemag: {0}'.format(bdo.mean_adjacency_differencemag))
+        print ('bdo.mean_adjacency_arrangement: {0}'.format(bdo.mean_adjacency_arrangement))
[…]

### 2020-06-13 paper/briefpaper/brief.tex
Last bit of Friday work
Updates the text for the section on modifying the noise applied to a_i at t=0

@@
 %
-  iii)~random noise (of various length scales) applied to the guidance fields;
+  iii)~random noise (at various length scales) applied to the guidance fields;
 %
@@
 %
-  We analysed the effect on the three pattern metrics, $\delta$ (the Honda
-  Voronoi metric), $\eta$ (the pattern-match metric) and $\omega$ (which
-  measures localization of connections, $c_i$) evaluated after 30000
-  simulation steps.}
+  In the following analysis, we always evaluated the three pattern metrics,
+  $\delta$ (the Honda Voronoi metric), $\eta$ (the pattern-match metric) and
+  $\omega$ (which measures localization of connections, $c_i$) after running
+  the model for 30000 simulation steps.}
 
@@
   Fig.\,\ref{fig:main}C to evaluate the mean and standard deviation of the
-  three metrics in the default model. The results (where deviations occur as a
-  result of the different random initial conditions for $a_i$) were
-  $\delta=0.089\pm{0.004}$, $\eta=113.5\pm{1.2}$ and
-  $\omega=0.2165\pm{0.0001}$. None of the three metrics were strongly affected
-  by the random initial conditions for $a_i$. The results give a baseline from
+  three metrics in the default model. The
+  results (where deviations occur as a result of the different random initial
+  conditions for $a_i$) were $\delta=0.089\,(0.004)$, $\eta=113.5\,(1.2)$ and
+  $\omega=0.2165\,(0.0001)$. None of the three metrics were strongly affected
+  by the random changes to $a_i$ on each run. The results give a baseline from
   which to evaluate whether changes in the other inputs have a statistically
@@
 
-\cmnt{Next, we changed the initial condition for $a_i$, changing the
-  parameters of the generating uniform distribution to set each hex to
-  $0.4+{U}(0,0.4)$. We ran this system 10 times and obtain the result
-  $\delta=0.0\pm{0.000}$, $\eta=100\pm{0}$ and $\omega=0.2ish\pm{0.0000}$. A
-  t-test demonstrates that this is ????. A final test of a noise-free initial
-  condition with $a_i=0.2$ for every hex also ????.}
+\cmnt{Next, we modified the parameters of the generating uniform distribution
+  to initialize $a_i$ for each hex to $0.1+{U}(0,0.4)$ (increasing the variability of
+  $a_i(\mb{x},t=0)$ without changing the mean initial activation). We ran this
+  system 10 times and obtained the results $\delta=0.090\,(0.005)$,
+  $\eta=115\,(1.9)$ and $\omega=0.2166\,(0.0003)$. Paired two-sample t-tests
+  demonstrated that the increase in noise amplitude made no significant
+  difference to the metric results ($\delta:
+  \mathrm{t}=-0.69,\mathrm{p}=0.49$,
+  $\eta: \mathrm{t}=-2.0,\mathrm{p}=0.06$,
+  $\omega: \mathrm{t}=-1.7,\mathrm{p}=0.11$). 10 runs of a
+  noise-free initial condition with $a_i=0.3$ for every hex showed that
+  the model is deterministic (all runs produced the results $\delta=0.0928$,
+  $\eta=114.31$ and $\omega=0.21649$) and that stochasticity in
+  $a_i(\mb{x},t=0)$ is not required by this model, as all three results lay within
+  one standard deviation of those for the runs in which $a_i$ was initialized stochastically.}
+
+% Quick check that noise free result is within 1 SD for the noisier initial conditions
+%         Noisefree    Normal noise    NF SD from Normal?   Noisy            NF SD from Noisy?
+% delta   0.0928       0.089(0.004)    <1                   0.090 (0.005)    <1
+% eta     114.31       113.5(1.2)      <1                   115 (1.9)        <1
+% omega   0.21649      0.2165(0.0001)  <1                   0.2166 (0.0003)  <1
 

### 2020-06-15 paper/briefpaper/brief.tex
Adds pattern metric section (eta)

@@
 \newcommand{\code}[1]{\textsf{#1}}
+% Nice norms:
+\usepackage{mathtools}
+\DeclarePairedDelimiter{\norm}{\lVert}{\rVert}
 
@@
 
-\cmnt{Add a description of $\eta$, the pattern-matching measure, here.}
+\cmnt{The pattern-matching metric, $\eta$, incorporates information about i)
+  the differences in the areas of the simulated and experimental barrels
+  ($\mathcal{A}_i^{\mathrm{sim}}$ and $\mathcal{A}_i^{\mathrm{exp}}$); and ii)
+  the information provided by `adjacency vectors', $\mathcal{V}_i$ (one for
+  each barrel) whose $N$ elements contain the length of the barrel boundary
+  which adjoins each other barrel in the pattern. The value of
+  $\mathcal{V}_{i,j}$ (the $j^{\mathrm{th}}$ element of $\mathcal{V}_i$) is
+  the length of the boundaries between regions of type $i$ and regions of type
+  $j$. In a well-formed barrel pattern, $\mathcal{V}_{i}$ is a sparse vector
+  and note that $\mathcal{V}_{i,i} = 0, \forall i$. $\mathcal{V}_{i}$ can be
+  computed for the experimental and simulated barrel patterns and compared. A
+  dimensionless measure of the pattern's arrangement and neighbour relations
+  can be obtained from the scalar product of the simulated and experimental
+  adjacency vectors:
+  $\frac{1}{N} \sum_i \frac { \mathcal{V}_{i}^{\mathrm{sim}}}
+  {b_{i}^{\mathrm{sim}} } \vcdot \frac{\mathcal{V}_{i}^{\mathrm{exp}}}{b_{i}^{\mathrm{exp}}}$, where
+  $b_i^{\mathrm{exp}}$($b_i^{\mathrm{sim}}$) is the total length of the boundary of
+  experimental(simulated) barrel $i$. This measure is low early in the simulation, when the regions do
+  not have the neighbour relations of the experimental pattern
+  (Fig.\,\ref{fig:main}C, $t=1000$) and maximal when the neighbour relations
+  have become established ($t=10000$). It does not distinguish between
+  $t=10000$ and $t=30000$.}
+%
+\cmnt{A second way to analyse the adjacency vectors is to consider
+  the mean magnitude of the difference between the simulated and experimental
+  vectors: $\frac{1}{N} \sum_i \norm{ \mathcal{V}_{i}^{\mathrm{sim}} -
+    \mathcal{V}_{i}^{\mathrm{exp}} } $. This tends to 0 for a perfect
+  match and can separate patterns with straight boundaries from those with
+  noise on the boundary shape (For an example of `noisy edges', see Fig.\,\ref{fig:sens}B.ii).}
+%
+\cmnt{We combined all of these measures into a single pattern quality metric:}
+%
+\begin{equation} \label{eq:eta}
+\color{blue}
+  \eta =
+  \frac
+  {
+    \frac{1}{N} \sum_i \norm[\big]{ \mathcal{A}_i^{\mathrm{sim}} - \mathcal{A}_i^{\mathrm{exp}} }
+    \times
+    \frac{1}{N} \sum_i \norm[\big]{ \mathcal{V}_{i}^{\mathrm{sim}} -
+      \mathcal{V}_{i}^{\mathrm{exp}} }
+  }
+  {
+    \frac{1}{N}
+    \sum_i \frac {\mathcal{V}_{i}^{\mathrm{sim}}} {b_{i}^{\mathrm{sim}}}
+    \vcdot
+    \frac {\mathcal{V}_{i}^{\mathrm{exp}}} {b_{i}^{\mathrm{exp}}}
+  }
+\color{black}
+\end{equation}
+%
+\cmnt{which has units of mm$^3$.}
 

### 2020-06-16 paper/briefpaper/brief.tex
plodding through with edits

--- 2020-06-13 paper/briefpaper/brief.tex
@@
 
-\documentclass[9pt,lineno]{elife}
+\documentclass[9pt,lineno,draft]{elife}
 
@@
 \newcommand{\cmnt}[1]{\textcolor{blue}{#1}}
+\newcommand{\stu}[1]{\textcolor{red}{#1}}
 \newcommand{\dvrg}{\nabla\vcdot\nabla}
@@
 simulation requires cells on the cortical sheet to communicate with immedately
-adjacent cells only (see \emph{Materials \& Methods}) The second term on the
+adjacent cells only (see \emph{Materials \& Methods}). The second term on the
 right quantifies the coupling between axon branching and synaptogenesis. Here
@@
       delineate boundaries (see Movie
-      S1). \textbf{D} Red dots show the Honda-$\delta$ metric obtained from
+      S1). \textbf{D} Red dots show the Honda-$\delta$ metric obtained from \stu{the}
       simulation approaching that obtained from \cmnt{real} barrels in \textbf{A} (dotted
-      line); black squares \cmnt{give a} measure \cmnt{of} the correspondence
-      between the real and
-      simulated barrel shapes, $\eta$ (the sum of differences in
+      line); black squares \stu{provide a} measure \cmnt{of} the correspondence
+      between the real and simulated barrel shapes, $\eta$ (the sum of differences in
       area; units \cmnt{mm$^2$});
-      \cmnt{grey hexagons show the measure of localization, $\omega =
-        \oiint_{S} \left( \mathrm{max}_i(c_i)/\sum_{i=1}^{N} c_i\right) \mathrm{d}S$. }
-      \textbf{E} \cmnt{The localization measure plotted across the domain
-        shows that the structure in the connection density which has developed
-        by $t=30000$ is largely absent at $t=1000$. The
-        colour map
-        indicates the values of  $\mathrm{max}_i(c_i)/\sum_{i=1}^{N} c_i$. Overlaid
+      \cmnt{grey hexagons \stu{provide a measure of how selectively each cortical site is innervated}, $\omega = \oiint_{S} \left( \mathrm{max}_i(c_i)/\sum_{i=1}^{N} c_i\right) \mathrm{d}S$. }
+      \textbf{E} \stu{Plotted across the cortical sheet, the selectivity develops to reveal an alignment with the emergent barrel boundary shapes}
+      \cmnt{The colour map indicates the values of  $\mathrm{max}_i(c_i)/\sum_{i=1}^{N} c_i$. Overlaid
         contours show $c_i > 0.95 \mathrm{max}_i(c_i)$.} All scale bars 1\,mm.}
@@
   \color{blue}
-  \chi_i(\mb{x}, t) = \frac{\epsilon a_i}{N-1} \nabla \textstyle\sum_{j\ne i}^{N}{a_j}
+  \chi_i(\mb{x}, t) = \frac{\epsilon a_i}{N-1} \nabla \textstyle\sum_{j\ne i}^{N}{a_j}.
   \color{black}
@@
 %
-\cmnt{which contributes to the \emph{flux of axonal branching} as an
-  additional diffusion-like term which makes the branching density for a given
-  projection diffuse away from regions} where the branches of other projections
-become dense. \cmnt{$\epsilon$ parameterizes the strength of this competitive
-  interaction.} Note that this operation is local to individual afferent
+\stu{This term contributes to the \emph{flux of axonal branching} as an
+ additional source of diffusion, which reduces the branching density for a given projection where the branches of other projections are dense.} \stu{The strength of this additional competition is parameterized by $\epsilon$}.
+Note that this operation is local to individual afferent
 projections.
@@
 determined by the concentration of a similar molecule at its thalamic origin,
-i.e., the putative barreloid center. \cmnt{Thus, we suggest that there exist a
-  pair of signalling molecules whose expression varies linearly (with
-  orthogonal gradients) across the thalamus, analgous to the ephrins, whose
-  existence is well known in the cortex. Thalamocortical axons growing from a
-  region in the thalamus with high expression of one of these molecules will
-  interact with a corresponding cortical signalling molecule more strongly
-  than an axon emanating from a region with low expression of the thalamic
-  signal.}  \cmnt{We defined two orthogonal, linear gradients oriented in the
-  anterior-posterior and medial-lateral directions and used the coordinates of
-  41 barreloid centers recreated from \cite{haidarliu_size_2001},
-  corresponding to the macrovibrissae, to populate 41 pairs of $\gamma$
-  values.}  (Fig.\,\ref{fig:main}B). \cmnt{In other words, we simply scaled
-  the barreloid coordinates in an anterior-posterior/medial-lateral coordinate
-  system to create the $\gamma$ values in the range [-2,2]. Note that positive
-  \emph{and} negative $\gamma$ values were generated; the origin of the
-  coordinate system was located at the approximate center of the thalamic
-  barreloid field. This prevented an overall drift of the cortical pattern
-  away from the centroid of the barrel field and may explain why
-  coordinate-giving transcription factors and the ephrins are often found as
-  pairs acting in opposing directions.}  \cmnt{We concentrated on
-  macrovibrissae (excluding microvibrissae) because it is easy to identify
-  their corresponding structures both in the thalamus and in the cortex. We
-  also excluded the 5 rhinal whiskers because their cortical representation is
-  in a separate island, distinct, and somewhat removed from the representation
-  of the macrovibirissae.}
+i.e., the putative barreloid center.
+\stu{As such, we consider linear gradients aligned to anterior-posterior and medial-lateral axes through the VPM, and as $\gamma$-values use the coordinates of the barreloid centers in these axes as they appear in \citep{haidarliu_size_2001} (Fig.\,\ref{fig:main}B).} \stu{Complementary thalamic and cortical molecular guidance gradients of this form have been well established experimentally \citep{vanderhaeghen_mapping_2000,miller_epha7-ephrin-a5_2006}, e.g., cells originating in the medial part of ventrobasal thalamus (comprising VPM and VPL) express high levels of EphA receptors and project to the lateral part of S1, which expresses low levels of ephrin-A5, and cells originating in the lateral part of ventrobasal thalamus express low levels of EphA receptors and project to the medial part of S1, which expresses high levels of ephrin-A5 (see Gao, 1998; Dufour et al., 2003; Vanderhaegen \& Polleux, 2004; Speer \& Chapman, 2005)}
+
+%\cmnt{In other words, we simply scaled the barreloid coordinates in an anterior-posterior/medial-lateral coordinate system to create the $\gamma$ values in the range [-2,2].
+Note that positive \emph{and} negative $\gamma$ values were generated; the origin of the coordinate system was located at the approximate center of the thalamic barreloid field.
+This prevented an overall drift of the cortical pattern away from the centroid of the barrel field and may explain why coordinate-giving transcription factors and the ephrins are often found as pairs acting in opposing directions.}  \cmnt{We concentrated on macrovibrissae (excluding microvibrissae) because it is easy to identify their corresponding structures both in the thalamus and in the cortex.
+We also excluded the 5 rhinal whiskers because their cortical representation is in a separate island, distinct, and somewhat removed from the representation of the macrovibirissae.}}
+
+%\cmnt{Thus, we suggest that there exist a pair of signalling molecules whose expression varies linearly (with orthogonal gradients) across the thalamus, analgous to the ephrins, whose existence is well known in the cortex.
+Thalamocortical axons growing from a region in the thalamus with high expression of one of these molecules will interact with a corresponding cortical signalling molecule more strongly than an axon emanating from a region with low expression of the thalamic signal.}
+
+%\cmnt{We defined two orthogonal, linear gradients oriented in the anterior-posterior and medial-lateral directions and used the coordinates of 41 barreloid centers recreated from \cite{haidarliu_size_2001}, corresponding to the macrovibrissae, to populate 41 pairs of $\gamma$ values.}  (Fig.\,\ref{fig:main}B).
+\cmnt{In other words, we simply scaled the barreloid coordinates in an anterior-posterior/medial-lateral coordinate system to create the $\gamma$ values in the range [-2,2].
+Note that positive \emph{and} negative $\gamma$ values were generated; the origin of the coordinate system was located at the approximate center of the thalamic barreloid field.
+This prevented an overall drift of the cortical pattern away from the centroid of the barrel field and may explain why coordinate-giving transcription factors and the ephrins are often found as pairs acting in opposing directions.}  \cmnt{We concentrated on macrovibrissae (excluding microvibrissae) because it is easy to identify their corresponding structures both in the thalamus and in the cortex.
+We also excluded the 5 rhinal whiskers because their cortical representation is in a separate island, distinct, and somewhat removed from the representation of the macrovibirissae.}
 

### 2020-06-16 paper/briefpaper/brief.tex
Merging one small conflict

@@
       simulation approaching that obtained from \cmnt{real} barrels in \textbf{A} (dotted
-      line); black squares \stu{provide a} measure \cmnt{of} the correspondence
-      between the real and simulated barrel shapes, $\eta$ (the sum of differences in
-      area; units \cmnt{mm$^2$});
-      \cmnt{grey hexagons \stu{provide a measure of how selectively each cortical site is innervated}, $\omega = \oiint_{S} \left( \mathrm{max}_i(c_i)/\sum_{i=1}^{N} c_i\right) \mathrm{d}S$. }
-      \textbf{E} \stu{Plotted across the cortical sheet, the selectivity develops to reveal an alignment with the emergent barrel boundary shapes}
-      \cmnt{The colour map indicates the values of  $\mathrm{max}_i(c_i)/\sum_{i=1}^{N} c_i$. Overlaid
+      line); black squares \cmnt{give a} measure \cmnt{of} the correspondence
+      between the real and
+      simulated barrel shapes, $\eta$ (the \cmnt{pattern quality metric---see
+        Materials \& Methods; units mm$^3$});
+      \cmnt{grey hexagons \stu{provide a measure of how selectively each
+          cortical site is innervated} OR: show the measure of localization, $\omega =
+        \oiint_{S} \left( \mathrm{max}_i(c_i)/\sum_{i=1}^{N} c_i\right) \mathrm{d}S$. }
+      \textbf{E} \stu{Plotted across the cortical sheet, the selectivity
+        develops to reveal an alignment with the emergent barrel boundary
+        shapes} OR: \cmnt{The localization measure plotted across the domain
+        shows that the structure in the connection density which has developed
+        by $t=30000$ is largely absent at $t=1000$. The
+        colour map
+        indicates the values of  $\mathrm{max}_i(c_i)/\sum_{i=1}^{N} c_i$. Overlaid
         contours show $c_i > 0.95 \mathrm{max}_i(c_i)$.} All scale bars 1\,mm.}
@@
   results (where deviations occur as a result of the different random initial
-  conditions for $a_i$) were $\delta=0.089\,(0.004)$, $\eta=113.5\,(1.2)$ and
+  conditions for $a_i$) were $\delta=0.089\,(0.004)$, $\eta=113.5\textbf{!!}\,(1.2)$ and
   $\omega=0.2165\,(0.0001)$. None of the three metrics were strongly affected
@@
   system 10 times and obtained the results $\delta=0.090\,(0.005)$,
-  $\eta=115\,(1.9)$ and $\omega=0.2166\,(0.0003)$. Paired two-sample t-tests
+  $\eta=115\,(1.9)\textbf{!!}$ and $\omega=0.2166\,(0.0003)$. Paired two-sample t-tests
   demonstrated that the increase in noise amplitude made no significant
@@
   \mathrm{t}=-0.69,\mathrm{p}=0.49$,
-  $\eta: \mathrm{t}=-2.0,\mathrm{p}=0.06$,
+  $\eta: \mathrm{t}=-2.0,\mathrm{p}=0.06\textbf{!!}$,
   $\omega: \mathrm{t}=-1.7,\mathrm{p}=0.11$). 10 runs of a
@@
   the model is deterministic (all runs produced the results $\delta=0.0928$,
-  $\eta=114.31$ and $\omega=0.21649$) and that stochasticity in
+  $\eta=114.31\textbf{!!}$ and $\omega=0.21649$) and that stochasticity in
   $a_i(\mb{x},t=0)$ is not required by this model, as all three results lay within
@@
     \frac {\mathcal{V}_{i}^{\mathrm{exp}}} {b_{i}^{\mathrm{exp}}}
-  }
+  },
 \color{black}

### 2020-06-16 analysis/include/BarrelData.py

@@ -8,6 +8,7 @@ import h5py
 import re
 import json
 import sys
+import math
 
 #
 # This is a class to load the data generated by the BarrelEmerge
@@ -145,8 +146,8 @@ class BarrelData:
         # in ID to the experimental map. Lower is better. Range [0,1].
         self.mapdiff = np.array([])
 
-        # The sum (in hexes) of the absolute difference in areas of each
-        # experimental domain and simulated domain. len(t). Range [0,nhex].
+        # The sum of the absolute difference in areas of each experimental
+        # domain and simulated domain. len(t). Range [0,nhex].
         self.area_diff = np.array([])
 
         # The divisions between domains. All those paths that the C++ code
@@ -591,9 +592,21 @@ class BarrelData:
 
             # The sum of squared distances between the simulated barrels and the experimentally determined pattern
             self.sos_dist[fi] = list(f['sos_distances'])[0]
-            # The sum of the square of the absolute differences in area (in num
-            # hexes) between the experimental and simulated barrel fields.
+
+            # From the h5 file, we get the sum of the absolute differences in
+            # area (*in num hexes*) between the experimental and simulated
+            # barrel fields. This is a total difference for all barrels:
+            # sum_i^N (abs(area_diff_i))...
             self.area_diff[fi] = list(f['area_diff'])[0]
+            # ...now convert the area_diff in hexes into an area by multiplying
+            # by the area of one hex
+            hex_area = self.hextohex_d * self.hextohex_d * math.sqrt(3) / 2.0
+            self.area_diff[fi] *= hex_area
+            # And divide by N, to get the area_diff per barrel, rather than the
+            # total area diff, as this is how the statistic is presented in the
+            # paper.
+            self.area_diff[fi] /= float(self.N)
+
             # Another metric to determine the difference between the current pattern and the
             # experimentally observed pattern, this one is based on traced barrel boundaries.
             self.mapdiff[fi] = list(f['mapdiff'])[0]

### 2020-06-16 analysis/include/paramplot.py

@@ -502,7 +502,7 @@ def paramplot_pub (sdata, F, column_tag, x_tag, y_tag, ktarg, ttarg, param_tuple
     sos_max = max(sdata[:]['sos_dist'])
     sos_min = min(sdata[:]['sos_dist'])
     eta_min = 0
-    eta_max = 600
+    eta_max = 1.5
     adj_diffmag_max = 62
     adj_diffmag_min = 15
     print ('Output to include in code for sensitivity graphs:')

### 2020-06-16 analysis/plot_FigSensitivity.py

@@ -29,7 +29,7 @@ locn_max=0.7319451998375573
 sos_min=2.9542248615230577
 sos_max=3.350204398214124
 eta_min=0
-eta_max=600
+eta_max=1.5
 
 # One figure:
 F1 = plt.figure (figsize=(16,12))
@@ -61,9 +61,10 @@ ax3 = F1.add_subplot(3,3,7)
 l3, = ax3.plot(gmn_noise_gain[:-mo], gmn_eta[:-mo], 's-', markersize=12, color=col.black, label='$\eta$')
 
 # this is an inset axes over the third axis
-ax3_1 = plt.axes([.17, .24, .1, .06])
+ax3_1 = plt.axes([.2, .24, .1, .06])
 ax3_1.plot(gmn_noise_gain[:-10], gmn_eta[:-10], 's-', markersize=8, color=col.black, label='$\eta$')
 ax3_1.tick_params(axis='both', which='major', labelsize=14)
+#ax3_1.set_yticks([0.0014,0.0017])
 
 #
 # Guidance noise

### 2020-06-16 analysis/plot_gamma_noise.py

@@ -35,7 +35,7 @@ locn_max=0.7319451998375573
 sos_min=2.9542248615230577
 sos_max=3.350204398214124
 eta_min=0
-eta_max=600
+eta_max=1.5
 
 # And plot this longhand here:
 F1 = plt.figure (figsize=(5,12))

### 2020-06-16 analysis/plot_guidance_noise.py

@@ -49,7 +49,7 @@ locn_max=0.7319451998375573
 sos_min=2.9542248615230577
 sos_max=3.350204398214124
 eta_min=0
-eta_max=600
+eta_max=1.5
 
 # Use same min/maxes as produced by the parameter search?
 im1 = ax1.imshow (honda, cmap='inferno_r', origin='lower', vmin=honda_min, vmax=honda_max, interpolation='nearest')

### 2020-06-16 analysis/plot_honda_quick.py

@@ -27,7 +27,7 @@ matplotlib.rc('font', **fnt)
 plt.rcParams['svg.fonttype'] = 'none'
 
 # And plot this longhand here:
-F1 = plt.figure (figsize=(9,12))
+F1 = plt.figure (figsize=(10,12))
 
 t1_masked = np.load ('postproc/honda_t.npy')
 hondadelta = np.load ('postproc/honda_delta.npy')
@@ -84,7 +84,8 @@ ax3.set_ylabel ('$\omega$', rotation=0, labelpad=30)
 ax2.tick_params (axis='y', labelcolor=col.black)
 ax1.set_xlim ((0,xmax))
 ax1.set_ylim ((0,0.35))
-ax2.set_ylim ((0,1000))
+eta_max = 1.5
+ax2.set_ylim ((0,eta_max))
 ax3.set_ylim ((0,0.28))
 ax2.set_xlim ((0,xmax))
 

### 2020-06-16 analysis/plot_sensitivity_gamma.py

@@ -42,7 +42,7 @@ locn_max=0.7319451998375573
 sos_min=2.9542248615230577
 sos_max=3.350204398214124
 eta_min=0
-eta_max=600
+eta_max=1.5
 
 bc = int(0) # 'barrel counter'
 for br in ['b', 'c', 'd']:

### 2020-06-16 analysis/plot_sensitivity_guide1.py

@@ -46,7 +46,7 @@ locn_max=0.7319451998375573
 sos_min=2.9542248615230577
 sos_max=3.350204398214124
 eta_min=0
-eta_max=600
+eta_max=1.5
 
 # Use same min/maxes as produced by the parameter search?
 im1 = ax1.imshow (honda, cmap='inferno_r', vmin=honda_min, vmax=honda_max, interpolation='nearest')

### 2020-06-16 analysis/process_gamma_noise.py

@@ -46,7 +46,6 @@ for logdirname in os.listdir(basedir):
         t1 = t1_masked[:,0]
         mapdiff = bdo.mapdiff[mask_combined].T
         area_diff = bdo.area_diff[mask_combined].T
-        area_diff = area_diff / bdo.nhex
         sos_dist = np.sqrt(sos_dist.compressed()/bdo.N)
         hondadelta = hondadelta.compressed()
         bdo.computeLocalization()

### 2020-06-16 analysis/process_guidance_noise.py

@@ -53,7 +53,6 @@ for logdirname in os.listdir(basedir):
         #print ('t1: {0}'.format (t1))
         mapdiff = bdo.mapdiff[mask_combined].T
         area_diff = bdo.area_diff[mask_combined].T
-        area_diff = area_diff / bdo.nhex
         sos_dist = np.sqrt(sos_dist.compressed()/bdo.N)
         hondadelta = hondadelta.compressed()
 

### 2020-06-16 analysis/process_honda.py

@@ -75,8 +75,6 @@ mask_combined = np.invert(hondadelta.mask | sos_dist.mask)
 t1_masked = bdo.t_steps[mask_combined].T
 mapdiff = bdo.mapdiff[mask_combined].T
 area_diff = bdo.area_diff[mask_combined].T
-# area_diff makes sense normalized by nhex
-area_diff = area_diff / bdo.nhex
 
 adjacency_arrangement = bdo.mean_adjacency_arrangement[mask_combined].T
 adjacency_differencemag = bdo.mean_adjacency_differencemag[mask_combined].T

### 2020-06-16 analysis/process_paramsearch_comp2.py

@@ -169,7 +169,6 @@ for logdirname in os.listdir(basedir):
     #print ('t1: {0}'.format (t1))
     mapdiff = bdo.mapdiff[mask_combined].T
     area_diff = bdo.area_diff[mask_combined].T
-    area_diff = area_diff / np.max(area_diff)
     adj_arrangement = bdo.mean_adjacency_arrangement[mask_combined].T
     adj_diffmag = bdo.mean_adjacency_differencemag[mask_combined].T
 

### 2020-06-16 analysis/process_sensitivity_guide1.py

@@ -58,7 +58,6 @@ for logdirname in os.listdir(basedir):
     #print ('t1: {0}'.format (t1))
     mapdiff = bdo.mapdiff[mask_combined].T
     area_diff = bdo.area_diff[mask_combined].T
-    area_diff = area_diff / bdo.nhex
     sos_dist = np.sqrt(sos_dist.compressed()/bdo.N)
     hondadelta = hondadelta.compressed()
     # So here, I have t, honda, sos at 1 time point, providing one table line.

### 2020-06-16 paper/briefpaper/brief.tex
Modifies Fig 1 caption as discussed today

@@
       simulation approaching that obtained from \cmnt{real} barrels in \textbf{A} (dotted
-      line); black squares \cmnt{give a} measure \cmnt{of} the correspondence
+      line); black squares \cmnt{provide a} measure \cmnt{of} the correspondence
       between the real and
@@
       \cmnt{grey hexagons \stu{provide a measure of how selectively each
-          cortical site is innervated} OR: show the measure of localization, $\omega =
-        \oiint_{S} \left( \mathrm{max}_i(c_i)/\sum_{i=1}^{N} c_i\right) \mathrm{d}S$. }
+          cortical site is innervated},
+        $\omega = \oiint_{S} \mu(\mb{x}) \mathrm{d}S$, where
+        $\mu(\mb{x}) \equiv \mathrm{max}_i\big(c_i(\mb{x})\big)\big/\sum_{j=1}^{N} c_j(\mb{x})$.}
       \textbf{E} \stu{Plotted across the cortical sheet, the selectivity
         develops to reveal an alignment with the emergent barrel boundary
-        shapes} OR: \cmnt{The localization measure plotted across the domain
-        shows that the structure in the connection density which has developed
-        by $t=30000$ is largely absent at $t=1000$. The
-        colour map
-        indicates the values of  $\mathrm{max}_i(c_i)/\sum_{i=1}^{N} c_i$. Overlaid
-        contours show $c_i > 0.95 \mathrm{max}_i(c_i)$.} All scale bars 1\,mm.}
+        shapes}. \cmnt{The colour map
+        indicates the values of $\mu(\mb{x})$. Overlaid
+        contours show $c_i > 0.95 \mathrm{max}_i\big(c_i(\mb{x})\big)$.} All scale bars 1\,mm.}
     \label{fig:main}

### 2020-06-16 paper/briefpaper/brief.tex
Localization mostly replaced with selectivity

@@
     connections} to \emph{the sum of all connection densities} gives a measure
-  of the localization of the connection density. We computed this measure
-  across the entire barrel domain and named it $\omega$ (see
+  of the selectivity of the connection density, $\mu(\mb{x},t)$. We summed this measure
+  across the entire barrel domain and named it $\omega(t)$ (see
   Fig.\,\ref{fig:main}D, grey hexagons). $\omega$ increases during the first
   20000 simulation steps and converges on a stable
-  value. Fig.\,\ref{fig:main}E shows the localization measure plotted as a
+  value. Fig.\,\ref{fig:main}E shows the selectivity measure plotted as a
   luminance map, giving a more graded impression of the structure which
@@
         barrel area difference and $\omega$ (bottom row) is a measure of
-        localization (separation) of the connection densities for the 41
+        selectivity of the connection densities for the 41
         thalamocortical axon types. For each row, lighter (orange/yellow)
@@
   area indicates good pattern matching. Higher values of $\omega$ indicate
-  better localization and are coloured yellow. White squares indicate that the
+  greater selectivity and are coloured yellow. White squares indicate that the
   simulation was not stable for that parameter set.}
@@
   and $\epsilon$ were high. The barrel area metric, $\eta$ indicates a strong
-  preference for $D\approx0.5$, although this does not give the maximally
-  localized solution, which for $D$ should be small. Thus, there is a
-  trade-off between good localization and good Voroni pattern
+  preference for $D\approx0.5$, although this does not give the
+  solution of maximum selectivity, which for $D$ should be small. Thus, there is a
+  trade-off between good selection and good Voroni pattern
   conformation. The
@@
   $\delta$ (the Honda Voronoi metric), $\eta$ (the pattern-match metric) and
-  $\omega$ (which measures localization of connections, $c_i$) after running
+  $\omega$ (which measures selectivity of connections, $c_i$) after running
   the model for 30000 simulation steps.}
@@
   $\gamma_{i,j}$, which is 4). The Voronoi nature of the pattern ($\delta$)
-  and the localization of the fields are not strongly affected by interaction
+  and the selectivity of the fields are not strongly affected by interaction
   parameter noise, but the pattern metric, $\eta$ shows a... ADD SUB GRAPH.}

### 2020-06-16 paper/briefpaper/brief.tex
More tweaks to fig caption; adds new references

@@
       delineate boundaries (see Movie
-      S1). \textbf{D} Red dots show the Honda-$\delta$ metric obtained from \stu{the}
+      S1). \textbf{D} Red dots show the Honda metric, \cmnt{$\delta(t)$}, obtained from \stu{the}
       simulation approaching that obtained from \cmnt{real} barrels in \textbf{A} (dotted
@@
       between the real and
-      simulated barrel shapes, $\eta$ (the \cmnt{pattern quality metric---see
+      simulated barrel shapes, $\eta(t)$ (the \cmnt{pattern quality metric---see
         Materials \& Methods; units mm$^3$});
       \cmnt{grey hexagons \stu{provide a measure of how selectively each
-          cortical site is innervated},
-        $\omega = \oiint_{S} \mu(\mb{x}) \mathrm{d}S$, where
-        $\mu(\mb{x}) \equiv \mathrm{max}_i\big(c_i(\mb{x})\big)\big/\sum_{j=1}^{N} c_j(\mb{x})$.}
+          cortical site is innervated};
+        $\omega(t) = \oiint_{S} \mu(\mb{x},t) \mathrm{d}S$, where
+        $\mu(\mb{x}) \equiv \mathrm{max}_i\big(c_i(\mb{x},t)\big)\big/\sum_{j=1}^{N} c_j(\mb{x},t)$.}
       \textbf{E} \stu{Plotted across the cortical sheet, the selectivity
@@
 i.e., the putative barreloid center.
-\stu{As such, we consider linear gradients aligned to anterior-posterior and medial-lateral axes through the VPM, and as $\gamma$-values use the coordinates of the barreloid centers in these axes as they appear in \citep{haidarliu_size_2001} (Fig.\,\ref{fig:main}B).} \stu{Complementary thalamic and cortical molecular guidance gradients of this form have been well established experimentally \citep{vanderhaeghen_mapping_2000,miller_epha7-ephrin-a5_2006}, e.g., cells originating in the medial part of ventrobasal thalamus (comprising VPM and VPL) express high levels of EphA receptors and project to the lateral part of S1, which expresses low levels of ephrin-A5, and cells originating in the lateral part of ventrobasal thalamus express low levels of EphA receptors and project to the medial part of S1, which expresses high levels of ephrin-A5 (see Gao, 1998; Dufour et al., 2003; Vanderhaegen \& Polleux, 2004; Speer \& Chapman, 2005)}
+\stu{As such, we consider linear gradients aligned to anterior-posterior and medial-lateral axes through the VPM, and as $\gamma$-values use the coordinates of the barreloid centers in these axes as they appear in \citep{haidarliu_size_2001} (Fig.\,\ref{fig:main}B).} \stu{Complementary thalamic and cortical molecular guidance gradients of this form have been well established experimentally} \citep{vanderhaeghen_mapping_2000,miller_epha7-ephrin-a5_2006}, \stu{e.g., cells originating in the medial part of ventrobasal thalamus (comprising VPM and VPL) express high levels of EphA receptors and project to the lateral part of S1, which expresses low levels of ephrin-A5, and cells originating in the lateral part of ventrobasal thalamus express low levels of EphA receptors and project to the medial part of S1, which expresses high levels of ephrin-A5 (see} \citealp{gao_regulation_1998,dufour_area_2003,vanderhaeghen_developmental_2004,speer_grading_2005}).
 

### 2020-06-16 paper/briefpaper/brief.tex
Reruns the stats for my now correct version of eta and updates the paper accordingly.

@@
   results (where deviations occur as a result of the different random initial
-  conditions for $a_i$) were $\delta=0.089\,(0.004)$, $\eta=113.5\textbf{!!}\,(1.2)$ and
+  conditions for $a_i$) were $\delta=0.089\,(0.004)$, $\eta=0.2108\,(0.002)$ and
   $\omega=0.2165\,(0.0001)$. None of the three metrics were strongly affected
@@
   system 10 times and obtained the results $\delta=0.090\,(0.005)$,
-  $\eta=115\,(1.9)\textbf{!!}$ and $\omega=0.2166\,(0.0003)$. Paired two-sample t-tests
+  $\eta=0.2136\,(0.004)$ and $\omega=0.2166\,(0.0003)$. Paired two-sample t-tests
   demonstrated that the increase in noise amplitude made no significant
@@
   \mathrm{t}=-0.69,\mathrm{p}=0.49$,
-  $\eta: \mathrm{t}=-2.0,\mathrm{p}=0.06\textbf{!!}$,
+  $\eta: \mathrm{t}=-2.0,\mathrm{p}=0.06$,
   $\omega: \mathrm{t}=-1.7,\mathrm{p}=0.11$). 10 runs of a
@@
   the model is deterministic (all runs produced the results $\delta=0.0928$,
-  $\eta=114.31\textbf{!!}$ and $\omega=0.21649$) and that stochasticity in
+  $\eta=0.21236$ and $\omega=0.21649$) and that stochasticity in
   $a_i(\mb{x},t=0)$ is not required by this model, as all three results lay within

### 2020-06-16 analysis/process_multirun_anoise.py

@@ -50,7 +50,7 @@ for logdirname in os.listdir(basedir):
         print ('bdo.locn_vs_t: {0}'.format(bdo.locn_vs_t))
         hondas.append(bdo.honda[0])
         sos.append(bdo.sos_dist[0])
-        area_diff = bdo.area_diff[0] / bdo.nhex
+        area_diff = bdo.area_diff[0]
         area.append(area_diff)
 
         mapdiffs.append(bdo.mapdiff[0])

### 2020-06-17 paper/briefpaper/brief.tex
Adds a caption for Fig 3, the sensitivity analysis

@@
 \DeclarePairedDelimiter{\norm}{\lVert}{\rVert}
+% Degrees symbol:
+\usepackage{textcomp}
 
@@
     \includegraphics[width=\linewidth]{./Fig3.png}
-    \caption{\cmnt{Sensitivity analysis. Column \textbf{A} shows...}}
+    \caption{
+      \cmnt{\textbf{A} Sensitivity of the system pattern
+        metrics $\delta(t)$ (top row), $\omega(t)$ (middle row) and $\eta(t)$
+        (bottom row), evaluated at $t=30000$ steps,
+        to three classes of perturbation. $y$-axes and colour scales have identical ranges
+        to the colour scales in Fig.\,\ref{fig:paramsweep}, to ease comparison
+        with the parameter exploration.
+        %
+        \emph{Column 1:} Random noise drawn from a uniform distribution,
+        $(\gamma_{max}-\gamma_{min})\,U(0,\nu_\gamma)$, added
+        to the values for the interaction parameters, $\gamma_{i,j}$ used in
+        Fig.\,\ref{fig:main}.
+        %
+        \emph{Column 2:} Random noise $(\rho_{max} - \rho_{min})\,U(0, \nu_{\rho})$
+        was added to each hex element of $\rho_1(\mb{x})$ and $\rho_2(\mb{x})$ and the result
+        was smoothed by convolution with a symmetric 2D
+        Gaussian kernel of width $\sigma_\rho$. Colour maps in this column
+        thus show the effect of changing the magnitude and the length scale of
+        noise in the guidance fields.
+        %
+        \emph{Column 3:} Colour maps show the result of setting the rotational angle
+        of the linearly varying guidance field, $\rho_1(\mb{x})$, to $\phi_{\rho_1}$,
+        and modifying its overall gain to $G_{\rho_1}$, whilst keeping the
+        parameters of $\rho_2(\mb{x})$
+        unchanged from those used in the simulation in
+        Fig.\,\ref{fig:main}, for which $\phi_{\rho_1}=6$\textdegree~and
+        $\phi_{\rho_2}=84$\textdegree.
+        %
+        \textbf{B} Four ways in which the perturbations in \textbf{A} affect
+        the patterns. i) The effect of significant interaction parameter noise
+        causes barrels to become swapped from the positions they would
+        otherwise take. ii) High magnitude, short lengthscale noise in
+        $\rho(\mb{x})$ leads to `wiggly' boundaries between barrels. iii)
+        Shows the effect of reducing the slope of $\rho_1$ (by a factor of 10)
+        which reduces the strength of the axons' interaction with the guidance
+        field in the left-right sense. Barrel rows B, C and D become `crushed'
+        down the center line, and edge barrels dominate. iv) The effect of
+        rotating $\rho_1$ by 20\textdegree~is a relatively mild distortion of
+        the pattern in the anticlockwise sense.
+      }
+    }
     \label{fig:sens}
@@
 % delta   0.0928       0.089(0.004)    <1                   0.090 (0.005)    <1
-% eta     114.31       113.5(1.2)      <1                   115 (1.9)        <1
+% eta     0.2124       0.2108(0.002)   <1                   0.2136 (0.004)   <1
 % omega   0.21649      0.2165(0.0001)  <1                   0.2166 (0.0003)  <1

### 2020-06-18 paper/briefpaper/brief.tex
Adds text for minor point 6

@@
 \newcommand*\dif{\mathop{}\!\mathrm{d}}
+
+% Colours for revisions:
+% Custom colours
+\definecolor{revgreen}{rgb}{0.220, 0.463, 0.114}
+\definecolor{revblue}{rgb}{0.235, 0.471, 0.847}
+\definecolor{revred}{rgb}{1.0, 0.0, 0.0}
+\definecolor{revgrey}{rgb}{0.6, 0.6, 0.63}
+\definecolor{revcomp}{rgb}{0.0, 0.0, 0.0}
+\definecolor{revpurp}{rgb}{0.8, 0.1, 0.8}
+
+% Blue changes are mostly major points 2 (competition mechanism) and 3 (parameter/sensitivity)
 \newcommand{\cmnt}[1]{\textcolor{blue}{#1}}
+% Red for Stuart's updates
 \newcommand{\stu}[1]{\textcolor{red}{#1}}
+% Minor point 6:
+\newcommand{\mpsix}[1]{\textcolor{revpurp}{#1}}
+
 \newcommand{\dvrg}{\nabla\vcdot\nabla}
@@
 thalamocortical axons, which are enclosed by borders a few neurons thick from
-postnatal day 3 \citep{erzurumlu_development_2012}. In the plane tangential to
+postnatal day 3 \citep{erzurumlu_development_2012}.
+%
+% These timings, from erzurumlu_development_2012 originate from
+% sehara_neuronal_2011, where they relate to mouse.
+\mpsix{In the mouse, axons from the trigeminal ganglion arrive in the principal sensory
+  nucleus of the trigeminal nerve (PrV) at E12, then axons from the PrV arrive
+  in the ventroposteromedial nucleus of the thalamus (VPM) at E17, then axons
+  from the VPM arrive in the cortical plate at E18/P0. Distinct whisker-related
+  clusters then become apparent in the PrV at P0-P1, in the VPM at P2-P3, and
+  in the cortex at P3-P5} \citep{erzurumlu_development_2012,sehara_neuronal_2011}.
+%
+In the plane tangential to
 the cortical surface the barrels constitute a somatotopic map of the whiskers,

### 2020-06-18 paper/briefpaper/brief.tex
Adds text and reference for minor point 4

@@
 \newcommand{\stu}[1]{\textcolor{red}{#1}}
+% Minor point 4:
+\newcommand{\mpfour}[1]{\textcolor{revgreen}{#1}}
 % Minor point 6:
@@
 \citep{shimogori_fibroblast_2005,bishop_regulation_2000}, which are expressed
-in gradients that mark orthogonal axes and can be manipulated to stretch,
+in gradients \mpfour{across the full cortical sheet} that mark orthogonal axes
+and can be manipulated to stretch,
 shrink, shift, and even duplicate barrels
@@
 for $N=41$ projections on the resulting domain, using $M=2$ linear signalling
-gradients aligned with the anterior-posterior and medial-lateral axes. From
+gradients aligned with the anterior-posterior and medial-lateral axes.
+%
+\mpfour{The signalling gradients are assumed not to interact with the barrel
+  field boundary and to continue well beyond its extent (see
+  \citealp{miller_epha7-ephrin-a5_2006}, their Fig.\,11) in order to
+  participate in other developmental processes, such as inter-area positional
+  specification \citep{sur_patterning_2005,oleary_area_2007}.} % Just oleary_area_2007 would be ok
+%
+From
 random initial conditions for $a(\mb{x},0)$, \cmnt{with each hexagonal element

### 2020-06-19 paper/briefpaper/brief.tex
sw edits

--- 2020-06-17 paper/briefpaper/brief.tex
@@
 \DeclarePairedDelimiter{\norm}{\lVert}{\rVert}
-% Degrees symbol:
-\usepackage{textcomp}
 
@@
       delineate boundaries (see Movie
-      S1). \textbf{D} Red dots show the Honda metric, \cmnt{$\delta(t)$}, obtained from \stu{the}
+      S1). \textbf{D} Red dots show the Honda-$\delta$ metric obtained from \stu{the}
       simulation approaching that obtained from \cmnt{real} barrels in \textbf{A} (dotted
-      line); black squares \cmnt{provide a} measure \cmnt{of} the correspondence
+      line); black squares \cmnt{give a} measure \cmnt{of} the correspondence
       between the real and
-      simulated barrel shapes, $\eta(t)$ (the \cmnt{pattern quality metric---see
+      simulated barrel shapes, $\eta$ (the \cmnt{pattern quality metric---see
         Materials \& Methods; units mm$^3$});
       \cmnt{grey hexagons \stu{provide a measure of how selectively each
-          cortical site is innervated};
-        $\omega(t) = \oiint_{S} \mu(\mb{x},t) \mathrm{d}S$, where
-        $\mu(\mb{x}) \equiv \mathrm{max}_i\big(c_i(\mb{x},t)\big)\big/\sum_{j=1}^{N} c_j(\mb{x},t)$.}
+          cortical site is innervated} OR: show the measure of localization, $\omega =
+        \oiint_{S} \left( \mathrm{max}_i(c_i)/\sum_{i=1}^{N} c_i\right) \mathrm{d}S$. }
       \textbf{E} \stu{Plotted across the cortical sheet, the selectivity
         develops to reveal an alignment with the emergent barrel boundary
-        shapes}. \cmnt{The colour map
-        indicates the values of $\mu(\mb{x})$. Overlaid
-        contours show $c_i > 0.95 \mathrm{max}_i\big(c_i(\mb{x})\big)$.} All scale bars 1\,mm.}
+        shapes} OR: \cmnt{The localization measure plotted across the domain
+        shows that the structure in the connection density which has developed
+        by $t=30000$ is largely absent at $t=1000$. The
+        colour map
+        indicates the values of  $\mathrm{max}_i(c_i)/\sum_{i=1}^{N} c_i$. Overlaid
+        contours show $c_i > 0.95 \mathrm{max}_i(c_i)$.} All scale bars 1\,mm.}
     \label{fig:main}
@@
 i.e., the putative barreloid center.
-\stu{As such, we consider linear gradients aligned to anterior-posterior and medial-lateral axes through the VPM, and as $\gamma$-values use the coordinates of the barreloid centers in these axes as they appear in \citep{haidarliu_size_2001} (Fig.\,\ref{fig:main}B).} \stu{Complementary thalamic and cortical molecular guidance gradients of this form have been well established experimentally} \citep{vanderhaeghen_mapping_2000,miller_epha7-ephrin-a5_2006}, \stu{e.g., cells originating in the medial part of ventrobasal thalamus (comprising VPM and VPL) express high levels of EphA receptors and project to the lateral part of S1, which expresses low levels of ephrin-A5, and cells originating in the lateral part of ventrobasal thalamus express low levels of EphA receptors and project to the medial part of S1, which expresses high levels of ephrin-A5 (see} \citealp{gao_regulation_1998,dufour_area_2003,vanderhaeghen_developmental_2004,speer_grading_2005}).
-
-%\cmnt{In other words, we simply scaled the barreloid coordinates in an anterior-posterior/medial-lateral coordinate system to create the $\gamma$ values in the range [-2,2].
-Note that positive \emph{and} negative $\gamma$ values were generated; the origin of the coordinate system was located at the approximate center of the thalamic barreloid field.
-This prevented an overall drift of the cortical pattern away from the centroid of the barrel field and may explain why coordinate-giving transcription factors and the ephrins are often found as pairs acting in opposing directions.}  \cmnt{We concentrated on macrovibrissae (excluding microvibrissae) because it is easy to identify their corresponding structures both in the thalamus and in the cortex.
-We also excluded the 5 rhinal whiskers because their cortical representation is in a separate island, distinct, and somewhat removed from the representation of the macrovibirissae.}}
-
-%\cmnt{Thus, we suggest that there exist a pair of signalling molecules whose expression varies linearly (with orthogonal gradients) across the thalamus, analgous to the ephrins, whose existence is well known in the cortex.
-Thalamocortical axons growing from a region in the thalamus with high expression of one of these molecules will interact with a corresponding cortical signalling molecule more strongly than an axon emanating from a region with low expression of the thalamic signal.}
-
-%\cmnt{We defined two orthogonal, linear gradients oriented in the anterior-posterior and medial-lateral directions and used the coordinates of 41 barreloid centers recreated from \cite{haidarliu_size_2001}, corresponding to the macrovibrissae, to populate 41 pairs of $\gamma$ values.}  (Fig.\,\ref{fig:main}B).
-\cmnt{In other words, we simply scaled the barreloid coordinates in an anterior-posterior/medial-lateral coordinate system to create the $\gamma$ values in the range [-2,2].
-Note that positive \emph{and} negative $\gamma$ values were generated; the origin of the coordinate system was located at the approximate center of the thalamic barreloid field.
-This prevented an overall drift of the cortical pattern away from the centroid of the barrel field and may explain why coordinate-giving transcription factors and the ephrins are often found as pairs acting in opposing directions.}  \cmnt{We concentrated on macrovibrissae (excluding microvibrissae) because it is easy to identify their corresponding structures both in the thalamus and in the cortex.
-We also excluded the 5 rhinal whiskers because their cortical representation is in a separate island, distinct, and somewhat removed from the representation of the macrovibirissae.}
-
-A cortical boundary enclosing 41 \cmnt{macrovibrissae} barrels was traced from
+\stu{As such, we consider linear gradients aligned to anterior-posterior and medial-lateral axes through the VPM, and as $\gamma$-values use the coordinates of the barreloid centers in these axes as they appear in \citep{haidarliu_size_2001} (Fig.\,\ref{fig:main}B).} \stu{Complementary thalamic and cortical molecular guidance gradients of this form have been well established experimentally \citep{vanderhaeghen_mapping_2000,miller_epha7-ephrin-a5_2006}, e.g., cells originating in the medial part of ventrobasal thalamus (comprising VPM and VPL) express high levels of EphA receptors and project to the lateral part of S1, which expresses low levels of ephrin-A5, and cells originating in the lateral part of ventrobasal thalamus express low levels of EphA receptors and project to the medial part of S1, which expresses high levels of ephrin-A5 (see Gao, 1998; Dufour et al., 2003; Vanderhaegen \& Polleux, 2004; Speer \& Chapman, 2005).} \stu{We concentrate on the representation of the 41 macrovibrissae because their thalamic and cortical correlates are easily identifiable and consistently positioned, we exclude the microvibrissae to limit the complexity of the simulations, and we exclude the five rhinal whiskers as their representation in S1 is isolated from the main barrel field.} A cortical boundary enclosing \stu{the} 41 \cmnt{macrovibrissae} barrels was traced from
 a cytochrome oxidase stain from \cite{zheng_signal_2001} \cmnt{(using the
@@
 for $N=41$ projections on the resulting domain, using $M=2$ linear signalling
-gradients aligned with the anterior-posterior and medial-lateral axes. From
+gradients aligned with the anterior-posterior and medial-lateral axes.
+
+\stu{From}
 random initial conditions for $a(\mb{x},0)$, \cmnt{with each hexagonal element
@@
     connections} to \emph{the sum of all connection densities} gives a measure
-  of the selectivity of the connection density, $\mu(\mb{x},t)$. We summed this measure
-  across the entire barrel domain and named it $\omega(t)$ (see
+  of the localization of the connection density. We computed this measure
+  across the entire barrel domain and named it $\omega$ (see
   Fig.\,\ref{fig:main}D, grey hexagons). $\omega$ increases during the first
   20000 simulation steps and converges on a stable
-  value. Fig.\,\ref{fig:main}E shows the selectivity measure plotted as a
+  value. Fig.\,\ref{fig:main}E shows the localization measure plotted as a
   luminance map, giving a more graded impression of the structure which
@@
         barrel area difference and $\omega$ (bottom row) is a measure of
-        selectivity of the connection densities for the 41
+        localization (separation) of the connection densities for the 41
         thalamocortical axon types. For each row, lighter (orange/yellow)
@@
   area indicates good pattern matching. Higher values of $\omega$ indicate
-  greater selectivity and are coloured yellow. White squares indicate that the
+  better localization and are coloured yellow. White squares indicate that the
   simulation was not stable for that parameter set.}
@@
   and $\epsilon$ were high. The barrel area metric, $\eta$ indicates a strong
-  preference for $D\approx0.5$, although this does not give the
-  solution of maximum selectivity, which for $D$ should be small. Thus, there is a
-  trade-off between good selection and good Voroni pattern
+  preference for $D\approx0.5$, although this does not give the maximally
+  localized solution, which for $D$ should be small. Thus, there is a
+  trade-off between good localization and good Voroni pattern
   conformation. The
@@
     \includegraphics[width=\linewidth]{./Fig3.png}
-    \caption{
-      \cmnt{\textbf{A} Sensitivity of the system pattern
-        metrics $\delta(t)$ (top row), $\omega(t)$ (middle row) and $\eta(t)$
-        (bottom row), evaluated at $t=30000$ steps,
-        to three classes of perturbation. $y$-axes and colour scales have identical ranges
-        to the colour scales in Fig.\,\ref{fig:paramsweep}, to ease comparison
-        with the parameter exploration.
-        %
-        \emph{Column 1:} Random noise drawn from a uniform distribution,
-        $(\gamma_{max}-\gamma_{min})\,U(0,\nu_\gamma)$, added
-        to the values for the interaction parameters, $\gamma_{i,j}$ used in
-        Fig.\,\ref{fig:main}.
-        %
-        \emph{Column 2:} Random noise $(\rho_{max} - \rho_{min})\,U(0, \nu_{\rho})$
-        was added to each hex element of $\rho_1(\mb{x})$ and $\rho_2(\mb{x})$ and the result
-        was smoothed by convolution with a symmetric 2D
-        Gaussian kernel of width $\sigma_\rho$. Colour maps in this column
-        thus show the effect of changing the magnitude and the length scale of
-        noise in the guidance fields.
-        %
-        \emph{Column 3:} Colour maps show the result of setting the rotational angle
-        of the linearly varying guidance field, $\rho_1(\mb{x})$, to $\phi_{\rho_1}$,
-        and modifying its overall gain to $G_{\rho_1}$, whilst keeping the
-        parameters of $\rho_2(\mb{x})$
-        unchanged from those used in the simulation in
-        Fig.\,\ref{fig:main}, for which $\phi_{\rho_1}=6$\textdegree~and
-        $\phi_{\rho_2}=84$\textdegree.
-        %
-        \textbf{B} Four ways in which the perturbations in \textbf{A} affect
-        the patterns. i) The effect of significant interaction parameter noise
-        causes barrels to become swapped from the positions they would
-        otherwise take. ii) High magnitude, short lengthscale noise in
-        $\rho(\mb{x})$ leads to `wiggly' boundaries between barrels. iii)
-        Shows the effect of reducing the slope of $\rho_1$ (by a factor of 10)
-        which reduces the strength of the axons' interaction with the guidance
-        field in the left-right sense. Barrel rows B, C and D become `crushed'
-        down the center line, and edge barrels dominate. iv) The effect of
-        rotating $\rho_1$ by 20\textdegree~is a relatively mild distortion of
-        the pattern in the anticlockwise sense.
-      }
-    }
+    \caption{\cmnt{Sensitivity analysis. Column \textbf{A} shows...}}
     \label{fig:sens}
@@
   $\delta$ (the Honda Voronoi metric), $\eta$ (the pattern-match metric) and
-  $\omega$ (which measures selectivity of connections, $c_i$) after running
+  $\omega$ (which measures localization of connections, $c_i$) after running
   the model for 30000 simulation steps.}
@@
   results (where deviations occur as a result of the different random initial
-  conditions for $a_i$) were $\delta=0.089\,(0.004)$, $\eta=0.2108\,(0.002)$ and
+  conditions for $a_i$) were $\delta=0.089\,(0.004)$, $\eta=113.5\textbf{!!}\,(1.2)$ and
   $\omega=0.2165\,(0.0001)$. None of the three metrics were strongly affected
@@
   system 10 times and obtained the results $\delta=0.090\,(0.005)$,
-  $\eta=0.2136\,(0.004)$ and $\omega=0.2166\,(0.0003)$. Paired two-sample t-tests
+  $\eta=115\,(1.9)\textbf{!!}$ and $\omega=0.2166\,(0.0003)$. Paired two-sample t-tests
   demonstrated that the increase in noise amplitude made no significant
@@
   the model is deterministic (all runs produced the results $\delta=0.0928$,
-  $\eta=0.21236$ and $\omega=0.21649$) and that stochasticity in
+  $\eta=114.31\textbf{!!}$ and $\omega=0.21649$) and that stochasticity in
   $a_i(\mb{x},t=0)$ is not required by this model, as all three results lay within
@@
 % delta   0.0928       0.089(0.004)    <1                   0.090 (0.005)    <1
-% eta     0.2124       0.2108(0.002)   <1                   0.2136 (0.004)   <1
+% eta     114.31       113.5(1.2)      <1                   115 (1.9)        <1
 % omega   0.21649      0.2165(0.0001)  <1                   0.2166 (0.0003)  <1
@@
   $\gamma_{i,j}$, which is 4). The Voronoi nature of the pattern ($\delta$)
-  and the selectivity of the fields are not strongly affected by interaction
+  and the localization of the fields are not strongly affected by interaction
   parameter noise, but the pattern metric, $\eta$ shows a... ADD SUB GRAPH.}
@@
 cortical space, as represented by $\chi$, with an implicit requirement for a
-self/other identifier amoungst projections. This latter form of competition may
-account for the absence of barrels in rodents with larger brains, such as
+self/other identifier amoungst projections. This latter form of competition may account for the absence of barrels in rodents with larger brains, such as
 capybara, for which competition for space is presumably weaker
@@
 sensory organs \citep{purves_iterated_1992}.
+
+\stu{In reality, scores of smaller Dirichletform barrels representing the microvibrissae form to the right of where the E-row barrels emerge in our simulations, presumably via the same competitive processes.
+Enforcing here the same boundary condition as used to represent the true edges of the barrel field is necessary to ensure the stability of the simulation, though we acknowledge that this region of the boundary is enforced primarily to keep the number of simulated projections, and hence the overall computational complexity of the simulation, manageable (simulating each additional projection introduces an extra \textbf{YYY} new dynamical variables).}
 
@@
 \end{document}
+
+% NOTES
+
+%\cmnt{In other words, we simply scaled the barreloid coordinates in an anterior-posterior/medial-lateral coordinate system to create the $\gamma$ values in the range [-2,2].
+Note that positive \emph{and} negative $\gamma$ values were generated; the origin of the coordinate system was located at the approximate center of the thalamic barreloid field.
+This prevented an overall drift of the cortical pattern away from the centroid of the barrel field and may explain why coordinate-giving transcription factors and the ephrins are often found as pairs acting in opposing directions.}  \cmnt{We concentrated on macrovibrissae (excluding microvibrissae) because it is easy to identify their corresponding structures both in the thalamus and in the cortex.
+We also excluded the 5 rhinal whiskers because their cortical representation is in a separate island, distinct, and somewhat removed from the representation of the macrovibirissae.}}
+
+%\cmnt{Thus, we suggest that there exist a pair of signalling molecules whose expression varies linearly (with orthogonal gradients) across the thalamus, analgous to the ephrins, whose existence is well known in the cortex.
+Thalamocortical axons growing from a region in the thalamus with high expression of one of these molecules will interact with a corresponding cortical signalling molecule more strongly than an axon emanating from a region with low expression of the thalamic signal.}
+
+%\cmnt{We defined two orthogonal, linear gradients oriented in the anterior-posterior and medial-lateral directions and used the coordinates of 41 barreloid centers recreated from \cite{haidarliu_size_2001}, corresponding to the macrovibrissae, to populate 41 pairs of $\gamma$ values.}  (Fig.\,\ref{fig:main}B).
+\cmnt{In other words, we simply scaled the barreloid coordinates in an anterior-posterior/medial-lateral coordinate system to create the $\gamma$ values in the range [-2,2].
+Note that positive \emph{and} negative $\gamma$ values were generated; the origin of the coordinate system was located at the approximate center of the thalamic barreloid field.
+This prevented an overall drift of the cortical pattern away from the centroid of the barrel field and may explain why coordinate-giving transcription factors and the ephrins are often found as pairs acting in opposing directions.}  \cmnt{We concentrated on macrovibrissae (excluding microvibrissae) because it is easy to identify their corresponding structures both in the thalamus and in the cortex.
+We also excluded the 5 rhinal whiskers because their cortical representation is in a separate island, distinct, and somewhat removed from the representation of the macrovibirissae.}

### 2020-06-19 paper/briefpaper/brief.tex

--- 2020-06-18 paper/briefpaper/brief.tex
@@
 determined by the concentration of a similar molecule at its thalamic origin,
+
 i.e., the putative barreloid center.
 \stu{As such, we consider linear gradients aligned to anterior-posterior and medial-lateral axes through the VPM, and as $\gamma$-values use the coordinates of the barreloid centers in these axes as they appear in \citep{haidarliu_size_2001} (Fig.\,\ref{fig:main}B).} \stu{Complementary thalamic and cortical molecular guidance gradients of this form have been well established experimentally} \citep{vanderhaeghen_mapping_2000,miller_epha7-ephrin-a5_2006}, \stu{e.g., cells originating in the medial part of ventrobasal thalamus (comprising VPM and VPL) express high levels of EphA receptors and project to the lateral part of S1, which expresses low levels of ephrin-A5, and cells originating in the lateral part of ventrobasal thalamus express low levels of EphA receptors and project to the medial part of S1, which expresses high levels of ephrin-A5 (see} \citealp{gao_regulation_1998,dufour_area_2003,vanderhaeghen_developmental_2004,speer_grading_2005}).
+\stu{We concentrate on the representation of the 41 macrovibrissae because their thalamic and cortical correlates are easily identifiable and consistently positioned, we exclude the microvibrissae to limit the complexity of the simulations, and we exclude the five rhinal whiskers as their representation in S1 is isolated from the main barrel field.} A cortical boundary enclosing \stu{the} 41 \cmnt{macrovibrissae} barrels was traced from
+a cytochrome oxidase stain from \cite{zheng_signal_2001} \cmnt{(using the
+  authors' original data)}, and Eqs.~\ref{eq:dc}--\ref{eq:comp} were solved
+for $N=41$ projections on the resulting domain, using $M=2$ linear signalling
+gradients aligned with the anterior-posterior and medial-lateral axes.
 
@@
 From
+>>>>>>> 7ffae6aecb56153dad7c6dd9a5aa658a6d9e293d
 random initial conditions for $a(\mb{x},0)$, \cmnt{with each hexagonal element

### 2020-06-19 paper/briefpaper/brief.tex
resolved conflict

@@
 velocity at which a projection migrates across the cortical subplate, is
-determined by the concentration of a similar molecule at its thalamic origin,
+determined by the concentration of a similar molecule at its thalamic origin,\stu{We concentrate on the representation of the 41 macrovibrissae because their thalamic and cortical correlates are easily identifiable and consistently positioned, we exclude the microvibrissae to limit the complexity of the simulations, and we exclude the five rhinal whiskers as their representation in S1 is isolated from the main barrel field.} A cortical boundary enclosing \stu{the} 41 \cmnt{macrovibrissae} barrels was traced from
+a cytochrome oxidase stain from \cite{zheng_signal_2001} \cmnt{(using the
+  authors' original data)}, and Eqs.~\ref{eq:dc}--\ref{eq:comp} were solved
+for $N=41$ projections on the resulting domain, using $M=2$ linear signalling
+gradients aligned with the anterior-posterior and medial-lateral axes.
 
@@
 
-A cortical boundary enclosing 41 \cmnt{macrovibrissae} barrels was traced from
-a cytochrome oxidase stain from \cite{zheng_signal_2001} \cmnt{(using the
-  authors' original data)}, and Eqs.~\ref{eq:dc}--\ref{eq:comp} were solved
-for $N=41$ projections on the resulting domain, using $M=2$ linear signalling
-gradients aligned with the anterior-posterior and medial-lateral axes.
-%
 \mpfour{The signalling gradients are assumed not to interact with the barrel
@@
 %
-From
->>>>>>> 7ffae6aecb56153dad7c6dd9a5aa658a6d9e293d
-random initial conditions for $a(\mb{x},0)$, \cmnt{with each hexagonal element
+From random initial conditions for $a(\mb{x},0)$, \cmnt{with each hexagonal element
   of $a_i$ set from a uniform distribution according to $0.2+{U}(0,0.2)$, and

### 2020-06-19 paper/briefpaper/brief.tex
more resolved conflicts

@@
 velocity at which a projection migrates across the cortical subplate, is
-determined by the concentration of a similar molecule at its thalamic origin,\stu{We concentrate on the representation of the 41 macrovibrissae because their thalamic and cortical correlates are easily identifiable and consistently positioned, we exclude the microvibrissae to limit the complexity of the simulations, and we exclude the five rhinal whiskers as their representation in S1 is isolated from the main barrel field.} A cortical boundary enclosing \stu{the} 41 \cmnt{macrovibrissae} barrels was traced from
-a cytochrome oxidase stain from \cite{zheng_signal_2001} \cmnt{(using the
-  authors' original data)}, and Eqs.~\ref{eq:dc}--\ref{eq:comp} were solved
-for $N=41$ projections on the resulting domain, using $M=2$ linear signalling
-gradients aligned with the anterior-posterior and medial-lateral axes.
+determined by the concentration of a similar molecule at its thalamic origin, i.e., the putative barreloid center.
 
-i.e., the putative barreloid center.
 \stu{As such, we consider linear gradients aligned to anterior-posterior and medial-lateral axes through the VPM, and as $\gamma$-values use the coordinates of the barreloid centers in these axes as they appear in \citep{haidarliu_size_2001} (Fig.\,\ref{fig:main}B).} \stu{Complementary thalamic and cortical molecular guidance gradients of this form have been well established experimentally} \citep{vanderhaeghen_mapping_2000,miller_epha7-ephrin-a5_2006}, \stu{e.g., cells originating in the medial part of ventrobasal thalamus (comprising VPM and VPL) express high levels of EphA receptors and project to the lateral part of S1, which expresses low levels of ephrin-A5, and cells originating in the lateral part of ventrobasal thalamus express low levels of EphA receptors and project to the medial part of S1, which expresses high levels of ephrin-A5 (see} \citealp{gao_regulation_1998,dufour_area_2003,vanderhaeghen_developmental_2004,speer_grading_2005}).
-\stu{We concentrate on the representation of the 41 macrovibrissae because their thalamic and cortical correlates are easily identifiable and consistently positioned, we exclude the microvibrissae to limit the complexity of the simulations, and we exclude the five rhinal whiskers as their representation in S1 is isolated from the main barrel field.} A cortical boundary enclosing \stu{the} 41 \cmnt{macrovibrissae} barrels was traced from
-a cytochrome oxidase stain from \cite{zheng_signal_2001} \cmnt{(using the
-  authors' original data)}, and Eqs.~\ref{eq:dc}--\ref{eq:comp} were solved
-for $N=41$ projections on the resulting domain, using $M=2$ linear signalling
-gradients aligned with the anterior-posterior and medial-lateral axes.
+\stu{We concentrate on the representation of the 41 macrovibrissae because their thalamic and cortical correlates are easily identifiable and consistently positioned, we exclude the microvibrissae to limit the complexity of the simulations, and we exclude the five rhinal whiskers as their representation in S1 is isolated from the main barrel field.}
+
+A cortical boundary enclosing \stu{the} 41 \cmnt{macrovibrissae} barrels was traced from
+a cytochrome oxidase stain from \cite{zheng_signal_2001} \cmnt{(using original data kindly supplied by the authors)}, and Eqs.~\ref{eq:dc}--\ref{eq:comp} were solved for $N=41$ projections on the resulting domain, using $M=2$ linear signalling gradients aligned with the anterior-posterior and medial-lateral axes.
 

### 2020-06-29 paper/briefpaper/brief.tex
Adds an initial Fig 5 (but I'm not happy with it yet) and does some word wrapping of brief.text

@@
 
-\documentclass[9pt,lineno,draft]{elife}
+\documentclass[9pt,lineno]{elife}
 
@@
 \stu{This term contributes to the \emph{flux of axonal branching} as an
- additional source of diffusion, which reduces the branching density for a given projection where the branches of other projections are dense.} \stu{The strength of this additional competition is parameterized by $\epsilon$}.
-Note that this operation is local to individual afferent
+  additional source of diffusion, which reduces the branching density for a
+  given projection where the branches of other projections are dense.}
+\stu{The strength of this additional competition is parameterized by
+  $\epsilon$}. Note that this operation is local to individual afferent
 projections.
@@
 
-\stu{As such, we consider linear gradients aligned to anterior-posterior and medial-lateral axes through the VPM, and as $\gamma$-values use the coordinates of the barreloid centers in these axes as they appear in \citep{haidarliu_size_2001} (Fig.\,\ref{fig:main}B).} \stu{Complementary thalamic and cortical molecular guidance gradients of this form have been well established experimentally} \citep{vanderhaeghen_mapping_2000,miller_epha7-ephrin-a5_2006}, \stu{e.g., cells originating in the medial part of ventrobasal thalamus (comprising VPM and VPL) express high levels of EphA receptors and project to the lateral part of S1, which expresses low levels of ephrin-A5, and cells originating in the lateral part of ventrobasal thalamus express low levels of EphA receptors and project to the medial part of S1, which expresses high levels of ephrin-A5 (see} \citealp{gao_regulation_1998,dufour_area_2003,vanderhaeghen_developmental_2004,speer_grading_2005}).
-\stu{We concentrate on the representation of the 41 macrovibrissae because their thalamic and cortical correlates are easily identifiable and consistently positioned, we exclude the microvibrissae to limit the complexity of the simulations, and we exclude the five rhinal whiskers as their representation in S1 is isolated from the main barrel field.}
+\stu{As such, we consider linear gradients aligned to anterior-posterior and
+  medial-lateral axes through the VPM, and as $\gamma$-values use the
+  coordinates of the barreloid centers in these axes as they appear in
+  \citep{haidarliu_size_2001} (Fig.\,\ref{fig:main}B).} \stu{Complementary
+  thalamic and cortical molecular guidance gradients of this form have been
+  well established experimentally}
+\citep{vanderhaeghen_mapping_2000,miller_epha7-ephrin-a5_2006}, \stu{e.g.,
+  cells originating in the medial part of ventrobasal thalamus (comprising VPM
+  and VPL) express high levels of EphA receptors and project to the lateral
+  part of S1, which expresses low levels of ephrin-A5, and cells originating
+  in the lateral part of ventrobasal thalamus express low levels of EphA
+  receptors and project to the medial part of S1, which expresses high levels
+  of ephrin-A5 (see}
+\citealp{gao_regulation_1998,dufour_area_2003,vanderhaeghen_developmental_2004,speer_grading_2005}). \stu{We
+  concentrate on the representation of the 41 macrovibrissae because their
+  thalamic and cortical correlates are easily identifiable and consistently
+  positioned, we exclude the microvibrissae to limit the complexity of the
+  simulations, and we exclude the five rhinal whiskers as their representation
+  in S1 is isolated from the main barrel field.}
 
@@
 comprising two separate, identically shaped cortical boundaries. In this case
-one distinct cortical field per thalamic projection emerged, within one of the
+\cmnt{NOTE: The following only holds for the divisive normalization system -
+  in the comp2 system, there's no `coupling' at all between the two regions,
+  so the mirrored region arranges itself as expected with the one reversed
+  guidance field.} one distinct cortical field per thalamic projection emerged, within one of the
 two boundaries only, i.e., no duplication. However, merging the two fields to
@@
   fields.}
+
+\begin{center}
+  \includegraphics[width=\linewidth]{./Fig5.png}
+  \captionof{figure}{\cmnt{Simulating whisker trimming by reduction of
+      $\epsilon$ for the thalamocortical projection for barrel C3, for which
+      $i=13$. All other model parameters match those in
+      Fig.\,\ref{fig:main}C. \textbf{A} Graph of individual barrel measures
+      plotted vs.~the multiplier applied to $\epsilon_{13}$. \textbf{B} The
+      pattern for $\epsilon_{m}=0.64$.}}
+  \label{fig:whiskertrim}
+\end{center}
 

### 2020-06-29 analysis/plot_trim.py
The rest of the changes for the whisker trimming, including necessary changes in the simulation to save out individual area diffs and so on

@@ -1,0 +1,118 @@
+# Make the whisker trimming line plot (barrel area vs epsilon, etc)
+
+# To access argv and also include the include dir
+import sys
+sys.path.insert (0, './include')
+import numpy as np
+# Import data loading code
+#import load as ld
+import BarrelData as bd
+# Import MY plotting code:
+import plot as pt
+import matplotlib
+import matplotlib.pyplot as plt
+# Direct HDF5 access
+import h5py
+# My domcentres linear fit code:
+import domcentres as dc
+import sebcolour
+col = sebcolour.Colour
+
+# Set plotting defaults
+fs = 26
+fnt = {'family' : 'Arial',
+       'weight' : 'regular',
+       'size'   : fs}
+matplotlib.rc('font', **fnt)
+
+# IMPORTANT for svg output of text as things that can be edited in inkscape
+plt.rcParams['svg.fonttype'] = 'none'
+
+# And plot this longhand here:
+F1 = plt.figure (figsize=(10,12))
+
+## load data from csv
+trimdata = np.genfromtxt ('postproc/whisker_trim_individual.csv', delimiter=",", names=True)
+td = np.sort (trimdata, order='barrel')
+
+# Show an example map for this multiplier
+map_mult = 0.64
+
+# Get per-barrel results
+# barrels of interest are: B3, C2, C3, C4, D2, D3
+#              TC indices:  8, 12, 13, 14, 22, 23
+tdc3 = td[td[:]['barrel_index'] == 13] # C3
+tdb3 = td[td[:]['barrel_index'] == 8]  # B3
+tdc2 = td[td[:]['barrel_index'] == 12]
+tdc4 = td[td[:]['barrel_index'] == 14]
+tdd2 = td[td[:]['barrel_index'] == 22]
+tdd3 = td[td[:]['barrel_index'] == 23]
+
+ax1 = F1.add_subplot(3,1,1)
+ax2 = F1.add_subplot(3,1,3)
+ax3 = F1.add_subplot(3,1,2)
+
+xmax = max(tdc3[:]['eps_mult'])
+xmin = min(tdc3[:]['eps_mult'])
+
+others_size = 6
+main_size = 12
+
+col_b3 = col.blue3
+col_c2 = col.navy
[…]

### 2020-06-29 analysis/process_trim.py

@@ -81,7 +81,8 @@ for logdirname in os.listdir(basedir):
             hond_dj = bdo.honda_delta_j[0][int(i)]
         except:
             hond_dj = math.nan
-        tableline = [eps_mult, epsilon, eps_modified, bdo.name_by_index[i], hond_dj, bdo.area_diffs[int(i)][0], bdo.barrel_areas[int(i)][0]]
+        print ('name_by_index: {0}'.format (bdo.name_by_index[i]))
+        tableline = [eps_mult, epsilon, eps_modified, bdo.name_by_index[i], i, hond_dj, bdo.area_diffs[int(i)][0], bdo.barrel_areas[int(i)][0]]
         print ('TL: {0}'.format(tableline))
         table_ind.append (tableline)
 
@@ -93,6 +94,6 @@ with open('postproc/whisker_trim_overall.csv', 'w', newline='\n') as csvfile:
         cw.writerow (tableline)
 with open('postproc/whisker_trim_individual.csv', 'w', newline='\n') as csvfile:
     cw = csv.writer (csvfile, delimiter=',')
-    cw.writerow (['eps_mult','epsilon','eps_modified','barrel','hondadelta','area_diff','area'])
+    cw.writerow (['eps_mult','epsilon','eps_modified','barrel','barrel_index','hondadelta','area_diff','area'])
     for tableline in table_ind:
         cw.writerow (tableline)

### 2020-06-29 sim/rd_james.h

@@ -349,9 +349,11 @@ public:
     //! The overall Honda 1983 Dirichlet approximation. 0.003 is a good fit. 0.05 not
     //! so good.
     Flt honda = 0.0;
+    std::map<Flt, Flt> honda_arr;
 
     //! Honda Dirichlet approx for the experimentally supplied barrels
     Flt expt_honda = 0.0;
+    std::map<Flt, Flt> expt_honda_arr;
     //@}
 
     /*!
@@ -366,6 +368,7 @@ public:
      * between the experimental and simulated barrel fields.
      */
     Flt area_diff = 0.0;
+    std::vector<Flt> area_diffs;
 
     /*!
      * Another metric to determine the difference between the current pattern and the
@@ -578,6 +581,8 @@ public:
         this->resize_vector_array_vector (this->J, this->N);
 
         this->resize_vector_variable (this->expt_barrel_id);
+
+        this->area_diffs.resize (this->N, Flt{0});
 
         // rhomethod is a vector of size M
         this->rhoMethod.resize (this->M);
@@ -895,10 +900,22 @@ public:
         }
         // Save the overall honda value for the system
         data.add_val ("/honda", this->honda);
+        // And individual honda delta_js:
+        // We have to make two vectors out of the map.
+        std::vector<Flt> ha_keys;
+        std::vector<Flt> ha_vals;
+        for (auto mi : this->honda_arr) {
+            ha_keys.push_back (mi.first);
+            ha_vals.push_back (mi.second);
+        }
+        data.add_contained_vals ("/honda_arr_keys", ha_keys);
+        data.add_contained_vals ("/honda_arr_vals", ha_vals);
+
         // Save sum of square distances
         data.add_val ("/sos_distances", this->sos_distances);
         // Save sum of square of area differences
         data.add_val ("/area_diff", this->area_diff);
+        data.add_contained_vals ("/area_diffs", this->area_diffs);
         // The difference between the experimental barrel map and the simulated map
         data.add_val ("/mapdiff", this->mapdiff);
         // Save the region centroids
@@ -1435,7 +1452,7 @@ public:
         this->domains = morph::ShapeAnalysis<Flt>::dirichlet_vertices (this->hg, this->expt_barrel_id, this->vertices);
         // Carry out the analysis.
         std::vector<std::pair<float, float>> d_centres;
[…]

### 2020-06-29 paper/briefpaper/brief.tex
Bringing in changes from Stuarts google doc, and preparing to shift as much paramsearch as possible into figure caption

@@
 \definecolor{revblue}{rgb}{0.235, 0.471, 0.847}
-\definecolor{revred}{rgb}{1.0, 0.0, 0.0}
+\definecolor{revred}{rgb}{0.8353, 0.1255, 0.1255}
+\definecolor{revorange}{rgb}{0.8235, 0.4275, 0.0}
 \definecolor{revgrey}{rgb}{0.6, 0.6, 0.63}
@@
 \definecolor{revpurp}{rgb}{0.8, 0.1, 0.8}
+\definecolor{revpurp2}{rgb}{0.3882, 0.0745, 0.7098}
+\definecolor{revpink}{rgb}{0.7098, 0.0745, 0.6431}
+\definecolor{revbblack}{rgb}{0, 0, 0.3}
+\definecolor{revgblack}{rgb}{0, 0.2, 0}
 
-% Blue changes are mostly major points 2 (competition mechanism) and 3 (parameter/sensitivity)
+% Blue changes are major point 2 (competition mechanism) and other, general
+% uncategorised changes.
 \newcommand{\cmnt}[1]{\textcolor{blue}{#1}}
-% Red for Stuart's updates
+% Red for Stuart's first updates
 \newcommand{\stu}[1]{\textcolor{red}{#1}}
@@
 \newcommand{\mpsix}[1]{\textcolor{revpurp}{#1}}
+% Justify competition:
+\newcommand{\justifycomp}[1]{\textcolor{revpurp2}{#1}}
+% Justify gamma/interaction parameter choice
+\newcommand{\justifygammas}[1]{\textcolor{revpink}{#1}}
+% Describe the metrics
+\newcommand{\metrics}[1]{\textcolor{revred}{#1}}
+% Parameter search. Major point 3
+\newcommand{\paramsearch}[1]{\textcolor{revbblack}{#1}}
+% Sensitivity analysis. Major point 3
+\newcommand{\sensitivity}[1]{\textcolor{revgblack}{#1}}
 
@@
       delineate boundaries (see Movie
-      S1). \textbf{D} Red dots show the Honda metric, \cmnt{$\delta(t)$}, obtained from \stu{the}
+      S1). \textbf{D} Red dots show the Honda metric, \cmnt{$\delta(t)$}, obtained from \cmnt{the}
       simulation approaching that obtained from \cmnt{real} barrels in \textbf{A} (dotted
@@
         Materials \& Methods; units mm$^3$});
-      \cmnt{grey hexagons \stu{provide a measure of how selectively each
+      \cmnt{grey hexagons \cmnt{provide a measure of how selectively each
           cortical site is innervated};
@@
         $\mu(\mb{x}) \equiv \mathrm{max}_i\big(c_i(\mb{x},t)\big)\big/\sum_{j=1}^{N} c_j(\mb{x},t)$.}
-      \textbf{E} \stu{Plotted across the cortical sheet, the selectivity
+      \textbf{E} \cmnt{Plotted across the cortical sheet, the selectivity
         develops to reveal an alignment with the emergent barrel boundary
@@
 projections are influenced by interactions with two or more signalling
-gradients that encourage migration in opposing directions. Note that these
-dynamics are quite unlike classic chemospecificity models
-\citep{sperry_chemoaffinity_1963}, which essentially assume center-points
-i.e., conditions in the target tissue that instruct pre-identified afferents
-to stop growing. As the number of distinct guidance fields is unlikely to
-approach the number of barrels, further extensions were required.
+gradients that encourage migration in opposing directions.
+%
+%Note that these dynamics are quite unlike classic chemospecificity models
+%\citep{sperry_chemoaffinity_1963}, which essentially assume center-points
+%i.e., conditions in the target tissue that instruct pre-identified afferents
+%to stop growing.
+%
+As the number of distinct guidance fields is unlikely to
+approach the number of barrels, \cmnt{modifications to the model} were required.
 
+% Justify competition term. This pulled from Stuart's Google doc
+\justifycomp{We reasoned that an arbitrary number of distinct field locations
+  may be specified using a minimum of two guidance fields, if the tendency for
+  the density of each projection to concentrate is appropriately balanced
+  against the tendency of projections that interact more strongly with the
+  guidance fields to migrate further in the directions specified by their
+  gradients. Accordingly, projections that interact most strongly with a given
+  guidance gradient would come to occupy cortical locations at which that
+  field has extreme values, leaving adjacent locations available to be
+  occupied by projections with the next strongest interactions, and so
+  forth. This would in principle allow the \emph{relative} locations of the
+  fields to be specified by the relative values of the interaction parameters,
+  $\gamma$, and hence for a topological map in the cortex to be specified by a
+  spatial ordering of the $\gamma$ values at the level of the thalamus.}
+
+\justifycomp{Such dynamics are quite unlike those described by classic
+  chemospecificity models} \citep{sperry_chemoaffinity_1963},
+\justifycomp{which essentially assume center-points by specifying conditions
+  in the target tissue that instruct pre-identified afferents to stop
+  growing. Consider, for example, that when simulated in isolation from
+  one-another, all projections in the model described would simply migrate to
+  the extrema of the cortical guidance fields.}
+
+\justifycomp{Testing this reasoning required increasing the strength of the
+  competition between simulated thalamocortical projections for cortical
+  territory, by increasing the tendency for the density of each projection to
+  concentrate. The major modification required was thus to introduce into the
+  model an additional source of competition between thalamic projections.}
+%
 The term in parentheses in Eq.\,\ref{eq:dc} represents competition between
@@
 %
-\stu{This term contributes to the \emph{flux of axonal branching} as an
-  additional source of diffusion, which reduces the branching density for a
-  given projection where the branches of other projections are dense.}
-\stu{The strength of this additional competition is parameterized by
+\justifycomp{This term contributes to the \emph{flux} of axonal branching as
+  an additional source of diffusion, which reduces the branching density for a
+  given projection where the branches of other projections are dense. The
+  strength of this additional competition is parameterized by
   $\epsilon$}. Note that this operation is local to individual afferent
@@
 
-The only differences between thalamic projections are their strengths of
-interaction with the guidance fields, $\gamma$, hence any reliable differences
-between emergent cortical fields must be due to differences in these values
-only. We speculate that the contribution of a given ephrin field, to the
-velocity at which a projection migrates across the cortical subplate, is
-determined by the concentration of a similar molecule at its thalamic origin, i.e., the putative barreloid center.
+\justifygammas{In addition, the model outlined requires that molecular
+  guidance gradients in the cortex are complemented by graded values of the
+  interaction strengths, $\gamma$, at the level of the thalamus. While the
+  precise mechanisms by thalamic and cortical gradients interact during
+  development have not been fully characterised, the presence of complementary
+  thalamic and cortical molecular guidance gradients has been well established
+  experimentally. In particular, the EphA4 receptor and its ligand ephrin-A5
+  are distributed in complementary gradients in the somatosensory thalamus and
+  cortex} (\citealp{vanderhaeghen_mapping_2000,miller_epha7-ephrin-a5_2006}).
+\justifygammas{Cells originating in the medial part of ventrobasal thalamus
+  (comprising VPM and VPL) express high levels of EphA receptors and project
+  to the lateral part of S1, which expresses low levels of ephrin-A5, and
+  cells originating in the lateral part of ventrobasal thalamus express low
+  levels of EphA receptors and project to the medial part of S1, which
+  expresses high levels of ephrin-A5} (see
+\citealp{gao_regulation_1998,dufour_area_2003,vanderhaeghen_developmental_2004,speer_grading_2005,torii_role_2013}).
 
-\stu{As such, we consider linear gradients aligned to anterior-posterior and
-  medial-lateral axes through the VPM, and as $\gamma$-values use the
-  coordinates of the barreloid centers in these axes as they appear in
-  \citep{haidarliu_size_2001} (Fig.\,\ref{fig:main}B).} \stu{Complementary
-  thalamic and cortical molecular guidance gradients of this form have been
-  well established experimentally}
-\citep{vanderhaeghen_mapping_2000,miller_epha7-ephrin-a5_2006}, \stu{e.g.,
-  cells originating in the medial part of ventrobasal thalamus (comprising VPM
-  and VPL) express high levels of EphA receptors and project to the lateral
-  part of S1, which expresses low levels of ephrin-A5, and cells originating
-  in the lateral part of ventrobasal thalamus express low levels of EphA
-  receptors and project to the medial part of S1, which expresses high levels
-  of ephrin-A5 (see}
-\citealp{gao_regulation_1998,dufour_area_2003,vanderhaeghen_developmental_2004,speer_grading_2005}). \stu{We
-  concentrate on the representation of the 41 macrovibrissae because their
-  thalamic and cortical correlates are easily identifiable and consistently
-  positioned, we exclude the microvibrissae to limit the complexity of the
-  simulations, and we exclude the five rhinal whiskers as their representation
-  in S1 is isolated from the main barrel field.}
+\justifygammas{For simplicity, both the thalamic interaction gradients and the
+  cortical guidance fields in the model were chosen in the first instance to
+  be linear and orthogonal. Hence a given pair of gamma values corresponds to
+  the coordinate of a barreloid center in the VPM (in a reference plane
+  defined by the anterior-posterior and medial-lateral axes. Coordinates were
+  estimated from Fig 5d of} \citep{haidarliu_size_2001}, \justifygammas{and
+  scaled (arbitrarily) such that $\gamma\in\pm 2$.}
 
-A cortical boundary enclosing \stu{the} 41 \cmnt{macrovibrissae} barrels was traced from
-a cytochrome oxidase stain from \cite{zheng_signal_2001} \cmnt{(using original data kindly supplied by the authors)}, and Eqs.~\ref{eq:dc}--\ref{eq:comp} were solved for $N=41$ projections on the resulting domain, using $M=2$ linear signalling gradients aligned with the anterior-posterior and medial-lateral axes.
+A cortical boundary enclosing \cmnt{the} 41 \cmnt{macrovibrissae} barrels was
+traced from a cytochrome oxidase stain from \cite{zheng_signal_2001}
+\cmnt{(using original data kindly supplied by the authors)}, and
+Eqs.~\ref{eq:dc}--\ref{eq:comp} were solved for $N=41$ projections on the
+resulting domain, using $M=2$ linear signalling gradients aligned with the
+anterior-posterior and medial-lateral axes.  \mpfour{These gradients may
+  extend well beyond the boundary of the barrel field (e.g. as ephrin-5 does;
+  see \citealp{miller_epha7-ephrin-a5_2006}, their Fig.\,11) but they are
+  shown within the barrel fields boundary in Fig.\,\ref{fig:main}A to aid the
+  visualization. Simulations were iterated through 30000 iterations of
+  Eqs.\,\ref{eq:dc}--\ref{eq:comp} ($\delta t=0.0001$)}
 
-%\cmnt{In other words, we simply scaled the barreloid coordinates in an anterior-posterior/medial-lateral coordinate system to create the $\gamma$ values in the range [-2,2].
-Note that positive \emph{and} negative $\gamma$ values were generated; the origin of the coordinate system was located at the approximate center of the thalamic barreloid field.
-This prevented an overall drift of the cortical pattern away from the centroid of the barrel field and may explain why coordinate-giving transcription factors and the ephrins are often found as pairs acting in opposing directions.}  \cmnt{We concentrated on macrovibrissae (excluding microvibrissae) because it is easy to identify their corresponding structures both in the thalamus and in the cortex.
-We also excluded the 5 rhinal whiskers because their cortical representation is in a separate island, distinct, and somewhat removed from the representation of the macrovibirissae.}}
+From initial conditions sample from a uniform random distribution for
+$a(\mb{x},0)\in(0.2,0.4)$ and $c(\mb{x},0)=0$, and across a wide range of
+parameter values,
+%From random initial conditions for $a(\mb{x},0)$, \cmnt{with each hexagonal element
+%  of $a_i$ set from a uniform distribution according to $0.2+{U}(0,0.2)$, and
+%  $c(\mb{x},0)=0$},
+a clear Voronoi-like tesselation \cmnt{of topographically organised
+  projections} emerged, \cmnt{confirming that barrel maps can self-organise in
+  the absence of pre-specified center points. Parameters for the example
+  simulation shown in Fig.\,\ref{fig:main}C (see Movie S1) were obtained by
+  conducting a full parameter sweep and choosing a combination of values that
+  scored well against three measures.}
 
-%\cmnt{Thus, we suggest that there exist a pair of signalling molecules whose expression varies linearly (with orthogonal gradients) across the thalamus, analgous to the ephrins, whose existence is well known in the cortex.
-Thalamocortical axons growing from a region in the thalamus with high expression of one of these molecules will interact with a corresponding cortical signalling molecule more strongly than an axon emanating from a region with low expression of the thalamic signal.}
+% New section to describe and introduce the metrics
+\metrics{First, we used an algorithm introduced by Honda to measure the
+  discrepancy of each barrel shape from a Dirichletform shape. Low
+  overall values of the \emph{Honda-}$\delta$ metric obtained from simulated
+  barrels indicate a close correspondence of the simulated barrel field with a
+  Voronoi tesselation, and thus with a biological barrel field. For the
+  % Run ./build/sim/honda boundaries/rat_barrels/wb_110405_Dirichlet.svg to
+  % get this result:
+  tesselation overlaid on the real barrel field in Fig.\,1A, $\delta=0.025$,
+  and a reduction in the Honda-$\delta$ metric in simulation
+  confirmed that an equivalent `good' Voronoi pattern} (see
+    \citealp{senft_mouse_1991}) \metrics{emerged within $\approx$~20000
+  iterations (Fig.\,\ref{fig:main}D, red squares). Second, we devised a
+  \emph{pattern difference} measure that is sensitive to deviations in the
+  component shapes and overall topographic registration of two tesselations,
+  $\eta$, and we used this measure to compare the simulated barrel fields to
+  the real barrel field from which the simulation boundary shape was obtained
+  (see Materials and Methods for details). A similar reduction in pattern
+  difference over simulation time confirmed that the shapes and arrangement of
+  emergent connection fields matched those of the real barrel field by around
+  20000 iterations. Third, we measured the \emph{connection selectivity} at
+  each location on the cortical sheet, as the connection density of the most
+  dense projection divided by the sum over all projection densities, $\omega$.}
 
-%\cmnt{We defined two orthogonal, linear gradients oriented in the anterior-posterior and medial-lateral directions and used the coordinates of 41 barreloid centers recreated from \cite{haidarliu_size_2001}, corresponding to the macrovibrissae, to populate 41 pairs of $\gamma$ values.}  (Fig.\,\ref{fig:main}B).
-\cmnt{In other words, we simply scaled the barreloid coordinates in an anterior-posterior/medial-lateral coordinate system to create the $\gamma$ values in the range [-2,2].
-Note that positive \emph{and} negative $\gamma$ values were generated; the origin of the coordinate system was located at the approximate center of the thalamic barreloid field.
-This prevented an overall drift of the cortical pattern away from the centroid of the barrel field and may explain why coordinate-giving transcription factors and the ephrins are often found as pairs acting in opposing directions.}  \cmnt{We concentrated on macrovibrissae (excluding microvibrissae) because it is easy to identify their corresponding structures both in the thalamus and in the cortex.
-We also excluded the 5 rhinal whiskers because their cortical representation is in a separate island, distinct, and somewhat removed from the representation of the macrovibirissae.}
+\metrics{Against these three metrics we are also able to characterise the
+  robustness of self-organization to the model parameters, and to investigate
+  the sensitivity of the model to variation in its inputs.}
 
-\mpfour{The signalling gradients are assumed not to interact with the barrel
-  field boundary and to continue well beyond its extent (see
-  \citealp{miller_epha7-ephrin-a5_2006}, their Fig.\,11) in order to
-  participate in other developmental processes, such as inter-area positional
-  specification \citep{sur_patterning_2005,oleary_area_2007}.} % Just oleary_area_2007 would be ok
+% Most of this commented out region is covered in the para above.
 %
-From random initial conditions for $a(\mb{x},0)$, \cmnt{with each hexagonal element
-  of $a_i$ set from a uniform distribution according to $0.2+{U}(0,0.2)$, and
-  $c(\mb{x},0)=0$}, a clear Voronoi-like tesselation
-emerged (Fig.\,\ref{fig:main}C; see Movie S1). A reduction in the Honda delta
-metric (see \citealp{senft_mouse_1991}) confirmed that a `good' Voronoi
-pattern emerged within \cmnt{$\approx$~20000} iterations
-(Fig.\,\ref{fig:main}D).  A measure of the difference in \cmnt{barrel area}
-between real and simulated barrels revealed a strong correspondence (see
-Fig.\,\ref{fig:main}D, squares). \cmnt{In a chosen region, the ratio of
-  \emph{the connection density of the TC type $i$ which makes the densest
-    connections} to \emph{the sum of all connection densities} gives a measure
-  of the selectivity of the connection density, $\mu(\mb{x},t)$. We summed this measure
-  across the entire barrel domain and named it $\omega(t)$ (see
-  Fig.\,\ref{fig:main}D, grey hexagons). $\omega$ increases during the first
-  20000 simulation steps and converges on a stable
-  value. Fig.\,\ref{fig:main}E shows the selectivity measure plotted as a
-  luminance map, giving a more graded impression of the structure which
-  develops in the system.}
+%A reduction in the Honda delta
+%metric (see \citealp{senft_mouse_1991}) confirmed that a `good' Voronoi
+%pattern emerged within \cmnt{$\approx$~20000} iterations
+%(Fig.\,\ref{fig:main}D).  A measure of the difference in \cmnt{barrel area}
+%between real and simulated barrels revealed a strong correspondence (see
+%Fig.\,\ref{fig:main}D, squares). \cmnt{In a chosen region, the ratio of
+%  \emph{the connection density of the TC type $i$ which makes the densest
+%    connections} to \emph{the sum of all connection densities} gives a measure
+%  of the selectivity of the connection density, $\mu(\mb{x},t)$. We summed this measure
+%  across the entire barrel domain and named it $\omega(t)$ (see
+%  Fig.\,\ref{fig:main}D, grey hexagons). $\omega$ increases during the first
+%  20000 simulation steps and converges on a stable
+%  value. Fig.\,\ref{fig:main}E shows the selectivity measure plotted as a
+%  luminance map, giving a more graded impression of the structure which
+%  develops in the system.}
 
@@
     \includegraphics[width=\linewidth]{./Fig2.png}
-    \caption{\cmnt{\textbf{A} Exploring the parameter space of the model given
+    \caption{\paramsearch{\textbf{A} Exploring the parameter space of the model given
         in Eqs.\,1--3. Graphs show the three pattern metrics for $t=30000$
@@
 
-\cmnt{To determine the robustness of the model to its parameters, we
+\paramsearch{To determine the robustness of the model to its parameters, we
   investigated its behaviour with respect to changes in the parameters $D$,
   $\epsilon$ and the ratio of $\alpha$ to $\beta$. Fig.\,\ref{fig:paramsweep}A
-  shows the value of the metrics $\delta$, $\eta$ and $\omega$ after $30000$
-  iterations plotted as colour maps for $216$ parameter sets. Each row of maps
-  shows the results for a separate metric. Each column of maps shows
-  results for a different value of the ratio $\alpha / \beta$. The diffusion
-  parameter, $D$ varies with the $x$ axes and the strength of competition,
-  $\epsilon$ varies with the $y$ axes. The colour maps are arranged so that
-  `yellow is good'. Lower values of $\delta$ indicate good Voroni patterns and
-  are yellow. Lower values of the difference between real and simulated barrel
-  area indicates good pattern matching. Higher values of $\omega$ indicate
-  greater selectivity and are coloured yellow. White squares indicate that the
-  simulation was not stable for that parameter set.}
+  shows the value of the metrics $\delta$, $\eta$ and $\omega$ after 30000
+  iterations plotted as colour maps for 216 parameter sets. Each row of maps
+  shows the results for a separate metric. Each column of maps shows results
+  for a different value of the ratio $\alpha / \beta$. The diffusion
+  parameter, $D$, varies with the $x$ axes and the strength of competition,
+  $\epsilon$ varies with the $y$ axes. The colour maps are chosen so that
+  yellow indicates desirable properties for the simulations: Lower values of
+  $\delta$ indicate good Voroni patterns and are yellow; Lower values of
+  $\eta$ indicate good pattern matching; Higher values of $\omega$ indicate
+  greater selectivity and are also coloured yellow. White squares indicate
+  that the simulation was not stable for that parameter set.}
 
-\cmnt{The maps indicate that the pattern is not strongly influenced by the
+\paramsearch{The maps indicate that the pattern is not strongly influenced by the
   choice of $\alpha/\beta$, which was varied by 5 orders of magnitude. The
@@
     \caption{
-      \cmnt{\textbf{A} Sensitivity of the system pattern
+      \sensitivity{\textbf{A} Sensitivity of the system pattern
         metrics $\delta(t)$ (top row), $\omega(t)$ (middle row) and $\eta(t)$
@@
 
-\cmnt{We performed a sensitivity analysis to determine if the quality of the
+\sensitivity{We performed a sensitivity analysis to determine if the quality of the
   pattern was affected by perturbations to any of the following inputs:
@@
 
-\cmnt{First, we conducted 10 runs of the simulation shown in
+\sensitivity{First, we conducted 10 runs of the simulation shown in
   Fig.\,\ref{fig:main}C to evaluate the mean and standard deviation of the
@@
 
-\cmnt{Next, we modified the parameters of the generating uniform distribution
+\sensitivity{Next, we modified the parameters of the generating uniform distribution
   to initialize $a_i$ for each hex to $0.1+{U}(0,0.4)$ (increasing the variability of
@@
 
-\cmnt{Fig\,\ref{fig:sens}A shows the results for perturbations ii) to iv). The
+\sensitivity{Fig\,\ref{fig:sens}A shows the results for perturbations ii) to iv). The
   first column shows the results for the three metrics where the interaction
@@
 
-\stu{In reality, scores of smaller Dirichletform barrels representing the microvibrissae form to the right of where the E-row barrels emerge in our simulations, presumably via the same competitive processes.
-Enforcing here the same boundary condition as used to represent the true edges of the barrel field is necessary to ensure the stability of the simulation, though we acknowledge that this region of the boundary is enforced primarily to keep the number of simulated projections, and hence the overall computational complexity of the simulation, manageable (simulating each additional projection introduces an extra \textbf{YYY} new dynamical variables).}
+\stu{In reality, scores of smaller Dirichletform barrels representing the
+  microvibrissae form to the right of where the E-row barrels emerge in our
+  simulations, presumably via the same competitive processes. Enforcing here
+  the same boundary condition as used to represent the true edges of the
+  barrel field is necessary to ensure the stability of the simulation, though
+  we acknowledge that this region of the boundary is enforced primarily to
+  keep the number of simulated projections, and hence the overall
+  computational complexity of the simulation, manageable (simulating each
+  additional projection introduces an extra \textbf{YYY} new dynamical
+  variables).}
 
@@
 boundaries, and overlaid contours show $c>0.5$. The final frame in the movie
-is step $30000$ of the simulation.
+is step 30000 of the simulation.
 

### 2020-06-30 paper/briefpaper/brief.tex
Changes to param search paragraphs, figure and figure caption

@@
     \includegraphics[width=\linewidth]{./Fig2.png}
-    \caption{\paramsearch{\textbf{A} Exploring the parameter space of the model given
-        in Eqs.\,1--3. Graphs show the three pattern metrics for $t=30000$
-        steps as colourmaps: Honda $\delta$ (top row) shows conformance to a
-        Voronoi pattern; $\eta$ (middle row) gives the simulated/experimental
-        barrel area difference and $\omega$ (bottom row) is a measure of
-        selectivity of the connection densities for the 41
-        thalamocortical axon types. For each row, lighter (orange/yellow)
-        colours imply a better model. The parameter space explored is three
-        dimensional with the parameter ratio $\alpha/\beta$ varied along the
-        columns of graphs, the competition parameter $\epsilon$ varied on the
-        $y$-axes and the diffusion constant, $D$ varied along the $x$-axes. An
-        asterisk (*) marks the parameter set displayed in
-        Fig.\,\ref{fig:main}. Boxes i), ii)~and iii)~mark parameter sets for which
-        corresponding patterns are shown in \textbf{B} (for $t=30000$ steps).}}
+    \caption{\paramsearch{\textbf{A} Exploring the parameter space of the
+        model given in Eqs.\,\ref{eq:dc}--\ref{eq:comp}. Graphs show the three
+        pattern metrics for $t=30000$ steps as colourmaps: Honda $\delta$ (top
+        row) shows conformance to a Voronoi pattern; $\eta$ (middle row) gives
+        the simulated/experimental barrel area difference and $\omega$ (bottom
+        row) is a measure of selectivity of the connection densities for the
+        41 thalamocortical axon types. The colour maps are chosen so that for
+        each row, lighter (orange and yellow) colours indicate desirable
+        properties for the simulations: Lower values of $\delta$ are yellow,
+        indicating good Voroni patterns; Lower values of $\eta$ indicate good
+        pattern matching; Higher values of $\omega$ indicate greater
+        selectivity. White squares indicate that the simulation was not stable
+        for that parameter set.
+%
+        The parameter space explored is three dimensional with the parameter
+        ratio $\alpha/\beta$ varied along the columns of graphs, the
+        competition parameter $\epsilon$ varied on the $y$-axes and the
+        diffusion constant, $D$ varied along the $x$-axes. An asterisk (*)
+        marks the parameter set displayed in Fig.\,\ref{fig:main}. Boxes i),
+        ii)~and iii)~mark parameter sets for which corresponding patterns are
+        shown in \textbf{B} (for $t=30000$ steps).
+%
+        \textbf{B} The effect of varying the diffusion constant,
+        $D$. i)~Increasing $D$ above the optimum value gives an expansion of
+        the outer barrels and a corresponding compression of the barrels
+        within. ii)~Conversely, an expansion of the central barrels and
+        compression of those near the boundary results from reducing $D$ from
+        its optimum. iii) Reducing the diffusion constant is equivalent to
+        increasing the size of the domain $S$ and so this simulation, in which
+        $D$ had the lowest stable value, can be thought to represent an animal
+        with a large cortical area but with unchanged cortical complexity. A
+        large proportion of the domain is occupied by the projections with
+        intermediate interaction parameters; those with strong interaction
+        parameters are compressed around the edge of the domain. A barrel
+        pattern fails to form.}}
     \label{fig:paramsweep}
@@
   shows the value of the metrics $\delta$, $\eta$ and $\omega$ after 30000
-  iterations plotted as colour maps for 216 parameter sets. Each row of maps
-  shows the results for a separate metric. Each column of maps shows results
-  for a different value of the ratio $\alpha / \beta$. The diffusion
-  parameter, $D$, varies with the $x$ axes and the strength of competition,
-  $\epsilon$ varies with the $y$ axes. The colour maps are chosen so that
-  yellow indicates desirable properties for the simulations: Lower values of
-  $\delta$ indicate good Voroni patterns and are yellow; Lower values of
-  $\eta$ indicate good pattern matching; Higher values of $\omega$ indicate
-  greater selectivity and are also coloured yellow. White squares indicate
-  that the simulation was not stable for that parameter set.}
+  iterations, plotted as colour maps, for 216 parameter sets. The maps
+  indicate that the pattern is not strongly influenced by the choice of
+  $\alpha/\beta$, which was varied by 5 orders of magnitude. The most
+  Voronoi-like patterns were achieved for parameter sets for which $D$ and
+  $\epsilon$ were high. The barrel area metric, $\eta$ indicates a strong
+  preference for $D\approx0.5$, although this does not give the solution of
+  maximum selectivity, which favours lower values of $D$. Thus, there is a
+  trade-off between good selection and good Voroni pattern conformation.  The
+  parameter set indicated by the asterisk in Fig.\,\ref{fig:paramsweep}A was
+  chosen for Fig.\,\ref{fig:main} because it provides good pattern
+  correspondence and a close approximation to a Voronoi tesselation.}
 
-\paramsearch{The maps indicate that the pattern is not strongly influenced by the
-  choice of $\alpha/\beta$, which was varied by 5 orders of magnitude. The
-  most Voronoi-like patterns were achieved for parameter sets for which $D$
-  and $\epsilon$ were high. The barrel area metric, $\eta$ indicates a strong
-  preference for $D\approx0.5$, although this does not give the
-  solution of maximum selectivity, which for $D$ should be small. Thus, there is a
-  trade-off between good selection and good Voroni pattern
-  conformation. The
-  parameter set indicated by the asterisk in
-  Fig.\,\ref{fig:paramsweep}A is presented in Fig.\,\ref{fig:main} because it
-  provides the best barrel area correspondence and produces a good Voronoi
-  pattern. In Fig.\,\ref{fig:paramsweep}B, we present some patterns
-  representing alternative parameter choices. Fig.\,\ref{fig:paramsweep}B i)
-  shows the simulation result with low diffusion constant, $D$. Reducing the
-  diffusion constant is equivalent to increasing the size of the domain $S$
-  and so this simulation can be thought to represent an animal with a large
-  cortical area but with unchanged cortical complexity. In this simulation, a
-  large proportion of the domain is occupied by the projections with
-  intermediate interaction parameters; those with strong interaction
-  parameters are bunched up towards the edge of the domain. A barrel pattern
-  fails to form. Fig.\,\ref{fig:paramsweep}B ii)~and iii)~indicate the effect
-  of varying $D$ in either direction from its optimum. ii)~is a less
-  exaggerated version of the pattern in i), whereas iii)~shows an expansion of
-  the outer barrels and a corresponding compression of the barrels within.}
+\paramsearch{In Fig.\,\ref{fig:paramsweep}B, we present some patterns
+  representing alternative parameter choices. Because the patterns are robust
+  to alterations in $\epsilon$ and the choices of $\alpha$ and $\beta$, we
+  concentrate our attention on the diffusion constant, $D$. Changing $D$,
+  whose value defines the rate at which a region of localized branching will
+  spread out across the domain, $S$, is equivalent to making changes to the
+  size of $S$ with a fixed branching diffusion rate. These simulations can
+  therefore be thought of as representing animals with a varying cortical
+  areas but with unchanged cortical complexity. The most extreme example
+  (Fig.\,\ref{fig:paramsweep}B iii) could represent an animal with a very
+  large cortical area (such as the South American capybara) which receives
+  sensory input from $N=41$ thalamic barreloids. Here, a large proportion of
+  the domain is occupied by the projections with intermediate interaction
+  parameters and those with strong interaction parameters are bunched up
+  towards the edge of the domain. In this case, a barrel pattern fails to
+  form, agreeing with the observation that cortical whisker barrels are absent
+  in the capybara.}
 
@@
 
+% This section needs to emphasize that the absolute values of gamma are not important
 \sensitivity{We performed a sensitivity analysis to determine if the quality of the
@@
 
-\cmnt{The pattern-matching metric, $\eta$, incorporates information about i)
+\metrics{The pattern-matching metric, $\eta$, incorporates information about i)
   the differences in the areas of the simulated and experimental barrels
@@
 %
-\cmnt{A second way to analyse the adjacency vectors is to consider
+\metrics{A second way to analyse the adjacency vectors is to consider
   the mean magnitude of the difference between the simulated and experimental
@@
 %
-\cmnt{We combined all of these measures into a single pattern quality metric:}
+\metrics{We combined all of these measures into a single pattern quality metric:}
 %
 \begin{equation} \label{eq:eta}
-\color{blue}
+\color{revred}
   \eta =
@@
 %
-\cmnt{which has units of mm$^3$.}
+\metrics{which has units of mm$^3$.}
 

### 2020-06-30 paper/briefpaper/brief.tex
Updates to sensitivity analysis paras

@@
 % Justify competition:
-\newcommand{\justifycomp}[1]{\textcolor{revpurp2}{#1}}
-% Justify gamma/interaction parameter choice
-\newcommand{\justifygammas}[1]{\textcolor{revpink}{#1}}
+\newcommand{\mpfive}[1]{\textcolor{revpurp2}{#1}}
+% Justify gamma/interaction parameter choice. major point 1.
+\newcommand{\MPone}[1]{\textcolor{revpink}{#1}}
 % Describe the metrics
@@
 % Justify competition term. This pulled from Stuart's Google doc
-\justifycomp{We reasoned that an arbitrary number of distinct field locations
+\mpfive{We reasoned that an arbitrary number of distinct field locations
   may be specified using a minimum of two guidance fields, if the tendency for
@@
 
-\justifycomp{Such dynamics are quite unlike those described by classic
+\mpfive{Such dynamics are quite unlike those described by classic
   chemospecificity models} \citep{sperry_chemoaffinity_1963},
-\justifycomp{which essentially assume center-points by specifying conditions
+\mpfive{which essentially assume center-points by specifying conditions
   in the target tissue that instruct pre-identified afferents to stop
@@
 
-\justifycomp{Testing this reasoning required increasing the strength of the
+\mpfive{Testing this reasoning required increasing the strength of the
   competition between simulated thalamocortical projections for cortical
@@
 %
-\justifycomp{This term contributes to the \emph{flux} of axonal branching as
+\mpfive{This term contributes to the \emph{flux} of axonal branching as
   an additional source of diffusion, which reduces the branching density for a
@@
 
-\justifygammas{In addition, the model outlined requires that molecular
+\MPone{In addition, the model outlined requires that molecular
   guidance gradients in the cortex are complemented by graded values of the
   interaction strengths, $\gamma$, at the level of the thalamus. While the
-  precise mechanisms by thalamic and cortical gradients interact during
+  precise mechanisms by which thalamic and cortical gradients interact during
   development have not been fully characterised, the presence of complementary
@@
   cortex} (\citealp{vanderhaeghen_mapping_2000,miller_epha7-ephrin-a5_2006}).
-\justifygammas{Cells originating in the medial part of ventrobasal thalamus
+\MPone{Cells originating in the medial part of ventrobasal thalamus
   (comprising VPM and VPL) express high levels of EphA receptors and project
@@
 
-\justifygammas{For simplicity, both the thalamic interaction gradients and the
+\MPone{For simplicity, both the thalamic interaction gradients and the
   cortical guidance fields in the model were chosen in the first instance to
@@
   defined by the anterior-posterior and medial-lateral axes. Coordinates were
-  estimated from Fig 5d of} \citep{haidarliu_size_2001}, \justifygammas{and
+  estimated from Fig 5d of} \citep{haidarliu_size_2001}, \MPone{and
   scaled (arbitrarily) such that $\gamma\in\pm 2$.}
@@
   extend well beyond the boundary of the barrel field (e.g. as ephrin-5 does;
-  see \citealp{miller_epha7-ephrin-a5_2006}, their Fig.\,11) but they are
-  shown within the barrel fields boundary in Fig.\,\ref{fig:main}A to aid the
-  visualization. Simulations were iterated through 30000 iterations of
-  Eqs.\,\ref{eq:dc}--\ref{eq:comp} ($\delta t=0.0001$)}
+  see \citealp{miller_epha7-ephrin-a5_2006}) but they are shown only within
+  the boundary in Fig.\,\ref{fig:main}A, to indicate that external values are
+  not used by the simulation. Simulations were stepped through 30000
+  iterations of Eqs.\,\ref{eq:dc}--\ref{eq:comp} ($\delta t=0.0001$)}
 
@@
     \citealp{senft_mouse_1991}) \metrics{emerged within $\approx$~20000
-  iterations (Fig.\,\ref{fig:main}D, red squares). Second, we devised a
+  iterations (Fig.\,\ref{fig:main}D, red circles). Second, we devised a
   \emph{pattern difference} measure that is sensitive to deviations in the
@@
   dense projection divided by the sum over all projection densities, $\omega$.}
-
+%
 \metrics{Against these three metrics we are also able to characterise the
@@
         unchanged from those used in the simulation in
-        Fig.\,\ref{fig:main}, for which $\phi_{\rho_1}=6$\textdegree~and
-        $\phi_{\rho_2}=84$\textdegree.
+        Fig.\,\ref{fig:main}, for which
+        $\phi_{\rho_2}=84$\textdegree and $G_{\rho_2}=1$.
         %
@@
 % This section needs to emphasize that the absolute values of gamma are not important
-\sensitivity{We performed a sensitivity analysis to determine if the quality of the
-  pattern was affected by perturbations to any of the following inputs:
+\sensitivity{We performed a sensitivity analysis to determine if the quality
+  of the pattern (i.e.~the values of $\delta$, $\eta$ \& $\omega$ after
+  $t=30000$ steps) was affected by perturbations to any of the following
+  inputs:
 %
@@
   Fig.\,\ref{fig:main}A).
-%
-  In the following analysis, we always evaluated the three pattern metrics,
-  $\delta$ (the Honda Voronoi metric), $\eta$ (the pattern-match metric) and
-  $\omega$ (which measures selectivity of connections, $c_i$) after running
-  the model for 30000 simulation steps.}
+}
 
@@
   Fig.\,\ref{fig:main}C to evaluate the mean and standard deviation of the
-  three metrics in the default model. The
-  results (where deviations occur as a result of the different random initial
-  conditions for $a_i$) were $\delta=0.089\,(0.004)$, $\eta=0.2108\,(0.002)$ and
-  $\omega=0.2165\,(0.0001)$. None of the three metrics were strongly affected
-  by the random changes to $a_i$ on each run. The results give a baseline from
-  which to evaluate whether changes in the other inputs have a statistically
-  significant effect.}
-
-\sensitivity{Next, we modified the parameters of the generating uniform distribution
-  to initialize $a_i$ for each hex to $0.1+{U}(0,0.4)$ (increasing the variability of
-  $a_i(\mb{x},t=0)$ without changing the mean initial activation). We ran this
-  system 10 times and obtained the results $\delta=0.090\,(0.005)$,
-  $\eta=0.2136\,(0.004)$ and $\omega=0.2166\,(0.0003)$. Paired two-sample t-tests
-  demonstrated that the increase in noise amplitude made no significant
-  difference to the metric results ($\delta:
-  \mathrm{t}=-0.69,\mathrm{p}=0.49$,
-  $\eta: \mathrm{t}=-2.0,\mathrm{p}=0.06$,
-  $\omega: \mathrm{t}=-1.7,\mathrm{p}=0.11$). 10 runs of a
-  noise-free initial condition with $a_i=0.3$ for every hex showed that
-  the model is deterministic (all runs produced the results $\delta=0.0928$,
-  $\eta=0.21236$ and $\omega=0.21649$) and that stochasticity in
-  $a_i(\mb{x},t=0)$ is not required by this model, as all three results lay within
-  one standard deviation of those for the runs in which $a_i$ was initialized stochastically.}
+  metrics in the default model. The results (where deviations occur as a
+  result of the different random initial conditions for $a_i$) were
+  $\delta=0.089\,(0.004)$, $\eta=0.2108\,(0.002)$ and
+  $\omega=0.2165\,(0.0001)$. The pattern was barely affected by the random
+  changes to $a_i(t=0)$ on each run.}
+%
+\sensitivity{Next, we modified the parameters of the generating uniform
+  distribution to initialize $a_i$ for each hex to $0.1+{U}(0,0.4)$,
+  increasing the variability of $a_i(\mb{x},t=0)$ without changing the mean
+  initial activation. We ran this system 10 times and obtained the results
+  $\delta=0.090\,(0.005)$, $\eta=0.2136\,(0.004)$ and
+  $\omega=0.2166\,(0.0003)$. Paired two-sample t-tests demonstrated that the
+  increase in noise amplitude made no significant difference to the metric
+  results ($\delta: \mathrm{t}=-0.69,\mathrm{p}=0.49$, $\eta:
+  \mathrm{t}=-2.0,\mathrm{p}=0.06$, $\omega:
+  \mathrm{t}=-1.7,\mathrm{p}=0.11$). 10 runs of a noise-free initial condition
+  with $a_i=0.3$ for every hex showed that the model is deterministic (all
+  runs produced the results $\delta=0.0928$, $\eta=0.21236$ and
+  $\omega=0.21649$) and that stochasticity in $a_i(\mb{x},t=0)$ is not
+  required by this model, as all three results lay within one standard
+  deviation of those for the runs in which $a_i$ was initialized
+  stochastically.}
 
@@
 
-\sensitivity{Fig\,\ref{fig:sens}A shows the results for perturbations ii) to iv). The
-  first column shows the results for the three metrics where the interaction
-  parameters $\gamma_{i,j}$ are replaced by $\gamma_{i,j}'$ according to
-  $\gamma_{i,j}' = \gamma_{i,j} + U(0,4\nu_\gamma)$, where $\nu_\gamma$ is the
-  chosen magnitude of the noise (the proportion of the full range of
-  $\gamma_{i,j}$, which is 4). The Voronoi nature of the pattern ($\delta$)
-  and the selectivity of the fields are not strongly affected by interaction
-  parameter noise, but the pattern metric, $\eta$ shows a... ADD SUB GRAPH.}
-
-% This figure is text width
-\begin{center}
-  \includegraphics[width=\linewidth]{./Fig4.png}
-  \captionof{figure}{Guidance
-    fields and emergent barrel pattern in a Fgf8 misexpression experiment
-    (c.f. \citealp{assimacopoulos_fibroblast_2012}), simulated by reflecting
-    $\rho_1$ from Fig.\,\ref{fig:main}A at the join. \cmnt{The model
-      parameters match those in Fig.\,\ref{fig:main}C. Scale
-      bars are 1\,mm.}}
-  \label{fig:fgf8}
-\end{center}
+\sensitivity{Fig\,\ref{fig:sens}A shows the results for perturbations ii) to
+  iv). The results show that the Voronoi nature of the pattern ($\delta$) and
+  the selectivity of the fields are not strongly affected by interaction
+  parameter noise, but the pattern metric, $\eta$ shows a degradation in the
+  pattern arrangement when a) noise in the interaction parameters causes their
+  disordering (find better way to express?), b) noise in the guidance field
+  with a high magnitude and short length scale causes disruption to the barrel
+  edges and c) the guidance fields tend towards a non-orthogonal arrangement,
+  or if one field becomes substantially stronger in its interaction with the
+  thalamocortical projections. Fig\,\ref{fig:sens}B shows perturbed examples.}
 
@@
   fields.}
+
+% This figure is text width
+\begin{center}
+  \includegraphics[width=\linewidth]{./Fig4.png}
+  \captionof{figure}{Guidance
+    fields and emergent barrel pattern in a Fgf8 misexpression experiment
+    (c.f. \citealp{assimacopoulos_fibroblast_2012}), simulated by reflecting
+    $\rho_1$ from Fig.\,\ref{fig:main}A at the join. \cmnt{The model
+      parameters match those in Fig.\,\ref{fig:main}C. Scale
+      bars are 1\,mm.}}
+  \label{fig:fgf8}
+\end{center}
+
+Add paragraph on whisker trimming here.
 

### 2020-06-30 paper/briefpaper/brief.tex
Colour wrangling, label changing

@@
 
+% Miscellaneous changes
+\newcommand{\cmnt}[1]{\textcolor{blue}{#1}}
+% Justify gamma/interaction parameter choice. major point 1.
+\newcommand{\MPone}[1]{\textcolor{revpink}{#1}}
 % Blue changes are major point 2 (competition mechanism) and other, general
 % uncategorised changes.
-\newcommand{\cmnt}[1]{\textcolor{blue}{#1}}
-% Red for Stuart's first updates
-\newcommand{\stu}[1]{\textcolor{red}{#1}}
+\newcommand{\MPtwo}[1]{\textcolor{revblue}{#1}}
+% Parameter search. Major point 3
+\newcommand{\MPthreePar}[1]{\textcolor{revbblack}{#1}}
+% Sensitivity analysis. Major point 3
+\newcommand{\MPthreeSens}[1]{\textcolor{revgblack}{#1}}
+% Justify competition:
+\newcommand{\mpfive}[1]{\textcolor{revpurp2}{#1}}
 % Minor point 4:
@@
 \newcommand{\mpsix}[1]{\textcolor{revpurp}{#1}}
-% Justify competition:
-\newcommand{\mpfive}[1]{\textcolor{revpurp2}{#1}}
-% Justify gamma/interaction parameter choice. major point 1.
-\newcommand{\MPone}[1]{\textcolor{revpink}{#1}}
 % Describe the metrics
 \newcommand{\metrics}[1]{\textcolor{revred}{#1}}
-% Parameter search. Major point 3
-\newcommand{\paramsearch}[1]{\textcolor{revbblack}{#1}}
-% Sensitivity analysis. Major point 3
-\newcommand{\sensitivity}[1]{\textcolor{revgblack}{#1}}
 
@@
 \begin{equation} \label{eq:da}
-\frac{\partial a_i}{\partial t} = \nabla\vcdot\left(D \nabla a_i-a_i\sum_{j=1}^{M} \gamma_{i,j}\nabla \rho_j(\mb{x}) \cmnt{+ \chi_i}\right) - \frac{\partial c_i}{\partial t}.
+\frac{\partial a_i}{\partial t} = \nabla\vcdot\left(D \nabla a_i-a_i\sum_{j=1}^{M} \gamma_{i,j}\nabla \rho_j(\mb{x}) \MPtwo{+ \chi_i}\right) - \frac{\partial c_i}{\partial t}.
 \end{equation}
@@
       the locations of putative barreloid centers (reconstructed from
-      \citealp{haidarliu_size_2001}, \cmnt{Fig.\,5b}). \textbf{C} Simulation results for parameters
-      \cmnt{$N=41$, $\alpha=3.6$, $\beta=16.67$, $k=3$, $D=0.5$, $\gamma\in\pm 2$,
+      \citealp{haidarliu_size_2001}, \MPone{Fig.\,5b}). \textbf{C} Simulation results for parameters
+      \MPtwo{$N=41$, $\alpha=3.6$, $\beta=16.67$, $k=3$, $D=0.5$, $\gamma\in\pm 2$,
       $\epsilon=1.2$} and $\delta{t}=0.0001$. Colours indicate the thalamic
@@
       delineate boundaries (see Movie
-      S1). \textbf{D} Red dots show the Honda metric, \cmnt{$\delta(t)$}, obtained from \cmnt{the}
-      simulation approaching that obtained from \cmnt{real} barrels in \textbf{A} (dotted
-      line); black squares \cmnt{provide a} measure \cmnt{of} the correspondence
+      S1). \textbf{D} Red dots show the Honda metric, \metrics{$\delta(t)$}, obtained from \metrics{the}
+      simulation approaching that obtained from \metrics{real} barrels in \textbf{A} (dotted
+      line); black squares \metrics{provide a} measure \metrics{of} the correspondence
       between the real and
-      simulated barrel shapes, $\eta(t)$ (the \cmnt{pattern quality metric---see
+      simulated barrel shapes, $\eta(t)$ (the \metrics{pattern quality metric---see
         Materials \& Methods; units mm$^3$});
-      \cmnt{grey hexagons \cmnt{provide a measure of how selectively each
+      \metrics{grey hexagons \metrics{provide a measure of how selectively each
           cortical site is innervated};
@@
         $\mu(\mb{x}) \equiv \mathrm{max}_i\big(c_i(\mb{x},t)\big)\big/\sum_{j=1}^{N} c_j(\mb{x},t)$.}
-      \textbf{E} \cmnt{Plotted across the cortical sheet, the selectivity
+      \textbf{E} \metrics{Plotted across the cortical sheet, the selectivity
         develops to reveal an alignment with the emergent barrel boundary
-        shapes}. \cmnt{The colour map
+        shapes}. \metrics{The colour map
         indicates the values of $\mu(\mb{x})$. Overlaid
@@
 connections. To introduce competition also in terms of axon branching,
-\cmnt{whilst ensuring that $a_i$ is conserved over time,} we
+\MPtwo{whilst ensuring that $a_i$ is conserved over time,} we
 redefined
@@
 \begin{equation} \label{eq:comp}
-  \color{blue}
+  \color{revblue}
   \chi_i(\mb{x}, t) = \frac{\epsilon a_i}{N-1} \nabla \textstyle\sum_{j\ne i}^{N}{a_j}.
@@
 parameter values,
-%From random initial conditions for $a(\mb{x},0)$, \cmnt{with each hexagonal element
+%From random initial conditions for $a(\mb{x},0)$, \MPtwo{with each hexagonal element
 %  of $a_i$ set from a uniform distribution according to $0.2+{U}(0,0.2)$, and
@@
     \includegraphics[width=\linewidth]{./Fig2.png}
-    \caption{\paramsearch{\textbf{A} Exploring the parameter space of the
+    \caption{\MPthreePar{\textbf{A} Exploring the parameter space of the
         model given in Eqs.\,\ref{eq:dc}--\ref{eq:comp}. Graphs show the three
@@
 
-\paramsearch{To determine the robustness of the model to its parameters, we
+\MPthreePar{To determine the robustness of the model to its parameters, we
   investigated its behaviour with respect to changes in the parameters $D$,
@@
 
-\paramsearch{In Fig.\,\ref{fig:paramsweep}B, we present some patterns
+\MPthreePar8u{In Fig.\,\ref{fig:paramsweep}B, we present some patterns
   representing alternative parameter choices. Because the patterns are robust
@@
     \caption{
-      \sensitivity{\textbf{A} Sensitivity of the system pattern
+      \MPthreeSens{\textbf{A} Sensitivity of the system pattern
         metrics $\delta(t)$ (top row), $\omega(t)$ (middle row) and $\eta(t)$
@@
 % This section needs to emphasize that the absolute values of gamma are not important
-\sensitivity{We performed a sensitivity analysis to determine if the quality
+\MPthreeSens{We performed a sensitivity analysis to determine if the quality
   of the pattern (i.e.~the values of $\delta$, $\eta$ \& $\omega$ after
@@
 
-\sensitivity{First, we conducted 10 runs of the simulation shown in
+\MPthreeSens{First, we conducted 10 runs of the simulation shown in
   Fig.\,\ref{fig:main}C to evaluate the mean and standard deviation of the
@@
 %
-\sensitivity{Next, we modified the parameters of the generating uniform
+\MPthreeSens{Next, we modified the parameters of the generating uniform
   distribution to initialize $a_i$ for each hex to $0.1+{U}(0,0.4)$,
@@
 
-\sensitivity{Fig\,\ref{fig:sens}A shows the results for perturbations ii) to
+\MPthreeSens{Fig\,\ref{fig:sens}A shows the results for perturbations ii) to
   iv). The results show that the Voronoi nature of the pattern ($\delta$) and
@@
   \includegraphics[width=\linewidth]{./Fig5.png}
-  \captionof{figure}{\cmnt{Simulating whisker trimming by reduction of
+  \captionof{figure}{\MPtwo{Simulating whisker trimming by reduction of
       $\epsilon$ for the thalamocortical projection for barrel C3, for which
@@
 
-\stu{In reality, scores of smaller Dirichletform barrels representing the
+\cmnt{In reality, scores of smaller Dirichletform barrels representing the
   microvibrissae form to the right of where the E-row barrels emerge in our
@@
   \nabla\vcdot\mb{J}_i(\mb{x},t) = \nabla\vcdot\left(D \nabla
-  a_i-a_i\sum_{j=1}^{M} \gamma_{i,j}\nabla \rho_j(\mb{x}) \color{blue}+
+  a_i-a_i\sum_{j=1}^{M} \gamma_{i,j}\nabla \rho_j(\mb{x}) \color{revblue}+
   \frac{\epsilon a_i}{N-1} \nabla \hat{a}_i \color{black}\right),
@@
 %
-\cmnt{where $\hat{a}_i\equiv\sum_{j\ne i}^{N}a_j$.} Note that the sum of the
+\MPtwo{where $\hat{a}_i\equiv\sum_{j\ne i}^{N}a_j$.} Note that the sum of the
 guidance gradients is time-independent and define $\mb{g}_i(\mb{x}) \equiv
@@
 \nabla\vcdot\mb{J}_i = \nabla\vcdot\big(D \nabla a_i\big) -
-\nabla\vcdot\big(a_i \mb{g}_i\big) \color{blue} + \frac{\epsilon}{N-1}\nabla\vcdot\big({a}_i\nabla\hat{a}_i\big)\color{black}.
+\nabla\vcdot\big(a_i \mb{g}_i\big) \color{revblue} + \frac{\epsilon}{N-1}\nabla\vcdot\big({a}_i\nabla\hat{a}_i\big)\color{black}.
 \end{equation}
@@
 \mb{g}_i\vcdot\nabla a_i % term3
-\color{blue}
+\color{revblue}
 + \frac{\epsilon a_i}{N-1} \dvrg \hat{a}_i % term1_1
@@
 %
-which has \cmnt{five} elements to compute: i)~$D \dvrg a_i$ (the Laplacian of
+which has \MPtwo{five} elements to compute: i)~$D \dvrg a_i$ (the Laplacian of
 $a_i$); ii)~a time-independent modulator of $a_i$ (because
@@
 product of the static vector field $\mb{g}_i$ and the gradient of $a_i$;
-\cmnt{iv)~the Laplacian of $\hat{a}_i$; and v)~a term involving the gradients
+\MPtwo{iv)~the Laplacian of $\hat{a}_i$; and v)~a term involving the gradients
   of $a_i$ and $\hat{a}_i$}. Each of the divergences can be simplified by
@@
 \noindent
-iii) The \cmnt{middle} term in Eq.\,\ref{eq:divJExpanded} is the scalar product of two
+iii) The \MPtwo{middle} term in Eq.\,\ref{eq:divJExpanded} is the scalar product of two
 vector fields which is straightforward to compute from their Cartesian
@@
 \noindent
-\cmnt{iv) The same method used to compute $\dvrg a_i$ in term (i) is used to compute
+\MPtwo{iv) The same method used to compute $\dvrg a_i$ in term (i) is used to compute
 $\dvrg \hat{a}_i$.}
@@
 \noindent
-\cmnt{v) The final term is the scalar product of the two vector fields $\nabla
+\MPtwo{v) The final term is the scalar product of the two vector fields $\nabla
   a_i$ and $\nabla \hat{a}_i$.}
@@
 ensuring that the second and third terms of Eq.\,\ref{eq:divJExpanded} also
-contribute nothing to $\mb{J}$. This is achieved by \cmnt{multiplying}
-$\mb{g}_i(\mb{x})$ \cmnt{by} a sharp logistic function of the distance, $d_b$,
-from $\mb{x}$ to the boundary, \cmnt{of the form $1/[1 +
+contribute nothing to $\mb{J}$. This is achieved by \MPtwo{multiplying}
+$\mb{g}_i(\mb{x})$ \MPtwo{by} a sharp logistic function of the distance, $d_b$,
+from $\mb{x}$ to the boundary, \MPtwo{of the form $1/[1 +
   \exp(100(d_f-d_b))]$, where $d_f=0.1$\,mm is the boundary fall-off distance,
@@
 described in (i)--(v) may be found in the class method
-\code{RD\_James\cmnt{\_comp2}::compute\_divJ()} which calculates \code{term1},
+\code{RD\_James\MPtwo{\_comp2}::compute\_divJ()} which calculates \code{term1},
 \code{term2}, \code{term3}, \code{term1\_1} and \code{term1\_2}, respectively.
@@
 
-%\cmnt{In other words, we simply scaled the barreloid coordinates in an anterior-posterior/medial-lateral coordinate system to create the $\gamma$ values in the range [-2,2].
+%\MPtwo{In other words, we simply scaled the barreloid coordinates in an anterior-posterior/medial-lateral coordinate system to create the $\gamma$ values in the range [-2,2].
 Note that positive \emph{and} negative $\gamma$ values were generated; the origin of the coordinate system was located at the approximate center of the thalamic barreloid field.
-This prevented an overall drift of the cortical pattern away from the centroid of the barrel field and may explain why coordinate-giving transcription factors and the ephrins are often found as pairs acting in opposing directions.}  \cmnt{We concentrated on macrovibrissae (excluding microvibrissae) because it is easy to identify their corresponding structures both in the thalamus and in the cortex.
+This prevented an overall drift of the cortical pattern away from the centroid of the barrel field and may explain why coordinate-giving transcription factors and the ephrins are often found as pairs acting in opposing directions.}  \MPtwo{We concentrated on macrovibrissae (excluding microvibrissae) because it is easy to identify their corresponding structures both in the thalamus and in the cortex.
 We also excluded the 5 rhinal whiskers because their cortical representation is in a separate island, distinct, and somewhat removed from the representation of the macrovibirissae.}}
 
-%\cmnt{Thus, we suggest that there exist a pair of signalling molecules whose expression varies linearly (with orthogonal gradients) across the thalamus, analgous to the ephrins, whose existence is well known in the cortex.
+%\MPtwo{Thus, we suggest that there exist a pair of signalling molecules whose expression varies linearly (with orthogonal gradients) across the thalamus, analgous to the ephrins, whose existence is well known in the cortex.
 Thalamocortical axons growing from a region in the thalamus with high expression of one of these molecules will interact with a corresponding cortical signalling molecule more strongly than an axon emanating from a region with low expression of the thalamic signal.}
 
-%\cmnt{We defined two orthogonal, linear gradients oriented in the anterior-posterior and medial-lateral directions and used the coordinates of 41 barreloid centers recreated from \cite{haidarliu_size_2001}, corresponding to the macrovibrissae, to populate 41 pairs of $\gamma$ values.}  (Fig.\,\ref{fig:main}B).
-\cmnt{In other words, we simply scaled the barreloid coordinates in an anterior-posterior/medial-lateral coordinate system to create the $\gamma$ values in the range [-2,2].
+%\MPtwo{We defined two orthogonal, linear gradients oriented in the anterior-posterior and medial-lateral directions and used the coordinates of 41 barreloid centers recreated from \cite{haidarliu_size_2001}, corresponding to the macrovibrissae, to populate 41 pairs of $\gamma$ values.}  (Fig.\,\ref{fig:main}B).
+\MPtwo{In other words, we simply scaled the barreloid coordinates in an anterior-posterior/medial-lateral coordinate system to create the $\gamma$ values in the range [-2,2].
 Note that positive \emph{and} negative $\gamma$ values were generated; the origin of the coordinate system was located at the approximate center of the thalamic barreloid field.
-This prevented an overall drift of the cortical pattern away from the centroid of the barrel field and may explain why coordinate-giving transcription factors and the ephrins are often found as pairs acting in opposing directions.}  \cmnt{We concentrated on macrovibrissae (excluding microvibrissae) because it is easy to identify their corresponding structures both in the thalamus and in the cortex.
+This prevented an overall drift of the cortical pattern away from the centroid of the barrel field and may explain why coordinate-giving transcription factors and the ephrins are often found as pairs acting in opposing directions.}  \MPtwo{We concentrated on macrovibrissae (excluding microvibrissae) because it is easy to identify their corresponding structures both in the thalamus and in the cortex.
 We also excluded the 5 rhinal whiskers because their cortical representation is in a separate island, distinct, and somewhat removed from the representation of the macrovibirissae.}

### 2020-06-30 paper/briefpaper/brief.tex
Less terse section on Fgf8 misexpression

@@
 \newcommand{\MPthreeSens}[1]{\textcolor{revgblack}{#1}}
+% More predictions. Major point 3
+\newcommand{\MPthreePred}[1]{\textcolor{revorange}{#1}}
 % Justify competition:
@@
 
-To further investigate the interplay of genetic and intrinsic mechanisms we
-simulated a seminal barrel duplication experiment
-\citep{shimogori_fibroblast_2005}, first by solving on a cortical domain
-comprising two separate, identically shaped cortical boundaries. In this case
-\cmnt{NOTE: The following only holds for the divisive normalization system -
-  in the comp2 system, there's no `coupling' at all between the two regions,
-  so the mirrored region arranges itself as expected with the one reversed
-  guidance field.} one distinct cortical field per thalamic projection emerged, within one of the
-two boundaries only, i.e., no duplication. However, merging the two fields to
-create an extended mirror-symmetric boundary shape, with an anterior-posterior
-guidance field reflected at the join, gave rise to two mirror-symmetrical
-barrel fields comprising 2N barrels, i.e., two cortical fields for each
-thalamic projection (Fig.\,\ref{fig:fgf8}). Together these results suggest that by the
-misexpression of Fgf8, \cite{shimogori_fibroblast_2005}
-effectively created one large barrel field rather than two distinct cortical
-fields. \cmnt{Add something about the slightly off-vertical arrangement of the
-  fields.}
+% This should be a little less terse than in the original.
+\MPthreePred{To further investigate the interplay of genetic and intrinsic
+  mechanisms we simulated two well known experimental manipulations of the
+  whisker barrel system. First, we simulated a seminal barrel duplication
+  experiments}
+\citep{shimogori_fibroblast_2005,assimacopoulos_fibroblast_2012}
+\MPthreePred{in which the growth factor Fgf8, which is normally expressed at
+  the anterior end of the cortical sub-plate from around E9.5}
+\citep{crossley_mouse_1995}, \MPthreePred{is ectopically expressed (by
+  electroporation) also at the posterior end. Because Fgf8 sets off a chain of
+  gene interactions which are involved in the control of the expression of
+  ephrin and cadherin molecules in gradients across the cortical plate, its
+  misexpression at the posterior pole of a cortical hemisphere could be
+  expected to mirror those gradients which usually vary along the
+  rostro-caudal axis. We assume that this results in a mirror of the primary
+  barrel cortex boundary along the rostro-caudal axis, to match the
+  experimental results of}
+\citealp{assimacopoulos_fibroblast_2012}. \MPthreePred{We also assume that a
+  further consequence of the posterior Fgf8 electroporation is the mirrored
+  anterior-posterior guidance gradient $\rho_1$ (Fig.\,\ref{fig:fgf8}A). We
+  mirrored this across a reflection line which is orthogonal to its angle
+  $\phi_{\rho_1}=6$\textdegree, and left $\rho_2$ unmodified. Although the
+  rostro-caudal boundary reflection is not perfectly aligned with $\rho_1$,
+  and we might expect some modification of both $\rho_1$ \emph{and} $\rho_2$,
+  mirroring only $\rho_1$ retains the simplicity of the guidance fields; they
+  remain either linear ($\rho_2$), or piece-wise linear ($\rho_1$).}
+
+\MPthreePred{The new boundary shape and $\rho_1$ guidance gradient were the
+  only changes made to the simulation of Fig.\,\ref{fig:main}; the simulation
+  was run with $N=41$ thalamocortical axon types as before and the initial
+  values for $a_i$ were randomly initialised according to the same
+  distribution. The result, after 30000 simulation steps, was two
+  mirror-symmetrical barrel fields comprising $2N$ barrels
+  (Fig\,\ref{fig:main}B). This result suggests that by the misexpression of
+  Fgf8,} \cite{shimogori_fibroblast_2005} and \cite{assimacopoulos_fibroblast_2012}
+  \MPthreePred{created the correct cortical
+  conditions for the self-organization of the two mirrored barrel fields. }
+
+%However, merging the two fields to
+%create an extended mirror-symmetric boundary shape, with an anterior-posterior
+%guidance field reflected at the join, gave rise to two mirror-symmetrical
+%barrel fields comprising 2N barrels, i.e., two cortical fields for each
+%thalamic projection (Fig.\,\ref{fig:fgf8}). Together these results suggest that by the
+%misexpression of Fgf8, \cite{shimogori_fibroblast_2005}
+%effectively created one large barrel field rather than two distinct cortical
+%fields.
 
@@
 
-Add paragraph on whisker trimming here.
+\MPthreePred{Add paragraph on whisker trimming here.}
 

### 2020-07-01 paper/briefpaper/brief.tex
Figure improvements, biggest ones for Fig 5

@@
 \metrics{First, we used an algorithm introduced by Honda to measure the
-  discrepancy of each barrel shape from a Dirichletform shape. Low
-  overall values of the \emph{Honda-}$\delta$ metric obtained from simulated
-  barrels indicate a close correspondence of the simulated barrel field with a
-  Voronoi tesselation, and thus with a biological barrel field. For the
+  discrepancy of each barrel shape from a Dirichletform shape. Low overall
+  values of the dimensionless \emph{Honda-}$\delta$ metric (called $\Delta$
+  in}
+\citealp{honda_geometrical_1983}) \metrics{obtained from simulated barrels indicate a close
+  correspondence of the simulated barrel field with a Voronoi tesselation, and
+  thus with a biological barrel field. For the
   % Run ./build/sim/honda boundaries/rat_barrels/wb_110405_Dirichlet.svg to
   % get this result:
-  tesselation overlaid on the real barrel field in Fig.\,1A, $\delta=0.025$,
-  and a reduction in the Honda-$\delta$ metric in simulation
+  tesselation overlaid on the real barrel field in Fig.\,\ref{fig:main}A,
+  $\delta=0.025$, and a reduction in the Honda-$\delta$ metric in simulation
   confirmed that an equivalent `good' Voronoi pattern} (see
-    \citealp{senft_mouse_1991}) \metrics{emerged within $\approx$~20000
-  iterations (Fig.\,\ref{fig:main}D, red circles). Second, we devised a
-  \emph{pattern difference} measure that is sensitive to deviations in the
-  component shapes and overall topographic registration of two tesselations,
-  $\eta$, and we used this measure to compare the simulated barrel fields to
-  the real barrel field from which the simulation boundary shape was obtained
-  (see Materials and Methods for details). A similar reduction in pattern
+\citealp{senft_mouse_1991}) \metrics{emerged within $\approx$~20000 iterations
+  (Fig.\,\ref{fig:main}D, red circles). Second, we devised a \emph{pattern
+    difference} measure that is sensitive to deviations in the component
+  shapes and overall topographic registration of two tesselations, $\eta$, and
+  we used this measure to compare the simulated barrel fields to the real
+  barrel field from which the simulation boundary shape was obtained (see
+  Materials and Methods for details). A similar reduction in pattern
   difference over simulation time confirmed that the shapes and arrangement of
@@
   each location on the cortical sheet, as the connection density of the most
-  dense projection divided by the sum over all projection densities, $\omega$.}
+  dense projection divided by the sum over all projection densities,
+  $\omega$.}
 %
@@
 
-\MPthreePar8u{In Fig.\,\ref{fig:paramsweep}B, we present some patterns
+\MPthreePar{In Fig.\,\ref{fig:paramsweep}B, we present some patterns
   representing alternative parameter choices. Because the patterns are robust
@@
   \includegraphics[width=\linewidth]{./Fig5.png}
-  \captionof{figure}{\MPtwo{Simulating whisker trimming by reduction of
+  \captionof{figure}{\MPthreePred{Simulating whisker trimming by reduction of
       $\epsilon$ for the thalamocortical projection for barrel C3, for which

### 2020-07-01 analysis/plot_trim.py

@@ -17,6 +17,7 @@ import h5py
 import domcentres as dc
 import sebcolour
 col = sebcolour.Colour
+import matplotlib.transforms
 
 # Set plotting defaults
 fs = 26
@@ -29,11 +30,14 @@ matplotlib.rc('font', **fnt)
 plt.rcParams['svg.fonttype'] = 'none'
 
 # And plot this longhand here:
-F1 = plt.figure (figsize=(10,12))
+F1 = plt.figure (figsize=(10,8.5))
 
 ## load data from csv
 trimdata = np.genfromtxt ('postproc/whisker_trim_individual.csv', delimiter=",", names=True)
 td = np.sort (trimdata, order='barrel')
+
+alldata = np.genfromtxt ('postproc/whisker_trim_overall.csv', delimiter=",", names=True)
+ad = np.sort (alldata, order='eps_mult')
 
 # Show an example map for this multiplier
 map_mult = 0.64
@@ -48,16 +52,17 @@ tdc4 = td[td[:]['barrel_index'] == 14]
 tdd2 = td[td[:]['barrel_index'] == 22]
 tdd3 = td[td[:]['barrel_index'] == 23]
 
-ax1 = F1.add_subplot(3,1,1)
-ax2 = F1.add_subplot(3,1,3)
-ax3 = F1.add_subplot(3,1,2)
+ax1 = F1.add_subplot(2,1,1)
+ax2 = F1.add_subplot(2,1,2)
 
 xmax = max(tdc3[:]['eps_mult'])
 xmin = min(tdc3[:]['eps_mult'])
 
 others_size = 6
 main_size = 12
+m_width = 3
 
+col_c3 = col.blue2
 col_b3 = col.blue3
 col_c2 = col.navy
 col_c4 = col.cobalt
@@ -65,51 +70,62 @@ col_d2 = col.royalblue2
 col_d3 = col.cornflowerblue
 
 
-l1_map = ax1.plot([map_mult,map_mult], [0, 0.26], '-.', color=col.black)
-l1_1, = ax1.plot(tdb3[:]['eps_mult'], tdb3[:]['area'], 'o', markersize=others_size, color=col_b3, label='B3')
-l1_1, = ax1.plot(tdb3[:]['eps_mult'], tdc2[:]['area'], 'o', markersize=others_size, color=col_c2, label='C2')
[…]

### 2020-07-01 analysis/process_trim.py

@@ -94,6 +94,6 @@ with open('postproc/whisker_trim_overall.csv', 'w', newline='\n') as csvfile:
         cw.writerow (tableline)
 with open('postproc/whisker_trim_individual.csv', 'w', newline='\n') as csvfile:
     cw = csv.writer (csvfile, delimiter=',')
-    cw.writerow (['eps_mult','epsilon','eps_modified','barrel','barrel_index','hondadelta','area_diff','area'])
+    cw.writerow (['eps_mult','epsilon','eps_modified','barrel','barrel_index','hondadeltaj','area_diff','area'])
     for tableline in table_ind:
         cw.writerow (tableline)

### 2020-07-01 paper/briefpaper/brief.tex
Fig 5 caption

@@
   \captionof{figure}{\MPthreePred{Simulating whisker trimming by reduction of
-      $\epsilon$ for the thalamocortical projection for barrel C3, for which
-      $i=13$. All other model parameters match those in
-      Fig.\,\ref{fig:main}C. \textbf{A} Graph of individual barrel measures
-      plotted vs.~the multiplier applied to $\epsilon_{13}$. \textbf{B} The
-      pattern for $\epsilon_{m}=0.64$.}}
+      $\epsilon_i$ for the thalamocortical projection for barrel
+      C3. $\epsilon_{C3} = m \epsilon$, where otherwise, all model parameters
+      match those in Fig.\,\ref{fig:main}C (thus $\epsilon=1.2$). \textbf{A}
+      The pattern formed after 30000 steps for $m=0.64$. Barrel C3 is reduced
+      in size as its axon branching competes less strongly with its
+      neighbours, whose area is slightly increased. \textbf{B} Top: Individual
+      barrel area, $A_i$, plotted vs.~$m$ shows that as $m$ is reduced, the
+      area of C3 (blue squares) goes down, disappearing for $m\approx0.5$,
+      whereas the areas of neighbouring barrels (triangles) increases. Bottom:
+      Honda measures of the Dirichliform nature of the pattern. Red circles
+      show the overall Honda-$\delta$ measure, with the red dotted line
+      showing the same `good' pattern threshold shown in
+      Fig.\,\ref{fig:main}D. Blue squares and triangles show the
+      \emph{individual} Honda metric ($\delta_i$ here; $\Delta_j$ in}
+    \citealp{honda_geometrical_1983}) \MPthreePred{divided by the barrel area,
+      making it comparable with the dimensionless, overall Honda-$\delta$
+      metric. The individual Honda measures do not vary systematically with
+      $m$.}}
   \label{fig:whiskertrim}

### 2020-07-02 paper/briefpaper/brief.tex
Updated caption for whisker trim and a paragraph about the experiment

--- 2020-07-01 paper/briefpaper/brief.tex
@@
 
-\MPthreePred{Add paragraph on whisker trimming here.}
+\MPthreePred{To investigate the response of the model to environmental
+  manipulation, we simulated a whisker deprivation experiment. In the first
+  few days after birth, the upper region of the rat's cortex develops into
+  distinct layers I--IV (layers V and VI are already distinct at
+  birth). During this critical period, removal of the mystacial whiskers by
+  electrocauterization, plucking or trimming leads to observable changes in
+  brain structures, including the barrel field in cortical layer
+  IV~\citep{jeanmonod_mouse_1981}. Amoung other changes, barrels associated
+  with deprived whiskers have a smaller area. This is found whether the animal
+  is deprived of an individual whisker~\citep{kossut_effects_1992}, or an
+  entire row~\citep{loos_somatosensory_1973}, and demonstrates that
+  thalamocortical activity contributes to the developmental process. We
+  simulated the trimming of the individual whisker C3 during this critical
+  period by reducing the amount by which its thalamocortical projections
+  compete for branching. This was achieved by reducing the value of $\epsilon$
+  in Eq.\,\ref{eq:comp} for the C3 projection, leaving other simulation
+  parameters unchanged. The reduced competition for branching limited the area
+  over which the C3 projection could make connections and its resulting barrel
+  size was diminished (Fig.\,\ref{fig:whiskertrim}A). Its neighbours, whose
+  ability to compete for branching was unchanged, were able to increase in
+  size compared to the control simulation presented in Fig\,\ref{fig:main}C
+  (Fig.\,\ref{fig:whiskertrim}B). The C3 barrel disappeared altogether if its
+  competition parameter was reduced below a half of the original value.}
 
@@
   \includegraphics[width=\linewidth]{./Fig5.png}
-  \captionof{figure}{\MPthreePred{Simulating whisker trimming by reduction of
-      $\epsilon$ for the thalamocortical projection for barrel C3, for which
-      $i=13$. All other model parameters match those in
-      Fig.\,\ref{fig:main}C. \textbf{A} Graph of individual barrel measures
-      plotted vs.~the multiplier applied to $\epsilon_{13}$. \textbf{B} The
-      pattern for $\epsilon_{m}=0.64$.}}
+  \captionof{figure}{\MPthreePred{Simulating whisker trimming by a reduction
+      of the competition parameter, $\epsilon$, in the expression $\chi_i$
+      (Eq.\,\ref{eq:comp}) for the C3 thalamocortical projection. For C3,
+      $\epsilon$ is replaced by $m \epsilon$, where $m$ is a scalar
+      multiplier. For the rest of the projections, $\chi_i$ is left unchanged
+      and all other parameters are identical to those reported in
+      Fig.\,\ref{fig:main}C. \textbf{A} The pattern formed after 30000 steps
+      for $m=0.64$. Barrel C3 is reduced in size as its axon branching
+      competes less strongly with its neighbours, whose area is slightly
+      increased. \textbf{B} Individual barrel area plotted vs.~$m$
+      shows that as $m$ is reduced, the area of C3 (blue squares) goes down,
+      disappearing for $m\approx0.5$, whereas the average area of neighbouring
+      barrels (triangles) increases.}}
   \label{fig:whiskertrim}

### 2020-07-02 paper/briefpaper/brief.tex
Updates to a combined Fig 4

@@
   \includegraphics[width=\linewidth]{./Fig4.png}
-  \captionof{figure}{Guidance
-    fields and emergent barrel pattern in a Fgf8 misexpression experiment
-    (c.f. \citealp{assimacopoulos_fibroblast_2012}), simulated by reflecting
-    $\rho_1$ from Fig.\,\ref{fig:main}A at the join. \cmnt{The model
-      parameters match those in Fig.\,\ref{fig:main}C. Scale
-      bars are 1\,mm.}}
+  \captionof{figure}{
+    \MPthreePred{
+      Two simulated whisker barrel manipulations.  Guidance fields
+      (\textbf{A}) and emergent barrel pattern (\textbf{B}) in a Fgf8
+      misexpression experiment
+      (c.f.~\citealp{assimacopoulos_fibroblast_2012}), simulated by reflecting
+      $\rho_1$ from Fig.\,\ref{fig:main}A at the join of the original boundary
+      with its mirror. The model parameters match those in
+      Fig.\,\ref{fig:main}C. \textbf{C} Simulating whisker trimming by a
+      reduction of the competition parameter, $\epsilon$, in the expression
+      $\chi_i$ (Eq.\,\ref{eq:comp}) for the C3 thalamocortical projection. For
+      C3, $\epsilon$ is replaced by $m\,\epsilon$, where $m$ is a scalar
+      multiplier. For the rest of the projections, $\chi_i$ is left unchanged
+      and all other parameters are identical to those reported in
+      Fig.\,\ref{fig:main}C. The pattern formed after 30000 steps with
+      $m=0.64$. Barrel C3 is reduced in size as its axon branching competes
+      less strongly with its neighbours, whose area is slightly
+      increased. \textbf{D} Individual barrel area plotted vs.~$m$ shows that
+      as $m$ is reduced, the area of C3 (blue squares) goes down, disappearing
+      for $m\approx0.5$, whereas the average area of neighbouring barrels
+      (triangles) increases. Scale bars are 1\,mm.
+    }
+  }
   \label{fig:fgf8}
@@
   over which the C3 projection could make connections and its resulting barrel
-  size was diminished (Fig.\,\ref{fig:whiskertrim}A). Its neighbours, whose
+  size was diminished (Fig.\,\ref{fig:fgf8}C). Its neighbours, whose
   ability to compete for branching was unchanged, were able to increase in
   size compared to the control simulation presented in Fig\,\ref{fig:main}C
-  (Fig.\,\ref{fig:whiskertrim}B). The C3 barrel disappeared altogether if its
+  (Fig.\,\ref{fig:fgf8}D). The C3 barrel disappeared altogether if its
   competition parameter was reduced below a half of the original value.}
-
-\begin{center}
-  \includegraphics[width=\linewidth]{./Fig5.png}
-  \captionof{figure}{\MPthreePred{Simulating whisker trimming by a reduction
-      of the competition parameter, $\epsilon$, in the expression $\chi_i$
-      (Eq.\,\ref{eq:comp}) for the C3 thalamocortical projection. For C3,
-      $\epsilon$ is replaced by $m \epsilon$, where $m$ is a scalar
-      multiplier. For the rest of the projections, $\chi_i$ is left unchanged
-      and all other parameters are identical to those reported in
-      Fig.\,\ref{fig:main}C. \textbf{A} The pattern formed after 30000 steps
-      for $m=0.64$. Barrel C3 is reduced in size as its axon branching
-      competes less strongly with its neighbours, whose area is slightly
-      increased. \textbf{B} Individual barrel area plotted vs.~$m$
-      shows that as $m$ is reduced, the area of C3 (blue squares) goes down,
-      disappearing for $m\approx0.5$, whereas the average area of neighbouring
-      barrels (triangles) increases.}}
-  \label{fig:whiskertrim}
-\end{center}
 

### 2020-07-02 analysis/plot_trim.py

@@ -108,9 +108,9 @@ ax2.set_ylim([0,0.11])
 ax1.set_xlim([0.5,1.0])
 ax2.set_xlim([0.5,1.0])
 
-ax1.set_xlabel ('$\epsilon_m$')
-ax1.set_ylabel ('$A_i$', labelpad=35, rotation=0)
-ax2.set_xlabel ('$\epsilon_m$')
+ax1.set_xlabel ('$m$')
+ax1.set_ylabel ('$Area$', labelpad=35)
+ax2.set_xlabel ('$m$')
 ax2.set_ylabel ('$\delta$', labelpad=22, rotation=0)
 
 lw = 2; ll = 6

### 2020-07-02 analysis/plot_trim_areaonly.py

@@ -1,0 +1,107 @@
+# Make the whisker trimming line plot (barrel area vs epsilon, etc)
+
+# To access argv and also include the include dir
+import sys
+sys.path.insert (0, './include')
+import numpy as np
+# Import data loading code
+#import load as ld
+import BarrelData as bd
+# Import MY plotting code:
+import plot as pt
+import matplotlib
+import matplotlib.pyplot as plt
+# Direct HDF5 access
+import h5py
+# My domcentres linear fit code:
+import domcentres as dc
+import sebcolour
+col = sebcolour.Colour
+import matplotlib.transforms
+
+# Set plotting defaults
+fs = 26
+fnt = {'family' : 'Arial',
+       'weight' : 'regular',
+       'size'   : fs}
+matplotlib.rc('font', **fnt)
+
+# IMPORTANT for svg output of text as things that can be edited in inkscape
+plt.rcParams['svg.fonttype'] = 'none'
+
+# And plot this longhand here:
+F1 = plt.figure (figsize=(10,6))
+
+## load data from csv
+trimdata = np.genfromtxt ('postproc/whisker_trim_individual.csv', delimiter=",", names=True)
+td = np.sort (trimdata, order='barrel')
+
+alldata = np.genfromtxt ('postproc/whisker_trim_overall.csv', delimiter=",", names=True)
+ad = np.sort (alldata, order='eps_mult')
+
+# Show an example map for this multiplier
+map_mult = 0.64
+
+# Get per-barrel results
+# barrels of interest are: B3, C2, C3, C4, D2, D3
+#              TC indices:  8, 12, 13, 14, 22, 23
+tdc3 = td[td[:]['barrel_index'] == 13] # C3
+tdb3 = td[td[:]['barrel_index'] == 8]  # B3
+tdc2 = td[td[:]['barrel_index'] == 12]
+tdc4 = td[td[:]['barrel_index'] == 14]
+tdd2 = td[td[:]['barrel_index'] == 22]
+tdd3 = td[td[:]['barrel_index'] == 23]
+
+ax1 = F1.add_subplot(1,1,1)
+
+xmax = max(tdc3[:]['eps_mult'])
+xmin = min(tdc3[:]['eps_mult'])
+
+others_size = 6
+main_size = 12
+m_width = 3
+
+col_c3 = col.black
+col_b3 = col.blue3
[…]

### 2020-07-02 paper/briefpaper/brief.tex
Correction to caption for the average neighbour barrel area

@@
       less strongly with its neighbours, whose area is slightly
-      increased. \textbf{D} Individual barrel area plotted vs.~$m$ shows that
-      as $m$ is reduced, the area of C3 (blue squares) goes down, disappearing
-      for $m\approx0.5$, whereas the average area of neighbouring barrels
-      (triangles) increases. Scale bars are 1\,mm.
+      increased. \textbf{D} Individual barrel area plotted vs.~$m$ for C3
+      (black squares) shows that
+      as $m$ is reduced, its area goes down, disappearing
+      for $m\approx0.5$, whereas the mean area of neighbouring barrels
+      B3, C2, C4, D2 \& D3
+      (grey circles) increases. Scale bars are 1\,mm.
     }

### 2020-07-02 paper/briefpaper/brief.tex
More caption tweaking

--- 2020-07-02 paper/briefpaper/brief.tex
@@
 \begin{center}
-  \includegraphics[width=\linewidth]{./Fig4.png}
-  \captionof{figure}{
-    \MPthreePred{
-      Two simulated whisker barrel manipulations.  Guidance fields
+  \includegraphics[width=\linewidth]{./Fig4.png} \captionof{figure}{
+    \MPthreePred{Two simulated whisker barrel manipulations.  Guidance fields
       (\textbf{A}) and emergent barrel pattern (\textbf{B}) in a Fgf8
@@
       with its mirror. The model parameters match those in
-      Fig.\,\ref{fig:main}C. \textbf{C} Simulating whisker trimming by a
-      reduction of the competition parameter, $\epsilon$, in the expression
-      $\chi_i$ (Eq.\,\ref{eq:comp}) for the C3 thalamocortical projection. For
-      C3, $\epsilon$ is replaced by $m\,\epsilon$, where $m$ is a scalar
-      multiplier. For the rest of the projections, $\chi_i$ is left unchanged
-      and all other parameters are identical to those reported in
-      Fig.\,\ref{fig:main}C. The pattern formed after 30000 steps with
-      $m=0.64$. Barrel C3 is reduced in size as its axon branching competes
-      less strongly with its neighbours, whose area is slightly
-      increased. \textbf{D} Individual barrel area plotted vs.~$m$ shows that
-      as $m$ is reduced, the area of C3 (blue squares) goes down, disappearing
-      for $m\approx0.5$, whereas the average area of neighbouring barrels
-      (triangles) increases. Scale bars are 1\,mm.
-    }
-  }
+      Fig.\,\ref{fig:main}C.  Scale bars are 1\,mm. \textbf{C} Simulating
+      whisker trimming by a reduction of the competition parameter,
+      $\epsilon$, in the expression $\chi_i$ (Eq.\,\ref{eq:comp}) for the C3
+      thalamocortical projection. For C3, $\epsilon$ is replaced by
+      $m\,\epsilon$, where $m$ is a scalar multiplier. For the rest of the
+      projections, $\chi_i$ is left unchanged and all other parameters are
+      identical to those reported in Fig.\,\ref{fig:main}C. The pattern formed
+      after 30000 steps with $m=0.64$. Barrel C3 is reduced in size as its
+      axon branching competes weakly with its neighbours, whose area is
+      slightly increased. \textbf{D} Individual barrel area (mm$^2$) plotted
+      vs.~$m$ for C3 (black squares) shows that as $m$ is reduced, its area
+      goes down, disappearing for $m\approx0.5$, whereas the mean area of
+      neighbouring barrels B3, C2, C4, D2 \& D3 (grey circles) increases. The
+      dotted grey line indicates $m=0.64$ for comparison with panel C.}}
   \label{fig:fgf8}

### 2020-07-02 paper/briefpaper/brief.tex
Brings in updated text from Stu's google doc

@@
 projections are influenced by interactions with two or more signalling
-gradients that encourage migration in opposing directions.
-%
-%Note that these dynamics are quite unlike classic chemospecificity models
-%\citep{sperry_chemoaffinity_1963}, which essentially assume center-points
-%i.e., conditions in the target tissue that instruct pre-identified afferents
-%to stop growing.
-%
-As the number of distinct guidance fields is unlikely to
-approach the number of barrels, \cmnt{modifications to the model} were required.
+gradients that encourage migration in opposing directions. As the number of
+guidance fields is unlikely to approach the number of individual barrels,
+\cmnt{modifications to the model} were required.
 
 % Justify competition term. This pulled from Stuart's Google doc
-\mpfive{We reasoned that an arbitrary number of distinct field locations
-  may be specified using a minimum of two guidance fields, if the tendency for
-  the density of each projection to concentrate is appropriately balanced
-  against the tendency of projections that interact more strongly with the
-  guidance fields to migrate further in the directions specified by their
+\mpfive{We reasoned that an arbitrary number of distinct field locations may
+  be specified using a minimum of two guidance fields, if the tendency for the
+  density of each projection to concentrate is appropriately balanced against
+  the tendency of projections that interact more strongly with the guidance
+  fields to migrate further in the directions specified by their
   gradients. Accordingly, projections that interact most strongly with a given
@@
 \mpfive{Such dynamics are quite unlike those described by classic
-  chemospecificity models} \citep{sperry_chemoaffinity_1963},
-\mpfive{which essentially assume center-points by specifying conditions
-  in the target tissue that instruct pre-identified afferents to stop
-  growing. Consider, for example, that when simulated in isolation from
-  one-another, all projections in the model described would simply migrate to
-  the extrema of the cortical guidance fields.}
+  chemospecificity models} \citep{sperry_chemoaffinity_1963}, \mpfive{which
+  essentially assume center-points by specifying conditions in the target
+  tissue that instruct pre-identified afferents to stop growing. Consider, for
+  example, that when simulated in isolation from one-another, all projections
+  in the model described would simply migrate to the extrema of the cortical
+  guidance fields.}
 
@@
   competition between simulated thalamocortical projections for cortical
-  territory, by increasing the tendency for the density of each projection to
-  concentrate. The major modification required was thus to introduce into the
-  model an additional source of competition between thalamic projections.}
+  territory, by increasing the tendency for each projection to compete for
+  cortical space into which to concentrate. The major modification required
+  was thus to introduce into the model an additional source of competition
+  between thalamic projections.}
 %
@@
 %
-\mpfive{This term contributes to the \emph{flux} of axonal branching as
-  an additional source of diffusion, which reduces the branching density for a
-  given projection where the branches of other projections are dense. The
-  strength of this additional competition is parameterized by
-  $\epsilon$}. Note that this operation is local to individual afferent
-projections.
+\mpfive{This term contributes to the \emph{flux} of axonal branching as an
+  additional source of diffusion, scaled by $\epsilon$, which reduces the
+  branching density for a given projection where the branches of other
+  projections are dense. Note that this operation is local to individual
+  afferent projections.}
 
-\MPone{In addition, the model outlined requires that molecular
+\MPone{In addition, the model we have outlined requires that molecular
   guidance gradients in the cortex are complemented by graded values of the
@@
 \citealp{gao_regulation_1998,dufour_area_2003,vanderhaeghen_developmental_2004,speer_grading_2005,torii_role_2013}).
+\MPone{We assume that such patterning arises because relative strengths of
+  interaction with guidance molecules (e.g., ephrin-A5) in the cortex are
+  correlated with the relative concentrations of complementary molecules
+  (e.g., EphA4) in the thalamus, and thus with thalamic position along the
+  axis to which their gradients are aligned.}
 
-\MPone{For simplicity, both the thalamic interaction gradients and the
-  cortical guidance fields in the model were chosen in the first instance to
-  be linear and orthogonal. Hence a given pair of gamma values corresponds to
-  the coordinate of a barreloid center in the VPM (in a reference plane
-  defined by the anterior-posterior and medial-lateral axes. Coordinates were
-  estimated from Fig 5d of} \citep{haidarliu_size_2001}, \MPone{and
-  scaled (arbitrarily) such that $\gamma\in\pm 2$.}
+\MPone{For simplicity, the two simulated thalamic interaction gradients, as
+  well as the two cortical guidance gradients, were initially chosen to be
+  linear and orthogonal. Hence a given pair of $\gamma$ values corresponds to
+  the coordinate of a barreloid center in the VPM. Coordinates, in a reference
+  plane defined by the anterior-posterior and medial-lateral axes, were
+  estimated from Fig 5d of} \citep{haidarliu_size_2001}, \MPone{and scaled
+  such that $\gamma\in\pm 2$. Note that this scaling is arbitrary because
+  according to the model the coordinates provide relative position information
+  only.}
 
@@
 Eqs.~\ref{eq:dc}--\ref{eq:comp} were solved for $N=41$ projections on the
-resulting domain, using $M=2$ linear signalling gradients aligned with the
-anterior-posterior and medial-lateral axes.  \mpfour{These gradients may
-  extend well beyond the boundary of the barrel field (e.g. as ephrin-5 does;
-  see \citealp{miller_epha7-ephrin-a5_2006}) but they are shown only within
-  the boundary in Fig.\,\ref{fig:main}A, to indicate that external values are
-  not used by the simulation. Simulations were stepped through 30000
-  iterations of Eqs.\,\ref{eq:dc}--\ref{eq:comp} ($\delta t=0.0001$)}
+resulting domain, $S$, using $M=2$ linear signalling gradients aligned with
+the anterior-posterior and medial-lateral axes.  \mpfour{These gradients are
+  shown with the barrel field boundary in Fig.\,\ref{fig:main}A for clarity,
+  though like ephrin-5 they may be thought of as extending across the cortical
+  hemisphere \citep{miller_epha7-ephrin-a5_2006}. Simulations were stepped
+  through 30000 iterations of Eqs.\,\ref{eq:dc}--\ref{eq:comp} ($\delta
+  t=0.0001$)}
 
-From initial conditions sample from a uniform random distribution for
-$a(\mb{x},0)\in(0.2,0.4)$ and $c(\mb{x},0)=0$, and across a wide range of
-parameter values,
-%From random initial conditions for $a(\mb{x},0)$, \MPtwo{with each hexagonal element
-%  of $a_i$ set from a uniform distribution according to $0.2+{U}(0,0.2)$, and
-%  $c(\mb{x},0)=0$},
-a clear Voronoi-like tesselation \cmnt{of topographically organised
-  projections} emerged, \cmnt{confirming that barrel maps can self-organise in
-  the absence of pre-specified center points. Parameters for the example
-  simulation shown in Fig.\,\ref{fig:main}C (see Movie S1) were obtained by
-  conducting a full parameter sweep and choosing a combination of values that
-  scored well against three measures.}
+\cmnt{Across a wide range of parameter values, random initial conditions (a
+  uniform random distribution for $a(\mb{x},0)\in(0.2,0.4)$, $c(\mb{x},0)=0$)
+  eventually yielded a clear Voronoi-like tessellation of topographically
+  organised thalamocortical projections, confirming that barrel maps can
+  self-organise in the absence of pre-specified center points. The
+  organization is apparent in a plot of the identity of the projection for
+  which the connection density is maximal at each simulated cortical location,
+  as shown in Fig.\,\ref{fig:main}C. Parameters for the example simulation
+  shown in Fig.\,\ref{fig:main}C (see also Movie S1) were obtained by
+  conducting a full parameter sweep and choosing a combination of parameter
+  values ($\alpha=3.6$, $\beta=16.67$, $k=3$, $D=0.5$, $\epsilon=1.2$) that
+  scored well against the following three measures.}
 
@@
 \metrics{First, we used an algorithm introduced by Honda to measure the
-  discrepancy of each barrel shape from a Dirichletform shape. Low overall
-  values of the dimensionless \emph{Honda-}$\delta$ metric (called $\Delta$
-  in}
-\citealp{honda_geometrical_1983}) \metrics{obtained from simulated barrels indicate a close
-  correspondence of the simulated barrel field with a Voronoi tesselation, and
-  thus with a biological barrel field. For the
-  % Run ./build/sim/honda boundaries/rat_barrels/wb_110405_Dirichlet.svg to
-  % get this result:
-  tesselation overlaid on the real barrel field in Fig.\,\ref{fig:main}A,
-  $\delta=0.025$, and a reduction in the Honda-$\delta$ metric in simulation
-  confirmed that an equivalent `good' Voronoi pattern} (see
-\citealp{senft_mouse_1991}) \metrics{emerged within $\approx$~20000 iterations
-  (Fig.\,\ref{fig:main}D, red circles). Second, we devised a \emph{pattern
-    difference} measure that is sensitive to deviations in the component
-  shapes and overall topographic registration of two tesselations, $\eta$, and
-  we used this measure to compare the simulated barrel fields to the real
-  barrel field from which the simulation boundary shape was obtained (see
-  Materials and Methods for details). A similar reduction in pattern
-  difference over simulation time confirmed that the shapes and arrangement of
-  emergent connection fields matched those of the real barrel field by around
-  20000 iterations. Third, we measured the \emph{connection selectivity} at
-  each location on the cortical sheet, as the connection density of the most
-  dense projection divided by the sum over all projection densities,
-  $\omega$.}
-%
+  discrepancy of each barrel shape from a Dirichletform shape}
+\citep{honda_geometrical_1983}.  \metrics{Low overall values of this
+  \emph{Honda-}$\delta$ metric obtained from simulated barrels indicate a
+  close correspondence of the simulated barrel field with a Voronoi
+  tesselation, and thus with a biological barrel field. For the tesselation
+  that is overlaid on the real barrel field in Fig.\,\ref{fig:main}A,
+  $\delta=0.025$, and a reduction in $\delta$ in the example simulation over
+  time confirmed that an equivalent `good' Voronoi pattern} (see
+\citealp{senft_mouse_1991}) \metrics{can emerge within $\approx$~20000
+  iterations (Fig.\,\ref{fig:main}D, red circles). Second, we devised a
+  \emph{pattern difference} measure that is sensitive to deviations in the
+  component shapes and overall topographic registration between two
+  tesselations, $\eta$, and we used this measure to compare the simulated
+  barrel fields to the real barrel field from which the boundary shape applied
+  to the simulation was obtained (see \emph{Materials and Methods} for
+  details). A similar reduction in $\eta$ in the development of the example
+  simulation confirmed that the shapes and arrangement of emergent connection
+  fields came to match those of the real barrel field by around 20000
+  iterations (Fig.\,\ref{fig:main}D, black squares). Third, we measured the
+  \emph{connection selectivity} at each location on the cortical sheet, as the
+  connection density of the most dense projection divided by the sum over all
+  projection densities, $\omega$. The connection selectivity increased as the
+  barrel map self-organised in the example simulation (Fig.\,\ref{fig:main}D,
+  grey hexagons), and the selectivity became concentrated in regions
+  overlapping with the emergent barrel centers (Fig.\,\ref{fig:main}E).}
+
 \metrics{Against these three metrics we are also able to characterise the
@@
   the sensitivity of the model to variation in its inputs.}
-
-% Most of this commented out region is covered in the para above.
-%
-%A reduction in the Honda delta
-%metric (see \citealp{senft_mouse_1991}) confirmed that a `good' Voronoi
-%pattern emerged within \cmnt{$\approx$~20000} iterations
-%(Fig.\,\ref{fig:main}D).  A measure of the difference in \cmnt{barrel area}
-%between real and simulated barrels revealed a strong correspondence (see
-%Fig.\,\ref{fig:main}D, squares). \cmnt{In a chosen region, the ratio of
-%  \emph{the connection density of the TC type $i$ which makes the densest
-%    connections} to \emph{the sum of all connection densities} gives a measure
-%  of the selectivity of the connection density, $\mu(\mb{x},t)$. We summed this measure
-%  across the entire barrel domain and named it $\omega(t)$ (see
-%  Fig.\,\ref{fig:main}D, grey hexagons). $\omega$ increases during the first
-%  20000 simulation steps and converges on a stable
-%  value. Fig.\,\ref{fig:main}E shows the selectivity measure plotted as a
-%  luminance map, giving a more graded impression of the structure which
-%  develops in the system.}
 
@@
 
-\MPthreePar{To determine the robustness of the model to its parameters, we
-  investigated its behaviour with respect to changes in the parameters $D$,
-  $\epsilon$ and the ratio of $\alpha$ to $\beta$. Fig.\,\ref{fig:paramsweep}A
-  shows the value of the metrics $\delta$, $\eta$ and $\omega$ after 30000
-  iterations, plotted as colour maps, for 216 parameter sets. The maps
-  indicate that the pattern is not strongly influenced by the choice of
-  $\alpha/\beta$, which was varied by 5 orders of magnitude. The most
-  Voronoi-like patterns were achieved for parameter sets for which $D$ and
-  $\epsilon$ were high. The barrel area metric, $\eta$ indicates a strong
-  preference for $D\approx0.5$, although this does not give the solution of
-  maximum selectivity, which favours lower values of $D$. Thus, there is a
-  trade-off between good selection and good Voroni pattern conformation.  The
-  parameter set indicated by the asterisk in Fig.\,\ref{fig:paramsweep}A was
-  chosen for Fig.\,\ref{fig:main} because it provides good pattern
-  correspondence and a close approximation to a Voronoi tesselation.}
+\MPthreePar{Fig.\,\ref{fig:paramsweep}A shows values of $\delta$, $\eta$ and
+  $\omega$ obtained after 30000 iterations, from 216 independent simulations,
+  each representing a unique combination of the model parameters $D$,
+  $\epsilon$, and the ratio $\alpha/\beta$. First we observe that
+  self-organization is highly robust to the ratio $\alpha/\beta$, across five
+  orders of magnitude, with respect to all three metrics. Second, the most
+  strongly Dirichletform patterns (low \emph{Honda}-$\delta$) were generated
+  by simulations in which the diffusion constant $D$ and the strength of
+  competition $\epsilon$ were high. Third, strongest connection selectivities,
+  $\omega$, were obtained for lower values of $D$. Fourth, variation in the
+  pattern difference metric, $\eta$, indicated that the alignment between real
+  and simulated patterns was greatest for intermediate rates of diffusion,
+  $D\approx0.5$. Together these results indicate that when competition is
+  strong, the rate of diffusion determines a trade-off such that fields emerge
+  to be barrel-shaped when diffusion is fast and they emerge to be more
+  selectively innervated when diffusion is slow.}
 
-\MPthreePar{In Fig.\,\ref{fig:paramsweep}B, we present some patterns
-  representing alternative parameter choices. Because the patterns are robust
-  to alterations in $\epsilon$ and the choices of $\alpha$ and $\beta$, we
-  concentrate our attention on the diffusion constant, $D$. Changing $D$,
-  whose value defines the rate at which a region of localized branching will
-  spread out across the domain, $S$, is equivalent to making changes to the
-  size of $S$ with a fixed branching diffusion rate. These simulations can
-  therefore be thought of as representing animals with a varying cortical
-  areas but with unchanged cortical complexity. The most extreme example
-  (Fig.\,\ref{fig:paramsweep}B iii) could represent an animal with a very
-  large cortical area (such as the South American capybara) which receives
-  sensory input from $N=41$ thalamic barreloids. Here, a large proportion of
-  the domain is occupied by the projections with intermediate interaction
-  parameters and those with strong interaction parameters are bunched up
-  towards the edge of the domain. In this case, a barrel pattern fails to
-  form, agreeing with the observation that cortical whisker barrels are absent
-  in the capybara.}
+\MPthreePar{The parameters of the example simulation are indicated in
+  Fig.\,\ref{fig:paramsweep}A using an asterisk. In
+  Fig.\,\ref{fig:paramsweep}B, we also present examples of alternative
+  patterns that emerge for different choices of $D$. Decreasing the rate of
+  diffusion may be considered equivalent to increasing the overall size of the
+  domain, $S$, hence insights into barrel development in species with
+  substantially larger barrel fields, such as capybara which do not have
+  clearly identifiable barrels, may be gained by studying pattern formation
+  when $D$ is small. In this context, it is interesting to note that for small
+  $D$, the organization is predicted to be topological but highly irregular,
+  with a general expansion in the territory occupied by the central versus
+  peripheral domains that would presumably manifest as an absence of
+  identifiable barrel fields (Fig.\,\ref{fig:paramsweep}Biii).}
 
@@
 
-% This section needs to emphasize that the absolute values of gamma are not important
-\MPthreeSens{We performed a sensitivity analysis to determine if the quality
-  of the pattern (i.e.~the values of $\delta$, $\eta$ \& $\omega$ after
-  $t=30000$ steps) was affected by perturbations to any of the following
-  inputs:
-%
-  i)~the magnitude and offset of the noise applied to $a_i$ at $t=0$;
-%
-  ii)~random noise applied to the interaction parameters, $\gamma_{i,j}$;
-%
-  iii)~random noise (at various length scales) applied to the guidance fields;
-%
-  and
-%
-  iv)~The angle and magnitude of one of the guidance fields, ($\rho_1$;
-  Fig.\,\ref{fig:main}A).
-}
+\MPthreeSens{Next we conducted a sensitivity analysis to determine the extent
+  to which the quality of the pattern (after $t=30000$ iterations) is affected
+  by perturbations to i)~the magnitude and offset of the noise applied to
+  $a_i$ at $t=0$; ii)~noise applied to the interaction parameters,
+  $\gamma_{i,j}$; iii)~noise (at various length scales) applied to the
+  guidance fields; and iv)~the magnitude and orientation of one cortical
+  guidance field relative to the other (Fig.\,\ref{fig:main}A).}
 
-\MPthreeSens{First, we conducted 10 runs of the simulation shown in
-  Fig.\,\ref{fig:main}C to evaluate the mean and standard deviation of the
-  metrics in the default model. The results (where deviations occur as a
-  result of the different random initial conditions for $a_i$) were
-  $\delta=0.089\,(0.004)$, $\eta=0.2108\,(0.002)$ and
-  $\omega=0.2165\,(0.0001)$. The pattern was barely affected by the random
-  changes to $a_i(t=0)$ on each run.}
-%
-\MPthreeSens{Next, we modified the parameters of the generating uniform
-  distribution to initialize $a_i$ for each hex to $0.1+{U}(0,0.4)$,
-  increasing the variability of $a_i(\mb{x},t=0)$ without changing the mean
-  initial activation. We ran this system 10 times and obtained the results
-  $\delta=0.090\,(0.005)$, $\eta=0.2136\,(0.004)$ and
-  $\omega=0.2166\,(0.0003)$. Paired two-sample t-tests demonstrated that the
-  increase in noise amplitude made no significant difference to the metric
-  results ($\delta: \mathrm{t}=-0.69,\mathrm{p}=0.49$, $\eta:
-  \mathrm{t}=-2.0,\mathrm{p}=0.06$, $\omega:
-  \mathrm{t}=-1.7,\mathrm{p}=0.11$). 10 runs of a noise-free initial condition
-  with $a_i=0.3$ for every hex showed that the model is deterministic (all
-  runs produced the results $\delta=0.0928$, $\eta=0.21236$ and
-  $\omega=0.21649$) and that stochasticity in $a_i(\mb{x},t=0)$ is not
-  required by this model, as all three results lay within one standard
-  deviation of those for the runs in which $a_i$ was initialized
-  stochastically.}
-
-% Quick check that noise free result is within 1 SD for the noisier initial conditions
-%         Noisefree    Normal noise    NF SD from Normal?   Noisy            NF SD from Noisy?
-% delta   0.0928       0.089(0.004)    <1                   0.090 (0.005)    <1
-% eta     0.2124       0.2108(0.002)   <1                   0.2136 (0.004)   <1
-% omega   0.21649      0.2165(0.0001)  <1                   0.2166 (0.0003)  <1
-
-\MPthreeSens{Fig\,\ref{fig:sens}A shows the results for perturbations ii) to
-  iv). The results show that the Voronoi nature of the pattern ($\delta$) and
-  the selectivity of the fields are not strongly affected by interaction
-  parameter noise, but the pattern metric, $\eta$ shows a degradation in the
-  pattern arrangement when a) noise in the interaction parameters causes their
-  disordering (find better way to express?), b) noise in the guidance field
-  with a high magnitude and short length scale causes disruption to the barrel
-  edges and c) the guidance fields tend towards a non-orthogonal arrangement,
-  or if one field becomes substantially stronger in its interaction with the
-  thalamocortical projections. Fig\,\ref{fig:sens}B shows perturbed examples.}
+\MPthreeSens{Using the parameters of the example simulation
+  (Fig.\,\ref{fig:main}C) we established baseline mean and standard deviations
+  from ten independent simulations with initial uniform random values for
+  $a(\mb{x},0)\in(0.2,0.4)$, to be $\delta=0.089\pm 0.004$, $\eta=0.2108\pm
+  0.002$, and $\omega=0.2165\pm 0.0001$. Repeating with the variation in the
+  initial noise doubled ($a(\mb{x},0)\in(0.1,0.5)$), or removed altogether
+  ($a(\mb{x},0)=0.3$), generated distributions of $\delta$, $\eta$, and
+  $\omega$ that were not statistically different, as established using paired
+  two-sample t-tests. Adding noise to the interaction parameters ($\gamma$)
+  affected neither the \emph{Honda-}$delta$ or the \emph{connection
+    selectivity} measures substantially (see Fig\,\ref{fig:sens}A), and an
+  increase in the \emph{pattern difference} reflected an increase in the
+  occurrence of topological defects only when perturbations become so large as
+  to cause the ordering of $\gamma$ values from neighbouring thalamic sites to
+  be switched (see example map Fig\,\ref{fig:sens}Bi). Adding noise to the
+  cortical guidance field values, $\rho_1(\mathbf{x})$ and
+  $\rho_2(\mathbf{x})$, disrupted pattern formation only for high levels of
+  noise applied at short length scales, which manifested as non-straight edges
+  at the domain boundaries (Fig\,\ref{fig:sens}Bii). Varying the slope of one
+  linear gradient $\rho_1(\mathbf{x})$ while keeping that of the other
+  constant caused elongation of the emergent domains along the corresponding
+  axis (Fig\,\ref{fig:sens}Biii), while pattern formation was not strongly
+  influenced by relaxing the assumption that the gradients of the two cortical
+  guidance fields are orthogonal (Fig\,\ref{fig:sens}Biv). Overall, the
+  sensitivity analysis revealed that self-organization of barrel-like fields
+  in the model is highly robust to a wide range of sources of perturbation.}
 
@@
   whisker barrel system. First, we simulated a seminal barrel duplication
-  experiments}
+  experimental paradigm}
 \citep{shimogori_fibroblast_2005,assimacopoulos_fibroblast_2012}
 \MPthreePred{in which the growth factor Fgf8, which is normally expressed at
-  the anterior end of the cortical sub-plate from around E9.5}
+  the anterior end of the cortical subplate from around E9.5}
 \citep{crossley_mouse_1995}, \MPthreePred{is ectopically expressed (by
-  electroporation) also at the posterior end. Because Fgf8 sets off a chain of
-  gene interactions which are involved in the control of the expression of
-  ephrin and cadherin molecules in gradients across the cortical plate, its
-  misexpression at the posterior pole of a cortical hemisphere could be
-  expected to mirror those gradients which usually vary along the
-  rostro-caudal axis. We assume that this results in a mirror of the primary
-  barrel cortex boundary along the rostro-caudal axis, to match the
-  experimental results of}
+  electroporation) also at the posterior pole of a cortical
+  hemisphere. Because Fgf8 sets off a chain of gene interactions involved in
+  controlling the graded expression of ephrin and cadherin molecules, its
+  misexpression at the posterior pole could be expected to mirror those
+  gradients which usually vary along the rostrocaudal axis. We assume that
+  this results in a mirror of the primary barrel cortex boundary along the
+  rostrocaudal axis, to match the experimental results of}
 \citealp{assimacopoulos_fibroblast_2012}. \MPthreePred{We also assume that a
   further consequence of the posterior Fgf8 electroporation is the mirrored
-  anterior-posterior guidance gradient $\rho_1$ (Fig.\,\ref{fig:fgf8}A). We
-  mirrored this across a reflection line which is orthogonal to its angle
-  $\phi_{\rho_1}=6$\textdegree, and left $\rho_2$ unmodified. Although the
-  rostro-caudal boundary reflection is not perfectly aligned with $\rho_1$,
-  and we might expect some modification of both $\rho_1$ \emph{and} $\rho_2$,
-  mirroring only $\rho_1$ retains the simplicity of the guidance fields; they
-  remain either linear ($\rho_2$), or piece-wise linear ($\rho_1$).}
-
-\MPthreePred{The new boundary shape and $\rho_1$ guidance gradient were the
-  only changes made to the simulation of Fig.\,\ref{fig:main}; the simulation
-  was run with $N=41$ thalamocortical axon types as before and the initial
-  values for $a_i$ were randomly initialised according to the same
-  distribution. The result, after 30000 simulation steps, was two
-  mirror-symmetrical barrel fields comprising $2N$ barrels
-  (Fig\,\ref{fig:main}B). This result suggests that by the misexpression of
-  Fgf8,} \cite{shimogori_fibroblast_2005} and \cite{assimacopoulos_fibroblast_2012}
-  \MPthreePred{created the correct cortical
-  conditions for the self-organization of the two mirrored barrel fields. }
-
-%However, merging the two fields to
-%create an extended mirror-symmetric boundary shape, with an anterior-posterior
-%guidance field reflected at the join, gave rise to two mirror-symmetrical
-%barrel fields comprising 2N barrels, i.e., two cortical fields for each
-%thalamic projection (Fig.\,\ref{fig:fgf8}). Together these results suggest that by the
-%misexpression of Fgf8, \cite{shimogori_fibroblast_2005}
-%effectively created one large barrel field rather than two distinct cortical
-%fields.
+  anterior-posterior guidance gradient $\rho_1$ (Fig.\,\ref{fig:fgf8}A). With
+  the modified $\rho_1$ guidance gradient the simulation was run with $N=41$
+  thalamocortical axon types, using the parameters of the example
+  simulation. The result, after 30000 iterations, was two mirror-symmetrical
+  barrel fields comprising $2N$ barrels (Fig\,\ref{fig:main}B), consistent
+  with the outcome of the original experiments.}
 
@@
   brain structures, including the barrel field in cortical layer
-  IV~\citep{jeanmonod_mouse_1981}. Amoung other changes, barrels associated
+  IV~\citep{jeanmonod_mouse_1981}. Amoungst other changes, barrels associated
   with deprived whiskers have a smaller area. This is found whether the animal
@@
   parameters unchanged. The reduced competition for branching limited the area
-  over which the C3 projection could make connections and its resulting barrel
-  size was diminished (Fig.\,\ref{fig:fgf8}C). Its neighbours, whose
-  ability to compete for branching was unchanged, were able to increase in
-  size compared to the control simulation presented in Fig\,\ref{fig:main}C
-  (Fig.\,\ref{fig:fgf8}D). The C3 barrel disappeared altogether if its
-  competition parameter was reduced below a half of the original value.}
+  over which the C3 projection could make connections and its resulting field
+  size was diminished (Fig.\,\ref{fig:fgf8}C). Emergent fields representing
+  the neighbours of the C3 projection, whose ability to compete for branching
+  was unchanged, increased in size. The C3 barrel disappeared altogether if
+  its competition parameter was reduced below a half of the original value. A
+  reduction in the area of barrel C3, comparable that induced by
+  \citep{kossut_effects_1992} (65\%) was obtained in simulation when the
+  competitiveness of projection C3, $\epsilon$, was reduced to 86\%, as
+  indicated in (Fig.\,\ref{fig:fgf8}D).}
+
+\MPthreePred{Together with the results of the misexpression experiment, the
+  consistency of the simulated whisker trimming results with those of the
+  original studies demonstrates how the model can be used to test specific
+  hypotheses about the contribution of intrinsic and extrinsic factors to the
+  development of cortical fields.}
 
@@
 
+% for ``end of methods''
+\MPone{We concentrated on the representation of the forty-six macrovibrissae
+  that constitute a given barrel field, because their thalamic and cortical
+  correlates are easily identifiable and consistently positioned. We exclude
+  the representation of the microvibrissae to limit the overall complexity of
+  the simulations, as well as the five rhinal whiskers as their representation
+  in S1 is isolated from the main barrel field. Consequently, we set $N=41$.}
+
 All code required to reproduce these results is available at

### 2020-07-02 analysis/include/BarrelData.py
Some changes to graphs for forthcoming updates to the figures

@@ -610,8 +610,12 @@ class BarrelData:
 
             # Get the Honda Dirichlet value (the overall value)
             self.honda[fi] = list(f['honda'])[0]
-            keylist = list(f['honda_arr_keys'])
-            vallist = list(f['honda_arr_vals'])
+            try:
+                keylist = list(f['honda_arr_keys'])
+                vallist = list(f['honda_arr_vals'])
+            except:
+                keylist = []
+                vallist = []
 
             delta_j_dict = {}
             for lcount in range(0, len(keylist)):
@@ -631,7 +635,10 @@ class BarrelData:
             # barrel fields. This is a total difference for all barrels:
             # sum_i^N (abs(area_diff_i))...
             self.area_diff[fi] = list(f['area_diff'])[0]
-            self.area_diffs[:,fi] = list(f['area_diffs'])
+            try:
+                self.area_diffs[:,fi] = list(f['area_diffs'])
+            except:
+                pass
 
             # ...now convert the area_diff in hexes into an area by multiplying
             # by the area of one hex

### 2020-07-02 analysis/plot_FigSensitivity.py

@@ -53,12 +53,12 @@ mo=2 # miss-out how many data points at end of the next three graphs?
 # Honda (Voronoi)
 ax1 = F1.add_subplot(3,3,1)
 l1, = ax1.plot(gmn_noise_gain[:-mo], gmn_honda[:-mo], 'o-', markersize=12, color=col.red, label='Honda $\delta$')
-# Omega (Localisation)
-ax2 = F1.add_subplot(3,3,4)
-l2, = ax2.plot(gmn_noise_gain[:-mo], gmn_locn[:-mo], 'h-', markersize=12, color=col.gray60, label='Locn $\omega$')
 # Eta (Pattern)
+ax2 = F1.add_subplot(3,3,4)
+l2, = ax2.plot(gmn_noise_gain[:-mo], gmn_eta[:-mo], 's-', markersize=12, color=col.black, label='$\eta$')
+# Omega (Localisation)
 ax3 = F1.add_subplot(3,3,7)
-l3, = ax3.plot(gmn_noise_gain[:-mo], gmn_eta[:-mo], 's-', markersize=12, color=col.black, label='$\eta$')
+l3, = ax3.plot(gmn_noise_gain[:-mo], gmn_locn[:-mo], 'h-', markersize=12, color=col.gray60, label='Locn $\omega$')
 
 # this is an inset axes over the third axis
 #ax3_1 = plt.axes([.2, .24, .1, .06])
@@ -95,12 +95,12 @@ for gdn in guidenoise:
 # Honda
 ax4 = F1.add_subplot(3,3,2)
 im4 = ax4.imshow (gdn_honda, cmap='inferno_r', origin='lower', vmin=honda_min, vmax=honda_max, interpolation='nearest')
-# Localization
-ax5 = F1.add_subplot(3,3,5)
-im5 = ax5.imshow (gdn_locn,  cmap='inferno', origin='lower', vmin=locn_min,  vmax=locn_max,  interpolation='nearest')
 # Eta
+ax5 = F1.add_subplot(3,3,5)
+im5 = ax5.imshow (gdn_eta,  cmap='inferno_r', origin='lower', vmin=eta_min,  vmax=eta_max,  interpolation='nearest')
+# Localization
 ax6 = F1.add_subplot(3,3,8)
[…]

### 2020-07-02 analysis/plot_trim_areaonly.py

@@ -5,7 +5,6 @@ import sys
 sys.path.insert (0, './include')
 import numpy as np
 # Import data loading code
-#import load as ld
 import BarrelData as bd
 # Import MY plotting code:
 import plot as pt
@@ -39,9 +38,6 @@ td = np.sort (trimdata, order='barrel')
 alldata = np.genfromtxt ('postproc/whisker_trim_overall.csv', delimiter=",", names=True)
 ad = np.sort (alldata, order='eps_mult')
 
-# Show an example map for this multiplier
-map_mult = 0.64
-
 # Get per-barrel results
 # barrels of interest are: B3, C2, C3, C4, D2, D3
 #              TC indices:  8, 12, 13, 14, 22, 23
@@ -53,6 +49,14 @@ tdd2 = td[td[:]['barrel_index'] == 22]
 tdd3 = td[td[:]['barrel_index'] == 23]
 
 ax1 = F1.add_subplot(1,1,1)
+
+# Max area
+ymax = max(tdc3[:]['area'])
+print ('Max area: {0}, 0.65 * max area (Kossut, 1992): {1}'.format (ymax, 0.65*ymax))
+# Show an example map for this multiplier. 0.86 is the area of the experimental manip
+map_mult = 0.64
+map_expt = 0.86
+
 
 xmax = max(tdc3[:]['eps_mult'])
 xmin = min(tdc3[:]['eps_mult'])
@@ -69,25 +73,27 @@ col_d2 = col.royalblue2
 col_d3 = col.cornflowerblue
 col_others = col.gray50
 
-
-l1_map = ax1.plot([map_mult,map_mult], [0, 0.16], '--', color=col.gray70)
+#l1_map = ax1.plot([map_mult,map_mult], [0, 0.16], '--', color=col.gray70)
+l1_expt = ax1.plot([map_expt,map_expt], [0, 0.16], '--', color=col.gray70)
+# The line of "shrunk by factor of 0.65" (Kossut, 1992)
+#l1_horz = ax1.plot([0, xmax], [0.65*ymax,0.65*ymax], '--', color=col.gray70)
 
 # Compute mean values of the neighbouring areas:
 others = np.vstack ((tdb3[:]['area'], tdc2[:]['area'], tdc4[:]['area'], tdd2[:]['area'], tdd3[:]['area']))
 mn_oth = np.mean(others, axis=0)
 
[…]

### 2020-07-03 paper/briefpaper/brief.tex
Tweaks from a read through a print out of the paper

@@
 
-% Colours for revisions:
-% Custom colours
-\definecolor{revgreen}{rgb}{0.220, 0.463, 0.114}
-\definecolor{revblue}{rgb}{0.235, 0.471, 0.847}
-\definecolor{revred}{rgb}{0.8353, 0.1255, 0.1255}
-\definecolor{revorange}{rgb}{0.8235, 0.4275, 0.0}
-\definecolor{revgrey}{rgb}{0.6, 0.6, 0.63}
-\definecolor{revcomp}{rgb}{0.0, 0.0, 0.0}
-\definecolor{revpurp}{rgb}{0.8, 0.1, 0.8}
-\definecolor{revpurp2}{rgb}{0.3882, 0.0745, 0.7098}
-\definecolor{revpink}{rgb}{0.7098, 0.0745, 0.6431}
-\definecolor{revbblack}{rgb}{0, 0, 0.3}
-\definecolor{revgblack}{rgb}{0, 0.2, 0}
+\input{sebcolour.tex}
 
 % Miscellaneous changes
-\newcommand{\cmnt}[1]{\textcolor{blue}{#1}}
+\newcommand{\cmnt}[1]{\textcolor{colcmnt}{#1}}
 % Justify gamma/interaction parameter choice. major point 1.
-\newcommand{\MPone}[1]{\textcolor{revpink}{#1}}
+\newcommand{\MPone}[1]{\textcolor{colmpone}{#1}}
 % Blue changes are major point 2 (competition mechanism) and other, general
 % uncategorised changes.
-\newcommand{\MPtwo}[1]{\textcolor{revblue}{#1}}
+\newcommand{\MPtwo}[1]{\textcolor{colmptwo}{#1}}
 % Parameter search. Major point 3
-\newcommand{\MPthreePar}[1]{\textcolor{revbblack}{#1}}
+\newcommand{\MPthreePar}[1]{\textcolor{colmpthreepar}{#1}}
 % Sensitivity analysis. Major point 3
-\newcommand{\MPthreeSens}[1]{\textcolor{revgblack}{#1}}
+\newcommand{\MPthreeSens}[1]{\textcolor{colmpthreesens}{#1}}
 % More predictions. Major point 3
-\newcommand{\MPthreePred}[1]{\textcolor{revorange}{#1}}
+\newcommand{\MPthreePred}[1]{\textcolor{colmpthreepred}{#1}}
 % Justify competition:
-\newcommand{\mpfive}[1]{\textcolor{revpurp2}{#1}}
+\newcommand{\mpfive}[1]{\textcolor{colmpfive}{#1}}
 % Minor point 4:
-\newcommand{\mpfour}[1]{\textcolor{revgreen}{#1}}
+\newcommand{\mpfour}[1]{\textcolor{colmpfour}{#1}}
 % Minor point 6:
-\newcommand{\mpsix}[1]{\textcolor{revpurp}{#1}}
+\newcommand{\mpsix}[1]{\textcolor{colmpsix}{#1}}
 % Describe the metrics
-\newcommand{\metrics}[1]{\textcolor{revred}{#1}}
+\newcommand{\metrics}[1]{\textcolor{colmetrics}{#1}}
 
@@
 bundle, \emph{prior} to barreloid formation \citep{agmon_organized_1993}.
-
-% Note: The key difference between this theory and the chemospecificity theory
-% is that of group membership.
 
@@
 
-% Justify competition term. This pulled from Stuart's Google doc
 \mpfive{We reasoned that an arbitrary number of distinct field locations may
@@
 \begin{equation} \label{eq:comp}
-  \color{revblue}
+  \color{colmptwo}
   \chi_i(\mb{x}, t) = \frac{\epsilon a_i}{N-1} \nabla \textstyle\sum_{j\ne i}^{N}{a_j}.
@@
   plane defined by the anterior-posterior and medial-lateral axes, were
-  estimated from Fig 5d of} \citep{haidarliu_size_2001}, \MPone{and scaled
+  estimated from Fig 5d of} \cite{haidarliu_size_2001}, \MPone{and scaled
   such that $\gamma\in\pm 2$. Note that this scaling is arbitrary because
@@
   through 30000 iterations of Eqs.\,\ref{eq:dc}--\ref{eq:comp} ($\delta
-  t=0.0001$)}
+  t=0.0001$).}
 
@@
   barrel fields to the real barrel field from which the boundary shape applied
-  to the simulation was obtained (see \emph{Materials and Methods} for
+  to the simulation was obtained (see \emph{Materials \& Methods} for
   details). A similar reduction in $\eta$ in the development of the example
@@
   peripheral domains that would presumably manifest as an absence of
-  identifiable barrel fields (Fig.\,\ref{fig:paramsweep}Biii).}
+  identifiable barrel fields (Fig.\,\ref{fig:paramsweep}B\,iii).}
 
@@
   two-sample t-tests. Adding noise to the interaction parameters ($\gamma$)
-  affected neither the \emph{Honda-}$delta$ or the \emph{connection
-    selectivity} measures substantially (see Fig\,\ref{fig:sens}A), and an
+  affected neither the \emph{Honda-}$\delta$ or the \emph{connection
+    selectivity} measures substantially (see Fig.\,\ref{fig:sens}A), and an
   increase in the \emph{pattern difference} reflected an increase in the
@@
   to cause the ordering of $\gamma$ values from neighbouring thalamic sites to
-  be switched (see example map Fig\,\ref{fig:sens}Bi). Adding noise to the
+  be switched (see example map Fig.\,\ref{fig:sens}B\,i). Adding noise to the
   cortical guidance field values, $\rho_1(\mathbf{x})$ and
@@
   noise applied at short length scales, which manifested as non-straight edges
-  at the domain boundaries (Fig\,\ref{fig:sens}Bii). Varying the slope of one
+  at the domain boundaries (Fig.\,\ref{fig:sens}B\,ii). Varying the slope of one
   linear gradient $\rho_1(\mathbf{x})$ while keeping that of the other
   constant caused elongation of the emergent domains along the corresponding
-  axis (Fig\,\ref{fig:sens}Biii), while pattern formation was not strongly
+  axis (Fig.\,\ref{fig:sens}B\,iii), while pattern formation was not strongly
   influenced by relaxing the assumption that the gradients of the two cortical
-  guidance fields are orthogonal (Fig\,\ref{fig:sens}Biv). Overall, the
+  guidance fields are orthogonal (Fig.\,\ref{fig:sens}B\,iv). Overall, the
   sensitivity analysis revealed that self-organization of barrel-like fields
@@
 
-% This should be a little less terse than in the original.
 \MPthreePred{To further investigate the interplay of genetic and intrinsic
@@
   rostrocaudal axis, to match the experimental results of}
-\citealp{assimacopoulos_fibroblast_2012}. \MPthreePred{We also assume that a
+\cite{assimacopoulos_fibroblast_2012}. \MPthreePred{We also assume that a
   further consequence of the posterior Fgf8 electroporation is the mirrored
@@
   simulation. The result, after 30000 iterations, was two mirror-symmetrical
-  barrel fields comprising $2N$ barrels (Fig\,\ref{fig:main}B), consistent
+  barrel fields comprising $2N$ barrels (Fig.\,\ref{fig:fgf8}B), consistent
   with the outcome of the original experiments.}
@@
       identical to those reported in Fig.\,\ref{fig:main}C. The pattern formed
-      after 30000 steps with $m=0.64$. Barrel C3 is reduced in size as its
-      axon branching competes weakly with its neighbours, whose area is
-      slightly increased. \textbf{D} Individual barrel area (mm$^2$) plotted
-      vs.~$m$ for C3 (black squares) shows that as $m$ is reduced, its area
-      goes down, disappearing for $m\approx0.5$, whereas the mean area of
-      neighbouring barrels B3, C2, C4, D2 \& D3 (grey circles) increases. The
-      dotted grey line indicates $m=0.64$ for comparison with panel C.}}
+      after 30000 steps with $m=0.86$. As its axon branching competes weakly
+      with its neighbours, field C3 is reduced to 65\% of its original size,
+      matching the average barrel area reduction observed by }
+    \citealp{kossut_effects_1992}. \MPthreePred{\textbf{D} Individual barrel
+      area (mm$^2$) plotted vs.~$m$ for C3 (black squares) shows that as $m$
+      is reduced, its area goes down, disappearing for $m\approx0.5$, whereas
+      the mean area of neighbouring barrels B3, C2, C4, D2 \& D3 (grey
+      circles) increases. The dotted grey line indicates $m=0.86$ for
+      comparison with panel C.}}
   \label{fig:fgf8}
@@
   reduction in the area of barrel C3, comparable that induced by
-  \citep{kossut_effects_1992} (65\%) was obtained in simulation when the
+  \citealp{kossut_effects_1992} (65\%) was obtained in simulation when the
   competitiveness of projection C3, $\epsilon$, was reduced to 86\%, as
-  indicated in (Fig.\,\ref{fig:fgf8}D).}
+  indicated in Fig.\,\ref{fig:fgf8}D.}
 
@@
   computational complexity of the simulation, manageable (simulating each
-  additional projection introduces an extra \textbf{YYY} new dynamical
+  additional projection introduces an extra 13030 new dynamical
   variables).}
@@
   \nabla\vcdot\mb{J}_i(\mb{x},t) = \nabla\vcdot\left(D \nabla
-  a_i-a_i\sum_{j=1}^{M} \gamma_{i,j}\nabla \rho_j(\mb{x}) \color{revblue}+
+  a_i-a_i\sum_{j=1}^{M} \gamma_{i,j}\nabla \rho_j(\mb{x}) \color{colmptwo}+
   \frac{\epsilon a_i}{N-1} \nabla \hat{a}_i \color{black}\right),
@@
 \nabla\vcdot\mb{J}_i = \nabla\vcdot\big(D \nabla a_i\big) -
-\nabla\vcdot\big(a_i \mb{g}_i\big) \color{revblue} + \frac{\epsilon}{N-1}\nabla\vcdot\big({a}_i\nabla\hat{a}_i\big)\color{black}.
+\nabla\vcdot\big(a_i \mb{g}_i\big) \color{colmptwo} + \frac{\epsilon}{N-1}\nabla\vcdot\big({a}_i\nabla\hat{a}_i\big)\color{black}.
 \end{equation}
@@
 \mb{g}_i\vcdot\nabla a_i % term3
-\color{revblue}
+\color{colmptwo}
 + \frac{\epsilon a_i}{N-1} \dvrg \hat{a}_i % term1_1
@@
   match and can separate patterns with straight boundaries from those with
-  noise on the boundary shape (For an example of `noisy edges', see Fig.\,\ref{fig:sens}B.ii).}
+  noise on the boundary shape (For an example of `noisy edges', see Fig.\,\ref{fig:sens}B\,ii).}
 %
@@
 \begin{equation} \label{eq:eta}
-\color{revred}
+\color{colmetrics}
   \eta =
@@
 \end{document}
-
-% NOTES
-
-%\MPtwo{In other words, we simply scaled the barreloid coordinates in an anterior-posterior/medial-lateral coordinate system to create the $\gamma$ values in the range [-2,2].
-Note that positive \emph{and} negative $\gamma$ values were generated; the origin of the coordinate system was located at the approximate center of the thalamic barreloid field.
-This prevented an overall drift of the cortical pattern away from the centroid of the barrel field and may explain why coordinate-giving transcription factors and the ephrins are often found as pairs acting in opposing directions.}  \MPtwo{We concentrated on macrovibrissae (excluding microvibrissae) because it is easy to identify their corresponding structures both in the thalamus and in the cortex.
-We also excluded the 5 rhinal whiskers because their cortical representation is in a separate island, distinct, and somewhat removed from the representation of the macrovibirissae.}}
-
-%\MPtwo{Thus, we suggest that there exist a pair of signalling molecules whose expression varies linearly (with orthogonal gradients) across the thalamus, analgous to the ephrins, whose existence is well known in the cortex.
-Thalamocortical axons growing from a region in the thalamus with high expression of one of these molecules will interact with a corresponding cortical signalling molecule more strongly than an axon emanating from a region with low expression of the thalamic signal.}
-
-%\MPtwo{We defined two orthogonal, linear gradients oriented in the anterior-posterior and medial-lateral directions and used the coordinates of 41 barreloid centers recreated from \cite{haidarliu_size_2001}, corresponding to the macrovibrissae, to populate 41 pairs of $\gamma$ values.}  (Fig.\,\ref{fig:main}B).
-\MPtwo{In other words, we simply scaled the barreloid coordinates in an anterior-posterior/medial-lateral coordinate system to create the $\gamma$ values in the range [-2,2].
-Note that positive \emph{and} negative $\gamma$ values were generated; the origin of the coordinate system was located at the approximate center of the thalamic barreloid field.
-This prevented an overall drift of the cortical pattern away from the centroid of the barrel field and may explain why coordinate-giving transcription factors and the ephrins are often found as pairs acting in opposing directions.}  \MPtwo{We concentrated on macrovibrissae (excluding microvibrissae) because it is easy to identify their corresponding structures both in the thalamus and in the cortex.
-We also excluded the 5 rhinal whiskers because their cortical representation is in a separate island, distinct, and somewhat removed from the representation of the macrovibirissae.}

### 2020-07-03 paper/briefpaper/brief.sh

@@ -1,5 +1,19 @@
 #!/bin/bash
 
+cat > sebcolour.tex <<EOF
+% Colours for revisions
+\definecolor{colcmnt}        {rgb} {1,      0,      0}      % blue
+\definecolor{colmpfour}      {rgb} {0.220,  0.463,  0.114}  % blue
+\definecolor{colmptwo}       {rgb} {0.235,  0.471,  0.847}  % green
+\definecolor{colmetrics}     {rgb} {0.8353, 0.1255, 0.1255} % red
+\definecolor{colmpthreepred} {rgb} {0.8235, 0.4275, 0.0}    % orange
+\definecolor{colmpsix}       {rgb} {0.8,    0.1,    0.8}    % purple
+\definecolor{colmpfive}      {rgb} {0.3882, 0.0745, 0.7098} % purple2
+\definecolor{colmpone}       {rgb} {0.7098, 0.0745, 0.6431} % pink
+\definecolor{colmpthreepar}  {rgb} {0,      0,      0.3}    % blueblack
+\definecolor{colmpthreesens} {rgb} {0,      0.2,    0}      % greenblack
+EOF
+
 xelatex brief.tex
 bibtex brief.aux
 xelatex brief.tex

### 2020-07-03 paper/briefpaper/brief_review.sh

@@ -1,0 +1,45 @@
+#!/bin/bash
+
+#
+# This script builds two extra versions of the PDF, with differing colour schemes.
+#
+
+# Now the one with all pink changes, for reviewers
+cat > sebcolour.tex <<EOF
+% Colours for reviewers
+\definecolor{colcmnt}        {rgb} {0.7098, 0.0745, 0.6431}
+\definecolor{colmpfour}      {rgb} {0.7098, 0.0745, 0.6431}
+\definecolor{colmptwo}       {rgb} {0.7098, 0.0745, 0.6431}
+\definecolor{colmetrics}     {rgb} {0.7098, 0.0745, 0.6431}
+\definecolor{colmpthreepred} {rgb} {0.7098, 0.0745, 0.6431}
+\definecolor{colmpsix}       {rgb} {0.7098, 0.0745, 0.6431}
+\definecolor{colmpfive}      {rgb} {0.7098, 0.0745, 0.6431}
+\definecolor{colmpone}       {rgb} {0.7098, 0.0745, 0.6431}
+\definecolor{colmpthreepar}  {rgb} {0.7098, 0.0745, 0.6431}
+\definecolor{colmpthreesens} {rgb} {0.7098, 0.0745, 0.6431}
+EOF
+
+xelatex -jobname=barrels_review brief.tex
+bibtex barrels_review.aux
+xelatex -jobname=barrels_review brief.tex
+xelatex -jobname=barrels_review brief.tex
+
+# Now the one with all black text, as final
+cat > sebcolour.tex <<EOF
+% Colours for the proof PDF
+\definecolor{colcmnt}        {rgb} {0, 0, 0}
+\definecolor{colmpfour}      {rgb} {0, 0, 0}
+\definecolor{colmptwo}       {rgb} {0, 0, 0}
+\definecolor{colmetrics}     {rgb} {0, 0, 0}
+\definecolor{colmpthreepred} {rgb} {0, 0, 0}
+\definecolor{colmpsix}       {rgb} {0, 0, 0}
+\definecolor{colmpfive}      {rgb} {0, 0, 0}
+\definecolor{colmpone}       {rgb} {0, 0, 0}
+\definecolor{colmpthreepar}  {rgb} {0, 0, 0}
+\definecolor{colmpthreesens} {rgb} {0, 0, 0}
+EOF
+
+xelatex -jobname=barrels_proof brief.tex
+bibtex barrels_proof.aux
[…]

### 2020-07-03 analysis/plot_FigSensitivity.py
Tweaks to scripts that generate the paper figure graphs

@@ -14,7 +14,7 @@ col = sc.Colour()
 
 # Set plotting defaults
 matplotlib.use('TkAgg') # cleaner likely to work
-fnt = {'family' : 'DejaVu Sans', 'weight' : 'regular', 'size' : 18 }
+fnt = {'family' : 'Arial', 'weight' : 'regular', 'size' : 18 }
 matplotlib.rc('font', **fnt)
 # IMPORTANT for svg output of text as things that can be edited in inkscape
 plt.rcParams['svg.fonttype'] = 'none'
@@ -190,10 +190,10 @@ ax3.set_xticklabels(['0','.2'])
 
 ax1.set_yticks ([0.1, 0.2, 0.3, 0.4])
 ax1.set_yticklabels (['.1', '.2', '.3', '.4'])
-ax2.set_yticks ([0.2, 0.4, 0.6])
-ax2.set_yticklabels (['.2', '.4', '.6'])
-ax3.set_yticks ([0, 0.5, 1.0, 1.5])
-ax3.set_yticklabels (['0', '.5', '1', '1.5'])
+ax2.set_yticks ([0, 0.5, 1.0, 1.5])
+ax2.set_yticklabels (['0', '.5', '1', '1.5'])
+ax3.set_yticks ([0.2, 0.4, 0.6])
+ax3.set_yticklabels (['.2', '.4', '.6'])
 
 
 equal_aspect (ax1)

### 2020-07-03 analysis/plot_trim_areaonly.py

@@ -19,7 +19,7 @@ col = sebcolour.Colour
 import matplotlib.transforms
 
 # Set plotting defaults
-fs = 26
+fs = 32
 fnt = {'family' : 'Arial',
        'weight' : 'regular',
        'size'   : fs}
@@ -88,8 +88,8 @@ l1, = ax1.plot(tdc3[:]['eps_mult'], tdc3[:]['area'], 's-', markersize=main_size,
 ax1.set_xlim([0.5,1.0])
 ax1.set_ylim([0,0.16])
 
-ax1.text (0.55, 0.022, 'C3', fontsize=24, horizontalalignment='left', color=col.black);
-ax1.text (0.67, 0.144, 'neighbours of C3', fontsize=24, horizontalalignment='left', color=col.black);
+ax1.text (0.55, 0.022, 'C3', fontsize=28, horizontalalignment='left', color=col.black);
+ax1.text (0.64, 0.144, 'neighbours of C3', fontsize=28, horizontalalignment='left', color=col.black);
 
 
 ax1.set_xlabel ('$m$')
@@ -101,10 +101,15 @@ for axis in ['top','bottom','left','right']:
   ax1.tick_params(length=ll, width=lw)
 
 # Shift x axis labels down a bit
-dx = 0/72.; dy = -10/72.
-offset = matplotlib.transforms.ScaledTranslation (dx, dy, F1.dpi_scale_trans)
+dx = 0/72.; dy = -14/72.
+offsetdown = matplotlib.transforms.ScaledTranslation (dx, dy, F1.dpi_scale_trans)
 for label in ax1.xaxis.get_majorticklabels():
-    label.set_transform(label.get_transform() + offset)
+    label.set_transform(label.get_transform() + offsetdown)
+
+dx = -8/72.; dy = 0/72.
+offsetleft = matplotlib.transforms.ScaledTranslation (dx, dy, F1.dpi_scale_trans)
+for label in ax1.yaxis.get_majorticklabels():
+    label.set_transform(label.get_transform() + offsetleft)
 
 plt.tight_layout()
 

### 2020-07-03 paper/briefpaper/brief.tex
unify spelling of organized

@@
   eventually yielded a clear Voronoi-like tessellation of topographically
-  organised thalamocortical projections, confirming that barrel maps can
-  self-organise in the absence of pre-specified center points. The
+  organized thalamocortical projections, confirming that barrel maps can
+  self-organize in the absence of pre-specified center points. The
   organization is apparent in a plot of the identity of the projection for
@@
   projection densities, $\omega$. The connection selectivity increased as the
-  barrel map self-organised in the example simulation (Fig.\,\ref{fig:main}D,
+  barrel map self-organized in the example simulation (Fig.\,\ref{fig:main}D,
   grey hexagons), and the selectivity became concentrated in regions

### 2020-07-03 paper/briefpaper/brief.tex
Corrects Fig 3 caption to change column naming

@@
         %
-        \emph{Column 1:} Random noise drawn from a uniform distribution,
+        \emph{Left column:} Random noise drawn from a uniform distribution,
         $(\gamma_{max}-\gamma_{min})\,U(0,\nu_\gamma)$, added
@@
         %
-        \emph{Column 2:} Random noise $(\rho_{max} - \rho_{min})\,U(0, \nu_{\rho})$
+        \emph{Middle column:} Random noise $(\rho_{max} - \rho_{min})\,U(0, \nu_{\rho})$
         was added to each hex element of $\rho_1(\mb{x})$ and $\rho_2(\mb{x})$ and the result
@@
         %
-        \emph{Column 3:} Colour maps show the result of setting the rotational angle
+        \emph{Right column:} Colour maps show the result of setting the rotational angle
         of the linearly varying guidance field, $\rho_1(\mb{x})$, to $\phi_{\rho_1}$,

### 2020-07-03 paper/briefpaper/brief.tex
Some improvements to Fig 3 caption

@@
         with the parameter exploration.
-        %
+        % $a(\mb{x},0)\in(0.2,0.4)$
         \emph{Left column:} Random noise drawn from a uniform distribution,
@@
         parameters of $\rho_2(\mb{x})$
-        unchanged from those used in the simulation in
-        Fig.\,\ref{fig:main}, for which
+        unchanged from those used in the example simulation, for which
         $\phi_{\rho_2}=84$\textdegree and $G_{\rho_2}=1$.
@@
         otherwise take. ii) High magnitude, short lengthscale noise in
-        $\rho(\mb{x})$ leads to `wiggly' boundaries between barrels. iii)
+        $\rho(\mb{x})$ leads to meandering boundaries between barrels. iii)
         Shows the effect of reducing the slope of $\rho_1$ (by a factor of 10)

### 2020-07-03 paper/briefpaper/brief.tex
Move a mis-placed paragraph
Move this short para agin, this time to the start of Materials and Methods

@@
 
+\MPone{We concentrated our simulation work on the representation of the
+  forty-six macrovibrissae that constitute a given barrel field, because their
+  thalamic and cortical correlates are easily identifiable and consistently
+  positioned. We excluded the representation of the microvibrissae to limit
+  the overall complexity of the simulations, as well as the five rhinal
+  whiskers as their representation in S1 is isolated from the main barrel
+  field. Consequently, we set $N=41$.}
+
 The cortical sheet was modelled as a two dimensional hexagonal lattice, which
@@
 
-% for ``end of methods''
-\MPone{We concentrated on the representation of the forty-six macrovibrissae
-  that constitute a given barrel field, because their thalamic and cortical
-  correlates are easily identifiable and consistently positioned. We exclude
-  the representation of the microvibrissae to limit the overall complexity of
-  the simulations, as well as the five rhinal whiskers as their representation
-  in S1 is isolated from the main barrel field. Consequently, we set $N=41$.}
-
 All code required to reproduce these results is available at

### 2020-07-04 paper/briefpaper/brief.tex
Adds a sentence saying what Honda delta is for mouse - Senft and Woolsey - and noting that I anaylsed Zheng data and found that the value is similar for the average rat map

@@
   close correspondence of the simulated barrel field with a Voronoi
-  tesselation, and thus with a biological barrel field. For the tesselation
+  tesselation, and thus with a biological barrel field.
+  (In mouse, $\delta\approx0.054$} \citep{senft_mouse_1991}\metrics{;
+  our analysis of data from} \cite{zheng_signal_2001} \metrics{ indicates that
+  the value for rat is similar). For the tesselation
   that is overlaid on the real barrel field in Fig.\,\ref{fig:main}A,
   $\delta=0.025$, and a reduction in $\delta$ in the example simulation over
-  time confirmed that an equivalent `good' Voronoi pattern} (see
-\citealp{senft_mouse_1991}) \metrics{can emerge within $\approx$~20000
+  time confirmed that an equivalent `good' Voronoi pattern can emerge
+  within $\approx$~20000
   iterations (Fig.\,\ref{fig:main}D, red circles). Second, we devised a

### 2020-07-04 paper/briefpaper/brief.sh
Make cmnts blue as intended

@@ -2,7 +2,7 @@
 
 cat > sebcolour.tex <<EOF
 % Colours for revisions
-\definecolor{colcmnt}        {rgb} {1,      0,      0}      % blue
+\definecolor{colcmnt}        {rgb} {0,      0,      1}      % blue
 \definecolor{colmpfour}      {rgb} {0.220,  0.463,  0.114}  % blue
 \definecolor{colmptwo}       {rgb} {0.235,  0.471,  0.847}  % green
 \definecolor{colmetrics}     {rgb} {0.8353, 0.1255, 0.1255} % red

### 2020-07-04 paper/briefpaper/brief.tex
Adds note about checking on contours in movie (need to remake the movie) and also a sentence noting how the label locns are determined, which one reviewer mentioned

@@
 Movie corresponding to Fig.\,\ref{fig:main}C in the main paper. Simulation parameters were
-$N=41$, $\alpha=3.6$, $\beta=16.67$, $k=3$, $D=0.5$, $\gamma\in\pm 2$,
-$\epsilon=1.2$ and $\delta{t}=0.0001$. Colours indicate the thalamic
+\MPtwo{$N=41$, $\alpha=3.6$, $\beta=16.67$, $k=3$, $D=0.5$, $\gamma\in\pm 2$,
+$\epsilon=1.2$ and $\delta{t}=0.0001$}. Colours indicate the thalamic
 projection for which the connection density is maximal, black lines delineate
-boundaries, and overlaid contours show $c>0.5$. The final frame in the movie
+boundaries, and overlaid contours show $c>0.5$ \textbf{FIXME: Will need to
+  check this statement after regenerating movie}. \cmnt{Barrel labels are
+  always located at the centroid of their field(s).} The final frame in the movie
 is step 30000 of the simulation.

### 2020-07-06 paper/briefpaper/brief.tex
Finished editting (I think) but note final fixme in movie caption still

@@
 %
-% These timings, from erzurumlu_development_2012 originate from
-% sehara_neuronal_2011, where they relate to mouse.
-\mpsix{In the mouse, axons from the trigeminal ganglion arrive in the principal sensory
-  nucleus of the trigeminal nerve (PrV) at E12, then axons from the PrV arrive
-  in the ventroposteromedial nucleus of the thalamus (VPM) at E17, then axons
-  from the VPM arrive in the cortical plate at E18/P0. Distinct whisker-related
-  clusters then become apparent in the PrV at P0-P1, in the VPM at P2-P3, and
-  in the cortex at P3-P5} \citep{erzurumlu_development_2012,sehara_neuronal_2011}.
-%
 In the plane tangential to
@@
 \citep{shimogori_fibroblast_2005,bishop_regulation_2000}, which are expressed
-in gradients \mpfour{across the full cortical sheet} that mark orthogonal axes
+in gradients \mpfour{spanning the cortical sheet} that mark orthogonal axes
 and can be manipulated to stretch,
@@
 
-The barrel boundaries form a Voronoi tessellation across the cortical sheet
+The barrel boundaries form a Voronoi tessellation
 \citep{senft_mouse_1991} (Fig.\,\ref{fig:main}A), suggesting that barreloid topology is
@@
 bundle, \emph{prior} to barreloid formation \citep{agmon_organized_1993}.
+\mpsix{In mice, axons from the trigeminal ganglion arrive in the principal sensory
+  nucleus of the trigeminal nerve (PrV) at E12, then axons from the PrV arrive
+  in the ventroposteromedial nucleus of the thalamus (VPM) at E17, then axons
+  from the VPM arrive in the cortical plate at E18/P0. Distinct whisker-related
+  clusters then become apparent in the PrV at P0-P1, in the VPM at P2-P3, and
+  in the cortex at P3-P5} \citep{erzurumlu_development_2012,sehara_neuronal_2011}.
 
@@
 adjacent cells only (see \emph{Materials \& Methods}). The second term on the
-right quantifies the coupling between axon branching and synaptogenesis. Here
+right represents the coupling between axon branching and synaptogenesis. Here
   $\chi_i=0$ is a placeholder.
@@
     \includegraphics[width=\linewidth]{./Fig1.png}
-    \caption{\textbf{A} Left shows a cytochrome oxidase stain obtained from rat S1
+    \caption{The emergence of whisker barrels. \textbf{A} Left shows a cytochrome oxidase stain obtained from rat S1
       by \cite{zheng_signal_2001}, with black lines to delineate barrels and to
@@
       the locations of putative barreloid centers (reconstructed from
-      \citealp{haidarliu_size_2001}, \MPone{Fig.\,5b}). \textbf{C} Simulation results for parameters
+      \citealp{haidarliu_size_2001}, their \MPone{Fig.\,5b}). \textbf{C} Results for the example simulation, with parameters
       \MPtwo{$N=41$, $\alpha=3.6$, $\beta=16.67$, $k=3$, $D=0.5$, $\gamma\in\pm 2$,
@@
       delineate boundaries (see Movie
-      S1). \textbf{D} Red dots show the Honda metric, \metrics{$\delta(t)$}, obtained from \metrics{the}
-      simulation approaching that obtained from \metrics{real} barrels in \textbf{A} (dotted
-      line); black squares \metrics{provide a} measure \metrics{of} the correspondence
-      between the real and
-      simulated barrel shapes, $\eta(t)$ (the \metrics{pattern quality metric---see
-        Materials \& Methods; units mm$^3$});
-      \metrics{grey hexagons \metrics{provide a measure of how selectively each
-          cortical site is innervated};
+      S1). \textbf{D} Red dots show the \emph{Honda}--\metrics{$\delta(t)$} metric obtained from \metrics{the}
+      simulation approaching that obtained from the \metrics{real} barrels in \textbf{A} (dotted
+      line); black squares show the \emph{pattern difference} metric $\eta(t)$, and reveal the emergence of a correspondence between the real and simulated barrel shapes (units mm$^3$); \metrics{grey hexagons show \metrics{how selectively each cortical site is innervated};
         $\omega(t) = \oiint_{S} \mu(\mb{x},t) \mathrm{d}S$, where
@@
         develops to reveal an alignment with the emergent barrel boundary
-        shapes}. \metrics{The colour map
-        indicates the values of $\mu(\mb{x})$. Overlaid
+        shapes}. \metrics{Greyscale colour indicates values of $\mu(\mb{x})$ and overlaid
         contours show $c_i > 0.95 \mathrm{max}_i\big(c_i(\mb{x})\big)$.} All scale bars 1\,mm.}
@@
 \mpfive{We reasoned that an arbitrary number of distinct field locations may
-  be specified using a minimum of two guidance fields, if the tendency for the
-  density of each projection to concentrate is appropriately balanced against
+  be determined by a minimum of two guidance fields, if a tendency for the
+  density of projections to concentrate is appropriately balanced with
   the tendency of projections that interact more strongly with the guidance
@@
   development have not been fully characterised, the presence of complementary
-  thalamic and cortical molecular guidance gradients has been well established
+  thalamic and cortical guidance gradients has been well established
   experimentally. In particular, the EphA4 receptor and its ligand ephrin-A5
@@
 \citealp{gao_regulation_1998,dufour_area_2003,vanderhaeghen_developmental_2004,speer_grading_2005,torii_role_2013}).
-\MPone{We assume that such patterning arises because relative strengths of
+\MPone{We assume that such patterning arises because the relative strengths of
   interaction with guidance molecules (e.g., ephrin-A5) in the cortex are
@@
 
-A cortical boundary enclosing \cmnt{the} 41 \cmnt{macrovibrissae} barrels was
+A cortical boundary enclosing barrels for 41 \cmnt{macrovibrissae} was
 traced from a cytochrome oxidase stain from \cite{zheng_signal_2001}
@@
   which the connection density is maximal at each simulated cortical location,
-  as shown in Fig.\,\ref{fig:main}C. Parameters for the example simulation
-  shown in Fig.\,\ref{fig:main}C (see also Movie S1) were obtained by
-  conducting a full parameter sweep and choosing a combination of parameter
-  values ($\alpha=3.6$, $\beta=16.67$, $k=3$, $D=0.5$, $\epsilon=1.2$) that
+  as shown in Fig.\,\ref{fig:main}C. Parameter values for this example simulation
+  (see also Movie S1) were obtained by
+  conducting a full parameter sweep and choosing a combination ($\alpha=3.6$, $\beta=16.67$, $k=3$, $D=0.5$, $\epsilon=1.2$) that
   scored well against the following three measures.}
@@
     \includegraphics[width=\linewidth]{./Fig2.png}
-    \caption{\MPthreePar{\textbf{A} Exploring the parameter space of the
-        model given in Eqs.\,\ref{eq:dc}--\ref{eq:comp}. Graphs show the three
-        pattern metrics for $t=30000$ steps as colourmaps: Honda $\delta$ (top
-        row) shows conformance to a Voronoi pattern; $\eta$ (middle row) gives
-        the simulated/experimental barrel area difference and $\omega$ (bottom
-        row) is a measure of selectivity of the connection densities for the
-        41 thalamocortical axon types. The colour maps are chosen so that for
-        each row, lighter (orange and yellow) colours indicate desirable
-        properties for the simulations: Lower values of $\delta$ are yellow,
-        indicating good Voroni patterns; Lower values of $\eta$ indicate good
-        pattern matching; Higher values of $\omega$ indicate greater
-        selectivity. White squares indicate that the simulation was not stable
-        for that parameter set.
+    \caption{\MPthreePar{Exploring the parameter space. \textbf{A} Colour indicates the quality of the pattern at $t=30000$ steps, against three measures: Low values of \emph{Honda}--$\delta$ (top
+        row) suggest a Voronoi-like pattern of fields.
+The \emph{pattern difference}, $\eta$ (middle row) measures the difference in the area and arrangement of barrels in real and simulated fields.
+The \emph{selectivity}, $\omega$ (bottom
+        row), measures the specificity with which the cortical sheet is innervated.
+Colour maps are chosen so that lighter (orange and yellow) colours indicate higher quality patterns against each measure.
+White squares indicate combinations of parameters for which simulations were numerically unstable.
 %
-        The parameter space explored is three dimensional with the parameter
-        ratio $\alpha/\beta$ varied along the columns of graphs, the
-        competition parameter $\epsilon$ varied on the $y$-axes and the
-        diffusion constant, $D$ varied along the $x$-axes. An asterisk (*)
-        marks the parameter set displayed in Fig.\,\ref{fig:main}. Boxes i),
+        The parameter space explored is three dimensional with the ratio $\alpha/\beta$ varying between plots, and the
+        competition parameter $\epsilon$ and the diffusion constant $D$ varying within plots. An asterisk (*)
+        marks the parameters used in Fig.\,\ref{fig:main}. Boxes i),
         ii)~and iii)~mark parameter sets for which corresponding patterns are
@@
 %
-        \textbf{B} The effect of varying the diffusion constant,
-        $D$. i)~Increasing $D$ above the optimum value gives an expansion of
-        the outer barrels and a corresponding compression of the barrels
-        within. ii)~Conversely, an expansion of the central barrels and
-        compression of those near the boundary results from reducing $D$ from
-        its optimum. iii) Reducing the diffusion constant is equivalent to
-        increasing the size of the domain $S$ and so this simulation, in which
-        $D$ had the lowest stable value, can be thought to represent an animal
-        with a large cortical area but with unchanged cortical complexity. A
-        large proportion of the domain is occupied by the projections with
-        intermediate interaction parameters; those with strong interaction
-        parameters are compressed around the edge of the domain. A barrel
-        pattern fails to form.}}
+        \textbf{B} Varying the diffusion constant $D$ generates qualitatively different patterns. Higher values 
+        cause expansions of the peripheral barrels and a corresponding compression of the inner barrels (i).
+Lower values instead cause an expansion of the central barrels and compression of the peripheral barrels (ii).
+Further reducing the rate of diffusion (iii), which is equivalent to increasing the size of the domain and hence simulating development in an animal with a larger cortex, causes a large area to be occupied by projections with intermediate interaction parameters; those with strong interaction parameters are compressed around the edge of the domain, and consequently a barrel pattern fails to form.}}
     \label{fig:paramsweep}
@@
     \caption{
-      \MPthreeSens{\textbf{A} Sensitivity of the system pattern
-        metrics $\delta(t)$ (top row), $\omega(t)$ (middle row) and $\eta(t)$
-        (bottom row), evaluated at $t=30000$ steps,
-        to three classes of perturbation. $y$-axes and colour scales have identical ranges
-        to the colour scales in Fig.\,\ref{fig:paramsweep}, to ease comparison
-        with the parameter exploration.
-        % $a(\mb{x},0)\in(0.2,0.4)$
-        \emph{Left column:} Random noise drawn from a uniform distribution,
-        $(\gamma_{max}-\gamma_{min})\,U(0,\nu_\gamma)$, added
-        to the values for the interaction parameters, $\gamma_{i,j}$ used in
+      \MPthreeSens{Sensitivity analysis.
+\textbf{A} Metrics of map quality $\delta(t)$ (top row), $\omega(t)$ (middle row) and $\eta(t)$ (bottom row), were evaluated at $t=30000$ steps.
+The $y$-axes and colour scales have identical ranges
+        to the colour scales in Fig.\,\ref{fig:paramsweep}, for easy comparison. \emph{Left column:} The effect of adding noise drawn from a uniform distribution,
+        $(\gamma_{max}-\gamma_{min})\,U(0,\nu_\gamma)$,
+        to the values of the interaction parameters, $\gamma$ used in
         Fig.\,\ref{fig:main}.
         %
-        \emph{Middle column:} Random noise $(\rho_{max} - \rho_{min})\,U(0, \nu_{\rho})$
-        was added to each hex element of $\rho_1(\mb{x})$ and $\rho_2(\mb{x})$ and the result
+        \emph{Middle column:} The effect of changing the magnitude and the length scale of
+        noise applied to the guidance fields. Uniform random noise $(\rho_{max} - \rho_{min})\,U(0, \nu_{\rho})$ was added
+        to each (hexagonal) element of $\rho_1(\mb{x})$ and $\rho_2(\mb{x})$ and the result
         was smoothed by convolution with a symmetric 2D
-        Gaussian kernel of width $\sigma_\rho$. Colour maps in this column
-        thus show the effect of changing the magnitude and the length scale of
-        noise in the guidance fields.
+        Gaussian kernel of width $\sigma_\rho$. 
         %
-        \emph{Right column:} Colour maps show the result of setting the rotational angle
+        \emph{Right column:} The effect of setting the rotational angle
         of the linearly varying guidance field, $\rho_1(\mb{x})$, to $\phi_{\rho_1}$,
@@
         the patterns. i) The effect of significant interaction parameter noise
-        causes barrels to become swapped from the positions they would
-        otherwise take. ii) High magnitude, short lengthscale noise in
-        $\rho(\mb{x})$ leads to meandering boundaries between barrels. iii)
-        Shows the effect of reducing the slope of $\rho_1$ (by a factor of 10)
-        which reduces the strength of the axons' interaction with the guidance
-        field in the left-right sense. Barrel rows B, C and D become `crushed'
-        down the center line, and edge barrels dominate. iv) The effect of
-        rotating $\rho_1$ by 20\textdegree~is a relatively mild distortion of
-        the pattern in the anticlockwise sense.
+        results in topological defects. ii) High magnitude, short lengthscale noise in
+        $\rho(\mb{x})$ leads to non-straight edges between adjacent barrels. iii)
+ Reducing the slope of gradient $\rho_1$ (by a factor of 10)
+causes barrel rows B, C and D to become `crushed'
+        down the center line, and edge barrels to dominate.
+iv) Rotating $\rho_1$ by 20\textdegree~causes a slight distortion of the pattern, resulting in an overall anticlockwise rotation of the field locations.
       }
@@
 \MPthreePred{To further investigate the interplay of genetic and intrinsic
-  mechanisms we simulated two well known experimental manipulations of the
-  whisker barrel system. First, we simulated a seminal barrel duplication
-  experimental paradigm}
+  mechanisms we simulated two well known experimental manipulations of 
+barrel development. First, we simulated a seminal barrel duplication
+  paradigm}
 \citep{shimogori_fibroblast_2005,assimacopoulos_fibroblast_2012}
@@
 \citep{crossley_mouse_1995}, \MPthreePred{is ectopically expressed (by
-  electroporation) also at the posterior pole of a cortical
-  hemisphere. Because Fgf8 sets off a chain of gene interactions involved in
-  controlling the graded expression of ephrin and cadherin molecules, its
-  misexpression at the posterior pole could be expected to mirror those
-  gradients which usually vary along the rostrocaudal axis. We assume that
+  electroporation) also at the posterior pole. We assume that
   this results in a mirror of the primary barrel cortex boundary along the
-  rostrocaudal axis, to match the experimental results of}
-\cite{assimacopoulos_fibroblast_2012}. \MPthreePred{We also assume that a
-  further consequence of the posterior Fgf8 electroporation is the mirrored
-  anterior-posterior guidance gradient $\rho_1$ (Fig.\,\ref{fig:fgf8}A). With
-  the modified $\rho_1$ guidance gradient the simulation was run with $N=41$
-  thalamocortical axon types, using the parameters of the example
-  simulation. The result, after 30000 iterations, was two mirror-symmetrical
+  rostrocaudal axis}
+\cite{assimacopoulos_fibroblast_2012} \MPthreePred{and a the mirroring of the
+  anterior-posterior guidance gradient $\rho_1$ at the border between them (Fig.\,\ref{fig:fgf8}A). 
+  The result after 30000 iterations, and otherwise using the parameters of the example simulation, was two mirror-symmetrical
   barrel fields comprising $2N$ barrels (Fig.\,\ref{fig:fgf8}B), consistent
@@
   \includegraphics[width=\linewidth]{./Fig4.png} \captionof{figure}{
-    \MPthreePred{Two simulated whisker barrel manipulations.  Guidance fields
+    \MPthreePred{Simulating altered barrel development.  Guidance fields
       (\textbf{A}) and emergent barrel pattern (\textbf{B}) in a Fgf8
@@
       $\rho_1$ from Fig.\,\ref{fig:main}A at the join of the original boundary
-      with its mirror. The model parameters match those in
-      Fig.\,\ref{fig:main}C.  Scale bars are 1\,mm. \textbf{C} Simulating
-      whisker trimming by a reduction of the competition parameter,
-      $\epsilon$, in the expression $\chi_i$ (Eq.\,\ref{eq:comp}) for the C3
-      thalamocortical projection. For C3, $\epsilon$ is replaced by
-      $m\,\epsilon$, where $m$ is a scalar multiplier. For the rest of the
-      projections, $\chi_i$ is left unchanged and all other parameters are
-      identical to those reported in Fig.\,\ref{fig:main}C. The pattern formed
-      after 30000 steps with $m=0.86$. As its axon branching competes weakly
-      with its neighbours, field C3 is reduced to 65\% of its original size,
+      with its mirror. All other model parameters match those in
+      Fig.\,\ref{fig:main}C. \textbf{C} Simulating
+      whisker trimming by reducing the competitiveness,
+      $\epsilon$, of the one projection. For C3 only, $\epsilon$ was mutiplied by
+      $m\in(0,1)$. The pattern shown is that formed
+      after 30000 steps with $m=0.86$, which reduces the size of the C3 field to 65\% of its original size,
       matching the average barrel area reduction observed by }
-    \citealp{kossut_effects_1992}. \MPthreePred{\textbf{D} Individual barrel
-      area (mm$^2$) plotted vs.~$m$ for C3 (black squares) shows that as $m$
-      is reduced, its area goes down, disappearing for $m\approx0.5$, whereas
-      the mean area of neighbouring barrels B3, C2, C4, D2 \& D3 (grey
+    \citealp{kossut_effects_1992}. \MPthreePred{\textbf{D} The area of the C3 barrel
+reduces as $m$ is reduced, whereas
+      the mean area of neighbouring barrels (B3, C2, C4, D2 \& D3, grey
       circles) increases. The dotted grey line indicates $m=0.86$ for
@@
 
-\MPthreePred{To investigate the response of the model to environmental
-  manipulation, we simulated a whisker deprivation experiment. In the first
-  few days after birth, the upper region of the rat's cortex develops into
-  distinct layers I--IV (layers V and VI are already distinct at
-  birth). During this critical period, removal of the mystacial whiskers by
-  electrocauterization, plucking or trimming leads to observable changes in
-  brain structures, including the barrel field in cortical layer
-  IV~\citep{jeanmonod_mouse_1981}. Amoungst other changes, barrels associated
-  with deprived whiskers have a smaller area. This is found whether the animal
-  is deprived of an individual whisker~\citep{kossut_effects_1992}, or an
-  entire row~\citep{loos_somatosensory_1973}, and demonstrates that
-  thalamocortical activity contributes to the developmental process. We
-  simulated the trimming of the individual whisker C3 during this critical
-  period by reducing the amount by which its thalamocortical projections
-  compete for branching. This was achieved by reducing the value of $\epsilon$
-  in Eq.\,\ref{eq:comp} for the C3 projection, leaving other simulation
-  parameters unchanged. The reduced competition for branching limited the area
-  over which the C3 projection could make connections and its resulting field
-  size was diminished (Fig.\,\ref{fig:fgf8}C). Emergent fields representing
-  the neighbours of the C3 projection, whose ability to compete for branching
-  was unchanged, increased in size. The C3 barrel disappeared altogether if
-  its competition parameter was reduced below a half of the original value. A
-  reduction in the area of barrel C3, comparable that induced by
-  \citealp{kossut_effects_1992} (65\%) was obtained in simulation when the
-  competitiveness of projection C3, $\epsilon$, was reduced to 86\%, as
-  indicated in Fig.\,\ref{fig:fgf8}D.}
+\MPthreePred{Finally, to investigate the response of the model to environmental
+  manipulation, we simulated a whisker deprivation experiment. In a critical period comprising the first postnatal days, removal of the whiskers by
+  electrocauterization, plucking, or trimming leads to observable changes in
+  brain structures, including the barrel field~\citep{jeanmonod_mouse_1981}.
+Amongst other changes, deprivation of individual whiskers, or an entire row of whiskers leads to smaller barrels~\citep{kossut_effects_1992,loos_somatosensory_1973}.
+We
+  simulated trimming of the individual whisker C3 during the critical
+  period by reducing the competitiveness of the C3 thalamocortical projection, $\epsilon$. As a result, the corresponding field
+  size was smaller (Fig.\,\ref{fig:fgf8}C), and the size of the fields representing
+  the neighbours of C3 increased in size. A reduction in area to 65\%, comparable to that induced by
+  \citealp{kossut_effects_1992}, was obtained in simulation when $\epsilon$ was reduced to 86\% (Fig.\,\ref{fig:fgf8}D.), and the C3 barrel disappeared altogether when $\epsilon$ was less than half of its original value.}
 
-\MPthreePred{Together with the results of the misexpression experiment, the
+\MPthreePred{Together with the results of simulated misexpression, the
   consistency of the simulated whisker trimming results with those of the
-  original studies demonstrates how the model can be used to test specific
-  hypotheses about the contribution of intrinsic and extrinsic factors to the
+  original studies demonstrates how the model can be used to investigate the contribution of intrinsic and extrinsic factors to the
   development of cortical fields.}
@@
 cortical space, as represented by $\chi$, with an implicit requirement for a
-self/other identifier amoungst projections. This latter form of competition may account for the absence of barrels in rodents with larger brains, such as
+self/other identifier amongst projections. This latter form of competition may account for the absence of barrels in rodents with larger brains, such as
 capybara, for which competition for space is presumably weaker
@@
   the same boundary condition as used to represent the true edges of the
-  barrel field is necessary to ensure the stability of the simulation, though
-  we acknowledge that this region of the boundary is enforced primarily to
+  barrel field was necessary to ensure the stability of the simulation, though
+  we acknowledge that this region of the boundary was enforced primarily to
   keep the number of simulated projections, and hence the overall
-  computational complexity of the simulation, manageable (simulating each
-  additional projection introduces an extra 13030 new dynamical
-  variables).}
+  computational complexity of the simulation, manageable (simulating an extra
+  projection introduces 13030 new dynamical variables).}
 
@@
 
-\MPone{We concentrated our simulation work on the representation of the
-  forty-six macrovibrissae that constitute a given barrel field, because their
+\MPone{We concentrated on the representation of the
+  forty-one macrovibrissae that constitute a given barrel field, because their
   thalamic and cortical correlates are easily identifiable and consistently
-  positioned. We excluded the representation of the microvibrissae to limit
-  the overall complexity of the simulations, as well as the five rhinal
-  whiskers as their representation in S1 is isolated from the main barrel
-  field. Consequently, we set $N=41$.}
+  located, excluding the five rhinal whiskers as their cortical representation is isolated from the main barrel
+  field. We excluded the representation of the microvibrissae to limit
+  the overall complexity of the simulations.}
 
@@
 flux of $\mb{J}$ across the boundary is 0. Then, $\mb{g}_i(\mb{x})$ can be
-tailored so that it, and its normal derivative approach 0 at the boundary,
+tailored so that it, and its normal derivative, approach 0 at the boundary,
 ensuring that the second and third terms of Eq.\,\ref{eq:divJExpanded} also
@@
 from $\mb{x}$ to the boundary, \MPtwo{of the form $1/[1 +
-  \exp(100(d_f-d_b))]$, where $d_f=0.1$\,mm is the boundary fall-off distance,
-  chosen to be approximately $3d$}.
+  \exp(100(d_f-d_b))]$, where $d_f=0.1$\,mm $\approx 3d$ is the boundary fall-off distance}.
 
-\metrics{The pattern-matching metric, $\eta$, incorporates information about i)
-  the differences in the areas of the simulated and experimental barrels
-  ($\mathcal{A}_i^{\mathrm{sim}}$ and $\mathcal{A}_i^{\mathrm{exp}}$); and ii)
-  the information provided by `adjacency vectors', $\mathcal{V}_i$ (one for
-  each barrel) whose $N$ elements contain the length of the barrel boundary
-  which adjoins each other barrel in the pattern. The value of
-  $\mathcal{V}_{i,j}$ (the $j^{\mathrm{th}}$ element of $\mathcal{V}_i$) is
-  the length of the boundaries between regions of type $i$ and regions of type
-  $j$. In a well-formed barrel pattern, $\mathcal{V}_{i}$ is a sparse vector
-  and note that $\mathcal{V}_{i,i} = 0, \forall i$. $\mathcal{V}_{i}$ can be
-  computed for the experimental and simulated barrel patterns and compared. A
-  dimensionless measure of the pattern's arrangement and neighbour relations
-  can be obtained from the scalar product of the simulated and experimental
+\metrics{The \emph{pattern difference} metric, $\eta$, incorporates information about
+  the differences in the areas of simulated and experimentally determined (real) barrels
+  ($\mathcal{A}_i^{\mathrm{sim}}$ and $\mathcal{A}_i^{\mathrm{exp}}$), as well as
+ information contained in the `adjacency vector' for each barrel, $\mathcal{V}_i$, the $j$-th element of which is the length of the border between barrel $i$ and $j$.
+For a well-formed barrel pattern, $\mathcal{V}_{i}$ is a sparse vector.
+A
+  dimensionless quantity can be obtained from the scalar product of the simulated and experimental
   adjacency vectors:
   $\frac{1}{N} \sum_i \frac { \mathcal{V}_{i}^{\mathrm{sim}}}
-  {b_{i}^{\mathrm{sim}} } \vcdot \frac{\mathcal{V}_{i}^{\mathrm{exp}}}{b_{i}^{\mathrm{exp}}}$, where
-  $b_i^{\mathrm{exp}}$($b_i^{\mathrm{sim}}$) is the total length of the boundary of
-  experimental(simulated) barrel $i$. This measure is low early in the simulation, when the regions do
-  not have the neighbour relations of the experimental pattern
-  (Fig.\,\ref{fig:main}C, $t=1000$) and maximal when the neighbour relations
-  have become established ($t=10000$). It does not distinguish between
-  $t=10000$ and $t=30000$.}
+  {b_{i}^{\mathrm{sim}} } \vcdot \frac{\mathcal{V}_{i}^{\mathrm{exp}}}{b_{i}^{\mathrm{exp}}}$, where e.g., 
+  $b_i^{\mathrm{exp}}$, is the total length of the border around 
+real barrel $i$. This quantity is small when the fields that form in simulation have dissimilar neighbour relations to those of the real barrels
+  (e.g., Fig.\,\ref{fig:main}C, $t=1000$), and maximal for a precise topological map ($t=10000$).}
 %
-\metrics{A second way to analyse the adjacency vectors is to consider
-  the mean magnitude of the difference between the simulated and experimental
+\metrics{A second comparison considers the mean magnitude of the difference between the simulated and experimental
   vectors: $\frac{1}{N} \sum_i \norm{ \mathcal{V}_{i}^{\mathrm{sim}} -
     \mathcal{V}_{i}^{\mathrm{exp}} } $. This tends to 0 for a perfect
-  match and can separate patterns with straight boundaries from those with
-  noise on the boundary shape (For an example of `noisy edges', see Fig.\,\ref{fig:sens}B\,ii).}
+  match and can separate patterns with straight boundaries from those with `noisy edges' (as in Fig.\,\ref{fig:sens}B\,ii).}
 %
-\metrics{We combined all of these measures into a single pattern quality metric:}
+\metrics{We combined these terms into a single metric:}
 %

### 2020-07-08 paper/briefpaper/brief.tex
some typos

@@
 and $c_i$ grow are reciprocally coupled. Extending the original 1D model to
-simulate arealization on a 2D cortical sheet, we use $a_i(\mb{x},t)$ and
+simulate arealisation on a 2D cortical sheet, we use $a_i(\mb{x},t)$ and
 $c_i(\mb{x},t)$, and model synaptogenesis as
@@
   rostrocaudal axis}
-\cite{assimacopoulos_fibroblast_2012} \MPthreePred{and a the mirroring of the
+\cite{assimacopoulos_fibroblast_2012} \MPthreePred{and a mirroring of the
   anterior-posterior guidance gradient $\rho_1$ at the border between them (Fig.\,\ref{fig:fgf8}A). 

### 2020-07-09 paper/briefpaper/brief.tex
Correctly refer to fraction of connections and point out why we can also use the language 'connection density'

@@
 extrinsic signalling gradients can constrain the emergence of distinct fields
-from intrinsic cortical dynamics. Their model defines how the density of
-connections $c(x,t)$ and axon branches $a(x,t)$ interact at time $t$, along a
+from intrinsic cortical dynamics. Their model defines how the \cmnt{fraction of
+occupied synapses} $c_i(x,t)$ and \cmnt{the density of} axon branches $a_i(x,t)$ interact at time $t$, along a
 1D anterior-posterior axis $x$, for $N$ thalamocortical projections indexed by
@@
 %
-Accordingly, where the total density of synaptic connections sums to one,
-connections decay at rate $\alpha$. Otherwise the connection density increases
+Accordingly, where the total \cmnt{fraction} of synaptic connections sums to one,
+connections decay at rate $\alpha$. Otherwise, \cmnt{$c_i(\mb{x},t)$} increases
 non-linearly ($k>1$) with the density of axon branching. Axon branching is
@@
 simulation requires cells on the cortical sheet to communicate with immedately
-adjacent cells only (see \emph{Materials \& Methods}). The second term on the
-right represents the coupling between axon branching and synaptogenesis. Here
-  $\chi_i=0$ is a placeholder.
+adjacent cells only (see \emph{Materials \& Methods}). Here $\chi_i=0$ is a
+placeholder. The second term on the right represents the coupling between axon
+branching and synaptogenesis \cmnt{and implies an assumption of spatially
+  homogeneous synaptic density which makes it possible to relate the quantity
+  $c_i$ to connection density}.
 

### 2020-07-09 paper/briefpaper/brief_proof.sh
Now each .sh script builds only one version of the manuscript

@@ -1,0 +1,24 @@
+#!/bin/bash
+
+#
+# This script builds the proof versions of the PDF: all black.
+#
+
+cat > sebcolour.tex <<EOF
+% Colours for the proof PDF
+\definecolor{colcmnt}        {rgb} {0, 0, 0}
+\definecolor{colmpfour}      {rgb} {0, 0, 0}
+\definecolor{colmptwo}       {rgb} {0, 0, 0}
+\definecolor{colmetrics}     {rgb} {0, 0, 0}
+\definecolor{colmpthreepred} {rgb} {0, 0, 0}
+\definecolor{colmpsix}       {rgb} {0, 0, 0}
+\definecolor{colmpfive}      {rgb} {0, 0, 0}
+\definecolor{colmpone}       {rgb} {0, 0, 0}
+\definecolor{colmpthreepar}  {rgb} {0, 0, 0}
+\definecolor{colmpthreesens} {rgb} {0, 0, 0}
+EOF
+
+xelatex -jobname=barrels_proof brief.tex
+bibtex barrels_proof.aux
+xelatex -jobname=barrels_proof brief.tex
+xelatex -jobname=barrels_proof brief.tex

### 2020-07-09 paper/briefpaper/brief_review.sh

@@ -1,9 +1,5 @@
 #!/bin/bash
 
-#
-# This script builds two extra versions of the PDF, with differing colour schemes.
-#
-
 # Now the one with all pink changes, for reviewers
 cat > sebcolour.tex <<EOF
 % Colours for reviewers
@@ -23,23 +19,3 @@ xelatex -jobname=barrels_review brief.tex
 bibtex barrels_review.aux
 xelatex -jobname=barrels_review brief.tex
 xelatex -jobname=barrels_review brief.tex
-
-# Now the one with all black text, as final
-cat > sebcolour.tex <<EOF
-% Colours for the proof PDF
-\definecolor{colcmnt}        {rgb} {0, 0, 0}
-\definecolor{colmpfour}      {rgb} {0, 0, 0}
-\definecolor{colmptwo}       {rgb} {0, 0, 0}
-\definecolor{colmetrics}     {rgb} {0, 0, 0}
-\definecolor{colmpthreepred} {rgb} {0, 0, 0}
-\definecolor{colmpsix}       {rgb} {0, 0, 0}
-\definecolor{colmpfive}      {rgb} {0, 0, 0}
-\definecolor{colmpone}       {rgb} {0, 0, 0}
-\definecolor{colmpthreepar}  {rgb} {0, 0, 0}
-\definecolor{colmpthreesens} {rgb} {0, 0, 0}
-EOF
-
-xelatex -jobname=barrels_proof brief.tex
-bibtex barrels_proof.aux
-xelatex -jobname=barrels_proof brief.tex
-xelatex -jobname=barrels_proof brief.tex

### 2020-07-09 paper/briefpaper/brief.tex
Removes trailing whitespace on a couple of lines, otherwise no changes.
Latex tidying (line lengths)

@@
 
-The barrel boundaries form a Voronoi tessellation
-\citep{senft_mouse_1991} (Fig.\,\ref{fig:main}A), suggesting that barreloid topology is
-preserved in the projection of thalamocortical axons into the cortex, and that
-a barrel forms by lateral axon branching from an initial center-point that
-ceases upon contact with axons branching from adjacent centers.  However, the
-assumption of pre-arranged center-points is difficult to resolve with the
-observation that axons arrive in the cortical plate as an undifferentiated
-bundle, \emph{prior} to barreloid formation \citep{agmon_organized_1993}.
-\mpsix{In mice, axons from the trigeminal ganglion arrive in the principal sensory
+The barrel boundaries form a Voronoi tessellation \citep{senft_mouse_1991}
+(Fig.\,\ref{fig:main}A), suggesting that barreloid topology is preserved in
+the projection of thalamocortical axons into the cortex, and that a barrel
+forms by lateral axon branching from an initial center-point that ceases upon
+contact with axons branching from adjacent centers.  However, the assumption
+of pre-arranged center-points is difficult to resolve with the observation
+that axons arrive in the cortical plate as an undifferentiated bundle,
+\emph{prior} to barreloid formation \citep{agmon_organized_1993}.  \mpsix{In
+  mice, axons from the trigeminal ganglion arrive in the principal sensory
   nucleus of the trigeminal nerve (PrV) at E12, then axons from the PrV arrive
   in the ventroposteromedial nucleus of the thalamus (VPM) at E17, then axons
-  from the VPM arrive in the cortical plate at E18/P0. Distinct whisker-related
-  clusters then become apparent in the PrV at P0-P1, in the VPM at P2-P3, and
-  in the cortex at P3-P5} \citep{erzurumlu_development_2012,sehara_neuronal_2011}.
+  from the VPM arrive in the cortical plate at E18/P0. Distinct
+  whisker-related clusters then become apparent in the PrV at P0-P1, in the
+  VPM at P2-P3, and in the cortex at P3-P5}
+\citep{erzurumlu_development_2012,sehara_neuronal_2011}.
 
@@
 extrinsic signalling gradients can constrain the emergence of distinct fields
-from intrinsic cortical dynamics. Their model defines how the \cmnt{fraction of
-occupied synapses} $c_i(x,t)$ and \cmnt{the density of} axon branches $a_i(x,t)$ interact at time $t$, along a
-1D anterior-posterior axis $x$, for $N$ thalamocortical projections indexed by
-$i$. The model was derived from the assumption that the rates at which $a_i$
-and $c_i$ grow are reciprocally coupled. Extending the original 1D model to
-simulate arealisation on a 2D cortical sheet, we use $a_i(\mb{x},t)$ and
-$c_i(\mb{x},t)$, and model synaptogenesis as
+from intrinsic cortical dynamics. Their model defines how the \cmnt{fraction
+  of occupied synapses} $c_i(x,t)$ and \cmnt{the density of} axon branches
+$a_i(x,t)$ interact at time $t$, along a 1D anterior-posterior axis $x$, for
+$N$ thalamocortical projections indexed by $i$. The model was derived from the
+assumption that the rates at which $a_i$ and $c_i$ grow are reciprocally
+coupled. Extending the original 1D model to simulate arealisation on a 2D
+cortical sheet, we use $a_i(\mb{x},t)$ and $c_i(\mb{x},t)$, and model
+synaptogenesis as
 %
@@
     \includegraphics[width=\linewidth]{./Fig1.png}
-    \caption{The emergence of whisker barrels. \textbf{A} Left shows a cytochrome oxidase stain obtained from rat S1
+    \caption{The emergence of whisker barrels.
+      %
+      \textbf{A} Left shows a cytochrome oxidase stain obtained from rat S1
       by \cite{zheng_signal_2001}, with black lines to delineate barrels and to
@@
       molecular guidance fields ($\rho$), \cmnt{where the domain $S$ has been
-        traced from \textbf{A}}. \textbf{B} The strengths of interaction
-      $\gamma$ with fields $\rho_1$ and $\rho_2$ are indicated for each of 41
-      projections by the lengths of green and blue arrows respectively, assuming
-      that similar fields aligned to the posterior-anterior and medial-lateral
-      axes in the ventroposterior medial nucleus of the thalamus are sampled at
-      the locations of putative barreloid centers (reconstructed from
-      \citealp{haidarliu_size_2001}, their \MPone{Fig.\,5b}). \textbf{C} Results for the example simulation, with parameters
-      \MPtwo{$N=41$, $\alpha=3.6$, $\beta=16.67$, $k=3$, $D=0.5$, $\gamma\in\pm 2$,
-      $\epsilon=1.2$} and $\delta{t}=0.0001$. Colours indicate the thalamic
-      projection for which the connection density is maximal and black lines
-      delineate boundaries (see Movie
-      S1). \textbf{D} Red dots show the \emph{Honda}--\metrics{$\delta(t)$} metric obtained from \metrics{the}
-      simulation approaching that obtained from the \metrics{real} barrels in \textbf{A} (dotted
-      line); black squares show the \emph{pattern difference} metric $\eta(t)$, and reveal the emergence of a correspondence between the real and simulated barrel shapes (units mm$^3$); \metrics{grey hexagons show \metrics{how selectively each cortical site is innervated};
+        traced from \textbf{A}}.
+      %
+      \textbf{B} The strengths of interaction $\gamma$ with fields $\rho_1$
+      and $\rho_2$ are indicated for each of 41 projections by the lengths of
+      green and blue arrows respectively, assuming that similar fields aligned
+      to the posterior-anterior and medial-lateral axes in the ventroposterior
+      medial nucleus of the thalamus are sampled at the locations of putative
+      barreloid centers (reconstructed from \citealp{haidarliu_size_2001},
+      their \MPone{Fig.\,5b}).
+      %
+      \textbf{C} Results for the example simulation, with parameters
+      \MPtwo{$N=41$, $\alpha=3.6$, $\beta=16.67$, $k=3$, $D=0.5$,
+        $\gamma\in\pm 2$, $\epsilon=1.2$} and $\delta{t}=0.0001$. Colours
+      indicate the thalamic projection for which the connection density is
+      maximal and black lines delineate boundaries (see Movie S1).
+      %
+      \textbf{D} Red dots show the \emph{Honda}--\metrics{$\delta(t)$} metric
+      obtained from \metrics{the} simulation approaching that obtained from
+      the \metrics{real} barrels in \textbf{A} (dotted line); black squares
+      show the \emph{pattern difference} metric $\eta(t)$, and reveal the
+      emergence of a correspondence between the real and simulated barrel
+      shapes (units mm$^3$); \metrics{grey hexagons show how selectively each
+        cortical site is innervated;
         $\omega(t) = \oiint_{S} \mu(\mb{x},t) \mathrm{d}S$, where
         $\mu(\mb{x}) \equiv \mathrm{max}_i\big(c_i(\mb{x},t)\big)\big/\sum_{j=1}^{N} c_j(\mb{x},t)$.}
+      %
       \textbf{E} \metrics{Plotted across the cortical sheet, the selectivity
         develops to reveal an alignment with the emergent barrel boundary
-        shapes}. \metrics{Greyscale colour indicates values of $\mu(\mb{x})$ and overlaid
-        contours show $c_i > 0.95 \mathrm{max}_i\big(c_i(\mb{x})\big)$.} All scale bars 1\,mm.}
+        shapes}. \metrics{Greyscale colour indicates values of $\mu(\mb{x})$
+        and overlaid contours show $c_i > 0.95
+        \mathrm{max}_i\big(c_i(\mb{x})\big)$.} All scale bars 1\,mm.}
     \label{fig:main}
@@
   (see also Movie S1) were obtained by
-  conducting a full parameter sweep and choosing a combination ($\alpha=3.6$, $\beta=16.67$, $k=3$, $D=0.5$, $\epsilon=1.2$) that
-  scored well against the following three measures.}
+  conducting a full parameter sweep and choosing a combination
+  ($\alpha=3.6$, $\beta=16.67$, $k=3$, $D=0.5$, $\epsilon=1.2$) that scored
+  well against the following three measures.}
 
@@
     \includegraphics[width=\linewidth]{./Fig2.png}
-    \caption{\MPthreePar{Exploring the parameter space. \textbf{A} Colour indicates the quality of the pattern at $t=30000$ steps, against three measures: Low values of \emph{Honda}--$\delta$ (top
-        row) suggest a Voronoi-like pattern of fields.
-The \emph{pattern difference}, $\eta$ (middle row) measures the difference in the area and arrangement of barrels in real and simulated fields.
-The \emph{selectivity}, $\omega$ (bottom
-        row), measures the specificity with which the cortical sheet is innervated.
-Colour maps are chosen so that lighter (orange and yellow) colours indicate higher quality patterns against each measure.
-White squares indicate combinations of parameters for which simulations were numerically unstable.
+    \caption{\MPthreePar{Exploring the parameter space.
         %
-        The parameter space explored is three dimensional with the ratio $\alpha/\beta$ varying between plots, and the
-        competition parameter $\epsilon$ and the diffusion constant $D$ varying within plots. An asterisk (*)
-        marks the parameters used in Fig.\,\ref{fig:main}. Boxes i),
-        ii)~and iii)~mark parameter sets for which corresponding patterns are
-        shown in \textbf{B} (for $t=30000$ steps).
+        \textbf{A} Colour indicates the quality of the pattern at $t=30000$
+        steps, against three measures: Low values of \emph{Honda}--$\delta$
+        (top row) suggest a Voronoi-like pattern of fields. The \emph{pattern
+          difference}, $\eta$ (middle row) measures the difference in the area
+        and arrangement of barrels in real and simulated fields. The
+        \emph{selectivity}, $\omega$ (bottom row), measures the specificity
+        with which the cortical sheet is innervated. Colour maps are chosen so
+        that lighter (orange and yellow) colours indicate higher quality
+        patterns against each measure. White squares indicate combinations of
+        parameters for which simulations were numerically unstable.
         %
-        \textbf{B} Varying the diffusion constant $D$ generates qualitatively different patterns. Higher values 
-        cause expansions of the peripheral barrels and a corresponding compression of the inner barrels (i).
-Lower values instead cause an expansion of the central barrels and compression of the peripheral barrels (ii).
-Further reducing the rate of diffusion (iii), which is equivalent to increasing the size of the domain and hence simulating development in an animal with a larger cortex, causes a large area to be occupied by projections with intermediate interaction parameters; those with strong interaction parameters are compressed around the edge of the domain, and consequently a barrel pattern fails to form.}}
+        The parameter space explored is three dimensional with the ratio
+        $\alpha/\beta$ varying between plots, and the competition parameter
+        $\epsilon$ and the diffusion constant $D$ varying within plots. An
+        asterisk (*) marks the parameters used in Fig.\,\ref{fig:main}. Boxes
+        i), ii)~and iii)~mark parameter sets for which corresponding patterns
+        are shown in \textbf{B} (for $t=30000$ steps).
+        %
+        \textbf{B} Varying the diffusion constant $D$ generates qualitatively
+        different patterns. Higher values cause expansions of the peripheral
+        barrels and a corresponding compression of the inner barrels
+        (i). Lower values instead cause an expansion of the central barrels
+        and compression of the peripheral barrels (ii). Further reducing the
+        rate of diffusion (iii), which is equivalent to increasing the size of
+        the domain and hence simulating development in an animal with a larger
+        cortex, causes a large area to be occupied by projections with
+        intermediate interaction parameters; those with strong interaction
+        parameters are compressed around the edge of the domain, and
+        consequently a barrel pattern fails to form.}}
     \label{fig:paramsweep}
@@
       \MPthreeSens{Sensitivity analysis.
-\textbf{A} Metrics of map quality $\delta(t)$ (top row), $\omega(t)$ (middle row) and $\eta(t)$ (bottom row), were evaluated at $t=30000$ steps.
-The $y$-axes and colour scales have identical ranges
-        to the colour scales in Fig.\,\ref{fig:paramsweep}, for easy comparison. \emph{Left column:} The effect of adding noise drawn from a uniform distribution,
-        $(\gamma_{max}-\gamma_{min})\,U(0,\nu_\gamma)$,
-        to the values of the interaction parameters, $\gamma$ used in
+        %
+        \textbf{A} Metrics of map quality $\delta(t)$ (top row), $\omega(t)$
+        (middle row) and $\eta(t)$ (bottom row), were evaluated at $t=30000$
+        steps. The $y$-axes and colour scales have identical ranges to the
+        colour scales in Fig.\,\ref{fig:paramsweep}, for easy comparison.
+        %
+        \emph{Left column:} The effect of adding noise drawn from a uniform
+        distribution, $(\gamma_{max}-\gamma_{min})\,U(0,\nu_\gamma)$, to the
+        values of the interaction parameters, $\gamma$ used in
         Fig.\,\ref{fig:main}.
         %
-        \emph{Middle column:} The effect of changing the magnitude and the length scale of
-        noise applied to the guidance fields. Uniform random noise $(\rho_{max} - \rho_{min})\,U(0, \nu_{\rho})$ was added
-        to each (hexagonal) element of $\rho_1(\mb{x})$ and $\rho_2(\mb{x})$ and the result
-        was smoothed by convolution with a symmetric 2D
-        Gaussian kernel of width $\sigma_\rho$. 
+        \emph{Middle column:} The effect of changing the magnitude and the
+        length scale of noise applied to the guidance fields. Uniform random
+        noise $(\rho_{max} - \rho_{min})\,U(0, \nu_{\rho})$ was added to each
+        (hexagonal) element of $\rho_1(\mb{x})$ and $\rho_2(\mb{x})$ and the
+        result was smoothed by convolution with a symmetric 2D Gaussian kernel
+        of width $\sigma_\rho$.
         %
-        \emph{Right column:} The effect of setting the rotational angle
-        of the linearly varying guidance field, $\rho_1(\mb{x})$, to $\phi_{\rho_1}$,
+        \emph{Right column:} The effect of setting the rotational angle of the
+        linearly varying guidance field, $\rho_1(\mb{x})$, to $\phi_{\rho_1}$,
         and modifying its overall gain to $G_{\rho_1}$, whilst keeping the
-        parameters of $\rho_2(\mb{x})$
-        unchanged from those used in the example simulation, for which
-        $\phi_{\rho_2}=84$\textdegree and $G_{\rho_2}=1$.
+        parameters of $\rho_2(\mb{x})$ unchanged from those used in the
+        example simulation, for which $\phi_{\rho_2}=84$\textdegree and
+        $G_{\rho_2}=1$.
         %
@@
         the patterns. i) The effect of significant interaction parameter noise
-        results in topological defects. ii) High magnitude, short lengthscale noise in
-        $\rho(\mb{x})$ leads to non-straight edges between adjacent barrels. iii)
- Reducing the slope of gradient $\rho_1$ (by a factor of 10)
-causes barrel rows B, C and D to become `crushed'
-        down the center line, and edge barrels to dominate.
-iv) Rotating $\rho_1$ by 20\textdegree~causes a slight distortion of the pattern, resulting in an overall anticlockwise rotation of the field locations.
-      }
-    }
+        results in topological defects. ii) High magnitude, short lengthscale
+        noise in $\rho(\mb{x})$ leads to non-straight edges between adjacent
+        barrels. iii) Reducing the slope of gradient $\rho_1$ (by a factor of
+        10) causes barrel rows B, C and D to become `crushed' down the center
+        line, and edge barrels to dominate. iv) Rotating $\rho_1$ by
+        20\textdegree~causes a slight distortion of the pattern, resulting in
+        an overall anticlockwise rotation of the field locations. }}
     \label{fig:sens}
@@
 \MPthreePred{To further investigate the interplay of genetic and intrinsic
-  mechanisms we simulated two well known experimental manipulations of 
-barrel development. First, we simulated a seminal barrel duplication
-  paradigm}
+  mechanisms we simulated two well known experimental manipulations of barrel
+  development. First, we simulated a seminal barrel duplication paradigm}
 \citep{shimogori_fibroblast_2005,assimacopoulos_fibroblast_2012}
@@
 \citep{crossley_mouse_1995}, \MPthreePred{is ectopically expressed (by
-  electroporation) also at the posterior pole. We assume that
-  this results in a mirror of the primary barrel cortex boundary along the
-  rostrocaudal axis}
+  electroporation) also at the posterior pole. We assume that this results in
+  a mirror of the primary barrel cortex boundary along the rostrocaudal axis}
 \cite{assimacopoulos_fibroblast_2012} \MPthreePred{and a mirroring of the
-  anterior-posterior guidance gradient $\rho_1$ at the border between them (Fig.\,\ref{fig:fgf8}A). 
-  The result after 30000 iterations, and otherwise using the parameters of the example simulation, was two mirror-symmetrical
+  anterior-posterior guidance gradient $\rho_1$ at the border between them
+  (Fig.\,\ref{fig:fgf8}A). The result after 30000 iterations, and otherwise
+  using the parameters of the example simulation, was two mirror-symmetrical
   barrel fields comprising $2N$ barrels (Fig.\,\ref{fig:fgf8}B), consistent
@@
       with its mirror. All other model parameters match those in
-      Fig.\,\ref{fig:main}C. \textbf{C} Simulating
-      whisker trimming by reducing the competitiveness,
-      $\epsilon$, of the one projection. For C3 only, $\epsilon$ was mutiplied by
-      $m\in(0,1)$. The pattern shown is that formed
-      after 30000 steps with $m=0.86$, which reduces the size of the C3 field to 65\% of its original size,
-      matching the average barrel area reduction observed by }
-    \citealp{kossut_effects_1992}. \MPthreePred{\textbf{D} The area of the C3 barrel
-reduces as $m$ is reduced, whereas
-      the mean area of neighbouring barrels (B3, C2, C4, D2 \& D3, grey
-      circles) increases. The dotted grey line indicates $m=0.86$ for
-      comparison with panel C.}}
+      Fig.\,\ref{fig:main}C. \textbf{C} Simulating whisker trimming by
+      reducing the competitiveness, $\epsilon$, of the one projection. For C3
+      only, $\epsilon$ was mutiplied by $m\in(0,1)$. The pattern shown is that
+      formed after 30000 steps with $m=0.86$, which reduces the size of the C3
+      field to 65\% of its original size, matching the average barrel area
+      reduction observed by }
+    \citealp{kossut_effects_1992}. \MPthreePred{\textbf{D} The area of the C3
+      barrel reduces as $m$ is reduced, whereas the mean area of neighbouring
+      barrels (B3, C2, C4, D2 \& D3, grey circles) increases. The dotted grey
+      line indicates $m=0.86$ for comparison with panel C.}}
   \label{fig:fgf8}
@@
 
-\MPthreePred{Finally, to investigate the response of the model to environmental
-  manipulation, we simulated a whisker deprivation experiment. In a critical period comprising the first postnatal days, removal of the whiskers by
-  electrocauterization, plucking, or trimming leads to observable changes in
-  brain structures, including the barrel field~\citep{jeanmonod_mouse_1981}.
-Amongst other changes, deprivation of individual whiskers, or an entire row of whiskers leads to smaller barrels~\citep{kossut_effects_1992,loos_somatosensory_1973}.
-We
-  simulated trimming of the individual whisker C3 during the critical
-  period by reducing the competitiveness of the C3 thalamocortical projection, $\epsilon$. As a result, the corresponding field
-  size was smaller (Fig.\,\ref{fig:fgf8}C), and the size of the fields representing
-  the neighbours of C3 increased in size. A reduction in area to 65\%, comparable to that induced by
-  \citealp{kossut_effects_1992}, was obtained in simulation when $\epsilon$ was reduced to 86\% (Fig.\,\ref{fig:fgf8}D.), and the C3 barrel disappeared altogether when $\epsilon$ was less than half of its original value.}
+\MPthreePred{Finally, to investigate the response of the model to
+  environmental manipulation, we simulated a whisker deprivation
+  experiment. In a critical period comprising the first postnatal days,
+  removal of the whiskers by electrocauterization, plucking, or trimming leads
+  to observable changes in brain structures, including the barrel
+  field~\citep{jeanmonod_mouse_1981}. Amongst other changes, deprivation of
+  individual whiskers, or an entire row of whiskers leads to smaller
+  barrels~\citep{kossut_effects_1992,loos_somatosensory_1973}. We simulated
+  trimming of the individual whisker C3 during the critical period by reducing
+  the competitiveness of the C3 thalamocortical projection, $\epsilon$. As a
+  result, the corresponding field size was smaller (Fig.\,\ref{fig:fgf8}C),
+  and the size of the fields representing the neighbours of C3 increased in
+  size. A reduction in area to 65\%, comparable to that induced by
+  \citealp{kossut_effects_1992}, was obtained in simulation when $\epsilon$
+  was reduced to 86\% (Fig.\,\ref{fig:fgf8}D.), and the C3 barrel disappeared
+  altogether when $\epsilon$ was less than half of its original value.}
 
@@
   consistency of the simulated whisker trimming results with those of the
-  original studies demonstrates how the model can be used to investigate the contribution of intrinsic and extrinsic factors to the
-  development of cortical fields.}
+  original studies demonstrates how the model can be used to investigate the
+  contribution of intrinsic and extrinsic factors to the development of
+  cortical fields.}
 
@@
 cortical space, as represented by $\chi$, with an implicit requirement for a
-self/other identifier amongst projections. This latter form of competition may account for the absence of barrels in rodents with larger brains, such as
+self/other identifier amongst projections. This latter form of competition may
+account for the absence of barrels in rodents with larger brains, such as
 capybara, for which competition for space is presumably weaker
@@
 
-\MPone{We concentrated on the representation of the
-  forty-one macrovibrissae that constitute a given barrel field, because their
-  thalamic and cortical correlates are easily identifiable and consistently
-  located, excluding the five rhinal whiskers as their cortical representation is isolated from the main barrel
-  field. We excluded the representation of the microvibrissae to limit
-  the overall complexity of the simulations.}
+\MPone{We concentrated on the representation of the forty-one macrovibrissae
+  that constitute a given barrel field, because their thalamic and cortical
+  correlates are easily identifiable and consistently located, excluding the
+  five rhinal whiskers as their cortical representation is isolated from the
+  main barrel field. We excluded the representation of the microvibrissae to
+  limit the overall complexity of the simulations.}
 
@@
 simplifies the computation of the 2D Laplacian. Within a boundary traced
-around the edge of a
-rat barrel field (Fig.\,\ref{fig:main}A) we set the hex-to-hex distance
-$d$ to 0.03\,mm, which resulted in a lattice containing 6515 hexes for the
-simulations shown in Figs.\,\ref{fig:main}A,C \& D and 12739 hexes for the Fgf8
-misexpression study shown in Fig.\,\ref{fig:fgf8}. Each hex contained 82 time-dependent
-variables: 41 branching densities ($a_i$) and 41 connection densities ($c_i$).
-The rate of change of each of the time-dependent variables (Eqs.\,\ref{eq:dc}
-\& \ref{eq:da}) was computed using a fourth-order Runge-Kutta method.
+around the edge of a rat barrel field (Fig.\,\ref{fig:main}A) we set the
+hex-to-hex distance $d$ to 0.03\,mm, which resulted in a lattice containing
+6515 hexes for the simulations shown in Figs.\,\ref{fig:main}A,C \& D and
+12739 hexes for the Fgf8 misexpression study shown in
+Fig.\,\ref{fig:fgf8}. Each hex contained 82 time-dependent variables: 41
+branching densities ($a_i$) and 41 connection densities ($c_i$).  The rate of
+change of each of the time-dependent variables (Eqs.\,\ref{eq:dc} \&
+\ref{eq:da}) was computed using a fourth-order Runge-Kutta method.
 
@@
 %
-where $v = d/\sqrt{3}$ is the length of each edge of the hexagon and $\dif\gamma$
-is an infinitesimally small distance along its perimeter.
+where $v = d/\sqrt{3}$ is the length of each edge of the hexagon and
+$\dif\gamma$ is an infinitesimally small distance along its perimeter.
 
@@
 ii) The computation of the second term in Eq.\,\ref{eq:divJExpanded},
-$\langle a_i(\mb{p}_0,t)\nabla\vcdot\mb{g}_i(\mb{p}_0)\rangle$, can be written out similarly:
+$\langle a_i(\mb{p}_0,t)\nabla\vcdot\mb{g}_i(\mb{p}_0)\rangle$, can be written
+out similarly:
 
@@
 %
-where $g_i^x$ and $g_i^y$ are the Cartesian components of $\mb{g}_i$. Both this
-last expression, and the final expression of Eq.\,\ref{eq:lapl} can
-be computed locally, by summing over values of the nearest neighbours.
+where $g_i^x$ and $g_i^y$ are the Cartesian components of $\mb{g}_i$. Both
+this last expression, and the final expression of Eq.\,\ref{eq:lapl} can be
+computed locally, by summing over values of the nearest neighbours.
 
 \noindent
-iii) The \MPtwo{middle} term in Eq.\,\ref{eq:divJExpanded} is the scalar product of two
-vector fields which is straightforward to compute from their Cartesian
-components.
+iii) The \MPtwo{middle} term in Eq.\,\ref{eq:divJExpanded} is the scalar
+product of two vector fields which is straightforward to compute from their
+Cartesian components.
 
 \noindent
-\MPtwo{iv) The same method used to compute $\dvrg a_i$ in term (i) is used to compute
-$\dvrg \hat{a}_i$.}
+\MPtwo{iv) The same method used to compute $\dvrg a_i$ in term (i) is used to
+  compute $\dvrg \hat{a}_i$.}
 
 \noindent
-\MPtwo{v) The final term is the scalar product of the two vector fields $\nabla
-  a_i$ and $\nabla \hat{a}_i$.}
+\MPtwo{v) The final term is the scalar product of the two vector fields
+  $\nabla a_i$ and $\nabla \hat{a}_i$.}
 
@@
 %
-can be fulfilled. On
-the boundary, the contribution to $\mb{J}$ resulting from the first term of
-Eq.\,\ref{eq:divJExpanded} can be fixed to 0 by the `ghost cell method' in
-which, during the evaluation of (i), a hex outside the boundary containing the
-same value as the hex inside the boundary is imagined to exist such that the
-flux of $\mb{J}$ across the boundary is 0. Then, $\mb{g}_i(\mb{x})$ can be
-tailored so that it, and its normal derivative, approach 0 at the boundary,
-ensuring that the second and third terms of Eq.\,\ref{eq:divJExpanded} also
-contribute nothing to $\mb{J}$. This is achieved by \MPtwo{multiplying}
-$\mb{g}_i(\mb{x})$ \MPtwo{by} a sharp logistic function of the distance, $d_b$,
-from $\mb{x}$ to the boundary, \MPtwo{of the form $1/[1 +
-  \exp(100(d_f-d_b))]$, where $d_f=0.1$\,mm $\approx 3d$ is the boundary fall-off distance}.
+can be fulfilled. On the boundary, the contribution to $\mb{J}$ resulting from
+the first term of Eq.\,\ref{eq:divJExpanded} can be fixed to 0 by the `ghost
+cell method' in which, during the evaluation of (i), a hex outside the
+boundary containing the same value as the hex inside the boundary is imagined
+to exist such that the flux of $\mb{J}$ across the boundary is 0. Then,
+$\mb{g}_i(\mb{x})$ can be tailored so that it, and its normal derivative,
+approach 0 at the boundary, ensuring that the second and third terms of
+Eq.\,\ref{eq:divJExpanded} also contribute nothing to $\mb{J}$. This is
+achieved by \MPtwo{multiplying} $\mb{g}_i(\mb{x})$ \MPtwo{by} a sharp logistic
+function of the distance, $d_b$, from $\mb{x}$ to the boundary, \MPtwo{of the
+  form $1/[1 + \exp(100(d_f-d_b))]$, where $d_f=0.1$\,mm $\approx 3d$ is the
+  boundary fall-off distance}.
 
-\metrics{The \emph{pattern difference} metric, $\eta$, incorporates information about
-  the differences in the areas of simulated and experimentally determined (real) barrels
-  ($\mathcal{A}_i^{\mathrm{sim}}$ and $\mathcal{A}_i^{\mathrm{exp}}$), as well as
- information contained in the `adjacency vector' for each barrel, $\mathcal{V}_i$, the $j$-th element of which is the length of the border between barrel $i$ and $j$.
-For a well-formed barrel pattern, $\mathcal{V}_{i}$ is a sparse vector.
-A
-  dimensionless quantity can be obtained from the scalar product of the simulated and experimental
-  adjacency vectors:
+\metrics{The \emph{pattern difference} metric, $\eta$, incorporates
+  information about the differences in the areas of simulated and
+  experimentally determined (real) barrels ($\mathcal{A}_i^{\mathrm{sim}}$ and
+  $\mathcal{A}_i^{\mathrm{exp}}$), as well as information contained in the
+  `adjacency vector' for each barrel, $\mathcal{V}_i$, the $j$-th element of
+  which is the length of the border between barrel $i$ and $j$. For a
+  well-formed barrel pattern, $\mathcal{V}_{i}$ is a sparse vector. A
+  dimensionless quantity can be obtained from the scalar product of the
+  simulated and experimental adjacency vectors:
+  %
   $\frac{1}{N} \sum_i \frac { \mathcal{V}_{i}^{\mathrm{sim}}}
-  {b_{i}^{\mathrm{sim}} } \vcdot \frac{\mathcal{V}_{i}^{\mathrm{exp}}}{b_{i}^{\mathrm{exp}}}$, where e.g., 
-  $b_i^{\mathrm{exp}}$, is the total length of the border around 
-real barrel $i$. This quantity is small when the fields that form in simulation have dissimilar neighbour relations to those of the real barrels
-  (e.g., Fig.\,\ref{fig:main}C, $t=1000$), and maximal for a precise topological map ($t=10000$).}
+  {b_{i}^{\mathrm{sim}} } \vcdot \frac{\mathcal{V}_{i}^{\mathrm{exp}}}{b_{i}^{\mathrm{exp}}}$,
   %
-\metrics{A second comparison considers the mean magnitude of the difference between the simulated and experimental
-  vectors: $\frac{1}{N} \sum_i \norm{ \mathcal{V}_{i}^{\mathrm{sim}} -
-    \mathcal{V}_{i}^{\mathrm{exp}} } $. This tends to 0 for a perfect
-  match and can separate patterns with straight boundaries from those with `noisy edges' (as in Fig.\,\ref{fig:sens}B\,ii).}
+  where e.g., $b_i^{\mathrm{exp}}$, is the total length of the border around
+  real barrel $i$. This quantity is small when the fields that form in
+  simulation have dissimilar neighbour relations to those of the real barrels
+  (e.g., Fig.\,\ref{fig:main}C, $t=1000$), and maximal for a precise
+  topological map ($t=10000$).}
+%
+\metrics{A second comparison considers the mean magnitude of the difference
+  between the simulated and experimental vectors: $\frac{1}{N} \sum_i \norm{
+    \mathcal{V}_{i}^{\mathrm{sim}} - \mathcal{V}_{i}^{\mathrm{exp}} } $. This
+  tends to 0 for a perfect match and can separate patterns with straight
+  boundaries from those with `noisy edges' (as in Fig.\,\ref{fig:sens}B\,ii).}
 %

### 2020-07-09 paper/briefpaper/brief.tex
Re-word the statement about dot(a_i) prop. dot(c_i)

@@
-% Brief report style paper
+% (This started life as a...) Brief report style paper
 
@@
 placeholder. The second term on the right represents the coupling between axon
-branching and synaptogenesis \cmnt{and implies an assumption of spatially
-  homogeneous synaptic density which makes it possible to relate the quantity
-  $c_i$ to connection density}.
+branching and synaptogenesis \cmnt{and implies the assumption that the spatial
+  distribution of synaptic density across the cortical sheet is broadly
+  homogeneous. This makes it possible to relate the quantity $c_i$ to
+  connection density}.
 

### 2020-07-09 paper/briefpaper/brief.tex
Changes one ref from a cite to a citep

@@
   a mirror of the primary barrel cortex boundary along the rostrocaudal axis}
-\cite{assimacopoulos_fibroblast_2012} \MPthreePred{and a mirroring of the
+\citep{assimacopoulos_fibroblast_2012} \MPthreePred{and a mirroring of the
   anterior-posterior guidance gradient $\rho_1$ at the border between them

### 2020-07-10 paper/briefpaper/brief.tex
Modifies paper with row trimming result

@@
     \citealp{kossut_effects_1992}. \MPthreePred{\textbf{D} The area of the C3
-      barrel reduces as $m$ is reduced, whereas the mean area of neighbouring
+      barrel (black squares) reduces as $m$ is reduced, whereas the mean area of neighbouring
       barrels (B3, C2, C4, D2 \& D3, grey circles) increases. The dotted grey
-      line indicates $m=0.86$ for comparison with panel C.}}
+      line indicates $m=0.86$ for comparison with panel C. If the $\epsilon$
+      value is instead reduced for \emph{all} row C projections (including the interstitial
+      $\gamma$ projection), the mean row C barrel area is only slightly
+      reduced (light grey triangles). In this case, the mean row C area at
+      $m=0.86$ is 91\% of that for $m=1$.}}
   \label{fig:fgf8}
@@
   field~\citep{jeanmonod_mouse_1981}. Amongst other changes, deprivation of
-  individual whiskers, or an entire row of whiskers leads to smaller
-  barrels~\citep{kossut_effects_1992,loos_somatosensory_1973}. We simulated
+  individual whiskers leads to smaller
+  % loos_somatosensory_1973 for row cauterization
+  barrels~\citep{kossut_effects_1992}. We simulated
   trimming of the individual whisker C3 during the critical period by reducing
@@
   size. A reduction in area to 65\%, comparable to that induced by
-  \citealp{kossut_effects_1992}, was obtained in simulation when $\epsilon$
-  was reduced to 86\% (Fig.\,\ref{fig:fgf8}D.), and the C3 barrel disappeared
+  \cite{kossut_effects_1992}, was obtained in simulation when $\epsilon$
+  was reduced to 86\% (Fig.\,\ref{fig:fgf8}D), and the C3 barrel disappeared
   altogether when $\epsilon$ was less than half of its original value.}
+
+\MPthreePred{Although individual whisker trimming reduces barrel size, if an
+  entire row is trimmed, barrel sizes for the trimmed row are not obviously
+  changed \citep{land_metabolic_1985}. We investigated with a simulation in
+  which we varied $\epsilon$ for all of the row C projections.  Although the
+  barrels which formed for row C did show some reduction in area
+  (Fig.\,\ref{fig:fgf8}D), this reduction was small when compared with the
+  individually trimmed C3 simulation. If the effect of trimming any whisker is
+  to reduce its $\epsilon$ (the competitiveness of its projection) to 86\% of
+  its original value, then the model predicts that the cortical barrels for
+  the trimmed C row will retain 91\% of their area.}
 

### 2020-07-10 analysis/plot_trim_areaonly.py
Final tweaks to Fig 4D graph

@@ -29,7 +29,7 @@ matplotlib.rc('font', **fnt)
 plt.rcParams['svg.fonttype'] = 'none'
 
 # And plot this longhand here:
-F1 = plt.figure (figsize=(10,6))
+F1 = plt.figure (figsize=(10,6.5))
 
 ## load data from csv
 trimdata = np.genfromtxt ('postproc/whisker_trim_individual.csv', delimiter=",", names=True)
@@ -109,9 +109,9 @@ l1_3, = ax1.plot(tdc3[:]['eps_mult'], tdc3[:]['area'], 's-', markersize=main_siz
 ax1.set_xlim([0.5,1.0])
 ax1.set_ylim([0,0.16])
 
-ax1.text (0.55, 0.022, 'C3', fontsize=28, horizontalalignment='left', color=col.black);
+ax1.text (0.54, 0.022, 'C3', fontsize=28, horizontalalignment='left', color=col.black);
 ax1.text (0.624, 0.144, 'neighbours of C3', fontsize=28, horizontalalignment='left', color=col.black);
-ax1.text (0.595, 0.085, 'row C (mean)', fontsize=28, horizontalalignment='left', color=col.black);
+ax1.text (0.595, 0.088, 'row C (mean)', fontsize=28, horizontalalignment='left', color=col.black);
 
 
 ax1.set_xlabel ('$m$')

### 2020-07-10 paper/briefpaper/brief.tex
Another tweak of the Fig 4 caption
Changes to Fig 4, a white background for Fig 3.

@@
     \citealp{kossut_effects_1992}. \MPthreePred{\textbf{D} The area of the C3
-      barrel (black squares) reduces as $m$ is reduced, whereas the mean area of neighbouring
-      barrels (B3, C2, C4, D2 \& D3, grey circles) increases. The dotted grey
-      line indicates $m=0.86$ for comparison with panel C. If the $\epsilon$
-      value is instead reduced for \emph{all} row C projections (including the interstitial
-      $\gamma$ projection), the mean row C barrel area is only slightly
-      reduced (light grey triangles). In this case, the mean row C area at
-      $m=0.86$ is 91\% of that for $m=1$.}}
+      barrel (black squares) reduces as $m$ is reduced, whereas the mean area
+      of neighbouring barrels (B3, C2, C4, D2 \& D3, grey circles)
+      increases. The dotted grey line indicates $m=0.86$ for comparison with
+      panel C. If the $\epsilon$ value is instead reduced for \emph{all} row C
+      projections (including the interstitial $\gamma$ projection), the mean
+      row C barrel area is only slightly reduced (light grey triangles). In
+      this case, the mean area of a simulated row C barrel at $m=0.86$ is 91\%
+      of that for $m=1$.}}
   \label{fig:fgf8}
@@
   its original value, then the model predicts that the cortical barrels for
-  the trimmed C row will retain 91\% of their area.}
+  the trimmed C row will retain 91\% of their area on average.}
 
@@
 projection for which the connection density is maximal, black lines delineate
-boundaries, and overlaid contours show $c>0.5$ \textbf{FIXME: Will need to
-  check this statement after regenerating movie}. \cmnt{Barrel labels are
+boundaries, and overlaid contours show
+\cmnt{$c_i > 0.95 \mathrm{max}_i\big(c_i(\mb{x})\big)$. Barrel labels are
   always located at the centroid of their field(s).} The final frame in the movie

### 2020-07-10 analysis/plot_trim_areaonly.py

@@ -109,9 +109,9 @@ l1_3, = ax1.plot(tdc3[:]['eps_mult'], tdc3[:]['area'], 's-', markersize=main_siz
 ax1.set_xlim([0.5,1.0])
 ax1.set_ylim([0,0.16])
 
-ax1.text (0.54, 0.022, 'C3', fontsize=28, horizontalalignment='left', color=col.black);
-ax1.text (0.624, 0.144, 'neighbours of C3', fontsize=28, horizontalalignment='left', color=col.black);
-ax1.text (0.595, 0.088, 'row C (mean)', fontsize=28, horizontalalignment='left', color=col.black);
+ax1.text (0.53, 0.144, 'C3 neighbours (C3 trimmed)', fontsize=24, horizontalalignment='left', color=col.black);
+ax1.text (0.52, 0.086, 'C-row mean (C-row trimmed)', fontsize=24, horizontalalignment='left', color=col.black);
+ax1.text (0.653, 0.017, 'C3 (C3 trimmed)', fontsize=24, horizontalalignment='left', color=col.black);
 
 
 ax1.set_xlabel ('$m$')

### 2020-07-11 paper/briefpaper/brief.tex
Makes changes suggested by Leah

@@
 \emph{prior} to barreloid formation \citep{agmon_organized_1993}.  \mpsix{In
-  mice, axons from the trigeminal ganglion arrive in the principal sensory
-  nucleus of the trigeminal nerve (PrV) at E12, then axons from the PrV arrive
+  mice, axons from the trigeminal ganglion arrive in the principal division of
+  the trigeminal nucleus (PrV) at E12, then axons from the PrV arrive
   in the ventroposteromedial nucleus of the thalamus (VPM) at E17, then axons
@@
   territory, by increasing the tendency for each projection to compete for
-  cortical space into which to concentrate. The major modification required
+  cortical space in which to branch and make connections. The major
+  modification required
   was thus to introduce into the model an additional source of competition
@@
   cortex} (\citealp{vanderhaeghen_mapping_2000,miller_epha7-ephrin-a5_2006}).
-\MPone{Cells originating in the medial part of ventrobasal thalamus
-  (comprising VPM and VPL) express high levels of EphA receptors and project
-  to the lateral part of S1, which expresses low levels of ephrin-A5, and
-  cells originating in the lateral part of ventrobasal thalamus express low
-  levels of EphA receptors and project to the medial part of S1, which
-  expresses high levels of ephrin-A5} (see
+
+\MPone{Cells originating in VPM express high levels of EphA receptors and
+  project to the lateral part of S1, which expresses low levels of ephrin-A5,
+  and cells originating in the VPL express low levels of EphA receptors and
+  project to the medial part of S1, which expresses high levels of ephrin-A5}
+(see
 \citealp{gao_regulation_1998,dufour_area_2003,vanderhaeghen_developmental_2004,speer_grading_2005,torii_role_2013}).
@@
   diffusion may be considered equivalent to increasing the overall size of the
-  domain, $S$, hence insights into barrel development in species with
-  substantially larger barrel fields, such as capybara which do not have
-  clearly identifiable barrels, may be gained by studying pattern formation
+  domain, $S$. Hence, insights into barrel development in species with a
+  larger representation of the vibrissae, which do not have barrel fields, may
+  be gained by studying pattern formation
   when $D$ is small. In this context, it is interesting to note that for small
@@
 \cmnt{In reality, scores of smaller Dirichletform barrels representing the
-  microvibrissae form to the right of where the E-row barrels emerge in our
-  simulations, presumably via the same competitive processes. Enforcing here
+  microvibrissae form alongside the E-row barrels, presumably via the same
+  competitive processes. Enforcing here
   the same boundary condition as used to represent the true edges of the

### 2020-07-11 paper/briefpaper/brief.tex
One more suggestion from Leah, not addressed in previous commmit

@@
 
-\MPthreePred{To further investigate the interplay of genetic and intrinsic
-  mechanisms we simulated two well known experimental manipulations of barrel
+\MPthreePred{To further investigate the interplay of genes intrinsic to the
+  developing neocortex and extrinsic factors such as thalamocortical input,
+  we simulated two well known experimental manipulations of barrel
   development. First, we simulated a seminal barrel duplication paradigm}

### 2020-07-11 paper/briefpaper/brief.tex
minor edits mainly to fix small typos and minor grammatical issues

@@
 placeholder. The second term on the right represents the coupling between axon
-branching and synaptogenesis \cmnt{and implies the assumption that the spatial
+branching and synaptogenesis, \cmnt{and an assumption that the spatial
   distribution of synaptic density across the cortical sheet is broadly
-  homogeneous. This makes it possible to relate the quantity $c_i$ to
+  homogeneous. As such, the quantity $c_i$ can be thought of as the
   connection density}.
@@
   close correspondence of the simulated barrel field with a Voronoi
-  tesselation, and thus with a biological barrel field.
-  (In mouse, $\delta\approx0.054$} \citep{senft_mouse_1991}\metrics{;
-  our analysis of data from} \cite{zheng_signal_2001} \metrics{ indicates that
-  the value for rat is similar). For the tesselation
+  tesselation, and thus with a biological barrel field 
+  (for mice, $\delta\approx0.054$} \citealp{senft_mouse_1991}\metrics{, and 
+  our analysis of data from} \citealp{zheng_signal_2001} \metrics{ indicates that
+  the value for rats is similar). For the tesselation
   that is overlaid on the real barrel field in Fig.\,\ref{fig:main}A,
@@
   connection density of the most dense projection divided by the sum over all
-  projection densities, $\omega$. The connection selectivity increased as the
+  projection densities, $\omega$. The overall connection selectivity increased as the
   barrel map self-organized in the example simulation (Fig.\,\ref{fig:main}D,
@@
   overlapping with the emergent barrel centers (Fig.\,\ref{fig:main}E).}
-
-\metrics{Against these three metrics we are also able to characterise the
-  robustness of self-organization to the model parameters, and to investigate
-  the sensitivity of the model to variation in its inputs.}
 
@@
 
-\MPthreePar{Fig.\,\ref{fig:paramsweep}A shows values of $\delta$, $\eta$ and
+\metrics{Against these three metrics we are also able to characterise the
+  robustness of self-organization to the model parameters, and to investigate
+  the sensitivity of the model to variation in its inputs.} \MPthreePar{Fig.\,\ref{fig:paramsweep}A shows values of $\delta$, $\eta$ and
   $\omega$ obtained after 30000 iterations, from 216 independent simulations,
@@
   by simulations in which the diffusion constant $D$ and the strength of
-  competition $\epsilon$ were high. Third, strongest connection selectivities,
+  competition $\epsilon$ were high. Third, strongest overall connection selectivities,
   $\omega$, were obtained for lower values of $D$. Fourth, variation in the
@@
       Fig.\,\ref{fig:main}C. \textbf{C} Simulating whisker trimming by
-      reducing the competitiveness, $\epsilon$, of the one projection. For C3
+      reducing the competitiveness, $\epsilon$, of one projection. For C3
       only, $\epsilon$ was mutiplied by $m\in(0,1)$. The pattern shown is that
@@
       reduction observed by }
-    \citealp{kossut_effects_1992}. \MPthreePred{\textbf{D} The area of the C3
+    \cite{kossut_effects_1992}. \MPthreePred{\textbf{D} The area of the C3
       barrel (black squares) reduces as $m$ is reduced, whereas the mean area
@@
       increases. The dotted grey line indicates $m=0.86$ for comparison with
-      panel C. If the $\epsilon$ value is instead reduced for \emph{all} row C
+      panel \textbf{C}. If the $\epsilon$ value is instead reduced for \emph{all} row C
       projections (including the interstitial $\gamma$ projection), the mean

### 2020-07-13 paper/briefpaper/brief.tex
Update to the We reasoned... sentence

@@
 \mpfive{We reasoned that an arbitrary number of distinct field locations may
-  be determined by a minimum of two guidance fields, if a tendency for the
-  density of projections to concentrate is appropriately balanced with
-  the tendency of projections that interact more strongly with the guidance
-  fields to migrate further in the directions specified by their
-  gradients. Accordingly, projections that interact most strongly with a given
+  be determined by a minimum of two guidance gradients, if the concentration
+  of the projection densities is influenced by competition between
+  projections, and if a projection that interacts more strongly with a given
+  guidance gradient migrates further in the direction of that gradient.
+  Accordingly, projections that interact most strongly with a given
   guidance gradient would come to occupy cortical locations at which that

### 2020-07-13 paper/briefpaper/brief.tex
Move a comma

@@
   tesselation, and thus with a biological barrel field
-  (for mice, $\delta\approx0.054$} \citealp{senft_mouse_1991}\metrics{, and
+  (for mice $\delta\approx0.054$,} \citealp{senft_mouse_1991}\metrics{, and
   our analysis of data from} \citealp{zheng_signal_2001} \metrics{ indicates that

### 2020-07-13 paper/briefpaper/brief.tex
Removes contours from image and changes movie script to generate frames without contours

@@
       indicate the thalamic projection for which the connection density is
-      maximal and black lines delineate boundaries (see Movie S1).
+      maximal, \cmnt{barrel labels are located at the centroid of each
+        region and} black lines delineate boundaries (see Movie S1).
       %
@@
         develops to reveal an alignment with the emergent barrel boundary
-        shapes}. \metrics{Greyscale colour indicates values of $\mu(\mb{x})$
-        and overlaid contours show $c_i > 0.95
-        \mathrm{max}_i\big(c_i(\mb{x})\big)$.} All scale bars 1\,mm.}
+        shapes}. \metrics{Greyscale colour indicates values of $\mu(\mb{x})$.}
+      All scale bars 1\,mm.}
     \label{fig:main}
@@
 $\epsilon=1.2$ and $\delta{t}=0.0001$}. Colours indicate the thalamic
-projection for which the connection density is maximal, black lines delineate
-boundaries, and overlaid contours show
-\cmnt{$c_i > 0.95 \mathrm{max}_i\big(c_i(\mb{x})\big)$. Barrel labels are
+projection for which the connection density is maximal and black lines delineate
+boundaries. \MPtwo{Barrel labels are
   always located at the centroid of their field(s).} The final frame in the movie

### 2020-07-13 analysis/plot_cid.py

@@ -71,7 +71,7 @@ def content():
         sf.showNames = True
     else:
         sf.showNames = False
-    sf.svgNames = False  # set True to put labels on separate svg
+    sf.svgNames = True  # set True to put labels on separate svg
     sf.showBarrelAreas = False # True to print barrel areas with labels
     sf.showBarrelHondas = False # True to print barrel Hondas with labels
     sf.domcentres = bdo.domcentres[0]
@@ -90,7 +90,7 @@ def content():
     #sf.addContour (maxc[:,0], 0.4, 'grey', 1.6);
 
     # Or single contour for each field
-    do_contour = 1
+    do_contour = 0
     if do_contour:
         for ii in range(0,bdo.N):
             c = bdo.c[ii,:,0]

### 2020-07-13 analysis/plot_cid_all.py

@@ -53,6 +53,9 @@ sf.boundarylw = 1.0
 sf.boundaryColour = col.black
 sf.boundaryOuterHexColour = col.gray50
 
+# No contours in the movie
+do_contours = 0
+
 for t in range(0,bdo.t_steps.size):
 
     print ('Printing frame {0}'.format(t))
@@ -79,12 +82,13 @@ for t in range(0,bdo.t_steps.size):
     print ('plotPoly...')
     sf.plotPoly()
 
-    print ('Contours...')
-    # A single contour for each field
-    for ii in range(0,bdo.N):
-        c = bdo.c[ii,:,t]
-        ccontour = 0.95*np.max(c)
-        sf.addContour (c, ccontour, 'white', 1.0, ii, False);
+    if do_contours:
+        print ('Contours...')
+        # A single contour for each field
+        for ii in range(0,bdo.N):
+            c = bdo.c[ii,:,t]
+            ccontour = 0.95*np.max(c)
+            sf.addContour (c, ccontour, 'white', 1.0, ii, False);
 
     print ('Outer boundary...')
     sf.addOuterBoundary()

### 2020-07-13 analysis/plot_locn.py

@@ -83,7 +83,7 @@ sf.plotPoly()
 #sf.addContour (maxc[:,0], 0.4, 'grey', 1.6);
 
 # Or single contour for each field
-do_contours = 1
+do_contours = 0
 if do_contours:
     for ii in range(0,bdo.N):
         c = bdo.c[ii,:,0]

### 2020-07-13 analysis/plot_locn_all.py

@@ -1,0 +1,104 @@
+import os
+import sys
+sys.path.insert (0, './include')
+import numpy as np
+import matplotlib.pyplot as plt
+# Import data loading code:
+import BarrelData as bd
+# Import my plotting code:
+import plot as pt
+import sebcolour
+
+# Get target x/y hex to show trace for and the time step to show the
+# map for from the arguments:
+if len(sys.argv) < 2:
+    print('Provide logdirname on cmd line.')
+    exit(1)
+logdirname = sys.argv[1].rstrip ('/')
+print ('basename: {0}'.format(os.path.basename(logdirname)))
+# time index
+ti = -1
+
+# Read the data
+bdo = bd.BarrelData()
+# Set True for inter-lines:
+bdo.loadAnalysisData = True
+bdo.loadDivisions = False
+# If loadGuidance is True, then expt id map will be plotted:
+bdo.loadGuidance = False
+bdo.loadHexFlags = True
+bdo.loadSimData = True
+bdo.loadTimeStep = ti
+bdo.load (logdirname)
+bdo.computeLocalization()
+
+# Plot a surface
+import Surface as surf
+sf = surf.Surface (19.2, 10.8)
+sf.associate(bdo)
+
+sf.showScalebar = False
+sf.showAxes = False
+sf.sb1 = [-1.3, -0.8]
+sf.sb2 = [-0.3, -0.8]
+sf.fs2 = 32
+sf.sbtext = '1 mm'
+sf.sbtpos = [-1.1, -1.1]
+sf.sblw = 5
+sf.sbfs = 48
+sf.showNames = True
+sf.svgNames = False
+sf.showBoundaries = True
+col = sebcolour.Colour()
+sf.cmap = plt.cm.get_cmap('Greys')
+sf.boundarylw = 1.0
+sf.boundaryColour = col.black
+sf.boundaryOuterHexColour = col.gray50
+
+do_contours = False
+
+for t in range(0,bdo.t_steps.size):
+
+    # Compute max of c
+    maxc = np.max (bdo.c[:,:,t], axis=0)
+
+    # Either use the precomputed ID map:
+    c_id = bdo.id_c[:,t]
+
+    # Compute the colour map
+    #colmap = np.zeros([bdo.nhex,3], dtype=float)
+    #ii = 0
+    #for oneid in c_id:
+    #    colmap[ii] = bdo.gammaColour_byid[oneid]
+    #    ii = ii + 1
+
[…]

### 2020-07-13 analysis/replot.sh

@@ -4,6 +4,12 @@
 # and SVG files which I then incorporate into Inkscape drawings, to
 # generate the final figures. Find output in ./plots/
 
+# Fig 1 C
+python plot_cid.py ../logs/41N2M_thalguide_Fig1  1000
+python plot_cid.py ../logs/41N2M_thalguide_Fig1 10000
+python plot_cid.py ../logs/41N2M_thalguide_Fig1 30000
+
+# Fig 1 D
 python plot_honda_quick.py
 
 python plot_paramsearch_paper.py

### 2020-07-14 paper/briefpaper/barrels_proof.tex
Adds the all-black final version

--- 2020-07-13 paper/briefpaper/brief.tex
@@
 
+% Black, so that updates following review do not show
+\definecolor{colcmnt}        {rgb} {0.0, 0.0, 0.0}
+% A command to highlight changes for reviewers
+\newcommand{\cmnt}[1]{\textcolor{colcmnt}{#1}}
+
 % Some document-defined commands
 \newcommand*\dif{\mathop{}\!\mathrm{d}}
-
-\input{sebcolour.tex}
-
-% Miscellaneous changes
-\newcommand{\cmnt}[1]{\textcolor{colcmnt}{#1}}
-% Justify gamma/interaction parameter choice. major point 1.
-\newcommand{\MPone}[1]{\textcolor{colmpone}{#1}}
-% Blue changes are major point 2 (competition mechanism) and other, general
-% uncategorised changes.
-\newcommand{\MPtwo}[1]{\textcolor{colmptwo}{#1}}
-% Parameter search. Major point 3
-\newcommand{\MPthreePar}[1]{\textcolor{colmpthreepar}{#1}}
-% Sensitivity analysis. Major point 3
-\newcommand{\MPthreeSens}[1]{\textcolor{colmpthreesens}{#1}}
-% More predictions. Major point 3
-\newcommand{\MPthreePred}[1]{\textcolor{colmpthreepred}{#1}}
-% Justify competition:
-\newcommand{\mpfive}[1]{\textcolor{colmpfive}{#1}}
-% Minor point 4:
-\newcommand{\mpfour}[1]{\textcolor{colmpfour}{#1}}
-% Minor point 6:
-\newcommand{\mpsix}[1]{\textcolor{colmpsix}{#1}}
-% Describe the metrics
-\newcommand{\metrics}[1]{\textcolor{colmetrics}{#1}}
-
 \newcommand{\dvrg}{\nabla\vcdot\nabla}
-\newcommand{\e}{\emph}
-\newcommand{\bol}{\textbf}
 \newcommand{\mb}[1]{\mathbf{#1}}
@@
 \citep{shimogori_fibroblast_2005,bishop_regulation_2000}, which are expressed
-in gradients \mpfour{spanning the cortical sheet} that mark orthogonal axes
+in gradients \cmnt{spanning the cortical sheet} that mark orthogonal axes
 and can be manipulated to stretch,
@@
 that axons arrive in the cortical plate as an undifferentiated bundle,
-\emph{prior} to barreloid formation \citep{agmon_organized_1993}.  \mpsix{In
+\emph{prior} to barreloid formation \citep{agmon_organized_1993}. \cmnt{In
   mice, axons from the trigeminal ganglion arrive in the principal division of
@@
 \begin{equation} \label{eq:da}
-\frac{\partial a_i}{\partial t} = \nabla\vcdot\left(D \nabla a_i-a_i\sum_{j=1}^{M} \gamma_{i,j}\nabla \rho_j(\mb{x}) \MPtwo{+ \chi_i}\right) - \frac{\partial c_i}{\partial t}.
+\frac{\partial a_i}{\partial t} = \nabla\vcdot\left(D \nabla a_i-a_i\sum_{j=1}^{M} \gamma_{i,j}\nabla \rho_j(\mb{x}) \cmnt{+ \chi_i}\right) - \frac{\partial c_i}{\partial t}.
 \end{equation}
@@
       barreloid centers (reconstructed from \citealp{haidarliu_size_2001},
-      their \MPone{Fig.\,5b}).
+      their \cmnt{Fig.\,5b}).
       %
       \textbf{C} Results for the example simulation, with parameters
-      \MPtwo{$N=41$, $\alpha=3.6$, $\beta=16.67$, $k=3$, $D=0.5$,
+      \cmnt{$N=41$, $\alpha=3.6$, $\beta=16.67$, $k=3$, $D=0.5$,
         $\gamma\in\pm 2$, $\epsilon=1.2$} and $\delta{t}=0.0001$. Colours
@@
       %
-      \textbf{D} Red dots show the \emph{Honda}--\metrics{$\delta(t)$} metric
-      obtained from \metrics{the} simulation approaching that obtained from
-      the \metrics{real} barrels in \textbf{A} (dotted line); black squares
+      \textbf{D} Red dots show the \emph{Honda}--\cmnt{$\delta(t)$} metric
+      obtained from \cmnt{the} simulation approaching that obtained from
+      the \cmnt{real} barrels in \textbf{A} (dotted line); black squares
       show the \emph{pattern difference} metric $\eta(t)$, and reveal the
       emergence of a correspondence between the real and simulated barrel
-      shapes (units mm$^3$); \metrics{grey hexagons show how selectively each
+      shapes (units mm$^3$); \cmnt{grey hexagons show how selectively each
         cortical site is innervated;
@@
       %
-      \textbf{E} \metrics{Plotted across the cortical sheet, the selectivity
+      \textbf{E} \cmnt{Plotted across the cortical sheet, the selectivity
         develops to reveal an alignment with the emergent barrel boundary
-        shapes}. \metrics{Greyscale colour indicates values of $\mu(\mb{x})$.}
+        shapes}. \cmnt{Greyscale colour indicates values of $\mu(\mb{x})$.}
       All scale bars 1\,mm.}
@@
 
-\mpfive{We reasoned that an arbitrary number of distinct field locations may
+\cmnt{We reasoned that an arbitrary number of distinct field locations may
   be determined by a minimum of two guidance gradients, if the concentration
@@
 
-\mpfive{Such dynamics are quite unlike those described by classic
-  chemospecificity models} \citep{sperry_chemoaffinity_1963}, \mpfive{which
+\cmnt{Such dynamics are quite unlike those described by classic
+  chemospecificity models} \citep{sperry_chemoaffinity_1963}, \cmnt{which
   essentially assume center-points by specifying conditions in the target
@@
 
-\mpfive{Testing this reasoning required increasing the strength of the
+\cmnt{Testing this reasoning required increasing the strength of the
   competition between simulated thalamocortical projections for cortical
@@
 connections. To introduce competition also in terms of axon branching,
-\MPtwo{whilst ensuring that $a_i$ is conserved over time,} we
+\cmnt{whilst ensuring that $a_i$ is conserved over time,} we
 redefined
@@
 \begin{equation} \label{eq:comp}
-  \color{colmptwo}
+  \color{colcmnt}
   \chi_i(\mb{x}, t) = \frac{\epsilon a_i}{N-1} \nabla \textstyle\sum_{j\ne i}^{N}{a_j}.
@@
 %
-\mpfive{This term contributes to the \emph{flux} of axonal branching as an
+\cmnt{This term contributes to the \emph{flux} of axonal branching as an
   additional source of diffusion, scaled by $\epsilon$, which reduces the
@@
 
-\MPone{In addition, the model we have outlined requires that molecular
+\cmnt{In addition, the model we have outlined requires that molecular
   guidance gradients in the cortex are complemented by graded values of the
@@
 
-\MPone{Cells originating in VPM express high levels of EphA receptors and
+\cmnt{Cells originating in VPM express high levels of EphA receptors and
   project to the lateral part of S1, which expresses low levels of ephrin-A5,
@@
 \citealp{gao_regulation_1998,dufour_area_2003,vanderhaeghen_developmental_2004,speer_grading_2005,torii_role_2013}).
-\MPone{We assume that such patterning arises because the relative strengths of
+\cmnt{We assume that such patterning arises because the relative strengths of
   interaction with guidance molecules (e.g., ephrin-A5) in the cortex are
@@
 
-\MPone{For simplicity, the two simulated thalamic interaction gradients, as
+\cmnt{For simplicity, the two simulated thalamic interaction gradients, as
   well as the two cortical guidance gradients, were initially chosen to be
@@
   plane defined by the anterior-posterior and medial-lateral axes, were
-  estimated from Fig 5d of} \cite{haidarliu_size_2001}, \MPone{and scaled
+  estimated from Fig 5d of} \cite{haidarliu_size_2001}, \cmnt{and scaled
   such that $\gamma\in\pm 2$. Note that this scaling is arbitrary because
@@
 resulting domain, $S$, using $M=2$ linear signalling gradients aligned with
-the anterior-posterior and medial-lateral axes.  \mpfour{These gradients are
+the anterior-posterior and medial-lateral axes.  \cmnt{These gradients are
   shown with the barrel field boundary in Fig.\,\ref{fig:main}A for clarity,
@@
 % New section to describe and introduce the metrics
-\metrics{First, we used an algorithm introduced by Honda to measure the
+\cmnt{First, we used an algorithm introduced by Honda to measure the
   discrepancy of each barrel shape from a Dirichletform shape}
-\citep{honda_geometrical_1983}.  \metrics{Low overall values of this
+\citep{honda_geometrical_1983}.  \cmnt{Low overall values of this
   \emph{Honda-}$\delta$ metric obtained from simulated barrels indicate a
@@
   tesselation, and thus with a biological barrel field
-  (for mice $\delta\approx0.054$,} \citealp{senft_mouse_1991}\metrics{, and
-  our analysis of data from} \citealp{zheng_signal_2001} \metrics{ indicates that
+  (for mice $\delta\approx0.054$,} \citealp{senft_mouse_1991}\cmnt{, and
+  our analysis of data from} \citealp{zheng_signal_2001} \cmnt{ indicates that
   the value for rats is similar). For the tesselation
@@
     \includegraphics[width=\linewidth]{./Fig2.png}
-    \caption{\MPthreePar{Exploring the parameter space.
+    \caption{\cmnt{Exploring the parameter space.
         %
@@
 
-\metrics{Against these three metrics we are also able to characterise the
+\cmnt{Against these three metrics we are also able to characterise the
   robustness of self-organization to the model parameters, and to investigate
-  the sensitivity of the model to variation in its inputs.} \MPthreePar{Fig.\,\ref{fig:paramsweep}A shows values of $\delta$, $\eta$ and
+  the sensitivity of the model to variation in its inputs.} \cmnt{Fig.\,\ref{fig:paramsweep}A shows values of $\delta$, $\eta$ and
   $\omega$ obtained after 30000 iterations, from 216 independent simulations,
@@
 
-\MPthreePar{The parameters of the example simulation are indicated in
+\cmnt{The parameters of the example simulation are indicated in
   Fig.\,\ref{fig:paramsweep}A using an asterisk. In
@@
     \caption{
-      \MPthreeSens{Sensitivity analysis.
+      \cmnt{Sensitivity analysis.
         %
@@
 
-\MPthreeSens{Next we conducted a sensitivity analysis to determine the extent
+\cmnt{Next we conducted a sensitivity analysis to determine the extent
   to which the quality of the pattern (after $t=30000$ iterations) is affected
@@
 
-\MPthreeSens{Using the parameters of the example simulation
+\cmnt{Using the parameters of the example simulation
   (Fig.\,\ref{fig:main}C) we established baseline mean and standard deviations
@@
 
-\MPthreePred{To further investigate the interplay of genes intrinsic to the
+\cmnt{To further investigate the interplay of genes intrinsic to the
   developing neocortex and extrinsic factors such as thalamocortical input,
@@
 \citep{shimogori_fibroblast_2005,assimacopoulos_fibroblast_2012}
-\MPthreePred{in which the growth factor Fgf8, which is normally expressed at
+\cmnt{in which the growth factor Fgf8, which is normally expressed at
   the anterior end of the cortical subplate from around E9.5}
-\citep{crossley_mouse_1995}, \MPthreePred{is ectopically expressed (by
+\citep{crossley_mouse_1995}, \cmnt{is ectopically expressed (by
   electroporation) also at the posterior pole. We assume that this results in
   a mirror of the primary barrel cortex boundary along the rostrocaudal axis}
-\citep{assimacopoulos_fibroblast_2012} \MPthreePred{and a mirroring of the
+\citep{assimacopoulos_fibroblast_2012} \cmnt{and a mirroring of the
   anterior-posterior guidance gradient $\rho_1$ at the border between them
@@
   \includegraphics[width=\linewidth]{./Fig4.png} \captionof{figure}{
-    \MPthreePred{Simulating altered barrel development.  Guidance fields
+    \cmnt{Simulating altered barrel development.  Guidance fields
       (\textbf{A}) and emergent barrel pattern (\textbf{B}) in a Fgf8
@@
       reduction observed by }
-    \cite{kossut_effects_1992}. \MPthreePred{\textbf{D} The area of the C3
+    \cite{kossut_effects_1992}. \cmnt{\textbf{D} The area of the C3
       barrel (black squares) reduces as $m$ is reduced, whereas the mean area
@@
 
-\MPthreePred{Finally, to investigate the response of the model to
+\cmnt{Finally, to investigate the response of the model to
   environmental manipulation, we simulated a whisker deprivation
@@
 
-\MPthreePred{Although individual whisker trimming reduces barrel size, if an
+\cmnt{Although individual whisker trimming reduces barrel size, if an
   entire row is trimmed, barrel sizes for the trimmed row are not obviously
@@
 
-\MPthreePred{Together with the results of simulated misexpression, the
+\cmnt{Together with the results of simulated misexpression, the
   consistency of the simulated whisker trimming results with those of the
@@
 
-\MPone{We concentrated on the representation of the forty-one macrovibrissae
+\cmnt{We concentrated on the representation of the forty-one macrovibrissae
   that constitute a given barrel field, because their thalamic and cortical
@@
   \nabla\vcdot\mb{J}_i(\mb{x},t) = \nabla\vcdot\left(D \nabla
-  a_i-a_i\sum_{j=1}^{M} \gamma_{i,j}\nabla \rho_j(\mb{x}) \color{colmptwo}+
+  a_i-a_i\sum_{j=1}^{M} \gamma_{i,j}\nabla \rho_j(\mb{x}) \color{colcmnt}+
   \frac{\epsilon a_i}{N-1} \nabla \hat{a}_i \color{black}\right),
@@
 %
-\MPtwo{where $\hat{a}_i\equiv\sum_{j\ne i}^{N}a_j$.} Note that the sum of the
+\cmnt{where $\hat{a}_i\equiv\sum_{j\ne i}^{N}a_j$.} Note that the sum of the
 guidance gradients is time-independent and define $\mb{g}_i(\mb{x}) \equiv
@@
 \nabla\vcdot\mb{J}_i = \nabla\vcdot\big(D \nabla a_i\big) -
-\nabla\vcdot\big(a_i \mb{g}_i\big) \color{colmptwo} + \frac{\epsilon}{N-1}\nabla\vcdot\big({a}_i\nabla\hat{a}_i\big)\color{black}.
+\nabla\vcdot\big(a_i \mb{g}_i\big) \color{colcmnt} + \frac{\epsilon}{N-1}\nabla\vcdot\big({a}_i\nabla\hat{a}_i\big)\color{black}.
 \end{equation}
@@
 \mb{g}_i\vcdot\nabla a_i % term3
-\color{colmptwo}
+\color{colcmnt}
 + \frac{\epsilon a_i}{N-1} \dvrg \hat{a}_i % term1_1
@@
 %
-which has \MPtwo{five} elements to compute: i)~$D \dvrg a_i$ (the Laplacian of
+which has \cmnt{five} elements to compute: i)~$D \dvrg a_i$ (the Laplacian of
 $a_i$); ii)~a time-independent modulator of $a_i$ (because
@@
 product of the static vector field $\mb{g}_i$ and the gradient of $a_i$;
-\MPtwo{iv)~the Laplacian of $\hat{a}_i$; and v)~a term involving the gradients
+\cmnt{iv)~the Laplacian of $\hat{a}_i$; and v)~a term involving the gradients
   of $a_i$ and $\hat{a}_i$}. Each of the divergences can be simplified by
@@
 \noindent
-iii) The \MPtwo{middle} term in Eq.\,\ref{eq:divJExpanded} is the scalar
+iii) The \cmnt{middle} term in Eq.\,\ref{eq:divJExpanded} is the scalar
 product of two vector fields which is straightforward to compute from their
@@
 \noindent
-\MPtwo{iv) The same method used to compute $\dvrg a_i$ in term (i) is used to
+\cmnt{iv) The same method used to compute $\dvrg a_i$ in term (i) is used to
   compute $\dvrg \hat{a}_i$.}
@@
 \noindent
-\MPtwo{v) The final term is the scalar product of the two vector fields
+\cmnt{v) The final term is the scalar product of the two vector fields
   $\nabla a_i$ and $\nabla \hat{a}_i$.}
@@
 Eq.\,\ref{eq:divJExpanded} also contribute nothing to $\mb{J}$. This is
-achieved by \MPtwo{multiplying} $\mb{g}_i(\mb{x})$ \MPtwo{by} a sharp logistic
-function of the distance, $d_b$, from $\mb{x}$ to the boundary, \MPtwo{of the
+achieved by \cmnt{multiplying} $\mb{g}_i(\mb{x})$ \cmnt{by} a sharp logistic
+function of the distance, $d_b$, from $\mb{x}$ to the boundary, \cmnt{of the
   form $1/[1 + \exp(100(d_f-d_b))]$, where $d_f=0.1$\,mm $\approx 3d$ is the
@@
 
-\metrics{The \emph{pattern difference} metric, $\eta$, incorporates
+\cmnt{The \emph{pattern difference} metric, $\eta$, incorporates
   information about the differences in the areas of simulated and
@@
 %
-\metrics{A second comparison considers the mean magnitude of the difference
+\cmnt{A second comparison considers the mean magnitude of the difference
   between the simulated and experimental vectors: $\frac{1}{N} \sum_i \norm{
@@
 %
-\metrics{We combined these terms into a single metric:}
+\cmnt{We combined these terms into a single metric:}
 %
 \begin{equation} \label{eq:eta}
-\color{colmetrics}
+\color{colcmnt}
   \eta =
@@
 %
-\metrics{which has units of mm$^3$.}
+\cmnt{which has units of mm$^3$.}
 
@@
 described in (i)--(v) may be found in the class method
-\code{RD\_James\MPtwo{\_comp2}::compute\_divJ()} which calculates \code{term1},
+\code{RD\_James\cmnt{\_comp2}::compute\_divJ()} which calculates \code{term1},
 \code{term2}, \code{term3}, \code{term1\_1} and \code{term1\_2}, respectively.
@@
 Movie corresponding to Fig.\,\ref{fig:main}C in the main paper. Simulation parameters were
-\MPtwo{$N=41$, $\alpha=3.6$, $\beta=16.67$, $k=3$, $D=0.5$, $\gamma\in\pm 2$,
+\cmnt{$N=41$, $\alpha=3.6$, $\beta=16.67$, $k=3$, $D=0.5$, $\gamma\in\pm 2$,
 $\epsilon=1.2$ and $\delta{t}=0.0001$}. Colours indicate the thalamic
 projection for which the connection density is maximal and black lines delineate
-boundaries. \MPtwo{Barrel labels are
+boundaries. \cmnt{Barrel labels are
   always located at the centroid of their field(s).} The final frame in the movie
@@
 
-\bibliography{../BarrelEmerge}
+\begin{thebibliography}{31}
+\providecommand{\natexlab}[1]{#1}
+\providecommand{\urlprefix}{}
+\providecommand{\doiprefix}{doi: }
+
+\bibitem[{Agmon et~al.(1993)Agmon, A. and Yang, L. T. and O'Dowd, D. K. and
+  Jones, E. G.}]{agmon_organized_1993}
+\textbf{\color{eLifeMediumGrey} Agmon A}, Yang LT, O'Dowd DK, Jones EG.
+\newblock Organized growth of thalamocortical axons from the deep tier of
+  terminations into layer {IV} of developing mouse barrel cortex.
+\newblock Journal of Neuroscience.  1993 Dec; 13(12):5365--5382.
+\newblock \urlprefix\url{http://www.jneurosci.org/content/13/12/5365},
+  \href{10.1523/JNEUROSCI.13-12-05365.1993}{\doiprefix
+  \detokenize{10.1523/JNEUROSCI.13-12-05365.1993}}.
+
+\bibitem[{Ant{\'o}n-Bola{\~n}os et~al.(2019)Ant{\'o}n-Bola{\~n}os, Noelia and
+  Sempere-Ferr{\`a}ndez, Alejandro and Guillam{\'o}n-Vivancos, Teresa and
+  Martini, Francisco J. and P{\'e}rez-Saiz, Leticia and Gezelius, Henrik and
+  Filipchuk, Anton and Valdeolmillos, Miguel and L{\'o}pez-Bendito,
+  Guillermina}]{anton-bolanos_prenatal_2019}
+\textbf{\color{eLifeMediumGrey} Ant{\'o}n-Bola{\~n}os N}, Sempere-Ferr{\`a}ndez
+  A, Guillam{\'o}n-Vivancos T, Martini FJ, P{\'e}rez-Saiz L, Gezelius H,
+  Filipchuk A, Valdeolmillos M, L{\'o}pez-Bendito G.
+\newblock Prenatal activity from thalamic neurons governs the emergence of
+  functional cortical maps in mice.
+\newblock Science.  2019 Jun; 364(6444):987--990.
+\newblock \urlprefix\url{https://science.sciencemag.org/content/364/6444/987},
+  \href{10.1126/science.aav7617}{\doiprefix
+  \detokenize{10.1126/science.aav7617}}.
+
+\bibitem[{Armstrong-James et~al.(1992)Armstrong-James, M. and Fox, K. and
+  Das-Gupta, A.}]{armstrong-james_flow_1992}
+\textbf{\color{eLifeMediumGrey} Armstrong-James M}, Fox K, Das-Gupta A.
+\newblock Flow of excitation within rat barrel cortex on striking a single
+  vibrissa.
+\newblock Journal of Neurophysiology.  1992 Oct; 68(4):1345--1358.
+\newblock
+  \urlprefix\url{https://www.physiology.org/doi/abs/10.1152/jn.1992.68.4.1345},
+  \href{10.1152/jn.1992.68.4.1345}{\doiprefix
+  \detokenize{10.1152/jn.1992.68.4.1345}}.
+
+\bibitem[{Assimacopoulos et~al.(2012)Assimacopoulos, Stavroula and Kao, Tina
+  and Issa, Naoum P. and Grove, Elizabeth A.}]{assimacopoulos_fibroblast_2012}
+\textbf{\color{eLifeMediumGrey} Assimacopoulos S}, Kao T, Issa NP, Grove EA.
+\newblock Fibroblast {Growth} {Factor} 8 {Organizes} the {Neocortical} {Area}
+  {Map} and {Regulates} {Sensory} {Map} {Topography}.
+\newblock Journal of Neuroscience.  2012 May; 32(21):7191--7201.
+\newblock \urlprefix\url{http://www.jneurosci.org/content/32/21/7191},
+  \href{10.1523/JNEUROSCI.0071-12.2012}{\doiprefix
+  \detokenize{10.1523/JNEUROSCI.0071-12.2012}}.
+
+\bibitem[{Bishop et~al.(2000)Bishop, Kathie M. and Goudreau, Guy and O'Leary,
+  Dennis D. M.}]{bishop_regulation_2000}
+\textbf{\color{eLifeMediumGrey} Bishop KM}, Goudreau G, O'Leary DDM.
+\newblock Regulation of {Area} {Identity} in the {Mammalian} {Neocortex} by
+  {Emx2} and {Pax6}.
+\newblock Science.  2000 Apr; 288(5464):344--349.
+\newblock \urlprefix\url{http://science.sciencemag.org/content/288/5464/344},
+  \href{10.1126/science.288.5464.344}{\doiprefix
+  \detokenize{10.1126/science.288.5464.344}}.
+
+\bibitem[{Crossley and Martin(1995)Crossley, P. H. and Martin, G.
+  R.}]{crossley_mouse_1995}
+\textbf{\color{eLifeMediumGrey} Crossley PH}, Martin GR.
+\newblock The mouse {Fgf8} gene encodes a family of polypeptides and is
+  expressed in regions that direct outgrowth and patterning in the developing
+  embryo.
+\newblock Development.  1995 Feb; 121(2):439--451.
+\newblock \urlprefix\url{https://dev.biologists.org/content/121/2/439},
+  publisher: The Company of Biologists Ltd Section: JOURNAL ARTICLES.
+
+\bibitem[{Dufour et~al.(2003)Dufour, Audrey and Seibt, Julie and Passante, Lara
+  and Depaepe, Vanessa and Ciossek, Thomas and Fris\'en, Jonas and Kullander,
+  Klas and Flanagan, John G. and Polleux, Franck and Vanderhaeghen,
+  Pierre}]{dufour_area_2003}
+\textbf{\color{eLifeMediumGrey} Dufour A}, Seibt J, Passante L, Depaepe V,
+  Ciossek T, Fris\'en J, Kullander K, Flanagan JG, Polleux F, Vanderhaeghen P.
+\newblock Area {Specificity} and {Topography} of {Thalamocortical}
+  {Projections} {Are} {Controlled} by ephrin/{Eph} {Genes}.
+\newblock Neuron.  2003 Jul; 39(3):453--465.
+\newblock
+  \urlprefix\url{http://www.sciencedirect.com/science/article/pii/S0896627303004409},
+  \href{10.1016/S0896-6273(03)00440-9}{\doiprefix
+  \detokenize{10.1016/S0896-6273(03)00440-9}}.
+
+\bibitem[{Erzurumlu and Gaspar(2012)Erzurumlu, Reha S. and Gaspar,
+  Patricia}]{erzurumlu_development_2012}
+\textbf{\color{eLifeMediumGrey} Erzurumlu RS}, Gaspar P.
+\newblock Development and critical period plasticity of the barrel cortex:
+  {Barrel} cortex plasticity.
+\newblock European Journal of Neuroscience.  2012 May; 35(10):1540--1553.
+\newblock
+  \urlprefix\url{http://doi.wiley.com/10.1111/j.1460-9568.2012.08075.x},
+  \href{10.1111/j.1460-9568.2012.08075.x}{\doiprefix
+  \detokenize{10.1111/j.1460-9568.2012.08075.x}}.
+
+\bibitem[{Gao et~al.(1998)Gao, Pan-Pan and Yue, Yong and Zhang, Jian-Hua and
+  Cerretti, Douglas Pat and Levitt, Pat and Zhou,
+  Renping}]{gao_regulation_1998}
+\textbf{\color{eLifeMediumGrey} Gao PP}, Yue Y, Zhang JH, Cerretti DP, Levitt
+  P, Zhou R.
+\newblock Regulation of thalamic neurite outgrowth by the {Eph} ligand
+  ephrin-{A5}: {Implications} in the development of thalamocortical
+  projections.
+\newblock Proceedings of the National Academy of Sciences.  1998 Apr;
+  95(9):5329--5334.
+\newblock \urlprefix\url{https://www.pnas.org/content/95/9/5329},
+  \href{10.1073/pnas.95.9.5329}{\doiprefix
+  \detokenize{10.1073/pnas.95.9.5329}}, publisher: National Academy of Sciences
+  Section: Biological Sciences.
+
+\bibitem[{Haidarliu and Ahissar(2001)Haidarliu, Sebastian and Ahissar,
+  Ehud}]{haidarliu_size_2001}
+\textbf{\color{eLifeMediumGrey} Haidarliu S}, Ahissar E.
+\newblock Size gradients of barreloids in the rat thalamus.
+\newblock Journal of Comparative Neurology.  2001; 429(3):372--387.
+\newblock
+  \urlprefix\url{https://onlinelibrary.wiley.com/doi/abs/10.1002/1096-9861%2820010115%29429%3A3%3C372%3A%3AAID-CNE2%3E3.0.CO%3B2-3},
+  \href{10.1002/1096-9861(20010115)429:3<372::AID-CNE2>3.0.CO;2-3}{\doiprefix
+  \detokenize{10.1002/1096-9861(20010115)429:3<372::AID-CNE2>3.0.CO;2-3}}.
+
+\bibitem[{Honda(1983)Honda, Hisao}]{honda_geometrical_1983}
+\textbf{\color{eLifeMediumGrey} Honda H}.
+\newblock Geometrical {Models} for {Cells} in {Tissues}.
+\newblock In: \emph{International {Review} of {Cytology}}, vol.~81 Elsevier;
+  1983.p. 191--248.
+\newblock
+  \urlprefix\url{https://linkinghub.elsevier.com/retrieve/pii/S0074769608623396},
+  \href{10.1016/S0074-7696(08)62339-6}{\doiprefix
+  \detokenize{10.1016/S0074-7696(08)62339-6}}.
+
+\bibitem[{Jeanmonod et~al.(1981)Jeanmonod, D. and Rice, F. L. and Van der Loos,
+  H.}]{jeanmonod_mouse_1981}
+\textbf{\color{eLifeMediumGrey} Jeanmonod D}, Rice FL, Van~der Loos H.
+\newblock Mouse somatosensory cortex: {Alterations} in the barrelfield
+  following receptor injury at different early postnatal ages.
+\newblock Neuroscience.  1981 Aug; 6(8):1503--1535.
+\newblock
+  \urlprefix\url{http://www.sciencedirect.com/science/article/pii/0306452281902220},
+  \href{10.1016/0306-4522(81)90222-0}{\doiprefix
+  \detokenize{10.1016/0306-4522(81)90222-0}}.
+
+\bibitem[{Karbowski and Ermentrout(2004)Karbowski, Jan and Ermentrout,
+  G.B.}]{karbowski_model_2004}
+\textbf{\color{eLifeMediumGrey} Karbowski J}, Ermentrout GB.
+\newblock Model of the {Early} {Development} of {Thalamo}-{Cortical}
+  {Connections} and {Area} {Patterning} via {Signaling} {Molecules}.
+\newblock Journal of Computational Neuroscience.  2004 Nov; 17(3):347--363.
+\newblock
+  \urlprefix\url{http://link.springer.com/10.1023/B:JCNS.0000044876.28268.18},
+  \href{10.1023/B:JCNS.0000044876.28268.18}{\doiprefix
+  \detokenize{10.1023/B:JCNS.0000044876.28268.18}}.
+
+\bibitem[{Kossut(1992)Kossut, Malgorzata}]{kossut_effects_1992}
+\textbf{\color{eLifeMediumGrey} Kossut M}.
+\newblock Effects of sensory deprivation upon a single cortical vibrissal
+  column: a {2DG} study.
+\newblock Experimental brain research.  1992; 90(3):639--642.
+\newblock Publisher: Springer.
+
+\bibitem[{Land and Simons(1985)Land, P. W. and Simons, D.
+  J.}]{land_metabolic_1985}
+\textbf{\color{eLifeMediumGrey} Land PW}, Simons DJ.
+\newblock Metabolic activity in {SmI} cortical barrels of adult rats is
+  dependent on patterned sensory stimulation of the mystacial vibrissae.
+\newblock Brain Research.  1985 Aug; 341(1):189--194.
+\newblock
+  \urlprefix\url{http://www.sciencedirect.com/science/article/pii/000689938591488X},
+  \href{10.1016/0006-8993(85)91488-X}{\doiprefix
+  \detokenize{10.1016/0006-8993(85)91488-X}}.
+
+\bibitem[{Lee et~al.(2014)Lee, D. and Tien, H.C. and Luo, C.P. and Luk,
+  H.-N.}]{lee_hexagonal_2014}
+\textbf{\color{eLifeMediumGrey} Lee D}, Tien HC, Luo CP, Luk HN.
+\newblock Hexagonal grid methods with applications to partial differential
+  equations.
+\newblock International Journal of Computer Mathematics.  2014 Sep;
+  91(9):1986--2009.
+\newblock
+  \urlprefix\url{http://www.tandfonline.com/doi/abs/10.1080/00207160.2013.864392},
+  \href{10.1080/00207160.2013.864392}{\doiprefix
+  \detokenize{10.1080/00207160.2013.864392}}.
+
+\bibitem[{Ma(1991)Ma, Pokay M.}]{ma_barrelettesarchitectonic_1991}
+\textbf{\color{eLifeMediumGrey} Ma PM}.
+\newblock The barrelettes{\textemdash}architectonic vibrissal representations
+  in the brainstem trigeminal complex of the mouse. {Normal} structural
+  organization.
+\newblock Journal of Comparative Neurology.  1991; 309(2):161--199.
+\newblock
+  \urlprefix\url{https://onlinelibrary.wiley.com/doi/abs/10.1002/cne.903090202},
+  \href{10.1002/cne.903090202}{\doiprefix \detokenize{10.1002/cne.903090202}}.
+
+\bibitem[{Miller et~al.(2006)Miller, Katherine and Kolk, Sharon M. and
+  Donoghue, Maria J.}]{miller_epha7-ephrin-a5_2006}
+\textbf{\color{eLifeMediumGrey} Miller K}, Kolk SM, Donoghue MJ.
+\newblock {EphA7}-ephrin-{A5} signaling in mouse somatosensory cortex:
+  {Developmental} restriction of molecular domains and postnatal maintenance of
+  functional compartments.
+\newblock The Journal of Comparative Neurology.  2006 Jun; 496(5):627--642.
+\newblock \urlprefix\url{http://doi.wiley.com/10.1002/cne.20926},
+  \href{10.1002/cne.20926}{\doiprefix \detokenize{10.1002/cne.20926}}.
+
+\bibitem[{Purves et~al.(1992)Purves, D. and Riddle, D.R. and LaMantia,
+  A-S.}]{purves_iterated_1992}
+\textbf{\color{eLifeMediumGrey} Purves D}, Riddle DR, LaMantia AS.
+\newblock Iterated patterns of brain circuitry (or how the cortex gets its
+  spots).
+\newblock Trends in Neurosciences.  1992 Jan; 15(10):362--368.
+\newblock
+  \urlprefix\url{https://linkinghub.elsevier.com/retrieve/pii/016622369290180G},
+  \href{10.1016/0166-2236(92)90180-G}{\doiprefix
+  \detokenize{10.1016/0166-2236(92)90180-G}}.
+
+\bibitem[{Sehara and Kawasaki(2011)Sehara, Keisuke and Kawasaki,
+  Hiroshi}]{sehara_neuronal_2011}
+\textbf{\color{eLifeMediumGrey} Sehara K}, Kawasaki H.
+\newblock Neuronal {Circuits} with {Whisker}-{Related} {Patterns}.
+\newblock Molecular Neurobiology.  2011 Jun; 43(3):155--162.
+\newblock \urlprefix\url{https://doi.org/10.1007/s12035-011-8170-8},
+  \href{10.1007/s12035-011-8170-8}{\doiprefix
+  \detokenize{10.1007/s12035-011-8170-8}}.
+
+\bibitem[{Senft and Woolsey(1991)Senft, Stephen L. and Woolsey, Thomas
+  A.}]{senft_mouse_1991}
+\textbf{\color{eLifeMediumGrey} Senft SL}, Woolsey TA.
+\newblock Mouse {Barrel} {Cortex} {Viewed} as {Dirichlet} {Domains}.
+\newblock Cerebral Cortex.  1991; 1(4):348--363.
+\newblock
+  \urlprefix\url{https://academic.oup.com/cercor/article-lookup/doi/10.1093/cercor/1.4.348},
+  \href{10.1093/cercor/1.4.348}{\doiprefix
+  \detokenize{10.1093/cercor/1.4.348}}.
+
+\bibitem[{Shimogori and Grove(2005)Shimogori, Tomomi and Grove, Elizabeth
+  A.}]{shimogori_fibroblast_2005}
+\textbf{\color{eLifeMediumGrey} Shimogori T}, Grove EA.
+\newblock Fibroblast {Growth} {Factor} 8 {Regulates} {Neocortical} {Guidance}
+  of {Area}-{Specific} {Thalamic} {Innervation}.
+\newblock Journal of Neuroscience.  2005 Jul; 25(28):6550--6560.
+\newblock \urlprefix\url{http://www.jneurosci.org/content/25/28/6550},
+  \href{10.1523/JNEUROSCI.0453-05.2005}{\doiprefix
+  \detokenize{10.1523/JNEUROSCI.0453-05.2005}}.
+
+\bibitem[{Speer and Chapman(2005)Speer, Colenso M. and Chapman,
+  Barbara}]{speer_grading_2005}
+\textbf{\color{eLifeMediumGrey} Speer CM}, Chapman B.
+\newblock Grading the thalamus: how can an `{Eph}' be excellent?
+\newblock Thalamus and Related Systems.  2005 Sep; 3(03):235.
+\newblock
+  \urlprefix\url{http://www.journals.cambridge.org/abstract_S1472928807000234},
+  \href{10.1017/S1472928807000234}{\doiprefix
+  \detokenize{10.1017/S1472928807000234}}.
+
+\bibitem[{Sperry(1963)Sperry, R. W.}]{sperry_chemoaffinity_1963}
+\textbf{\color{eLifeMediumGrey} Sperry RW}.
+\newblock Chemoaffinity in the {Orderly} {Growth} of {Nerve} {Fiber} {Patterns}
+  and {Connections}.
+\newblock Proceedings of the National Academy of Sciences.  1963 Oct;
+  50(4):703--710.
+\newblock \urlprefix\url{https://www.pnas.org/content/50/4/703},
+  \href{10.1073/pnas.50.4.703}{\doiprefix \detokenize{10.1073/pnas.50.4.703}}.
+
+\bibitem[{Torii et~al.(2013)Torii, Masaaki and Rakic, Pasko and Levitt,
+  Pat}]{torii_role_2013}
+\textbf{\color{eLifeMediumGrey} Torii M}, Rakic P, Levitt P.
+\newblock Role of {EphA}/ephrin-{A} signaling in the development of topographic
+  maps in mouse corticothalamic projections.
+\newblock Journal of Comparative Neurology.  2013; 521(3):626--637.
+\newblock
+  \urlprefix\url{https://onlinelibrary.wiley.com/doi/abs/10.1002/cne.23195},
+  \href{10.1002/cne.23195}{\doiprefix \detokenize{10.1002/cne.23195}},
+  \_eprint: https://onlinelibrary.wiley.com/doi/pdf/10.1002/cne.23195.
+
+\bibitem[{Van Der~Loos(1976)Van Der Loos,
+  Hendrik}]{van_der_loos_barreloids_1976}
+\textbf{\color{eLifeMediumGrey} Van Der~Loos H}.
+\newblock Barreloids in mouse somatosensory thalamus.
+\newblock Neuroscience Letters.  1976 Mar; 2(1):1--6.
+\newblock
+  \urlprefix\url{http://www.sciencedirect.com/science/article/pii/0304394076900367},
+  \href{10.1016/0304-3940(76)90036-7}{\doiprefix
+  \detokenize{10.1016/0304-3940(76)90036-7}}.
+
+\bibitem[{Vanderhaeghen et~al.(2000)Vanderhaeghen, Pierre and Lu, Qiang and
+  Prakash, Neal and Fris{\'e}n, Jonas and Walsh, Christopher A. and Frostig,
+  Ron D. and Flanagan, John G.}]{vanderhaeghen_mapping_2000}
+\textbf{\color{eLifeMediumGrey} Vanderhaeghen P}, Lu Q, Prakash N, Fris{\'e}n
+  J, Walsh CA, Frostig RD, Flanagan JG.
+\newblock A mapping label required for normal scale of body representation in
+  the cortex.
+\newblock Nature Neuroscience.  2000 Apr; 3(4):358--365.
+\newblock \urlprefix\url{https://www.nature.com/articles/nn0400_358},
+  \href{10.1038/73929}{\doiprefix \detokenize{10.1038/73929}}.
+
+\bibitem[{Vanderhaeghen and Polleux(2004)Vanderhaeghen, Pierre and Polleux,
+  Franck}]{vanderhaeghen_developmental_2004}
+\textbf{\color{eLifeMediumGrey} Vanderhaeghen P}, Polleux F.
+\newblock Developmental mechanisms patterning thalamocortical projections:
+  intrinsic, extrinsic and in between.
+\newblock Trends in Neurosciences.  2004 Jul; 27(7):384--391.
+\newblock
+  \urlprefix\url{http://www.sciencedirect.com/science/article/pii/S0166223604001638},
+  \href{10.1016/j.tins.2004.05.009}{\doiprefix
+  \detokenize{10.1016/j.tins.2004.05.009}}.
+
+\bibitem[{Woolsey and Van~der Loos(1970)Woolsey, Thomas A. and Van der Loos,
+  Hendrik}]{woolsey_structural_1970}
+\textbf{\color{eLifeMediumGrey} Woolsey TA}, Van~der Loos H.
+\newblock The structural organization of layer {IV} in the somatosensory region
+  ({S} {I}) of mouse cerebral cortex: {The} description of a cortical field
+  composed of discrete cytoarchitectonic units.
+\newblock Brain Research.  1970 Jan; 17(2):205--242.
+\newblock
+  \urlprefix\url{http://www.sciencedirect.com/science/article/pii/000689937090079X},
+  \href{10.1016/0006-8993(70)90079-X}{\doiprefix
+  \detokenize{10.1016/0006-8993(70)90079-X}}.
+
+\bibitem[{Woolsey et~al.(1975)Woolsey, Thomas A. and Welker, Carol and
+  Schwartz, Richard H.}]{woolsey_comparative_1975}
+\textbf{\color{eLifeMediumGrey} Woolsey TA}, Welker C, Schwartz RH.
+\newblock Comparative anatomical studies of the {Sml} face cortex with special
+  reference to the occurrence of {\textquotedblleft}barrels{\textquotedblright}
+  in layer {IV}.
+\newblock The Journal of Comparative Neurology.  1975 Nov; 164(1):79--94.
+\newblock \urlprefix\url{http://doi.wiley.com/10.1002/cne.901640107},
+  \href{10.1002/cne.901640107}{\doiprefix \detokenize{10.1002/cne.901640107}}.
+
+\bibitem[{Zheng et~al.(2001)Zheng, Ying and Johnston, David and Berwick, Jason
+  and Mayhew, John}]{zheng_signal_2001}
+\textbf{\color{eLifeMediumGrey} Zheng Y}, Johnston D, Berwick J, Mayhew J.
+\newblock Signal {Source} {Separation} in the {Analysis} of {Neural} {Activity}
+  in {Brain}.
+\newblock NeuroImage.  2001 Mar; 13(3):447--458.
+\newblock
+  \urlprefix\url{https://linkinghub.elsevier.com/retrieve/pii/S1053811900907055},
+  \href{10.1006/nimg.2000.0705}{\doiprefix
+  \detokenize{10.1006/nimg.2000.0705}}.
+
+\end{thebibliography}
 

### 2020-07-16 paper/briefpaper/barrels_proof.tex
Renames Movie S1 to Video 1 in line with eLife standards

@@
       maximal, \cmnt{barrel labels are located at the centroid of each
-        region and} black lines delineate boundaries (see Movie S1).
+        region and} black lines delineate boundaries (see Video 1).
       %
@@
   as shown in Fig.\,\ref{fig:main}C. Parameter values for this example simulation
-  (see also Movie S1) were obtained by
+  (see also Video 1) were obtained by
   conducting a full parameter sweep and choosing a combination
@@
 
-\section{Movie S1 caption}
-
-Movie corresponding to Fig.\,\ref{fig:main}C in the main paper. Simulation parameters were
+\section{Video 1 caption}
+
+Video corresponding to Fig.\,\ref{fig:main}C in the main paper. Simulation parameters were
 \cmnt{$N=41$, $\alpha=3.6$, $\beta=16.67$, $k=3$, $D=0.5$, $\gamma\in\pm 2$,

### 2020-07-16 paper/briefpaper/barrels_review.tex
Improved naming
Renames Movie S1 to Video 1 in line with eLife standards

--- 2020-07-14 paper/briefpaper/barrels_proof.tex
@@
 
-% Black, so that updates following review do not show
-\definecolor{colcmnt}        {rgb} {0.0, 0.0, 0.0}
+% Custom pink for comments
+\definecolor{colcmnt}        {rgb} {0.7098, 0.0745, 0.6431}
 % A command to highlight changes for reviewers
